### 2021-07-13 move_schedules.tex
Initial commit

\documentclass[10pt]{article}
\usepackage[margin=0.75in]{geometry}
\usepackage{amsmath, amsthm, amssymb}
\usepackage{mathtools}
\usepackage{url}
\usepackage{enumitem}
\usepackage{setspace}
\usepackage{scalefnt}
\usepackage[utf8]{inputenc}
\usepackage[english]{babel}
\usepackage{algorithm}
%\usepackage{algorithm2e}
%\usepackage{algorithmicx}
\usepackage[noend]{algpseudocode}
\usepackage{bbm}
\usepackage{graphicx}

\newcommand\sbullet[1][.5]{\mathbin{\vcenter{\hbox{\scalebox{#1}{$\bullet$}}}}}
\usepackage{blkarray, bigstrut}
\usepackage{tikz}
\usetikzlibrary{arrows.meta, calc, quotes, tikzmark,shapes}

\usepackage{caption} 
\captionsetup[table]{skip=4pt, position=above}

\usepackage{array}
\newcolumntype{P}[1]{>{\centering\arraybackslash}p{#1}}

\newcommand\topstrut[1][1.1ex]{\setlength\bigstrutjot{#1}{\bigstrut[t]}}
\newcommand\botstrut[1][0.95ex]{\setlength\bigstrutjot{#1}{\bigstrut[b]}}

\makeatletter
\newcommand{\customlabel}[2]{%
   \protected@write \@auxout {}{\string \newlabel {#1}{{#2}{\thepage}{#2}{#1}{}} }%
   \hypertarget{#1}{#2}
}
\makeatother

\usepackage{hyperref} % load hyperref last

\newtheorem{theorem}{Theorem}
\newtheorem{definition}{Definition}
\newtheorem{proposition}{Proposition}
\newtheorem{lemma}{Lemma}
\newtheorem{corollary}{Corollary}
% \newcommand{\CommentX}[1]{\unskip~\%~#1~\%}

\DeclareMathOperator*{\argmax}{arg\,max}
\DeclareMathOperator*{\argmin}{arg\,min}
\DeclarePairedDelimiter\ceil{\lceil}{\rceil}
\DeclarePairedDelimiter\floor{\lfloor}{\rfloor}

\usepackage{outlines}
\usepackage{enumitem}
\setenumerate[1]{label=\arabic*}
\setenumerate[2]{label*=.\arabic*}
\setenumerate[3]{label*=.\arabic*}
\setenumerate[4]{label*=.\arabic*}

%Unraveling
\title{\vspace{-2.0em} Move Schedules: \\ Fast persistence computations in sparse dynamic settings\thanks{This work was partially supported by the National Science Foundation through grants CCF-2006661
and CAREER award  DMS-1943758.}\vspace{-0.5em}}
\author{Matt Piekenbrock\thanks{Department of Computational Mathematics, Science \& Engineering, Michigan State University}\;\; and Jose A. Perea$^\dagger$\thanks{Department of Mathematics, Michigan State University}}
\date{}

\begin{document} \vspace{-4em} \maketitle
\begin{abstract}
	The standard procedure for computing the persistent homology of a filtered simplicial complex is the matrix reduction algorithm. Its output is a particular decomposition of the total boundary matrix, from which the persistence diagrams and generating cycles can be derived. 
	Persistence diagrams are known to vary continuously with respect to their input; this motivates the algorithmic study of persistence computations for time-varying filtered complexes. Computationally, simulating persistence dynamically can be reduced to maintaining a valid decomposition under adjacent transpositions in the filtration order. In practice, the quadratic scaling in the number of  transpositions often makes this maintenance procedure slower than simply computing the decomposition from scratch, effectively limiting the application of dynamic persistence to relatively small data sets. In this work, we propose a coarser strategy for maintaining the decomposition over a discrete 1-parameter family of filtrations. Our first result is an analysis of a simple linear-time strategy which reduces the number of column operations needed to simulate persistence across a fixed homotopy by at most a factor of 2. We then show a modification of this technique which maintains only a sublinear number of valid states, as opposed to a quadratic number of states, and we provide tight lower bounds for this technique.
	Finally, we provide empirical results suggesting that the decrease in operations needed to compute diagrams across a family of filtrations is proportional to the difference between the expected quadratic number of states, and the proposed sublinear coarsening.
\end{abstract}

\section{Introduction} 
\subsection{Overview}\label{sec:overview} 
The starting point to the work described in this paper is the stability results  established in~\cite{cohen2007stability, cohen2006vines}. 
Indeed, given a triangulable topological space equipped with a sufficiently tame continuous function, persistent homology captures the changes in topology across the sublevel sets of the space, and encodes them in a persistence diagram. The stability of persistence contends 
that small changes in the underlying function produce small changes in the corresponding diagrams:  if the function changes continuously, so too will the points on the persistence diagram. 
This motivates the study and application of persistence to time-varying or dynamic settings, like that of dynamic metric spaces~\cite{kim2020spatiotemporal}.
Focusing on the computational aspects of this problem has far-reaching consequences: methods that vectorize  persistence diagrams for machine learning purposes, like adaptive template functions~\cite{polanco2019adaptive} and persistence images~\cite{adams2017persistence}, or summary statistics like $\alpha$-smoothed Betti curves~\cite{ulmer2019topological}, all immediately become  computationally viable tools for time-varying settings. 
%This analogy provides a strong interpretation at gaining insight on the subtle topological changes that may occur in studying a continuous process.

Cohen-Steiner et al. refer to a continuous 1-parameter family of persistence diagrams as a \emph{vineyard}, and they give in \cite{cohen2006vines} an efficient algorithm for their computation
given a time-varying filtration. 
Computing vineyards can be interpreted as an extension of the \emph{matrix reduction} algorithm~\cite{zomorodian2005computing}, which computes the persistence diagram of a fixed filtration $K$ with $m$ simplices in $O(m^3)$ time.
The main step in the reduction algorithm is  a specific   decomposition $R = D V$ of the  boundary matrix $D$ of $K$. 
Since $V$ is always full rank (and upper-triangular), an alternative is to compute a $D = RU$ decomposition, where $U = V^{-1}$. 
The vineyards algorithm transforms a time-varying filtration into an ordered set of adjacent transpositions to be applied to the  rows and columns of the decomposition $D = RU$, each of which take at most $O(m)$ time to execute. If one is interested in understanding how the persistent homology of a continuous function changes over time, then this algorithm is sufficient, for  homological critical points can only occur when the filtration order changes. 
If there are $d$ time-points where this order changes, then the vineyards algorithm takes $O(m^3 + md)$ time: one $O(m^3)$-time reduction upfront for the filtration at time $t_0$ followed by one $O(m)$ operation to update the decomposition for the remaining time points $(t_1, t_2, \dots, t_d)$. When $d >> m$, this $O(md)$ approach is far more efficient than the $O(dm^3)$  naive strategy of computing the diagrams independently at every time point.

Despite the vineyards algorithm being efficient in theory, it is often not the method of choice in practice in dynamic settings. 
While there is an increasingly rich ecosystem of software packages offering variations of the standard reduction algorithm (e.g. Ripser, PHAT, Dionysus, etc... see~\cite{otter2017roadmap} for an overview), implementations of the vineyards algorithm are relatively uncommon.\footnote{Dionysus 1 has an implementation of vineyards, but the algorithm was never ported to the 2nd version.} The reason for this disparity is perhaps explained by Lesnick and Wright~\cite{lesnick2015interactive}: ``While an update to an $RU$ decomposition involving few transpositions is very fast in practice... many transpositions can be quite slow... it is sometimes much faster to simply recompute the $RU$-decomposition from scratch using the standard persistence algorithm.'' Indeed, whether performing the reductions independently or updating the decomposition dynamically using vineyards, they observed that obtaining the decomposition is the most computationally demanding aspect of their 2-D persistence algorithm.

This work seeks to further understand and remedy this discrepancy: building on the work presented in~\cite{busaryev2010tracking}, we introduce a coarser approach to the vineyards algorithm.
While the vineyards algorithm is efficient at constructing a \emph{continuous} time family of diagrams, it is not necessarily efficient when the time parameter is coarsely discretized.
Our methodology is based on the observation that practitioners often don't need (or want!) all $d$ persistence diagrams generated by a continuous 1-parameter of filtrations; usually just $n << d$ of them are sufficient.   
% in many time varying settings, a valid decomposition is often only sought after at a relatively sparse number of time points relative to the size of the (finite) set needed to generated by the family. 
By making a tradeoff in the number of times we restore the decomposition to a valid state, we are able to reduce the total number of column operations needed to apply an arbitrary permutation to the filtration. This tradeoff, paired with a fast greedy heuristic explained in section~\ref{sec:proxy_objective}, yields an algorithm that can update a $R = DV$ decomposition more efficiently than vineyards in sparse contexts, making dynamic persistence more computationally tractable for a wider class of use-cases.

 %TODO fill in section reference above.

 \subsection{A Motivating Example and Contributions}\label{sec:motivation} 
As a motivating example, we describe a simple experiment illustrating why the vineyards algorithm does not always yield an efficient method in time-varying applications. 
Consider a series of grayscale images (i.e. a video) depicting a fixed-width annulus expanding about the center of a $9 \times 9$ grid, where scale of the annulus is parameterized by a single parameter $\tau$. Thus, the data set is a 1-parameter family of images. Each image contains the same number of pixels whose intensities---these are determined by the parameter---vary with time. For each image, we build a filtered simplicial complex using the \emph{Freudenthal} triangulation, with the ordering of inclusions obtained by lower stars.  
 \begin{figure*}[t]
 	\centering
 	\includegraphics[width=0.36\textwidth, height = 1.65in]{circle_vineyards.png}
 	\includegraphics[width=0.63\textwidth]{circle_complex.png}
 	\caption{Grayscale image data set depicting an expanding annulus (top right), each image's corresponding lower star filtration (bottom right), and the cumulative number of column operations required to compute persistence for this time-varying filtration (left). The persistent Betti numbers for each filtration from left to right are: $\beta_1 = (1,0,0), \beta_2 = (1,1,0), \beta_3 = (1,1,0), \beta_4 = (1,1,0), \beta_5 = (4,0,0)$. Observe computing $10$ diagrams independently at evenly spaced time points (the green line) both captures these major topological changes and is the most computationally efficient approach shown.}
 	\label{fig:vineyards}
 \end{figure*}
This time-varying filtration is depicted in Figure~\ref{fig:vineyards} (bottom right), which we use as a baseline for comparing the standard algorithm $\texttt{pHcol}$ with vineyards. Note the size of the filtration is fixed--no simplices are ever added or removed over time. 
There are two events which dramatically alter the persistence diagrams extracted from these filtrations: the first occurs when the central connected component splits to form a cycle, and the second occurs when the annulus approaches the outer dimensions of the grid, splitting into four components. As a consequence, in this example, only a few persistence diagrams are needed to capture the major topological changes over time.

  Consider the situation where one is concerned with capturing these dramatic changes in homology.
  On the left side of Figure~\ref{fig:vineyards}, we compare the cumulative cost (in total number of column operations) of various approaches which one can use to simulate persistence dynamically. 
  Since it is unknown a-priori at what points in  time the persistent homology will change in application settings, one solution is to discretize the time domain at $n$ evenly spaced points and compute the persistence diagram at each of these time points independently. An alternative approach is to construct a homotopy between two time-points, and then decompose this homotopy into adjacent transpositions--the vineyards approach. The former is often used in practice, but it is only an approximation; the latter is guaranteed to capture all homological changes in persistence that occur in simulating the 1-parameter family. The graph in Figure~\ref{fig:vineyards} depicts two different size approximations (purple, $n=100$ and green, $n=10$), wherein the reduction algorithm (\texttt{pHcol}) was applied at $n$ time points, and two exact strategies (black and blue), which compute all $\approx 7 \times 10^{4}$ diagrams given by the homotopy. As one can see from the figure, if one needs to simulate the homotopy exactly, the vineyards approach is indeed far more efficient than naively  applying the reduction algorithm independently at all time points.
  However,  when the discretization of the time domain is sparse enough, the naive approach   performs less column operations than the vineyards strategy, while still capturing the main events.

  The existence of a time discretization that is more efficient to compute than continually updating the decomposition indicates that the vineyards algorithm must incur some amount of overhead (in terms of column operations) to perform the transpositions. That is, even in situations where one is updating a decomposition to a filtration that is  relatively ``close'', e.g. any two adjacent filtrations in the bottom left of Figure~\ref{fig:vineyards}, it is still more efficient to apply \texttt{pHcol} directly than  to iteratively update the decomposition (see the case where $n = 10$). 
 Indeed, any optimizations to the reduction algorithm (for example, the clearing optimization~\cite{chen2011persistent}) would only increase this disparity, making the vineyards algorithm less computationally attractive even in settings where $n$ is larger. Since dynamic or time-varying are common settings where persistence is applied (see ~\cite{topaz2015topological, xian2020capturing, lesnick2015interactive, kim2020spatiotemporal}), it is worth studying this phenomenon.

\textbf{Our results and approach are as follows:} First, we leverage work of Busaryev et. al.~\cite{busaryev2010tracking} to include a coarser type of operation to the dynamic setting, called \emph{moves}. We give a tight lower bound on the number of moves needed to perform an arbitrary permutation to the $R = D V$ decomposition, and give a proof of optimality by a reduction to the permutation edit distance problem.
We also give worst-case bounds in expectation as well as efficient algorithms for achieving these bounds---both of which are derived from a reduction to the Longest Increasing Subsequence (LIS) problem. 
This reduction yields an efficient algorithm for generating sequences of moves $( s_1, s_2, \dots, s_d )$, which we call \emph{schedules}. 
We investigate the feasibility of optimizing the cost 
(i.e., number of column operations) of these schedules directly, including the existence of a greedy-type approach, which we show can lead to arbitrarily bad behavior. 
In light of these results, we give an alternative proxy-objective to minimize, provide bounds justifying its relevance to the original objective, and give an efficient $O(d \log d)$ algorithm for approximately solving this proxy minimization. 
A performance comparison with other reduction-based persistent homology computations is given, wherein our approach is demonstrated to be an order of magnitude more efficient than existing approaches in sparse dynamic settings. We conclude with a discussion of situations where our strategy is most applicable and future directions for research.

\subsection{Related Work}\label{sec:related_work} 
To the authors knowledge,   work focused on ways of updating a  decomposition $R = DV$, for all homological dimensions, is limited---there is the vineyards algorithm~\cite{cohen2006vines} and the moves algorithm~\cite{busaryev2010tracking}, both of which are discussed extensively in section~\ref{sec:background}.

In contrast, there is extensive work  on improving the efficiency of computing a (static) $R = DV$ decomposition. Chen~\cite{chen2011persistent} proposed \emph{persistence with a twist}, also called the \emph{clearing optimization}, which exploits a boundary/cycle relationship to ``kill'' columns early in the reduction rather than reducing them. 
Another popular optimization is to utilize the duality between   homology and cohomology~\cite{de2011dualities}, which dramatically improves the effectiveness of the clearing optimization~\cite{bauer2021ripser}. 
There are many other optimizations on the implementation side: the use of ranking functions defined on the combinatorial number system enables implicit cofacet enumeration, removing the need to store the boundary matrix explciitly; the apparent/emergent pairs optimization identifies columns whose pivot entries are unaffected by the reduction algorithm, reducing the total number of columns which need be reduced; sparse data structures such as bit-trees and lazy heaps allow for efficient column-wise additions with $\mathbb{Z}_2 = \mathbb{Z}/2\mathbb{Z}$ coefficients and effective $O(1)$ pivot entry retrieval, and so on~\cite{bauer2017phat, bauer2021ripser}.

By making stronger assumptions on the underlying topological space, restricting the homological dimension, or targeting  a weaker invariant (e.g. Betti numbers), one can usually obtain  faster or alternative approaches  recording information related to persistence.
For example, Attali et al.~\cite{attali2009persistence} give a linear time algorithm (under the $\texttt{word RAM}$ model of computation) for computing persistence on graphs.
In the same paper, they describe how to obtain $\epsilon$-simplifications of $1$-dimensional persistence diagrams for filtered $2$-manifolds by using duality and symmetry theorems. 
Along a similar vein, Edelsbrunner et. al.~\cite{edelsbrunner2000topological} give a fast incremental algorithm for computing persistent Betti numbers up to dimension $2$, again by utilizing symmetry, duality, and ``time-reversal''~\cite{delfinado1995incremental}. Chen et. al.~\cite{chen2013output} give an output-sensitive method for computing persistent homology,  utilizing the property that certain submatrices of  $D$ have the same rank as  $R$, which they exploit through fast sub-quadratic sparse-matrix rank algorithms.

If  zeroth homology is the only dimension of interest, computing and updating both the persistence and rank information of the persistent homology groups is greatly simplified. For example, if the relations are available a-priori, obtaining a tree representation fully characterizing the connectivity of the underlying space (also known as the \emph{incremental connectivity} problem) takes just $O(\alpha(n) n)$ time using the disjoint-set data structure, where $\alpha(n)$ is the extremely slow-growing inverse Ackermann function. 
Adapting this approach to the time-varying setting, Oesterling et al.~\cite{oesterling2015computing} give a $O(e)$-per-update algorithm that allows one to maintain a \emph{merge tree} with $e$ edges,
associated to the filtration parameter changing continuously over time. 
If only the zeroth-dimensional Betti numbers are needed for a particular application, this problem reduces even further to tracking the connected components of a dynamic graph---sometimes referred to as the \emph{dynamic connectivity problem}, which can be efficiently  solved in amortized $O(\log n)$ query and update times using either Link-cut trees or multi-level Euler tour trees~\cite{kapron2013dynamic}.

\subsection{Outline} The remainder of the paper is organized as follows: we review and establish the notation we will use to describe simplicial complexes, persistent homology, and dynamic persistence. We also cover the reduction algorithm (designated here as \texttt{pHcol}), the vineyards algorithm, and the set of \emph{move}-related algorithms introduced in~\cite{busaryev2010tracking}, which serves as the starting point of this work. 
In section 3 we introduce our proposed alternative to vineyards and provide efficient algorithms to make this alternative viable. 
In section 4 we demonstrate a few applications of the proposed method. In section 5 we conclude the paper by discussing other possible applications, future work. 
We also include an appendix which introduces certain topics in more detail, such as the vineyards algorithm itself, for readers who may be unfamiliar but would like a more in-depth understanding.

%\section*{Outline}
%\begin{outline}[enumerate]
%   \1 Background
%      \2 Point A1
%      \2 Point A2
%         \3 ...
%            \4 ...
%   \1 Point B
%      \2 Point B1
%         \3 Point B1a
%            \4 ...
%\end{outline}
% Interesting+ hard + solved + novel.

\section{Background}\label{sec:background} 
\subsection{Notation}\label{sec:notation}

Suppose one has a family $\{K_i\}_{i\in I}$ of simplicial complexes indexed by a totally ordered index set $I$, and so that for any $i< j \in I$ we have $K_i \subseteq K_j$. There are two index sets of interest in this effort: $\mathbb{R}$ and $[n] = \{ 1, \dots, n\}$. 
Such a family is called a \emph{filtration},
which is deemed to be \emph{essential} if $i \neq j$ implies $K_i \neq K_j$.
Moreover, an essential filtration is said to be  \emph{simplexwise} if $K_j \smallsetminus K_i = \{\sigma_j\}$ for all $i< j \in I$ with $K_j \neq \emptyset$.
Any filtration may be trivially converted into an essential, simplexwise filtration via a set of \emph{condensing}, \emph{refining}, and \emph{reindexing} maps (see~\cite{bauer2021ripser} for more details). 
As a result, and without loss of generality, here we exclusively consider essential simplexwise filtrations. For brevity, we will simply refer to them as filtrations.

Let $K$ be an abstract simplicial complex. With respect to some field $\mathbb{F}$, a $p$-chain is a
formal $\mathbb{F}$-linear combination  of $p$-simplices of $K$. The collection of $p$-chains under addition yields an abelian group called the $p$-th chain group of $K$, denoted   $C_p(K)$. 
The $p$-boundary $\partial_p(\sigma)$  of a $p$-simplex $\sigma$ is the alternating sum of its oriented co-dimension 1 faces; the $p$-boundary of a $p$-chain is the sum of the boundaries of its simplices. A $p$-cycle is a $p$-chain with zero boundary. The collection of $p$-cycles forms the group $Z_p(K) = \mathrm{Ker}\,\partial_p$, the collection of $p$-boundaries forms the group $B_p(K) = \mathrm{Im}\,\partial_{p+1}$, and the quotient of these two yields the $p$-th homology group, $H_p(K) = Z_p(K) / B_{p}(K)$. 
If $\{K_i\}_{i\in [m]}$ is a filtration, then the inclusion maps  $K_i\subset K_j$  for $i < j$, induce homomorphisms $f_p^{i,j}: H_p(K_i) \to H_p(K_j)$ between their corresponding homology groups:
\begin{equation}
	H_p(K_1) \to H_p(K_2) \to \dots \to H_p(K_m)
\end{equation}
The $p$-th persistent homology groups are the images of these homomorphisms: $H_{p}^{i,j} = \mathrm{Im}\,f_p^{i,j}$. Note that if $i = j$, then $H_{p}^{i,j} = H_{p}(K_i) = H_{p}(K_i)$ is   just the ``standard'' homology. It is common for simplices which create new homology classes to be called \emph{creators} and simplices that destroy homology classes to be called \emph{destroyers}. 
Similarly, the filtration indices of 
these creators/destroyers are referred to as \emph{birth} and \emph{death} times, respectively. 
If a homology class $[c]$ is born at time $i$ and dies entering time $j$, the difference $\lvert i - j \rvert$ is called the \emph{persistence} of that class. In practice, filtrations often arise from triangulations parameterized by geometric scaling parameters, and the ``persistence'' of a homology class actually refers to its lifetime with respect to the scaling parameter. 
%For this reason, we'll refer to the difference $\lvert i - j \rvert$ of a homology class $[c]$ as its \emph{index persistence}.
\\
\\
\noindent \textbf{Example 2.1:} Non-essential filtrations often arise in geometrical contexts. For example, given a finite metric space $(X, d)$, the \emph{Vietoris-Rips} complex at scale $\epsilon \in \mathbb{R}$ is the abstract simplicial complex given by: 
\begin{equation}\label{eq:rips}
	\mathrm{Rips_\epsilon}(X) = \{ S \subseteq X : S \neq \emptyset \;\mbox{ and }\;\mathrm{diam}(S) \leq \epsilon \}
\end{equation} 
The corresponding filtration is called the \emph{Vietoris-Rips} (VR) filtration, indexed by $I = \mathbb{R}$. 
A VR filtration indexed over $\mathbb{R}$  can be condensed to an essential filtration indexed over the set of pairwise distances $\{\, d(x,x') \; | \; x,x' \in X \, \}$, 
which can then be further reindexed to an essential simplexwise filtration by extending a total order of $X$ to its power set 
using the lexicographical ordering. 
%This is often necessary in the computational setting, where it is easier to work with a filtration with $m$ simplices indexed over $[m]$ instead of $\mathbb{R}$. 
\\
\\
Given a triangulable topological space $\mathbb{X}$ equipped with a
real-valued function $f: \mathbb{X} \to \mathbb{R}$, 
we write $\mathbb{X}_a = f^{-1}(-\infty, a]$ to denote the sublevel sets of $\mathbb{X}$ defined by the value $a$. 
A \emph{homological critical value} of $f$ is any value $a \in \mathbb{R}$ such that the homology of the sublevel sets of $f$ changes at $a$, i.e. the inclusion-induced map $H(\mathbb{X}_{a - \epsilon}) \to H(\mathbb{X}_a)$ is not an isomorphism. If there are only finitely many of these homological critical values, $f$ is said to be \emph{tame}.

Consider a homotopy $F(x,\tau) : \mathbb{X} \times [0,1] \to \mathbb{R}$ and denote it's ``snapshot'' at a given time-point $\tau$ by $f_\tau(x) = F(x,\tau)$.
The snapshot $f_0$ denotes the initial function at time $\tau = 0$ and $f_1$ denotes the function at the last time step. 
As $\tau$ varies in $[0,1]$, the points in $\mathrm{dgm}_p(f_\tau)$ trace  curves in $\bar{\mathbb{R}}^3$, called \emph{vines} and which together form a \emph{vineyard}. 
The stability of persistence implies that these curves will be continuous if the homotopy is continuous.
The vineyard analogy acts as a guidepost for practitioners seeking to understand how subtle changes occurring to the topological structure over time reveal intrinsic information  about the underlying continuous process. We discuss the vineyards approach more in detail in section~\ref{sec:vineyards}.
%Let $\mathbb{M}$ be a smooth $d$-manifold and $f : \mathbb{M} \to \mathbb{R}$ a smooth function. 
%A critical point $x$ is called \emph{non-degenerate} if the Hessian matrix of second partial derivatives is non-singular, otherwise if the Hessian is zero $x$ is said to be \emph{degenerate}. A \emph{Morse function} is a smooth function that has only non-degenerate critical points with distinct critical values. 
%and two Morse functions $f_0, f_1: \mathbb{M} \to \mathbb{R}$, and further suppose we equip $f_0, f_1$ with with a smooth homotopy $F : \mathbb{M}: [0,1] \to \mathbb{R}$ such that $F(x, 0) = f_0(x)$ and $F(x, 1) = f_1(x)$ for $x \in \mathbb{M}$. We use $f_\tau$ to denote the function at $F(x, \tau)$, where $\tau \in [0, 1]$. 
\subsection{The Reduction Algorithm}\label{sec:reduction}
In this section we briefly recount the original reduction algorithm introduced in~\cite{zomorodian2005computing}, also sometimes called the \emph{standard} algorithm or more explicitly \texttt{pHcol}~\cite{de2011dualities}. 
The pseudocode is outlined in Algorithm~\ref{alg:reduce}. Without optimizations, like the clearing optimization or the use of implicit matrix reduction, the standard  algorithm is very inefficient. Nonetheless, the reduction algorithm serves as the foundation of most implementations that compute persistent homology, and its invariants are necessary before introducing both vineyards in section~\ref{sec:vineyards} and our approach in section~\ref{sec:move_schedules}.

\begin{algorithm}[t]
	\setstretch{1.15}
	\caption{Reduction Algorithm (\texttt{pHcol}) }
	\begin{algorithmic}[1]
		\Require{$D = (m \times m)$ filtration boundary matrix }
		\Ensure{$R$ is reduced, $V$ is full rank upper triangular, and $R = D V$}
		\Function{Reduction}{$D$}
		\State $(R, V) \gets (D, I)$
		\For{$j = 1$ \textbf{to} $m$} 
			\While{$\exists \, i < j$ \textbf{such that} $\mathrm{low}_R(i) = \mathrm{low}_R(j)$ }
				\State $\lambda \gets \mathrm{pivot}_R(j)/\mathrm{pivot}_R(i)$
				\State $(\mathrm{col}_R(j), \mathrm{col}_V(j)) \mathrel{-}= \left ( \lambda \cdot \mathrm{col}_R(i), \lambda \cdot \mathrm{col}_V(i) \right )$
			\EndWhile
		\EndFor 
		\State \Return $(R, V)$
		\EndFunction
	\end{algorithmic}	\label{alg:reduce}
\end{algorithm}

 The main output of the reduction algorithm is a matrix decomposition $R = D V$, where the persistence diagram is encoded in $R$. 
 Given a  filtration $K$ with $m$ simplices and maximal dimension $d$, 
 one assembles the elementary boundary chains $\partial(\sigma)$ as columns ordered according to the filtration order, forming the \emph{filtration boundary matrix} $D$. 
 The reduction algorithm can  compute the persistent homology for all dimensions up to $d - 1$.
 In this case,  $D$ is a square matrix of size $m \times m$. Alternatively, a single dimension $i \leq d - 1$ can be computed, in which case one reduces a pair of matrices $D_i, D_{i+1}$, where the former has dimension $d_{i-1} \times d_i$ and the latter $d_{i} \times d_{i+1}$. If one is interested in a particular dimension of homology then the latter approach is preferred. However, the algorithmic description of the algorithms is simpler when considering the full matrix, so $D$ will always be $(m \times m)$ in the descriptions that follow. All algorithms discussed generalize to both situations. We give an example using the matrix pair $(R_1, V_1)$  below. 
\\
\\
\noindent
\textbf{Example 2.1: Reduction} Consider a triangle with vertices $u,v,w$ and edges $a = (u,w)$, $b = (v,w)$, $c = (u, v)$ whose filtration order, read as inclusions from left to right, is given as $K = (u, v, w, a, b, c)$. The reduction algorithm begins by setting $R = D$ and $V = I$, followed by left-to-right column operations until invariant (2) below is satisfied.
If the $i$-th column of $R$ is nonzero, then $\mathrm{low}_R(i)$ 
is the row index of its lowest nonzero entry.
Using $\mathbb{Z}_2$ coefficients and homology in dimension 1 to simplify the presentation, the steps look at follows:
\begin{displaymath}
	\def\arraystretch{0.95}
	\makeatletter\setlength\BA@colsep{2.8pt}\makeatother
	%
	\begin{blockarray}{cccc}{R_1}
	 \kern-0.2em & a & b & c  \\
		\begin{block}{c[ccc]}
  		u \kern-0.1em  & \; 1 &    &  1 \; \topstrut \\
  		v \kern-0.1em & \;     &  1 & 1 \; \\
  		w \kern-0.1em & \; 1 &  1 &    \; \botstrut \\
		\end{block}
	\end{blockarray}
	\; , \;
	\begin{blockarray}{cccc}{V_1}
	\kern-0.2em & a & b & c  \\
		\begin{block}{c[ccc]}
        a \kern-0.1em  & \; 1 &  &  \; \topstrut \\
  		b \kern-0.1em & \;  & 1 & \; \\
  		c \kern-0.1em & \;  &  & 1 \; \botstrut \\
		\end{block}
	\end{blockarray}
	\quad \xrightarrow{} \,
	\begin{blockarray}{cccc}{R_1}
	\kern-0.2em & a & b & c  \\
		\begin{block}{c[ccc]}
  		u \kern-0.1em  & \; 1 & 1 & 1 \; \topstrut \\
  		v \kern-0.1em & \;     & 1 & 1 \; \\
  		w \kern-0.1em & \; 1 &    &    \; \botstrut \\
		\end{block}
	\end{blockarray}
	\; , \;
	\begin{blockarray}{cccc}{V_1}
	\kern-0.2em & a & b & c  \\
		\begin{block}{c[ccc]}
  		a \kern-0.1em  & \; 1 & 1 &    \; \topstrut \\
  		b \kern-0.1em & \;     & 1 &    \; \\
  		c \kern-0.1em & \;     &    & 1 \; \botstrut \\
		\end{block}
	\end{blockarray}
	\quad \xrightarrow{} \,
	\begin{blockarray}{cccc}{R_1}
	\kern-0.2em & b & a & c  \\
		\begin{block}{c[ccc]}
  		u \kern-0.1em  & \;     & 1  &  \; \topstrut \\
  		v \kern-0.1em &  \; 1  & 1  &  \; \\
  		w \kern-0.1em & \; 1  &     &  \; \botstrut \\
		\end{block}
	\end{blockarray}
	\; , \;
	\begin{blockarray}{cccc}{V_1}
	\kern-0.2em & b & a & c  \\
		\begin{block}{c[ccc]}
  		b \kern-0.1em  & \; 1 & 1 & 1 \; \topstrut \\
  		a \kern-0.1em & \;     & 1 & 1 \; \\
  		c \kern-0.1em & \;     &    & 1 \; \botstrut \\
		\end{block}
	\end{blockarray}
	\vspace*{-1em}
\end{displaymath}
The column $c$ in $R$ indicates that dimension $1$ homology is born, which in this filtration is never killed since the dimension of the complex is $2$. Similarly, the columns at $u, v, w$ in $R_0$ (not shown) are all zero, indicating three $0$-dimensional homology classes  are born, which are then killed by the pivot entries in columns $a$ and $b$ in $R_1$.
\\
\\
\noindent
Its clear from Algorithm~\ref{alg:reduce} that a loose upper bound for this reduction is $O(m^3)$, where $m$ is the number of simplices of the filtration; it turns out that this bound is in fact tight, see~\cite{morozov2005persistence} for more details. 
There are actually many variations and optimizations proposed over the past decade to Algorithm~\ref{alg:reduce}; for example, one can also use the \emph{row algorithm}~\cite{de2011dualities}, or reverse the order of the filtration and compute cohomology~\cite{de2011dualities}, etc. Despite these variations in computation, regardless of how the decomposition is obtained, it must obey the following invariants:
  \vspace*{0.8em}
 \begin{enumerate}[leftmargin=2\parindent, align=left, labelsep=-5pt, topsep=0pt,itemsep=-0.25ex,parsep=1.2ex]
	\item[\hspace{-1.5em}\textbf{ Decomposition Invariants:}]
 	\item $R = D V$ where $D$ is the boundary matrix of the filtration $K$
 	\item $V$ is full-rank upper-triangular, and $R$ is \emph{reduced}: $\mathrm{low}_R(i) \neq \mathrm{low}_R(j)$ if its $i$-th and $j$-th columns are nonzero
 	\end{enumerate} 
 \vspace*{1em}
 If these two invariants are maintained, then we call the decomposition \emph{valid}. Given any $R = D V$ decomposition, it was shown in~\cite{zomorodian2005computing} that the following proposition is true:
\begin{proposition}[Persistent Pairing~\cite{zomorodian2005computing}]
	Let $R = D V$, where $R$ is in reduced form and $V$ is full-rank upper triangular (invariants (1) and (2)  hold). Then the pairing $(\sigma_i, \sigma_j)$ defined by $i = \mathrm{low}_R(j)$ is a \emph{creator/destroyer (persistent) pair}. While the matrices $R,V$ satisfying (1) and (2) are not unique in general,  the persistent pairs $(i,j)$ are unique and define the persistence diagrams of $K$.
\end{proposition}
\noindent
Amongst other important corollaries, one immediate consequence of this proposition is that the persistent pairing is independent of the particular entries in $R$ and $V$. When reducing the boundary matrix, we can therefore perform column operations in any order, as long as columns are added from left to right, and so long as every column operation performed in $R$ is also performed on $V$. Once $R$ is reduced, if both of the invariants are respected, then the low entries in $R$ yield the correct pairing encoding the persistence diagram of the corresponding filtration $K$.
% Important corollaries of the persistence pairing lemma is that the 
% The pairing given by the lowest non-zero entries of $R$ is uniqueness, even thought the decomposition itself not.  
% \section*{Computational Perspective} To convey a meaningful computational perspective, the continuous notion of \emph{dynamic persistence} needs to be translated to a discrete, combinatorial definition. Such a description requires a few prerequisite definitions, accompanied by some simplifying assumptions. 
%\subsection*{A General Matrix Framework} Perhaps the most frequently used approach to computing the persistence diagram of a given filtration $K$ is with the \emph{reduction} algorithm. There have been many advancements in the past decade in improving the reduction algorithm---a comprehensive overview is beyond the scope of this effort. The majority of these computational improvements were developed to improve the efficiency of performing the \emph{initial} reduction. A concise summary of these advancements can be found in~\cite{bauer2019ripser}. In contrast, we focus here on improving the efficiency of maintaining a \emph{given} decomposition under an arbitrary reordering.  
\subsection{Vineyards}\label{sec:vineyards}
The original purpose of the vineyards algorithm, as described in~\cite{cohen2006vines}, was to compute a continuous 1-parameter family of persistence diagrams given a time-varying filtration, and to detect homological critical events during the construction.
Notice homological critical events can only occur when the filtration order changes, though not all changes in the filtration order result in homology changes.
It follows that detecting these homological events reduces to computing matrix decompositions at a finite set of time points interleaved between homology-altering changes. 
At the finest scale, changes to the filtration order manifest as adjacent transpositions. 
As a result, a fixed set of rules for maintaining a valid $R = D V$ decomposition under the transposition of adjacent simplices in the filtration is enough to simulate persistence dynamically. 
These rules prescribe certain column and row operations which must be applied to the matrix decomposition either before, during, or after the transposition, in order to ensure decomposition invariants (1) and (2) are respected.

Let $S_{i}^j$ represent the upper-triangular matrix such that multiplying by it on the right is equivalent to adding column $i$ to column $j$. 
Multiplication by $S_{i}^j$ on the left is equivalent to adding row $j$ to row $i$, and thus $S_{i}^j S_{i}^j = I$. 
Similarly, let $P$ denote the permutation matrix so that multiplication from the right applies some permutation $p$ to the columns of some matrix $A$. 
The inverse of any permutation is also its transpose, i.e. $P^{-1} = P^T$, and thus one would write $P A P^T$ to denote the application of the permutation $P$ to both the columns and rows of $A$. In the special case where $P$ represents a transposition, we have $P = P^T$ and may instead simply write $P A P$. 
The goal of the vineyards algorithm can now be described explicitly: to prescribe a set of rules, written as matrices $S_{i}^{j}$, such that if $R = D V$ is a valid decomposition, then $(\ast P \ast R \ast P \ast) = (PDP)(\ast P\ast V \ast P \ast)$ is also a valid decomposition, where $\ast$ is some number (possibly zero) of matrices encoding elementary column or row operations. 
%We refer to the algorithm that maintains the $R = \partial V$ decomposition under adjacent permutations $P$ as the \emph{transposition framework}. 
The explicit pseudocode describing these rules is given in the appendix by Algorithm~\ref{alg:tr}. 
%	\; = 
%	\begin{blockarray}{cccc}
%	\tiny{D_1} \kern-0.2em & a & b & c  \\
%		\begin{block}{c[ccc]}
%  		u \kern-0.1em & \; 1 & 1 & 0 \; \topstrut \\
%  		v \kern-0.1em & \; 0 & 1 & 0 \; \\
%  		w \kern-0.1em & \; 1 & 0 & 0 \; \botstrut \\
%		\end{block}
%	\end{blockarray}	
%	\;\, 
%	+
\\
\\
\noindent 
\textbf{Example 2.2} To illustrate the basic principles on which vineyards works, we re-use the running example introduced in the previous section. Below, we illustrate the case of exchanging simplices $a$ and $b$ in the filtration order. 
\begin{displaymath}
	\def\arraystretch{0.95}
	\makeatletter\setlength\BA@colsep{2.8pt}\makeatother
	%
	\begin{blockarray}{cccc}{R_1}
	 \kern-0.2em & a & b & c  \\
		\begin{block}{c[ccc]}
  		u \kern-0.1em  & \; 1 & 1 &  \; \topstrut \\
  		v \kern-0.1em & \;  & 1 &  \; \\
  		w \kern-0.1em & \; 1 &  &  \; \botstrut \\
		\end{block}
	\end{blockarray}
	\, , 
	\begin{blockarray}{cccc}{V_1}
	\kern-0.2em & a & b & c  \\
		\begin{block}{c[ccc]}
        a \kern-0.1em  & \; 1 & \tikz[baseline]{
            \node[fill=orange!20,anchor=base] (t1)
            {$ 1 $};
        } & 1 \; \topstrut \\
  		b \kern-0.1em & \;  & 1 & 1 \; \\
  		c \kern-0.1em & \;  &  & 1 \; \botstrut \\
		\end{block}
	\end{blockarray}
	\xrightarrow{S_{1}^2} 
	\hspace{-0.5em}
	\begin{blockarray}{cccc}{R_1}
	\kern-0.2em & a & b & c  \\
		\begin{block}{c[ccc]}
  		u \kern-0.1em  & \; 1 &  &  \; \topstrut \\
  		v \kern-0.1em & \;  & 1 &  \; \\
  		w \kern-0.1em & \; 1 & \tikz[baseline]{
            \node[fill=red!20,anchor=base] (t1)
            {$ 1 $};
        } &  \; \botstrut \\
		\end{block}
	\end{blockarray}
	\, , 
	\begin{blockarray}{cccc}{V_1}
	\kern-0.2em & a & b & c  \\
		\begin{block}{c[ccc]}
  		a \kern-0.1em  & \; 1 &  & 1 \; \topstrut \\
  		b \kern-0.1em & \;  & 1 & 1 \; \\
  		c \kern-0.1em & \;  &  & 1 \; \botstrut \\
		\end{block}
	\end{blockarray}
	\xrightarrow{P}
	\hspace{-0.5em}
	\begin{blockarray}{cccc}{R_1}
	\kern-0.2em & b & a & c  \\
		\begin{block}{c[ccc]}
  		u \kern-0.1em  & \;  & 1 &  \; \topstrut \\
  		v \kern-0.1em & \; 1 &  &  \; \\
  		w \kern-0.1em & \; 1 & \tikz[baseline]{
            \node[fill=red!20,anchor=base] (t1)
            {$ 1 $};
        } &  \; \botstrut \\
		\end{block}
	\end{blockarray}
	\, , 
	\begin{blockarray}{cccc}{V_1}
	\kern-0.2em & b & a & c  \\
		\begin{block}{c[ccc]}
  		b \kern-0.1em  & \; 1 &  & 1 \; \topstrut \\
  		a \kern-0.1em & \;  & 1 & 1 \; \\
  		c \kern-0.1em & \;  &  & 1 \; \botstrut \\
		\end{block}
	\end{blockarray}
	\xrightarrow{S_1^2}
	\hspace{-0.5em}
	\begin{blockarray}{cccc}{R_1}
	\kern-0.2em & b & a & c  \\
		\begin{block}{c[ccc]}
  		u \kern-0.1em  & \;  & 1 &  \; \topstrut \\
  		v \kern-0.1em & \; 1 & 1 &  \; \\
  		w \kern-0.1em & \; 1 &  &  \; \botstrut \\
		\end{block}
	\end{blockarray}
	\, ,
	\begin{blockarray}{cccc}{V_1}
	\kern-0.2em & b & a & c  \\
		\begin{block}{c[ccc]}
  		b \kern-0.1em  & \; 1 & 1 & 1 \; \topstrut \\
  		a \kern-0.1em & \;  & 1 & 1 \; \\
  		c \kern-0.1em & \;  &  & 1 \; \botstrut \\
		\end{block}
	\end{blockarray}
\end{displaymath}
Prior to performing the exchange, observe that that the highlighted entry in $V_1$ would render $V_1$ non-upper triangular after the exchange. This entry is removed by a left-to-right column operation, given by applying the $S_1^{2}$ on the right to $R_1$ and $V_1$. After this operation, the permutation may be safely applied to $V_1$. In this case, both before and after the permutation, $R_1$ is rendered non-reduced, requiring another column operation to restore the decomposition to a valid state.

 The time complexity of vineyards is determined entirely by the complexity of performing a single adjacent transposition, and its complexity is often stated in this way. Informally, since column operations are the largest complexity operations needed and each column can have potentially $m$ entries, the complexity of vineyards is $O(m)$ per transposition. 
 However, there are a number of subtleties involved in this statement. 
 The first subtlety is that both the $V$ and $R$ matrices tend to be quite sparse, and thus column and row operations rarely require $m$ operations---in fact, as a rule of thumb, most transpositions require no column operations~\cite{edelsbrunner2000topological}. A second subtlety, due to the necessity of sparse matrix data structures, is that querying the non-zero status of any particular entry typically takes $O(\log m)$ time. This hidden cost may seem minor, however, the quadratic scaling of the number of transpositions vineyards typically requires makes this cost non-trivial.   
In fact, achieving $O(m)$ complexity per transposition requires a special sparse matrix representation that allows swapping any two rows and columns in $O(1)$---see the appendix for more details and discussion. Finally, note the constant factor associated with the stated complexity: inspection of the individual cases of Algorithm~\ref{alg:tr} shows that there are at most two $O(m)$ operations, on both $R$ and $V$, that may be needed for any single transposition.

 % As a rule of thumb, due to the sparsity of both $R$ and $V$, most transpositions actually require no $O(m)$ operations. For example, exchanging two simplices of differing dimensions never requires column operations. A simple analysis given by Chen et al.~\cite{chen2013output} quantifies the order of magnitude we can expect the sparsity of these matrices to be.

\subsection{Moves}\label{sec:moves} Busaryev et al.~\cite{busaryev2010tracking} introduced an algorithm which maintains an $R = D V$ decomposition under \emph{move operations}. A move operation $\mathrm{Move}(i,j)$ is a set of rules for maintaining a valid decomposition under the application of a permutation $P$ that moves a simplex $\sigma_i$ at position $i$ to position $j$. If $j = i \pm 1$, this operation obviously is just an adjacent transposition, and thus in some sense Busaryev's move algorithm is a generalization of the vineyards algorithm.

Though they are similar, move operations confer a few computational advantages compared to vineyards. Informally, a move operation has two convenient properties:
\\	
 \begin{enumerate}[leftmargin=2\parindent, align=left, labelsep=-5pt, topsep=0pt,itemsep=-0.25ex,parsep=1.2ex]
 	\item Querying the non-zero status of entries in any matrix representation occurs once per move.
 	\item The decomposition $R = D V$ is not maintained during the movement of $\sigma_i \mapsto \sigma_j$.
 \end{enumerate} 
 \vspace*{1em}
First, consider property (1). Prior to applying any permutation $P$ to the decomposition, it is necessary to remove non-zero entries in $V$ which render $PVP^T$ non-upper triangular, to maintain invariant (2). 
If one uses the (augmented) sparse matrix representation described in the original vineyards publication~\cite{cohen2006vines}, checking whether these entries are non-zero takes $O(\log m)$ per column. Since moving a simplex from $i$ to $j$ using vineyards performs $\lvert i - j \rvert - 1$ transpositions, this query operation is required at most $O(\lvert i - j \rvert)$ times, which is excessive. If a row-oriented sparse matrix implementation is used, a move operation can perform these non-zero entry checks in just one $O(m)$ pass, prior to performing any column operations.

Property (2) indicates move operations only guarantee the decomposition is valid after the operation completes, implying that the decomposition is not fully maintained during the execution of \textit{Restore}$\ast$. 
This is a significant difference compared to the vineyards algorithm, which was designed specifically to maintain a valid decomposition after each transposition. 
Thus, move operations can be thought of as making a tradeoff in granularity: whereas a sequence of adjacent transpositions $(i, i{+}1), (i{+}1, i{+}2), \dots, (j{-}1, j)$ generates $\lvert i - j \rvert $ valid decompositions in vineyards, a move operation $\mathrm{Move}(i,j)$ generates only one.

\begin{algorithm}
	\caption{Move Right Algorithm}\label{alg:mr}
	\setstretch{1.15}
    \begin{algorithmic}[1]
        \Function{RestoreRight}{$R$, $V$, $\mathbb{I} = \{I_1, I_2, \dots, I_s \}$}
            \State $(\, d_{low}, \, d_R, \, d_V \,) \gets ( \, \mathrm{low}_R(I_1), \, \mathrm{col}_R(I_1), \, \mathrm{col}_V(I_1) \, )$
        	\For{$k$ \textbf{in} $I_2, \dots, I_s$} 
        		\State $( \, d_{low}', \, d_R', \, d_V' \, ) \gets ( \, \mathrm{low}_R(k), \, \mathrm{col}_R(k), \, \mathrm{col}_V(k) \, )$
        		\State $( \, \mathrm{col}_R(k), \, \mathrm{col}_V(k) \, ) \mathrel{+}= (\, d_R, \, d_V \,) $
        		\If {$d_{low}' < d_{low}$}
        			\State $(\, d_{low}, \, d_R, \, d_V \,) \gets (\, d_{low}', \, d_R', \, d_V' \, )$
        		\EndIf
        	\EndFor 
        	\State \Return $(\, R, \, V, \, d_R, \, d_V \, )$
        \EndFunction
    \end{algorithmic}
    \begin{algorithmic}[1]
        \Function{MoveRight}{$R$, $V$, $i$, $j$} 
            \State $\mathbb{I} =$ columns satisfying $V[i, i:j] \neq 0$ %\Comment{$O(m)$}
            \State $\mathbb{J} = $ columns satisfying $\mathrm{low}_R \in [i,j]$ and $\mathrm{row}_R(i) \neq 0$ %\Comment{$O(m)$}
            \State $(\, R, \, V, d_R, d_V \,) \gets $ \Call{RestoreRight}{$R$, $V$, $\mathbb{I}$} 
            %\Comment{Restores $p$-simplices} \Comment{$O(\lvert \mathbb{I} \rvert)$}
            \State $(\, R, \, V \,) \gets $ \Call{RestoreRight}{$R$, $V$, $\mathbb{J}$}
            	%\Comment{Restores $(p+1)$-simplices}
            	%\Comment{$O(m \lvert \mathbb{I} \rvert)$}
            \State $(\, R, \, V\, ) \gets (\, P R P^T, \, P V P^T\,)$  % \Comment{$P$ is a permutation given by (i,j)}
            \State $(\, \mathrm{col}_R(j), \, \mathrm{col}_V(j) \, ) \gets (\, P d_R, \, P d_V \,)$
            \State \Return $(\, R, \, V\,)$
        \EndFunction
    \end{algorithmic}
\end{algorithm}

\begin{algorithm}
	\caption{Move Left Algorithm}\label{alg:ml}
	\setstretch{1.15}
    \begin{algorithmic}[1]
        \Function{RestoreLeft}{$R$, $V$, $\mathbb{K} = \{k_1, k_2, \dots, k_s \}$}
            % \State $\mathrm{low}_K \gets \{ \mathrm{low}_R(k_1), \mathrm{low}_R(k_2), \dots, \mathrm{low}_R(k_s) \}$
            \State $(l, r) \gets $ indices $l, r \in \mathbb{K}$ satisfying $l < r$, $\mathrm{low}_R(l) = \mathrm{low}_R(r)$ maximal 
            \While{$\mathrm{low}_R(l) \neq 0$ \textbf{and} $\mathrm{low}_R(r) \neq 0$}
            	\State $(\mathrm{col}_R(r), \mathrm{col}_V(r)) \mathrel{+}= (\mathrm{col}_R(l), \mathrm{col}_V(l))$
            	\State $\mathbb{K} \gets \mathbb{K} \setminus l$
            	\State $(l, r) \gets $ indices $l, r \in \mathbb{K}$ satisfying $l < r$, $\mathrm{low}_R(l) = \mathrm{low}_R(r)$ maximal
            \EndWhile
        	\State \Return $(\, R, \, V \, )$
        \EndFunction
    \end{algorithmic}

    \begin{algorithmic}[1]
        \Function{MoveLeft}{$R$, $V$, $i$, $j$} 
        	\State $\mathbb{J} \gets \{ j \}$
            % \State $(\mathbb{K}, d_R, d_V) \gets (\emptyset, \mathrm{col}_R(i), \mathrm{col}_V(i))$
            \While{$V(k, i) \neq 0 $\textbf{ for }$ k = \mathrm{low}_V(i)$\textbf{ where }$j \leq k < i$}
            	\State $(\mathrm{col}_R(i), \mathrm{col}_V(i)) \mathrel{+}= (\mathrm{col}_R(k), \mathrm{col}_V(k))$
            	\State $\mathbb{K} \gets \mathbb{K} \cup k + 1 $
            \EndWhile
            \State $(R, V) \gets (P R P^T, P V P^T)$
            \State $\mathbb{J} = $ columns satisfying $\mathrm{low}_R \in [i,j]$ and $\mathrm{row}_R(i) \neq 0$ %\Comment{$O(m)$}

            \State $(R, V) \gets$ \Call{RestoreLeft}{$R, V, \mathbb{K}$}
            \State $(R, V) \gets$ \Call{RestoreLeft}{$R, V, \mathbb{J}$}
            % \State $(\mathrm{col}_R(i), \mathrm{col}_V(i)) \gets (P d_R, P d_V )$
			\State \Return $(\, R, \, V\,)$
        \EndFunction
    \end{algorithmic}

\end{algorithm}

If one has a pair of filtrations $(K_0, K_1)$ each with $m$ simplices, and we assume that each simplex $\sigma_i$ switches its relative ordering with another simplex $\sigma_j$ at most once in some ordered schedule of transpositions, then the number of intermediate persistence diagrams is bounded above by $O(m^2)$ using transpositions, whereas with move operations this bound clearly is at most $O(m)$.  
\\
\\
\noindent \textbf{Example:} We continue the example used in sections~\ref{sec:vineyards} and~\ref{sec:reduction} to illustrate moves. Consider moving the edge $a$ to the position taken by the edge $c$ in the filtration. The donor columns introduced in~\cite{busaryev2010tracking} are shown on the left side of each matrix. Note that using vineyards, the equivalent permutation to the decomposition would require $4$ column operations on both $R_1$ and $V_1$, respectively, whereas a single move operation accomplishes this permutation using only $2$ column operations per matrix. 
\begin{displaymath}
	\def\arraystretch{0.95}
	\makeatletter\setlength\BA@colsep{2.8pt}\makeatother
	%
	\begin{blockarray}{cc}{d_R}
	\kern-0.2em & a \\
		\begin{block}{c[c]}
  		u \kern-0.1em  & \; 1 \; \topstrut \\
  		v \kern-0.1em & \;  \; \\
  		w \kern-0.1em & \; 1 \; \botstrut \\
		\end{block}
	\end{blockarray}
	\; \;
	\begin{blockarray}{cccc}{R_1}
	 \kern-0.2em & a & b & c  \\
		\begin{block}{c[ccc]}
  		u \kern-0.1em  & \; 1 & 1 &  \; \topstrut \\
  		v \kern-0.1em & \;  & 1 &  \; \\
  		w \kern-0.1em & \; 1 &  &  \; \botstrut \\
		\end{block}
	\end{blockarray}
	%
\; \to
	%
\begin{blockarray}{cc}{d_R}
	\kern-0.2em & b \\
		\begin{block}{c[c]}
  		u \kern-0.1em  & \; 1 \; \topstrut \\
  		v \kern-0.1em & \;  1 \; \\
  		w \kern-0.1em & \; \; \botstrut \\
		\end{block}
	\end{blockarray}
	\; \;
\begin{blockarray}{cccc}{R_1}
	\kern-0.2em & a & b & c  \\
		\begin{block}{c[ccc]}
  		u \kern-0.1em  & \; 1 &    &  \; \topstrut \\
  		v \kern-0.1em  & \;    & 1 &  \; \\
  		w \kern-0.1em & \; 1 & 1 &  \; \botstrut \\
		\end{block}
	\end{blockarray}
	%
\; \to
	%
	\begin{blockarray}{cc}{d_R}
	\kern-0.2em & c \\
		\begin{block}{c[c]}
  		u \kern-0.1em  & \; \hphantom{1} \; \topstrut \\
  		v \kern-0.1em & \;  \hphantom{1} \; \\
  		w \kern-0.1em & \; \hphantom{1} \; \botstrut \\
		\end{block}
	\end{blockarray}
	\; \; 
	\begin{blockarray}{cccc}{R_1}
	\kern-0.2em & a & b &  c \\
		\begin{block}{c[ccc]}
  		u \kern-0.1em  & \; 1   &    &   1  \; \topstrut \\
  		v \kern-0.1em & \;       &  1  &  1  \; \\
  		w \kern-0.1em & \; 1   &   1   &     \; \botstrut \\
		\end{block}
	\end{blockarray}
%
\; \xrightarrow{P}
% 
\begin{blockarray}{cc}{d_R}
	\kern-0.2em & c \\
		\begin{block}{c[c]}
  		u \kern-0.1em  & \; \hphantom{1} \; \topstrut \\
  		v \kern-0.1em & \; \hphantom{1} \; \\
  		w \kern-0.1em & \; \hphantom{1} \; \botstrut \\
		\end{block}
	\end{blockarray}
\; \;
\begin{blockarray}{cccc}{R_1}
	\kern-0.2em & b & c & a  \\
		\begin{block}{c[ccc]}
  		u \kern-0.1em  & \;  & 1 &  1\; \topstrut \\
  		v \kern-0.1em & \; 1 & 1 &  \; \\
  		w \kern-0.1em & \; 1 &  & 1 \; \botstrut \\
		\end{block}
	\end{blockarray}
%
\; \xrightarrow{d_R}
% 
	\begin{blockarray}{cccc}{R_1}
	\kern-0.2em & b & c & a  \\
		\begin{block}{c[ccc]}
  		u \kern-0.1em  & \;  & 1 &  \; \topstrut \\
  		v \kern-0.1em & \; 1 & 1 &  \; \\
  		w \kern-0.1em & \; 1 &  &  \; \botstrut \\
		\end{block}
	\end{blockarray}
\end{displaymath}
\vspace{-1.0em}
% Start V example 
\begin{displaymath}
	\def\arraystretch{0.95}
	\makeatletter\setlength\BA@colsep{2.8pt}\makeatother
	%
	\begin{blockarray}{cc}{d_V}
	\kern-0.2em & a \\
		\begin{block}{c[c]}
  		a \kern-0.1em  & \; 1 \; \topstrut \\
  		b \kern-0.1em & \;  \; \\
  		c \kern-0.1em & \;  \; \botstrut \\
		\end{block}
	\end{blockarray}
	\; \;
	\begin{blockarray}{cccc}{V_1}
	 \kern-0.2em & a & b & c  \\
		\begin{block}{c[ccc]}
  		a \kern-0.1em  & \; 1 & 1 &  1\; \topstrut \\
  		b \kern-0.1em & \;  & 1 &  1\; \\
  		c \kern-0.1em & \; &  & 1 \; \botstrut \\
		\end{block}
	\end{blockarray}
	%
\; \to
	%
\begin{blockarray}{cc}{d_V}
	\kern-0.2em & b \\
		\begin{block}{c[c]}
  		a \kern-0.1em  & \; 1 \; \topstrut \\
  		b \kern-0.1em & \;  1 \; \\
  		c \kern-0.1em & \; \; \botstrut \\
		\end{block}
	\end{blockarray}
	\; \;
\begin{blockarray}{cccc}{V_1}
	 \kern-0.2em & a & b & c  \\
		\begin{block}{c[ccc]}
  		a \kern-0.1em  & \; 1 & &  1\; \topstrut \\
  		b \kern-0.1em & \;  & 1 &  1\; \\
  		c \kern-0.1em & \; &   & 1 \; \botstrut \\
		\end{block}
	\end{blockarray}
	%
\; \to
	%
	\begin{blockarray}{cc}{d_V}
	\kern-0.2em & c \\
		\begin{block}{c[c]}
  		a \kern-0.1em  & \; 1 \; \topstrut \\
  		b \kern-0.1em & \;  1 \; \\
  		c \kern-0.1em & \; 1 \; \botstrut \\
		\end{block}
	\end{blockarray}
	\; \;
	\begin{blockarray}{cccc}{V_1}
	\kern-0.2em & a & b &  c \\
		\begin{block}{c[ccc]}
  		a \kern-0.1em  & \; 1 & &   \; \topstrut \\
  		b \kern-0.1em & \;  & 1 &   \; \\
  		c \kern-0.1em & \;  &   & 1 \; \botstrut \\
		\end{block}
	\end{blockarray}
%
\; \xrightarrow{P}
% 
\begin{blockarray}{cc}{d_V}
	\kern-0.2em & c \\
		\begin{block}{c[c]}
  		b \kern-0.1em  & \; 1 \; \topstrut \\
  		c \kern-0.1em & \;  1 \; \\
  		a \kern-0.1em & \; 1 \; \botstrut \\
		\end{block}
	\end{blockarray}
\; \;
\begin{blockarray}{cccc}{V_1}
	\kern-0.2em & b & c & a  \\
		\begin{block}{c[ccc]}
  		b \kern-0.1em  & \; 1 & &\; \topstrut \\
  		c \kern-0.1em & \;  & 1 &  \; \\
  		a \kern-0.1em & \; &  & 1 \; \botstrut \\
		\end{block}
	\end{blockarray}
%
\; \xrightarrow{d_V}
% 
	\begin{blockarray}{cccc}{V_1}
	\kern-0.2em & b & c & a  \\
		\begin{block}{c[ccc]}
  		b \kern-0.1em  & \; 1 & & 1 \; \topstrut \\
  		c \kern-0.1em & \; & 1 &  1 \; \\
  		a \kern-0.1em & \; &  & 1  \; \botstrut \\
		\end{block}
	\end{blockarray}
\end{displaymath}
\\
\noindent
We recall an important claim given by~\cite{busaryev2010tracking} on the effect that move operations have on the status of simplices in the pairing. Recall from section~\ref{sec:background} that simplices which create new homology classes are called \emph{creators} and simplices that destroy homology classes are called \emph{destroyers}.
\begin{proposition}[Locality of moves~\cite{busaryev2010tracking}]
	Other than the simplex $\sigma_i$ being moved, $\mathrm{MoveRight}(i,j)$ cannot change a destroyer to a creator, $\mathrm{MoveLeft}(i,j)$ cannot change a creator to destroyer, and neither move operation changes the status of any simplex in $K$ outside of the interval $[i,j]$.  
\end{proposition}
\noindent The effect of the movement on intermediate simplices depends on the direction of the movement. If $i < j$ (respectively, $j < i$), all simplices at positions $k \in [i,j] \smallsetminus i$ are shifted down (respectively, up) by $1$. 
The pseudo-code for the \emph{MoveRight} operation is given in Algorithm~\ref{alg:mr}, and the pseudo-code for the \emph{MoveLeft} is given in Algorithm~\ref{alg:ml}. 
We note that while the pseudocode for \emph{MoveRight} is included in~\cite{busaryev2010tracking}, the one for \emph{MoveLeft} is not, and Algorithm~\ref{alg:ml} is the one we propose and leverage in this paper.

Although conceptually similar, there is an asymmetry between \emph{MoveRight} and \emph{MoveLeft}. Moving a simplex upwards in the filtration requires removing non-zero entries along several columns of a particular row in $V$ so that the corresponding permutation does not render $V$ non-upper triangular. The key insight of the algorithm presented in~\cite{busaryev2010tracking} is that $R$ can actually be maintained in all but one column during this procedure (by employing the \emph{donor} column). In contrast, moving a simplex to an earlier time in the filtration requires removing non-zero entries along several rows of a particular column of $V$. As before, $R$ stays reduced during this cancellation procedure in all but one column, however the subsequent permutation to $R$ requires reducing a pair of columns which may cascade into a larger chain of column operations to keep $R$ reduced. Since these operations always occur in a left-to-right fashion, its not immediately clear how to apply a donor column kind of concept. Fortunately, we can still separate out these two stages. 
% with a logarithmic loss in efficiency, although executing the movement has a higher constant factor. Nonetheless, we can still move a simplex arbitrarily within the given filtration with such a procedure,  which when paired with the augmented sparse matrix structure from~\cite{} implies one can move simplices within the filtration with just O(|i - j|) time permutation, as opposed the |i - j| separate O(1) permutations. Moreover, the ability to move a simplex earlier in the filtration enables a key preprocessing step which we introduce next which, as we shall prove, may greatly reduce the number of times the decomposition need be maintained.

%% TODO: prove R stays reduced by re-reducing only columns in K: should be easy, first show that $R$ is reduced un until time T = i - delta, where delta = | i - j |. Then show that that the only columns in [ j, j + delta ] which have low entries that change are precisely the ones in R[i:j, i]
%% TODO: sparsity of boundary matrix \approx O(n \log n)
%% TODO: mention Morozov note on how reduction requires \Omega(n^3), which is tight

\section{Our contribution: Move Schedules}\label{sec:move_schedules}
%\subsection{Overview}
%In this section, we introduce our modification to the moves algorithm, deferring more detailed discussions or proofs to the appendix. 
%First, we give a tight lower bound on the number of moves needed to perform an arbitrary permutation to the $R = \partial V$ decomposition. We also give bounds in expectation, as well as efficient algorithms for achieving these bounds; the proof of optimality for the former comes directly by reduction to the permutation edit distance problem, and the latter comes from a reduction to the Longest Increasing Subsequence (LIS) problem. 
%
%Thus, the output of the first phase is a tight bound $d \leq m$ and an efficient algorithm for generating sequences of move permutations $( s_1, s_2, \dots, s_d )$, which we call \emph{schedules}. In the second phase, we investigate the feasibility of optimizing the \emph{cost} of these schedules directly, measured in terms of column operations. We establish two negative results: the first result indicates that obtaining the cost of a given schedule is about as expensive as executing the schedule, precluding the use of  combinatorial optimization techniques like dynamic programming; our second result investigates the feasibility of an optimal step strategy, to which we give a counter example demonstrating a greedy-type approach can lead to arbitrarily bad behavior. In light of these results, we give an alternative proxy-objective to minimize instead and provide bounds justifying its relevance to the original objective. We conclude this section by giving a greedy strategy for minimizing this proxy objective that performs well in practice, along with an efficient algorithm.

\subsection{Proposed approach}
Let us begin by briefly describing the overall pipeline of our proposed approach, which is outlined in Algorithm~\ref{alg:schedule}. As before, we assume as input  a discrete $1$-parameter family of filtrations $\mathcal{K} = (K_1, K_2, \dots, K_n)$, 
and the goal being to compute a persistence diagram for each filtration.

Without loss of generality, we may assume $\lvert K_i \rvert = m$ for every $i \in [n]$, 
such that there are  bijections $f_i : K_i \to K_{i+1}$. 
Otherwise, if $\lvert K_i \rvert \neq \lvert K_{i+1} \rvert$,   
we may append $K_{i+1} \smallsetminus K_i$ to the end of $K_i$ and likewise append $K_{i} \smallsetminus K_{i+1}$ to the end of $K_{i+1}$ prior to constructing the bijection. 
At the matrix level, both concatenations can be achieved by appending the boundary chains of the simplices being inserted to $R$, and then running Algorithm~\ref{alg:reduce} only on the appended columns. 
Each $f_i$ induces a bijection $f_i^\ast : [m] \to [m]$, or equivalently, a permutation of the index set $[m]$. For each pair of filtrations $(K_i, K_{i+1})$, we assign labels to the simplices of $K_i$ using the index set $[m]$ and relabel $K_{i+1}$ accordingly using $f_i$. Denote these permutations by $p$ and $q$. 
We compute the longest increasing subsequence (LIS) of $q$ and use this subsequence to recover a longest common subsequence (LCS) $\mathrm{LCS}(p,q)$. We pass $p$, $q$, and this LCS to our scheduling algorithm, which returns as output an ordered set of move permutations. We collect all of these permutations into a single schedule $\mathcal{S}$, obtain an $R=DV$ decomposition for the first filtration $K_1$, and then execute the permutations stored in $\mathcal{S}$ in sequence. During execution, we obtain a valid decomposition for each filtration $K_1, K_2, \dots, K_n$. 
It turns out that not all valid schedules are created equal:
some incur a much larger number of column operations,
even if they have a minimal number of moves. 
Minimizing schedule costs will be addressed in 
Section~\ref{sec:schedule_cost}.

\begin{algorithm}[h]
	\caption{Scheduling algorithm}\label{alg:schedule}
	\setstretch{1.05}
    \begin{algorithmic}[1]
    	\Require Ordered set of filtrations $\mathcal{K}$, each of size $m$, with bijections $f_{i}: K_i \to K_{i+1}$
    	\Ensure Valid $R = D V$ decompositions are computed for each filtration $K_1, K_2, \dots, K_n$
    	\Procedure{SparseSchedule}{$\mathcal{K} = ( K_1, K_2, \dots, K_n )$} 
    		\State $\mathcal{S} = \emptyset$
    		\For{$i = 1$\textbf{ to }$n-1$}
    		    \State $(p, q) \gets ([m], \, \mathrm{Im}\, f^\ast_{i})$ 
    			\State $\mathrm{lis}_q \gets \mathrm{LIS}(q)$ \Comment{$O(m \log \log m)$}
    			\State $\mathcal{S} \gets \mathcal{S} \cup \mathrm{GreedySchedule}(p, q, \mathrm{lis}_q)$ \Comment{$O(d\log d)$, see section~\ref{sec:proxy_objective}}
    		\EndFor
	    	\State $(R, V) \gets$ \Call{Reduction}{$D = \partial K_1$}
	    	\For{$(i,j)$\textbf{ in }$\mathcal{S}$}
	    		\If{$i < j$}
	    			\State $(R, V) \gets$ \Call{MoveRight}{i, j}
	    		\Else
	    			\State $(R, V) \gets$ \Call{MoveLeft}{i, j}
	    		\EndIf
	    	\EndFor
    	\EndProcedure
	\end{algorithmic}
\end{algorithm}
\noindent Algorithm~\ref{alg:schedule} is purely illustrative at this point, 
and is meant to serve as a guidepost before discussing its other components in depth. 
Specifically, lines (4-5) are discussed next in section~\ref{sec:schedule_sizes}, and the algorithm for line (6) is given in section~\ref{sec:schedule_cost}. 
Note that the loop starting at line (3) in Algorithm~\ref{alg:schedule} could always be merged with the loop starting at line (8), i.e. the individual schedules $S_i$ may be computed on the fly. Due to the combinatorial aspect of our approach, there is actually no need to have access to---or explicitly store---the entire family of filtrations and the schedules between them. The entire approach can be performed online, keeping at most two filtrations and one decomposition in memory at any given time.

\subsection{Minimizing schedule size}\label{sec:schedule_sizes}
Here we describe the reasoning for lines (4-5) in Algorithm~\ref{alg:schedule}. Recall in section~\ref{sec:motivation}, an example was given showing the vineyards algorithm efficiency compared to a naïve approach at three varying levels of coarseness. It was hypothesized that the reason the vineyards algorithm was more expensive than the naïve approach is due to the extra overhead of maintaining the decomposition at each transposition. Thus, decreasing the number of times the decomposition is restored to a valid state ought to reduce the total number of column operations needed to update a decomposition. This motivates the following question: 
in order to apply an \emph{arbitrary} permutation $P$ to a given $R = DV$ decomposition, 
what is the minimal number of times the decomposition needs to be restored to a valid state?

\subsubsection{The Continuous Setting}\label{sec:continuous_setting}
In the vineyards case, we are given a homotopy $F(x, \tau) = \mathbb{X} \times [0,1] \to \mathbb{R}$ and a pair of filtrations $(K_0, K_1)$ whose simplices are ordered according to $f_0(x) = F(x, 0)$ and $f_1(x) = F(x, 1)$, respectively. The choice of homotopy $F$ completely determines the number of adjacent transpositions, so to get a meaningful bound we must know something about $F$. If we assume the curves traced by the homotopy are in general position and we assume each pair of curves cross each other at most once, then the set of adjacent transpositions $S_F$ that $F$ decomposes into is given by the pairs of simplices which change in their relative ordering:
\begin{align}\label{eq:sf_schedule}
	S_F &= \{ (i,j) \, | \, f_0(\sigma_i) < f_0(\sigma_j) \land  f_1(\sigma_i) > f_1(\sigma_j) \quad \forall \; i,j \in [m] \} \\
	&= ( s_1, s_2, \dots, s_k )
\end{align}
We assume, without loss of generality, that the underlying complexes $K_0$ and $K_1$ have the same cardinality. 
Since we deal exclusively with essential filtrations, we may think of $K_1$ as simply a reordering of $K_0$ (or vice versa). Let $p = [m]$ and let $q = \mathrm{Im}\,f^\ast$, where $f^\ast : [m] \to [m]$ is the one-to-one correspondence induced by the bijection $f: K_0 \to K_1$. Then $S_F$ is given by inversions between $p$ and $q$:
$$ \mathrm{Inv}(p, q) = \{ \, (i,j) \, | \, p(i) < p(j) \land q(i) > q(j)  \, \}$$
The associated distance, measuring $\lvert S_F \rvert$, is the well-known \emph{Kendall}-$\tau$ distance $K_\tau(p,q)$: 
\begin{equation}\label{eq:kendall_dist}
	K_\tau(p, q) = \lvert \, \mathrm{Inv}(p, q) \, \rvert
\end{equation}
Geometrically, our assumptions on $F$ defines a class of $x$-monotone curves called \emph{pseudo-segments}, i.e. parameterized curves whose behavior with respect to a certain restricted set of geometric predicates is invariant. This family includes e.g. the straight-line homotopy $f_\tau(\sigma) = (1 - \tau) f_0(\sigma) + \tau f_1(\sigma)$, which was studied in the original vineyards paper~\cite{cohen2006vines}. Detecting all $k$ intersections of $m$ pseudo-segments is a well-studied problem in computational geometry that can be optimally solved in output-sensitive $O(m \log m + k)$ time by several algorithms~\cite{boissonnat2000efficient}, where $k$ is the output-sensitive term. 
Since $k \sim O(m^2)$ in the worst case, achieved when $f_1 = - f_0$, we have the following upper bound on $\lvert S_F \rvert$
\begin{equation}\label{eq:sf_m2}
	\lvert S_F \rvert \approx O(m^2)
\end{equation}   
 This bound is exhibited in the grayscale image data example in section~\ref{sec:motivation}: each 9x9 patch contains $(81, 208, 128)$ simplices of dimensions $(0, 1, 2)$ yielding $m=417$ simplices in total which, as the homotopy is simulated, generate $\approx 70,\!000$ transpositions. The upper bound size on $\lvert S_F \rvert$ is thus $\binom{417}{2} \approx 87,\!000$, which nearly occurs in this case, since the last filtration is nearly the reverse of the filtration at the beginning.   
 This quadratic scaling can induce a number of issues in practical implementations of vineyards, e.g. although many transpositions require $0$ column reductions in practice, detecting the existence of non-zero entries at specific positions in the sparse matrix takes $O(\log(m))$ time. Since this check is required after each permutations are applied, it is in some sense unavoidable in the vineyards algorithm.

The ordered set of inversions $(i,j) \in S_F$ can be interpreted as permutations to apply to the decomposition. We call $S_F$ a \emph{schedule} with respect to $F$. If our goal is to decrease the size of $\lvert S_F \rvert$, one option is to \emph{coarsen} $S_F$ by collapsing contiguous sequences of adjacent transpositions to moves, via the map:
\begin{equation}\label{eq:tr_to_mv}
	 (i, i+1)(i+1, i+2)\cdots(j-1, j) \mapsto (i, i+1, \cdots, j-1, j)  \quad \text{ if } i < j
\end{equation}
We expect a coarsened schedule to be cheaper to execute than the original schedule it was derived from, with the tradeoff being that less diagrams are produced, which we now prove
\begin{proposition}\label{prop:factor2}
	For some given decomposition $R = D V$ where each matrix is $(m \times m)$, let $C$ denote the number of $O(m)$ operations required to execute schedule $S  = (t_1, t_2, \dots, t_h)$, and similarly let $L$ be the number of $O(m)$ operations required to execute a coarsened schedule $\widetilde{S}$ of $S$. We have the following relationship between $C$ and $L$:
	$$ \frac{C}{2} \leq L \leq C$$ 
\end{proposition}
\begin{proof}
	Consider Algorithms~\ref{alg:tr} and~\ref{alg:mr}, where it's assumed Algorithms~\ref{alg:tr} is used to execute a given schedule $S$ and Algorithms~\ref{alg:mr} is used to execute its coarsened schedule $\widetilde{S}$. If maps are created providing $O(1)$ access to the low entries needed by lines (5) and (12), and the non-zero entries in each column are sorted according to some fixed order providing $O(\log(m))$ time for lines (5), (11), (17), and (20), then the dominating cost of each transposition in Algorithm~\ref{alg:tr} are the column operations, each of which take $O(m)$ time. There are at most $2$ column operations for any given case. For any contiguous sequence of transpositions $(t_i, t_{i{+}1}, \dots, t_j)$, this implies we have at most $2\lvert i - j \rvert$ operations each requiring $O(m)$ time. Now consider a single move operation, Move($i$,$j$), as a replacement for the sequence of transpositions above. Identifying non-zero entries in the matrix occurs once, which takes $O(\lvert i - j \rvert \log(m))$ time. If we maintain a map providing low entries in $O(1)$ time, since permutations take $O(1)$ time and lines (7) and (7) in both Restore* and Move* takes just $O(1)$ via pointer swapping, the dominant cost again are the column operations (line 5). Since a move operation requires at most a single column operation for each index between $[i, j]$, the complexity of $\mathrm{Move}(i,j)$ is bounded above by $O(\lvert i - j \rvert m)$, and the claimed inequality follows. 
\end{proof}
In terms of bounds, we clearly have $\lvert \widetilde{S}_F \rvert \leq \lvert S_F \rvert$, and the associated coarsened schedule $\widetilde{S}_F$ can be computed in $O(m)$ time, so this is a viable approach to improving the efficiency of the vineyards algorithm. However, the coarsening depends entirely on the choice of $F$ and the upper bound from~\ref{eq:sf_m2} remains, as its always possible that there are no contiguous subsequences to collapse. If a straight line homotopy is used, the number of collapsible subsequences may depend on the simplices categorization as critical points; transpositions between pairs of simplices which trigger a changes in homology tend to be more expensive. Cohen-Steiner~\cite{cohen2006vines} referred to these as \emph{switches}, see~\cite{edelsbrunner2000topological} for more details.

\subsubsection{The Discrete Setting} 
Since moves do not maintain a valid decomposition during movement, it is worth asking if we can achieve better bounds than vineyards, by removing the assumption that the goal is to simulate persistence across a supplied homotopy $F(x, \tau) = \mathbb{X} \times [0,1] \to \mathbb{R}$. In the case of vineyards, the bound is clear: whether assuming a homotopy between a pair of filtrations $(K_0, K_1)$ or not, if $\kappa = K_\tau(K_0, K_1)$, then $\Omega(\kappa)$ adjacent transpositions are required to transform $K_0 \mapsto K_1$. 
Allowing move operations in any order, we have $O(m)$ as a trivial upper bound: simply move each simplex in $K_0$ into its position in the filtration given by $K_1$ in the order given by the latter. However, it's not immediately clear whether this bound is tight. To establish this formally, we require a few definitions.

 A \emph{permutation}  is a bijection of a set onto itself. The symmetric group $S_n$ is the set of all permutations of  $[n]$ under function composition $\circ$, applied from right to left. Given two fixed permutations $p, q \in S_n$ and a set $S \subseteq S_n$, a few goals that appear commonly in the literature are the following:
\begin{enumerate}
	\item Find an ordered sequence of permutations $s_1, s_2, \dots, s_d$ from $S$ whose composition transforms $p$ into $q$:
	\begin{displaymath}
		s_d \circ \cdots \circ s_2 \circ s_1 \circ p = q
		% p \sbullet s_1 \sbullet s_2 \sbullet \cdots \sbullet s_d = q
	\end{displaymath}
	\item Find a sequence satisfying (1) of minimal length ($d$)
	\item Find the minimal length $d$, referred to as the \emph{distance} between $p$ and $q$ with respect to $S$. 
\end{enumerate}
A sequence $s_1, s_2, \dots, s_d \in S \subseteq S_n$ of operations mapping $p \mapsto q$ is sometimes called a \emph{sorting of $p$} in the literature. In the context of performing updates to the $R = D V$ decomposition, we call such sequences \emph{schedules} . There are typically many ways to solve (1), and any solution to (2) necessarily solves (1) and (3). When $p, q$ are interpreted as strings, distances defined with respect to a fixed $S \subseteq S_n$ are commonly referred to as \emph{edit distances}~\cite{labarre2013lower}, which are denoted as $d_S(\cdot, \cdot)$. Note that here we describe distances between permutations (e.g. $p,q \in S_n$) using operations that can themselves be expressed as permutations (e.g. $s \in S \subseteq S_n$), and that the choice of $S$ defines the universe of operations and thus the type of edit distance being measured---otherwise if $S = S_n$, then $d_S(p, q) = 1$ for any $p, q \in S_n$ and the problem is trivial!

Many types of edit distances can be described succinctly. Let $\beta(i,j,k,l)$ with $1 \leq i < j \leq k < l \leq n$ denote the permutation that exchanges the two closed intervals given by $i, j-1$ and $k,l-1$:
\[
\scalefont{1.0}{
\setcounter{MaxMatrixCols}{20}
\left (
\begin{array}{*3c|*4c|*4c|*4c|*{5}c } 
\cline{4-7} \cline{12-15}  \rule{0pt}{2ex}
1 & \cdots & i -1 & i & i + 1 & \cdots & j - 1 & j & j +1 & \cdots & k - 1  & k & k+1 & \cdots & l - 1 & l & l + 1 & \cdots & n \\
\cline{4-7} \cline{12-15} 
\cline{4-7} \cline{12-15}  \rule{0pt}{2ex}
1 & \cdots & i -1 & k & k+1 & \cdots & l - 1 & j & j +1 & \cdots & k - 1  & i & i + 1 & \cdots & j - 1 & l & l + 1 & \cdots & n \\
\cline{4-7} \cline{12-15} 
\end{array}
\right ) 
}
\]
Common edit distances found in the literature expressed as specializations of $\beta$ are given below:
\begin{enumerate}
	\item (no restriction on $i,j,k,l$) $\implies$ $\beta$ is called a \emph{block interchange}
	\item $j = k$ $\implies$ $\beta$ swaps adjacent intervals, called a \emph{transposition}
	\item $j = i+1$ and $l = k+1$ $\implies$ $\beta$ swaps two not necessarily adjacent elements, called an \emph{exchange}\footnote{In comparative genomics, exchanging two elements is sometimes called a \emph{super short transposition}.}
\end{enumerate}
Restricting the operations given above further to the case where $i = 1$ gives the so-called \emph{prefix} versions of the described distances. 
Although each edit distance can be described using common notation, both the difficulty of obtaining a minimal sorting and the size $d_S(p,q)$ varies dramatically in the choice of $S$. For example, while sorting by transpositions and reversals is NP-hard and sorting by prefix transpositions is unknown, there are polynomial time algorithms for sorting by block interchanges, exchanges, and prefix exchanges~\cite{labarre2013lower}. Sorting by adjacent transpositions can be achieved in many ways: any sorting algorithm that exchanges two adjacent elements during its execution (e.g. bubble sort, insertion sort, merge sort) yields a sorting of size $K_\tau(p, q)$.

Here we consider sorting by move permutations. 
A \emph{move} on the set $[m]$ is 
a permutation given by a pair $(i,j)$, 
and its effect is to cyclically rotate 
the interval $[i,j]$. For example, if $i < j$, this permutation takes the form: 
\[
\scalefont{1.0}{
\setcounter{MaxMatrixCols}{20}
\left (
\begin{array}{*3c | *5c | *{3}c } 
\cline{4-8} \rule{0pt}{2ex}
1 & \cdots & i -1 & i & i + 1 & \cdots & j - 1 & j & j +1 &  \cdots & m \\
\cline{4-8}
\cline{4-8}\rule{0pt}{2ex}
1 & \cdots & i -1 & i + 1 & \cdots & j - 1 & j & i & j +1 &  \cdots & m \\
\cline{4-8}
\end{array}
\right ) 
}
\]
The complexity of sorting by moves is related to the well-known \emph{Levenshtein distance}~\cite{bar2004approximating}, also known as the edit distance on strings. The Levenshtein distance between two strings $A$ and $B$ is the minimum number of character insertions, deletions, and substitutions needed to transform one string into another. 
Through a dynamic programming type approach, the Wagner-Fischer algorithm can compute the Levenstein distance in $O(m^2)$, and there are several sub-quadratic approximations available~\cite{bar2004approximating} A related edit distance is the \emph{LCS} distance, which is the minimum number of character insertions or deletions needed to transform one set of symbols into another. 
%In terms of hardness, there is evidence to suggest that the lower bound on computing Levenshtein distance in full generality is near quadratic time.

Like the \emph{LCS} distance, sorting by move operations can be interpreted as finding a minimal sequence of edit operations where the only operations allowed are insertions and deletions. If these operations each take uniform cost and the universe of strings is restricted to \emph{permutations}, the corresponding edit distance has a special name called the \emph{permutation edit distance}. We obtain the following lower bound size of a sorting (i.e., a schedule) using moves.

\begin{proposition}[Schedule size]
Given a pair of filtrations $(K_0, K_1)$ and a one-to-one correspondence $f: K_0 \to K_1$, the coarsest move schedule $S^\ast$ is of size $\Omega(d)$, where:
$$ d = m - \lvert \mathrm{LCS}(K_0, K_1) \rvert $$ 
Moreover, any size $d$ schedule $S$ can be computed in $O(m \log \log m)$ time. 
\end{proposition}
\begin{proof}
Recall our definition of  edit distance given above, 
depending on the choice of $S \subseteq S_m$ of allowable edit operations, and that in order for any edit distance to be symmetric, if $s \in S$ then $s^{-1} \in S$. This implies that $d_S(p,q) = d_S(p^{-1}, q)$ for any choice of  $p,q \in S_m$. 
Moreover, edit distances are \emph{left-invariant}, i.e.
\[
d_S(p,q) = d_S(r \circ p, r \circ q) \quad \text{ for all } p,q,r \in S_m
\]
Conceptually, left-invariance implies that the edit distance between any pair of permutations $p,q$ is invariant under an arbitrary relabeling $p,q$--as long as the relabeling is consistent. Thus, the following identity always holds: 
$$ d_S(p,q) = d_S(\iota, p^{-1} \circ q) = d_S(q^{-1} \circ p, \iota) $$
where $\iota = [m]$, the identity permutation. Suppose we are given permutations $p = K_0, q = K_1$, and we seek to compute $\mathrm{LCS}(p, q)$. Consider the permutation $r = p^{-1} \circ q$. Since the LCS distance is a valid edit distance, if $\lvert \mathrm{LCS}(p, q) \rvert = k$, then $\lvert \mathrm{LCS}(\iota, r) \rvert = k$ as well. Notice that $\iota$ is strictly increasing--thus any common subsequence $\iota$ has with $r$ must also be strictly increasing. Thus, the problem of computing $\mathrm{LCS}(p, q)$ reduces to the problem of computing the \emph{longest increasing subsequence} (LIS) of $r$, which can done in $O( m \log \log m)$ time~\cite{bespamyatnikh2000enumerating}. The optimality of $d$ follows from the optimality of the well-studied LCS problem~\cite{kumar1987linear}. 
\end{proof}

\begin{corollary}\label{cor:expectation}
	If $K_0, K_1$ are random distinct filtrations each of size $m$, then $d$ is upper bounded by $O(m - \sqrt{m})$ in expectation, with probability $1$ as $m \to \infty$.
\end{corollary}
\begin{proof}
	The proof of this result reduces to showing the average length of the LIS for random permutations. Let $L(p) \in [1,m]$ denote the maximal length of a increasing subsequence of $p \in S_m$. 
	The essential quantity to show the expected length of $L(p)$ over all permutations: 
	$$ \ell_m = \mathbb{E} \, L(p) = \frac{1}{m!} \sum\limits_{p \in S_m} L(p)$$
	The history of the mathematics of estimating this quantity dates back at least 50 years, and is sometimes called the \emph{Ulam-Hammersley} problem. It was shown by Baik et al.~\cite{} that as $m \to \infty$:
	$$ \displaystyle \ell_m = 2 \sqrt{m} + c m^{1/6} + o(m^{1/6}) $$
where $c \approx -1.77108$. As a result, we have: 
$$ \frac{\ell_m}{\sqrt{m}} \to 2\rlap{\quad \text{as } $m \to \infty$} $$	 
Thus, if $p \in S_m$ denotes a uniformly random permutation in $S_m$, then $L(p)/\sqrt{m} \to 2$ in probability as $m \to \infty$. 
Since the schedule size is defined as $d = m - \lvert \mathrm{LCS}(K_0, K_1) \rvert$, the claimed bound follows.
\end{proof}
\noindent The upper bound from Corollary~\ref{cor:expectation} captures the size of $S^\ast$ under worst-case conditions. It is worth noting that this bound applies to pairs of uniformly sampled permutations, as opposed to uniformly sampled distinct filtrations---intuitively, simplicial filtrations have more structure due to the subcomplex relation that must be respected in the filtration. Thus, while $O(m - \sqrt{m})$ in expectation is an improvement over $O(m)$ it is still not tight. However, Boissonnat~\cite{boissonnat2018efficient} has shown that for any $d$-dimensional simplicial complex $K$ with $m$ simplices, the number of distinct filtrations built from $K$ is \emph{at least} $\floor{\frac{t+1}{d+1}}^m$, where $t$ is the number of distinct filtration values associated with the complex $K$. Since this lower bound grows similarly to $m!$ when $t \approx O(m)$ and $d << m$ fixed, we expect $O(m - \sqrt{m})$ is not too pessimistic of an upper bound when the pair of filtrations are truly random. In practice, when one has a time-varying filtration and the sampling points are relatively close [in time], the LCS between adjacent filtrations is expected to be much larger, subsequently making $d$ much smaller. 
This had been confirmed empirically, e.g. the average size of the LCS across the 10 filtrations from Section~\ref{sec:motivation} each having $417$ simplices sampled at evenly spaced time points was $343$, implying $d \approx 70$ permutations needed on average to update the decomposition between adjacent time points.

\subsection{Minimizing schedule cost}\label{sec:schedule_cost}
\subsubsection{Motivation}
The results from the previous section yield a sufficient algorithm for generating move schedules of minimal cardinality: simply compute a $\mathrm{LCS}(K_0, K_1)$ for a pair $(K_0, K_1)$ and fix an order of simplices in the set $\mathrm{LCS}(K_0, K_1) \smallsetminus K_0$ to move. Any sequence of moves which increases the LCS on each move is guaranteed to have size $m - \lvert \mathrm{LCS}(K_0, K_1) \rvert$, and the reduction to the permutation edit distance problem ensures this size is optimal. 
However, like in the vineyards algorithm, certain pairs of simplices cost more to exchange depending on whether they are critical pairs in the sense described~\cite{cohen2006vines}, making specific move operations more expensive than others. 
%This variation in cost exaggerates the variability of the cost of scheduling move operations.

One advantage of Algorithm~\ref{alg:mr} is that the cost of applying a permutation to some given $R = D V$ decomposition can be determined efficiently prior to performing any column operations. 
Specifically, the cost of determining $\lvert \mathbb{I}\rvert + \lvert \mathbb{J}\rvert$ in lines (2) and (3) of Algorithm~\ref{alg:mr} cannot take more than $O(m)$ time. If we assume a similar complexity to determine the cost of applying a permutation with Algorithm~\ref{alg:ml} and we know to execute $d$ moves, an initial thought might be to use an optimization method like dynamic programming. 
That is, suppose $O(dm)$ move costs are computed ahead of time and these costs are stored in a table providing $O(1)$ element access and modification. Using a sparse row representation for $V$, this cannot take more than $O(dm)$ time, which is relatively efficient when $d << m$. With this structure, determining the cost of the cheapest move takes just $O(d)$ time. However, moving a simplex from position $i$ to position $j$ may require updating $O(\kappa^2)$ entries in the structure, where $\kappa = \lvert i - j \rvert$. Since $\kappa \sim O(m)$ in the worst case, direct optimization may require upwards of $\approx O(dm^2)$. Since this time complexity is potentially more expensive than executing a schedule to begin with, this direct strategy does not seem viable for practical settings.

% Specifically, if a compressed sparse row (CSR) representation is used for $V$ and a compressed sparse column (CSC) representation is used for $R$, then determining $\lvert \mathbb{I}\rvert + \lvert \mathbb{J}\rvert$ takes $O(\lvert \mathbb{I}\rvert \log m + \lvert \mathbb{J}\rvert \log m ) \approx O(m)$ time.

% To explain this discrepancy, fixed any two permutations $p,q \in S^n$ and consider a few different sorting strategies for transforming $p \mapsto q$: linear homotopy $S_{lh}$, insertion sort $S_{ins}$, selection sort $S_{sel}$, the coarsened version of selection sort $\tilde{S}_{sel}$, and a schedule derived from $\mathrm{LCS}(p,q)$, $S_{lcs}$. To give some context for these choices, $S_{lh}$ is the strategy used in~\cite{cohen2006vines}, $S_{ins}$ the strategy used in~\cite{lesnick2015interactive}, $\tilde{S}_{sel}$ the strategy used in~\cite{busaryev2010tracking}, and $S_{lcs}$ the strategy discussed above.

%Also, let $\mathbf{s}_i =  s_i \circ \cdots \circ s_2 \circ s_1 \circ p$ denote the composition of the first $i$ permutations for some given schedule $S_{\ast} = (s_1, \dots, s_d)$, where $i < d$. 
%The following identities are true: 
%\begin{align*}
%	K_\tau(p, q) &= \sum\limits_{i = 1}^{d-1} K_\tau(\boldsymbol{s}_i, \boldsymbol{s}_{i+1}) 
%	% = \sum\limits_{i = 1}^{d-1} K_\tau(\tilde{\boldsymbol{s}}_i, \tilde{\boldsymbol{s}}_{i+1})
%%	&= \lvert S_{lh} \rvert = \lvert S_{ins} \rvert = \lvert S_{sel} \rvert \\
%%	&= \sum\limits_{i}^{d-1} \lvert \tilde{S}_{sel} \rvert \\
%%	&\leq \sum\limits_{i}^{d-1} \lvert S_{lcs} \rvert
%\end{align*}
%Define $\Delta(S) = \sum\limits_{i = 1}^{d-1} K_\tau(\boldsymbol{s}_i, \boldsymbol{s}_{i+1})  $

%A famous result from Diaconis et al.~\cite{} is that following inequality relating the Spearman footrule distance and the Kendall-$\tau$ distances:
%$$ K(\sigma, \tau) \leq F(\sigma, \tau) \leq 2 K(\sigma, \tau)$$
%The Kendall-$\tau$ distance between two permutations of size $n$ can be computed in $O(n \log n)$, while the Spearman footrule distance can be computed in $O(n)$ time.

% However, in terms of cost, moves are not commutative and thus the choice of order affects the total cost of the schedule. That is, since the cost of a move is depends on the non-zero entries in the decomposition, and because the decomposition is not unique, we cannot simply use e.g. a weighted permutation edit distance.  
%After all, at least with $\mathrm{MoveRight}$, the cost of a move determined ahead of time by the sizes of $\mathbb{I}$ and $\mathbb{J}$ in Algorithm~\ref{alg:mr}, both of which may be computed in just $O(m)$ time. If we assume a similar complexity to determine the cost of   $\mathrm{MoveLeft}$ and we need to execute $d$ moves, a greedy strategy would take $O(m \omega) \approx O(m d^2 )$ to determine, where $\omega$ is the $d^{th}$ triangular number, or $\approx O(m)$ if $d << m$, which is often the case.

\subsubsection{Greedy approach}\label{sec:greedy} If an optimization procedure like dynamic programming is too expensive, one would hope that a greedy solution which chooses the minimal cost choice at every step could lead be an efficient strategy.  We give a counter-example below demonstrating that a greedy procedure may lead to arbitrarily bad behavior. 
\\
\\
\textbf{Counter-Example} A pair of filtrations is given below, each filtration contains $m = 10$ simplices comprising the $1$-skeletons of a $3$-simplex with a particular choice of ordering. Relabeling the simplices of $K_0$ to the index set $[m]$ and modifying the labels of $K_1$ correspondingly yields permutations given below: 
%$$ K_0 = \{ {\color{red} (1) (2) (3) (4) (1,4) (2,4) (3,4) } (1,2) (2,3) (1,3) \} $$ 
%$$ K_1 = \{ {\color{red} (1) (2) (3) (4) } (1,2) (2,3) (1,3) {\color{red} (1,4) (2,4) (3,4) }\} $$ 
$$ K_0 = \{ {\color{red} a \; b \; c \; d \; u \; v \; w} \; x \; y \; z  \} = { {\color{red} 1 \; 2 \; 3 \; 4 \; 5 \; 6 \; 7 } \; 8 \; 9 \; 10 } $$
$$ K_1 = \{ {\color{red} a \; b \; c \; d } \; x \; y \; z \; {\color{red} u \; v \; w } \}  = { {\color{red} 1 \; 2 \; 3 \; 4 } \;8 \; 9 \; 10 \; {\color{red} 5 \; 6 \; 7 } }$$
The subset of the filtration which corresponds to the simplices which lie in the LCS between these permutations is colored in red. For this example, the permutation edit distance is $d = m - \mathrm{LCS}(K_0, K_1)$ implies exactly $3$ moves are needed to apply the mapping $K_0 \mapsto K_1$. There are six possible valid schedules of moves: 
\begin{alignat*}{3}
	S_1 = m_{x u}, m_{y u}, m_{z u} \quad & S_3 = m_{y u}, m_{x y}, m_{z u} \quad  & S_5 = m_{z u}, m_{x z}, m_{y z} \\
	S_2 = m_{x u}, m_{z u}, m_{y z} \quad  & S_4 = m_{y u}, m_{z u}, m_{x y} \quad  & S_6 = m_{z u}, m_{y z}, m_{x z} 
\end{alignat*}
The cost of each successive move operation for each of these schedules is recorded in Table~\ref{table:move_costs}.
\begin{table}[h]
\caption{Move schedule costs}
\centering
\begin{tabular}{ |P{0.4cm}|P{0.8cm}|P{0.8cm}|P{0.8cm}|P{0.8cm}|  }
 \hline
 \multicolumn{5}{|c|}{Cost of each permutation} \\
 \hline
 & 1st & 2nd & 3rd & Total\\
 \hline
 $S_1$ & 2 & 3 & 1 & 6 \\
 \hline 
 $S_2$ & 2 & 2 & 4 & 8 \\
  \hline 
 $S_3$ & 4 & 2 & 2 & 8 \\
  \hline 
 $S_4$ & 4 & 3 & 3 & 10 \\
  \hline 
 $S_5$ & 2 & 2 & 4 & 8  \\
  \hline 
 $S_6$ & 2 & 5 & 3 & 10\\
 \hline
\end{tabular}
\label{table:move_costs}
\end{table}
A greedy strategy which always selects the cheapest move in succession would begin by moving $x$ or $z$ first, since these are the cheapest moves available, which implies one of $S_1, S_2, S_5, S_6$ would be picked depending on the choice of a tie-breaker. While the cheapest schedule $S_1$ is in this candidate set, so is $S_6$, the most expensive schedule. Moreover, a tie-breaker that keeps track of previous cheapest moves would end up picking either schedule $S_2$ or $S_5$ as the final greedy schedule, both of which are suboptimal cost-wise.  
\\
\\
\noindent
Nonetheless, greedy strategies can be effective in practice, even if they are suboptimal. In the previous example, since evaluating the cost of move is $\approx O(m)$, a greedy strategy that chooses the least expensive move would require $O(d^2 m)$ time, which is computationally viable when $d << m$.

\subsubsection{Proxy objective}\label{sec:proxy_objective}
In light of the direct optimization being too expensive and a greedy procedure leading to suboptimal behavior, we consider whether there exists an alternative, or \emph{proxy objective} to optimize that is related to the cost of a schedule but not dependent on the entries in the decomposition. In this section, we present such an objective, showing its relation to the problem at hand as well its use in other known computational problems. 
We show empirical results demonstrating its practical effectiveness as a heuristic in section~\ref{sec:results}.

As before, given two permutations $p, q \in S_m$, let $S = ( s_1, s_2, \dots, s_d )$ denote a sorting such that $S \circ p = q$. 
In general, there are exponentially many ways to generate such a sorting $S$. 
The example given in section~\ref{sec:greedy} shows all (minimal size) $d!$ possible choices of $S$ for a fixed pair of permutations and choice of $\mathrm{LCS}(p,q)$. 
Let $\tau = \mathrm{LCS}(p,q)$ and let $\sigma \in S_d$ denote a choice of ordering   for the symbols in  $p \smallsetminus \tau$, 
such that $S(\sigma)$ denotes a schedule generated from this ordering. Ideally, we would like to achieve something similar to the following objective: 
\begin{equation}\label{eq:schedule_opt}
	S_\sigma^\ast = \argmin_{\sigma \in S_d} \mathrm{cost}(S(\sigma) \circ p)
\end{equation}
where $\mathrm{cost}(S(\sigma) \circ p)$ measures the number of column operations require to execute $S(\sigma)$ on an existing decomposition. 
The schedule $S_1$ shown in Table~\ref{table:move_costs} would be an example of a minimal cost schedule. Since optimizing this objective combinatorially is inefficient, we would like an alternative objective that is cheaper and performs well in practice. 
Intuitively, one alternative objective might be the Kendall-$\tau$ distance $K_\tau$, 
which can be computed in $O(m \log m)$ time---indeed, this is the weighting measure minimized in a certain path optimization in the RIVET software~\cite{lesnick2015interactive} to ensure adjacent filtrations are similar. Motivated by this, define $\hat{s}_i =  s_i \circ \cdots \circ s_2 \circ s_1 \circ p$ to be the result of applying the composition of the first $i$ permutations for some given schedule $S = (s_1, \dots, s_d)$, where $i < d$. 
A possible inequality to utilize is the following: 
\begin{equation}\label{eq:kendall_inequality}
K_\tau(p, q) = K_{\tau}(p, S \circ p) \leq \sum\limits_{i=1}^{d-1} K_{\tau}(\hat{s}_i, \hat{s}_{i+1})
\end{equation}
When the set of permutations $S$ are adjacent transpositions, equality is achieved above since $K_\tau(\hat{s}_i, \hat{s}_{i+1}) = 1$ for every inversion between $p$ and $q$. However, this is not guaranteed if $S$ was derived using the strategy from section~\ref{sec:move_schedules}. Obtaining $S_\sigma^\ast$ via direct optimization over $S_d$ isn't practical, but experimental tests suggest that a greedy-type procedure performs well as a heuristic, which yields a $O(d^2 m \log m)$ minimization procedure. When $d <<  m$, this strategy of successively minimizing the Kendall-$\tau$ distance is efficient. 
%where the set and order of pairs $\hat{s}_\ast$ depend on $S(\sigma)$. 
% In the vineyards algorithm, permutations take $O(1)$ time if a special sparse matrix structure is used~\cite{cohen2006vines}; others permutations $PRP^T = (PDP^T)(PVP^T)$ are non-trivial operations computationally. 
% \begin{displaymath}
% 	S_\sigma^\ast = \argmin_{\sigma \in S^d} \sum\limits_{i=1}^{d-1} K_{\tau}(\hat{s}_i, \hat{s}_{i+1})
%	%S_\sigma^\ast = \argmin_{\sigma \in S^d} \sum\limits_{i=1}^{d-1} K_{\tau}(\hat{s}_i, \hat{s}_{i+1})
%\end{displaymath}
%Let $p,q \in S^n$ be fixed permutations of $[n]$ and let $S = (s_1, \dots, s_d)$ denote a sorting $p \mapsto q$ such that $q = S \circ p = s_d \circ \cdots \circ s_1 \circ p$, and 
\\
\\
However, when $d \sim O(m)$, the quadratic nature of this minimization is too expensive. It turns out there is a similar distance that is both cheaper to compute and also amenable to a greedy type optimization. The \emph{Spearman distance} $F$ is a common distance for measuring the disarrangement between permutations. It is defined as:
\begin{equation}\label{eq:spearman_dist}
F(p, q) = \sum\limits_{i \in [m]} \lvert p(i) - q(i) \rvert
\end{equation}
Intuitively, the Spearman distance can be interpreted as $\ell_1$-type distance between two permutations. Like the Kendall-$\tau$ distance, it is a metric, it is invariant under relabeling, and the distance between pairs of uniformly random permutations $p, q \sim S_m$ are asymptotically normal. 
Indeed, Diaconis et al.~\cite{diaconis1977spearman} showed the following inequality relating the Spearman distance and the Kendall-$\tau$ distances:
\begin{equation}\label{eq:diaconis_inequality}
K_\tau(p, q) \leq F(p, q) \leq 2 K_\tau(p, q)
\end{equation}
Although very similarity to the Kendall-$\tau$ distance, the Spearman distance offers some convenient computational advantages: it can be computed in $O(m)$ time, and the variance of its asymptotic distribution is larger. Consider the following minimization: 
\begin{equation}\label{eq:schedule_opt_succ}
S_{\sigma}^\ast = \argmin_{\sigma \in S_d} \frac{1}{2} \sum\limits_{i = 1}^{d - 1} F(\hat{s}_i, \hat{s}_{i+1})
\end{equation}
The minimizer of this optimization can be interpreted as a sorting which minimizes the net displacement amongst all possible sortings, again in the $\ell_1$ sense. 
%\\
%\\
%\noindent
%\textbf{Example:} We reconsider the greedy optimization example from section~\ref{sec:greedy}. 
%Table~\ref{table:greedy_costs} lists the \emph{successive} Kendall-$\tau$ and Spearman distances, respectively, between adjacent LCS edit (move) operations. Observe that, in this example, both greedy approaches lead to the schedule with minimal cost ($S_1$) (though this is not guaranteed in the general case).  
%\begin{table}[h]
%\caption{Move schedule costs (kendall and spearman)}
%\centering
%\begin{tabular}{ |P{0.4cm}|P{0.8cm}|P{0.8cm}|P{0.8cm}|P{0.8cm}|  }
% \hline
% \multicolumn{5}{|c|}{ Successive $K_\tau$ cost} \\
% \hline
% & 1st & 2nd & 3rd & Total\\
% \hline
% $S_1$ & 3 & 5 & 7 & 15 \\
% \hline 
% $S_2$ & 3 & 6 & 6 & 15 \\
%  \hline 
% $S_3$ & 4 & 4 & 7 & 15 \\
%  \hline 
% $S_4$ & 4 & 6 & 5 & 15 \\
%  \hline 
% $S_5$ & 5 & 6 & 6 & 17 \\
%  \hline 
% $S_6$ & 5 & 5 & 5 & 15 \\
% \hline
%\end{tabular}
%\quad 
%\begin{tabular}{ |P{0.4cm}|P{0.8cm}|P{0.8cm}|P{0.8cm}|P{0.8cm}|  }
% \hline
% \multicolumn{5}{|c|}{ Successive $F$ cost} \\
% \hline
% & 1st & 2nd & 3rd & Total\\
% \hline
% $S_1$ & 3 & 3 & 3 & 9 \\
% \hline 
% $S_2$ & 3 & 4 & 4 & 11 \\
%  \hline 
% $S_3$ & 4 & 4 & 3 & 11 \\
%  \hline 
% $S_4$ & 4 & 4 & 5 & 13 \\
%  \hline 
% $S_5$ & 5 & 4 & 4 & 13 \\
%  \hline 
% $S_6$ & 5 & 5 & 5 & 15 \\
% \hline
%\end{tabular}
%\label{table:greedy_costs}
%\end{table}  
%\\
%% TODO: elaborate on monotonicity, describe segment tree
Given two permutations $p,q \in S_m$ and a LCS $\tau = \mathrm{LCS}(p,q)$ between them, the analysis in section~\ref{sec:move_schedules} shows we must move each symbol $x \in \mathcal{D}$ where $\mathcal{D} = p \smallsetminus \tau$ in a way that extends $\tau$ after each move. Thus we have $d = \lvert \mathcal{D} \rvert$ symbols to choose from initially. Each symbol $x \in \mathcal{D}$ yields a permutation $s$ to apply to $p$. Unlike the Kendall distance, the Spearman distance satisfies: 
$$ F(p, q) \geq F(s_1 \circ p, q) \geq F( s_2 \circ s_1 \circ p, q) \geq F( s_d \circ \dots \circ s_2 \circ s_1 \circ p, q)  = F(q,q) = 0 $$
For any sorting $S$ such that $s_{i}$ that increases the size of longest common subsequence between $s_{i-1} \circ \dots \circ s_1 \circ p$ and $q$. Since $\lvert \mathcal{D} \rvert$ is fixed, we would like to choose $x \in \mathcal{D}$ such that the corresponding permutation $s_x$ reduces the total displacement of every symbol $x \in \mathcal{D}$ relative to  $q$. A greedy strategy chooses $x$ such that $F(\hat{s}_i, \hat{s}_{i+1})$ is as small as possible, removes that symbol from $\mathcal{D}$, and recurses. Thus, a greedy strategy depends on our ability to quickly compute $F(\hat{s}_i, \hat{s}_{i+1})$. 
\\
\\
\noindent \textbf{Example:} An example of three possible schedules, $S_1$, $S_2$, and $S_3$ is given in Figure~\ref{fig:crossings}. Each schedule, from left to right, has $21$, $31$, and $35$ crossings, respectively. Each column represents the successive application of permutations from the three fixed schedules. Black/red vertices correspond to symbols in and outside of $\mathrm{LCS}(p,q)$, respectively. 
\begin{figure}[ht]
    \centering
    \includegraphics[width=0.50\textwidth]{crossings.png}
    \caption{Example of three possible schedules, $S_1$, $S_2$, and $S_3$.}
    %. Each schedule, from left to right, has $21$, $31$, and $35$ crossings, respectively. The greedy heuristic proposed to minimize~\ref{eq:schedule_opt_succ} was used to produce $S_1$, which has the minimal number of crossings, since $K_\tau(p,q) = 21$ in this example.
    \label{fig:crossings}
\end{figure}
All three schedules were generated from the same $\mathrm{LCS}(p,q) = (4\,7\,8)$ and all three schedules transform $p \mapsto q$ in $d = 6$ moves. The greedy heuristic proposed to minimize equation~\ref{eq:schedule_opt_succ} was used to produce $S_1$, which has the minimal number of crossings amongst all possible schedules in this example, since $K_\tau(p,q) = 21$. This minimization problem is related to a constrained version of the bipartite crossing minimization problem for $k$ sets of permutations, which is NP-hard for $k \geq 4$~\cite{biedl2009complexity}.
\\
\\
\noindent 
The monotonic nature of successively applying the Spearman distance enables certain computational advantages. Suppose we begin with an array $\mathcal{A}$ of size $m$ which provides $O(1)$ access and modification, initialized to $0$. Summing the entries in this array yields the Spearman distance in $O(m)$ time, which at initialization represents $F(p, p)$. Each symbol $x \in \mathcal{D}$ induces a  permutation $s_x$ such that $F(p, q) \geq F(s_x \circ p, q)$. Since there are $d = \lvert \mathcal{D} \rvert$ symbols to move and computing $F$ takes $O(m)$ time, a na\"ive greedy strategy requires $O((d+(d-1)+(d-2)+\dots+1)m) \approx O(d^2 m)$ time to complete. However, observe that each permutation $s_x$ changes the displacement of every symbol in a very predictable way: if $s_x$ represents a move permutation that moves $x$ from position $i$ to position $j$, then relative to the previous permutation $s_{i}$, the displacement of $x$ in $s_{i+1}$ changes by at most  $\lvert i - j \rvert$ and the displacement of any symbol between $i$ and $j$ changes by $\pm 1$. All other symbols are unaffected. Since the Spearman distance is simply the sum of these displacements, replacing $\mathcal{A}$ with a structure that supports both $O(\log m)$ access time to aggregate information (such as the sum) and $O(\log m)$ modification time ability to up to $\lvert i - j \vert$ entries. 
The former problem reduces to the problem of efficiently calculating and updating \emph{prefix sums}, however since $\lvert i - j \rvert \leq m$ is potentially larger than $O(\log m)$, it's not immediately clear how to achieve the latter modification complexity in $O(\log m)$ time. Fortunately, since every elements displacement other than $x$ changes by $\pm 1$ per update, both queries can be efficiently done in $O(\log m)$ time using a Segment Tree that supports \emph{lazy propagation}~\cite{de1997computational}. 
%We give an example and a more detailed discussion for those interested in the appendix\ref{sec:appendix}.

%This cumulative Spearman distance enables certain computational advantages. Suppose we begin with an array $\mathcal{D}$ of size $m$ which provides $O(1)$ access and modification, initialized to $0$, whose entries represent the summands of equation~\ref{eq:schedule_opt_succ}. 
%observe that each entry 
%
%%\begin{align}
%%	
%%\end{align}
%%$$ \argmin_{\sigma \in S_d} \frac{1}{2} \sum\limits_{i = 1}^{d - 1} F(\hat{s}_i, \hat{s}_{i+1}) = \argmin_{\sigma \in S_d} \frac{1}{2} \sum\limits_{i = 1}^{d - 1}  \sum\limits_{i \in [m]} \lvert \hat{s}(i) - \hat{s}_{i+1}(i) \rvert$$ 
%%
%% = 
%
%such that summing the entries in this array yields the Spearman distance in $O(m)$ time, which at initialization represents $F(p, p)$. Each symbol $x \in \mathcal{D}$ induces a  permutation $s_x$ such that $F(p, q) \geq F(s_x \circ p, q)$. Since there are $d = \lvert \mathcal{D} \rvert$ symbols to move and computing $F$ takes $O(m)$ time, a na\"ive greedy strategy requires $O((d+(d-1)+(d-2)+\dots+1)m) \approx O(d^2 m)$ time to complete. However, observe that each permutation $s_x$ changes the displacement of every symbol in a very predictable way: if $s_x$ represents a move permutation that moves $x$ from position $i$ to position $j$, then relative to the previous permutation $s_{i}$, the displacement of $x$ in $s_{i+1}$ changes by at most  $\lvert i - j \rvert$ and the displacement of any symbol between $i$ and $j$ changes by $\pm 1$. All other symbols are unaffected. Since the Spearman distance is simply the sum of these displacements, replacing $\mathcal{A}$ with a structure that supports both $O(\log m)$ access time to aggregate information (such as the sum) and $O(\log m)$ modification time ability to up to $\lvert i - j \vert$ entries. 
%The former problem reduces to the problem of efficiently calculating and updating \emph{prefix sums}, however since $\lvert i - j \rvert \leq m$ is potentially larger than $O(\log m)$, it's not immediately clear how to achieve the latter modification complexity in $O(\log m)$ time. Fortunately, since every elements displacement other than $x$ changes by $\pm 1$ per update, both queries can be efficiently done in $O(\log m)$ time using a Segment Tree that supports \emph{lazy propagation}~\cite{de1997computational}.

%Suppose we precompute for each symbol $\sigma \in p$ its displacement relative to $q$, $| p(i) - q(i) |$, and store these displacements in an array providing $O(1)$ access and modification. 
%each LCS edit operation changes the distance by one of {-1, 0, +1}.  
% To obtain the first greedy step, this must done for all $d$ potential symbols to move, making the first step $O(dm)$. Upon applying an operation s_1 \circ p, every symbol that was moved must have its displacement updated. A naive solution would be to update the vector as-is, recomputing the new spearman distance for each candidate symbol to move. However, notice that the change in displacement for each symbol is limited to one of three options: {-1, 0, +1}. We may update this vector by utilizing a clever trick called \emph{lazy propagation}, a trick to a well known data structure called a segment tree. Updates take O(\log m) time, as do queries, reducing the complexity to $O(m + d log d)$ time.

% TODO: Move computational results to subsection, add applications
% \subsection{Empirical results}\label{sec:results}
% \section{Applications and Discussion}\label{sec:applications}
\section{Empirical results}\label{sec:results}
\subsection*{Crocker stacks}
There are many challenges to depicting topological behavior in dynamic settings. One approach is to trace out the curves constituting a continuous family of persistence diagrams in $\mathbb{R}^3$---the vineyards approach---however this visualization can be cumbersome to work with as there are potentially many such vines tangled together, making topological critical events with low persistence difficult to detect. Moreover, the vineyards visualization does not admit a natural simplification utilizing the stability properties of persistence, as individual vines are not stable: if two vines move near each other and then pull apart without touching, then a pairing in their corresponding persistence diagrams may cross under a small perturbation, signaling the presence of an erroneous topological critical event~\cite{topaz2015topological, xian2020capturing}.

\begin{figure}[t]
	\centering
	\includegraphics[width=0.95\textwidth]{crocker_combo_1.png}
	\caption{An example of a crocker plot (right) depicting the evolution of dimension $p = 1$ Betti curves over time. The green $\mathrm{X}$ marks correspond chronologically to the complexes (left), in row-wise order. The large orange and purple areas depict $1$-cycles persisting in space (the y-axis) and time (x-axis). }\label{fig:crocker1}
\end{figure}

Acknowledging this, Topaz et al.~\cite{topaz2015topological} proposed the use of a 2-dimensional summary visualization, called a \emph{crocker}\footnote{\emph{crocker} stands for ``Contour Realization Of Computed k-dimensional hole Evolution in the Rips complex.'' Although the acronym includes \emph{Rips complexes} in the name, in principle a crocker plot could just as easily be created using other types of triangulations (e.g. \v{C}ech filtrations).} plot. 
In brief, a crocker plot is a contour plot on a family of Betti curves. Formally, given a filtration $K = K_0 \subseteq K_1 \subseteq \dots \subseteq K_m$, a $p$-dimensional \emph{Betti curve} $\beta_p^{\bullet}$ is defined as the ordered sequence of $p$-th dimensional Betti numbers:
$$ \beta_p^\bullet = \{ \, \mathrm{rank}(H_p(K_0)), \, \mathrm{rank}(H_p(K_1)), \, \dots, \, \mathrm{rank}(H_p(K_m))\, \}$$
Given a time-varying filtration $K(\tau)$, a crocker plot can be interpreted as a contour plot on the 1-parameter family of Betti curves $\beta_p^\bullet(\tau)$. A example of a crocker plot generated from the simulation described below is given in Figure~\ref{fig:crocker1}. Since only the Betti numbers at each simplex in the filtration are needed to generate these Betti curves, the persistence diagram is not directly needed to generate a crocker plot; it is sufficient to use e.g. any of the specialized methods discussed in~\ref{sec:related_work}. This dependence only on the Betti numbers makes crocker plots easier to compute than standard persistence, however what one gains in efficiency one loses in stability; it is known that Betti curves are inherently unstable with respect to small fluctuations about the diagonal of the persistence diagram.

Acknowledging this, Xian et al.~\cite{xian2020capturing} showed that crocker plots may be \emph{smoothed} to inherit the stability property of persistence diagrams and reduce noise in the visualization. That is, when applied to a time-varying persistence module $M = \{M_t\}_{t \in [0, T]}$ an $\alpha$-smoothed crocker plot for $\alpha \geq 0$ is the rank of the map $M_t(\epsilon - \alpha) \to M_t(\epsilon + \alpha)$ at time $t$ and scale $\epsilon$. For example, the standard crock plot is a $0$-smoothed crocker plot. Allowing all three parameters ($t, \epsilon, \alpha$) to vary continuously leads to 3D visualization called an $\alpha$\emph{-smoothed crocker stack}.
\begin{definition}[crocker stack]
	A crocker stack is a family of $\alpha$-smoothed crock plots which summarizes the topological information of a time-varying persistence module $M$ via the function $f_M : [0, T] \times [0, \infty) \times [0, \infty) \to \mathbb{N}$, where:
	$$ f_M(t,\epsilon, \alpha) = \mathrm{rank}(M_t(\epsilon - \alpha) \to M_t(\epsilon + \alpha)) $$
	Thus, a crocker stack is a sequence of $\alpha$-smoothed crocker plots that vary over $\alpha \geq 0$, satisfying $f_M(t,\epsilon,\alpha) \leq f_M(t,\epsilon, \alpha)$ for all $\alpha \geq \alpha'$. 
\end{definition}
\noindent Note that, unlike crocker plots, applying this $\alpha$ smoothing efficiently \emph{requires} the persistence pairing. Indeed, it has been shown that crocker stacks and stacked persistence diagrams (i.e. vineyards) are equivalent to each other in the sense that either one contains the information needed to reconstruct the other~\cite{xian2020capturing}.  Thus, computing crocker stacks reduces to computing the persistence of a (time-varying) family of filtrations.

%To describe this formally, let $K_\tau^\epsilon$ denote a simplicial filtration with simplices $\sigma \in K_\tau^\epsilon$ whose function value $f_\tau(\sigma) \leq \epsilon$. The $\alpha$

We test the efficiency of computing the necessary information to generate these crocker stacks using a spatiotemporal data set to illustrate the applicability of our method. Specifically, we ran a \emph{flocking} simulation similar to the simulation run in~\cite{topaz2015topological} with $m = 20$ vertices moving around on the unit square equipped with periodic boundary conditions (i.e. $S^1 \times S^1$). We simulated movement by equipping the vertices with a simple set of rules which control how the individual vertices position change over time. Such simulations are also called \emph{boid} simulations, and they have been extensively used as models to describe how the evolution of collective behavior over time can be described by simple sets of rules.
The simulation is initialized with every vertex positioned randomly in the space; the positions of vertices over time is updated according to a set of rules related to the vertices acceleration, distance to other vertices, etc. To get a sense of the time domain, we ran the simulation until a vertex made at least 5 rotations around the torus. 
\begin{figure}[ht]
	\centering
	\includegraphics[width=0.65\textwidth]{boid_sim_results.png}
	\caption{ The cumulative number of $O(m)$ column operations (left, log-scale) of the three approaches tested and the distribution of the normalized Kendall distance between adjacent filtration pairs. The histogram on the right depicts the coarseness of the discretization; most filtration pairs have $\approx 5\%$ of their $\approx O(m^2)$ simplex pairs inverted between adjacent time steps.}\label{fig:boid_sim_results}
\end{figure}

Given this time-evolving data set, we computed the persistence diagram of the Rips filtration up to $\epsilon = 0.30$ at 60 evenly spaced time points using three approaches: the standard algorithm \texttt{pHcol} applied naïvely at each of the 60 time steps, the vineyards algorithm applied to (linear) homotopy connecting filtrations adjacent in time, and our approach using moves.   
The cumulative number of $O(m)$ column operations executed by three different approaches. Note again that vineyards requires generating many decompositions by design (in this case, $\approx 1.8M$). The standard algorithm \texttt{pHcol} and our move strategy were computed at 60 evenly spaced time points of the simulations. As depicted in Figure~\ref{fig:boid_sim_results}, our move strategy is far more efficient then both vineyards and the naive \texttt{pHcol} strategies.

\subsection*{Multidimensional persistence}
A natural question is whether or not there exists a multi-dimensional generalization of persistence. The starting point for this generalization is a structure which filters the space in multiple dimensions simultaneously, a \emph{multifiltration}. Given this structure, the goal of multi-dimensional persistence is to identify persistent features by examining the entire multifiltration. Such a generalization has been thought to have enormous potential in terms of applications~\cite{}. 
For example, one of the often noted drawbacks of persistence is its instability with respect to strong outliers. That is, although persistence diagrams are stable with respect to their input and thus are robust to noise, the presence of outliers can artificially introduce topological features such as loops, obscuring the persistence of significant topological features~\cite{buchet2015topological}. Indeed, this obfuscation can hinder the ability of 1D persistence to uncover ``significant'' topological features, one of the original purposes of persistence. In theory, this formalism enables the practitioner to analyze even noisy data sets directly without resorting to sophisticated data preprocessing techniques which may or may not preserve the underlying topology of the data.

Unfortunately, multi-dimensional persistent homology comes with its own set of new computational and theoretical challenges not necessarily inherited by the 1-D case. The multi-dimensional persistent homology of a multifiltration cannot be captured by complete discrete invariant, and incomplete invariants have proven challenging to use in practice due to the high algorithmic complexity of their corresponding computation. A number of computational strategies have been proposed to counter this, such as homology preserving multi-filtration simplification or parallelization schemes using shared-memory computation~\cite{fugacci2019chunk, lesnick2019computing}. Another strategy is to focus specifically on a fixed dimension, such as 2-D persistence. Utilizing the equivalence between the rank and fibered barcode invariants, Lesnick and Wright~\cite{lesnick2015interactive} developed an elegant way of computing the former via a reparameterization of the latter using standard point-line duality. This clever reparameterization effectively reduces the rank invariant computation to sequence of 1-D barcode computations at ``template points'' lying within the 2-cells of particular planar subdivision of the half-plane $[0, \infty) \times \mathbb{R}$. Indeed, since algorithm~\ref{alg:schedule} was designed for precisely such a computation, the algorithm proposed by~\cite{lesnick2015interactive} is prototypical of the class of methods that stand to benefit from move scheduling.

The purpose of the following sections is to uncover the intrinsic bottlenecks impeding the computation of the rank invariant in practical settings. Towards this goal, we briefly summarize the theory established by several researchers~\cite{lesnick2015interactive, carlsson2010computing, landi2014rank}, focusing on their computational and algorithmic details. 
Our goal is to show that the highest complexity subcomputation of the rank invariant algorithm proposed by~\cite{lesnick2015interactive} is precisely the type of computation we focus on in this effort. 
In doing so, we substantiate our claims from section~\ref{sec:motivation} that our scheduling approach can improve the efficiency of a real-world application of multidimensional persistence by an order of magnitude by comparing Algorithm~\ref{alg:schedule} to several alternatives on a real-world data set and problem: the problem of identifying the underlying topological space of a data set extracting from high-contrast natural images. 
% In what follows, we briefly recount an incomplete but nonetheless very useful invariant of multi-dimensional persistence modules, the \emph{rank invariant}, as well as another invariant which in some sense determines the rank invariant, the \emph{fibered barcode}. It turns out that computing the latter invariant

% --- For each section of the paper, consider writing a mini-introduction that says what its organization is, what is in each subpart, and how the parts relate to one another.

%To demonstrate how our scheduling approach can accelerate the multidimensional persistence computation, we establish the computational framework we will be considering. 
% These transforms end up being crucial in decomposing the definition of the fibered barcode (and thus, the rank invariant) into a discrete computation.

\subsubsection*{The Rank Invariant}

%We begin with a few definitions. An $n-D$ \emph{filtration} is a functor $\mathcal{F} : \mathbb{R}^n \to \textbf{Simp}$, the category of simplicial complexes, such that for all $a \leq b$, $\mathcal{F}(a,b) : F_a \to F_b$ is an inclusion. We will focus exclusively on the case where $n = 2$, where the object of study is a \emph{bifiltration}. Multifiltrations can be broadly categorized into two types: 1-critical and multi-critical filtrations. A \emph{1-critical filtration} $\mathcal{F}$ is a finite $n-D$ filtration each simplex $s \in K$ is associated with only one multi-grade (denoting its appearance in each of the filtered dimensions), otherwise $\mathcal{F}$ is said to \emph{multi-critical}. 
%An $n-D$ persistence module is a functor $M : \mathbb{R}^n \to \textbf{Vect}$, where $\textbf{Vect}$ denotes the category of $\mathbf{F}$-vector spaces and linear maps for some suitable field $\mathbb{F}$. A module $M$ is said to be \emph{pointwise finite dimensional} (p.f.d) if $\mathrm{dim}(M_a) < \infty$ for all $a \in \mathbb{R}^n$.
%For $i \geq 0$, let $H_i : \textbf{Simp} \to \textbf{Vect}$ denote the $i^{\text{th}}$ simplicial homology functor with coefficients in $\mathbb{F}$. 
%Using this notation, given a 1D filtration $\mathcal{F}$ of size $m$, each p.f.d 1D persistence module $M = H_i(\mathcal{F})$ for  may be associated with a multiset $\mathcal{B}(M)$ of intervals in $\mathbb{R}$ which record the isomorphism classes of indecomposable summands of $M$. The set $\mathcal{B}(M)$ is called the \emph{barcode}\footnote{Note that \emph{barcodes} and \emph{persistence diagrams} are interchangeable representations of the same multiset.} of $M$.

An \emph{invariant} of $n$-parameter persistence modules is a function from the collection of persistence modules to some set $S$ with the property that $f(M) = f(M')$ when $M \cong M'$. An invariant is said to be \emph{complete} if $M \cong M'$ whenever $f(M) = f(M')$. Thus, incomplete invariants are invariants which may have the same value on two non-isomorphic modules. 
\emph{The rank invariant} proposed by Carlsson et al~\cite{carlsson2009theory} is one such invariant that is complete when $n=1$ but is otherwise incomplete when $n \geq 2$. 
% The rank invariant is computable in polynomial time~\cite{carlsson2010computing}, that has been the subject of much interest lately. 
It is defined as follows: 
\begin{definition}[Rank Invariant]
	For $n \geq 1$, the \emph{rank invariant} of a $\mathbb{R}^n$-indexed persistence module $M$ over $\mathcal{H}^{n} \subseteq \mathbb{R}^n \times \mathbb{R}^n$ is the function:
	\begin{equation*}
		\begin{aligned}
		& \mathrm{rank}(M)  \; : & \mathcal{H}^n & \to \mathbb{N}  \\
		& & (a,b) & \mapsto \mathrm{rank} \, M(a,b) 
		\end{aligned}
	\end{equation*}
	for $a \leq b \in \mathcal{H}^n$. When $n = 1$, the rank invariant is complete, thus $\mathrm{rank}(M)$ and the persistence barcodes $\mathcal{B}(M)$ determine each other.
\end{definition}
\noindent Thus, in one dimension, persistent homology provides a complete invariant that fully captures the lifetime of topological features in a given filtration: 1-D persistent homology modules decompose in an essentially unique way into indecomposable summands~\cite{zomorodian2005computing}. In contrast, the theory of multidimensional persistence shows that no complete discrete invariant exists---the structure of the target for the invariant depends on the structure of the underlying field, and although multiparameter persistence modules also decompose in a similar way as the 1-D case, the set of isomorphism classes is known to be extremely complicated~\cite{carlsson2009theory, lesnick2015interactive}. 
Nonetheless, there are invariants related to multi-parameter persistence that can be quite useful in practice, even though they are incomplete. 
\noindent For example, even though the rank invariant does not encode the isomorphism type of the underlying module $M$ when $n > 1$, it does captures important ``first-order'' information about the modules structure.

Along this vein of research, Lesnick and Wright~\cite{lesnick2015interactive} identify three simple invariants associated with a multidimensional persistence module $M$: the \emph{dimension function of $M$}, the \emph{multigraded Betti numbers of $M$}, and the \emph{fibered barcode of $M$}. 
% The dimension function is simply the the function which maps every point $a \in \mathbb{R}^2$ to $\mathrm{dim}(M_a)$. It is a simple and easy to visualize invariant, but is unstable and yields no information about the persistent features of $M$. 
% The multigraded Betti numbers constitute a natural and important class of invariants associated with a multidimensional persistence module. There are many ways of defining these, with perhaps the simplest definition stemming from minimal resolutions: for a finitely generated $n$-parameter persistence module $M$ and $i \in \{0, 1, \dots, d \}$, the $i^{\text{th}}$ graded Betti number of $M$ at grade $z$, denoted as $\xi_i^M(z)$, is defined as the number of elements at grade $z$ in a basis of the $i^{\text{th}}$ module in a free resolution for $M$. 
% Thus, the $i^{\text{th}}$-graded Betti number of $M$ is a function $\xi_i(M): \mathbb{R}^n \to \mathbb{N}$. Being central objects in commutative algebra and algebraic geometry, both the theory and the computation of multi-graded Betti numbers is extensive and beyond the scope of this effort---the interested reader is referred to~\cite{carlsson2009theory, lesnick2015interactive} for more details. 
The fibered barcode, denoted as $\mathcal{B}(M)$, is defined as follows: 
\begin{definition}[Fibered barcode]
	%\in \overline{\mathcal{L}}
	The fibered barcode $\mathcal{B}(M)$ of a bipersistence module $M$ is the map which sends each line $L$ with non-negative slope to the barcode $\mathcal{B}_L(M)$: 
	$$ L \mapsto \mathcal{B}_L(M) $$
	Thus, the fibered barcode is simply a 2-parameter family of barcodes given by the collection of 1-D affine slices of $M$. 
\end{definition} 
\noindent Each line $L$ induces a 1-parameter restriction on $M$. The fibered barcode enjoys a number of stability properties readily expressed using interleaving distances~\cite{landi2014rank}. More important to this work, the fibered barcode $\mathcal{B}(M)$ and the rank invariant $\mathrm{rank}(M)$ determine each other~\cite{lesnick2015interactive}.

\subsubsection*{Fibered barcode reparameterization}\label{sec:fibered_barcode_reparam}
Let $\overline{\mathcal{L}}$ denote the collection of all lines in $\mathbb{R}^2$ with non-negative slope, $\mathcal{L} \subset \overline{\mathcal{L}}$ the collection of all lines in $\mathbb{R}^2$ with non-negative, finite slope, and $\mathcal{L}^\circ$ the collection of all affine lines in $\mathbb{R}^2$ with positive, finite slope. There is a standard point-line duality that gives a parameterization of $\mathcal{L}$ with the half-plane $[0, \infty) \times \mathbb{R}$ that is convenient to work with here. Define the \emph{line} and \emph{point} dual transforms $\mathcal{D}_{\ell}$ and $\mathcal{D}_p$, respectively, as follows: 
\begin{equation}
	\begin{aligned}
\mathcal{D}_{\ell}: \mathcal{L} \rightarrow[0, \infty) \times \mathbb{R} & \quad \hfill \quad & \mathcal{D}_{p}:[0, \infty) \times \mathbb{R} \rightarrow \mathcal{L} \\
y=a x+b \mapsto(a,-b) &\quad \hfill \quad & (c, d) \mapsto y=c x-d
\end{aligned}
\end{equation}
% Use \multicolumn{1}{|r|}{text} to align
The transforms $\mathcal{D}_{\ell}$ and $\mathcal{D}_p$ are \emph{dual} to each other in the sense that for any point $a \in [0, \infty) \times \mathbb{R}$ and any line $L \in \mathcal{L}$, $a \in L$ if and only if $D_\ell(L) \in D_{p}(a)$. Now, for some fixed line $L$, define  the \emph{push map} $\mathrm{push}_L(a):  \mathbb{R}^2 \to L \cup \infty$ as: 
\begin{equation}
	\mathrm{push}_L(a) \mapsto \mathrm{min}\{ v \in L \mid a \leq v \}
\end{equation}
%\begin{equation*}
%	\begin{aligned}
%	& \mathrm{push}_L : & \mathbb{R}^2 & \to L \cup \infty \\
%	& & a \in \mathbb{R}^2 & \mapsto \mathrm{min}\{ v \in L \mid a \leq v \}
%	\end{aligned}
%\end{equation*}
The push map satisfies a number of useful properties. Namely: 
\begin{enumerate}
	\item For $r < s \in \mathbb{R}^2$, $\mathrm{push}_L(r) \leq \mathrm{push}_L(s)$
	\item For each $a \in \mathbb{R}^2$, $\mathrm{push}_L(a)$ is continuous on $\mathcal{L}^\circ$
	\item For $L \in \mathcal{L}^\circ$ and $S \subset \mathbb{R}^2$, $\mathrm{push}_L$ induces a totally ordered partition $S_L$ on $S$ 
\end{enumerate}
Property (1) asserts that for any $L \in \overline{\mathcal{L}}$, the partial order on $\mathbb{R}^2$ restricts to a total order on $L$. Properties (2) and (3) qualify the following definition:
\begin{definition}[Critical Lines]
	For some fixed $S \subset \mathbb{R}^2$, a line $L \in L^\circ$ is defined to be \emph{regular} if there is an open ball $B \in L^\circ$ containing $L$ such that $S_L = S_{L'}$ for all $L' \in B$. Otherwise, the line $L$ is defined as \emph{critical}. 
\end{definition}
\noindent The set of critical lines $\mathrm{crit}(M)$ with respect to some fixed set $S \subset \mathbb{R}^2$ fully characterizes a certain planar subdivision of the half plane $[0, \infty) \times \mathbb{R}$. Explicitly, define $\mathcal{A}^1(M)$ to the 1-skeleton of the line arrangement defined by: 
% \{ \mathcal{D}_p(\alpha) \text{ for } \alpha = \mathrm{LUB}(u,v), u,v \in S \} 
\begin{equation}\label{eq:arrangement_crit}
	\mathcal{A}^1(M) = \{ \mathcal{D}_{\ell}(\mathrm{crit}(M)) \cup (\{0\} \times \mathbb{R}) \}
\end{equation}
Expanding the 1-skeleton to a 2-D cell complex $\mathcal{A}(M)$ fully partitions the upper half plane.  
A consequence of equation~\ref{eq:arrangement_crit} that if the duals of two lines $L, L' \in \mathcal{L}$ are contained in the same $2$-cell in $\mathcal{A}(M)$, then $S_L = S_{L'}$, i.e. the partitions induced by $\mathrm{push}_L$ are equivalent. Indeed, the total order on $S_L$ is the pullback of the total order on $L$ with respect to the push map.
Since $\mathcal{A}(M)$ partitions the entire half-plane, the dual to every line $L \in \mathcal{L}$ is contained within $\mathcal{A}(M)$---the desired reparameterization.

Let $S = \mathrm{supp}\,\xi_0(M) \, \cup \, \mathrm{supp}\,\xi_1(M)$, where the functions $\xi_0(M), \xi_1(M)$ are $0^{\text{th}}$ and $1^{\text{st}}$ bigraded Betti numbers of $M$, respectively. 
The main mathematical result from~\cite{lesnick2015interactive} is a characterization of the barcodes $\mathcal{B}_L(M)$, for any $L \in \mathcal{L}$, in terms of a set of \emph{barcode templates} $\mathcal{T}$ computed at every 2-cell in $\mathcal{A}(M)$.
 More formally, for any line $L \in \overline{\mathcal{L}}$ and $e$ any 2-cell in $\mathcal{A}(M)$ whose closure contains the dual of $L$ under point-line duality, the 1-parameter restriction of the persistence module $M$ induced by $L$ is given by: 
 \begin{equation}
 \mathcal{B}_L(M) = \{ [\,\mathrm{push}_L(a), \, \mathrm{push}_L(b)\,) \mid (a,b) \in \mathcal{T}^e, \mathrm{push}_L(a) < \mathrm{push}_L(b)  \} 
\end{equation}
Minor additional conditions are needed for handling completely horizontal and vertical lines. 
The importance of this theorem for this effort lies in the fact that the fibered barcodes are completely defined from the precomputed barcode templates $\mathcal{T}$: once $\mathcal{T}$ has been computed and augmented onto $\mathcal{A}(M)$, $\mathcal{B}(M)$ is completely characterized, and the barcodes $\mathcal{B}_L(M)$ of any $L$-induced 1-D filtration can be efficiently computed via a point-location query on $\mathcal{A}(M)$ and a $O(\lvert \mathcal{B}_L(M) \rvert)$ application of the push map.

%For some fixed set $S \subset \mathbb{R}^2$ and any pair $u, v \in S$ of points which are incomparable with respect to the partial order on $\mathbb{R}^2$
% The definition of \emph{critical} lines provides an analogue to the

%This is the main mathematical result from~\cite{lesnick2015interactive},

% Indeed, it was shown in~\cite{lesnick2015interactive} that the fibered barcodes $$ of $M$ (i.e. the collection of barcodes of 1D affine slices of $M$) and the rank invariant determine each other.     
\subsubsection*{Computing the fibered barcode}
%Lesnick and Wright introduced the Rank Invariant Visualization and Exploration Tool (RIVET)~\cite{lesnick2015interactive}, which is both a tool and a collection of computational and mathematical theory which enables exploratory analysis of 2D persistence modules built from arbitrary bifiltrations. From a computational perspective, the core novelty of the RIVET approach is its ability to compute the fibered barcodes efficiently from precomputed information. 
The three dominant subcomputations needed to compute the fibered barcode using the reparameratization discussed above are the following: 
\begin{enumerate}
	\item Computing the bigraded Betti numbers $\xi(M)$ of a given bipersistence module $M$ 
	\item Constructing a line arrangement $\mathcal{A}(M)$ induced by critical lines derived from (1) 
	\item Augmenting $\mathcal{A}(M)$ with a \emph{barcode template} $\mathcal{T}^e$ at every 2-cell $e \in \mathcal{A}(M)$
	%, yielding the \emph{augmented arrangement} $\mathcal{A}^{\sbullet}(M)$ 
\end{enumerate}
It is shown in~\cite{lesnick2019computing} how to compute (1) in approximately $\approx O(m^3)$ using a matrix algorithm similar to Algorithm~\ref{alg:reduce}. Constructing and storing the line arrangement $\mathcal{A}(M)$ with $n$ lines and $k$ vertices can be achieved using known algorithms in computational geometry (e.g. the Benteley-Ottmann algorithm) in output-sensitive $O((n+k) \log n)$ time and $O(\kappa^2)$ memory, where $\kappa$ is a coarseness parameter associated with the computation of $\xi(M)$. 
Finally, the computation of the barcodes templates $\mathcal{T}$ (step 3) takes on the order of $O(m^3 \kappa + m \kappa^2 \log \kappa)$ elementary operations and $O(m \kappa^2)$ storage.

Of the three steps above, the barcode template computation is both the highest complexity and most demanding computation required by the algorithm. The number of 2-cells in $\mathcal{A}(M)$ is on the order $O(\kappa^2)$, and $\kappa$ itself is on the order of $O(m^2)$ in the worst case. 
This suggests the computation of the barcode templates is on the order $O(m^5)$, which is effectively intractable for moderately-sized filtrations. In practice, the external stability result from~\cite{landi2014rank} justifies the use of a simple procedure which approximates the module $M$ with a smaller module $M'$ using a grid-like coarsening procedure. Two binning parameters associated with the coarsening procedure, $x_{bin}$ and $y_{bin}$, together determine the degree of the approximation. The coarsening procedure allows the practitioner to restrict the size of $\kappa$ to a constant, which in-turn can dramatically reduce the size of $\mathcal{A}(M)$ when $\kappa << m^2$. Keeping in mind this constant factor, the entire computation for approximating the rank invariant given a multifiltration with $m$ simplices is $\approx O(m^3)$.

There are several approaches one can use to compute $\mathcal{T}$, the simplest being to run Algorithm~\ref{alg:reduce} independently on the 1-D filtration induced by the duals of some set of points (e.g. the barycenters) contained within the 2-cells of $\mathcal{A}(M)$.  
The approach taken by~\cite{lesnick2015interactive} is to use an $R = DV$ decomposition previously computed at some 2-cell $e \in \mathbb{A}(M)$ to speed up the computation of a new decomposition at some adjacent cell $e' \in \mathbb{A}(M)$. More explicitly, define the \emph{dual graph} of $\mathcal{A}(M)$ to be the undirected graph $G$ which has a vertex for every 2-cell $e \in \mathcal{A}(M)$ and an edge for each pair of cells $e, e' \in \mathcal{A}(M)$ which are adjacent in $\mathcal{A}(M)$. 
Each vertex in $G$ is associated with a barcode template $\mathcal{T}^e$, and the computation of $\mathcal{T}$ now reduces to computing a path $\Gamma$ on $G$ which visits each vertex at least once. To minimize the computation time, assume the $n$ edges of $G$ are endowed with a weights $w_1, w_2, \dots, w_n$ whose values represent some notion of distance between barcode templates.  The optimal path $\Gamma^\ast$ would be one of minimal length which visits every vertex of $G$ at least once. A $\frac{3}{2}$-approximation for computing an approximately minimal path is known, which reduces the problem to the traveling salesman problem on a metric graph~\cite{christofides1976worst}. 
\\
\\
\textbf{Noisy circle example}
\begin{figure}[h]
	\includegraphics[width=0.98\textwidth]{bifiltration_ex2}
	\caption{Bipersistence example on an $8 \times 8$ coarsened grid. On left, the input data, colored by density. In the middle, the bigraded Betti numbers $\xi_0(M)$ and $\xi_1(M)$ (green and red, respectively), the Hilbert function (gray), and a line $L$ emphasizing the persistence of features with high density. On the right, the line arrangement $\mathcal{A}(M)$ lying in the dual space $D_p(\alpha)$ derived from the $\xi(M)$. The barycenters of the 2-cells in $\mathcal{A}(M)$ depicting where the barcodes templates $\mathcal{T}$ are computed are drawn as orange points, and the dual graph $G$ of $\mathcal{A}(M)$ is drawn using dotted blue lines. The shortest path $\Gamma^\ast$ spanning the vertices of $G$ is shown by the solid colored lines. }
\end{figure}
\\

%To an along an Eulerian path spanning the dual graph $G$ of the arrangement $\mathcal{A}(M)$. The
%As input, one has a set of points $A \subset \mathbb{R}^2$, whose duals induce filtrations ww
%Since each template point corresponds to a dual 
%TODO: (Discuss path along dual graph of $\mathcal{A}(M)$)
% It is worth briefly discussing just how difficult computationally step (3) is. For a bifiltration with $m$ simplices, \emph{without any coarsening}, $\kappa$ is on the order of $O(m^2)$ in the worst case. Thus, the uncoarsened complexity of the template computation barcode is $\approx O(m^5)$, where again $m$ is the number of simplices in the bifiltration. A more common practice in data science is to report the complexity of algorithm in terms of the number of data points. Given some input point cloud data (PCD) with $n$ points, the standard rips filtration up to dimension 2 can have upwards of $O(n^3)$ 2-simplices. Thus, given a 2-dimensional bifiltration $\mathcal{F}$ with $m = n^3$ simplices derived from a rips filtration on PCD with $n$ data points, the complexity of computing all barcode templates for every $2$-cell in $\mathcal{A}(M)$ for $M = H_1(\mathcal{F})$ is on the order of $O((n^3)^5) \approx O(n^{15})$. This is clearly computationally intractable for even relatively small data sets.

%There are a few ways of tackling this computational barrier. Recent work has focused on using alternative filtration whose size is much smaller than that of the rips (or cech) filtration, but whose homology is approximately the same, for example by using a sparsified rips filtration~\cite{} or the chunk-based approach to multi-parameter persistence~\cite{fugacci2019chunk}. It was shown in~\cite{} that one can alternatively opt to approximate a bipersistence module $M$ by a \emph{coarsened} module $M'$ in a way that bounds the the interleaving distance $d_I(M, M')$ between the two modules. The external stability result of Landi~\cite{} implies that if the two modules $M$ and $M'$ are close in their interleaving distance, then the fibered barcodes $\mathcal{B}(M)$ and $\mathcal{B}(M)$ will be close as well. In practice, this coarsening procedure is essential to enabling the practical computation of the fibered barcodes.

% However, if a give bipersistence module $M$ is coarsened at all, the 
%  Specifically, assembling the barcode takes just $O(\log \kappa + \lvert \mathcal{B}_L(M) \rvert ) \approx O(\lvert \mathcal{B}_L(M) \rvert )$ time when $\kappa$, a coarsening parameter discussed below, is treated as a constant.

 \subsubsection*{Natural images dataset} 
 In the previous section, we outlined the computational theory of multidimensional persistence, including its relevance to our proposed move scheduling approach. In this section, we demonstrate the efficiency of our method on practical use case of multidimensional persistence: detecting the presence of a low-dimensional topological space which well-approximates the distribution of high dimensional data.

 \begin{figure}[t]
	\includegraphics[width=0.98\textwidth]{natural_images}
	\caption{Bipersistence example of natural images data set on an $12 \times 16$ coarsened grid. On the left, a projection the full data set onto the first two DCT basis vectors is shown, along with the 15 landmark patches. In the middle, the bigraded Betti numbers, the Hilbert function, and a chosen line $L$ are shown for the coarsened grid. As before, the $0/1/2$ dimension bigraded Betti numbers are shown in green/red/yellow, respectively. The blue regions highlights where $\mathrm{dim}(M) = 5$. On the right, the persistence diagram of $L$, with larger points indicating a multiplicity of 2. Observe five persistent features are revealed in both the middle and right plots, matching $\beta_1$ of the three-circle model. }
	\label{fig:patch_data_dgm}
\end{figure}

A common hypothesis is that high dimensional data tend to lie in the vicinity of an embedded, low dimensional manifold or topological space. An exemplary demonstration of this is given in~\cite{lee2003nonlinear}, who explored the space of $3 \times 3$ high-contrast patches extracted from Van Hatern's still image collection, which consists of approximately 4,000 monochrome images depicting various areas outside around Groningen (Holland)~\cite{}. The study of how empirical data from natural images is distributed with respect to predicted spaces and manifolds is motivated by applications in sparse coding and efficient image representations. In particular, Lee et al.~\cite{} was interested in analyzing the empirical probability distribution of high-contrast $3 \times 3$ patches of natural images, in pixel-space, towards understanding whether there were any clear qualitative differences the distributions of patches extracted from from images of different modalities, e.g. optical versus range images. 
Experiments conducted by Carllsson et al.~\cite{} using persistent homology confirmed that the distribution of high-contrast $3 \times 3$ patches is indeed well-approximated by a Klein bottle---around 60\% of the high-contrast patches from the still image data set are concentrated within a small neighborhood around a model of the Klein bottle which accounts for $\approx$ 21\% of the volume $S^7$, compared to the $\approx$ 84\% of volume of $S^7$ accounted for by the entire data set. In a subsequent effort by Perea et al.~\cite{}, a framework for estimating and representing the distribution around low dimensional submanifolds of pixel space is given, establishing a principled way of obtaining multi-scale rotation invariant descriptors useful for classification tasks.

Using the same high-contrast patch data set studied in~\cite{lee2003nonlinear}, we evaluate the performance of various methods at computing the rank invariant using the parameterization from~\ref{sec:fibered_barcode_reparam}. Specifically, we begin by extracting a representative subset of the set $X \subset \mathcal{X}$ of patches representing the top 30\% densest patches of the space of all patches $\mathcal{X}$. Carlsson et al.~\cite{} studied a variety of choices of $k$ for a k-nearest neighbor (knn) density estimator $\rho$---which determines which points are the ``densest''---in order to understand how the data were distributed around the Klein bottle model. It was determined empirically that when $k \approx 15$, the top 30\% captures three types of directional shifts in the grayscale gradients of the image patches. These three gradients were shown to be captured well by the three-circle model $C_3$---the 1-skeleton of the Klein bottle. In this space, there are three circles, two of which (say, $S_v$ and $S_h$) intersect the third (say, $S_\mathrm{lin}$) in exactly two points, but they themselves do not intersect. One of the circles, $S_{\mathrm{lin}}$, parameterizes a the rotational gradient, whereas the other two $S_v$ and $S_h$ parameterize vertical and horizontal gradients. A projection of the data $X$ onto the first two DCT basis, is shown in the left-most image in Figure~\ref{fig:patch_data_dgm}, along with 15 landmark points. It was shown in~\cite{} that when $k \approx 300$, the top 30\% of data extracted from the knn density estimator $\rho_{300}$ `fills out' the two-manifold, yielding Betti numbers $\beta_0 = \beta_1 = 1$, whereas when $k$ is decreases to a more localized estimate, the three-circle model is recovered, which has Betti numbers $\beta_0 = 1, \beta_1 = 5$.

We sampled landmarks from $X \subset \mathcal{X}$ using a stratified sampling approach with $n = 25$ strata, where each strata contains the $(1/n)$-thick level set given by the knn density estimate $\rho_{15}$. The use of landmarks gives us certain coverage guarantees that the geometry is preserved within each level set, whereas the stratification ensures the original density of is approximated preserved as well. From this data set, we construct a Rips-density bifiltration $\mathcal{F}$ using $\rho_{15}$ equipped with the geodesic metric computed over the $k = 15$ knn graph of $X$. In theory, this bifiltration should be well approximated by $C_3$, which we ought to be able to determine by a careful inspection of the rank invariant. That is, since $C_3$ has $\beta_1  = 5$, there should be some line $L$ which induces a 1-parameter filtration whose corresponding persistence diagram (for dimension $1$) displays five highly persistent features. We show an example of the data, the bigraded Betti numbers, and the persistence diagram extracted from a particular choice of line $L$ in Figure~\ref{fig:patch_data_dgm}.

The example from~\ref{fig:patch_data_dgm} demonstrates a practical application of the rank invariant to uncovering the underlying topological structure characterizing a ``real-world'' data set. Indeed, this is just one of the many applications of multidimensional persistence. As discussed, one of the main issues inhibiting the use of rank invariant in practice is computation. In Table~\ref{tab:barcode_templates}, we record a variety of statistics related to the computation of the rank invariant on the bifiltration constructed from the natural images data set. 
TODO: finish explanation
%An increasingly refined

\begin{table}[h]
\caption{Cost to compute all barcode templates $\mathcal{T}$ }\label{tab:barcode_templates}
\centering
\begin{tabular}{ |P{1.6cm}|P{1.6cm}|P{1.2cm}|P{2.8cm}|P{2.4cm}|}
 \hline
 %\multicolumn{7}{|c|}{ Size and cost of constructing $\mathcal{A}^\ast(M)$ } \\
 %\hline
  \multicolumn{2}{|c|}{ $\xi$ and $\mathcal{A}(M)$ } & \multicolumn{3}{c|}{ Column reductions / Permutations } \\
 \hline
 Coarsening & \# 2-cells & \texttt{phCol} & Vineyards / $d_{K}$ & Moves / $d_{\text{LCS}}$ \\
 \hline
 8 x 8 & 39 & 94.9K & 245K / 1.53M & 38.0K / 11.6K   \\
 \hline
 12 x 12 & 127 & 318K & 439K / 2.66M & 81.9K / 33.0K \\
 \hline 
 16 x 16 & 425 & 1.07M & 825K / 4.75M & 114K / 87.4K \\
 \hline 
  20 x 20 & 926 & 2.32M & 1.15M / 6.77M  & 148K / 154K\\
 \hline 
  24 x 24 & 1.53K & 3.92M & 1.50M / 8.70M & 184K / 232K  \\
   \hline
 \end{tabular}
\label{table:dual_graph_costs}
\end{table}

%[1,]         8   94945  245297  37999
%[2,]        12  318828  438932  81894
%[3,]        16 1073995  824639 114005
%[4,]        20 2328432 1153742 148434
%[5,]        24 3916177 1506480 184125

% permutations  (b, vineyards, moves) 
% 8, 1525826, 11677
% 12, 2657899, 33056
% 16, 4749576, 87417
% 20, 6772406, 154029
% 24, 8697797, 231603

%To obtain family of filtrations, we computed the Rips filtration up to a fixed $\epsilon = 0.20$ at $d = 150$ snapshots in time. The simulation length is normalized such that the duration of the entire simulation is $1$ unit.  Each time point is evenly sampled between $[0,1]$, and 
%simulated 
%We tested our approach against a variety of different methods.
%\begin{figure}[ht]
%	\includegraphics[width=0.95\textwidth, height=2.5in]{prelim_results}
%	\caption{[Placeholder] 	Cumulative number of $O(m)$ operations (left, middle), and the distribution of $K_\tau$ distances between each adjacent permutation (right). The left plot shows the cumulative number of column operations (in log-scale) for various strategies, and the middle plot shows the cumulative number of permutations.}
%	\label{fig:res1}
%\end{figure}

\section{Conclusion and Future Work}\label{sec:conclusion}
In conclusion, we presented a scheduling algorithm for efficiently updating a decomposition in sparse dynamic settings. Our approach is simple, easy to implement, and fully general: it does not depend on the geometry of underlying space nor the choice of triangulation. Moreover, we supplied efficient algorithms for our scheduling strategy, provided tight bounds where applicable, and demonstrated our algorithms performance with several real world use cases.

There are many possible applications of our work beyond the example application of generating crocker stacks and simulating multidimensional persistence. As discussed in sections~\ref{sec:related_work}, examples include: 
\begin{enumerate}
	\item Computing persistent images for time-varying systems 
	\item Optimization procedures involving persistence diagrams
	\item Detecting homological critical points in time-varying filtrations
\end{enumerate}
Indeed, we envision our approach as potentially useful to any situation where the structure of interest is a parameterized family of filtrations and the goal involves computing the persistent homology for each filtration in the family. Areas of particular interest include time-series analysis and dynamic metric spaces~\cite{kim2020spatiotemporal}.

The simple and combinatorial nature of approach does pose some limitations to its applicability. There may be better bounds or algorithms available if stronger assumptions can be made on how the filtration is changing. If the filtration $K_0$ shares little similarity to the ``target'' filtration $K_1$, then the overhead of reducing the simplices from $K_1 \setminus K_0$ appended to the decomposition derived from $K_0$ may be large enough to motivate simply computing the decomposition at $K_1$ independently, especially if parallel processors are available. This overhead does not exist for the multidimensional persistence use-case: the entire decomposition is needed at each template 2-cell.

From an implementation perspective, one non-trivial complication of our approach is its heavy dependence on a particular sparse matrix data structure which permits permuting both the row and columns of a given matrix in at most $O(m)$ time. Our use of this data structure is motivated by the fact that, like vineyards and unlike the standard persistence algorithm, the $R = DV$ decomposition is permuted with every $O(m)$ operation. Indeed, as shown with the natural images example in section~\ref{sec:results}, there are often more permutation operations being applied than there are column reductions. In the more standard \emph{compressed} sparse matrix representations (i.e. CSC, CSR, etc.), permuting both the rows and columns generally takes at most $O(Z)$ time, where $Z$ is the number of non-zero entries. If the particular filtration has many cycles, permuting these compressed representations can be quite expensive, bottlenecking the computation both in theory and in practice. As a result, the more complex sparse matrix representation from~\cite{cohen2006vines} is necessary to be efficient in practice.

Moving forward, our results suggest there are many aspects of computing persistence in dynamic settings yet to be explored. Do existing optimizations to the standard reduction algorithm have natural extensions to dynamic setting? Are there other edit distances that yield efficient algorithms to update $R = DV$ decompositions, beyond moves and transpositions? An ideal method would be one that allows arbitrary permutations to a previously computed decomposition that is provably more efficient than performing the reduction independently in all cases.

%\[
%R_1 = 
%\begin{blockarray}{ccccccc}
%& u & v & w & x & y & z \\
%\begin{block}{c[cccccc]}
%  a & 1 & 1  & 1 &  & & \topstrut \\
%  b &    & 1 &     &  & &  \\
%  c &    &    &  1 &  & &  \\
%  d & 1 &    &     &  & & \botstrut  \\  
%\end{block}
%\end{blockarray}, \quad 
%V_1 = 
%\begin{blockarray}{ccccccc}
%& u & v & w & x & y & z \\
%\begin{block}{c[cccccc]}
%  u & \, 1 & 1  & 1 & 1 & & 1 \, \topstrut \\
%  v &    & 1 &     & 1 & 1 &  \\
%  w &    &    &  1 &  & 1 & 1  \\
%  x &  &    &     &  1 & &  \\  
%  y &  &    &     &  & 1 &  \\  
%  z &  \, &    &     &  & & 1 \, \botstrut \\  
%\end{block}
%\end{blockarray}
%\]
%
%\subsection*{To remove: Rank aggregation}
% \sum\limits_{1 \leq i \leq j \leq n} \mathbbm{1}_{\sigma(i) < \sigma(j)} \land  \mathbbm{1}_{\tau(i) < \tau(j)} 
%$$ F(\sigma, \tau) = \sum\limits_{i \in [n]} \lvert \sigma(i) - \tau(i) \rvert $$ 
%
%Given a fixed permutation $\sigma$ and a multiset $\mathcal{P} = \{ p_1, p_2, \dots, p_k \}$, the \emph{rank aggregation} problem is the following: 
%$$ \argmin_{\sigma^\ast \in S_n} \sum\limits_{i = 1}^k K(\sigma^\ast, p_i) $$
%An ordering $\sigma^\ast$ satisfying the above is said be a \emph{Kemeny optimal} ordering. The problem is NP-hard. However, a variety of approximations are available, and there is an $O(n^3)$ time algorithm for computing the the optimal spearman ranking:
%$$  \argmin_{\sigma^\ast \in S_n} \sum\limits_{i = 1}^k F(p_i, \sigma^\ast) $$ 
%A famous result from Diaconis et al.~\cite{} is that following inequality relating the Spearman footrule distance and the Kendall-$\tau$ distances:
%$$ K(\sigma, \tau) \leq F(\sigma, \tau) \leq 2 K(\sigma, \tau)$$
%The Kendall-$\tau$ distance between two permutations of size $n$ can be computed in $O(n \log n)$, while the Spearman footrule distance can be computed in $O(n)$ time. Weighted versions of both exist, as well as proofs that such versions are valid distance metrics~\cite{}. Additionally, the Kendall distance lends to a positive definite kernel on $\mathbb{S}_n$.
%\noindent Given two permutations $p, q \in S_n$, let $\mathcal{S}(\sigma) = ( s_1, s_2, \dots, s_d )$ denote a sorting that achieves $p \mapsto q$. By the reduction to the permutation edit distance, we know that $d = n - \lvert \mathrm{LCS}(p, q) \rvert$.
%Let $\hat{p}_i = p \sbullet s_1 \sbullet s_2 \sbullet \cdots \sbullet s_i$ denote the composition of the first $i$ permutations for some given schedule $\mathcal{S}_{\sigma}$. The objective is to minimize the following:
%\begin{displaymath}
%\mathcal{S}_{\sigma^\ast} = \argmin_{\sigma \in S_d} \sum\limits_{i = 1}^{d - 1} F(\hat{p}_i, \hat{p}_{i+1})
%\end{displaymath}
%Each permutation $\sigma \in S_d$ generates an order of symbols to move to transform $P \mapsto Q$, so for each $\sigma \in S_d$ we have a unique $\mathcal{P}_\sigma$. This objective can be interpreted as finding a schedule which minimizes the total displacement distance. 
%\[
%\scalefont{1.8}{
%\left(
%\begin{smallmatrix}{| *19c |}
%	\cline{1-7}
%	1 & \cdots & i -1 & i & i + 1 & \cdots & j - 1 & j & j +1 & \cdots & k - 1  & k & k+1 & \cdots & l - 1 & l & l + 1 & \cdots & n \\
%	\cline{1-7}
%	1 & \cdots & i -1 & i & i + 1 & \cdots & j - 1 & j & j + 1 & \cdots & k - 1  & k & k + 1 & \cdots & l-1 & l & l + 1 & \cdots & n
%\end{smallmatrix}
%\right)
%}
%\]
%
%\begin{gather*}
%    A = \left[\,\begin{matrix} a & b & c \\ d & e & f \\ g & h & i \end{matrix}\,\right]
%\end{gather*}
%
%\scalefont{0.9}{
%\begin{gather*}
%	\setcounter{MaxMatrixCols}{20}
%    A = \left[\,\begin{matrix} 1 & \cdots &\mkern-15mub i -1 & i & i + 1\cdots & j - 1 & j & j + 1 & \cdots & k - 1  & k & k+1 & \cdots & l - 1 & l & l + 1 & \cdots & n \end{matrix}\,\right]
%\end{gather*}}

%\begin{displaymath}
%	\setcounter{MaxMatrixCols}{20}
%	\left( \begin{matrix} 
%	1 & \cdots & i -1 & i & \cdots & j - 1 & j & j + 1 & \cdots & k - 1  & k & \cdots & l & l + 1 & \cdots & n
%	\end{matrix}\right)
%\end{displaymath}

%
%\subsubsection*{Move Schedules}
%
%First, assuming a fixed order of $d$ simplices to move is imposed. We need to characterize how many possible schedules transforming $P_s$ into $P_t$ are valid. This is dependent on the particular choice of $\mathrm{LCS}(P_s, P_t)$. 
%
%
%In what follows, we always  use a triangulation $K$ of the manifold and a filtration that changes with time. 
%
%\noindent Let $(f_1, f_2)$ denote the function values associated with two chain complexes $(K_{1}, K_{2})$ filtered according to $(f_1, f_2)$. Assume that if $\sigma \in K_1 \implies \sigma \in K_2$, i.e. $K_1$ and $K_2$ both contain the same simplices, and thus $K_2$ represents some permutation of the order given on $K_1$. $\dots$. 
%Suppose the simplices in $K_1$ are labeled according to $[n]$ and the simplices in $K_2$ are given labels corresponding to the matched simplices in $K_1$. Denote these labelings with $P_1$ and $P_2$.  
%\\
%\\
%\noindent One may perform the transformation $P_1 \mapsto P_2$ via a set of \emph{move} operations. Minimizing the number of such operations reduces to computing the \emph{permutation edit distance}. The permutation edit distance $d(P_1, P_2)$ is defined as the minimum number of moves necessary to transform $P_1$ into $P_2$. Computing the distance reduces computing a Longest Common Subsequence (LCS) between two permutations. Like the Levenshtein edit distance on strings, any set of move operations which preserves the LCS yields a set of edit operations transforming $P_1$ to $P_2$. 
%
%Since the objects of interest here are permutations, computing $LCS(P_1, P_2)$ reduces to computing a Longest Increasing Subsequence (LIS) of a single permutation. The reduction comes from the fact that the $LCS(P_1, P_2)$ is invariant under re-labeling: for any pair of permutations $P_1, P_2 \in S_n$, we may re-label any entries in $P_1$ as long as we perform the same re-labeling in $P_2$. If $P_1$ is re-labeled to the identity permutation, it is a strictly increasing sequence, then the LCS between $P_1$ and $P_2$ corresponds to the LIS of $P_2$. 
%
% This is not a unique computation---for a given permutation, their may be several LIS's with the same cardinality---however it may be shown that edit distance derived from a LIS is indeed optimal. Computing a LIS can be done in $O(n \log n)$ time via a dynamic programming strategy. Given a permutation $P$ of length $n$ and a corresponding LIS of length $k$, there are exactly $n - k$ move operations needed to transform $P$ to the identity. There are many such sequences of move operations---any move that is said to \emph{preserve} the LIS of the permutation is a valid operation, and after exactly $k$ such operations the identity permutation is recovered. 
%\\
%\\
%\noindent Any sequence of move operations transforming $P_1 \mapsto P_2$ that preserves $\mathrm{LCS}(P_1, P_2)$ has a cardinality of $d = d(P_1, P_2) = n - k$, where $k$ is the length of the given $\mathrm{LCS}(P_1, P_2)$. By interpreting the vineyards problem in the context of edit distances, we can guarantee we need only $d$ operations for any two permutations $P_1, P_2$, and by reduction to the permutation edit distance, this is indeed optimal. 
%However, unlike the permutation edit distance, the edit operations for the vineyards problem have non-uniform cost and are time-varying---each move $\mathrm{Move}(i,j)$ requires a number of column operations at most proportional to $\lvert i - j\rvert$, and this cost changes as the decomposition is modified. Ideally, we want an ordered set of move operations that permutes the given decomposition $R = \partial V$ into the target one $PRP^T=(P \partial P^T)(P V P^T)$ as efficiently as possible. In what follows, we outline various roadblocks that prevent us from doing this directly, concluding with a heuristic and greedy procedure that works well in practice.    
%
%
%An ordered set of move operations transforming $[n]$ to a given permutation $P$ is called a \emph{move schedule}. 
%
%Without loss of generality, let $P_1 = [n]$ and let $P_2 = P$ be fixed permutations. Denote the length $k$ LIS with $\mathcal{L} = \mathrm{LIS}(P_1, P_2) = \mathrm{LIS}([n], P)$. There are $d = n - k$ move operations move operations needed to transform the identity permutation to $P$. Let $L_1 = ( \ell_1, \dots, \ell_k)$ denote the positions of the elements of $\mathcal{L}$ in $P_1$, and similarly for $L_2$. By definition of the LIS, both $L_1$ and $L_2$ are strictly increasing, and $(P_1[\ell_1], P_1[\ell_2], \dots, P_1[\ell_k])$ represents a LCS between $P_1$ and $P_2$. 
%
%Move operations can thought of two special types permutations: the operation $\mathrm{Move}(i,j)$ where $i < j$ is equivalent to moving the element at position $i$ up to position $j$, shifting down every element between them. Similarly, if $i > j$ then the element at position $i$ is moved down to position $j$, shifting up every element between them. In both cases, every element outside of $[i,j]$ is mapped by the identity. 
%\begin{displaymath}
%\mathrm{Move}(i,j) = 
%\begin{cases}
% 	(1, 2, \dots, i-1, i+1, \dots, j-1, j, i, j+1, \dots, n) & \text{ if } i < j \\ 
% 	(1, 2, \dots, j-1, i,  j, j + 1\dots, i - 1,  i + 1, \dots, n) & \text{ if } i > j
%\end{cases} 
%\end{displaymath} 
%Fix an order of elements to move $\mathcal{S} = I_n \setminus \mathcal{L}$. Let $s \in \mathcal{S}$ represent the first symbol to move. A move sequence $\mathcal{M}_{\mathcal{S}} = ((i_1, j_1), (i_2, j_2), \dots, (i_d, j_d))$ with respect to $(P_1, P_2)$ is said to be \emph{valid} if the set of move operations preserves the LCS with each move operation. There are potentially many such move sequences; to see this, consider partitioning $I_n$ into the following set of equivalence classes based on $L_1$:
%\begin{displaymath}
%	E_1 = \Big( \underbrace{[1, 2, \dots, \ell_1)}_{e_1}, \underbrace{(\ell_1, \ell_1 + 1, \dots, \ell_2)}_{e_2}, \dots, \underbrace{(\ell_k, \ell_k + 1, \dots, n]}_{e_k} \Big ) 
%\end{displaymath}
%If a move operation $\mathrm{Move}(i,j)$ is valid with respect to $(P_1, P_2)$, then $i \notin L_1$ and $j$ must lie in one of the classes given by $E_1$. 
%%Per the discussion above, the set $\mathrm{LCA}(P_1, P_2) = (P_2[\ell_1], P_2[\ell_2], \dots, P_2[\ell_k])$ forms a common subsequence between $P_1$ and $P_2$, any edit (move) operation must \emph{preserve} this LCS upon performing the edit. 
%After performing the edit, we obtain a new permutation $P_1'$ with $\lvert \mathrm{LCS}(P_1', P_2) \rvert = \lvert \mathrm{LCS}(P_1, P_2) \rvert + 1$. The new set $L_1'$ representing the positions of the computed $\mathrm{LCS}(P_1', P_2)$ forms a new set of equivalence classes of possible moves: 
%\begin{displaymath}
%	E_1' = \Big( [1, 2, \dots, \ell_1), (\ell_1, \ell_1 + 1, \dots, \ell_2), \dots, (\ell_{k+1}, \ell_{k+1} + 1, \dots, n] \Big ) 
%\end{displaymath}
%The goal is to find the ordered set of move operations with minimal cost. 
%\\
%\\
%\noindent
%To make this optimization more concrete, consider performing a single move operation with symbol $s$. Three question come to mind: 
%\begin{enumerate}
%	\item Given a symbol to move, how expensive is the corresponding edit operation? 
%	\item Given a set of symbols to move, how expensive is it to find the symbol whose edit operation has minimal cost? 
%	\item Suppose questions (1) and (2) are answerable. How viable would a greedy-type algorithm be?
%\end{enumerate}
%For a given symbol $s$ at position $i$, we know what position we wish to move the symbol to $j$. While $i \notin L_1$, $j$ may or may not be in $L_1$. The target position $j$ defines the equivalence class $c_j$ of possible move operations. Each move operation creates a new equivalence class of remaining positions, and the move operations continue until $\lvert \mathrm{LCS}(P_1, P_2) \rvert = n \Rightarrow P_1 = P_2$. Viewed from the equivalence class perspective, this process of splitting a set into equivalence classes is synonymous with the \emph{partition refinement} technique.  
%\\
%\\
%\noindent \textbf{Defining the cost of a move}:  The cost of a move operation $\mathrm{Move}(i,j)$ depends on the direction of the movement. If $i < j$, each non-zero entry in the $i^\mathrm{th}$ row of $V$ with a column index in $(i,j]^{\mathrm{th}}$ contributes a column addition in both $R$ and $V$. If $i > j$, similarly, non-zero entries in the $j^{th}$ column of $V$ between rows $[i,j)$ determine the cost of the operation, although the exact cost may be less. Given a decomposition $\partial V = R$ of $m$ simplices stored using a sparse matrix representation, determining the cost of a move right operation takes $O(\ell\log(m))$ time where $\ell = \lvert i - j \rvert$. Determining move left is more expensive, since column operations are required, but can be bounded from above by $O(\ell m )$, since each column operation takes $O(m)$ time, and there are (again) at most $\ell = \lvert i - j \rvert$ such operations needed. 
%
%%Let $s$ be some fixed symbol to move. A move operation $m_s$ is said to be \emph{free} if its corresponding permutation matrix $P$ induces no column operations in the decomposition $(P\partial P^T)(P V P^T) = PRP^T$. When is a move free? 
%%If moving left and all of the entries in rows $[i,j)$ or column $i$ of $V$ are zero, then the movement is free. If moving right and all the entries in columns $(i, j]$ of row $i$ of $V$ are zero, then the movement is free. 
%
%
%\subsection*{Enumerating move schedules}
%Consider the space of all possible move sequences. 
%
%\begin{algorithm}
%    \caption{Generate Move Schedules}
%    \label{ms}
%    \begin{algorithmic}[1] 
%    \Require $p := \text{ source permutation }, q := \text{ target permutation }$
%    \Require $\tau := \mathrm{LCS}(p, q)$ 
%        \Procedure{GenerateMS}{$p$, $q$, $\tau$, $S = \emptyset$}
%        	\If{$p = q$} 
%        	\State $\text{ Report } S \text{ as valid schedule}$ \Comment{ Base case } 
%        	\Else 
%        		\For{$e \in p \setminus \tau$} 
%        			\State $I \gets \mathrm{ValidMoves(e, p, q, \tau)}$ \Comment{ Any move that preserves $\mathrm{LCS}(p, q)$ is valid }
%        			\For{$(i, j) \in I$} 
%        			% \State $(i,j) \gets \mathrm{isolate\_move}(e, P_s, P_t)$
%        			\State $p' \gets \mathrm{Move}(p, i, j)$
%        			\State $\tau' \gets  \mathrm{order}_{<_p}(\tau \cup e)$ 
%        			\State $\mathrm{GenerateMS(p', q, \tau', S \cup (i,j))}$	
%        			\EndFor
%        		\EndFor 
%        	\EndIf
%%            \State $r\gets a \bmod b$
%%            \While{$r\not=0$} \Comment{We have the answer if r is 0}
%%                \State $a \gets b$
%%                \State $b \gets r$
%%                \State $r \gets a \bmod b$
%%            \EndWhile\label{euclidendwhile}
%%            \State \textbf{return} $b$\Comment{The gcd is b}
%        \EndProcedure
%    \end{algorithmic}
%\end{algorithm}
%\noindent How many possible move sequences are there? 
%\\
%\\
%\noindent 
%Suppose $d = n - k = n - \lvert LCS(P_1, P_2) \rvert$. Fix and order of symbols to move. Each symbol $s$ can move to any set of positions that preserves the LCS between the two permutations. Call this set $I_s$. After $s$ is moved, the size of the LCS increases by 1, and $I_s$ decreases by 1. If each symbol $s \in  P_s \setminus \mathrm{LCS}(P_s, P_t)$ contains only one possible movement, then there is only one such move sequence that preserves the LCS. Alternatively, suppose every symbol to move has the identical target set $I$. Before the first move, $\lvert I \rvert = d$, and there are $d$ possible choices for the first symbol $s^{(1)}$ to move to. After $s^{(1)}$, $I$ decreases by one, and the next symbol to move has $d - 1$ locations that are valid. Thus, given a fixed ordering of symbols to move, there is at least $1$ and at most $d!$ possible move sequences that transform $P_s \mapsto P_t$ in $d$ move operations. This accounts for (1). 
%\\
%\\
% \noindent There are also $d!$ number of ways to order the move set $S = P_s \setminus \mathrm{LCS}(P_s, P_t)$, which accounts for (2). 
%\\
%\\
% \noindent How large is the LCS, $k$, expected to be? if $P_s$ and $P_t$ are random permutations of size $n$, then it turns out that $k \sim O(\sqrt{n})$. In fact, the probability that $k \approx \sqrt(n)$ is not just the expected value, but the length of the LCS that is attained with with probability 1 as $n \to \infty$. See [1] for more details. 
% \\
% \\
% \noindent Given any LCS of size $k$, it is true that any move sequence that preserves said LCS under each operation has size exactly $n - k$. However, even if one could obtain an optimal move sequence for a given LCS, it's worth asking whether or not that cost changes if the LCS changes. The answer is ``yes'': since individual move operations have non-uniform cost, it's possible that the optimal sequence of moves produced by one LCS is suboptimal relative to another LCS. This begs the question of how many LCs  
%
%
%
%\subsection*{Applications:} \emph{Moments} i.e. summaries of persistence diagrams, CROCKER stacks, multidimensional persistence

\bibliography{move_schedules}
\bibliographystyle{ieeetr}

\section*{Appendix}\label{sec:appendix}
\appendix

\subsection*{Vineyard Algorithm}
\begin{algorithm}[ht]
	\caption{Vineyards: Tranposition Framework}\label{alg:tr}
	\setstretch{1.15}
	\begin{algorithmic}[1]
		\Require Matrices $R, V$ satisfying $R = D V$, $1 \leq i \leq m - 1$
		\Ensure Output $(R, V)$ maintains the decomposition invariants
        \Function{Transpose}{$R$, $V$, $i$}
            \State $\mathrm{pos} \gets $ columns satisfying $\mathrm{col}_R = 0$
            \If {$\mathrm{pos}[i]$ \textbf{and} $\mathrm{pos}[i+1]$}
	            \If{$V[i,i+1] \neq 0$}
            		\State $\mathrm{col}_V(i+1)  \mathrel{+}= \mathrm{col}_V(i)$	
            	\EndIf
            	\If {$\exists \; k, l$ satisfying $\mathrm{low}_R(k) = i, \mathrm{low}_R(l) = i+1$ and $R[i,l] \neq 0$} \Comment{O(m)}
            		\If {$k < l$}
            			\State \Return $(\, R, \, V \,) \gets (\, PRPS_k^l, \, PVP S_k^l \,)$
            		\Else
            			\State \Return $(\, R, \, V \,) \gets  (\, PRPS_l^k, \, PVP S_l^k\,) $
            		\EndIf
            	\EndIf 
            \ElsIf{$!\mathrm{pos}[i]$ \textbf{and} $!\mathrm{pos}[i+1]$}
            	\If{$V[i,i+1] \neq 0$}\Comment{O(m)}
            		\If{$\mathrm{low}_R(i) < \mathrm{low}_R(i+1)$}
            			\State \Return  $(\, R, V \,) \gets (P R S^{i+1}_i P, \, P V S^{i+1}_i  P) $
            		\Else 
            			\State \Return  $(\, R, V \,) \gets (P R S^{i+1}_i P S^{i+1}_i, \, P V S^{i+1}_i P S^{i+1}_i ) $
            		\EndIf 
            	\EndIf
           	\ElsIf{$!\mathrm{pos}[i]$ \textbf{and} $\mathrm{pos}[i+1]$} 
            	\If{$V[i,i+1] \neq 0$}\Comment{O(m)}
            		\State \Return  $(\, R, V \,) \gets (P R S^{i+1}_i P S^{i+1}_i, \, P V S^{i+1}_i P S^{i+1}_i ) $
            	\EndIf
            \ElsIf{$\mathrm{pos}[i]$ \textbf{and} $!\mathrm{pos}[i+1]$}
            	\If{$V[i,i+1] \neq 0$}\Comment{O(m)}
            		\State $\mathrm{col}_V(i+1) \mathrel{+}= \mathrm{col}_V(i)$
            	\EndIf 
            \EndIf
            \State \Return $(\, R , V \,) \gets (\, P R P , P V P \,)$
        \EndFunction
    \end{algorithmic}
\end{algorithm}

\subsection*{Transposition complexity details}
Consider Algorithm~\ref{alg:tr}. The matrices $R, V$ are both $(m \times m)$, where $m$ is the number of simplices in the underlying filtration. Each transposition $\mathrm{Tr}(i, i+1)$ requires at most 2 column operations (denoted with $S_{i}^{j}$ in the algorithm) in both $R$ and $V$, a constant number of queries to the matrix representation, and exactly 1 row and column exchange. Column operations take $O(m)$ time. Querying to determine whether an entry at position $(i,i+1)$ is non-zero depends on the sparse matrix implementation, but is generally either $O(\log m)$ or $O(m)$. In compressed sparse matrix representation, column and row exchanges take upwards of $O(z)$ time, where $z$ is the number of non-zero entries in the matrix, however this can be made $O(1)$ time by using the data structure described in~\cite{cohen2006vines}. This data structure requires storing an index using at most $O(m \log m)$ bits with each non-zero entry, as well as several $O(m)$-sized auxiliary arrays to protect the matrices from row exchanges.

%The principle motivation  
%Upon moving $s$ 
%
%Denote this class with $c_s$.
%Now if the position of $s$ in $P_1$ is $i$ and the position in $P_2$ is $j$, and $i < j$, then the

%% Note: patience sorting gives means of enumerating all LIS's
%% Note #2: Consider having the identity permutation be the target and going in reverse--at the end, there is only one symbol to move into place
%% prims-MST like approach possible?

%\begin{definition}[Coarsening]
%	Given a schedule $\mathcal{S} = ( s_1, s_2, \dots, s_h)$ of permutations, a \textbf{\emph{coarsening}} of $\mathcal{S}$ is a schedule $\widetilde{S} = ( \tilde{s}_1, \tilde{s}_2, \dots, \tilde{s}_l)$ obtained by collapsing contiguous sequences in $S$ via the map
%	$ (t_i, t_{i{+}1}, \dots, t_j) \mapsto (m_{ij}) $
%	where, in cycle notation, $t_i$ is a permutation of the form $(i, i{+}1)$, and $m_{ij}$ is a permutation of the form $(i{+}1, \dots, j, i)$.
%\end{definition}
%\begin{lemma}
%	For some given decomposition $R = \partial V$ where each matrix is $(m \times m)$, let $C$ denote the number of $O(m)$ operations required to execute schedule $\mathcal{S} = (t_1, t_2, \dots, t_h)$. If a coarsened schedule $\widetilde{S}$ of $\mathcal{S}$ takes $K$ operations requiring $O(m)$, we have:
%	$$ \frac{C}{2} \leq K \leq C$$ 
%\end{lemma}
%\begin{proof}
%	Consider Algorithms~\ref{alg:tr} and~\ref{alg:mr}, where it's assumed Algorithms~\ref{alg:tr} is used to execute a given schedule $\mathcal{S}$ and Algorithms~\ref{alg:mr} is used to execute its coarsened schedule $\widetilde{\mathcal{S}}$. If maps are created providing $O(1)$ access to the low entries needed by lines (5) and (12), and the non-zero entries in each column are sorted according to some fixed providing $O(\log(m))$ time for lines (5), (11), (17), and (20), then the dominating cost of each transposition in Algorithm~\ref{alg:tr} are the column operations, each of which take $O(m)$ time. There are at most $2$ column operations for any given case. For any contiguous sequence of transpositions $(t_i, t_{i{+}1}, \dots, t_j)$, this implies we have at most $2\lvert i - j \rvert$ operations each requiring $O(m)$ time. Now consider a single move operation, Move($i$,$j$), as a replacement for the sequence of transpositions above. Identifying non-zero entries in the matrix occurs once, which takes $O(\lvert i - j \rvert \log(m))$ time. If we maintain a map providing low entries in $O(1)$ time, since permutations take $O(1)$ time and lines (7) and (7) in both Restore* and Move* takes just $O(1)$ via pointer swapping, the dominant cost again are the column operations (line 5). Since a move operation requires at most a single column operation for each index between $[i, j]$, the complexity of $\mathrm{Move}(i,j)$ is bounded above by $O(\lvert i - j \rvert m)$, and the claimed inequality follows. 
%\end{proof}
%\noindent \textbf{Free Schedules:} Previously, we assumed one had a homotopy $F(x,\tau) : \mathbb{X} \times [0,1] \to \mathbb{R}$ mapping the functions associated with two fixed filtrations $K_0, K_1$ each of size $m$, and we showed that simulating persistence dynamically reduced to determining a set of ordered ``crossings.'' If these crossings occur at time points $(\tilde{t}_1, \tilde{t}_2, \dots, \tilde{t}_m)$, then obtaining the updating decomposition at all times $t_i$ where $\tilde{t}_i < t_i < \tilde{t}_{i{+}1}$ fully characterizes how the persistent homology changes across the homotopy. If we again assume both general position and that the curves traced out by $F$ intersect at most once, then we obtain an ordered schedule of transpositions $\mathcal{S}_F$. The size of this schedule is equal to the number of inversions between $K_0$ and $K_1$, which is at most $O(m^2)$.
%
%There are many applications wherein a practitioner is interested some notion of dynamic persistence, but only at specific time points---how the persistence changes between these time points may not be needed. More generally, it's been observed in practice that many applications of persistence homology involve computing persistence across discrete 1-parameter family of (usually related) filtrations. One example is that one has set of point clouds sampled from some unknown continuous process at discrete time points $\mathcal{T} = t_1, t_2, \dots, t_n$, and one wishes to compute the persistence homology of some filtration derived from the geometry of each point cloud. One solution is to assume a straight-line homotopy $F(x, \tau)$ between each adjacent filtrations $K_{t_i}, K_{t_{i {+} 1}}$ and proceed using any of the methods described above. 
%
%If one \emph{only} needs the persistence diagrams at exactly the given time points $\mathcal{T}$, i.e. the intermediate persistence diagrams between each $K_{t_i}, K_{t_{i {+} 1}}$ aren't required, then intuitively we ought to coarsen each schedule as much as possible to reduce the number of diagrams generated. Moreover, since we can control the ordering of any such schedule altogether, it's worth exploring the degree to which a schedule can be coarsened under \emph{any} reordering. We call a schedule $\mathcal{S}$ that can be arbritrarily reordered a \emph{free schedule}.
%
%\subsection{Minimizing schedule cost}\label{sec:schedule_cost}
%\textbf{Motivation:}
%Using the notation presented in the previous section, we now have two goals to consider: 
%\begin{enumerate}
%	\item Given an ordered schedule, can we produce a more efficient schedule using \emph{move} operations?
%	\item If we allow any order of transpositions or moves, can we product an optimally efficient schedule?
%\end{enumerate}
%We address each of these in order.
% \textbf{Example: Coarsening around homological critical points}
%( Insert more on detecting homological critical points, coarsening around critical points, etc. )
% Although move operations can be thought of as a generalization of the rules given for the vineyards algorithm, there is a significant departure philosophically between what the two algorithms seek to accomplish. The vineyards algorithm seeks to provide an efficient solution at computing persistent homology across a continuous, time-varying filtration, and to detect homological critical values as they arise. Moves, on the other hand, were originally designed to speed up simplex insertions and deletions into a filtration\footnote{In fact, the move algorithm was originally introduced solely to assist an accompanying algorithm which seeks to track a homology generator stably over time, see~\cite{busaryev2010tracking}.}; there is no focus on detecting critical values. 
\end{document}

### 2021-08-04 move_schedules.tex
Updates from Overleaf
Testing github overleaf merge
Reworded the intro paragraphs to multidimensional persistence

@@
 \subsection*{Multidimensional persistence}
-A natural question is whether or not there exists a multi-dimensional generalization of persistence.
-The starting point for this generalization is a structure which filters the space in multiple dimensions simultaneously, a \emph{multifiltration}.
-Given this structure, the goal of multi-dimensional persistence is to identify persistent features by examining the entire multifiltration.
-Such a generalization has been thought to have enormous potential in terms of applications~\cite{}.
+A natural question is whether or not there exists a multidimensional generalization of persistence.
+Given a structure which filters the space in multiple dimensions simultaneously, a \emph{multifiltration}, the goal of multidimensional persistence is to identify persistent features by examining the entire multifiltration.
+Such a generalization has appeared naturally in many application contexts, showing potential as a tool for exploratory data analysis~\cite{lesnick2012multidimensional}.
 For example, one of the often noted drawbacks of persistence is its instability with respect to strong outliers.
-That is, although persistence diagrams are stable with respect to their input and thus are robust to noise, the presence of outliers can artificially introduce topological features such as loops, obscuring the persistence of significant topological features~\cite{buchet2015topological}.
-Indeed, this obfuscation can hinder the ability of 1D persistence to uncover ``significant'' topological features, one of the original purposes of persistence.
-In theory, this formalism enables the practitioner to analyze even noisy data sets directly without resorting to sophisticated data preprocessing techniques which may or may not preserve the underlying topology of the data.
-
-Unfortunately, multi-dimensional persistent homology comes with its own set of new computational and theoretical challenges not necessarily inherited by the 1-D case.
-The multi-dimensional persistent homology of a multifiltration cannot be captured by complete discrete invariant, and incomplete invariants have proven challenging to use in practice due to the high algorithmic complexity of their corresponding computation.
-A number of computational strategies have been proposed to counter this, such as homology preserving multi-filtration simplification or parallelization schemes using shared-memory computation~\cite{fugacci2019chunk, lesnick2019computing}.
-Another strategy is to focus specifically on a fixed dimension, such as 2-D persistence.
-Utilizing the equivalence between the rank and fibered barcode invariants, Lesnick and Wright~\cite{lesnick2015interactive} developed an elegant way of computing the former via a reparameterization of the latter using standard point-line duality.
-This clever reparameterization effectively reduces the rank invariant computation to sequence of 1-D barcode computations at ``template points'' lying within the 2-cells of particular planar subdivision of the half-plane $[0, \infty) \times \mathbb{R}$.
+Although persistence diagrams are stable with respect to their input and thus are robust to noise, the presence of strong outliers can artificially modify the connectivity of the underlying filtration, obscuring the detection of what would otherwise be significant topological structures~\cite{buchet2015topological}.
+Indeed, since inferring the homology of some underlying topological space around which data are distributed was one of the original purposes of persistence\footnote{One of the first uses of persistence was to infer the topology of attractors in dynamical systems from experimental data\cite{perea2018brief}.}, strong outliers pose a significant barrier to the widespread adoption of persistent homology.
+By filtering the space according to e.g.
+some notion of density, multidimensional persistence enables the practitioner to handle outliers directly, without resorting to sophisticated data preprocessing techniques which may or may not preserve the underlying topology of the data.
+
+Unfortunately, multidimensional persistent homology comes with its own set of computational and theoretical challenges not necessarily inherited by the 1-D case.
+There is not complete discrete invariant for multidimensional persistence, and the computations associated to other incomplete invariants exhibit high algorithmic complexity.
+A number of computational strategies have been proposed to counter this, such as homology preserving multifiltration simplification or parallelization schemes using shared-memory computation~\cite{fugacci2019chunk}.
+Rather than attempting to tackle multiple dimensions simultaneously, another approach is to focus on a fixed dimension, such as 2-D persistence.
+Utilizing the equivalence between the rank and fibered barcode invariants, Lesnick and Wright~\cite{lesnick2015interactive} developed an elegant way of visualizing the former via a reparameterization of the latter using standard point-line duality.
+This clever reparameterization effectively reduces the fibered barcode computation to a sequence of 1-D barcode computations at ``template points'' lying within the 2-cells of particular planar subdivision of the half-plane $[0, \infty) \times \mathbb{R}$.
 Indeed, since algorithm~\ref{alg:schedule} was designed for precisely such a computation, the algorithm proposed by~\cite{lesnick2015interactive} is prototypical of the class of methods that stand to benefit from move scheduling.
 
-The purpose of the following sections is to uncover the intrinsic bottlenecks impeding the computation of the rank invariant in practical settings.
+The purpose of the following sections is to uncover the intrinsic bottlenecks impeding the computation of the fibered barcode invariant in practical settings.
 Towards this goal, we briefly summarize the theory established by several researchers~\cite{lesnick2015interactive, carlsson2010computing, landi2014rank}, focusing on their computational and algorithmic details.
-Our goal is to show that the highest complexity subcomputation of the rank invariant algorithm proposed by~\cite{lesnick2015interactive} is precisely the type of computation we focus on in this effort.
+Our goal is to show that the highest complexity subcomputation of the fibered barcode invariant algorithm proposed by~\cite{lesnick2015interactive} is precisely the type of computation we focus on in this effort.
 In doing so, we substantiate our claims from section~\ref{sec:motivation} that our scheduling approach can improve the efficiency of a real-world application of multidimensional persistence by an order of magnitude by comparing Algorithm~\ref{alg:schedule} to several alternatives on a real-world data set and problem: the problem of identifying the underlying topological space of a data set extracting from high-contrast natural images.
@@
 It is shown in~\cite{lesnick2019computing} how to compute (1) in approximately $\approx O(m^3)$ using a matrix algorithm similar to Algorithm~\ref{alg:reduce}.
-Constructing and storing the line arrangement $\mathcal{A}(M)$ with $n$ lines and $k$ vertices can be achieved using known algorithms in computational geometry (e.g.
-the Benteley-Ottmann algorithm) in output-sensitive $O((n+k) \log n)$ time and $O(\kappa^2)$ memory, where $\kappa$ is a coarseness parameter associated with the computation of $\xi(M)$.
+Constructing and storing the line arrangement $\mathcal{A}(M)$ with $n$ lines and $k$ vertices is related to the \emph{line segment intersection problem}, which known algorithms in computational geometry can solve in output-sensitive $O((n+k) \log n)$ time (see~\cite{boissonnat2000robust} for an overview).
+The number of 2-cells in $\mathcal{A}(M)$ and $O(\kappa^2)$ memory, where $\kappa$ is a coarseness parameter associated with the computation of $\xi(M)$.
 Finally, the computation of the barcodes templates $\mathcal{T}$ (step 3) takes on the order of $O(m^3 \kappa + m \kappa^2 \log \kappa)$ elementary operations and $O(m \kappa^2)$ storage.

### 2021-08-04 move_schedules.tex
Reworked and removed rank invariant intro paragraphs, small changes to fibered barcode paragraph. Switching notation of bigraded betti numbers!

@@
 For example, one of the often noted drawbacks of persistence is its instability with respect to strong outliers.
-Although persistence diagrams are stable with respect to their input and thus are robust to noise, the presence of strong outliers can artificially modify the connectivity of the underlying filtration, obscuring the detection of what would otherwise be significant topological structures~\cite{buchet2015topological}.
+Although persistence diagrams are stable with respect to their input and thus are robust to noise, the presence of strong outliers can artificially modify the connectivity of the underlying filtration, obscuring the detection of what wogiuld otherwise be significant topological structures~\cite{buchet2015topological}.
 Indeed, since inferring the homology of some underlying topological space around which data are distributed was one of the original purposes of persistence\footnote{One of the first uses of persistence was to infer the topology of attractors in dynamical systems from experimental data\cite{perea2018brief}.}, strong outliers pose a significant barrier to the widespread adoption of persistent homology.
@@
 % These transforms end up being crucial in decomposing the definition of the fibered barcode (and thus, the rank invariant) into a discrete computation.
-
-\subsubsection*{The Rank Invariant}
-
 %We begin with a few definitions.
@@
 The set $\mathcal{B}(M)$ is called the \emph{barcode}\footnote{Note that \emph{barcodes} and \emph{persistence diagrams} are interchangeable representations of the same multiset.} of $M$.
-
+%\emph{The rank invariant} proposed by Carlsson et al~\cite{carlsson2009theory} is one such invariant that is complete when $n=1$ but is otherwise incomplete when $n \geq 2$. 
+%% The rank invariant is computable in polynomial time~\cite{carlsson2010computing}, that has been the subject of much interest lately. 
+%It is defined as follows: 
+%\begin{definition}[Rank Invariant]
+%	For $n \geq 1$, the \emph{rank invariant} of a $\mathbb{R}^n$-indexed persistence module $M$ over $\mathcal{H}^{n} \subseteq \mathbb{R}^n \times \mathbb{R}^n$ is the function:
+%	\begin{equation*}
+%		\begin{aligned}
+%		& \mathrm{rank}(M)  \; : & \mathcal{H}^n & \to \mathbb{N}  \\
+%		& & (a,b) & \mapsto \mathrm{rank} \, M(a,b) 
+%		\end{aligned}
+%	\end{equation*}
+%	for $a \leq b \in \mathcal{H}^n$. When $n = 1$, the rank invariant is complete, thus $\mathrm{rank}(M)$ and the persistence barcodes $\mathcal{B}(M)$ determine each other.
+%\end{definition}
+%\noindent Thus, in one dimension, persistent homology provides a complete invariant that fully captures the lifetime of topological features in a given filtration: 1-D persistent homology modules decompose in an essentially unique way into indecomposable summands~\cite{zomorodian2005computing}.
+In contrast, the theory of multidimensional persistence shows that no complete discrete invariant exists---the structure of the target for the invariant depends on the structure of the underlying field, and although multiparameter persistence modules also decompose in a similar way as the 1-D case, the set of isomorphism classes is known to be extremely complicated~\cite{carlsson2009theory, lesnick2015interactive}.
+%Nonetheless, there are invariants related to multi-parameter persistence that can be quite useful in practice, even though they are incomplete. 
+%\noindent For example, even though the rank invariant does not encode the isomorphism type of the underlying module $M$ when $n > 1$, it does captures important ``first-order'' information about the modules structure.
+ %The multigraded Betti numbers constitute a natural and important class of invariants often studied in algebraic geometry commutative algebra.
+There are many ways of defining these, with perhaps the simplest definition stemming from minimal resolutions: for a finitely generated $n$-parameter persistence module $M$ and $i \in \{0, 1, \dots, d \}$, the $i^{\text{th}}$ graded Betti number of $M$ at grade $z$, denoted as $\beta_i^M(z)$, is defined as the number of elements at grade $z$ in a basis of the $i^{\text{th}}$ module in a free resolution for $M$.
+
+\subsubsection*{The Fibered Barcode}
 An \emph{invariant} of $n$-parameter persistence modules is a function from the collection of persistence modules to some set $S$ with the property that $f(M) = f(M')$ when $M \cong M'$.
@@
 Thus, incomplete invariants are invariants which may have the same value on two non-isomorphic modules.
-\emph{The rank invariant} proposed by Carlsson et al~\cite{carlsson2009theory} is one such invariant that is complete when $n=1$ but is otherwise incomplete when $n \geq 2$. 
-% The rank invariant is computable in polynomial time~\cite{carlsson2010computing}, that has been the subject of much interest lately. 
-It is defined as follows: 
-\begin{definition}[Rank Invariant]
-	For $n \geq 1$, the \emph{rank invariant} of a $\mathbb{R}^n$-indexed persistence module $M$ over $\mathcal{H}^{n} \subseteq \mathbb{R}^n \times \mathbb{R}^n$ is the function:
-	\begin{equation*}
-		\begin{aligned}
-		& \mathrm{rank}(M)  \; : & \mathcal{H}^n & \to \mathbb{N}  \\
-		& & (a,b) & \mapsto \mathrm{rank} \, M(a,b) 
-		\end{aligned}
-	\end{equation*}
-	for $a \leq b \in \mathcal{H}^n$. When $n = 1$, the rank invariant is complete, thus $\mathrm{rank}(M)$ and the persistence barcodes $\mathcal{B}(M)$ determine each other.
-\end{definition}
-\noindent Thus, in one dimension, persistent homology provides a complete invariant that fully captures the lifetime of topological features in a given filtration: 1-D persistent homology modules decompose in an essentially unique way into indecomposable summands~\cite{zomorodian2005computing}.
-In contrast, the theory of multidimensional persistence shows that no complete discrete invariant exists---the structure of the target for the invariant depends on the structure of the underlying field, and although multiparameter persistence modules also decompose in a similar way as the 1-D case, the set of isomorphism classes is known to be extremely complicated~\cite{carlsson2009theory, lesnick2015interactive}.
-Nonetheless, there are invariants related to multi-parameter persistence that can be quite useful in practice, even though they are incomplete. 
-\noindent For example, even though the rank invariant does not encode the isomorphism type of the underlying module $M$ when $n > 1$, it does captures important ``first-order'' information about the modules structure.
-
-Along this vein of research, Lesnick and Wright~\cite{lesnick2015interactive} identify three simple invariants associated with a multidimensional persistence module $M$: the \emph{dimension function of $M$}, the \emph{multigraded Betti numbers of $M$}, and the \emph{fibered barcode of $M$}.
-% The dimension function is simply the the function which maps every point $a \in \mathbb{R}^2$ to $\mathrm{dim}(M_a)$.
+Lesnick and Wright~\cite{lesnick2015interactive} associate three simple invariants to a 2-dimensional persistence module $M$: the \emph{dimension function}, the \emph{multigraded Betti numbers}, and the \emph{fibered barcode}.
+
+The dimension function is simply the the function which maps every point $a \in \mathbb{R}^2$ to $\mathrm{dim}(M_a)$.
 It is a simple and easy to visualize invariant, but is unstable and yields no information about the persistent features of $M$.
-% The multigraded Betti numbers constitute a natural and important class of invariants associated with a multidimensional persistence module.
-There are many ways of defining these, with perhaps the simplest definition stemming from minimal resolutions: for a finitely generated $n$-parameter persistence module $M$ and $i \in \{0, 1, \dots, d \}$, the $i^{\text{th}}$ graded Betti number of $M$ at grade $z$, denoted as $\xi_i^M(z)$, is defined as the number of elements at grade $z$ in a basis of the $i^{\text{th}}$ module in a free resolution for $M$.
-% Thus, the $i^{\text{th}}$-graded Betti number of $M$ is a function $\xi_i(M): \mathbb{R}^n \to \mathbb{N}$.
-Being central objects in commutative algebra and algebraic geometry, both the theory and the computation of multi-graded Betti numbers is extensive and beyond the scope of this effort---the interested reader is referred to~\cite{carlsson2009theory, lesnick2015interactive} for more details.
+The multigraded Betti numbers constitute a natural and important class of invariants often studied in algebraic geometry commutative algebra, though both their theory and computation is beyond the scope of this effort---the interested reader is referred to~\cite{lesnick2015interactive, carlsson2009theory} for more details.
+For our purposes, the the $i^{\text{th}}$-graded Betti number of $M$ is simply a function $\beta_i(M): \mathbb{R}^n \to \mathbb{N}$.
 The fibered barcode, denoted as $\mathcal{B}(M)$, is defined as follows:
@@
 	The fibered barcode $\mathcal{B}(M)$ of a bipersistence module $M$ is the map which sends each line $L$ with non-negative slope to the barcode $\mathcal{B}_L(M)$: 
-	$$ L \mapsto \mathcal{B}_L(M) $$
-	Thus, the fibered barcode is simply a 2-parameter family of barcodes given by the collection of 1-D affine slices of $M$. 
+%	$$ L \mapsto \mathcal{B}_L(M) $$
+$$ \mathcal{B}(M) = \{ \; B_L(M) : L \in \mathbb{R} \times \mathbb{R}^{+} \; \}$$
+	Thus, for a bipersistence modules $M$, the fibered barcode is simply a 2-parameter family of barcodes given by restricting $M$ to a set affine lines in $\mathbb{R}^2$. 
 \end{definition} 
-\noindent Each line $L$ induces a 1-parameter restriction on $M$.
-The fibered barcode enjoys a number of stability properties readily expressed using interleaving distances~\cite{landi2014rank}.
-More important to this work, the fibered barcode $\mathcal{B}(M)$ and the rank invariant $\mathrm{rank}(M)$ determine each other~\cite{lesnick2015interactive}.
-
-\subsubsection*{Fibered barcode reparameterization}\label{sec:fibered_barcode_reparam}
+\noindent Each line $L$ induces a 1-parameter restriction on $M$. The fibered barcode enjoys a number of stability properties readily expressed using interleaving distances~\cite{landi2014rank}. 
+% More important to this work, the fibered barcode $\mathcal{B}(M)$ and the rank invariant $\mathrm{rank}(M)$ determine each other~\cite{lesnick2015interactive}.
+
+\subsubsection*{Reparameterization using point-line duality}\label{sec:fibered_barcode_reparam}
 Let $\overline{\mathcal{L}}$ denote the collection of all lines in $\mathbb{R}^2$ with non-negative slope, $\mathcal{L} \subset \overline{\mathcal{L}}$ the collection of all lines in $\mathbb{R}^2$ with non-negative, finite slope, and $\mathcal{L}^\circ$ the collection of all affine lines in $\mathbb{R}^2$ with positive, finite slope.
@@
 
-Let $S = \mathrm{supp}\,\xi_0(M) \, \cup \, \mathrm{supp}\,\xi_1(M)$, where the functions $\xi_0(M), \xi_1(M)$ are $0^{\text{th}}$ and $1^{\text{st}}$ bigraded Betti numbers of $M$, respectively. 
+Let $S = \mathrm{supp}\,\beta_0(M) \, \cup \, \mathrm{supp}\,\beta_1(M)$, where the functions $\beta_0(M), \beta_1(M)$ are $0^{\text{th}}$ and $1^{\text{st}}$ bigraded Betti numbers of $M$, respectively.
 The main mathematical result from~\cite{lesnick2015interactive} is a characterization of the barcodes $\mathcal{B}_L(M)$, for any $L \in \mathcal{L}$, in terms of a set of \emph{barcode templates} $\mathcal{T}$ computed at every 2-cell in $\mathcal{A}(M)$.
@@
 \begin{enumerate}
-	\item Computing the bigraded Betti numbers $\xi(M)$ of a given bipersistence module $M$ 
+	\item Computing the bigraded Betti numbers $\beta(M)$ of a given bipersistence module $M$ 
 	\item Constructing a line arrangement $\mathcal{A}(M)$ induced by critical lines derived from (1) 
@@
 \end{enumerate}
-It is shown in~\cite{lesnick2019computing} how to compute (1) in approximately $\approx O(m^3)$ using a matrix algorithm similar to Algorithm~\ref{alg:reduce}.
-Constructing and storing the line arrangement $\mathcal{A}(M)$ with $n$ lines and $k$ vertices is related to the \emph{line segment intersection problem}, which known algorithms in computational geometry can solve in output-sensitive $O((n+k) \log n)$ time (see~\cite{boissonnat2000robust} for an overview).
-The number of 2-cells in $\mathcal{A}(M)$ and $O(\kappa^2)$ memory, where $\kappa$ is a coarseness parameter associated with the computation of $\xi(M)$.
-Finally, the computation of the barcodes templates $\mathcal{T}$ (step 3) takes on the order of $O(m^3 \kappa + m \kappa^2 \log \kappa)$ elementary operations and $O(m \kappa^2)$ storage.
-
-Of the three steps above, the barcode template computation is both the highest complexity and most demanding computation required by the algorithm.
+Computing (1) takes approximately $\approx O(m^3)$ using a matrix algorithm similar to Algorithm~\ref{alg:reduce}~\cite{lesnick2019computing}.
+Constructing and storing the line arrangement $\mathcal{A}(M)$ with $n$ lines and $k$ vertices is related to the \emph{line segment intersection problem}, which known algorithms in computational geometry can solve in (optimal) output-sensitive $O((n+k) \log n)$ time (see~\cite{boissonnat2000robust} for an overview).
+In terms of memory complexity, the number of 2-cells in $\mathcal{A}(M)$ is upper bounded by $O(\kappa^2)$, where $\kappa$ is a coarseness parameter associated with the computation of $\beta(M)$.
+By combining this bound with a bound on the number of transpositions needed to traverse a particular sequence of 2-cells in $\mathcal{A}(M)$,~\cite{lesnick2015interactive} show that the computation of the barcodes templates $\mathcal{T}$ from step (3) using vineyard updates requires on the order of $O(m^3 \kappa + m \kappa^2 \log \kappa)$ elementary operations and $O(m \kappa^2)$ storage.
+
+Of the three steps above, the barcode template computation is both the highest complexity and most demanding computation in practice.
 The number of 2-cells in $\mathcal{A}(M)$ is on the order $O(\kappa^2)$, and $\kappa$ itself is on the order of $O(m^2)$ in the worst case.
@@
 Two binning parameters associated with the coarsening procedure, $x_{bin}$ and $y_{bin}$, together determine the degree of the approximation.
-The coarsening procedure allows the practitioner to restrict the size of $\kappa$ to a constant, which in-turn can dramatically reduce the size of $\mathcal{A}(M)$ when $\kappa << m^2$.
-Keeping in mind this constant factor, the entire computation for approximating the rank invariant given a multifiltration with $m$ simplices is $\approx O(m^3)$.
+The coarsening procedure allows the practitioner to restrict the size of $\kappa$ to a relatively small constant, which in-turn dramatically reduces the size of $\mathcal{A}(M)$.
+With this in mind, given a multifiltration with $m$ simplices, the entire computation for approximating the fibered barcode invariant reduces to an $\approx O(m^3)$ algorithm.
 
 There are several approaches one can use to compute $\mathcal{T}$, the simplest being to run Algorithm~\ref{alg:reduce} independently on the 1-D filtration induced by the duals of some set of points (e.g.
-the barycenters) contained within the 2-cells of $\mathcal{A}(M)$.
-The approach taken by~\cite{lesnick2015interactive} is to use an $R = DV$ decomposition previously computed at some 2-cell $e \in \mathbb{A}(M)$ to speed up the computation of a new decomposition at some adjacent cell $e' \in \mathbb{A}(M)$.
+the barycenters) lying in the interior of the 2-cells of $\mathcal{A}(M)$.
+The approach taken by~\cite{lesnick2015interactive} is to use the $R = DV$ decomposition computed at some adjacent 2-cell $e \in \mathcal{A}(M)$ to speed up the computation of an adjacent cell $e' \in \mathcal{A}(M)$.
 More explicitly, define the \emph{dual graph} of $\mathcal{A}(M)$ to be the undirected graph $G$ which has a vertex for every 2-cell $e \in \mathcal{A}(M)$ and an edge for each pair of cells $e, e' \in \mathcal{A}(M)$ which are adjacent in $\mathcal{A}(M)$.
@@
 On left, the input data, colored by density.
-In the middle, the bigraded Betti numbers $\xi_0(M)$ and $\xi_1(M)$ (green and red, respectively), the Hilbert function (gray), and a line $L$ emphasizing the persistence of features with high density.
-On the right, the line arrangement $\mathcal{A}(M)$ lying in the dual space $D_p(\alpha)$ derived from the $\xi(M)$.
+In the middle, the bigraded Betti numbers $\beta_0(M)$ and $\beta_1(M)$ (green and red, respectively), the Hilbert function (gray), and a line $L$ emphasizing the persistence of features with high density.
+On the right, the line arrangement $\mathcal{A}(M)$ lying in the dual space $D_p(\alpha)$ derived from the $\beta(M)$.
 The barycenters of the 2-cells in $\mathcal{A}(M)$ depicting where the barcodes templates $\mathcal{T}$ are computed are drawn as orange points, and the dual graph $G$ of $\mathcal{A}(M)$ is drawn using dotted blue lines.
@@
  %\hline
-  \multicolumn{2}{|c|}{ $\xi$ and $\mathcal{A}(M)$ } & \multicolumn{3}{c|}{ Column reductions / Permutations } \\
+  \multicolumn{2}{|c|}{ $\beta$ and $\mathcal{A}(M)$ } & \multicolumn{3}{c|}{ Column reductions / Permutations } \\
  \hline

### 2021-08-04 move_schedules.tex
Updated paragraph transitioning to reparameterization

@@
 	$$ f_M(t,\epsilon, \alpha) = \mathrm{rank}(M_t(\epsilon - \alpha) \to M_t(\epsilon + \alpha)) $$
-	Thus, a crocker stack is a sequence of $\alpha$-smoothed crocker plots that vary over $\alpha \geq 0$, satisfying $f_M(t,\epsilon,\alpha) \leq f_M(t,\epsilon, \alpha)$ for all $\alpha \geq \alpha'$. 
 \end{definition}
-\noindent Note that, unlike crocker plots, applying this $\alpha$ smoothing efficiently \emph{requires} the persistence pairing.
+\noindent A crocker stack is a sequence of $\alpha$-smoothed crocker plots that vary over $\alpha \geq 0$, satisfying $f_M(t,\epsilon,\alpha) \leq f_M(t,\epsilon, \alpha)$ for all $\alpha \geq \alpha'$.
+Note that, unlike crocker plots, applying this $\alpha$ smoothing efficiently \emph{requires} the persistence pairing.
 Indeed, it has been shown that crocker stacks and stacked persistence diagrams (i.e.
@@
 $$ \mathcal{B}(M) = \{ \; B_L(M) : L \in \mathbb{R} \times \mathbb{R}^{+} \; \}$$
-	Thus, for a bipersistence modules $M$, the fibered barcode is simply a 2-parameter family of barcodes given by restricting $M$ to a set affine lines in $\mathbb{R}^2$. 
 \end{definition} 
-\noindent Each line $L$ induces a 1-parameter restriction on $M$. The fibered barcode enjoys a number of stability properties readily expressed using interleaving distances~\cite{landi2014rank}. 
+\noindent Thus, given a bipersistence module $M$, the fibered barcode of $M$ is the 2-parameter family of barcodes given by restricting $M$ to the of set affine lines with non-negative slope in $\mathbb{R}^2$.
+
+Although the fiber barcode is a very intuitive invariant, it is not clear how one might go about computing it efficiently.
+One could compute the 1-D persistence associated to some fixed linear combination of two given filter functions; each linear combination parameterizes a line $L \in \mathbb{R} \times \mathbb{R}^{+}$ which $M$ restricts too.
+However, this an $O(m^3)$ time computation, which is prohibitive for exploratory data analysis purposes.
+In what follows, we describe the reparameterization introduced by~\cite{lesnick2015interactive} that fully characterizes $\mathcal{B}(M)$.
+
+%The fibered barcode enjoys a number of stability properties readily expressed using interleaving distances~\cite{landi2014rank}. 
 % More important to this work, the fibered barcode $\mathcal{B}(M)$ and the rank invariant $\mathrm{rank}(M)$ determine each other~\cite{lesnick2015interactive}.
@@
 \end{enumerate}
-Property (1) asserts that for any $L \in \overline{\mathcal{L}}$, the partial order on $\mathbb{R}^2$ restricts to a total order on $L$. Properties (2) and (3) qualify the following definition:
+Property (1) asserts that for any $L \in \overline{\mathcal{L}}$, the standard partial order on $\mathbb{R}^2$ restricts to a total order on $L$.
+Properties (2) and (3) qualify the following definition:
 \begin{definition}[Critical Lines]
@@
 Minor additional conditions are needed for handling completely horizontal and vertical lines. 
-The importance of this theorem for this effort lies in the fact that the fibered barcodes are completely defined from the precomputed barcode templates $\mathcal{T}$: once $\mathcal{T}$ has been computed and augmented onto $\mathcal{A}(M)$, $\mathcal{B}(M)$ is completely characterized, and the barcodes $\mathcal{B}_L(M)$ of any $L$-induced 1-D filtration can be efficiently computed via a point-location query on $\mathcal{A}(M)$ and a $O(\lvert \mathcal{B}_L(M) \rvert)$ application of the push map.
+The importance of this theorem for this effort lies in the fact that the fibered barcodes are completely defined from the precomputed barcode templates $\mathcal{T}$---once every barcode template $\mathcal{T}^e$ has been computed and augmented onto $\mathcal{A}(M)$, $\mathcal{B}(M)$ is completely characterized, and the barcodes $\mathcal{B}_L(M)$ associated to a 1-D filtration induced by \emph{any} choice of $L$ can be efficiently computed via a point-location query on $\mathcal{A}(M)$ and a $O(\lvert \mathcal{B}_L(M) \rvert)$ application of the push map.
 
@@
 The approach taken by~\cite{lesnick2015interactive} is to use the $R = DV$ decomposition computed at some adjacent 2-cell $e \in \mathcal{A}(M)$ to speed up the computation of an adjacent cell $e' \in \mathcal{A}(M)$.
-More explicitly, define the \emph{dual graph} of $\mathcal{A}(M)$ to be the undirected graph $G$ which has a vertex for every 2-cell $e \in \mathcal{A}(M)$ and an edge for each pair of cells $e, e' \in \mathcal{A}(M)$ which are adjacent in $\mathcal{A}(M)$.
+More explicitly, define the \emph{dual graph} of $\mathcal{A}(M)$ to be the undirected graph $G$ which has a vertex for every 2-cell $e \in \mathcal{A}(M)$ and an edge for each adjacent pair of cells $e, e' \in \mathcal{A}(M)$.
 Each vertex in $G$ is associated with a barcode template $\mathcal{T}^e$, and the computation of $\mathcal{T}$ now reduces to computing a path $\Gamma$ on $G$ which visits each vertex at least once.
-To minimize the computation time, assume the $n$ edges of $G$ are endowed with a weights $w_1, w_2, \dots, w_n$ whose values represent some notion of distance between barcode templates.
-The optimal path $\Gamma^\ast$ would be one of minimal length which visits every vertex of $G$ at least once.
-A $\frac{3}{2}$-approximation for computing an approximately minimal path is known, which reduces the problem to the traveling salesman problem on a metric graph~\cite{christofides1976worst}.
+To minimize the computation time, assume the $n$ edges of $G$ are endowed with non-negative weights $W = w_1, w_2, \dots, w_n$ whose values $w_i \in \mathbb{R}_{+}$ represent some notion of distance.
+The optimal path $\Gamma^\ast$ is the one of minimal length with respect to $W$ which visits every vertex of $G$ at least once.
+There is a known $\frac{3}{2}$-approximation that can be computed efficiently which reduces the problem to the traveling salesman problem on a metric graph~\cite{christofides1976worst}.
+In the next paragraph, we include a small example to illustrate many of the concepts and definitions discussed so far below.
 \\
@@
 \textbf{Noisy circle example}
-\begin{figure}[h]
+Consider a small set of points distributed noisily around $S^1$ that contains a few strong outliers, shown on the left side of Figure~\ref{fig:bifiltration_ex2}.
+Filtering this data set with respect to the Rips parameter and the complement of a kernel density estimate yields a bifiltration whose dimension function and bigraded Betti numbers are shown in the middle figure.
+The gray areas indicate the region where the dimension function is constant.
+On the right side, the space dual to the space in the middle is shown.
+The black lines represent the lines of $\mathcal{A}(M)$, the blue dashed-lines the edges of the dual graph $\mathcal{G}$, and the barycenters of the 2-cells in $\mathcal{A}(M)$ depicting where the barcodes templates $\mathcal{T}^e$ are computed are drawn as orange points.
+\begin{figure}[t]
 	\includegraphics[width=0.98\textwidth]{bifiltration_ex2}
@@
 On the right, the line arrangement $\mathcal{A}(M)$ lying in the dual space $D_p(\alpha)$ derived from the $\beta(M)$.
-The barycenters of the 2-cells in $\mathcal{A}(M)$ depicting where the barcodes templates $\mathcal{T}$ are computed are drawn as orange points, and the dual graph $G$ of $\mathcal{A}(M)$ is drawn using dotted blue lines.
-The shortest path $\Gamma^\ast$ spanning the vertices of $G$ is shown by the solid colored lines.
 }
+	\label{fig:bifiltration_ex2}
 \end{figure}
+The shortest path $\Gamma^\ast$ spanning the vertices of $G$ is shown by the solid rainbow colored lines, with the traversal beginning in the top-left cell (at the red edge) and ending at the bottom-right cell (the indigo edge).
+Traversing $\Gamma^\ast$ can be done via a simple depth-first search, though note that a $R=DV$ decomposition needs to be stored at each degree-3 or higher vertex in $\Gamma^\ast$ to avoid redundant computations as the traversal recurses.
 \\

### 2021-08-04 move_schedules.tex
Reworking the last section on image patches results

@@
  In the previous section, we outlined the computational theory of multidimensional persistence, including its relevance to our proposed move scheduling approach.
-In this section, we demonstrate the efficiency of our method on practical use case of multidimensional persistence: detecting the presence of a low-dimensional topological space which well-approximates the distribution of high dimensional data.
+In this section, we demonstrate the efficiency of our method on practical use case of multidimensional persistence: detecting the presence of a low-dimensional topological space which well-approximates the distribution of texture patches extracted from real-world image data.
 
@@
 	\caption{Bipersistence example of natural images data set on an $12 \times 16$ coarsened grid.
-On the left, a projection the full data set onto the first two DCT basis vectors is shown, along with the 15 landmark patches.
+On the left, a projection the full data set is shown, along with the 15 landmark patches.
 In the middle, the bigraded Betti numbers, the Hilbert function, and a chosen line $L$ are shown for the coarsened grid.
@@
 A common hypothesis is that high dimensional data tend to lie in the vicinity of an embedded, low dimensional manifold or topological space.
-An exemplary demonstration of this is given in~\cite{lee2003nonlinear}, who explored the space of $3 \times 3$ high-contrast patches extracted from Van Hatern's still image collection, which consists of approximately 4,000 monochrome images depicting various areas outside around Groningen (Holland)~\cite{}.
-The study of how empirical data from natural images is distributed with respect to predicted spaces and manifolds is motivated by applications in sparse coding and efficient image representations.
-In particular, Lee et al.~\cite{} was interested in analyzing the empirical probability distribution of high-contrast $3 \times 3$ patches of natural images, in pixel-space, towards understanding whether there were any clear qualitative differences the distributions of patches extracted from from images of different modalities, e.g.
+An exemplary demonstration of this is given in~\cite{lee2003nonlinear}, who explored the space of $3 \times 3$ high-contrast patches extracted from Hans van Hateren's still image collection~\cite{hateren_schaaf_1998}, which consists of approximately 4,000 monochrome images\footnote{See \url{http://bethgelab.org/datasets/vanhateren/} for details on the image collection.} depicting various areas outside around Groningen (Holland).
+The study of how empirical data extracted from natural images is distributed with respect to predicted spaces and manifolds is motivated by applications in sparse coding and efficient image representations.
+In particular, Lee et al.~\cite{lee2003nonlinear} was interested in analyzing the empirical probability distribution of high-contrast $3 \times 3$ patches of natural images, in pixel-space, towards understanding whether there were any clear qualitative differences the distributions of patches extracted from from images of different modalities, e.g.
 optical versus range images.
-Experiments conducted by Carllsson et al.~\cite{} using persistent homology confirmed that the distribution of high-contrast $3 \times 3$ patches is indeed well-approximated by a Klein bottle---around 60\% of the high-contrast patches from the still image data set are concentrated within a small neighborhood around a model of the Klein bottle which accounts for $\approx$ 21\% of the volume $S^7$, compared to the $\approx$ 84\% of volume of $S^7$ accounted for by the entire data set.
-In a subsequent effort by Perea et al.~\cite{}, a framework for estimating and representing the distribution around low dimensional submanifolds of pixel space is given, establishing a principled way of obtaining multi-scale rotation invariant descriptors useful for classification tasks.
-
-Using the same high-contrast patch data set studied in~\cite{lee2003nonlinear}, we evaluate the performance of various methods at computing the rank invariant using the parameterization from~\ref{sec:fibered_barcode_reparam}.
-Specifically, we begin by extracting a representative subset of the set $X \subset \mathcal{X}$ of patches representing the top 30\% densest patches of the space of all patches $\mathcal{X}$.
-Carlsson et al.~\cite{} studied a variety of choices of $k$ for a k-nearest neighbor (knn) density estimator $\rho$---which determines which points are the ``densest''---in order to understand how the data were distributed around the Klein bottle model.
-It was determined empirically that when $k \approx 15$, the top 30\% captures three types of directional shifts in the grayscale gradients of the image patches.
+Experiments conducted by Carllsson et al.~\cite{carlsson2008local} using persistent homology confirmed that the distribution of high-contrast $3 \times 3$ patches is indeed well-approximated by a Klein bottle---around 60\% of the high-contrast patches from the still image data set are concentrated within a small neighborhood around a model of a Klein bottle which accounts for $\approx$ 21\% of the volume $S^7$, compared to the $\approx$ 84\% of volume of $S^7$ accounted for by the entire data set.
+Along a similar vein but from a sparse coding perspective, Perea et al.~\cite{perea2014klein} introduced dictionary learning framework for estimating and representing the distribution of patches from texture images.
+
+A priori, had one not been aware of the analysis done by~\cite{lee2003nonlinear, hateren_schaaf_1998, carlsson2008local, perea2014klein}, there is no specific reason to assume the data are well-approximated by the Klein bottle specifically.
+
+Using the same high-contrast patch data set studied in~\cite{lee2003nonlinear}, we evaluate the performance of various methods at computing the fibered barcode invariant using the parameterization from~\ref{sec:fibered_barcode_reparam}.
+First, we give a summary of the data set.
+Specifically, we begin by extracting a representative subset of the set $X \subset \mathcal{X}$ of patches representing the top 30\% densest patches of the space of all patches $\mathcal{X}$, where the notion of density is approximated with a k-nearest neighbor (knn) density estimator $\rho$.
+Carlsson et al.~\cite{carlsson2008local} studied a variety of choices of $k$ in order to understand how localized the data were distributed around the Klein bottle model.
+He determined empirically that when $k \approx 15$, the top 30\% captures three types of directional shifts in the grayscale gradients of the image patches.
 These three gradients were shown to be captured well by the three-circle model $C_3$---the 1-skeleton of the Klein bottle.
-In this space, there are three circles, two of which (say, $S_v$ and $S_h$) intersect the third (say, $S_\mathrm{lin}$) in exactly two points, but they themselves do not intersect.
+The $C_3$ model consists of three circles, two of which (say, $S_v$ and $S_h$) intersect the third (say, $S_\mathrm{lin}$) in exactly two points, but themselves do not intersect.
 One of the circles, $S_{\mathrm{lin}}$, parameterizes a the rotational gradient, whereas the other two $S_v$ and $S_h$ parameterize vertical and horizontal gradients.
-A projection of the data $X$ onto the first two DCT basis, is shown in the left-most image in Figure~\ref{fig:patch_data_dgm}, along with 15 landmark points.
-It was shown in~\cite{} that when $k \approx 300$, the top 30\% of data extracted from the knn density estimator $\rho_{300}$ `fills out' the two-manifold, yielding Betti numbers $\beta_0 = \beta_1 = 1$, whereas when $k$ is decreases to a more localized estimate, the three-circle model is recovered, which has Betti numbers $\beta_0 = 1, \beta_1 = 5$.
-
-We sampled landmarks from $X \subset \mathcal{X}$ using a stratified sampling approach with $n = 25$ strata, where each strata contains the $(1/n)$-thick level set given by the knn density estimate $\rho_{15}$.
-The use of landmarks gives us certain coverage guarantees that the geometry is preserved within each level set, whereas the stratification ensures the original density of is approximated preserved as well.
-From this data set, we construct a Rips-density bifiltration $\mathcal{F}$ using $\rho_{15}$ equipped with the geodesic metric computed over the $k = 15$ knn graph of $X$.
-In theory, this bifiltration should be well approximated by $C_3$, which we ought to be able to determine by a careful inspection of the rank invariant.
+A projection of the data $X$ is shown in the left-most image in Figure~\ref{fig:patch_data_dgm}, along with 15 landmark points.
+It was shown in~\cite{carlsson2008local} that when $k \approx 300$, the top 30\% of data extracted from the knn density estimator $\rho_{300}$ `fills out' the two-manifold, yielding Betti numbers $\beta_0 = \beta_1 = 1$, whereas when $k$ is decreases to a more localized estimate, the three-circle model is recovered, which has Betti numbers $\beta_0 = 1, \beta_1 = 5$.
+
+Due to the aformentioned high complexity of the fibered barcode computation, we begin by working with a subset of the image patch data $\mathcal{X}$.
+Specifically, we combine the use furthest-point sampling and proportionate allocation (stratified) sampling to sample landmarks $X \subset \mathcal{X}$ distributed within $n = 25$ strata.
+Each strata consists of the $(1/n)$-thick level set given by $\rho_{15}$, the density estimator.
+The use of furthest-point sampling gives us certain coverage guarantees that the geometry is approximately preserved within each level set, whereas the stratification ensures the original density of is approximated preserved as well.
+From this data set, we construct a Rips-(co)density bifiltration $\mathcal{F}$ using $\rho_{15}$ equipped with the geodesic metric computed over the $k = 15$ knn graph on $X$.
+
+In theory, this bifiltration should be well approximated by the $C_3$ model, which one ought to be able to determine by a careful inspection of the fibered barcode.
 That is, since $C_3$ has $\beta_1  = 5$, there should be some line $L$ which induces a 1-parameter filtration whose corresponding persistence diagram (for dimension $1$) displays five highly persistent features.
@@
 \begin{table}[h]
-\caption{Cost to compute all barcode templates $\mathcal{T}$ }\label{tab:barcode_templates}
+\caption{Cost to compute all barcode templates $\mathcal{T}$ for various coarsening choices $\beta(M)$.}\label{tab:barcode_templates}
 \centering
@@
  %\hline
-  \multicolumn{2}{|c|}{ $\beta$ and $\mathcal{A}(M)$ } & \multicolumn{3}{c|}{ Column reductions / Permutations } \\
+  \multicolumn{1}{|c|}{ $\beta(M)$ } & \multicolumn{1}{|c|}{$\mathcal{A}(M)$ } & \multicolumn{3}{c|}{ Column reductions / Permutations } \\
  \hline

### 2021-08-04 move_schedules.tex
Wrapping up changes to the last paragraphs

@@
 
-A priori, had one not been aware of the analysis done by~\cite{lee2003nonlinear, hateren_schaaf_1998, carlsson2008local, perea2014klein}, there is no specific reason to assume the data are well-approximated by the Klein bottle specifically.
+If one was not aware of the analysis done by~\cite{lee2003nonlinear, hateren_schaaf_1998, carlsson2008local, perea2014klein}, there is no specific reason apriori to assume the data are well-approximated by the Klein bottle.
+Consider the (coarsened) fibered barcode shown in Figure~\ref{fig:patch_data_dgm} which was computed from a standard Rips/codensity bifiltration on a constructed on a representative sample of the image data from~\cite{hateren_schaaf_1998} (how we obtained this sample is described further down).
+Upon inspection of the bigraded Betti number and the dimension function, one finds that a large area of dimension function is constant (highlighted as the blue portion in the middle of Figure~\ref{fig:patch_data_dgm}), wherein $\mathrm{dim}(M) = 5$.
+This indicates
 

### 2021-08-05 move_schedules.tex
Beginning the bridge to describe the detection example better

@@
 The gray areas indicate the region where the dimension function is constant.
+In particular, the light gray area of the middle plot shows where $\mathrm{dim}_1(M) = 1$, indicating where in the multifiltration a persistent 1-cycle was detected.
 On the right side, the space dual to the space in the middle is shown.
@@
 The blue regions highlights where $\mathrm{dim}(M) = 5$.
-On the right, the persistence diagram of $L$, with larger points indicating a multiplicity of 2.
+On the right, the persistence diagram of $M$ restricted to $L$, with larger points indicating a multiplicity of 2.
 Observe five persistent features are revealed in both the middle and right plots, matching $\beta_1$ of the three-circle model.
@@
 An exemplary demonstration of this is given in~\cite{lee2003nonlinear}, who explored the space of $3 \times 3$ high-contrast patches extracted from Hans van Hateren's still image collection~\cite{hateren_schaaf_1998}, which consists of approximately 4,000 monochrome images\footnote{See \url{http://bethgelab.org/datasets/vanhateren/} for details on the image collection.} depicting various areas outside around Groningen (Holland).
-The study of how empirical data extracted from natural images is distributed with respect to predicted spaces and manifolds is motivated by applications in sparse coding and efficient image representations.
-In particular, Lee et al.~\cite{lee2003nonlinear} was interested in analyzing the empirical probability distribution of high-contrast $3 \times 3$ patches of natural images, in pixel-space, towards understanding whether there were any clear qualitative differences the distributions of patches extracted from from images of different modalities, e.g.
+Motivated by applications to sparse coding applications, Lee et al.~\cite{lee2003nonlinear} was interested in how these natural image patches were distributed, in pixel-space, with respect to predicted spaces and manifolds.
+For example, one of the motivating questions of their research was whether there were any clear qualitative differences the distributions of patches extracted from from images of different modalities, e.g.
 optical versus range images.
-Experiments conducted by Carllsson et al.~\cite{carlsson2008local} using persistent homology confirmed that the distribution of high-contrast $3 \times 3$ patches is indeed well-approximated by a Klein bottle---around 60\% of the high-contrast patches from the still image data set are concentrated within a small neighborhood around a model of a Klein bottle which accounts for $\approx$ 21\% of the volume $S^7$, compared to the $\approx$ 84\% of volume of $S^7$ accounted for by the entire data set.
-Along a similar vein but from a sparse coding perspective, Perea et al.~\cite{perea2014klein} introduced dictionary learning framework for estimating and representing the distribution of patches from texture images.
-
-If one was not aware of the analysis done by~\cite{lee2003nonlinear, hateren_schaaf_1998, carlsson2008local, perea2014klein}, there is no specific reason apriori to assume the data are well-approximated by the Klein bottle.
+A major result from their work is that the majority of high-contrast image patches are concentrated around a 2-dimensional submanifold embedded in $S^7$.
+Supplemental experiments conducted by Carllsson et al.~\cite{carlsson2008local} using persistent homology confirmed that the distribution of high-contrast $3 \times 3$ patches is indeed well-approximated by a Klein bottle---around 60\% of the high-contrast patches from the still image data set are concentrated within a small neighborhood around a model of a Klein bottle which accounts for $\approx$ 21\% of the volume $S^7$, compared to the $\approx$ 84\% of volume of $S^7$ accounted for by the entire data set.
+Along a similar vein, in the context of sparse coding, Perea et al.~\cite{perea2014klein} introduced a dictionary learning framework for estimating and representing the distribution of patches from texture images.
+
+If one was not aware of the analysis done by~\cite{lee2003nonlinear, hateren_schaaf_1998, carlsson2008local, perea2014klein}, it is not immediately clear a priori that the Klein bottle model is indeed an excellent candidate for capturing the non-linearity of the data.
+Nonetheless, as we have mentioned, applications of persistent homology detecting the presence of a topological space persistent homology is designed to reveal.
 Consider the (coarsened) fibered barcode shown in Figure~\ref{fig:patch_data_dgm} which was computed from a standard Rips/codensity bifiltration on a constructed on a representative sample of the image data from~\cite{hateren_schaaf_1998} (how we obtained this sample is described further down).

### 2021-08-05 move_schedules.tex
Reworked overview!

@@
 Indeed, given a triangulable topological space equipped with a sufficiently tame continuous function, persistent homology captures the changes in topology across the sublevel sets of the space, and encodes them in a persistence diagram.
-The stability of persistence contends
-that small changes in the underlying function produce small changes in the corresponding diagrams:  if the function changes continuously, so too will the points on the persistence diagram. 
+The stability of persistence contends that small changes in the underlying function produce small changes in the corresponding diagrams:  if the function changes continuously, so too will the points on the persistence diagram.
 This motivates the study and application of persistence to time-varying or dynamic settings, like that of dynamic metric spaces~\cite{kim2020spatiotemporal}. 
-Focusing on the computational aspects of this problem has far-reaching consequences: methods that vectorize  persistence diagrams for machine learning purposes, like adaptive template functions~\cite{polanco2019adaptive} and persistence images~\cite{adams2017persistence}, or summary statistics like $\alpha$-smoothed Betti curves~\cite{ulmer2019topological}, all immediately become  computationally viable tools for time-varying settings.
+As persistence-related computations tend to exhibit high algorithmic complexity (e.g.
+essentially cubic\footnote{For finite fields, it is known that the persistence computation reduces to the PLU factorization problem, which takes $O(m^\omega)$ where $\omega \approx 2.373$ is the matrix multiplication constant.} in the size of the underlying filtration~\cite{morozov2005persistence}), their adoption to dynamic settings poses a challenging computational problem.
+With the current state of the art tools, there is no recourse to the problem of computing persistence from a time-varying filtration containing millions of simplices at thousands of snapshots in time.
+An efficient means of updating the persistence of a [time-varying] filtration has far-reaching consequences: methods that vectorize persistence diagrams for machine learning purposes, like adaptive template functions~\cite{polanco2019adaptive} and persistence images~\cite{adams2017persistence}, or summary statistics like $\alpha$-smoothed Betti curves~\cite{ulmer2019topological}, all immediately become computationally viable tools in dynamic  settings.
 %This analogy provides a strong interpretation at gaining insight on the subtle topological changes that may occur in studying a continuous process.
 
-Cohen-Steiner et al. refer to a continuous 1-parameter family of persistence diagrams as a \emph{vineyard}, and they give in \cite{cohen2006vines} an efficient algorithm for their computation
-given a time-varying filtration. 
-Computing vineyards can be interpreted as an extension of the \emph{matrix reduction} algorithm~\cite{zomorodian2005computing}, which computes the persistence diagram of a fixed filtration $K$ with $m$ simplices in $O(m^3)$ time.
+Cohen-Steiner et al.
+refer to a continuous 1-parameter family of persistence diagrams as a \emph{vineyard}, and they give in \cite{cohen2006vines} an efficient algorithm for their computation given a time-varying filtration.
+The vineyards approach can be interpreted as an extension of the \emph{matrix reduction} algorithm~\cite{zomorodian2005computing}, which computes the persistence diagram of a fixed filtration $K$ with $m$ simplices in $O(m^3)$ time.
 The main step in the reduction algorithm is a specific decomposition $R = D V$ of the boundary matrix $D$ of $K$. 
@@
 If one is interested in understanding how the persistent homology of a continuous function changes over time, then this algorithm is sufficient, for  homological critical points can only occur when the filtration order changes.
-If there are $d$ time-points where this order changes, then the vineyards algorithm takes $O(m^3 + md)$ time: one $O(m^3)$-time reduction upfront for the filtration at time $t_0$ followed by one $O(m)$ operation to update the decomposition for the remaining time points $(t_1, t_2, \dots, t_d)$.
-When $d >> m$, this $O(md)$ approach is far more efficient than the $O(dm^3)$  naive strategy of computing the diagrams independently at every time point.
-
-Despite the vineyards algorithm being efficient in theory, it is often not the method of choice in practice in dynamic settings. 
+Indeed, this algorithm is also efficient  asymptotically: if there are $d$ time-points where this filtration order changes, then the vineyards algorithm takes $O(m^3 + md)$ time: one $O(m^3)$-time reduction upfront for the filtration at time $t_0$ followed by one $O(m)$ operation to update the decomposition for the remaining time points $(t_1, t_2, \dots, t_d)$.
+When $d >> m$, this $O(md)$ approach is far more efficient than the $O(dm^3)$  strategy of computing the diagrams independently at every time point, what we call the ``naive'' approach.
+
+Despite its efficient in theory, vineyards is often not the method of choice in practice in dynamic settings. 
 While there is an increasingly rich ecosystem of software packages offering variations of the standard reduction algorithm (e.g.
@@
 many transpositions can be quite slow...
-it is sometimes much faster to simply recompute the $RU$-decomposition from scratch using the standard persistence algorithm.'' Indeed, whether performing the reductions independently or updating the decomposition dynamically using vineyards, they observed that obtaining the decomposition is the most computationally demanding aspect of their 2-D persistence algorithm.
+it is sometimes much faster to simply recompute the $RU$-decomposition from scratch using the standard persistence algorithm.'' Indeed, whether performing the reductions naively or updating the decomposition dynamically using vineyards, they observed that obtaining the decomposition is the most computationally demanding aspect of their 2-D persistence algorithm.
 
@@
 % in many time varying settings, a valid decomposition is often only sought after at a relatively sparse number of time points relative to the size of the (finite) set needed to generated by the family.
-By making a tradeoff in the number of times we restore the decomposition to a valid state, we are able to reduce the total number of column operations needed to apply an arbitrary permutation to the filtration.
+By exploiting the ``donor'' concept introduced in in~\cite{busaryev2010tracking}, we are able to make a tradeoff in the number of times the decomposition is restored to a valid state, dramatically reducing the total number of column operations needed to apply an arbitrary permutation to the filtration.
 This tradeoff, paired with a fast greedy heuristic explained in section~\ref{sec:proxy_objective}, yields an algorithm that can update a $R = DV$ decomposition more efficiently than vineyards in sparse contexts, making dynamic persistence more computationally tractable for a wider class of use-cases.
@@
  \subsection{A Motivating Example and Contributions}\label{sec:motivation} 
-As a motivating example, we describe a simple experiment illustrating why the vineyards algorithm does not always yield an efficient method in time-varying applications. 
+As a motivating example, we describe a simple experiment illustrating why the vineyards algorithm does not always yield an efficient method in many time-varying applications. 
 Consider a series of grayscale images (i.e.
@@
  \end{figure*}
-This time-varying filtration is depicted in Figure~\ref{fig:vineyards} (bottom right), which we use as a baseline for comparing the standard algorithm $\texttt{pHcol}$ with vineyards.
+Snapshots of the image data and its corresponding time-varying filtration is depicted in Figure~\ref{fig:vineyards} (top right and bottom right, respectively), which we use as a baseline for comparing the standard algorithm $\texttt{pHcol}$ with vineyards.
 Note the size of the filtration is fixed--no simplices are ever added or removed over time.
@@
   On the left side of Figure~\ref{fig:vineyards}, we compare the cumulative cost (in total number of column operations) of various approaches which one can use to simulate persistence dynamically. 
-  Since it is unknown a-priori at what points in  time the persistent homology will change in application settings, one solution is to discretize the time domain at $n$ evenly spaced points and compute the persistence diagram at each of these time points independently.
+  Since it is unknown a priori at what points in time the persistent homology will change in application settings, one solution is to discretize the time domain at $n$ evenly spaced points and compute the persistence diagram at each of these time points independently.
 An alternative approach is to construct a homotopy between two time-points, and then decompose this homotopy into adjacent transpositions--the vineyards approach.
@@
 As one can see from the figure, if one needs to simulate the homotopy exactly, the vineyards approach is indeed far more efficient than naively applying the reduction algorithm independently at all time points.
-  However,  when the discretization of the time domain is sparse enough, the naive approach   performs less column operations than the vineyards strategy, while still capturing the main events.
+  However, when the discretization of the time domain is sparse enough, the naive approach actually performs less column operations than the vineyards strategy, while still capturing the main events.
 
@@
 That is, even in situations where one is updating a decomposition to a filtration that is  relatively ``close'', e.g.
-any two adjacent filtrations in the bottom left of Figure~\ref{fig:vineyards}, it is still more efficient to apply \texttt{pHcol} directly than  to iteratively update the decomposition (see the case where $n = 10$).
+any two adjacent filtrations in the bottom right of Figure~\ref{fig:vineyards}, it is still more efficient to apply \texttt{pHcol} directly than  to iteratively update the decomposition (see the case where $n = 10$).
  Indeed, any optimizations to the reduction algorithm (for example, the clearing optimization~\cite{chen2011persistent}) would only increase this disparity, making the vineyards algorithm less computationally attractive even in settings where $n$ is larger.
@@
 In light of these results, we give an alternative proxy-objective to minimize, provide bounds justifying its relevance to the original objective, and give an efficient $O(d \log d)$ algorithm for approximately solving this proxy minimization.
-A performance comparison with other reduction-based persistent homology computations is given, wherein our approach is demonstrated to be an order of magnitude more efficient than existing approaches in sparse dynamic settings.
+A performance comparison with other reduction-based persistent homology computations is given, wherein our approach is demonstrated to be an order of magnitude more efficient than existing approaches in a few common sparse dynamic settings, such as flock simulation and 2D persistence computation.
 We conclude with a discussion of situations where our strategy is most applicable and future directions for research.
@@
 Another popular optimization is to utilize the duality between   homology and cohomology~\cite{de2011dualities}, which dramatically improves the effectiveness of the clearing optimization~\cite{bauer2021ripser}.
-There are many other optimizations on the implementation side: the use of ranking functions defined on the combinatorial number system enables implicit cofacet enumeration, removing the need to store the boundary matrix explciitly; the apparent/emergent pairs optimization identifies columns whose pivot entries are unaffected by the reduction algorithm, reducing the total number of columns which need be reduced; sparse data structures such as bit-trees and lazy heaps allow for efficient column-wise additions with $\mathbb{Z}_2 = \mathbb{Z}/2\mathbb{Z}$ coefficients and effective $O(1)$ pivot entry retrieval, and so on~\cite{bauer2017phat, bauer2021ripser}.
+There are many other optimizations on the implementation side: the use of ranking functions defined on the combinatorial number system enables implicit cofacet enumeration, removing the need to store the boundary matrix explciitly; the apparent/emergent pairs optimization identifies columns whose pivot entries are unaffected by the reduction algorithm, reducing the total number of columns which need be reduced; sparse data structures such as bit-trees and lazy heaps allow for efficient column-wise additions with $\mathbb{Z}_2 = \mathbb{Z}/2\mathbb{Z}$ coefficients and effective $O(1)$ pivot entry retrieval, and so on~\cite{bauer2021ripser, bauer2017phat}.
 
@@
 In section 3 we introduce our proposed alternative to vineyards and provide efficient algorithms to make this alternative viable. 
-In section 4 we demonstrate a few applications of the proposed method. In section 5 we conclude the paper by discussing other possible applications, future work. 
+In section 4 we demonstrate a few applications of the proposed method. In section 5 we conclude the paper by discussing other possible applications and future work. 
 We also include an appendix which introduces certain topics in more detail, such as the vineyards algorithm itself, for readers who may be unfamiliar but would like a more in-depth understanding.
@@
 It is worth noting that this bound applies to pairs of uniformly sampled permutations, as opposed to uniformly sampled distinct filtrations---intuitively, simplicial filtrations have more structure due to the subcomplex relation that must be respected in the filtration.
-Thus, while $O(m - \sqrt{m})$ in expectation is an improvement over $O(m)$ it is still not tight.
+Thus, while $O(m - \sqrt{m})$ in expectation is an improvement over $O(m)$, it is still not tight.
 However, Boissonnat~\cite{boissonnat2018efficient} has shown that for any $d$-dimensional simplicial complex $K$ with $m$ simplices, the number of distinct filtrations built from $K$ is \emph{at least} $\floor{\frac{t+1}{d+1}}^m$, where $t$ is the number of distinct filtration values associated with the complex $K$.
@@
 In practice, when one has a time-varying filtration and the sampling points are relatively close [in time], the LCS between adjacent filtrations is expected to be much larger, subsequently making $d$ much smaller.
-This had been confirmed empirically, e.g.
-the average size of the LCS across the 10 filtrations from Section~\ref{sec:motivation} each having $417$ simplices sampled at evenly spaced time points was $343$, implying $d \approx 70$ permutations needed on average to update the decomposition between adjacent time points.
+For example, the average size of the LCS across the 10 filtrations from Section~\ref{sec:motivation} each having $m=417$ simplices sampled at evenly spaced time points was $343$, implying $d \approx 70$ permutations needed on average to update the decomposition between adjacent time points.
 

### 2021-08-05 move_schedules.tex
various edits throughout the main body of the paper

@@
 Due to the combinatorial aspect of our approach, there is actually no need to have access to---or explicitly store---the entire family of filtrations and the schedules between them.
-The entire approach can be performed online, keeping at most two filtrations and one decomposition in memory at any given time.
+That is, Algorithm~\ref{alg:schedule} can be easily modified to be completely online, keeping at most two filtrations and one decomposition in memory at any given time.
 
@@
 In the case of vineyards, the bound is clear: whether assuming a homotopy between a pair of filtrations $(K_0, K_1)$ or not, if $\kappa = K_\tau(K_0, K_1)$, then $\Omega(\kappa)$ adjacent transpositions are required to transform $K_0 \mapsto K_1$.
-Allowing move operations in any order, we have $O(m)$ as a trivial upper bound: simply move each simplex in $K_0$ into its position in the filtration given by $K_1$ in the order given by the latter.
+In contrast, if we allow move operations in any order, we have $O(m)$ as a trivial upper bound: simply move each simplex in $K_0$ into its position in the filtration given by $K_1$ in the order given by the latter.
 However, it's not immediately clear whether this bound is tight.
@@
 	The history of the mathematics of estimating this quantity dates back at least 50 years, and is sometimes called the \emph{Ulam-Hammersley} problem.
-It was shown by Baik et al.~\cite{} that as $m \to \infty$:
+The seminal work by Baik et al.~\cite{baik1999distribution} established that as $m \to \infty$:
 	$$ \displaystyle \ell_m = 2 \sqrt{m} + c m^{1/6} + o(m^{1/6}) $$
-where $c \approx -1.77108$. As a result, we have: 
+where $c = -1.77108...$. As a result, we have: 
 $$ \frac{\ell_m}{\sqrt{m}} \to 2\rlap{\quad \text{as } $m \to \infty$} $$	 
@@
 However, Boissonnat~\cite{boissonnat2018efficient} has shown that for any $d$-dimensional simplicial complex $K$ with $m$ simplices, the number of distinct filtrations built from $K$ is \emph{at least} $\floor{\frac{t+1}{d+1}}^m$, where $t$ is the number of distinct filtration values associated with the complex $K$.
-Since this lower bound grows similarly to $m!$ when $t \approx O(m)$ and $d << m$ fixed, we expect $O(m - \sqrt{m})$ is not too pessimistic of an upper bound when the pair of filtrations are truly random.
+Since this lower bound grows similarly to $m!$ when $t \sim O(m)$ and $d << m$ fixed, we expect $O(m - \sqrt{m})$ is not too pessimistic of an upper bound when the pair of filtrations are truly random.
 In practice, when one has a time-varying filtration and the sampling points are relatively close [in time], the LCS between adjacent filtrations is expected to be much larger, subsequently making $d$ much smaller.

### 2021-08-05 move_schedules.tex
Minor changes to fibered barcode section

@@
 The multigraded Betti numbers constitute a natural and important class of invariants often studied in algebraic geometry commutative algebra, though both their theory and computation is beyond the scope of this effort---the interested reader is referred to~\cite{lesnick2015interactive, carlsson2009theory} for more details.
-For our purposes, the the $i^{\text{th}}$-graded Betti number of $M$ is simply a function $\beta_i(M): \mathbb{R}^n \to \mathbb{N}$.
+For our purposes, the $i^{\text{th}}$-graded Betti number of $M$ is simply a function $\beta_i(M): \mathbb{R}^n \to \mathbb{N}$ whose values $\beta_i^z(M)$ indicate the the number of elements at grade $z$ in a basis of the $i^{\text{th}}$ module in a free resolution for $M$.
 The fibered barcode, denoted as $\mathcal{B}(M)$, is defined as follows:
@@
 However, this an $O(m^3)$ time computation, which is prohibitive for exploratory data analysis purposes.
-In what follows, we describe the reparameterization introduced by~\cite{lesnick2015interactive} that fully characterizes $\mathcal{B}(M)$.
+Ideally, one would want to continuously vary $L \in \mathbb{R} \times \mathbb{R}^{+}$ interactively, inspecting the corresponding changes to the barcodes immediately.
+To enable this, ~\cite{lesnick2015interactive} introduced an elegant reparameterization that not only fully characterizes $\mathcal{B}(M)$, but also yields an efficient method for rendering $\mathcal{B}_L(M)$ in real time.
+In the next section, we briefly summarize this reparameterization.
 
@@
 Minor additional conditions are needed for handling completely horizontal and vertical lines. 
-The importance of this theorem for this effort lies in the fact that the fibered barcodes are completely defined from the precomputed barcode templates $\mathcal{T}$---once every barcode template $\mathcal{T}^e$ has been computed and augmented onto $\mathcal{A}(M)$, $\mathcal{B}(M)$ is completely characterized, and the barcodes $\mathcal{B}_L(M)$ associated to a 1-D filtration induced by \emph{any} choice of $L$ can be efficiently computed via a point-location query on $\mathcal{A}(M)$ and a $O(\lvert \mathcal{B}_L(M) \rvert)$ application of the push map.
+The importance of this theorem lies in the fact that the fibered barcodes are completely defined from the precomputed barcode templates $\mathcal{T}$---once every barcode template $\mathcal{T}^e$ has been computed and augmented onto $\mathcal{A}(M)$, $\mathcal{B}(M)$ is completely characterized, and the barcodes $\mathcal{B}_L(M)$ associated to a 1-D filtration induced by \emph{any} choice of $L$ can be efficiently computed via a point-location query on $\mathcal{A}(M)$ and a $O(\lvert \mathcal{B}_L(M) \rvert)$ application of the push map.
 
@@
 In terms of memory complexity, the number of 2-cells in $\mathcal{A}(M)$ is upper bounded by $O(\kappa^2)$, where $\kappa$ is a coarseness parameter associated with the computation of $\beta(M)$.
-By combining this bound with a bound on the number of transpositions needed to traverse a particular sequence of 2-cells in $\mathcal{A}(M)$,~\cite{lesnick2015interactive} show that the computation of the barcodes templates $\mathcal{T}$ from step (3) using vineyard updates requires on the order of $O(m^3 \kappa + m \kappa^2 \log \kappa)$ elementary operations and $O(m \kappa^2)$ storage.
+By combining this bound with a bound on the number of transpositions needed to traverse a particular sequence of 2-cells in $\mathcal{A}(M)$, the analysis from~\cite{lesnick2015interactive} shows that the computation of the barcodes templates $\mathcal{T}$ from step (3) using vineyard updates requires on the order of $O(m^3 \kappa + m \kappa^2 \log \kappa)$ elementary operations and $O(m \kappa^2)$ storage.
 
@@
 The number of 2-cells in $\mathcal{A}(M)$ is on the order $O(\kappa^2)$, and $\kappa$ itself is on the order of $O(m^2)$ in the worst case.
-This suggests the computation of the barcode templates is on the order $O(m^5)$, which is effectively intractable for moderately-sized filtrations.
+This suggests the computation of the barcode templates is on the order $O(m^5)$, which is effectively intractable for even moderately-sized filtrations.
 In practice, the external stability result from~\cite{landi2014rank} justifies the use of a simple procedure which approximates the module $M$ with a smaller module $M'$ using a grid-like coarsening procedure.
@@
 The coarsening procedure allows the practitioner to restrict the size of $\kappa$ to a relatively small constant, which in-turn dramatically reduces the size of $\mathcal{A}(M)$.
-With this in mind, given a multifiltration with $m$ simplices, the entire computation for approximating the fibered barcode invariant reduces to an $\approx O(m^3)$ algorithm.
 
@@
 There is a known $\frac{3}{2}$-approximation that can be computed efficiently which reduces the problem to the traveling salesman problem on a metric graph~\cite{christofides1976worst}.
-In the next paragraph, we include a small example to illustrate many of the concepts and definitions discussed so far below.
-\\
-\\
-\textbf{Noisy circle example}
+\\
+\\
+\textbf{Noisy circle example} We include a small example to illustrate many of the concepts and definitions discussed so far. 
 Consider a small set of points distributed noisily around $S^1$ that contains a few strong outliers, shown on the left side of Figure~\ref{fig:bifiltration_ex2}.
 Filtering this data set with respect to the Rips parameter and the complement of a kernel density estimate yields a bifiltration whose dimension function and bigraded Betti numbers are shown in the middle figure.
-The gray areas indicate the region where the dimension function is constant.
-In particular, the light gray area of the middle plot shows where $\mathrm{dim}_1(M) = 1$, indicating where in the multifiltration a persistent 1-cycle was detected.
+The gray areas indicate the region where the dimension function is constant; the lighter gray area where $\mathrm{dim}_1(M) = 1$, indicating where a persistent loop was detected.
 On the right side, the space dual to the space in the middle is shown.
-The black lines represent the lines of $\mathcal{A}(M)$, the blue dashed-lines the edges of the dual graph $\mathcal{G}$, and the barycenters of the 2-cells in $\mathcal{A}(M)$ depicting where the barcodes templates $\mathcal{T}^e$ are computed are drawn as orange points.
+The black lines represent the lines of $\mathcal{A}(M)$, the blue dashed-lines the edges of the dual graph $\mathcal{G}$, and the orange points the barycenters of the 2-cells in $\mathcal{A}(M)$, depicting where the barcodes templates $\mathcal{T}^e$ are parameterized.
 \begin{figure}[t]

### 2021-08-05 move_schedules.tex
Finished the natural images section! Looking much better

@@
 
-If one was not aware of the analysis done by~\cite{lee2003nonlinear, hateren_schaaf_1998, carlsson2008local, perea2014klein}, it is not immediately clear a priori that the Klein bottle model is indeed an excellent candidate for capturing the non-linearity of the data.
-Nonetheless, as we have mentioned, applications of persistent homology detecting the presence of a topological space persistent homology is designed to reveal.
-Consider the (coarsened) fibered barcode shown in Figure~\ref{fig:patch_data_dgm} which was computed from a standard Rips/codensity bifiltration on a constructed on a representative sample of the image data from~\cite{hateren_schaaf_1998} (how we obtained this sample is described further down).
-Upon inspection of the bigraded Betti number and the dimension function, one finds that a large area of dimension function is constant (highlighted as the blue portion in the middle of Figure~\ref{fig:patch_data_dgm}), wherein $\mathrm{dim}(M) = 5$.
-This indicates
+If one was not aware of the analysis done by~\cite{lee2003nonlinear, hateren_schaaf_1998, carlsson2008local, perea2014klein}, it is not immediately clear a priori that the Klein bottle model is indeed a good candidate for capturing non-linearity of image patches.
+Nonetheless, as we have mentioned, among the first uses of persistent homology was the so-called ``homology inference problem''~\cite{perea2018brief}; given a suitable treatment of strong outliers using by e.g.
+filtering both on the geometry and the density of the image patches, the homology of the Klein bottle should become apparent through a persistence computation.
+
+Consider the (coarsened) fibered barcode computed from a standard Rips/codensity bifiltration on a representative sample of the image data from~\cite{hateren_schaaf_1998} (details to follow), shown in Figure~\ref{fig:patch_data_dgm}.
+Upon inspection of the bigraded Betti number and the dimension function, one finds that a large area of dimension function is constant (highlighted as the blue portion in the middle of Figure~\ref{fig:patch_data_dgm}), wherein the first Betti number is 5.
+There are many spaces whose  meets this criterion, though upon further inspection, one candidate space that seems particularly plausible is the three-circle model $C_3$.
+The $C_3$ model consists of three circles, two of which (say, $S_v$ and $S_h$) intersect the third (say, $S_\mathrm{lin}$) in exactly two points, but themselves do not intersect.
+Upon performing an analysis on the density of the patches analogous to the one done by~\cite{lee2003nonlinear}, one finds that the mean-centered and contrast normalized $3 \times 3$ patches like on a 7-dimensional ellipsoid $\tilde{S}^7 \subset \mathbb{R}^9$.
+Contrast is typically measured using a discrete version of the Dirichlet semi-norm, or the ``$D$-norm'' $\lVert \cdot \rVert_D$ for short, which is a unique scale-invariant norm on images:
+$$ \lVert x \rVert_D = \sqrt{\sum_{i \sim j}(x_i - x_j)^2} = \sqrt{x^T D x}$$
+where $D$ is a fixed matrix which upon application to an image $x \in \mathbb{R}^9$ yields a value proportional to the sum of the differences between the 4-connected neighbors (described by the relation $i \sim j$).
+It is advantageous to ``whiten'' the data via a change of coordinates to Euclidean sphere.
+This can be achieved by expressing the patch data as an expansion of a certain convenient basis of 8 non-constant eigenvectors, often called the Discrete Cosine Transform (DCT) basis~\cite{lee2003nonlinear}, which are given below:
+\vspace{-0.2em}
+\begin{figure}[h]
+	\centering
+	\includegraphics[width=0.60\textwidth]{dct_basis}
+\end{figure}
+\vspace{-0.5em}
+
+\noindent Projecting the image data onto the first two basis vectors of leads to the projection shown in the top left of Figure~\ref{fig:patch_data_dgm}.
+15 landmark points are also shown.
+Observe the data are distributed well around three ``circles''---the outside circle capturing the rotation gradient of the image patches ($S_{\mathrm{lin}}$), and the other two circles capturing the vertical and horizontal gradients ($S_v$ and $S_h$, respectively).
+Since the three circle model is the 1-skeleton of the Klein bottle~\cite{carlsson2008local}, one may conclude that the Klein bottle might be a reasonable candidate upon which the image data are distributed.
+This can be further verified empirically by e.g.
+using the iterative, meanshift-like procedure described in~\cite{carlsson2008local} to fit a parameterization of the model in question to the data.
 
 Using the same high-contrast patch data set studied in~\cite{lee2003nonlinear}, we evaluate the performance of various methods at computing the fibered barcode invariant using the parameterization from~\ref{sec:fibered_barcode_reparam}.
-First, we give a summary of the data set.
-Specifically, we begin by extracting a representative subset of the set $X \subset \mathcal{X}$ of patches representing the top 30\% densest patches of the space of all patches $\mathcal{X}$, where the notion of density is approximated with a k-nearest neighbor (knn) density estimator $\rho$.
-Carlsson et al.~\cite{carlsson2008local} studied a variety of choices of $k$ in order to understand how localized the data were distributed around the Klein bottle model.
-He determined empirically that when $k \approx 15$, the top 30\% captures three types of directional shifts in the grayscale gradients of the image patches.
-These three gradients were shown to be captured well by the three-circle model $C_3$---the 1-skeleton of the Klein bottle.
-The $C_3$ model consists of three circles, two of which (say, $S_v$ and $S_h$) intersect the third (say, $S_\mathrm{lin}$) in exactly two points, but themselves do not intersect.
-One of the circles, $S_{\mathrm{lin}}$, parameterizes a the rotational gradient, whereas the other two $S_v$ and $S_h$ parameterize vertical and horizontal gradients.
-A projection of the data $X$ is shown in the left-most image in Figure~\ref{fig:patch_data_dgm}, along with 15 landmark points.
-It was shown in~\cite{carlsson2008local} that when $k \approx 300$, the top 30\% of data extracted from the knn density estimator $\rho_{300}$ `fills out' the two-manifold, yielding Betti numbers $\beta_0 = \beta_1 = 1$, whereas when $k$ is decreases to a more localized estimate, the three-circle model is recovered, which has Betti numbers $\beta_0 = 1, \beta_1 = 5$.
-
 Due to the aformentioned high complexity of the fibered barcode computation, we begin by working with a subset of the image patch data $\mathcal{X}$.
@@
 The use of furthest-point sampling gives us certain coverage guarantees that the geometry is approximately preserved within each level set, whereas the stratification ensures the original density of is approximated preserved as well.
-From this data set, we construct a Rips-(co)density bifiltration $\mathcal{F}$ using $\rho_{15}$ equipped with the geodesic metric computed over the $k = 15$ knn graph on $X$.
-
-In theory, this bifiltration should be well approximated by the $C_3$ model, which one ought to be able to determine by a careful inspection of the fibered barcode.
+From this data set, we construct a Rips-(co)density bifiltration using $\rho_{15}$ equipped with the geodesic metric computed over the same $k = 15$ knn graph on $X$.
+Finally, we record the number of column reductions needed to compute the fibered barcode at a variety of levels of coarsening using $\mathtt{pHcol}$, vineyards, and our moves approach.
+The results are summarized in Table~\ref{tab:barcode_templates}.
+We also record the number of 2-cells encountered along the path $\Gamma$ permutations that were applied throughout the computation of the barcode templates for both vineyards and moves, denoted in the table as $d_K$ and $d_{\mathrm{LCS}}$, respectively.
+%Specifically, we begin by extracting a representative subset of the set $X \subset \mathcal{X}$ of patches representing the top 30\% densest patches of the space of all patches $\mathcal{X}$, where the notion of density is approximated with a k-nearest neighbor (knn) density estimator $\rho$.
+Carlsson et al.~\cite{carlsson2008local} studied a variety of choices of $k$ in order to understand how localized the data were distributed around the Klein bottle model.
+He determined empirically that when $k \approx 15$, the top 30\% captures three types of directional shifts in the grayscale gradients of the image patches.
+These three gradients were shown to be captured well by the three-circle model $C_3$---the 1-skeleton of the Klein bottle.
+One of the circles, $S_{\mathrm{lin}}$, parameterizes a the rotational gradient, whereas the other two $S_v$ and $S_h$ parameterize vertical and horizontal gradients.
+%It was shown in~\cite{carlsson2008local} that when $k \approx 300$, the top 30\% of data extracted from the knn density estimator $\rho_{300}$ `fills out' the two-manifold, yielding Betti numbers $\beta_0 = \beta_1 = 1$, whereas when $k$ is decreases to a more localized estimate, the three-circle model is recovered, which has Betti numbers $\beta_0 = 1, \beta_1 = 5$.
+%In theory, this bifiltration should be well approximated by the $C_3$ model, which one ought to be able to determine by a careful inspection of the fibered barcode.
 That is, since $C_3$ has $\beta_1  = 5$, there should be some line $L$ which induces a 1-parameter filtration whose corresponding persistence diagram (for dimension $1$) displays five highly persistent features.
 We show an example of the data, the bigraded Betti numbers, and the persistence diagram extracted from a particular choice of line $L$ in Figure~\ref{fig:patch_data_dgm}.
-
-The example from~\ref{fig:patch_data_dgm} demonstrates a practical application of the rank invariant to uncovering the underlying topological structure characterizing a ``real-world'' data set.
+%The example from~\ref{fig:patch_data_dgm} demonstrates a practical application of the fiber barcode  to uncovering the underlying topological structure characterizing a ``real-world'' data set.
 Indeed, this is just one of the many applications of multidimensional persistence.
@@
 In Table~\ref{tab:barcode_templates}, we record a variety of statistics related to the computation of the rank invariant on the bifiltration constructed from the natural images data set.
-TODO: finish explanation
+%TODO: finish explanation
 %An increasingly refined 
-
 \begin{table}[h]
-\caption{Cost to compute all barcode templates $\mathcal{T}$ for various coarsening choices $\beta(M)$.}\label{tab:barcode_templates}
+\caption{Cost to computing $\mathcal{T}$ for various coarsening choices of $\beta(M)$.}\label{tab:barcode_templates}
 \centering
@@
 \end{table}
+
+\noindent
+As shown on the table, we're able to achieve a significant reduction in the number of total column operations needed to compute $\mathcal{T}$ compared to both vineyards and $\mathtt{pHcol}$, though it's worth noting that permutations are constant time for vineyards while they require at most $O(m)$ elementary operations for moves.
+Observe that initially, when there is a high degree of coarsening, vineyards is particularly inefficient and moves requires only about 3x less column operations that naively computing $\mathcal{T}$ independently.
+As the coarsening becomes more refined and more 2-cells are added to $\mathcal{A}(M)$, however, vineyards quickly becomes a much more viable option compared to $\mathtt{pHcol}$, though even at the highest coarsening we tested the gain in efficient was relatively small.
+In contrast, our proposed moves approach scales quite well with this refinement, requiring about $12\%$  and $\approx 5\%$ of the number of column operations as vineyards and $\mathrm{pHcol}$, respectively.
 

### 2021-08-20 move_schedules.tex
Updates from Overleaf

--- 2021-08-05 move_schedules.tex
@@
 	Persistence diagrams are known to vary continuously with respect to their input; this motivates the algorithmic study of persistence computations for time-varying filtered complexes.
-Computationally, simulating persistence dynamically can be reduced to maintaining a valid decomposition under adjacent transpositions in the filtration order.
+Computationally, simulating persistence dynamically can be reduced to maintaining a valid decomposition of the boundary matrix, under adjacent transpositions in the filtration order.
 In practice, the quadratic scaling in the number of  transpositions often makes this maintenance procedure slower than simply computing the decomposition from scratch, effectively limiting the application of dynamic persistence to relatively small data sets.
@@
 \subsection{Overview}\label{sec:overview} 
-The starting point to the work described in this paper is the stability results established in~\cite{cohen2007stability, cohen2006vines}. 
-Indeed, given a triangulable topological space equipped with a sufficiently tame continuous function, persistent homology captures the changes in topology across the sublevel sets of the space, and encodes them in a persistence diagram.
-The stability of persistence contends that small changes in the underlying function produce small changes in the corresponding diagrams:  if the function changes continuously, so too will the points on the persistence diagram.
+Given a triangulable topological space equipped with a sufficiently tame continuous function, persistent homology captures the changes in topology across the sublevel sets of the space, and encodes them in a persistence diagram.
+The stability of persistence contends
+that small changes in the underlying function produce small changes in the corresponding diagrams~\cite{cohen2007stability, cohen2006vines}:  if the function changes continuously, so too will the points on the persistence diagram.
 This motivates the study and application of persistence to time-varying or dynamic settings, like that of dynamic metric spaces~\cite{kim2020spatiotemporal}.
-As persistence-related computations tend to exhibit high algorithmic complexity (e.g.
-essentially cubic\footnote{For finite fields, it is known that the persistence computation reduces to the PLU factorization problem, which takes $O(m^\omega)$ where $\omega \approx 2.373$ is the matrix multiplication constant.} in the size of the underlying filtration~\cite{morozov2005persistence}), their adoption to dynamic settings poses a challenging computational problem.
-With the current state of the art tools, there is no recourse to the problem of computing persistence from a time-varying filtration containing millions of simplices at thousands of snapshots in time.
-An efficient means of updating the persistence of a [time-varying] filtration has far-reaching consequences: methods that vectorize persistence diagrams for machine learning purposes, like adaptive template functions~\cite{polanco2019adaptive} and persistence images~\cite{adams2017persistence}, or summary statistics like $\alpha$-smoothed Betti curves~\cite{ulmer2019topological}, all immediately become computationally viable tools in dynamic  settings.
+Focusing on the computational aspects of this problem has far-reaching consequences: methods that vectorize  persistence diagrams for machine learning purposes, like adaptive template functions~\cite{polanco2019adaptive} and persistence images~\cite{adams2017persistence}, or summary statistics like $\alpha$-smoothed Betti curves~\cite{ulmer2019topological}, all immediately become  computationally viable tools for time-varying settings.
 %This analogy provides a strong interpretation at gaining insight on the subtle topological changes that may occur in studying a continuous process.
 
-Cohen-Steiner et al.
-refer to a continuous 1-parameter family of persistence diagrams as a \emph{vineyard}, and they give in \cite{cohen2006vines} an efficient algorithm for their computation given a time-varying filtration.
-The vineyards approach can be interpreted as an extension of the \emph{matrix reduction} algorithm~\cite{zomorodian2005computing}, which computes the persistence diagram of a fixed filtration $K$ with $m$ simplices in $O(m^3)$ time.
+Cohen-Steiner et al. refer to a continuous 1-parameter family of persistence diagrams as a \emph{vineyard}, and they give in \cite{cohen2006vines} an efficient algorithm for their computation
+given a time-varying filtration. 
+Computing vineyards can be interpreted as an extension of the \emph{matrix reduction} algorithm~\cite{zomorodian2005computing}, which computes the persistence diagram of a fixed filtration $K$ with $m$ simplices in $O(m^3)$ time.
 The main step in the reduction algorithm is  a specific   decomposition $R = D V$ of the  boundary matrix $D$ of $K$. 
@@
 If one is interested in understanding how the persistent homology of a continuous function changes over time, then this algorithm is sufficient, for  homological critical points can only occur when the filtration order changes.
-Indeed, this algorithm is also efficient  asymptotically: if there are $d$ time-points where this filtration order changes, then the vineyards algorithm takes $O(m^3 + md)$ time: one $O(m^3)$-time reduction upfront for the filtration at time $t_0$ followed by one $O(m)$ operation to update the decomposition for the remaining time points $(t_1, t_2, \dots, t_d)$.
-When $d >> m$, this $O(md)$ approach is far more efficient than the $O(dm^3)$  strategy of computing the diagrams independently at every time point, what we call the ``naive'' approach.
-
-Despite its efficient in theory, vineyards is often not the method of choice in practice in dynamic settings. 
+If there are $d$ time-points where this order changes, then the vineyards algorithm takes $O(m^3 + md)$ time: one $O(m^3)$-time reduction upfront for the filtration at time $t_0$ followed by one $O(m)$ operation to update the decomposition for the remaining time points $(t_1, t_2, \dots, t_d)$.
+When $d >> m$, this $O(md)$ approach is far more efficient than the $O(dm^3)$  naive strategy of computing the diagrams independently at every time point.
+
+Despite the vineyards algorithm being efficient in theory, it is often not the method of choice in practice in dynamic settings. 
 While there is an increasingly rich ecosystem of software packages offering variations of the standard reduction algorithm (e.g.
@@
 many transpositions can be quite slow...
-it is sometimes much faster to simply recompute the $RU$-decomposition from scratch using the standard persistence algorithm.'' Indeed, whether performing the reductions naively or updating the decomposition dynamically using vineyards, they observed that obtaining the decomposition is the most computationally demanding aspect of their 2-D persistence algorithm.
+it is sometimes much faster to simply recompute the $RU$-decomposition from scratch using the standard persistence algorithm.'' Indeed, whether performing the reductions independently or updating the decomposition dynamically using vineyards, they observed that obtaining the decomposition is the most computationally demanding aspect of their 2-D persistence algorithm.
 
@@
 % in many time varying settings, a valid decomposition is often only sought after at a relatively sparse number of time points relative to the size of the (finite) set needed to generated by the family.
-By exploiting the ``donor'' concept introduced in in~\cite{busaryev2010tracking}, we are able to make a tradeoff in the number of times the decomposition is restored to a valid state, dramatically reducing the total number of column operations needed to apply an arbitrary permutation to the filtration.
+By making a tradeoff in the number of times we restore the decomposition to a valid state, we are able to reduce the total number of column operations needed to apply an arbitrary permutation to the filtration.
 This tradeoff, paired with a fast greedy heuristic explained in section~\ref{sec:proxy_objective}, yields an algorithm that can update a $R = DV$ decomposition more efficiently than vineyards in sparse contexts, making dynamic persistence more computationally tractable for a wider class of use-cases.
@@
  \subsection{A Motivating Example and Contributions}\label{sec:motivation} 
-As a motivating example, we describe a simple experiment illustrating why the vineyards algorithm does not always yield an efficient method in many time-varying applications. 
+As a motivating example, we describe a simple experiment illustrating why the vineyards algorithm does not always yield an efficient method in time-varying applications. 
 Consider a series of grayscale images (i.e.
@@
  \end{figure*}
-Snapshots of the image data and its corresponding time-varying filtration is depicted in Figure~\ref{fig:vineyards} (top right and bottom right, respectively), which we use as a baseline for comparing the standard algorithm $\texttt{pHcol}$ with vineyards.
+This time-varying filtration is depicted in Figure~\ref{fig:vineyards} (bottom right), which we use as a baseline for comparing the standard algorithm $\texttt{pHcol}$ with vineyards.
 Note the size of the filtration is fixed--no simplices are ever added or removed over time.
@@
   On the left side of Figure~\ref{fig:vineyards}, we compare the cumulative cost (in total number of column operations) of various approaches which one can use to simulate persistence dynamically. 
-  Since it is unknown a priori at what points in time the persistent homology will change in application settings, one solution is to discretize the time domain at $n$ evenly spaced points and compute the persistence diagram at each of these time points independently.
+  Since it is unknown a-priori at what points in  time the persistent homology will change in application settings, one solution is to discretize the time domain at $n$ evenly spaced points and compute the persistence diagram at each of these time points independently.
 An alternative approach is to construct a homotopy between two time-points, and then decompose this homotopy into adjacent transpositions--the vineyards approach.
 The former is often used in practice, but it is only an approximation; the latter is guaranteed to capture all homological changes in persistence that occur in simulating the 1-parameter family.
-The graph in Figure~\ref{fig:vineyards} depicts two different size approximations (purple, $n=100$ and green, $n=10$), wherein the reduction algorithm (\texttt{pHcol}) was applied at $n$ time points, and two exact strategies (black and blue), which compute all $\approx 7 \times 10^{4}$ diagrams given by the homotopy.
+  The graph in Figure~\ref{fig:vineyards} (left) depicts two different size approximations (purple, $n=100$ and green, $n=10$), wherein the reduction algorithm (\texttt{pHcol}) was applied at $n$ time points, and two exact strategies (black and blue), which compute all $\approx 7 \times 10^{4}$ diagrams given by the homotopy.
 As one can see from the figure, if one needs to simulate the homotopy exactly, the vineyards approach is indeed far more efficient than naively  applying the reduction algorithm independently at all time points.
-  However, when the discretization of the time domain is sparse enough, the naive approach actually performs less column operations than the vineyards strategy, while still capturing the main events.
+  However,  when the discretization of the time domain is sparse enough, the naive approach   performs less column operations than the vineyards strategy, while still capturing the main events.
 
@@
 That is, even in situations where one is updating a decomposition to a filtration that is  relatively ``close'', e.g.
-any two adjacent filtrations in the bottom right of Figure~\ref{fig:vineyards}, it is still more efficient to apply \texttt{pHcol} directly than  to iteratively update the decomposition (see the case where $n = 10$).
+any two adjacent filtrations in the bottom left of Figure~\ref{fig:vineyards}, it is still more efficient to apply \texttt{pHcol} directly than  to iteratively update the decomposition (see the case where $n = 10$).
  Indeed, any optimizations to the reduction algorithm (for example, the clearing optimization~\cite{chen2011persistent}) would only increase this disparity, making the vineyards algorithm less computationally attractive even in settings where $n$ is larger.
@@
 In light of these results, we give an alternative proxy-objective to minimize, provide bounds justifying its relevance to the original objective, and give an efficient $O(d \log d)$ algorithm for approximately solving this proxy minimization.
-A performance comparison with other reduction-based persistent homology computations is given, wherein our approach is demonstrated to be an order of magnitude more efficient than existing approaches in a few common sparse dynamic settings, such as flock simulation and 2D persistence computation.
+A performance comparison with other reduction-based persistent homology computations is given, wherein our approach is demonstrated to be an order of magnitude more efficient than existing approaches in sparse dynamic settings.
 We conclude with a discussion of situations where our strategy is most applicable and future directions for research.
@@
 Another popular optimization is to utilize the duality between   homology and cohomology~\cite{de2011dualities}, which dramatically improves the effectiveness of the clearing optimization~\cite{bauer2021ripser}.
-There are many other optimizations on the implementation side: the use of ranking functions defined on the combinatorial number system enables implicit cofacet enumeration, removing the need to store the boundary matrix explciitly; the apparent/emergent pairs optimization identifies columns whose pivot entries are unaffected by the reduction algorithm, reducing the total number of columns which need be reduced; sparse data structures such as bit-trees and lazy heaps allow for efficient column-wise additions with $\mathbb{Z}_2 = \mathbb{Z}/2\mathbb{Z}$ coefficients and effective $O(1)$ pivot entry retrieval, and so on~\cite{bauer2021ripser, bauer2017phat}.
+There are many other optimizations on the implementation side: the use of ranking functions defined on the combinatorial number system enables implicit cofacet enumeration, removing the need to store the boundary matrix explciitly; the apparent/emergent pairs optimization identifies columns whose pivot entries are unaffected by the reduction algorithm, reducing the total number of columns which need be reduced; sparse data structures such as bit-trees and lazy heaps allow for efficient column-wise additions with $\mathbb{Z}_2 = \mathbb{Z}/2\mathbb{Z}$ coefficients and effective $O(1)$ pivot entry retrieval, and so on~\cite{bauer2017phat, bauer2021ripser}.
 
@@
 In section 3 we introduce our proposed alternative to vineyards and provide efficient algorithms to make this alternative viable. 
-In section 4 we demonstrate a few applications of the proposed method. In section 5 we conclude the paper by discussing other possible applications and future work. 
+In section 4 we demonstrate a few applications of the proposed method. In section 5 we conclude the paper by discussing other possible applications, future work. 
 We also include an appendix which introduces certain topics in more detail, such as the vineyards algorithm itself, for readers who may be unfamiliar but would like a more in-depth understanding.
@@
 Due to the combinatorial aspect of our approach, there is actually no need to have access to---or explicitly store---the entire family of filtrations and the schedules between them.
-That is, Algorithm~\ref{alg:schedule} can be easily modified to be completely online, keeping at most two filtrations and one decomposition in memory at any given time.
+The entire approach can be performed online, keeping at most two filtrations and one decomposition in memory at any given time.
 
@@
 In the case of vineyards, the bound is clear: whether assuming a homotopy between a pair of filtrations $(K_0, K_1)$ or not, if $\kappa = K_\tau(K_0, K_1)$, then $\Omega(\kappa)$ adjacent transpositions are required to transform $K_0 \mapsto K_1$.
-In contrast, if we allow move operations in any order, we have $O(m)$ as a trivial upper bound: simply move each simplex in $K_0$ into its position in the filtration given by $K_1$ in the order given by the latter.
+Allowing move operations in any order, we have $O(m)$ as a trivial upper bound: simply move each simplex in $K_0$ into its position in the filtration given by $K_1$ in the order given by the latter.
 However, it's not immediately clear whether this bound is tight.
@@
 	The history of the mathematics of estimating this quantity dates back at least 50 years, and is sometimes called the \emph{Ulam-Hammersley} problem.
-The seminal work by Baik et al.~\cite{baik1999distribution} established that as $m \to \infty$:
+It was shown by Baik et al.~\cite{} that as $m \to \infty$:
 	$$ \displaystyle \ell_m = 2 \sqrt{m} + c m^{1/6} + o(m^{1/6}) $$
-where $c = -1.77108...$. As a result, we have: 
+where $c \approx -1.77108$. As a result, we have: 
 $$ \frac{\ell_m}{\sqrt{m}} \to 2\rlap{\quad \text{as } $m \to \infty$} $$	 
@@
 It is worth noting that this bound applies to pairs of uniformly sampled permutations, as opposed to uniformly sampled distinct filtrations---intuitively, simplicial filtrations have more structure due to the subcomplex relation that must be respected in the filtration.
-Thus, while $O(m - \sqrt{m})$ in expectation is an improvement over $O(m)$, it is still not tight.
+Thus, while $O(m - \sqrt{m})$ in expectation is an improvement over $O(m)$ it is still not tight.
 However, Boissonnat~\cite{boissonnat2018efficient} has shown that for any $d$-dimensional simplicial complex $K$ with $m$ simplices, the number of distinct filtrations built from $K$ is \emph{at least} $\floor{\frac{t+1}{d+1}}^m$, where $t$ is the number of distinct filtration values associated with the complex $K$.
-Since this lower bound grows similarly to $m!$ when $t \sim O(m)$ and $d << m$ fixed, we expect $O(m - \sqrt{m})$ is not too pessimistic of an upper bound when the pair of filtrations are truly random.
+Since this lower bound grows similarly to $m!$ when $t \approx O(m)$ and $d << m$ fixed, we expect $O(m - \sqrt{m})$ is not too pessimistic of an upper bound when the pair of filtrations are truly random.
 In practice, when one has a time-varying filtration and the sampling points are relatively close [in time], the LCS between adjacent filtrations is expected to be much larger, subsequently making $d$ much smaller.
-For example, the average size of the LCS across the 10 filtrations from Section~\ref{sec:motivation} each having $m=417$ simplices sampled at evenly spaced time points was $343$, implying $d \approx 70$ permutations needed on average to update the decomposition between adjacent time points.
+This had been confirmed empirically, e.g.
+the average size of the LCS across the 10 filtrations from Section~\ref{sec:motivation} each having $417$ simplices sampled at evenly spaced time points was $343$, implying $d \approx 70$ permutations needed on average to update the decomposition between adjacent time points.
 
@@
 The gray areas indicate the region where the dimension function is constant.
-In particular, the light gray area of the middle plot shows where $\mathrm{dim}_1(M) = 1$, indicating where in the multifiltration a persistent 1-cycle was detected.
 On the right side, the space dual to the space in the middle is shown.
@@
 The blue regions highlights where $\mathrm{dim}(M) = 5$.
-On the right, the persistence diagram of $M$ restricted to $L$, with larger points indicating a multiplicity of 2.
+On the right, the persistence diagram of $L$, with larger points indicating a multiplicity of 2.
 Observe five persistent features are revealed in both the middle and right plots, matching $\beta_1$ of the three-circle model.
@@
 An exemplary demonstration of this is given in~\cite{lee2003nonlinear}, who explored the space of $3 \times 3$ high-contrast patches extracted from Hans van Hateren's still image collection~\cite{hateren_schaaf_1998}, which consists of approximately 4,000 monochrome images\footnote{See \url{http://bethgelab.org/datasets/vanhateren/} for details on the image collection.} depicting various areas outside around Groningen (Holland).
-Motivated by applications to sparse coding applications, Lee et al.~\cite{lee2003nonlinear} was interested in how these natural image patches were distributed, in pixel-space, with respect to predicted spaces and manifolds.
-For example, one of the motivating questions of their research was whether there were any clear qualitative differences the distributions of patches extracted from from images of different modalities, e.g.
+The study of how empirical data extracted from natural images is distributed with respect to predicted spaces and manifolds is motivated by applications in sparse coding and efficient image representations.
+In particular, Lee et al.~\cite{lee2003nonlinear} was interested in analyzing the empirical probability distribution of high-contrast $3 \times 3$ patches of natural images, in pixel-space, towards understanding whether there were any clear qualitative differences the distributions of patches extracted from from images of different modalities, e.g.
 optical versus range images.
-A major result from their work is that the majority of high-contrast image patches are concentrated around a 2-dimensional submanifold embedded in $S^7$.
-Supplemental experiments conducted by Carllsson et al.~\cite{carlsson2008local} using persistent homology confirmed that the distribution of high-contrast $3 \times 3$ patches is indeed well-approximated by a Klein bottle---around 60\% of the high-contrast patches from the still image data set are concentrated within a small neighborhood around a model of a Klein bottle which accounts for $\approx$ 21\% of the volume $S^7$, compared to the $\approx$ 84\% of volume of $S^7$ accounted for by the entire data set.
-Along a similar vein, in the context of sparse coding, Perea et al.~\cite{perea2014klein} introduced a dictionary learning framework for estimating and representing the distribution of patches from texture images.
-
-If one was not aware of the analysis done by~\cite{lee2003nonlinear, hateren_schaaf_1998, carlsson2008local, perea2014klein}, it is not immediately clear a priori that the Klein bottle model is indeed an excellent candidate for capturing the non-linearity of the data.
-Nonetheless, as we have mentioned, applications of persistent homology detecting the presence of a topological space persistent homology is designed to reveal.
+Experiments conducted by Carllsson et al.~\cite{carlsson2008local} using persistent homology confirmed that the distribution of high-contrast $3 \times 3$ patches is indeed well-approximated by a Klein bottle---around 60\% of the high-contrast patches from the still image data set are concentrated within a small neighborhood around a model of a Klein bottle which accounts for $\approx$ 21\% of the volume $S^7$, compared to the $\approx$ 84\% of volume of $S^7$ accounted for by the entire data set.
+Along a similar vein but from a sparse coding perspective, Perea et al.~\cite{perea2014klein} introduced dictionary learning framework for estimating and representing the distribution of patches from texture images.
+
+If one was not aware of the analysis done by~\cite{lee2003nonlinear, hateren_schaaf_1998, carlsson2008local, perea2014klein}, there is no specific reason apriori to assume the data are well-approximated by the Klein bottle.
 Consider the (coarsened) fibered barcode shown in Figure~\ref{fig:patch_data_dgm} which was computed from a standard Rips/codensity bifiltration on a constructed on a representative sample of the image data from~\cite{hateren_schaaf_1998} (how we obtained this sample is described further down).

### 2021-08-21 move_schedules.tex
Updated for SIAM format

--- 2021-08-05 move_schedules.tex
@@
-\documentclass[10pt]{article}
-\usepackage[margin=0.75in]{geometry}
-\usepackage{amsmath, amsthm, amssymb}
+\documentclass{siamart190516}
+
+%\usepackage[margin=0.75in]{geometry}
+\usepackage{amsmath, amssymb}
 \usepackage{mathtools}
@@
 
+\newcommand\undermat[2]{%
+  	\makebox[0pt][l]{$\smash{\underbrace{\phantom{%
+    \begin{matrix}#2\end{matrix}}}_{\text{$#1$}}}$}#2}
+
 \usepackage{hyperref} % load hyperref last
 
-\newtheorem{theorem}{Theorem}
-\newtheorem{definition}{Definition}
-\newtheorem{proposition}{Proposition}
-\newtheorem{lemma}{Lemma}
-\newtheorem{corollary}{Corollary}
+%\newtheorem{theorem}{Theorem}
+%\newtheorem{definition}{Definition}
+%\newtheorem{proposition}{Proposition}
+%\newtheorem{lemma}{Lemma}
+%\newtheorem{corollary}{Corollary}
 % \newcommand{\CommentX}[1]{\unskip~\%~#1~\%}
@@
 
-%Unraveling
-\title{\vspace{-2.0em} Move Schedules: \\ Fast persistence computations in sparse dynamic settings\thanks{This work was partially supported by the National Science Foundation through grants CCF-2006661
-and CAREER award  DMS-1943758.}\vspace{-0.5em}}
-\author{Matt Piekenbrock\thanks{Department of Computational Mathematics, Science \& Engineering, Michigan State University}\;\; and Jose A.
+% Sets running headers as well as PDF title and authors
+\headers{An Example Article}{D. Doe, P. T. Frank, and J. E. Smith}
+
+% Title. If the supplement option is on, then "Supplementary Material"
+% is automatically inserted before the title.
+\title{Move Schedules: Fast persistence computations in sparse dynamic settings
+	\thanks{Submitted to the editors DATE.
+		\funding{This work was partially supported by the National Science Foundation through grants CCF-2006661 and CAREER award DMS-1943758.}
+	}
+}
+
+% Authors: full names plus addresses.
+\author{
+	Matt Piekenbrock\thanks{Department of Computational Mathematics, Science \& Engineering, Michigan State University.}
+	\and Jose A. Perea\thanks{Department of Mathematics, Michigan State University.}
+}
+
+%\title{\vspace{-2.0em} Move Schedules: \\ Fast persistence computations in sparse dynamic settings\thanks{This work was partially supported by the National Science Foundation through grants CCF-2006661 and CAREER award DMS-1943758.}\vspace{-0.5em}}
+%\author{Matt Piekenbrock\thanks{Department of Computational Mathematics, Science \& Engineering, Michigan State University}\;\; and Jose A.
 Perea$^\dagger$\thanks{Department of Mathematics, Michigan State University}}
-\date{}
-
-\begin{document} \vspace{-4em} \maketitle
+%\date{}
+
+% Optional PDF information
+\ifpdf
+\hypersetup{
+  pdftitle={An Example Article},
+  pdfauthor={Matt Piekenbrock, P. T. Frank, and J. E. Smith}
+}
+\fi
+
+% ------  Begin document ------- 
+\begin{document}
+
+\maketitle
+
+% REQUIRED - "Abstracts must be able to stand alone and so cannot contain citations to the paper's references, equations, etc. 
 \begin{abstract}
@@
 
+% REQUIRED
+\begin{keywords}
+  example, \LaTeX
+\end{keywords}
+
+% REQUIRED
+\begin{AMS}
+  68Q25, 68R10, 68U05
+\end{AMS}
+
 \section{Introduction} 
@@
  	\centering
- 	\includegraphics[width=0.36\textwidth, height = 1.65in]{circle_vineyards.png}
- 	\includegraphics[width=0.63\textwidth]{circle_complex.png}
- 	\caption{Grayscale image data set depicting an expanding annulus (top right), each image's corresponding lower star filtration (bottom right), and the cumulative number of column operations required to compute persistence for this time-varying filtration (left).
+ 	\includegraphics[width=0.36\textwidth, height = 1.45in]{circle_vineyards.png}
+ 	\includegraphics[width=0.63\textwidth, height = 1.45in]{circle_complex.png}
+ 	\caption{On the right, the grayscale image data set and corresponding lower star filtrations.
+On the left, the cumulative number of column operations required to compute persistence for this time-varying filtration.
 The persistent Betti numbers for each filtration from left to right are: $\beta_1 = (1,0,0), \beta_2 = (1,1,0), \beta_3 = (1,1,0), \beta_4 = (1,1,0), \beta_5 = (4,0,0)$.
@@
 In light of these results, we give an alternative proxy-objective to minimize, provide bounds justifying its relevance to the original objective, and give an efficient $O(d \log d)$ algorithm for approximately solving this proxy minimization.
-A performance comparison with other reduction-based persistent homology computations is given, wherein our approach is demonstrated to be an order of magnitude more efficient than existing approaches in a few common sparse dynamic settings, such as flock simulation and 2D persistence computation.
+A performance comparison with other reduction-based persistent homology computations is given, wherein our approach is demonstrated to be an order of magnitude more efficient than existing approaches in applications such as flock simulation and 2D persistence computation.
 We conclude with a discussion of situations where our strategy is most applicable and future directions for research.
@@
 \begin{displaymath}
+	\def\rowsp{0.4em}
 	\def\arraystretch{0.95}
-	\makeatletter\setlength\BA@colsep{2.8pt}\makeatother
+	\makeatletter\setlength\BA@colsep{1.2pt}\makeatother
 	%
@@
 		\begin{block}{c[ccc]}
-  		u \kern-0.1em  & \; 1 &    &  1 \; \topstrut \\
-  		v \kern-0.1em & \;     &  1 & 1 \; \\
-  		w \kern-0.1em & \; 1 &  1 &    \; \botstrut \\
+  		u \kern\rowsp  & \; 1 &    &  1 \; \topstrut \\
+  		v \kern\rowsp & \;     &  1 & 1 \; \\
+  		w \kern\rowsp & \; 1 &  1 &    \; \botstrut \\
 		\end{block}
@@
 		\begin{block}{c[ccc]}
-        a \kern-0.1em  & \; 1 &  &  \; \topstrut \\
-  		b \kern-0.1em & \;  & 1 & \; \\
-  		c \kern-0.1em & \;  &  & 1 \; \botstrut \\
+        a \kern\rowsp  & \; 1 &  &  \; \topstrut \\
+  		b \kern\rowsp & \;  & 1 & \; \\
+  		c \kern\rowsp & \;  &  & 1 \; \botstrut \\
 		\end{block}
@@
 		\begin{block}{c[ccc]}
-  		u \kern-0.1em  & \; 1 & 1 & 1 \; \topstrut \\
-  		v \kern-0.1em & \;     & 1 & 1 \; \\
-  		w \kern-0.1em & \; 1 &    &    \; \botstrut \\
+  		u \kern\rowsp  & \; 1 & 1 & 1 \; \topstrut \\
+  		v \kern\rowsp & \;     & 1 & 1 \; \\
+  		w \kern\rowsp & \; 1 &    &    \; \botstrut \\
 		\end{block}
@@
 		\begin{block}{c[ccc]}
-  		a \kern-0.1em  & \; 1 & 1 &    \; \topstrut \\
-  		b \kern-0.1em & \;     & 1 &    \; \\
-  		c \kern-0.1em & \;     &    & 1 \; \botstrut \\
+  		a \kern\rowsp  & \; 1 & 1 &    \; \topstrut \\
+  		b \kern\rowsp & \;     & 1 &    \; \\
+  		c \kern\rowsp & \;     &    & 1 \; \botstrut \\
 		\end{block}
@@
 		\begin{block}{c[ccc]}
-  		u \kern-0.1em  & \;     & 1  &  \; \topstrut \\
-  		v \kern-0.1em &  \; 1  & 1  &  \; \\
-  		w \kern-0.1em & \; 1  &     &  \; \botstrut \\
+  		u \kern\rowsp  & \;     & 1  &  \; \topstrut \\
+  		v \kern\rowsp &  \; 1  & 1  &  \; \\
+  		w \kern\rowsp & \; 1  &     &  \; \botstrut \\
 		\end{block}
@@
 		\begin{block}{c[ccc]}
-  		b \kern-0.1em  & \; 1 & 1 & 1 \; \topstrut \\
-  		a \kern-0.1em & \;     & 1 & 1 \; \\
-  		c \kern-0.1em & \;     &    & 1 \; \botstrut \\
+  		b \kern\rowsp  & \; 1 & 1 & 1 \; \topstrut \\
+  		a \kern\rowsp & \;     & 1 & 1 \; \\
+  		c \kern\rowsp & \;     &    & 1 \; \botstrut \\
 		\end{block}
@@
 \end{displaymath}
+
 The column $c$ in $R$ indicates that dimension $1$ homology is born, which in this filtration is never killed since the dimension of the complex is $2$.
@@
 Given any $R = D V$ decomposition, it was shown in~\cite{zomorodian2005computing} that the following proposition is true:
-\begin{proposition}[Persistent Pairing~\cite{zomorodian2005computing}]
-	Let $R = D V$, where $R$ is in reduced form and $V$ is full-rank upper triangular (invariants (1) and (2)  hold).
+%\begin{proposition}[Persistent Pairing~\cite{zomorodian2005computing}]
+%	Let $R = D V$, where $R$ is in reduced form and $V$ is full-rank upper triangular (invariants (1) and (2)  hold).
 Then the pairing $(\sigma_i, \sigma_j)$ defined by $i = \mathrm{low}_R(j)$ is a \emph{creator/destroyer (persistent) pair}.
 While the matrices $R,V$ satisfying (1) and (2) are not unique in general,  the persistent pairs $(i,j)$ are unique and define the persistence diagrams of $K$.
-\end{proposition}
+%\end{proposition}
 \noindent
@@
 		\begin{block}{c[ccc]}
-  		u \kern-0.1em  & \; 1 & 1 &  \; \topstrut \\
-  		v \kern-0.1em & \;  & 1 &  \; \\
-  		w \kern-0.1em & \; 1 &  &  \; \botstrut \\
-		\end{block}
-	\end{blockarray}
-	\, , 
+  		u \kern\rowsp  & \; 1 & 1 &  \; \topstrut \\
+  		v \kern\rowsp & \;  & 1 &  \; \\
+  		w \kern\rowsp & \; 1 &  &  \; \botstrut \\
+		\end{block}
+	\end{blockarray}
+	\,\,, 
 	\begin{blockarray}{cccc}{V_1}
@@
 		\begin{block}{c[ccc]}
-        a \kern-0.1em  & \; 1 & \tikz[baseline]{
-            \node[fill=orange!20,anchor=base] (t1)
-            {$ 1 $};
-        } & 1 \; \topstrut \\
-  		b \kern-0.1em & \;  & 1 & 1 \; \\
-  		c \kern-0.1em & \;  &  & 1 \; \botstrut \\
+        a \kern\rowsp  & \; 1 & 1 & 1 \; \topstrut \\
+  		b \kern\rowsp & \;  & 1 & 1 \; \\
+  		c \kern\rowsp & \;  &  & 1 \; \botstrut \\
 		\end{block}
@@
 	\xrightarrow{S_{1}^2} 
-	\hspace{-0.5em}
-	\begin{blockarray}{cccc}{R_1}
+	\hspace{-0.1em}
+	\begin{blockarray}{cccc}{}
 	\kern-0.2em & a & b & c  \\
 		\begin{block}{c[ccc]}
-  		u \kern-0.1em  & \; 1 &  &  \; \topstrut \\
-  		v \kern-0.1em & \;  & 1 &  \; \\
-  		w \kern-0.1em & \; 1 & \tikz[baseline]{
-            \node[fill=red!20,anchor=base] (t1)
-            {$ 1 $};
-        } &  \; \botstrut \\
-		\end{block}
-	\end{blockarray}
-	\, , 
-	\begin{blockarray}{cccc}{V_1}
+  		 & \; 1 &  &  \; \topstrut \\
+  		 & \;  & 1 &  \; \\
+  		 & \; 1 & 1 & \; \botstrut \\
+		\end{block}
+	\end{blockarray}
+	\,\,, 
+	\begin{blockarray}{cccc}{}
 	\kern-0.2em & a & b & c  \\
 		\begin{block}{c[ccc]}
-  		a \kern-0.1em  & \; 1 &  & 1 \; \topstrut \\
-  		b \kern-0.1em & \;  & 1 & 1 \; \\
-  		c \kern-0.1em & \;  &  & 1 \; \botstrut \\
+  		 & \; 1 &  & 1 \; \topstrut \\
+  		 & \;  & 1 & 1 \; \\
+  		 & \;  &  & 1 \; \botstrut \\
 		\end{block}
@@
 	\xrightarrow{P}
-	\hspace{-0.5em}
+	\hspace{-0.2em}
+	\begin{blockarray}{cccc}{}
+	\kern-0.2em & b & a & c  \\
+		\begin{block}{c[ccc]}
+  		& \;  & 1 &  \; \topstrut \\
+  		& \; 1 &  &  \; \\
+  		& \; 1 & 1 &  \; \botstrut \\
+		\end{block}
+	\end{blockarray}
+	\,\,, 
+	\begin{blockarray}{cccc}{}
+	\kern-0.2em & b & a & c  \\
+		\begin{block}{c[ccc]}
+  		& \; 1 &  & 1 \; \topstrut \\
+  		& \;  & 1 & 1 \; \\
+  		& \;  &  & 1 \; \botstrut \\
+		\end{block}
+	\end{blockarray}
+	\xrightarrow{S_1^2}
+	\hspace{-0.1em}
 	\begin{blockarray}{cccc}{R_1}
@@
 		\begin{block}{c[ccc]}
-  		u \kern-0.1em  & \;  & 1 &  \; \topstrut \\
-  		v \kern-0.1em & \; 1 &  &  \; \\
-  		w \kern-0.1em & \; 1 & \tikz[baseline]{
-            \node[fill=red!20,anchor=base] (t1)
-            {$ 1 $};
-        } &  \; \botstrut \\
-		\end{block}
-	\end{blockarray}
-	\, , 
+  		u \kern\rowsp  & \;  & 1 &  \; \topstrut \\
+  		v \kern\rowsp & \; 1 & 1 &  \; \\
+  		w \kern\rowsp & \; 1 &  &  \; \botstrut \\
+		\end{block}
+	\end{blockarray}
+	\,\,,
 	\begin{blockarray}{cccc}{V_1}
@@
 		\begin{block}{c[ccc]}
-  		b \kern-0.1em  & \; 1 &  & 1 \; \topstrut \\
-  		a \kern-0.1em & \;  & 1 & 1 \; \\
-  		c \kern-0.1em & \;  &  & 1 \; \botstrut \\
-		\end{block}
-	\end{blockarray}
-	\xrightarrow{S_1^2}
-	\hspace{-0.5em}
-	\begin{blockarray}{cccc}{R_1}
-	\kern-0.2em & b & a & c  \\
-		\begin{block}{c[ccc]}
-  		u \kern-0.1em  & \;  & 1 &  \; \topstrut \\
-  		v \kern-0.1em & \; 1 & 1 &  \; \\
-  		w \kern-0.1em & \; 1 &  &  \; \botstrut \\
-		\end{block}
-	\end{blockarray}
-	\, ,
-	\begin{blockarray}{cccc}{V_1}
-	\kern-0.2em & b & a & c  \\
-		\begin{block}{c[ccc]}
-  		b \kern-0.1em  & \; 1 & 1 & 1 \; \topstrut \\
-  		a \kern-0.1em & \;  & 1 & 1 \; \\
-  		c \kern-0.1em & \;  &  & 1 \; \botstrut \\
+  		b \kern\rowsp  & \; 1 & 1 & 1 \; \topstrut \\
+  		a \kern\rowsp & \;  & 1 & 1 \; \\
+  		c \kern\rowsp & \;  &  & 1 \; \botstrut \\
 		\end{block}
@@
 		\begin{block}{c[c]}
-  		u \kern-0.1em  & \; 1 \; \topstrut \\
-  		v \kern-0.1em & \;  \; \\
-  		w \kern-0.1em & \; 1 \; \botstrut \\
+  		u \kern \rowsp & \; 1 \; \topstrut \\
+  		v \kern \rowsp & \;  \; \\
+  		w \kern \rowsp & \; 1 \; \botstrut \\
 		\end{block}
@@
 		\begin{block}{c[ccc]}
-  		u \kern-0.1em  & \; 1 & 1 &  \; \topstrut \\
-  		v \kern-0.1em & \;  & 1 &  \; \\
-  		w \kern-0.1em & \; 1 &  &  \; \botstrut \\
+  		u \kern \rowsp & \; 1 & 1 &  \; \topstrut \\
+  		v \kern \rowsp & \;  & 1 &  \; \\
+  		w \kern \rowsp & \; 1 &  &  \; \botstrut \\
 		\end{block}
@@
 	%
-\begin{blockarray}{cc}{d_R}
+\begin{blockarray}{cc}{}
 	\kern-0.2em & b \\
 		\begin{block}{c[c]}
-  		u \kern-0.1em  & \; 1 \; \topstrut \\
-  		v \kern-0.1em & \;  1 \; \\
-  		w \kern-0.1em & \; \; \botstrut \\
+  		  & \; 1 \; \topstrut \\
+  		 & \;  1 \; \\
+  		 & \; \; \botstrut \\
 		\end{block}
@@
 	\; \;
-\begin{blockarray}{cccc}{R_1}
+\begin{blockarray}{cccc}{}
 	\kern-0.2em & a & b & c  \\
 		\begin{block}{c[ccc]}
-  		u \kern-0.1em  & \; 1 &    &  \; \topstrut \\
-  		v \kern-0.1em  & \;    & 1 &  \; \\
-  		w \kern-0.1em & \; 1 & 1 &  \; \botstrut \\
+  		  & \; 1 &    &  \; \topstrut \\
+  		  & \;    & 1 &  \; \\
+  		 & \; 1 & 1 &  \; \botstrut \\
 		\end{block}
@@
 	%
-	\begin{blockarray}{cc}{d_R}
+	\begin{blockarray}{cc}{}
 	\kern-0.2em & c \\
 		\begin{block}{c[c]}
-  		u \kern-0.1em  & \; \hphantom{1} \; \topstrut \\
-  		v \kern-0.1em & \;  \hphantom{1} \; \\
-  		w \kern-0.1em & \; \hphantom{1} \; \botstrut \\
+  		  & \; \hphantom{1} \; \topstrut \\
+  		 & \;  \hphantom{1} \; \\
+  		 & \; \hphantom{1} \; \botstrut \\
 		\end{block}
@@
 	\; \; 
-	\begin{blockarray}{cccc}{R_1}
+	\begin{blockarray}{cccc}{}
 	\kern-0.2em & a & b &  c \\
 		\begin{block}{c[ccc]}
-  		u \kern-0.1em  & \; 1   &    &   1  \; \topstrut \\
-  		v \kern-0.1em & \;       &  1  &  1  \; \\
-  		w \kern-0.1em & \; 1   &   1   &     \; \botstrut \\
+  		  & \; 1   &    &   1  \; \topstrut \\
+  		 & \;       &  1  &  1  \; \\
+  		 & \; 1   &   1   &     \; \botstrut \\
 		\end{block}
@@
 \; \xrightarrow{P} 
-% 
-\begin{blockarray}{cc}{d_R}
+\begin{blockarray}{cc}{}
 	\kern-0.2em & c \\
 		\begin{block}{c[c]}
-  		u \kern-0.1em  & \; \hphantom{1} \; \topstrut \\
-  		v \kern-0.1em & \; \hphantom{1} \; \\
-  		w \kern-0.1em & \; \hphantom{1} \; \botstrut \\
+  		  & \; \hphantom{1} \; \topstrut \\
+  		 & \; \hphantom{1} \; \\
+  		 & \; \hphantom{1} \; \botstrut \\
 		\end{block}
@@
 \; \;
+\begin{blockarray}{cccc}{}
+	\kern-0.2em & b & c & a  \\
+		\begin{block}{c[ccc]}
+  		  & \;  & 1 &  1\; \topstrut \\
+  		 & \; 1 & 1 &  \; \\
+  		 & \; 1 &  & 1 \; \botstrut \\
+		\end{block}
+	\end{blockarray}
+%
+\; \xrightarrow{d_R} 
 \begin{blockarray}{cccc}{R_1}
@@
 	\begin{block}{c[ccc]}
-  		u \kern-0.1em  & \;  & 1 &  1\; \topstrut \\
-  		v \kern-0.1em & \; 1 & 1 &  \; \\
-  		w \kern-0.1em & \; 1 &  & 1 \; \botstrut \\
-		\end{block}
-	\end{blockarray}
-%
-\; \xrightarrow{d_R}
-% 
-	\begin{blockarray}{cccc}{R_1}
-	\kern-0.2em & b & c & a  \\
-		\begin{block}{c[ccc]}
-  		u \kern-0.1em  & \;  & 1 &  \; \topstrut \\
-  		v \kern-0.1em & \; 1 & 1 &  \; \\
-  		w \kern-0.1em & \; 1 &  &  \; \botstrut \\
-		\end{block}
+		u \kern \rowsp & \;  & 1 &  \; \topstrut \\
+		v \kern \rowsp & \; 1 & 1 &  \; \\
+		w \kern \rowsp & \; 1 &  &  \; \botstrut \\
+	\end{block}
+	% \undermat{B}{ & & }  
 \end{blockarray}
 \end{displaymath}
+
 \vspace{-1.0em}
+
 % Start V example 
 \begin{displaymath}
+	\def\rowsp{0.4em}
 	\def\arraystretch{0.95}
-	\makeatletter\setlength\BA@colsep{2.8pt}\makeatother
+	\makeatletter\setlength\BA@colsep{1.2pt}\makeatother
 	%
@@
 		\begin{block}{c[c]}
-  		a \kern-0.1em  & \; 1 \; \topstrut \\
-  		b \kern-0.1em & \;  \; \\
-  		c \kern-0.1em & \;  \; \botstrut \\
+  		 & \; 1 \; \topstrut \\
+  		& \;  \; \\
+  		& \;  \; \botstrut \\
 		\end{block}
@@
 		\begin{block}{c[ccc]}
-  		a \kern-0.1em  & \; 1 & 1 &  1\; \topstrut \\
-  		b \kern-0.1em & \;  & 1 &  1\; \\
-  		c \kern-0.1em & \; &  & 1 \; \botstrut \\
+  		 & \; 1 & 1 &  1\; \topstrut \\
+  		& \;  & 1 &  1\; \\
+  		& \; &  & 1 \; \botstrut \\
 		\end{block}
@@
 	%
-\begin{blockarray}{cc}{d_V}
+\begin{blockarray}{cc}{}
 	\kern-0.2em & b \\
 		\begin{block}{c[c]}
-  		a \kern-0.1em  & \; 1 \; \topstrut \\
-  		b \kern-0.1em & \;  1 \; \\
-  		c \kern-0.1em & \; \; \botstrut \\
+  		 & \; 1 \; \topstrut \\
+  		& \;  1 \; \\
+  		& \; \; \botstrut \\
 		\end{block}
@@
 	\; \;
-\begin{blockarray}{cccc}{V_1}
+\begin{blockarray}{cccc}{}
 	 \kern-0.2em & a & b & c  \\
 		\begin{block}{c[ccc]}
-  		a \kern-0.1em  & \; 1 & &  1\; \topstrut \\
-  		b \kern-0.1em & \;  & 1 &  1\; \\
-  		c \kern-0.1em & \; &   & 1 \; \botstrut \\
+  		 & \; 1 & &  1\; \topstrut \\
+  		& \;  & 1 &  1\; \\
+  		& \; &   & 1 \; \botstrut \\
 		\end{block}
@@
 	%
-	\begin{blockarray}{cc}{d_V}
+	\begin{blockarray}{cc}{}
 	\kern-0.2em & c \\
 		\begin{block}{c[c]}
-  		a \kern-0.1em  & \; 1 \; \topstrut \\
-  		b \kern-0.1em & \;  1 \; \\
-  		c \kern-0.1em & \; 1 \; \botstrut \\
+  		 & \; 1 \; \topstrut \\
+  		& \;  1 \; \\
+  		& \; 1 \; \botstrut \\
 		\end{block}
@@
 	\; \;
-	\begin{blockarray}{cccc}{V_1}
+	\begin{blockarray}{cccc}{}
 	\kern-0.2em & a & b &  c \\
 		\begin{block}{c[ccc]}
-  		a \kern-0.1em  & \; 1 & &   \; \topstrut \\
-  		b \kern-0.1em & \;  & 1 &   \; \\
-  		c \kern-0.1em & \;  &   & 1 \; \botstrut \\
-		\end{block}
-	\end{blockarray}
-%
-\; \xrightarrow{P}
-% 
-\begin{blockarray}{cc}{d_V}
+  		 & \; 1 & &   \; \topstrut \\
+  		& \;  & 1 &   \; \\
+  		& \;  &   & 1 \; \botstrut \\
+		\end{block}
+	\end{blockarray}
+%
+\; \xrightarrow{P} \\
+% 
+\begin{blockarray}{cc}{}
 	\kern-0.2em & c \\
 		\begin{block}{c[c]}
-  		b \kern-0.1em  & \; 1 \; \topstrut \\
-  		c \kern-0.1em & \;  1 \; \\
-  		a \kern-0.1em & \; 1 \; \botstrut \\
+  		 & \; 1 \; \topstrut \\
+  		& \;  1 \; \\
+  		& \; 1 \; \botstrut \\
 		\end{block}
@@
 \; \;
+\begin{blockarray}{cccc}{}
+	\kern-0.2em & b & c & a  \\
+		\begin{block}{c[ccc]}
+  		 & \; 1 & &\; \topstrut \\
+  		& \;  & 1 &  \; \\
+  		& \; &  & 1 \; \botstrut \\
+		\end{block}
+	\end{blockarray}
+%
+\; \xrightarrow{d_V}
+% 
 	\begin{blockarray}{cccc}{V_1}
@@
 		\begin{block}{c[ccc]}
-  		b \kern-0.1em  & \; 1 & &\; \topstrut \\
-  		c \kern-0.1em & \;  & 1 &  \; \\
-  		a \kern-0.1em & \; &  & 1 \; \botstrut \\
-		\end{block}
-	\end{blockarray}
-%
-\; \xrightarrow{d_V}
-% 
-	\begin{blockarray}{cccc}{V_1}
-	\kern-0.2em & b & c & a  \\
-		\begin{block}{c[ccc]}
-  		b \kern-0.1em  & \; 1 & & 1 \; \topstrut \\
-  		c \kern-0.1em & \; & 1 &  1 \; \\
-  		a \kern-0.1em & \; &  & 1  \; \botstrut \\
+  		b \kern \rowsp & \; 1 & & 1 \; \topstrut \\
+  		c \kern \rowsp & \; & 1 &  1 \; \\
+  		a \kern \rowsp & \; &  & 1  \; \botstrut \\
 		\end{block}
@@
 Recall from section~\ref{sec:background} that simplices which create new homology classes are called \emph{creators} and simplices that destroy homology classes are called \emph{destroyers}.
-\begin{proposition}[Locality of moves~\cite{busaryev2010tracking}]
-	Other than the simplex $\sigma_i$ being moved, $\mathrm{MoveRight}(i,j)$ cannot change a destroyer to a creator, $\mathrm{MoveLeft}(i,j)$ cannot change a creator to destroyer, and neither move operation changes the status of any simplex in $K$ outside of the interval $[i,j]$.
-\end{proposition}
+%\begin{proposition}[Locality of moves~\cite{busaryev2010tracking}]
+%	Other than the simplex $\sigma_i$ being moved, $\mathrm{MoveRight}(i,j)$ cannot change a destroyer to a creator, $\mathrm{MoveLeft}(i,j)$ cannot change a creator to destroyer, and neither move operation changes the status of any simplex in $K$ outside of the interval $[i,j]$.
+%\end{proposition}
 \noindent The effect of the movement on intermediate simplices depends on the direction of the movement.
@@
 $$ \mathrm{Inv}(p, q) = \{ \, (i,j) \, | \, p(i) < p(j) \land q(i) > q(j)  \, \}$$
-The associated distance, measuring $\lvert S_F \rvert$, is the well-known \emph{Kendall}-$\tau$ distance $K_\tau(p,q)$: 
+The associated distance, measuring $\lvert S_F \rvert$, is the \emph{Kendall}-$\tau$ distance $K_\tau(p,q)$: 
 \begin{equation}\label{eq:kendall_dist}
@@
 \[
+\setlength\arraycolsep{2.3pt}
 \scalefont{1.0}{
@@
 	\caption{An example of a crocker plot (right) depicting the evolution of dimension $p = 1$ Betti curves over time.
-The green $\mathrm{X}$ marks correspond chronologically to the complexes (left), in row-wise order.
-The large orange and purple areas depict $1$-cycles persisting in space (the y-axis) and time (x-axis).
+The green $\mathrm{X}$ marks correspond chronologically to the complexes (left), in row-major order.
+The large orange and purple areas depict $1$-cycles persisting in both space (the y-axis) and time (x-axis).
 }\label{fig:crocker1}
@@
 For example, one of the often noted drawbacks of persistence is its instability with respect to strong outliers.
-Although persistence diagrams are stable with respect to their input and thus are robust to noise, the presence of strong outliers can artificially modify the connectivity of the underlying filtration, obscuring the detection of what wogiuld otherwise be significant topological structures~\cite{buchet2015topological}.
+Although persistence diagrams are stable with respect to their input and thus are robust to noise, the presence of strong outliers can artificially modify the connectivity of the underlying filtration, obscuring the detection of what would otherwise be significant topological structures~\cite{buchet2015topological}.
 Indeed, since inferring the homology of some underlying topological space around which data are distributed was one of the original purposes of persistence\footnote{One of the first uses of persistence was to infer the topology of attractors in dynamical systems from experimental data\cite{perea2018brief}.}, strong outliers pose a significant barrier to the widespread adoption of persistent homology.
@@
 \subsubsection*{The Fibered Barcode}
-An \emph{invariant} of $n$-parameter persistence modules is a function from the collection of persistence modules to some set $S$ with the property that $f(M) = f(M')$ when $M \cong M'$.
+%In the context of persistence, and \emph{invariant} is a function from the collection of persistence modules to some set $S$ with the property that $f(M) = f(M')$ when $M \cong M'$.
 An invariant is said to be \emph{complete} if $M \cong M'$ whenever $f(M) = f(M')$.
@@
 Lesnick and Wright~\cite{lesnick2015interactive} associate three simple invariants to a 2-dimensional persistence module $M$: the \emph{dimension function}, the \emph{multigraded Betti numbers}, and the \emph{fibered barcode}.
-
 The dimension function is simply the the function which maps every point $a \in \mathbb{R}^2$ to $\mathrm{dim}(M_a)$.
@@
 From a computational perspective, the core novelty of the RIVET approach is its ability to compute the fibered barcodes efficiently from precomputed information.
-The three dominant subcomputations needed to compute the fibered barcode using the reparameratization discussed above are the following: 
+Given bipersistence module $M$, the three dominant computations needed to compute the fibered barcode using the reparameteratization discussed above are the following: 
 \begin{enumerate}
-	\item Computing the bigraded Betti numbers $\beta(M)$ of a given bipersistence module $M$ 
-	\item Constructing a line arrangement $\mathcal{A}(M)$ induced by critical lines derived from (1) 
+	\item Computing the bigraded Betti numbers $\beta(M)$ of $M$
+	\item Constructing a line arrangement $\mathcal{A}(M)$ induced by critical lines from (1) 
 	\item Augmenting $\mathcal{A}(M)$ with a \emph{barcode template} $\mathcal{T}^e$ at every 2-cell $e \in \mathcal{A}(M)$
@@
 
-Consider the (coarsened) fibered barcode computed from a standard Rips/codensity bifiltration on a representative sample of the image data from~\cite{hateren_schaaf_1998} (details to follow), shown in Figure~\ref{fig:patch_data_dgm}.
+Consider the (coarsened) fibered barcode computed from a standard Rips / codensity bifiltration on a representative sample of the image data from~\cite{hateren_schaaf_1998} (details to follow), shown in Figure~\ref{fig:patch_data_dgm}.
 Upon inspection of the bigraded Betti number and the dimension function, one finds that a large area of dimension function is constant (highlighted as the blue portion in the middle of Figure~\ref{fig:patch_data_dgm}), wherein the first Betti number is 5.
@@
 This can be achieved by expressing the patch data as an expansion of a certain convenient basis of 8 non-constant eigenvectors, often called the Discrete Cosine Transform (DCT) basis~\cite{lee2003nonlinear}, which are given below:
-\vspace{-0.2em}
-\begin{figure}[h]
-	\centering
-	\includegraphics[width=0.60\textwidth]{dct_basis}
-\end{figure}
-\vspace{-0.5em}
-
+%\vspace{-0.2em}
+\begin{center}
+	\includegraphics[width=0.80\textwidth]{dct_basis} 
+\end{center}
+%\begin{figure}[h]
+%	\captionsetup{labelformat=empty}
+%	\centering
+%	\includegraphics[width=0.90\textwidth]{dct_basis}
+%	\caption{}
+%\end{figure}
+%\vspace{-1.5em}
 \noindent Projecting the image data onto the first two basis vectors of leads to the projection shown in the top left of Figure~\ref{fig:patch_data_dgm}.
@@
  \hline
- Coarsening & \# 2-cells & \texttt{phCol} & Vineyards / $d_{K}$ & Moves / $d_{\text{LCS}}$ \\
+\small{Coarsening} & \# 2-cells & \texttt{phCol} & Vineyards / $d_{K}$ & Moves / $d_{\text{LCS}}$ \\
  \hline
@@
 Moving forward, our results suggest there are many aspects of computing persistence in dynamic settings yet to be explored.
-Do existing optimizations to the standard reduction algorithm have natural extensions to dynamic setting?
-Are there other edit distances that yield efficient algorithms to update $R = DV$ decompositions, beyond moves and transpositions?
-An ideal method would be one that allows arbitrary permutations to a previously computed decomposition that is provably more efficient than performing the reduction independently in all cases.
+For example, it's not immediately clear whether one could adopt, for example, the twist optimization used in the reduction algorithm to the dynamic setting.
+Another avenue of interest is the setting where one has knowledge a priori of the time-varying...
+particular spaces
 
@@
 
+\bibliographystyle{siamplain}
 \bibliography{move_schedules}
-\bibliographystyle{ieeetr}
 

### 2021-08-23 move_schedules.tex
More tweaks to SIAM format

@@
 % Sets running headers as well as PDF title and authors
-\headers{An Example Article}{D. Doe, P. T. Frank, and J. E. Smith}
+\headers{Fast Persistence Computations}{M. Piekenbrock and J. Perea}
 
@@
 \title{Move Schedules: Fast persistence computations in sparse dynamic settings
-	\thanks{Submitted to the editors DATE.
-		\funding{This work was partially supported by the National Science Foundation through grants CCF-2006661 and CAREER award DMS-1943758.}
-	}
+	\thanks{\textbf{Funding:} This work was partially supported by the National Science Foundation through grants CCF-2006661 and CAREER award DMS-1943758.}
 }
@@
 \hypersetup{
-  pdftitle={An Example Article},
-  pdfauthor={Matt Piekenbrock, P. T. Frank, and J. E. Smith}
+  pdftitle={Fast Persistence Computations},
+  pdfauthor={Matt Piekenbrock and Jose A. Perea}
 }
@@
 \end{displaymath}
-
-The column $c$ in $R$ indicates that dimension $1$ homology is born, which in this filtration is never killed since the dimension of the complex is $2$.
+\noindent The column $c$ in $R$ indicates that dimension $1$ homology is born, which in this filtration is never killed since the dimension of the complex is $2$.
 Similarly, the columns at $u, v, w$ in $R_0$ (not shown) are all zero, indicating three $0$-dimensional homology classes  are born, which are then killed by the pivot entries in columns $a$ and $b$ in $R_1$.

### 2021-08-25 move_schedules.tex
Reorganized and editted section 3.3

@@
 \subsection{Vineyards}\label{sec:vineyards}
-The original purpose of the vineyards algorithm, as described in~\cite{cohen2006vines}, was to compute a continuous 1-parameter family of persistence diagrams given a time-varying filtration, and to detect homological critical events during the construction.
+The original purpose of the vineyards algorithm, as described in~\cite{cohen2006vines}, was to compute a continuous 1-parameter family of persistence diagrams over a time-varying filtration, and to detect homological critical events during the construction.
 Notice homological critical events can only occur when the filtration order changes, though not all changes in the filtration order result in homology changes.
@@
 However, Boissonnat~\cite{boissonnat2018efficient} has shown that for any $d$-dimensional simplicial complex $K$ with $m$ simplices, the number of distinct filtrations built from $K$ is \emph{at least} $\floor{\frac{t+1}{d+1}}^m$, where $t$ is the number of distinct filtration values associated with the complex $K$.
-Since this lower bound grows similarly to $m!$ when $t \sim O(m)$ and $d << m$ fixed, we expect $O(m - \sqrt{m})$ is not too pessimistic of an upper bound when the pair of filtrations are truly random.
+In the case where $t \sim O(m)$ and $d << m$ fixed, this lower bound grows similarly to $m!$---thus, $O(m - \sqrt{m})$ is not too pessimistic of an upper bound when the pair of filtrations are truly random.
 In practice, when one has a time-varying filtration and the sampling points are relatively close [in time], the LCS between adjacent filtrations is expected to be much larger, subsequently making $d$ much smaller.
@@
 However, like in the vineyards algorithm, certain pairs of simplices cost more to exchange depending on whether they are critical pairs in the sense described~\cite{cohen2006vines}, making specific move operations more expensive than others.
-%This variation in cost exaggerates the variability of the cost of scheduling move operations.
-
+This small variation in cost-per-transposition induces a large variability in the cost of executing any given schedule, which is undesirable computationally.
+Ideally, we would like to generate a schedule which not only is minimal in size, but also cost.
+Formally, given two permutations $p, q \in S_m$, let $S = ( s_1, s_2, \dots, s_d )$ denote a sorting such that $S \circ p = q$.
+%The example given in section~\ref{sec:greedy} shows all (minimal size) $d!$ possible choices of $S$ for a fixed pair of permutations and choice of $\mathrm{LCS}(p,q)$. 
+Let $\nu = \mathrm{LCS}(p,q)$ and let $\sigma \in S_d$ denote a choice of ordering for the symbols in  $p \smallsetminus \nu$, such that $S(\sigma)$ denotes a schedule generated from this ordering.
+Rephrasing from the description above, we would like to achieve something similar to the following objective:
+\begin{equation}\label{eq:schedule_opt}
+	S_\sigma^\ast = \argmin_{\sigma \in S_d} \mathrm{cost}(S(\sigma) \circ p)
+\end{equation}
+where $\mathrm{cost}(S(\sigma) \circ p)$ measures the number of column operations require to execute $S(\sigma)$ on an existing decomposition.
+There are exponentially many ways to generate such a sorting $S$.
+% Dynamic programming discussion 
 One advantage of Algorithm~\ref{alg:mr} is that the cost of applying a permutation to some given $R = D V$ decomposition can be determined efficiently prior to performing any column operations. 
 Specifically, the cost of determining $\lvert \mathbb{I}\rvert + \lvert \mathbb{J}\rvert$ in lines (2) and (3) of Algorithm~\ref{alg:mr} cannot take more than $O(m)$ time.
-If we assume a similar complexity to determine the cost of applying a permutation with Algorithm~\ref{alg:ml} and we know to execute $d$ moves, an initial thought might be to use an optimization method like dynamic programming.
+If we assume a similar complexity to determine the cost of applying a permutation with Algorithm~\ref{alg:ml} and we know to execute $d$ moves, one might expect be to use an optimization method like dynamic programming.
 That is, suppose $O(dm)$ move costs are computed ahead of time and these costs are stored in a table providing $O(1)$ element access and modification.
@@
 
-As before, given two permutations $p, q \in S_m$, let $S = ( s_1, s_2, \dots, s_d )$ denote a sorting such that $S \circ p = q$. 
-In general, there are exponentially many ways to generate such a sorting $S$. 
-The example given in section~\ref{sec:greedy} shows all (minimal size) $d!$ possible choices of $S$ for a fixed pair of permutations and choice of $\mathrm{LCS}(p,q)$. 
-Let $\tau = \mathrm{LCS}(p,q)$ and let $\sigma \in S_d$ denote a choice of ordering   for the symbols in  $p \smallsetminus \tau$, 
-such that $S(\sigma)$ denotes a schedule generated from this ordering. Ideally, we would like to achieve something similar to the following objective: 
-\begin{equation}\label{eq:schedule_opt}
-	S_\sigma^\ast = \argmin_{\sigma \in S_d} \mathrm{cost}(S(\sigma) \circ p)
-\end{equation}
-where $\mathrm{cost}(S(\sigma) \circ p)$ measures the number of column operations require to execute $S(\sigma)$ on an existing decomposition. 
 The schedule $S_1$ shown in Table~\ref{table:move_costs} would be an example of a minimal cost schedule.
@@
 \end{equation}
-When the set of permutations $S$ are adjacent transpositions, equality is achieved above since $K_\tau(\hat{s}_i, \hat{s}_{i+1}) = 1$ for every inversion between $p$ and $q$.
-However, this is not guaranteed if $S$ was derived using the strategy from section~\ref{sec:move_schedules}.
-Obtaining $S_\sigma^\ast$ via direct optimization over $S_d$ isn't practical, but experimental tests suggest that a greedy-type procedure performs well as a heuristic, which yields a $O(d^2 m \log m)$ minimization procedure.
-When $d <<  m$, this strategy of successively minimizing the Kendall-$\tau$ distance is efficient.
+When the set of permutations $S$ are adjacent transpositions, equality is achieved above since $K_\tau(\hat{s}_i, \hat{s}_{i+1}) = 1$ for every inversion between $p$ and $q$, though this not guaranteed if $S$ was derived using the strategy from section~\ref{sec:move_schedules}.
+Obtaining $S_\sigma^\ast$ via direct optimization over $S_d$ isn't practical, but experimental tests suggest that a greedy-type procedure performs well as a heuristic, which yields a $O(d^2 m \log m)$ minimization procedure, which is efficient when $d <<  m$.
 %where the set and order of pairs $\hat{s}_\ast$ depend on $S(\sigma)$. 
@@
 \end{equation}
-Although very similarity to the Kendall-$\tau$ distance, the Spearman distance offers some convenient computational advantages: it can be computed in $O(m)$ time, and the variance of its asymptotic distribution is larger.
-Consider the following minimization:
+Although very similar to the Kendall-$\tau$ distance, the Spearman distance offers some convenient computational advantages: it can be computed in $O(m)$ time, the variance of its asymptotic distribution is larger than Kendalls for uniformly random permutations, and it satisfies a certain monotonicity property under LCS edit operations.
+Specifically, given two permutations $p,q \in S_m$ with $\nu = \mathrm{LCS}(p,q)$ between them, the analysis in section~\ref{sec:move_schedules} shows we must move each symbol $x \in \mathcal{D}$ where $\mathcal{D} = p \smallsetminus \nu $ in a way that extends $\nu$ after each move.
+Each symbol $x \in \mathcal{D}$ induces a permutation $s_x$ such that $F(p, q) \geq F(s_x \circ p, q)$.
+That is, unlike the Kendall distance, the Spearman distance satisfies:
+$$ F(p, q) \geq F(s_1 \circ p, q) \geq F( s_2 \circ s_1 \circ p, q) \geq F( s_d \circ \dots \circ s_2 \circ s_1 \circ p, q)  = F(q,q) = 0 $$
+For any sorting $S$ such that $s_{i}$ that increases the size of longest common subsequence between $s_{i-1} \circ \dots \circ s_1 \circ p$ and $q$.
+Since the complexity of the move operations discussed in section~\ref{sec:moves} depend on the size of the movement, it is advantageous to construct a schedule which minimizes this movement.
+%Since $\lvert \mathcal{D} \rvert$ is fixed, we would like to choose $x \in \mathcal{D}$ such that the corresponding permutation $s_x$ reduces the total displacement of every symbol $x \in \mathcal{D}$ relative to  $q$.
+We consider the following objective modified from equation~\ref{eq:schedule_opt}: 
 \begin{equation}\label{eq:schedule_opt_succ}
@@
 \end{equation}
-The minimizer of this optimization can be interpreted as a sorting which minimizes the net displacement amongst all possible sortings, again in the $\ell_1$ sense. 
+The minimizer of this optimization can be interpreted as a sorting which minimizes the net displacement amongst all possible sortings, in the $\ell_1$ sense.
+This minimization problem is related to a constrained version of the bipartite crossing minimization problem for $k$ sets of permutations, which is NP-hard for $k \geq 4$~\cite{biedl2009complexity}.
+As a result, we instead consider a greedy approach.
+A greedy strategy requires $O((d+(d-1)+(d-2)+\dots+1)m) \approx O(d^2 m)$ time to complete, since there are $O(d)$ symbols to choose from and minimizing $F(\hat{s}_i, \hat{s}_{i+1})$ at each step takes $O(md)$.
+Since $\mathcal{D}$ is fixed, a greedy strategy thus depends on our ability to quickly compute $F(\hat{s}_i, \hat{s}_{i+1})$.
 %\\
@@
 %% TODO: elaborate on monotonicity, describe segment tree
-Given two permutations $p,q \in S_m$ and a LCS $\tau = \mathrm{LCS}(p,q)$ between them, the analysis in section~\ref{sec:move_schedules} shows we must move each symbol $x \in \mathcal{D}$ where $\mathcal{D} = p \smallsetminus \tau$ in a way that extends $\tau$ after each move.
-Thus we have $d = \lvert \mathcal{D} \rvert$ symbols to choose from initially.
-Each symbol $x \in \mathcal{D}$ yields a permutation $s$ to apply to $p$.
-Unlike the Kendall distance, the Spearman distance satisfies:
-$$ F(p, q) \geq F(s_1 \circ p, q) \geq F( s_2 \circ s_1 \circ p, q) \geq F( s_d \circ \dots \circ s_2 \circ s_1 \circ p, q)  = F(q,q) = 0 $$
-For any sorting $S$ such that $s_{i}$ that increases the size of longest common subsequence between $s_{i-1} \circ \dots \circ s_1 \circ p$ and $q$.
-Since $\lvert \mathcal{D} \rvert$ is fixed, we would like to choose $x \in \mathcal{D}$ such that the corresponding permutation $s_x$ reduces the total displacement of every symbol $x \in \mathcal{D}$ relative to  $q$.
-A greedy strategy chooses $x$ such that $F(\hat{s}_i, \hat{s}_{i+1})$ is as small as possible, removes that symbol from $\mathcal{D}$, and recurses.
-Thus, a greedy strategy depends on our ability to quickly compute $F(\hat{s}_i, \hat{s}_{i+1})$.
 \\
@@
 The greedy heuristic proposed to minimize equation~\ref{eq:schedule_opt_succ} was used to produce $S_1$, which has the minimal number of crossings amongst all possible schedules in this example, since $K_\tau(p,q) = 21$.
-This minimization problem is related to a constrained version of the bipartite crossing minimization problem for $k$ sets of permutations, which is NP-hard for $k \geq 4$~\cite{biedl2009complexity}.
 \\
@@
 The monotonic nature of successively applying the Spearman distance enables certain computational advantages.
-Suppose we begin with an array $\mathcal{A}$ of size $m$ which provides $O(1)$ access and modification, initialized to $0$.
-Summing the entries in this array yields the Spearman distance in $O(m)$ time, which at initialization represents $F(p, p)$.
-Each symbol $x \in \mathcal{D}$ induces a  permutation $s_x$ such that $F(p, q) \geq F(s_x \circ p, q)$.
-Since there are $d = \lvert \mathcal{D} \rvert$ symbols to move and computing $F$ takes $O(m)$ time, a na\"ive greedy strategy requires $O((d+(d-1)+(d-2)+\dots+1)m) \approx O(d^2 m)$ time to complete.
-However, observe that each permutation $s_x$ changes the displacement of every symbol in a very predictable way: if $s_x$ represents a move permutation that moves $x$ from position $i$ to position $j$, then relative to the previous permutation $s_{i}$, the displacement of $x$ in $s_{i+1}$ changes by at most  $\lvert i - j \rvert$ and the displacement of any symbol between $i$ and $j$ changes by $\pm 1$.
-All other symbols are unaffected.
-Since the Spearman distance is simply the sum of these displacements, replacing $\mathcal{A}$ with a structure that supports both $O(\log m)$ access time to aggregate information (such as the sum) and $O(\log m)$ modification time ability to up to $\lvert i - j \vert$ entries.
-The former problem reduces to the problem of efficiently calculating and updating \emph{prefix sums}, however since $\lvert i - j \rvert \leq m$ is potentially larger than $O(\log m)$, it's not immediately clear how to achieve the latter modification complexity in $O(\log m)$ time.
-Fortunately, since every elements displacement other than $x$ changes by $\pm 1$ per update, both queries can be efficiently done in $O(\log m)$ time using a Segment Tree that supports \emph{lazy propagation}~\cite{de1997computational}.
+Suppose we begin with an array $\mathcal{A}$ of size $m$ which provides $O(1)$ access and modification, initialized with the (signed) \emph{displacement} of every element in $p$ to its corresponding position in $q$.
+Since the Spearman distance is simply the sum of these displacements, if we could replace $\mathcal{A}$ with a data structure that supports $O(\log m)$ access time to aggregate information (i.e.
+the sum) and $O(\log m)$ modification time ability on $\lvert i - j \vert$ entries (simultaneously), we would have an efficient $O(d \log m)$ greedy solution.
+The former problem reduces to the problem of efficiently calculating \emph{prefix sums}, which is easily solved.
+It's not immediately clear how to achieve the latter modification complexity in $O(\log m)$ time, since $\lvert i - j \rvert \leq m$ is potentially larger than $O(\log m)$.
+However, observe that each permutation $s_x$ changes the displacement of every symbol in a very predictable way: if $s_x$ represents a move permutation that moves $x$ from position $i$ to position $j$, then relative to the previous permutation $s_{i}$, the displacement of $x$ in $s_{i+1}$ changes by $\lvert i - j \rvert$ and the displacement of any symbol between $i$ and $j$ changes by $\pm 1$.
+All other symbol displacements are unchanged.
+%, and 
+It is known that one can apply constant-factor updates to multiple values in a prefix sum in $O(\log m)$ time by any prefix-sum data structure that also supports \emph{range updates}, e.g.
+a Fenwick tree or a Segment Tree that supports lazy propagation~\cite{de1997computational}.
+Single element prefix sum modifications can also be achieved $O(\log m)$ time with these data structures, the time to compute $F(\hat{s}_i, \hat{s}_{i+1})$ is bounded above by $O(\log m)$ and thus the greedy solution to equation~\ref{eq:schedule_opt_succ} can be computed in $O(d \log m)$ time.
 %We give an example and a more detailed discussion for those interested in the appendix\ref{sec:appendix}. 
-
 %This cumulative Spearman distance enables certain computational advantages.

### 2021-08-25 move_schedules.tex
Again updated 3.3, is looking much better now!

@@
 where $\mathrm{cost}(S(\sigma) \circ p)$ measures the number of column operations require to execute $S(\sigma)$ on an existing decomposition. 
-There are exponentially many ways to generate such a sorting $S$.
+%There are exponentially many ways to generate such a sorting $S$. 
 % Dynamic programming discussion 
@@
 However, moving a simplex from position $i$ to position $j$ may require updating $O(\kappa^2)$ entries in the structure, where $\kappa = \lvert i - j \rvert$.
-Since $\kappa \sim O(m)$ in the worst case, direct optimization may require upwards of $\approx O(dm^2)$.
-Since this time complexity is potentially more expensive than executing a schedule to begin with, this direct strategy does not seem viable for practical settings.
+Since $\kappa \sim O(m)$ in the worst case, direct optimization may require upwards of $\approx O(dm^2)$, which is potentially more expensive than executing a schedule to begin with and thus not viable in practical settings.
 
@@
 \subsubsection{Proxy objective}\label{sec:proxy_objective}
-In light of the direct optimization being too expensive and a greedy procedure leading to suboptimal behavior, we consider whether there exists an alternative, or \emph{proxy objective} to optimize that is related to the cost of a schedule but not dependent on the entries in the decomposition.
-In this section, we present such an objective, showing its relation to the problem at hand as well its use in other known computational problems.
-We show empirical results demonstrating its practical effectiveness as a heuristic in section~\ref{sec:results}.
-
-The schedule $S_1$ shown in Table~\ref{table:move_costs} would be an example of a minimal cost schedule.
+When $d \sim O(m)$, the quadratic nature of the aforementioned greedy minimization is too expensive.
+In order for Algorithm~\ref{alg:schedule} to be considered as an efficient alternative to vineyards---which requires $O(m \log m + k)$ preprocessing time to compute $k$ transpositions---a solution which is subquadratic in $m$ is desirable.
+%In what follows, we present such an objective, showing its relation to the problem at hand as well its use in other known computational problems.
+% We show empirical results demonstrating its practical effectiveness as a heuristic in section~\ref{sec:results}.
+
+%The schedule $S_1$ shown in Table~\ref{table:move_costs} would be an example of a minimal cost schedule.
 Since optimizing this objective combinatorially is inefficient, we would like an alternative objective that is cheaper and performs well in practice.
-Intuitively, one alternative objective might be the Kendall-$\tau$ distance $K_\tau$, 
-which can be computed in $O(m \log m)$ time---indeed, this is the weighting measure minimized in a certain path optimization in the RIVET software~\cite{lesnick2015interactive} to ensure adjacent filtrations are similar.
+%Intuitively, one alternative objective might be the Kendall-$\tau$ distance $K_\tau$, 
+%which can be computed in $O(m \log m)$ time---indeed, this is the weighting measure minimized in a certain path optimization in the RIVET software~\cite{lesnick2015interactive} to ensure adjacent filtrations are similar.
 Motivated by this, define $\hat{s}_i =  s_i \circ \cdots \circ s_2 \circ s_1 \circ p$ to be the result of applying the composition of the first $i$ permutations for some given schedule $S = (s_1, \dots, s_d)$, where $i < d$.
-A possible inequality to utilize is the following: 
-\begin{equation}\label{eq:kendall_inequality}
-K_\tau(p, q) = K_{\tau}(p, S \circ p) \leq \sum\limits_{i=1}^{d-1} K_{\tau}(\hat{s}_i, \hat{s}_{i+1})
-\end{equation}
-When the set of permutations $S$ are adjacent transpositions, equality is achieved above since $K_\tau(\hat{s}_i, \hat{s}_{i+1}) = 1$ for every inversion between $p$ and $q$, though this not guaranteed if $S$ was derived using the strategy from section~\ref{sec:move_schedules}.
+%A possible inequality to utilize is the following: 
+%\begin{equation}\label{eq:kendall_inequality}
+%K_\tau(p, q) = K_{\tau}(p, S \circ p) \leq \sum\limits_{i=1}^{d-1} K_{\tau}(\hat{s}_i, \hat{s}_{i+1})
+%\end{equation}
+%When the set of permutations $S$ are adjacent transpositions, equality is achieved above since $K_\tau(\hat{s}_i, \hat{s}_{i+1}) = 1$ for every inversion between $p$ and $q$, though this not guaranteed if $S$ was derived using the strategy from section~\ref{sec:move_schedules}.
 Obtaining $S_\sigma^\ast$ via direct optimization over $S_d$ isn't practical, but experimental tests suggest that a greedy-type procedure performs well as a heuristic, which yields a $O(d^2 m \log m)$ minimization procedure, which is efficient when $d <<  m$.
@@
 %Let $p,q \in S^n$ be fixed permutations of $[n]$ and let $S = (s_1, \dots, s_d)$ denote a sorting $p \mapsto q$ such that $q = S \circ p = s_d \circ \cdots \circ s_1 \circ p$, and 
-\\
-\\
-However, when $d \sim O(m)$, the quadratic nature of this minimization is too expensive.
-It turns out there is a similar distance that is both cheaper to compute and also amenable to a greedy type optimization.
-The \emph{Spearman distance} $F$ is a common distance for measuring the disarrangement between permutations.
+In light of the discussion above, we consider whether there exists an alternative \emph{proxy objective} to optimize that is related to the cost of a schedule but not dependent on the entries in the decomposition.
+Consider again the objective:
+\begin{equation}\label{eq:schedule_opt2}
+	S_\sigma^\ast = \argmin_{\sigma \in S_d} \mathrm{cost}(S(\sigma) \circ p)
+\end{equation}
+where now $\mathrm{cost}(\cdot)$ is to be replaced by a proxy objective function.
+There are exponentially many ways to choose $\sigma \in S_d$, so we opt for a greedy strategy as before, which takes $O(d \, C(m))$ time where $C(m)$ denotes the cost of evaluating $\mathrm{cost}(\cdot)$.
+For this minimization to be subquadratic in $m$, we need $C(m) \sim O(\log m)$ at each of the $d$ steps of the greedy optimization, and for the minimization to yield an efficient solution in practice, $\mathrm{cost}(\cdot)$ must be related to the problem of minimizing move costs.
+
+A natural upper bound on the number of column operations needed to move a simplex from position $i$ in the filtration to position $j$ is the displacement between these positions, $\lvert i - j \rvert$.
+In practice, due to the sparsity of both $R$ and $V$, this is a very loose upper bound---the number of column operations is generally much smaller.
+Nonetheless, the complexities of the move operations discussed in Section~\ref{sec:moves} are dependent on the magnitude of $\lvert i - j \rvert$, and thus it's worth considering a proxy measure.
+If one forms bijections between two given filtrations $p: K_0 \to [m], q: K_1 \to [m]$ to permutations of $[m]$ as before, then an upper bound on the number of column reductions needed to perform any given schedule is given the \emph{Spearman distance}, a common distance for measuring the disarrangement between permutations.
 It is defined as:
@@
 For any sorting $S$ such that $s_{i}$ that increases the size of longest common subsequence between $s_{i-1} \circ \dots \circ s_1 \circ p$ and $q$.
-Since the complexity of the move operations discussed in section~\ref{sec:moves} depend on the size of the movement, it is advantageous to construct a schedule which minimizes this movement.
+Given $p, q$ and some schedule $S = (s_1, \dots, s_d)$, let $\hat{s}_i =  s_i \circ \cdots \circ s_2 \circ s_1 \circ p$ denote the composition of the first $i$ permutations in $S$, where $i < d$.
+An upper bound on the Spearman distance can be decomposed into a sum of $d-1$ local distances with the following inequality:
+%A possible inequality to utilize is the following: 
+\begin{equation}\label{eq:kendall_inequality}
+F(p, q) = F(p, S(\sigma) \circ p) \leq \sum\limits_{i=1}^{d-1} F_\sigma(\hat{s}_i, \hat{s}_{i+1})
+\end{equation}
+where equality is only achieved with schedules $S$ for which each symbols displacement between its initial position in $p$ and its final position in $q$ is non-increasing with every application of $s_i$.
+Note this not guaranteed if $S$ was derived using the strategy from section~\ref{sec:move_schedules}.
+%Since the complexity of the move operations discussed in section~\ref{sec:moves} depend on the size of the movement, it is advantageous to construct a schedule which minimizes this movement. 
 %Since $\lvert \mathcal{D} \rvert$ is fixed, we would like to choose $x \in \mathcal{D}$ such that the corresponding permutation $s_x$ reduces the total displacement of every symbol $x \in \mathcal{D}$ relative to  $q$.
-We consider the following objective modified from equation~\ref{eq:schedule_opt}: 
+Substituting objective~\ref{eq:kendall_inequality} for $\mathrm{cost}(\cdot)$ in equation~\ref{eq:schedule_opt2} yields the following global optimization problem over $S_d$: 
 \begin{equation}\label{eq:schedule_opt_succ}
@@
 \end{equation}
-The minimizer of this optimization can be interpreted as a sorting which minimizes the net displacement amongst all possible sortings, in the $\ell_1$ sense.
-This minimization problem is related to a constrained version of the bipartite crossing minimization problem for $k$ sets of permutations, which is NP-hard for $k \geq 4$~\cite{biedl2009complexity}.
-As a result, we instead consider a greedy approach.
-A greedy strategy requires $O((d+(d-1)+(d-2)+\dots+1)m) \approx O(d^2 m)$ time to complete, since there are $O(d)$ symbols to choose from and minimizing $F(\hat{s}_i, \hat{s}_{i+1})$ at each step takes $O(md)$.
-Since $\mathcal{D}$ is fixed, a greedy strategy thus depends on our ability to quickly compute $F(\hat{s}_i, \hat{s}_{i+1})$.
+The minimizer of this optimization can be interpreted as a sorting which minimizes the net displacement amongst all possible sortings $p \mapsto q$, in the $\ell_1$ sense.
+It is related to a constrained version of the crossing minimization problem for $k$ sets of permutations, which is NP-hard for $k \geq 4$~\cite{biedl2009complexity}, see below for an example.
+As a result, we do not expect a greedy solution to yield the optimal solution exactly, but simply a fast approximation.
+%A greedy strategy requires $O((d+(d-1)+(d-2)+\dots+1)m) \approx O(d^2 m)$ time to complete, since there are $O(d)$ symbols to choose from and minimizing $F(\hat{s}_i, \hat{s}_{i+1})$ at each step takes $O(md)$.
+Since $\mathcal{D}$ is fixed, the complexity of a greedy strategy depends entirely on the complexity of minimizing $F(\hat{s}_i, \hat{s}_{i+1})$ at each greedy step.  
 %\\

### 2021-09-03 move_schedules.tex
Tweaks to fibered barcode description

@@
 
+% Journal name: SICOMP
 %\usepackage[margin=0.75in]{geometry}
@@
 %We refer to the algorithm that maintains the $R = \partial V$ decomposition under adjacent permutations $P$ as the \emph{transposition framework}. 
-The explicit pseudocode describing these rules is given in the appendix by Algorithm~\ref{alg:tr}. 
+% The explicit pseudocode describing these rules is given in the appendix by Algorithm~\ref{alg:tr}. 
 %	\; = 
@@
 \noindent 
-\textbf{Example 2.2} To illustrate the basic principles on which vineyards works, we re-use the running example introduced in the previous section.
-Below, we illustrate the case of exchanging simplices $a$ and $b$ in the filtration order.
+\textbf{Example 2.2} Below, we illustrate using vineyards to exchange simplices $a$ and $b$ in the filtration order using the running example introduced in the previous section. 
 \begin{displaymath}
@@
 		\begin{block}{c[ccc]}
-        a \kern\rowsp  & \; 1 & 1 & 1 \; \topstrut \\
+        a \kern\rowsp  & \; 1 & \textbf{1} & 1 \; \topstrut \\
   		b \kern\rowsp & \;  & 1 & 1 \; \\
@@
   		 & \;  & 1 &  \; \\
-  		 & \; 1 & 1 & \; \botstrut \\
+  		 & \; 1 & \textbf{1} & \; \botstrut \\
 		\end{block}
@@
   		& \; 1 &  &  \; \\
-  		& \; 1 & 1 &  \; \botstrut \\
+  		& \; 1 & \textbf{1} &  \; \botstrut \\
 		\end{block}
@@
 \end{displaymath}
-Prior to performing the exchange, observe that that the highlighted entry in $V_1$ would render $V_1$ non-upper triangular after the exchange.
+Prior to performing the exchange, observe that that the bold entry in $V_1$ would render $V_1$ non-upper triangular after the exchange.
 This entry is removed by a left-to-right column operation, given by applying the $S_1^{2}$ on the right to $R_1$ and $V_1$.
@@
 In fact, achieving $O(m)$ complexity per transposition requires a special sparse matrix representation that allows swapping any two rows and columns in $O(1)$---see the appendix for more details and discussion.
-Finally, note the constant factor associated with the stated complexity: inspection of the individual cases of Algorithm~\ref{alg:tr} shows that there are at most two $O(m)$ operations, on both $R$ and $V$, that may be needed for any single transposition.
+Finally, note the constant factor associated with the stated complexity: inspection of the individual cases of the algorihtm from~\cite{cohen2006vines} shows that there are at most two $O(m)$ operations, on both $R$ and $V$, that may be needed for any single transposition.
 
@@
  \begin{enumerate}[leftmargin=2\parindent, align=left, labelsep=-5pt, topsep=0pt,itemsep=-0.25ex,parsep=1.2ex]
- 	\item Querying the non-zero status of entries in any matrix representation occurs once per move.
- 	\item The decomposition $R = D V$ is not maintained during the movement of $\sigma_i \mapsto \sigma_j$.
+ 	\item Checking whether an entry is non-zero occurs once per move.
+ 	\item $R = D V$ is not maintained during the movement of $\sigma_i \mapsto \sigma_j$.
  \end{enumerate} 
@@
 Since moving a simplex from $i$ to $j$ using vineyards performs $\lvert i - j \rvert - 1$ transpositions, this query operation is required at most $O(\lvert i - j \rvert)$ times, which is excessive.
-If a row-oriented sparse matrix implementation is used, a move operation can perform these non-zero entry checks in just one $O(m)$ pass, prior to performing any column operations.
+If a row-oriented sparse matrix implementation is used, a move operation can perform these non-zero checks in just one $O(\lvert i - j \rvert)$ pass, prior to performing any column operations.
 
@@
 This is a significant difference compared to the vineyards algorithm, which was designed specifically to maintain a valid decomposition after each transposition. 
-Thus, move operations can be thought of as making a tradeoff in granularity: whereas a sequence of adjacent transpositions $(i, i{+}1), (i{+}1, i{+}2), \dots, (j{-}1, j)$ generates $\lvert i - j \rvert $ valid decompositions in vineyards, a move operation $\mathrm{Move}(i,j)$ generates only one.
-
+In this way, we interpret move operations as making a tradeoff in granularity: whereas a sequence of adjacent transpositions $(i, i{+}1), (i{+}1, i{+}2), \dots, (j{-}1, j)$ generates $\lvert i - j \rvert $ valid decompositions in vineyards, a move operation $\mathrm{Move}(i,j)$ generates only one.
+Indeed, if one has a pair of filtrations $(K_0, K_1)$ each with $m$ simplices, and we assume that each simplex $\sigma_i$ switches its relative ordering with another simplex $\sigma_j$ at most once in some ordered schedule of transpositions, then the number of intermediate persistence diagrams is bounded above by $O(m^2)$ using transpositions, whereas with move operations this bound clearly is at most $O(m)$.
 \begin{algorithm}
@@
 \end{algorithm}
-
-If one has a pair of filtrations $(K_0, K_1)$ each with $m$ simplices, and we assume that each simplex $\sigma_i$ switches its relative ordering with another simplex $\sigma_j$ at most once in some ordered schedule of transpositions, then the number of intermediate persistence diagrams is bounded above by $O(m^2)$ using transpositions, whereas with move operations this bound clearly is at most $O(m)$.
-\\
-\\
-\noindent \textbf{Example:} We continue the example used in sections~\ref{sec:vineyards} and~\ref{sec:reduction} to illustrate moves.
+\noindent \textbf{Example:} We continue the example used in sections~\ref{sec:reduction} and~\ref{sec:vineyards} to illustrate moves.
 Consider moving the edge $a$ to the position taken by the edge $c$ in the filtration.
 The donor columns introduced in~\cite{busaryev2010tracking} are shown on the left side of each matrix.
-Note that using vineyards, the equivalent permutation to the decomposition would require $4$ column operations on both $R_1$ and $V_1$, respectively, whereas a single move operation accomplishes this permutation using only $2$ column operations per matrix.
+A single move operation is able to apply this permutation using only $2$ column operations per matrix, assuming donor columns may be assigned in $O(1)$\footnote{The assumption that donor columns can be assigned in $O(1)$ is easily achieved under the wordRAM model of computation using any representation that allows swapping word-sized pointers each pointing to columns of field entries, e.g.
+the sparse matrix representation from~\cite{cohen2006vines}.}.
+Note that using vineyards, the equivalent permutation to the decomposition would require $4$ column operations on both $R_1$ and $V_1$, respectively.
 \begin{displaymath}
@@
 	S_F &= \{ (i,j) \, | \, f_0(\sigma_i) < f_0(\sigma_j) \land  f_1(\sigma_i) > f_1(\sigma_j) \quad \forall \; i,j \in [m] \} \\
-	&= ( s_1, s_2, \dots, s_k )
+	\nonumber &= ( s_1, s_2, \dots, s_k )
 \end{align}
@@
 This family includes e.g.
-the straight-line homotopy $f_\tau(\sigma) = (1 - \tau) f_0(\sigma) + \tau f_1(\sigma)$, which was studied in the original vineyards paper~\cite{cohen2006vines}.
+straight-line homotopies $f_\tau(\sigma) = (1 - \tau) f_0(\sigma) + \tau f_1(\sigma)$, as used in the original vineyards implementation~\cite{cohen2006vines}.
 Detecting all $k$ intersections of $m$ pseudo-segments is a well-studied problem in computational geometry that can be optimally solved in output-sensitive $O(m \log m + k)$ time by several algorithms~\cite{boissonnat2000efficient}, where $k$ is the output-sensitive term.
@@
 A sequence $s_1, s_2, \dots, s_d \in S \subseteq S_n$ of operations mapping $p \mapsto q$ is sometimes called a \emph{sorting of $p$} in the literature.
-In the context of performing updates to the $R = D V$ decomposition, we call such sequences \emph{schedules} .
+In the context of performing updates to the $R = D V$ decomposition, we call such ordered sequences \emph{schedules}.
 There are typically many ways to solve (1), and any solution to (2) necessarily solves (1) and (3).
@@
 \noindent The set of critical lines $\mathrm{crit}(M)$ with respect to some fixed set $S \subset \mathbb{R}^2$ fully characterizes a certain planar subdivision of the half plane $[0, \infty) \times \mathbb{R}$.
-Explicitly, define $\mathcal{A}^1(M)$ to the 1-skeleton of the line arrangement defined by:
+This planar subdivision, denoted by $\mathcal{A}(M)$, is thus entirely determined by $S$ under point line duality.
+%It has been shown~\cite{lesnick2015interactive} that the 1-skeleton $\mathcal{A}^1(M)$ of the line arrangement $\mathcal{A}(M)$ can be characterized b
+%Explicitly, define $\mathcal{A}^1(M)$ to the 1-skeleton of the line arrangement defined by: 
 % \{ \mathcal{D}_p(\alpha) \text{ for } \alpha = \mathrm{LUB}(u,v), u,v \in S \} 
-\begin{equation}\label{eq:arrangement_crit}
-	\mathcal{A}^1(M) = \{ \mathcal{D}_{\ell}(\mathrm{crit}(M)) \cup (\{0\} \times \mathbb{R}) \}
-\end{equation}
-Expanding the 1-skeleton to a 2-D cell complex $\mathcal{A}(M)$ fully partitions the upper half plane.  
-A consequence of equation~\ref{eq:arrangement_crit} that if the duals of two lines $L, L' \in \mathcal{L}$ are contained in the same $2$-cell in $\mathcal{A}(M)$, then $S_L = S_{L'}$, i.e.
+%\begin{equation}\label{eq:arrangement_crit}
+%	\mathcal{A}^1(M) = \{ \mathcal{D}_{\ell}(\mathrm{crit}(M)) \cup (\{0\} \times \mathbb{R}) \}
+%\end{equation}
+%Expanding the 1-skeleton to a 2-D cell complex $\mathcal{A}(M)$ fully partitions the upper half plane.  
+A corollary from~\cite{lesnick2015interactive} shows that if the duals of two lines $L, L' \in \mathcal{L}$ are contained in the same $2$-cell in $\mathcal{A}(M)$, then $S_L = S_{L'}$, i.e.
 the partitions induced by $\mathrm{push}_L$ are equivalent.
-Indeed, the total order on $S_L$ is the pullback of the total order on $L$ with respect to the push map.
+Indeed, the total order on $S_L$ is simply the pullback of the total order on $L$ with respect to the push map.
 Since $\mathcal{A}(M)$ partitions the entire half-plane, the dual to every line $L \in \mathcal{L}$ is contained within $\mathcal{A}(M)$---the desired reparameterization.
@@
 
-\section*{Appendix}\label{sec:appendix}
-\appendix
-
-\subsection*{Vineyard Algorithm}
-\begin{algorithm}[ht]
-	\caption{Vineyards: Tranposition Framework}\label{alg:tr}
-	\setstretch{1.15}
-	\begin{algorithmic}[1]
-		\Require Matrices $R, V$ satisfying $R = D V$, $1 \leq i \leq m - 1$
-		\Ensure Output $(R, V)$ maintains the decomposition invariants
-        \Function{Transpose}{$R$, $V$, $i$}
-            \State $\mathrm{pos} \gets $ columns satisfying $\mathrm{col}_R = 0$
-            \If {$\mathrm{pos}[i]$ \textbf{and} $\mathrm{pos}[i+1]$}
-	            \If{$V[i,i+1] \neq 0$}
-            		\State $\mathrm{col}_V(i+1)  \mathrel{+}= \mathrm{col}_V(i)$	
-            	\EndIf
-            	\If {$\exists \; k, l$ satisfying $\mathrm{low}_R(k) = i, \mathrm{low}_R(l) = i+1$ and $R[i,l] \neq 0$} \Comment{O(m)}
-            		\If {$k < l$}
-            			\State \Return $(\, R, \, V \,) \gets (\, PRPS_k^l, \, PVP S_k^l \,)$
-            		\Else
-            			\State \Return $(\, R, \, V \,) \gets  (\, PRPS_l^k, \, PVP S_l^k\,) $
-            		\EndIf
-            	\EndIf 
-            \ElsIf{$!\mathrm{pos}[i]$ \textbf{and} $!\mathrm{pos}[i+1]$}
-            	\If{$V[i,i+1] \neq 0$}\Comment{O(m)}
-            		\If{$\mathrm{low}_R(i) < \mathrm{low}_R(i+1)$}
-            			\State \Return  $(\, R, V \,) \gets (P R S^{i+1}_i P, \, P V S^{i+1}_i  P) $
-            		\Else 
-            			\State \Return  $(\, R, V \,) \gets (P R S^{i+1}_i P S^{i+1}_i, \, P V S^{i+1}_i P S^{i+1}_i ) $
-            		\EndIf 
-            	\EndIf
-           	\ElsIf{$!\mathrm{pos}[i]$ \textbf{and} $\mathrm{pos}[i+1]$} 
-            	\If{$V[i,i+1] \neq 0$}\Comment{O(m)}
-            		\State \Return  $(\, R, V \,) \gets (P R S^{i+1}_i P S^{i+1}_i, \, P V S^{i+1}_i P S^{i+1}_i ) $
-            	\EndIf
-            \ElsIf{$\mathrm{pos}[i]$ \textbf{and} $!\mathrm{pos}[i+1]$}
-            	\If{$V[i,i+1] \neq 0$}\Comment{O(m)}
-            		\State $\mathrm{col}_V(i+1) \mathrel{+}= \mathrm{col}_V(i)$
-            	\EndIf 
-            \EndIf
-            \State \Return $(\, R , V \,) \gets (\, P R P , P V P \,)$
-        \EndFunction
-    \end{algorithmic}
-\end{algorithm}
-
-\subsection*{Transposition complexity details}
-Consider Algorithm~\ref{alg:tr}.
+%
+%\appendix 
+%\section{}\label{sec:appendix}
+%\subsection*{Vineyard Algorithm}
+%\begin{algorithm}[ht]
+%	\caption{Vineyards: Tranposition Framework}\label{alg:tr}
+%	\setstretch{1.15}
+%	\begin{algorithmic}[1]
+%		\Require Matrices $R, V$ satisfying $R = D V$, $1 \leq i \leq m - 1$
+%		\Ensure Output $(R, V)$ maintains the decomposition invariants
+%        \Function{Transpose}{$R$, $V$, $i$}
+%            \State $\mathrm{pos} \gets $ columns satisfying $\mathrm{col}_R = 0$
+%            \If {$\mathrm{pos}[i]$ \textbf{and} $\mathrm{pos}[i+1]$}
+%	            \If{$V[i,i+1] \neq 0$}
+%            		\State $\mathrm{col}_V(i+1)  \mathrel{+}= \mathrm{col}_V(i)$	
+%            	\EndIf
+%            	\If {$\exists \; k, l$ satisfying $\mathrm{low}_R(k) = i, \mathrm{low}_R(l) = i+1$ and $R[i,l] \neq 0$} \Comment{O(m)}
+%            		\If {$k < l$}
+%            			\State \Return $(\, R, \, V \,) \gets (\, PRPS_k^l, \, PVP S_k^l \,)$
+%            		\Else
+%            			\State \Return $(\, R, \, V \,) \gets  (\, PRPS_l^k, \, PVP S_l^k\,) $
+%            		\EndIf
+%            	\EndIf 
+%            \ElsIf{$!\mathrm{pos}[i]$ \textbf{and} $!\mathrm{pos}[i+1]$}
+%            	\If{$V[i,i+1] \neq 0$}\Comment{O(m)}
+%            		\If{$\mathrm{low}_R(i) < \mathrm{low}_R(i+1)$}
+%            			\State \Return  $(\, R, V \,) \gets (P R S^{i+1}_i P, \, P V S^{i+1}_i  P) $
+%            		\Else 
+%            			\State \Return  $(\, R, V \,) \gets (P R S^{i+1}_i P S^{i+1}_i, \, P V S^{i+1}_i P S^{i+1}_i ) $
+%            		\EndIf 
+%            	\EndIf
+%           	\ElsIf{$!\mathrm{pos}[i]$ \textbf{and} $\mathrm{pos}[i+1]$} 
+%            	\If{$V[i,i+1] \neq 0$}\Comment{O(m)}
+%            		\State \Return  $(\, R, V \,) \gets (P R S^{i+1}_i P S^{i+1}_i, \, P V S^{i+1}_i P S^{i+1}_i ) $
+%            	\EndIf
+%            \ElsIf{$\mathrm{pos}[i]$ \textbf{and} $!\mathrm{pos}[i+1]$}
+%            	\If{$V[i,i+1] \neq 0$}\Comment{O(m)}
+%            		\State $\mathrm{col}_V(i+1) \mathrel{+}= \mathrm{col}_V(i)$
+%            	\EndIf 
+%            \EndIf
+%            \State \Return $(\, R , V \,) \gets (\, P R P , P V P \,)$
+%        \EndFunction
+%    \end{algorithmic}
+%\end{algorithm}
+%
+%
+%\subsection*{Transposition complexity details}
+%Consider Algorithm~\ref{alg:tr}.
 The matrices $R, V$ are both $(m \times m)$, where $m$ is the number of simplices in the underlying filtration.

### 2021-09-03 move_schedules.tex
Updates from Overleaf

--- 2021-08-25 move_schedules.tex
@@
 Its output is a particular decomposition of the total boundary matrix, from which the persistence diagrams and generating cycles can be derived.
-	Persistence diagrams are known to vary continuously with respect to their input; this motivates the algorithmic study of persistence computations for time-varying filtered complexes.
+	Persistence diagrams are known to vary continuously with respect to their input, which motivates the algorithmic study of persistence computations for time-varying filtered complexes.
 Computationally, simulating persistence dynamically can be reduced to maintaining a valid decomposition under adjacent transpositions in the filtration order.
@@
 	Finally, we provide empirical results suggesting that the decrease in operations needed to compute diagrams across a family of filtrations is proportional to the difference between the expected quadratic number of states, and the proposed sublinear coarsening.
+	Applications to multi-dimenisonal persistence and crocker stacks are also presented. 
 \end{abstract}
@@
 \subsection{Overview}\label{sec:overview} 
-The starting point to the work described in this paper is the stability results established in~\cite{cohen2007stability, cohen2006vines}. 
-Indeed, given a triangulable topological space equipped with a sufficiently tame continuous function, persistent homology captures the changes in topology across the sublevel sets of the space, and encodes them in a persistence diagram.
-The stability of persistence contends that small changes in the underlying function produce small changes in the corresponding diagrams:  if the function changes continuously, so too will the points on the persistence diagram.
+Given a triangulable topological space equipped with a sufficiently tame continuous function, persistent homology captures the changes in topology across the sublevel sets of the space, and encodes them in a persistence diagram.
+The stability of persistence contends that small changes in the underlying function produce small changes in the corresponding diagrams:  if the function changes continuously, so too will the points on the persistence diagram~\cite{cohen2007stability, cohen2006vines}.
 This motivates the study and application of persistence to time-varying or dynamic settings, like that of dynamic metric spaces~\cite{kim2020spatiotemporal}. 
@@
 If one is interested in understanding how the persistent homology of a continuous function changes over time, then this algorithm is sufficient, for  homological critical points can only occur when the filtration order changes.
-Indeed, this algorithm is also efficient  asymptotically: if there are $d$ time-points where this filtration order changes, then the vineyards algorithm takes $O(m^3 + md)$ time: one $O(m^3)$-time reduction upfront for the filtration at time $t_0$ followed by one $O(m)$ operation to update the decomposition for the remaining time points $(t_1, t_2, \dots, t_d)$.
-When $d >> m$, this $O(md)$ approach is far more efficient than the $O(dm^3)$  strategy of computing the diagrams independently at every time point, what we call the ``naive'' approach.
-
-Despite its efficient in theory, vineyards is often not the method of choice in practice in dynamic settings. 
-While there is an increasingly rich ecosystem of software packages offering variations of the standard reduction algorithm (e.g.
+This algorithm is also efficient  asymptotically: if there are $d$ time-points where this filtration order changes, then the vineyards algorithm takes $O(m^3 + md)$ time: one $O(m^3)$-time reduction upfront for the filtration at time $t_0$ followed by one $O(m)$ operation to update the decomposition for the remaining time points $(t_1, t_2, \dots, t_d)$.
+When $d >> m$, this $O(md)$ approach is far more efficient than the $O(dm^3)$  ``naive'' strategy of computing the diagrams independently at every time point.
+
+Despite its theoretical efficiency, vineyards is often not the method of choice in practical settings. 
+Indeed, while there is an increasingly rich ecosystem of software packages offering variations of the standard reduction algorithm (e.g.
 Ripser, PHAT, Dionysus, etc...
@@
 This work seeks to further understand and remedy this discrepancy: building on the work presented in~\cite{busaryev2010tracking}, we introduce a coarser approach to the vineyards algorithm.
-While the vineyards algorithm is efficient at constructing a \emph{continuous} time family of diagrams, it is not necessarily efficient when the time parameter is coarsely discretized.
+While vineyards are  efficient for   \emph{continuous} time families of diagrams, it is not necessarily efficient when the time parameter is coarsely discretized.
 Our methodology is based on the observation that practitioners often don't need (or want!) all $d$ persistence diagrams generated by a continuous 1-parameter of filtrations; usually just $n << d$ of them are sufficient.
@@
 By exploiting the ``donor'' concept introduced in in~\cite{busaryev2010tracking}, we are able to make a tradeoff in the number of times the decomposition is restored to a valid state, dramatically reducing the total number of column operations needed to apply an arbitrary permutation to the filtration.
-This tradeoff, paired with a fast greedy heuristic explained in section~\ref{sec:proxy_objective}, yields an algorithm that can update a $R = DV$ decomposition more efficiently than vineyards in sparse contexts, making dynamic persistence more computationally tractable for a wider class of use-cases.
+This trade off, paired with a fast greedy heuristic explained in section~\ref{sec:proxy_objective}, yields an algorithm that can update a $R = DV$ decomposition more efficiently than vineyards in coarse contexts, making dynamic persistence more computationally tractable for a wider class of use-cases.
 
@@
  \subsection{A Motivating Example and Contributions}\label{sec:motivation} 
-As a motivating example, we describe a simple experiment illustrating why the vineyards algorithm does not always yield an efficient method in many time-varying applications. 
+As a motivating example, we describe a simple experiment illustrating why the vineyards algorithm does not always yield an efficient method for time-varying applications. 
 Consider a series of grayscale images (i.e.
@@
 Thus, the data set is a 1-parameter family of images.
-Each image contains the same number of pixels whose intensities---these are determined by the parameter---vary with time.
+Each image contains the same number of pixels whose intensities---these are determined by $\tau$---vary with time.
 For each image, we build a filtered simplicial complex using the \emph{Freudenthal} triangulation, with the ordering of inclusions obtained by lower stars.
@@
   On the left side of Figure~\ref{fig:vineyards}, we compare the cumulative cost (in total number of column operations) of various approaches which one can use to simulate persistence dynamically. 
-  Since it is unknown a priori at what points in time the persistent homology will change in application settings, one solution is to discretize the time domain at $n$ evenly spaced points and compute the persistence diagram at each of these time points independently.
+  Since it is unknown a priori at which times  the persistent homology will change, one solution is to discretize the time domain at $n$ evenly spaced points and compute the persistence diagrams at each one of them  independently.
 An alternative approach is to construct a homotopy between two time-points, and then decompose this homotopy into adjacent transpositions--the vineyards approach.
@@
 \subsection{Vineyards}\label{sec:vineyards}
-The original purpose of the vineyards algorithm, as described in~\cite{cohen2006vines}, was to compute a continuous 1-parameter family of persistence diagrams over a time-varying filtration, and to detect homological critical events during the construction.
+The original purpose of the vineyards algorithm, as described in~\cite{cohen2006vines}, was to compute a continuous 1-parameter family of persistence diagrams given a time-varying filtration, and to detect homological critical events during the construction.
 Notice homological critical events can only occur when the filtration order changes, though not all changes in the filtration order result in homology changes.
@@
 However, Boissonnat~\cite{boissonnat2018efficient} has shown that for any $d$-dimensional simplicial complex $K$ with $m$ simplices, the number of distinct filtrations built from $K$ is \emph{at least} $\floor{\frac{t+1}{d+1}}^m$, where $t$ is the number of distinct filtration values associated with the complex $K$.
-In the case where $t \sim O(m)$ and $d << m$ fixed, this lower bound grows similarly to $m!$---thus, $O(m - \sqrt{m})$ is not too pessimistic of an upper bound when the pair of filtrations are truly random.
+Since this lower bound grows similarly to $m!$ when $t \sim O(m)$ and $d << m$ fixed, we expect $O(m - \sqrt{m})$ is not too pessimistic of an upper bound when the pair of filtrations are truly random.
 In practice, when one has a time-varying filtration and the sampling points are relatively close [in time], the LCS between adjacent filtrations is expected to be much larger, subsequently making $d$ much smaller.
@@
 However, like in the vineyards algorithm, certain pairs of simplices cost more to exchange depending on whether they are critical pairs in the sense described~\cite{cohen2006vines}, making specific move operations more expensive than others.
-This small variation in cost-per-transposition induces a large variability in the cost of executing any given schedule, which is undesirable computationally.
-Ideally, we would like to generate a schedule which not only is minimal in size, but also cost.
-Formally, given two permutations $p, q \in S_m$, let $S = ( s_1, s_2, \dots, s_d )$ denote a sorting such that $S \circ p = q$.
-%The example given in section~\ref{sec:greedy} shows all (minimal size) $d!$ possible choices of $S$ for a fixed pair of permutations and choice of $\mathrm{LCS}(p,q)$. 
-Let $\nu = \mathrm{LCS}(p,q)$ and let $\sigma \in S_d$ denote a choice of ordering for the symbols in  $p \smallsetminus \nu$, such that $S(\sigma)$ denotes a schedule generated from this ordering.
-Rephrasing from the description above, we would like to achieve something similar to the following objective:
-\begin{equation}\label{eq:schedule_opt}
-	S_\sigma^\ast = \argmin_{\sigma \in S_d} \mathrm{cost}(S(\sigma) \circ p)
-\end{equation}
-where $\mathrm{cost}(S(\sigma) \circ p)$ measures the number of column operations require to execute $S(\sigma)$ on an existing decomposition.
-There are exponentially many ways to generate such a sorting $S$.
-% Dynamic programming discussion 
+%This variation in cost exaggerates the variability of the cost of scheduling move operations.
+
 One advantage of Algorithm~\ref{alg:mr} is that the cost of applying a permutation to some given $R = D V$ decomposition can be determined efficiently prior to performing any column operations. 
 Specifically, the cost of determining $\lvert \mathbb{I}\rvert + \lvert \mathbb{J}\rvert$ in lines (2) and (3) of Algorithm~\ref{alg:mr} cannot take more than $O(m)$ time.
-If we assume a similar complexity to determine the cost of applying a permutation with Algorithm~\ref{alg:ml} and we know to execute $d$ moves, one might expect be to use an optimization method like dynamic programming.
+If we assume a similar complexity to determine the cost of applying a permutation with Algorithm~\ref{alg:ml} and we know to execute $d$ moves, an initial thought might be to use an optimization method like dynamic programming.
 That is, suppose $O(dm)$ move costs are computed ahead of time and these costs are stored in a table providing $O(1)$ element access and modification.
@@
 
+As before, given two permutations $p, q \in S_m$, let $S = ( s_1, s_2, \dots, s_d )$ denote a sorting such that $S \circ p = q$. 
+In general, there are exponentially many ways to generate such a sorting $S$. 
+The example given in section~\ref{sec:greedy} shows all (minimal size) $d!$ possible choices of $S$ for a fixed pair of permutations and choice of $\mathrm{LCS}(p,q)$. 
+Let $\tau = \mathrm{LCS}(p,q)$ and let $\sigma \in S_d$ denote a choice of ordering   for the symbols in  $p \smallsetminus \tau$, 
+such that $S(\sigma)$ denotes a schedule generated from this ordering. Ideally, we would like to achieve something similar to the following objective: 
+\begin{equation}\label{eq:schedule_opt}
+	S_\sigma^\ast = \argmin_{\sigma \in S_d} \mathrm{cost}(S(\sigma) \circ p)
+\end{equation}
+where $\mathrm{cost}(S(\sigma) \circ p)$ measures the number of column operations require to execute $S(\sigma)$ on an existing decomposition. 
 The schedule $S_1$ shown in Table~\ref{table:move_costs} would be an example of a minimal cost schedule.
@@
 \end{equation}
-When the set of permutations $S$ are adjacent transpositions, equality is achieved above since $K_\tau(\hat{s}_i, \hat{s}_{i+1}) = 1$ for every inversion between $p$ and $q$, though this not guaranteed if $S$ was derived using the strategy from section~\ref{sec:move_schedules}.
-Obtaining $S_\sigma^\ast$ via direct optimization over $S_d$ isn't practical, but experimental tests suggest that a greedy-type procedure performs well as a heuristic, which yields a $O(d^2 m \log m)$ minimization procedure, which is efficient when $d <<  m$.
+When the set of permutations $S$ are adjacent transpositions, equality is achieved above since $K_\tau(\hat{s}_i, \hat{s}_{i+1}) = 1$ for every inversion between $p$ and $q$.
+However, this is not guaranteed if $S$ was derived using the strategy from section~\ref{sec:move_schedules}.
+Obtaining $S_\sigma^\ast$ via direct optimization over $S_d$ isn't practical, but experimental tests suggest that a greedy-type procedure performs well as a heuristic, which yields a $O(d^2 m \log m)$ minimization procedure.
+When $d <<  m$, this strategy of successively minimizing the Kendall-$\tau$ distance is efficient.
 %where the set and order of pairs $\hat{s}_\ast$ depend on $S(\sigma)$. 
@@
 \end{equation}
-Although very similar to the Kendall-$\tau$ distance, the Spearman distance offers some convenient computational advantages: it can be computed in $O(m)$ time, the variance of its asymptotic distribution is larger than Kendalls for uniformly random permutations, and it satisfies a certain monotonicity property under LCS edit operations.
-Specifically, given two permutations $p,q \in S_m$ with $\nu = \mathrm{LCS}(p,q)$ between them, the analysis in section~\ref{sec:move_schedules} shows we must move each symbol $x \in \mathcal{D}$ where $\mathcal{D} = p \smallsetminus \nu $ in a way that extends $\nu$ after each move.
-Each symbol $x \in \mathcal{D}$ induces a permutation $s_x$ such that $F(p, q) \geq F(s_x \circ p, q)$.
-That is, unlike the Kendall distance, the Spearman distance satisfies:
-$$ F(p, q) \geq F(s_1 \circ p, q) \geq F( s_2 \circ s_1 \circ p, q) \geq F( s_d \circ \dots \circ s_2 \circ s_1 \circ p, q)  = F(q,q) = 0 $$
-For any sorting $S$ such that $s_{i}$ that increases the size of longest common subsequence between $s_{i-1} \circ \dots \circ s_1 \circ p$ and $q$.
-Since the complexity of the move operations discussed in section~\ref{sec:moves} depend on the size of the movement, it is advantageous to construct a schedule which minimizes this movement.
-%Since $\lvert \mathcal{D} \rvert$ is fixed, we would like to choose $x \in \mathcal{D}$ such that the corresponding permutation $s_x$ reduces the total displacement of every symbol $x \in \mathcal{D}$ relative to  $q$.
-We consider the following objective modified from equation~\ref{eq:schedule_opt}: 
+Although very similarity to the Kendall-$\tau$ distance, the Spearman distance offers some convenient computational advantages: it can be computed in $O(m)$ time, and the variance of its asymptotic distribution is larger.
+Consider the following minimization:
 \begin{equation}\label{eq:schedule_opt_succ}
@@
 \end{equation}
-The minimizer of this optimization can be interpreted as a sorting which minimizes the net displacement amongst all possible sortings, in the $\ell_1$ sense.
-This minimization problem is related to a constrained version of the bipartite crossing minimization problem for $k$ sets of permutations, which is NP-hard for $k \geq 4$~\cite{biedl2009complexity}.
-As a result, we instead consider a greedy approach.
-A greedy strategy requires $O((d+(d-1)+(d-2)+\dots+1)m) \approx O(d^2 m)$ time to complete, since there are $O(d)$ symbols to choose from and minimizing $F(\hat{s}_i, \hat{s}_{i+1})$ at each step takes $O(md)$.
-Since $\mathcal{D}$ is fixed, a greedy strategy thus depends on our ability to quickly compute $F(\hat{s}_i, \hat{s}_{i+1})$.
+The minimizer of this optimization can be interpreted as a sorting which minimizes the net displacement amongst all possible sortings, again in the $\ell_1$ sense. 
 %\\
@@
 %% TODO: elaborate on monotonicity, describe segment tree
+Given two permutations $p,q \in S_m$ and a LCS $\tau = \mathrm{LCS}(p,q)$ between them, the analysis in section~\ref{sec:move_schedules} shows we must move each symbol $x \in \mathcal{D}$ where $\mathcal{D} = p \smallsetminus \tau$ in a way that extends $\tau$ after each move.
+Thus we have $d = \lvert \mathcal{D} \rvert$ symbols to choose from initially.
+Each symbol $x \in \mathcal{D}$ yields a permutation $s$ to apply to $p$.
+Unlike the Kendall distance, the Spearman distance satisfies:
+$$ F(p, q) \geq F(s_1 \circ p, q) \geq F( s_2 \circ s_1 \circ p, q) \geq F( s_d \circ \dots \circ s_2 \circ s_1 \circ p, q)  = F(q,q) = 0 $$
+For any sorting $S$ such that $s_{i}$ that increases the size of longest common subsequence between $s_{i-1} \circ \dots \circ s_1 \circ p$ and $q$.
+Since $\lvert \mathcal{D} \rvert$ is fixed, we would like to choose $x \in \mathcal{D}$ such that the corresponding permutation $s_x$ reduces the total displacement of every symbol $x \in \mathcal{D}$ relative to  $q$.
+A greedy strategy chooses $x$ such that $F(\hat{s}_i, \hat{s}_{i+1})$ is as small as possible, removes that symbol from $\mathcal{D}$, and recurses.
+Thus, a greedy strategy depends on our ability to quickly compute $F(\hat{s}_i, \hat{s}_{i+1})$.
 \\
@@
 The greedy heuristic proposed to minimize equation~\ref{eq:schedule_opt_succ} was used to produce $S_1$, which has the minimal number of crossings amongst all possible schedules in this example, since $K_\tau(p,q) = 21$.
+This minimization problem is related to a constrained version of the bipartite crossing minimization problem for $k$ sets of permutations, which is NP-hard for $k \geq 4$~\cite{biedl2009complexity}.
 \\
@@
 The monotonic nature of successively applying the Spearman distance enables certain computational advantages.
-Suppose we begin with an array $\mathcal{A}$ of size $m$ which provides $O(1)$ access and modification, initialized with the (signed) \emph{displacement} of every element in $p$ to its corresponding position in $q$.
-Since the Spearman distance is simply the sum of these displacements, if we could replace $\mathcal{A}$ with a data structure that supports $O(\log m)$ access time to aggregate information (i.e.
-the sum) and $O(\log m)$ modification time ability on $\lvert i - j \vert$ entries (simultaneously), we would have an efficient $O(d \log m)$ greedy solution.
-The former problem reduces to the problem of efficiently calculating \emph{prefix sums}, which is easily solved.
-It's not immediately clear how to achieve the latter modification complexity in $O(\log m)$ time, since $\lvert i - j \rvert \leq m$ is potentially larger than $O(\log m)$.
-However, observe that each permutation $s_x$ changes the displacement of every symbol in a very predictable way: if $s_x$ represents a move permutation that moves $x$ from position $i$ to position $j$, then relative to the previous permutation $s_{i}$, the displacement of $x$ in $s_{i+1}$ changes by $\lvert i - j \rvert$ and the displacement of any symbol between $i$ and $j$ changes by $\pm 1$.
-All other symbol displacements are unchanged.
-%, and 
-It is known that one can apply constant-factor updates to multiple values in a prefix sum in $O(\log m)$ time by any prefix-sum data structure that also supports \emph{range updates}, e.g.
-a Fenwick tree or a Segment Tree that supports lazy propagation~\cite{de1997computational}.
-Single element prefix sum modifications can also be achieved $O(\log m)$ time with these data structures, the time to compute $F(\hat{s}_i, \hat{s}_{i+1})$ is bounded above by $O(\log m)$ and thus the greedy solution to equation~\ref{eq:schedule_opt_succ} can be computed in $O(d \log m)$ time.
+Suppose we begin with an array $\mathcal{A}$ of size $m$ which provides $O(1)$ access and modification, initialized to $0$.
+Summing the entries in this array yields the Spearman distance in $O(m)$ time, which at initialization represents $F(p, p)$.
+Each symbol $x \in \mathcal{D}$ induces a  permutation $s_x$ such that $F(p, q) \geq F(s_x \circ p, q)$.
+Since there are $d = \lvert \mathcal{D} \rvert$ symbols to move and computing $F$ takes $O(m)$ time, a na\"ive greedy strategy requires $O((d+(d-1)+(d-2)+\dots+1)m) \approx O(d^2 m)$ time to complete.
+However, observe that each permutation $s_x$ changes the displacement of every symbol in a very predictable way: if $s_x$ represents a move permutation that moves $x$ from position $i$ to position $j$, then relative to the previous permutation $s_{i}$, the displacement of $x$ in $s_{i+1}$ changes by at most  $\lvert i - j \rvert$ and the displacement of any symbol between $i$ and $j$ changes by $\pm 1$.
+All other symbols are unaffected.
+Since the Spearman distance is simply the sum of these displacements, replacing $\mathcal{A}$ with a structure that supports both $O(\log m)$ access time to aggregate information (such as the sum) and $O(\log m)$ modification time ability to up to $\lvert i - j \vert$ entries.
+The former problem reduces to the problem of efficiently calculating and updating \emph{prefix sums}, however since $\lvert i - j \rvert \leq m$ is potentially larger than $O(\log m)$, it's not immediately clear how to achieve the latter modification complexity in $O(\log m)$ time.
+Fortunately, since every elements displacement other than $x$ changes by $\pm 1$ per update, both queries can be efficiently done in $O(\log m)$ time using a Segment Tree that supports \emph{lazy propagation}~\cite{de1997computational}.
 %We give an example and a more detailed discussion for those interested in the appendix\ref{sec:appendix}.
+
 %This cumulative Spearman distance enables certain computational advantages.

### 2021-09-22 move_schedules.tex
Changes from jperea

@@
 
+\newsiamremark{remark}{Remark}
+
 \DeclareMathOperator*{\argmax}{arg\,max}
@@
 % is automatically inserted before the title.
-\title{Move Schedules: Fast persistence computations in sparse dynamic settings
-	\thanks{\textbf{Funding:} This work was partially supported by the National Science Foundation through grants CCF-2006661 and CAREER award DMS-1943758.}
+\title{Move Schedules: Fast persistence computations in coarse dynamic settings\thanks{\textbf{Funding:} This work was partially supported by the National Science Foundation through grants CCF-2006661 and CAREER award DMS-1943758.}
 }
@@
 	Matt Piekenbrock\thanks{Department of Computational Mathematics, Science \& Engineering, Michigan State University.}
-	\and Jose A. Perea\thanks{Department of Mathematics, Michigan State University.}
+	\and Jose A. Perea\thanks{
+	Department of Computational Mathematics, Science \& Engineering, 
+	Department of Mathematics, Michigan State University.}
 }
@@
 Computationally, simulating persistence dynamically can be reduced to maintaining a valid decomposition under adjacent transpositions in the filtration order.
-In practice, the quadratic scaling in the number of  transpositions often makes this maintenance procedure slower than simply computing the decomposition from scratch, effectively limiting the application of dynamic persistence to relatively small data sets.
+	In practice, the quadratic scaling in the number of  such transpositions often makes this maintenance procedure slower than simply computing the decomposition from scratch,  limiting the application of dynamic persistence to relatively small data sets.
 In this work, we propose a coarser strategy for maintaining the decomposition over a discrete 1-parameter family of filtrations.
@@
 We then show a modification of this technique which maintains only a sublinear number of valid states, as opposed to a quadratic number of states, and we provide tight lower bounds for this technique.
-	Finally, we provide empirical results suggesting that the decrease in operations needed to compute diagrams across a family of filtrations is proportional to the difference between the expected quadratic number of states, and the proposed sublinear coarsening.
+	Finally, we present empirical results suggesting that the decrease in operations needed to compute diagrams across a family of filtrations is proportional to the difference between the expected quadratic number of states, and the proposed sublinear coarsening.
 	Applications to multi-dimenisonal persistence and crocker stacks are also presented. 
@@
 This motivates the study and application of persistence to time-varying or dynamic settings, like that of dynamic metric spaces~\cite{kim2020spatiotemporal}. 
-As persistence-related computations tend to exhibit high algorithmic complexity (e.g.
-essentially cubic\footnote{For finite fields, it is known that the persistence computation reduces to the PLU factorization problem, which takes $O(m^\omega)$ where $\omega \approx 2.373$ is the matrix multiplication constant.} in the size of the underlying filtration~\cite{morozov2005persistence}), their adoption to dynamic settings poses a challenging computational problem.
+As persistence-related computations tend to exhibit high algorithmic complexity---essentially cubic\footnote{For finite fields, it is known that the persistence computation reduces to the PLU factorization problem, which takes $O(m^\omega)$ where $\omega \approx 2.373$ is the matrix multiplication constant.} in the size of the underlying filtration~\cite{morozov2005persistence}---their adoption to dynamic settings poses a challenging computational problem.
 With the current state of the art tools, there is no recourse to the problem of computing persistence from a time-varying filtration containing millions of simplices at thousands of snapshots in time.
-An efficient means of updating the persistence of a [time-varying] filtration has far-reaching consequences: methods that vectorize persistence diagrams for machine learning purposes, like adaptive template functions~\cite{polanco2019adaptive} and persistence images~\cite{adams2017persistence}, or summary statistics like $\alpha$-smoothed Betti curves~\cite{ulmer2019topological}, all immediately become computationally viable tools in dynamic  settings.
+Efficiently  updating the persistence of a time-varying filtration has far-reaching consequences: methods that vectorize persistence diagrams for machine learning purposes, like adaptive template functions~\cite{polanco2019adaptive} and persistence images~\cite{adams2017persistence}, or summary statistics like $\alpha$-smoothed Betti curves~\cite{ulmer2019topological}, all immediately become computationally viable tools in dynamic  settings.
 %This analogy provides a strong interpretation at gaining insight on the subtle topological changes that may occur in studying a continuous process.
@@
 The main step in the reduction algorithm is a specific decomposition $R = D V$ of the boundary matrix $D$ of $K$. 
-Since $V$ is always full rank (and upper-triangular), an alternative is to compute a $D = RU$ decomposition, where $U = V^{-1}$. 
-The vineyards algorithm transforms a time-varying filtration into an ordered set of adjacent transpositions to be applied to the rows and columns of the decomposition $D = RU$, each of which take at most $O(m)$ time to execute.
+Since $V$ is always full rank and upper-triangular, an alternative is to compute a $D = RU$ decomposition, where $U = V^{-1}$. 
+The vineyards algorithm transforms a time-varying filtration into an ordered set of permutations to be applied to adjacent rows and columns of the decomposition $D = RU$, each of which takes at most $O(m)$ time to execute.
 If one is interested in understanding how the persistent homology of a continuous function changes over time, then this algorithm is sufficient, for  homological critical points can only occur when the filtration order changes.
-This algorithm is also efficient  asymptotically: if there are $d$ time-points where this filtration order changes, then the vineyards algorithm takes $O(m^3 + md)$ time: one $O(m^3)$-time reduction upfront for the filtration at time $t_0$ followed by one $O(m)$ operation to update the decomposition for the remaining time points $(t_1, t_2, \dots, t_d)$.
+This algorithm is also efficient  asymptotically: if there are $d$ time-points where the filtration order changes, then   vineyards   takes $O(m^3 + md)$ time; one $O(m^3)$-time reduction upfront for the filtration at time $t_0$ followed by one $O(m)$ operation to update the decomposition for the remaining time points $(t_1, t_2, \dots, t_d)$.
 When $d >> m$, this $O(md)$ approach is far more efficient than the $O(dm^3)$  ``naive'' strategy of computing the diagrams independently at every time point.
@@
 Ripser, PHAT, Dionysus, etc...
-see~\cite{otter2017roadmap} for an overview), implementations of the vineyards algorithm are relatively uncommon.\footnote{Dionysus 1 has an implementation of vineyards, but the algorithm was never ported to the 2nd version.} The reason for this disparity is perhaps explained by Lesnick and Wright~\cite{lesnick2015interactive}: ``While an update to an $RU$ decomposition involving few transpositions is very fast in practice...
+see~\cite{otter2017roadmap} for an overview), implementations of the vineyards algorithm are relatively uncommon.\footnote{Dionysus 1 has an implementation of vineyards, but the algorithm was never ported to the 2nd version.}
+The reason for this disparity is perhaps explained by Lesnick and Wright~\cite{lesnick2015interactive}: ``While an update to an $RU$ decomposition involving few transpositions is very fast in practice...
 many transpositions can be quite slow...
-it is sometimes much faster to simply recompute the $RU$-decomposition from scratch using the standard persistence algorithm.'' Indeed, whether performing the reductions naively or updating the decomposition dynamically using vineyards, they observed that obtaining the decomposition is the most computationally demanding aspect of their 2-D persistence algorithm.
+it is sometimes much faster to simply recompute the $RU$-decomposition from scratch using the standard persistence algorithm.''
+Whether performing the reductions naively or updating the decomposition dynamically using vineyards, they observe that this is the most computationally demanding aspect of  RIVET \cite{rivet}.
 
 This work seeks to further understand and remedy this discrepancy: building on the work presented in~\cite{busaryev2010tracking}, we introduce a coarser approach to the vineyards algorithm.
-While vineyards are  efficient for   \emph{continuous} time families of diagrams, it is not necessarily efficient when the time parameter is coarsely discretized.
+While vineyards is  designed  for handling   \emph{continuous} time families of diagrams, it is not necessarily efficient when the time parameter is coarsely discretized.
 Our methodology is based on the observation that practitioners often don't need (or want!) all $d$ persistence diagrams generated by a continuous 1-parameter of filtrations; usually just $n << d$ of them are sufficient.
@@
 As a motivating example, we describe a simple experiment illustrating why the vineyards algorithm does not always yield an efficient method for time-varying applications. 
-Consider a series of grayscale images (i.e.
-a video) depicting a fixed-width annulus expanding about the center of a $9 \times 9$ grid, where scale of the annulus is parameterized by a single parameter $\tau$.
-Thus, the data set is a 1-parameter family of images.
+Consider a series of grayscale images (i.e., a video) depicting a fixed-width annulus expanding about the center of a $9 \times 9$ grid, where the scale of the annulus is parameterized by a single parameter $\tau$.
 Each image contains the same number of pixels whose intensities---these are determined by $\tau$---vary with time.
-For each image, we build a filtered simplicial complex using the \emph{Freudenthal} triangulation, with the ordering of inclusions obtained by lower stars.
+For each image, we build a filtered simplicial complex using the \emph{Freudenthal} triangulation of the plane, with the order of inclusions obtained by lower stars.
  \begin{figure*}[t]
@@
 On the left, the cumulative number of column operations required to compute persistence for this time-varying filtration.
-The persistent Betti numbers for each filtration from left to right are: $\beta_1 = (1,0,0), \beta_2 = (1,1,0), \beta_3 = (1,1,0), \beta_4 = (1,1,0), \beta_5 = (4,0,0)$.
-Observe computing $10$ diagrams independently at evenly spaced time points (the green line) both captures these major topological changes and is the most computationally efficient approach shown.}
+The   Betti numbers for each of the shown  steps in the filtration, from left to right, are: $(\beta_0,\beta_1, \beta_2)  = (1,0,0), \; (1,1,0),  \; (1,1,0), \; (1,1,0), \; (4,0,0)$.
+Observe that computing the diagrams independently at 10 evenly spaced time points (the green line) both captures these major topological changes and is the most computationally efficient approach shown.}
  	\label{fig:vineyards}
  \end{figure*}
-Snapshots of the image data and its corresponding time-varying filtration is depicted in Figure~\ref{fig:vineyards} (top right and bottom right, respectively), which we use as a baseline for comparing the standard algorithm $\texttt{pHcol}$ with vineyards.
-Note the size of the filtration is fixed--no simplices are ever added or removed over time.
+Snapshots of the image data and its corresponding time-varying filtration is depicted in Figure~\ref{fig:vineyards} (top right and bottom right, respectively), which we use as a baseline for comparing the standard reduction algorithm $\texttt{pHcol}$ (Algorithm \ref{alg:reduce}) with vineyards.
+Note that the size of the filtration is fixed--no simplices are ever added or removed over time.
 There are two events which dramatically alter the persistence diagrams extracted from these filtrations: the first occurs when the central connected component splits to form a cycle, and the second occurs when the annulus approaches the outer dimensions of the grid, splitting into four components.
@@
   Consider the situation where one is concerned with capturing these dramatic changes in homology.
-  On the left side of Figure~\ref{fig:vineyards}, we compare the cumulative cost (in total number of column operations) of various approaches which one can use to simulate persistence dynamically. 
+  On the left side of Figure~\ref{fig:vineyards}, we compare the cumulative cost (in total number of column operations) of various approaches.
   Since it is unknown a priori at which times  the persistent homology will change, one solution is to discretize the time domain at $n$ evenly spaced points and compute the persistence diagrams at each one of them  independently.
-An alternative approach is to construct a homotopy between two time-points, and then decompose this homotopy into adjacent transpositions--the vineyards approach.
+An alternative approach is to construct a homotopy between two contiguous time-points, and then decompose this homotopy into adjacent transpositions--the vineyards approach.
 The former is often used in practice, but it is only an approximation; the latter is guaranteed to capture all homological changes in persistence that occur in simulating the 1-parameter family.
   The graph in Figure~\ref{fig:vineyards} depicts two different size approximations (purple, $n=100$ and green, $n=10$), wherein the reduction algorithm (\texttt{pHcol}) was applied at $n$ time points, and two exact strategies (black and blue), which compute all $\approx 7 \times 10^{4}$ diagrams given by the homotopy.
-As one can see from the figure, if one needs to simulate the homotopy exactly, the vineyards approach is indeed far more efficient than naively applying the reduction algorithm independently at all time points.
+  As one can see from the figure, if one needs to simulate the homotopy exactly, then the vineyards approach is indeed far more efficient than naively applying the reduction algorithm independently at all time points.
   However, when the discretization of the time domain is sparse enough, the naive approach actually performs less column operations than the vineyards strategy, while still capturing the main events.
 
-  The existence of a time discretization that is more efficient to compute than continually updating the decomposition indicates that the vineyards algorithm must incur some amount of overhead (in terms of column operations) to perform the transpositions.
-That is, even in situations where one is updating a decomposition to a filtration that is  relatively ``close'', e.g.
+  The existence of a time discretization that is more efficient to compute than continually updating the decomposition, indicates that the vineyards algorithm must incur some amount of overhead (in terms of column operations) to perform transpositions, even if the persistence diagrams remain unchanged.
+  Even in situations where one is updating a decomposition to a filtration that is  relatively ``close'', e.g.
 any two adjacent filtrations in the bottom right of Figure~\ref{fig:vineyards}, it is still more efficient to apply \texttt{pHcol} directly than  to iteratively update the decomposition (see the case where $n = 10$).
- Indeed, any optimizations to the reduction algorithm (for example, the clearing optimization~\cite{chen2011persistent}) would only increase this disparity, making the vineyards algorithm less computationally attractive even in settings where $n$ is larger.
-Since dynamic or time-varying are common settings where persistence is applied (see ~\cite{topaz2015topological, xian2020capturing, lesnick2015interactive, kim2020spatiotemporal}), it is worth studying this phenomenon.
-
-\textbf{Our results and approach are as follows:} First, we leverage work of Busaryev et.
-al.~\cite{busaryev2010tracking} to include a coarser type of operation to the dynamic setting, called \emph{moves}.
+ Moreover, any optimization to the reduction algorithm (e.g.,  clearing~\cite{chen2011persistent}) would only increase this disparity.
+ See~\cite{topaz2015topological, xian2020capturing, lesnick2015interactive, kim2020spatiotemporal} 
+ for relevant scenarios and applications.
+
+\textbf{Our approach and contributions are as follows:} First, we leverage the \emph{moves} framework of Busaryev et.
+al.~\cite{busaryev2010tracking} to include  coarser  operations---in terms of valid intermediate decomposition states, as compared to vineyards---for the dynamic setting.
 We give a tight lower bound on the number of moves needed to perform an arbitrary permutation to the $R = D V$ decomposition, and give a proof of optimality by a reduction to the permutation edit distance problem.
 We also give worst-case bounds in expectation as well as efficient algorithms for achieving these bounds---both of which are derived from a reduction to the Longest Increasing Subsequence (LIS) problem.
-This reduction yields an efficient algorithm for generating sequences of moves $( s_1, s_2, \dots, s_d )$, which we call \emph{schedules}. 
-We investigate the feasibility of optimizing the cost 
-(i.e., number of column operations) of these schedules directly, including the existence of a greedy-type approach, which we show can lead to arbitrarily bad behavior. 
-In light of these results, we give an alternative proxy-objective to minimize, provide bounds justifying its relevance to the original objective, and give an efficient $O(d \log d)$ algorithm for approximately solving this proxy minimization.
-A performance comparison with other reduction-based persistent homology computations is given, wherein our approach is demonstrated to be an order of magnitude more efficient than existing approaches in applications such as flock simulation and 2D persistence computation.
-We conclude with a discussion of situations where our strategy is most applicable and future directions for research.
+This reduction yields an efficient algorithm for generating sequences of moves $( s_1, s_2, \dots, s_d )$ of minimal size $d$, which we call \emph{schedules}. 
+However, not all valid schedules incur the same permutation costs 
+(i.e., number of column operations).
+We investigate the feasibility of choosing optimal cost schedules, and show that 
+greedy-type approaches can lead to arbitrarily bad behavior. 
+In light of these results, we give an alternative proxy-objective for cost minimization, provide bounds justifying its relevance to the original problem, and give an efficient $O(d \log d)$ algorithm for approximately solving this proxy minimization.
+A performance comparison with other reduction-based persistent homology computations is given; 
+our move schedules are demonstrated to be an order of magnitude more efficient than existing approaches in applications such as flock simulation and 2D persistence computations. 
+%We conclude with a discussion of situations where our strategy is most applicable and future directions for research.
 
@@
 Another popular optimization is to utilize the duality between   homology and cohomology~\cite{de2011dualities}, which dramatically improves the effectiveness of the clearing optimization~\cite{bauer2021ripser}.
-There are many other optimizations on the implementation side: the use of ranking functions defined on the combinatorial number system enables implicit cofacet enumeration, removing the need to store the boundary matrix explciitly; the apparent/emergent pairs optimization identifies columns whose pivot entries are unaffected by the reduction algorithm, reducing the total number of columns which need be reduced; sparse data structures such as bit-trees and lazy heaps allow for efficient column-wise additions with $\mathbb{Z}_2 = \mathbb{Z}/2\mathbb{Z}$ coefficients and effective $O(1)$ pivot entry retrieval, and so on~\cite{bauer2021ripser, bauer2017phat}.
+There are many other optimizations on the implementation side: the use of ranking functions defined on the combinatorial number system enables implicit cofacet enumeration, removing the need to store the boundary matrix explicitly; the apparent/emergent pairs optimization identifies columns whose pivot entries are unaffected by the reduction algorithm, reducing the total number of columns which need be reduced; sparse data structures such as bit-trees and lazy heaps allow for efficient column-wise additions with $\mathbb{Z}_2 = \mathbb{Z}/2\mathbb{Z}$ coefficients and effective $O(1)$ pivot entry retrieval, and so on~\cite{bauer2021ripser, bauer2017phat}.
 
@@
 
-If  zeroth homology is the only dimension of interest, computing and updating both the persistence and rank information of the persistent homology groups is greatly simplified.
+If  zeroth homology is the only dimension of interest, computing and updating both the persistence and rank information  is greatly simplified.
 For example, if the relations are available a-priori, obtaining a tree representation fully characterizing the connectivity of the underlying space (also known as the \emph{incremental connectivity} problem) takes just $O(\alpha(n) n)$ time using the disjoint-set data structure, where $\alpha(n)$ is the extremely slow-growing inverse Ackermann function.
@@
 
-\subsection{Outline} The remainder of the paper is organized as follows: we review and establish the notation we will use to describe simplicial complexes, persistent homology, and dynamic persistence.
+\subsection{Outline} The remainder of the paper is organized as follows: we review and establish the notations we will use to describe simplicial complexes, persistent homology, and dynamic persistence in Section \ref{sec:background}.
 We also cover the reduction algorithm (designated here as \texttt{pHcol}), the vineyards algorithm, and the set of \emph{move}-related algorithms introduced in~\cite{busaryev2010tracking}, which serves as the starting point of this work.
-In section 3 we introduce our proposed alternative to vineyards and provide efficient algorithms to make this alternative viable. 
-In section 4 we demonstrate a few applications of the proposed method. In section 5 we conclude the paper by discussing other possible applications and future work. 
+In Section \ref{sec:move_schedules} we introduce our move schedules  and provide efficient algorithms to make this alternative viable. 
+In Section \ref{sec:results} we demonstrate a few applications of the proposed method. 
+In particular, to the computation of crocker stacks from flock simulations, and to 2-dimensional persistence on the Mumford data set of image patches from natural images. 
+In Section \ref{sec:conclusion} we conclude the paper by discussing other possible applications and future work. 
 We also include an appendix which introduces certain topics in more detail, such as the vineyards algorithm itself, for readers who may be unfamiliar but would like a more in-depth understanding.
@@
 
-Suppose one has a family $\{K_i\}_{i\in I}$ of simplicial complexes indexed by a totally ordered index set $I$, and so that for any $i< j \in I$ we have $K_i \subseteq K_j$.
-There are two index sets of interest in this effort: $\mathbb{R}$ and $[n] = \{ 1, \dots, n\}$.
+Suppose one has a family $\{K_i\}_{i\in I}$ of simplicial complexes indexed by a totally ordered   set $I$, and so that for any $i< j \in I$ we have $K_i \subseteq K_j$.
+There are two index sets of interest in what follows: $\mathbb{R}$ and $[n] = \{ 1, \dots, n\}$.
 Such a family is called a \emph{filtration},
 which is deemed to be \emph{essential} if $i \neq j$ implies $K_i \neq K_j$.
-Moreover, an essential filtration is said to be  \emph{simplexwise} if $K_j \smallsetminus K_i = \{\sigma_j\}$ for all $i< j \in I$ with $K_j \neq \emptyset$.
-Any filtration may be trivially converted into an essential, simplexwise filtration via a set of \emph{condensing}, \emph{refining}, and \emph{reindexing} maps (see~\cite{bauer2021ripser} for more details).
+Moreover, an essential filtration is said to be  \emph{simplexwise} if $K_j \smallsetminus K_i = \{\sigma_j\}$ 
+whenever $j$ is the immediate successor of $i$ in $I$.
+Any finite filtration may be trivially converted into an essential, simplexwise filtration via a set of \emph{condensing}, \emph{refining}, and \emph{reindexing} maps (see~\cite{bauer2021ripser} for more details).
 As a result, and without loss of generality, here we exclusively consider essential simplexwise filtrations. For brevity, we will simply refer to them as filtrations.
 
-Let $K$ be an abstract simplicial complex. With respect to some field $\mathbb{F}$, a $p$-chain is a
-formal $\mathbb{F}$-linear combination  of $p$-simplices of $K$. The collection of $p$-chains under addition yields an abelian group called the $p$-th chain group of $K$, denoted   $C_p(K)$. 
-The $p$-boundary $\partial_p(\sigma)$  of a $p$-simplex $\sigma$ is the alternating sum of its oriented co-dimension 1 faces; the $p$-boundary of a $p$-chain is the sum of the boundaries of its simplices.
-A $p$-cycle is a $p$-chain with zero boundary.
-The collection of $p$-cycles forms the group $Z_p(K) = \mathrm{Ker}\,\partial_p$, the collection of $p$-boundaries forms the group $B_p(K) = \mathrm{Im}\,\partial_{p+1}$, and the quotient of these two yields the $p$-th homology group, $H_p(K) = Z_p(K) / B_{p}(K)$.
-If $\{K_i\}_{i\in [m]}$ is a filtration, then the inclusion maps  $K_i\subset K_j$  for $i < j$, induce homomorphisms $f_p^{i,j}: H_p(K_i) \to H_p(K_j)$ between their corresponding homology groups:
+Let $K$ be an abstract simplicial complex and $\mathbb{F}$ a field.
+A $p$-chain is a
+formal $\mathbb{F}$-linear combination  of $p$-simplices of $K$. The collection of $p$-chains under addition yields an 
+$\mathbb{F}$-vector space  denoted   $C_p(K)$. 
+The $p$-boundary $\partial_p(\sigma)$  of a $p$-simplex $\sigma\in K$ is the alternating sum of its oriented co-dimension 1 faces; the $p$-boundary of a $p$-chain is defined 
+linearly in terms of its constitutive simplices. 
+A $p$-chain with zero boundary is called a $p$-cycle, and together they form $Z_p(K) = \mathrm{Ker}\,\partial_p$. 
+Similarly, the collection of $p$-boundaries forms  $B_p(K) = \mathrm{Im}\,\partial_{p+1}$.
+Since $\partial_p \circ \partial_{p+1} = 0$ for all $p\geq 0$, 
+then the quotient space $H_p(K) = Z_p(K) / B_{p}(K)$ is well-defined, and called the 
+$p$-th homology of $K$ with coefficients in $\mathbb{F}$. 
+If $\{K_i\}_{i\in [m]}$ is a filtration, then the inclusion maps  $K_i\subset K_j$  for $i < j$, induce linear transformations $f_p^{i,j}: H_p(K_i) \to H_p(K_j)$ at the level of homology:
 \begin{equation}
@@
 \end{equation}
-The $p$-th persistent homology groups are the images of these homomorphisms: $H_{p}^{i,j} = \mathrm{Im}\,f_p^{i,j}$.
-Note that if $i = j$, then $H_{p}^{i,j} = H_{p}(K_i) = H_{p}(K_i)$ is   just the ``standard'' homology.
-It is common for simplices which create new homology classes to be called \emph{creators} and simplices that destroy homology classes to be called \emph{destroyers}.
-Similarly, the filtration indices of 
+The $p$-th persistent homology groups are the images of these transformations: $H_{p}^{i,j} = \mathrm{Im}\,f_p^{i,j}$. 
+%Note that if $i = j$, then $H_{p}^{i,j} = H_{p}(K_i) = H_{p}(K_i)$ is   just the ``standard'' homology. 
+Simplices whose inclusion in the filtration creates a new homology class are   called \emph{creators}, and simplices that destroy homology classes are   called \emph{destroyers}. 
+The filtration indices of 
 these creators/destroyers are referred to as \emph{birth} and \emph{death} times, respectively. 
-If a homology class $[c]$ is born at time $i$ and dies entering time $j$, the difference $\lvert i - j \rvert$ is called the \emph{persistence} of that class.
+The collection of birth/death  pairs 
+$(i,j)$ is denoted $\mathrm{dgm}_p(K)$, 
+and referred to as the $p$-th \emph{persistence diagram} of $K$.
+If a homology class is born at time $i$ and dies entering time $j$, the difference $\lvert i - j \rvert$ is called the \emph{persistence} of that class.
 In practice, filtrations often arise from triangulations parameterized by geometric scaling parameters, and the ``persistence'' of a homology class actually refers to its lifetime with respect to the scaling parameter.
@@
 A VR filtration indexed over $\mathbb{R}$  can be condensed to an essential filtration indexed over the set of pairwise distances $\{\, d(x,x') \; | \; x,x' \in X \, \}$, 
-which can then be further reindexed to an essential simplexwise filtration by extending a total order of $X$ to its power set 
+which can then be further reindexed to an essential simplexwise filtration by extending a total order of $X$ to its power set---e.g., 
 using the lexicographical ordering. 
@@
 Given a triangulable topological space $\mathbb{X}$ equipped with a
-real-valued function $f: \mathbb{X} \to \mathbb{R}$, 
+  function $f: \mathbb{X} \to \mathbb{R}$, 
 we write $\mathbb{X}_a = f^{-1}(-\infty, a]$ to denote the sublevel sets of $\mathbb{X}$ defined by the value $a$. 
 A \emph{homological critical value} of $f$ is any value $a \in \mathbb{R}$ such that the homology of the sublevel sets of $f$ changes at $a$, i.e.
-the inclusion-induced map $H(\mathbb{X}_{a - \epsilon}) \to H(\mathbb{X}_a)$ is not an isomorphism.
-If there are only finitely many of these homological critical values, $f$ is said to be \emph{tame}.
-
-Consider a homotopy $F(x,\tau) : \mathbb{X} \times [0,1] \to \mathbb{R}$ and denote it's ``snapshot'' at a given time-point $\tau$ by $f_\tau(x) = F(x,\tau)$.
+if for some $p$ the inclusion-induced homomorphism  $H_p(\mathbb{X}_{a - \epsilon}) \to H_p(\mathbb{X}_{a+\epsilon})$ is not an isomorphism for  any $\epsilon >0$.
+If there are only finitely many of these homological critical values, then $f$ is said to be \emph{tame}.
+
+Consider a homotopy $F(x,\tau) : \mathbb{X} \times [0,1] \to \mathbb{R}$ and denote its ``snapshot'' at a given time-point $\tau$ by $f_\tau(x) = F(x,\tau)$.
 The snapshot $f_0$ denotes the initial function at time $\tau = 0$ and $f_1$ denotes the function at the last time step. 
-As $\tau$ varies in $[0,1]$, the points in $\mathrm{dgm}_p(f_\tau)$ trace  curves in $\bar{\mathbb{R}}^3$, called \emph{vines} and which together form a \emph{vineyard}. 
+As $\tau$ varies in $[0,1]$, the points in $\mathrm{dgm}_p(f_\tau)$ trace  curves in $\mathbb{R}^3$, called \emph{vines} and which together form a \emph{vineyard}. 
 The stability of persistence implies that these curves will be continuous if the homotopy is continuous.
-The vineyard analogy acts as a guidepost for practitioners seeking to understand how subtle changes occurring to the topological structure over time reveal intrinsic information  about the underlying continuous process.
-We discuss the vineyards approach more in detail in section~\ref{sec:vineyards}.
+The vineyard analogy acts as a guidepost for practitioners seeking to understand how subtle changes occurring to the topological structure over time.
+A more detailed discussion of the vineyards approach is provided in section~\ref{sec:vineyards}.
 %Let $\mathbb{M}$ be a smooth $d$-manifold and $f : \mathbb{M} \to \mathbb{R}$ a smooth function. 
@@
 The pseudocode is outlined in Algorithm~\ref{alg:reduce}.
-Without optimizations, like the clearing optimization or the use of implicit matrix reduction, the standard  algorithm is very inefficient.
-Nonetheless, the reduction algorithm serves as the foundation of most implementations that compute persistent homology, and its invariants are necessary before introducing both vineyards in section~\ref{sec:vineyards} and our approach in section~\ref{sec:move_schedules}.
+Without optimizations, like   clearing   or       implicit matrix reduction, the standard  algorithm is very inefficient.
+Nonetheless, it serves as the foundation of most implementations that compute persistent homology, and its invariants are necessary before introducing both vineyards in section~\ref{sec:vineyards} and our move schedules  in section~\ref{sec:move_schedules}.
 
@@
 
- The main output of the reduction algorithm is a matrix decomposition $R = D V$, where the persistence diagram is encoded in $R$. 
+ The main output of the reduction algorithm is a matrix decomposition $R = D V$, where the persistence diagram is encoded in $R$, and the generating cycles in the columns of $V$. 
  Given a  filtration $K$ with $m$ simplices and maximal dimension $d$, 
- one assembles the elementary boundary chains $\partial(\sigma)$ as columns ordered according to the filtration order, forming the \emph{filtration boundary matrix} $D$. 
+ one assembles the elementary boundary chains $\partial(\sigma)$ as columns ordered according to the filtration, forming the \emph{filtration boundary matrix} $D$. 
  The reduction algorithm can  compute the persistent homology for all dimensions up to $d - 1$.
  In this case,  $D$ is a square matrix of size $m \times m$.
-Alternatively, a single dimension $i \leq d - 1$ can be computed, in which case one reduces a pair of matrices $D_i, D_{i+1}$, where the former has dimension $d_{i-1} \times d_i$ and the latter $d_{i} \times d_{i+1}$.
-If one is interested in a particular dimension of homology then the latter approach is preferred.
-However, the algorithmic description of the algorithms is simpler when considering the full matrix, so $D$ will always be $(m \times m)$ in the descriptions that follow.
-All algorithms discussed generalize to both situations.
-We give an example using the matrix pair $(R_1, V_1)$  below.
+Alternatively, a single dimension $i \leq d - 1$ can be computed, in which case one reduces a pair of sub-matrices  $D_i$ and $ D_{i+1}$ of $D$,  where $D_i$ has dimension $d_{i-1} \times d_i$ and $D_{i+1}$ is  $d_{i} \times d_{i+1}$.
+ If one is interested in a particular dimension of homology, then the latter approach is preferred.
+However, the  description of the algorithms is simpler when considering the full matrix, so $D$ will always be $(m \times m)$ in what follows.
+ All algorithms discussed generalize to both situations. We give an example using the matrix pair $(R_1, V_1)$  below. 
 \\
@@
 \noindent
-\textbf{Example 2.1: Reduction} Consider a triangle with vertices $u,v,w$ and edges $a = (u,w)$, $b = (v,w)$, $c = (u, v)$ whose filtration order, read as inclusions from left to right, is given as $K = (u, v, w, a, b, c)$.
-The reduction algorithm begins by setting $R = D$ and $V = I$, followed by left-to-right column operations until invariant (2) below is satisfied.
+\textbf{Example 2.1: Reduction} Consider a triangle with vertices $u,v,w$, edges $a = (u,w)$, $b = (v,w)$, $c = (u, v)$,
+and whose filtration order, read as inclusions from left to right, is given as $K = (u, v, w, a, b, c)$.
+The reduction algorithm begins by setting $R = D$ and $V = I$, followed by left-to-right column operations until invariant I2 below is satisfied.
 If the $i$-th column of $R$ is nonzero, then $\mathrm{low}_R(i)$ 
@@
 	\begin{blockarray}{cccc}{R_1}
-	\kern-0.2em & b & a & c  \\
+	\kern-0.2em & a & b & c  \\
 		\begin{block}{c[ccc]}
-  		u \kern\rowsp  & \;     & 1  &  \; \topstrut \\
-  		v \kern\rowsp &  \; 1  & 1  &  \; \\
+  		u \kern\rowsp  & \; 1    & 1  &  \; \topstrut \\
+  		v \kern\rowsp &  \;    & 1  &  \; \\
   		w \kern\rowsp & \; 1  &     &  \; \botstrut \\
@@
 	\begin{blockarray}{cccc}{V_1}
-	\kern-0.2em & b & a & c  \\
+	\kern-0.2em & a & b & c  \\
 		\begin{block}{c[ccc]}
@@
 Its clear from Algorithm~\ref{alg:reduce} that a loose upper bound for this reduction is $O(m^3)$, where $m$ is the number of simplices of the filtration; it turns out that this bound is in fact tight, see~\cite{morozov2005persistence} for more details.
-There are actually many variations and optimizations proposed over the past decade to Algorithm~\ref{alg:reduce}; for example, one can also use the \emph{row algorithm}~\cite{de2011dualities}, or reverse the order of the filtration and compute cohomology~\cite{de2011dualities}, etc.
+There are actually many variations and optimizations proposed over the past decade to Algorithm~\ref{alg:reduce}; for example, one can   reverse the order of the filtration and compute cohomology~\cite{de2011dualities} instead.
 Despite these variations in computation, regardless of how the decomposition is obtained, it must obey the following invariants:
@@
 	\item[\hspace{-1.5em}\textbf{ Decomposition Invariants:}]
- 	\item $R = D V$ where $D$ is the boundary matrix of the filtration $K$
- 	\item $V$ is full-rank upper-triangular, and $R$ is \emph{reduced}: $\mathrm{low}_R(i) \neq \mathrm{low}_R(j)$ if its $i$-th and $j$-th columns are nonzero
+ 	\item[I1] $R = D V$ where $D$ is the boundary matrix of the filtration $K$
+ 	\item[I2] $V$ is full-rank upper-triangular, and $R$ is \emph{reduced}: if its $i$-th and $j$-th columns are nonzero, then $\mathrm{low}_R(i) \neq \mathrm{low}_R(j)$ 
  	\end{enumerate} 
  \vspace*{1em}
- If these two invariants are maintained, then we call the decomposition \emph{valid}.
-Given any $R = D V$ decomposition, it was shown in~\cite{zomorodian2005computing} that the following proposition is true:
+ If these two invariants are satisfied, then we call the decomposition \emph{valid}. Given any valid $R = D V$ decomposition, then~\cite{zomorodian2005computing}:
 %\begin{proposition}[Persistent Pairing~\cite{zomorodian2005computing}]
@@
 \noindent
-Amongst other important corollaries, one immediate consequence of this proposition is that the persistent pairing is independent of the particular entries in $R$ and $V$.
+The collection of persistent pairings is independent of the choice of $R$ and $V$.
 When reducing the boundary matrix, we can therefore perform column operations in any order, as long as columns are added from left to right, and so long as every column operation performed in $R$ is also performed on $V$.
-Once $R$ is reduced, if both of the invariants are respected, then the low entries in $R$ yield the correct pairing encoding the persistence diagram of the corresponding filtration $K$.
+Once $R$ is reduced, then the low entries in $R$ yield the correct pairings encoding the persistence diagram of the corresponding filtration $K$.
 % Important corollaries of the persistence pairing lemma is that the 
@@
 The original purpose of the vineyards algorithm, as described in~\cite{cohen2006vines}, was to compute a continuous 1-parameter family of persistence diagrams over a time-varying filtration, and to detect homological critical events during the construction.
-Notice homological critical events can only occur when the filtration order changes, though not all changes in the filtration order result in homology changes.
-It follows that detecting these homological events reduces to computing matrix decompositions at a finite set of time points interleaved between homology-altering changes. 
+Notice that homological critical events can only occur when the filtration order changes, though not all filtration changes result in homological critical events.
+It follows that detecting these homological events reduces to computing matrix decompositions at a finite set of time points interleaved between filtration changes. 
 At the finest scale, changes to the filtration order manifest as adjacent transpositions. 
 As a result, a fixed set of rules for maintaining a valid $R = D V$ decomposition under the transposition of adjacent simplices in the filtration is enough to simulate persistence dynamically. 
-These rules prescribe certain column and row operations which must be applied to the matrix decomposition either before, during, or after the transposition, in order to ensure decomposition invariants (1) and (2) are respected.
+These rules prescribe certain column and row operations which must be applied to the matrix decomposition either before, during, or after the transposition, in order to ensure decomposition invariants I1 and I2 are respected.
 
 Let $S_{i}^j$ represent the upper-triangular matrix such that multiplying by it on the right is equivalent to adding column $i$ to column $j$. 
-Multiplication by $S_{i}^j$ on the left is equivalent to adding row $j$ to row $i$, and thus $S_{i}^j S_{i}^j = I$. 
-Similarly, let $P$ denote the permutation matrix so that multiplication from the right applies some permutation $p$ to the columns of some matrix $A$. 
-The inverse of any permutation is also its transpose, i.e.
-$P^{-1} = P^T$, and thus one would write $P A P^T$ to denote the application of the permutation $P$ to both the columns and rows of $A$.
+Multiplication by $S_{i}^j$ on the left is equivalent to adding row $j$ to row $i$.
+% NOT SURE about this:, and thus $S_{i}^j S_{i}^j = I$. 
+Similarly, let $P$ denote the matrix so that multiplication from the right applies a fixed permutation to the columns of $A$.
+Since the columns of $P$ are orthonormal, then  $P^{-1} = P^T$, and thus one would write $P^T A P$ to denote the application of the same permutation  to both the columns and rows of $A$.
 In the special case where $P$ represents a transposition, we have $P = P^T$ and may instead simply write $P A P$.
@@
 \textbf{Example 2.2} To illustrate the basic principles on which vineyards works, we re-use the running example introduced in the previous section.
-Below, we illustrate the case of exchanging simplices $a$ and $b$ in the filtration order.
+Below, we illustrate the case of exchanging simplices $a$ and $b$ in the filtration order, and restoring $RV$ to a valid decomposition.
 \begin{displaymath}
@@
 		\begin{block}{c[ccc]}
-        a \kern\rowsp  & \; 1 & 1 & 1 \; \topstrut \\
+        a \kern\rowsp  & \; 1 &  \fbox{1} & 1 \; \topstrut \\
   		b \kern\rowsp & \;  & 1 & 1 \; \\
@@
 \end{displaymath}
-Prior to performing the exchange, observe that that the highlighted entry in $V_1$ would render $V_1$ non-upper triangular after the exchange.
+Starting with a valid reduction $R = DV$ and prior to performing the exchange, observe that that the highlighted entry in $V_1$ would render $V_1$ non-upper triangular after the exchange.
 This entry is removed by a left-to-right column operation, given by applying the $S_1^{2}$ on the right to $R_1$ and $V_1$.
 After this operation, the permutation may be safely applied to $V_1$.
-In this case, both before and after the permutation, $R_1$ is rendered non-reduced, requiring another column operation to restore the decomposition to a valid state.
+Both before and after the permutation $P$, $R_1$ is rendered non-reduced, requiring another column operation to restore the decomposition to a valid state.
 
@@
 A move operation $\mathrm{Move}(i,j)$ is a set of rules for maintaining a valid decomposition under the application of a permutation $P$ that moves a simplex $\sigma_i$ at position $i$ to position $j$.
-If $j = i \pm 1$, this operation obviously is just an adjacent transposition, and thus in some sense Busaryev's move algorithm is a generalization of the vineyards algorithm.
+If $j = i \pm 1$, this operation is an adjacent transposition, and thus in some sense Busaryev's move algorithm is a generalization of  vineyards.
 
@@
  \begin{enumerate}[leftmargin=2\parindent, align=left, labelsep=-5pt, topsep=0pt,itemsep=-0.25ex,parsep=1.2ex]
- 	\item Querying the non-zero status of entries in any matrix representation occurs once per move.
- 	\item The decomposition $R = D V$ is not maintained during the movement of $\sigma_i \mapsto \sigma_j$.
+ 	\item[M1:] Querying the non-zero status of entries in any matrix representation occurs once per move.
+ 	\item[M2:] The decomposition $R = D V$ is not maintained during the movement of $\sigma_i \mapsto \sigma_j$, but only when the move is completed.
  \end{enumerate} 
  \vspace*{1em}
-First, consider property (1).
-Prior to applying any permutation $P$ to the decomposition, it is necessary to remove non-zero entries in $V$ which render $PVP^T$ non-upper triangular, to maintain invariant (2).
+First, consider property M1.
+Prior to applying any permutation $P$ to the decomposition, it is necessary to remove non-zero entries in $V$ which render $PVP^T$ non-upper triangular, to maintain invariant I2.
 If one uses the (augmented) sparse matrix representation described in the original vineyards publication~\cite{cohen2006vines}, checking whether these entries are non-zero takes $O(\log m)$ per column.
@@
 
-Property (2) indicates move operations only guarantee the decomposition is valid after the operation completes, implying that the decomposition is not fully maintained during the execution of \textit{Restore}$\ast$.
+Property M2 indicates move operations only guarantee the decomposition is valid after the operation completes, implying that the decomposition is not fully maintained during the execution of \textit{RestoreRight} and \textit{RestoreLeft} below.
 This is a significant difference compared to the vineyards algorithm, which was designed specifically to maintain a valid decomposition after each transposition. 
@@
             \State $\mathbb{I} =$ columns satisfying $V[i, i:j] \neq 0$ %\Comment{$O(m)$}
-            \State $\mathbb{J} = $ columns satisfying $\mathrm{low}_R \in [i,j]$ and $\mathrm{row}_R(i) \neq 0$ %\Comment{$O(m)$}
+            \State $\mathbb{J} = $ columns satisfying $\mathrm{low}_R \in [i:j]$ and $\mathrm{row}_R(i) \neq 0$ %\Comment{$O(m)$}
             \State $(\, R, \, V, d_R, d_V \,) \gets $ \Call{RestoreRight}{$R$, $V$, $\mathbb{I}$} 
@@
             \State $(R, V) \gets (P R P^T, P V P^T)$
-            \State $\mathbb{J} = $ columns satisfying $\mathrm{low}_R \in [i,j]$ and $\mathrm{row}_R(i) \neq 0$ %\Comment{$O(m)$}
+            \State $\mathbb{J} = $ columns satisfying $\mathrm{low}_R \in [i:j]$ and $\mathrm{row}_R(i) \neq 0$ %\Comment{$O(m)$}
 
@@
 
-If one has a pair of filtrations $(K_0, K_1)$ each with $m$ simplices, and we assume that each simplex $\sigma_i$ switches its relative ordering with another simplex $\sigma_j$ at most once in some ordered schedule of transpositions, then the number of intermediate persistence diagrams is bounded above by $O(m^2)$ using transpositions, whereas with move operations this bound clearly is at most $O(m)$.
+If one has a pair of filtrations $(K_0, K_1)$ each with $m$ simplices, and we assume that each simplex $\sigma_i$ switches its relative ordering with another simplex $\sigma_j$ at most once in some ordered schedule of transpositions, then the number of intermediate persistence diagrams is bounded above by $O(m^2)$ using transpositions, whereas with move operations this bound  is at most $O(m)$.
 \\
@@
 \noindent \textbf{Example:} We continue the example used in sections~\ref{sec:vineyards} and~\ref{sec:reduction} to illustrate moves.
-Consider moving the edge $a$ to the position taken by the edge $c$ in the filtration.
+Consider moving the edge $a$ to the position of edge $c$ in the filtration.
 The donor columns introduced in~\cite{busaryev2010tracking} are shown on the left side of each matrix.
@@
 \noindent
-We recall an important claim given by~\cite{busaryev2010tracking} on the effect that move operations have on the status of simplices in the pairing.
+We recall an important claim given in~\cite{busaryev2010tracking} on the effect that move operations have on the status of simplices in the pairing.
 Recall from section~\ref{sec:background} that simplices which create new homology classes are called \emph{creators} and simplices that destroy homology classes are called \emph{destroyers}.
@@
 \noindent The effect of the movement on intermediate simplices depends on the direction of the movement.
-If $i < j$ (respectively, $j < i$), all simplices at positions $k \in [i,j] \smallsetminus i$ are shifted down (respectively, up) by $1$.
+If $i < j$ (respectively, $j < i$), all simplices at positions $k \in [i+1\;:\;j] $ are shifted down (respectively, up) by $1$.
 The pseudo-code for the \emph{MoveRight} operation is given in Algorithm~\ref{alg:mr}, and the pseudo-code for the \emph{MoveLeft} is given in Algorithm~\ref{alg:ml}.
+
+\begin{remark}
 We note that while the pseudocode for \emph{MoveRight} is included in~\cite{busaryev2010tracking}, the one for \emph{MoveLeft} is not, and Algorithm~\ref{alg:ml} is the one we propose and leverage in this paper.
+\end{remark}
 
@@
 Since these operations always occur in a left-to-right fashion, its not immediately clear how to apply a donor column kind of concept.
-Fortunately, we can still separate out these two stages.
+Fortunately, we can still separate  these two stages.
 % with a logarithmic loss in efficiency, although executing the movement has a higher constant factor.
@@
 \subsection{Proposed approach}
-Let us begin by briefly describing the overall pipeline of our proposed approach, which is outlined in Algorithm~\ref{alg:schedule}.
+Let us begin with a briefly overview of the pipeline, which we outline in Algorithm~\ref{alg:schedule}.
 As before, we assume as input  a discrete $1$-parameter family of filtrations $\mathcal{K} = (K_1, K_2, \dots, K_n)$,
-and the goal being to compute a persistence diagram for each filtration.
+and the goal being to compute their persistence diagrams.
 
@@
 \noindent Algorithm~\ref{alg:schedule} is purely illustrative at this point, 
-and is meant to serve as a guidepost before discussing its other components in depth. 
+and is meant to serve as a guidepost before discussing its  components in depth. 
 Specifically, lines (4-5) are discussed next in section~\ref{sec:schedule_sizes}, and the algorithm for line (6) is given in section~\ref{sec:schedule_cost}. 
@@
 Due to the combinatorial aspect of our approach, there is actually no need to have access to---or explicitly store---the entire family of filtrations and the schedules between them.
-That is, Algorithm~\ref{alg:schedule} can be easily modified to be completely online, keeping at most two filtrations and one decomposition in memory at any given time.
+That is, Algorithm~\ref{alg:schedule} can be  modified to be completely online, keeping at most two filtrations and one decomposition in memory at any given time.
 
@@
 \subsubsection{The Continuous Setting}\label{sec:continuous_setting}
-In the vineyards case, we are given a homotopy $F(x, \tau) = \mathbb{X} \times [0,1] \to \mathbb{R}$ and a pair of filtrations $(K_0, K_1)$ whose simplices are ordered according to $f_0(x) = F(x, 0)$ and $f_1(x) = F(x, 1)$, respectively.
+In the vineyards case, we are given a homotopy $F : K \times [0,1] \to \mathbb{R}$ and a pair of filtrations $K_0, K_1$ of $K$, with simplices  ordered according to $f_0 = F(\cdot, 0)$ and $f_1 = F(\cdot, 1)$, respectively.
 The choice of homotopy $F$ completely determines the number of adjacent transpositions, so to get a meaningful bound we must know something about $F$.
-If we assume the curves traced by the homotopy are in general position and we assume each pair of curves cross each other at most once, then the set of adjacent transpositions $S_F$ that $F$ decomposes into is given by the pairs of simplices which change in their relative ordering:
+If we assume that the curves $\tau \mapsto F(\cdot, \tau)$  are in general position in $[0,1]\times \mathbb{R}$, and that each pair of curves cross  at most once, then the set of adjacent transpositions $S_F$ that $F$ decomposes into is given by the pairs of simplices with changes in their relative ordering:
 \begin{align}\label{eq:sf_schedule}
-	S_F &= \{ (i,j) \, | \, f_0(\sigma_i) < f_0(\sigma_j) \land  f_1(\sigma_i) > f_1(\sigma_j) \quad \forall \; i,j \in [m] \} \\
-	&= ( s_1, s_2, \dots, s_k )
+	S_F &= \{ (i,j) \, | \, f_0(\sigma_i) < f_0(\sigma_j) \mbox{ and }  f_1(\sigma_i) > f_1(\sigma_j) \quad  \; i,j \in [m] \} \\
+	&= \{ s_1, s_2, \dots, s_k \}
 \end{align}
-We assume, without loss of generality, that the underlying complexes $K_0$ and $K_1$ have the same cardinality. 
+%We assume, without loss of generality, that the underlying complexes $K_0$ and $K_1$ have the same cardinality. 
 Since we deal exclusively with essential filtrations, we may think of $K_1$ as simply a reordering of $K_0$ (or vice versa).
 Let $p = [m]$ and let $q = \mathrm{Im}\,f^\ast$, where $f^\ast : [m] \to [m]$ is the one-to-one correspondence induced by the bijection $f: K_0 \to K_1$.
-Then $S_F$ is given by inversions between $p$ and $q$:
-$$ \mathrm{Inv}(p, q) = \{ \, (i,j) \, | \, p(i) < p(j) \land q(i) > q(j)  \, \}$$
-The associated distance, measuring $\lvert S_F \rvert$, is the \emph{Kendall}-$\tau$ distance $K_\tau(p,q)$: 
+Then $S_F$ is given by the inversions between $p$ and $q$:
+\[
+\mathrm{Inv}(p, q) = \{ \, (i,j) \, | \, p(i) < p(j) \mbox{ and } q(i) > q(j)  \quad \; i,j \in [m]\, \}
+\]
+The cardinality of $  S_F $ is exactly the \emph{Kendall}-$\tau$ distance $K_\tau(p,q)$ between $p$ and $q$: 
 \begin{equation}\label{eq:kendall_dist}
@@
 parameterized curves whose behavior with respect to a certain restricted set of geometric predicates is invariant.
-This family includes e.g.
-the straight-line homotopy $f_\tau(\sigma) = (1 - \tau) f_0(\sigma) + \tau f_1(\sigma)$, which was studied in the original vineyards paper~\cite{cohen2006vines}.
+This family includes
+the straight-line homotopy $F(\sigma, \tau) = (1 - \tau) f_0(\sigma) + \tau f_1(\sigma)$,   studied in the original vineyards paper~\cite{cohen2006vines}.
 Detecting all $k$ intersections of $m$ pseudo-segments is a well-studied problem in computational geometry that can be optimally solved in output-sensitive $O(m \log m + k)$ time by several algorithms~\cite{boissonnat2000efficient}, where $k$ is the output-sensitive term.
@@
  This bound is exhibited in the grayscale image data example in section~\ref{sec:motivation}: each 9x9 patch contains $(81, 208, 128)$ simplices of dimensions $(0, 1, 2)$ yielding $m=417$ simplices in total which, as the homotopy is simulated, generate $\approx 70,\!000$ transpositions.
-The upper bound size on $\lvert S_F \rvert$ is thus $\binom{417}{2} \approx 87,\!000$, which nearly occurs in this case, since the last filtration is nearly the reverse of the filtration at the beginning.
- This quadratic scaling can induce a number of issues in practical implementations of vineyards, e.g.
-although many transpositions require $0$ column reductions in practice, detecting the existence of non-zero entries at specific positions in the sparse matrix takes $O(\log(m))$ time.
+The upper bound size on $\lvert S_F \rvert$ is thus $\binom{417}{2} \approx 87,\!000$, which nearly occurs in this case, since the last filtration is almost the reverse of the filtration at the beginning.
+ This quadratic scaling can induce a number of issues in practical implementations of vineyards.
+ Indeed, although many transpositions require $0$ column reductions in practice, detecting the existence of non-zero entries at specific positions in the sparse matrix takes $O(\log(m))$ time.
 Since this check is required after each permutations are applied, it is in some sense unavoidable in the vineyards algorithm.
@@
 We call $S_F$ a \emph{schedule} with respect to $F$.
-If our goal is to decrease the size of $\lvert S_F \rvert$, one option is to \emph{coarsen} $S_F$ by collapsing contiguous sequences of adjacent transpositions to moves, via the map:
+If our goal is to decrease the size of $\lvert S_F \rvert$, one option is to \emph{coarsen} $S_F$ to a new schedule $\widetilde{S}_F$ by collapsing contiguous sequences of adjacent transpositions to moves, via the map:
 \begin{equation}\label{eq:tr_to_mv}
-	 (i, i+1)(i+1, i+2)\cdots(j-1, j) \mapsto (i, i+1, \cdots, j-1, j)  \quad \text{ if } i < j
+	 (i, i+1)(i+1, i+2)\cdots(j-1, j) \mapsto (j, i+1, \cdots, j-1, i)  \quad \text{ if } i < j
 \end{equation}
-We expect a coarsened schedule to be cheaper to execute than the original schedule it was derived from, with the tradeoff being that less diagrams are produced, which we now prove
+We expect $\widetilde{S}_F$ to be cheaper to execute than $S_F$, with the tradeoff  that less diagrams are produced.
+We make this precise next:
 \begin{proposition}\label{prop:factor2}
-	For some given decomposition $R = D V$ where each matrix is $(m \times m)$, let $C$ denote the number of $O(m)$ operations required to execute schedule $S  = (t_1, t_2, \dots, t_h)$, and similarly let $L$ be the number of $O(m)$ operations required to execute a coarsened schedule $\widetilde{S}$ of $S$.
+	Let $R = D V$ be a valid decomposition of size $(m \times m)$, let $C$ denote the number of $O(m)$ operations required to execute schedule $S$, and similarly let $L$ be the number of $O(m)$ operations required to execute a coarsened schedule $\widetilde{S}$ of $S$.
 We have the following relationship between $C$ and $L$:
@@
 \begin{proof}
-	Consider Algorithms~\ref{alg:tr} and~\ref{alg:mr}, where it's assumed Algorithms~\ref{alg:tr} is used to execute a given schedule $S$ and Algorithms~\ref{alg:mr} is used to execute its coarsened schedule $\widetilde{S}$.
+	Consider Algorithms~\ref{alg:tr} and~\ref{alg:mr}, where   Algorithm~\ref{alg:tr} is used to execute a given schedule $S$ and Algorithm~\ref{alg:mr} is used to execute its coarsened version $\widetilde{S}$.
 If maps are created providing $O(1)$ access to the low entries needed by lines (5) and (12), and the non-zero entries in each column are sorted according to some fixed order providing $O(\log(m))$ time for lines (5), (11), (17), and (20), then the dominating cost of each transposition in Algorithm~\ref{alg:tr} are the column operations, each of which take $O(m)$ time.
@@
 If we maintain a map providing low entries in $O(1)$ time, since permutations take $O(1)$ time and lines (7) and (7) in both Restore* and Move* takes just $O(1)$ via pointer swapping, the dominant cost again are the column operations (line 5).
-Since a move operation requires at most a single column operation for each index between $[i, j]$, the complexity of $\mathrm{Move}(i,j)$ is bounded above by $O(\lvert i - j \rvert m)$, and the claimed inequality follows.
+Since a move operation requires at most a single column operation for each index in $[i: j]$, the complexity of $\mathrm{Move}(i,j)$ is bounded above by $O(\lvert i - j \rvert m)$, and the claimed inequality follows.
 \end{proof}
-In terms of bounds, we clearly have $\lvert \widetilde{S}_F \rvert \leq \lvert S_F \rvert$, and the associated coarsened schedule $\widetilde{S}_F$ can be computed in $O(m)$ time, so this is a viable approach to improving the efficiency of the vineyards algorithm.
+In terms of bounds, we  have $\lvert \widetilde{S}_F \rvert \leq \lvert S_F \rvert$, and the associated coarsened schedule $\widetilde{S}_F$ can be computed in $O(m)$ time. 
 However, the coarsening depends entirely on the choice of $F$ and the upper bound from~\ref{eq:sf_m2} remains, as its always possible that there are no contiguous subsequences to collapse.
-If a straight line homotopy is used, the number of collapsible subsequences may depend on the simplices categorization as critical points; transpositions between pairs of simplices which trigger a changes in homology tend to be more expensive.
+If a straight line homotopy is used, the number of collapsible subsequences may depend on the simplices categorization as critical points; transpositions between pairs of simplices which trigger  changes in homology tend to be more expensive.
 Cohen-Steiner~\cite{cohen2006vines} referred to these as \emph{switches}, see~\cite{edelsbrunner2000topological} for more details.
@@
 \subsubsection{The Discrete Setting} 
-Since moves do not maintain a valid decomposition during movement, it is worth asking if we can achieve better bounds than vineyards, by removing the assumption that the goal is to simulate persistence across a supplied homotopy $F(x, \tau) = \mathbb{X} \times [0,1] \to \mathbb{R}$.
-In the case of vineyards, the bound is clear: whether assuming a homotopy between a pair of filtrations $(K_0, K_1)$ or not, if $\kappa = K_\tau(K_0, K_1)$, then $\Omega(\kappa)$ adjacent transpositions are required to transform $K_0 \mapsto K_1$.
+Since moves do not maintain a valid decomposition during intermediate steps, it is worth asking if  the vineyard bounds can be improved by 
+relaxing the goal of simulating persistence across  $F : K \times [0,1] \to \mathbb{R}$. 
+In the case of vineyards, the bound is clear: 
+%whether assuming a homotopy between a pair of filtrations $(K_0, K_1)$ or not, 
+if $\kappa = K_\tau(K_0, K_1)$, then $\Omega(\kappa)$ adjacent transpositions are required to transform $K_0 \mapsto K_1$. 
 In contrast, if we allow move operations in any order, we have $O(m)$ as a trivial upper bound: simply move each simplex in $K_0$ into its position in the filtration given by $K_1$ in the order given by the latter.
@@
 
- A \emph{permutation}  is a bijection of a set onto itself.
-The symmetric group $S_n$ is the set of all permutations of  $[n]$ under function composition $\circ$, applied from right to left.
-Given two fixed permutations $p, q \in S_n$ and a set $S \subseteq S_n$, a few goals that appear commonly in the literature are the following:
+Let $S_m$ denote the symmetric group  of   permutations of  $[m]$ under function composition $\circ$.
+Given two fixed permutations $p, q \in S_m$ and a set $\Sigma \subseteq S_m$,  the following 
+are common problems in the literature:
 \begin{enumerate}
-	\item Find an ordered sequence of permutations $s_1, s_2, \dots, s_d$ from $S$ whose composition transforms $p$ into $q$:
+	\item Find an ordered sequence of permutations $s_1, s_2, \dots, s_d \in \Sigma$ whose composition transforms $p$ into $q$:
 	\begin{displaymath}
@@
 	\item Find a sequence satisfying (1) of minimal length ($d$)
-	\item Find the minimal length $d$, referred to as the \emph{distance} between $p$ and $q$ with respect to $S$. 
+	\item Find the minimal length $d$, referred to as the \emph{distance} between $p$ and $q$ with respect to $\Sigma$. 
 \end{enumerate}
-A sequence $s_1, s_2, \dots, s_d \in S \subseteq S_n$ of operations mapping $p \mapsto q$ is sometimes called a \emph{sorting of $p$} in the literature.
-In the context of performing updates to the $R = D V$ decomposition, we call such sequences \emph{schedules} .
+A sequence $S = \{s_1, s_2, \dots, s_d\} \subseteq \Sigma \subseteq S_m$ of operations mapping $p \mapsto q$ is sometimes called a \emph{sorting of $p$}.
+In the context of performing updates to the $R = D V$ decomposition using moves,  such sequences will form schedules. 
 There are typically many ways to solve (1), and any solution to (2) necessarily solves (1) and (3).
-When $p, q$ are interpreted as strings, distances defined with respect to a fixed $S \subseteq S_n$ are commonly referred to as \emph{edit distances}~\cite{labarre2013lower}, which are denoted as $d_S(\cdot, \cdot)$.
-Note that here we describe distances between permutations (e.g.
-$p,q \in S_n$) using operations that can themselves be expressed as permutations (e.g.
-$s \in S \subseteq S_n$), and that the choice of $S$ defines the universe of operations and thus the type of edit distance being measured---otherwise if $S = S_n$, then $d_S(p, q) = 1$ for any $p, q \in S_n$ and the problem is trivial!
+When $p, q$ are interpreted as strings, distances defined with respect to a fixed $\Sigma \subseteq S_m$ are commonly referred to as \emph{edit distances}~\cite{labarre2013lower}, which are denoted as $d_\Sigma(\cdot, \cdot)$.
+%Note that here we describe distances between permutations (e.g. $p,q \in S_n$) using operations that can themselves be expressed as permutations (e.g. $s \in S \subseteq S_n$), and that 
+The choice of $\Sigma$ defines the universe of operations and thus the type of edit distance being measured---otherwise if $\Sigma = S_m$, then $d_\Sigma(p, q) = 1$ for any $p\neq  q \in S_m$.
 
 Many types of edit distances can be described succinctly. 
-Let $\beta(i,j,k,l)$ with $1 \leq i < j \leq k < l \leq n$ denote the permutation that exchanges the two closed intervals given by $i, j-1$ and $k,l-1$:
+Let $\beta(i,j,k,l)$ with $1 \leq i < j \leq k < l \leq m$ denote the permutation that exchanges the two closed intervals given by $i, j-1$ and $k,l-1$:
 \[
@@
 \cline{4-7} \cline{12-15}  \rule{0pt}{2ex}
-1 & \cdots & i -1 & i & i + 1 & \cdots & j - 1 & j & j +1 & \cdots & k - 1  & k & k+1 & \cdots & l - 1 & l & l + 1 & \cdots & n \\
+1 & \cdots & i -1 & i & i + 1 & \cdots & j - 1 & j & j +1 & \cdots & k - 1  & k & k+1 & \cdots & l - 1 & l & l + 1 & \cdots & m \\
 \cline{4-7} \cline{12-15} 
 \cline{4-7} \cline{12-15}  \rule{0pt}{2ex}
-1 & \cdots & i -1 & k & k+1 & \cdots & l - 1 & j & j +1 & \cdots & k - 1  & i & i + 1 & \cdots & j - 1 & l & l + 1 & \cdots & n \\
+1 & \cdots & i -1 & k & k+1 & \cdots & l - 1 & j & j +1 & \cdots & k - 1  & i & i + 1 & \cdots & j - 1 & l & l + 1 & \cdots & m \\
 \cline{4-7} \cline{12-15} 
@@
 \end{enumerate}
-Restricting the operations given above further to the case where $i = 1$ gives the so-called \emph{prefix} versions of the described distances. 
-Although each edit distance can be described using common notation, both the difficulty of obtaining a minimal sorting and the size $d_S(p,q)$ varies dramatically in the choice of $S$.
-For example, while sorting by transpositions and reversals is NP-hard and sorting by prefix transpositions is unknown, there are polynomial time algorithms for sorting by block interchanges, exchanges, and prefix exchanges~\cite{labarre2013lower}.
+%Restricting the operations given above further to the case where $i = 1$ gives the so-called \emph{prefix} versions of the described distances. 
+%Although each edit distance can be described using common notation, 
+Both the difficulty of obtaining a minimal sorting, and the size of $d_\Sigma(p,q)$, vary dramatically with the choice of $\Sigma$. 
+For example, while sorting by transpositions and reversals is NP-hard, and sorting by prefix  transpositions (i.e., when $i=1$ always) is unknown, there are polynomial time algorithms for sorting by block interchanges, exchanges, and prefix exchanges~\cite{labarre2013lower}.
 Sorting by adjacent transpositions can be achieved in many ways: any sorting algorithm that exchanges two adjacent elements during its execution (e.g.
@@
 
-Here we consider sorting by move permutations. 
-A \emph{move} on the set $[m]$ is 
-a permutation given by a pair $(i,j)$, 
-and its effect is to cyclically rotate 
-the interval $[i,j]$. For example, if $i < j$, this permutation takes the form: 
+Here we consider sorting by moves.
+In terms of permutations, 
+moving  $i$ to $j$ in 
+$[m]$, for $i < j$,
+corresponds to 
 \[
@@
 \]
-The complexity of sorting by moves is related to the well-known \emph{Levenshtein distance}~\cite{bar2004approximating}, also known as the edit distance on strings.
-The Levenshtein distance between two strings $A$ and $B$ is the minimum number of character insertions, deletions, and substitutions needed to transform one string into another.
-Through a dynamic programming type approach, the Wagner-Fischer algorithm can compute the Levenstein distance in $O(m^2)$, and there are several sub-quadratic approximations available~\cite{bar2004approximating} A related edit distance is the \emph{LCS} distance, which is the minimum number of character insertions or deletions needed to transform one set of symbols into another.
+while moving $j$ to $i$ is the inverse permutation.
+%The complexity of sorting by moves %is related to the well-known \emph{Levenshtein distance}~\cite{bar2004approximating}, which is the minimum number of character insertions, deletions, and substitutions needed to transform one string into another.
+%Through a dynamic programming type approach, the Wagner-Fischer algorithm can compute the Levenstein distance in $O(m^2)$, and there are several sub-quadratic approximations~\cite{bar2004approximating}.
+%is related to the \emph{LCS} distance, which is the minimum number of character insertions or deletions needed to transform one string into another. 
 %In terms of hardness, there is evidence to suggest that the lower bound on computing Levenshtein distance in full generality is near quadratic time. 
-
-Like the \emph{LCS} distance, sorting by move operations can be interpreted as finding a minimal sequence of edit operations where the only operations allowed are insertions and deletions.
-If these operations each take uniform cost and the universe of strings is restricted to \emph{permutations}, the corresponding edit distance has a special name called the \emph{permutation edit distance}.
+%
+%Like the \emph{LCS} distance, 
+Sorting by move operations 
+can be interpreted as finding a minimal sequence of edits  where the only operations allowed are insertions and deletions. 
+This is exactly the LCS (Longest Common Subsequence) distance. 
+If these operations each take uniform cost and the universe of strings is restricted to $S_m$, the corresponding edit distance has a special name: the \emph{permutation edit distance}.
 We obtain the following lower bound size of a sorting (i.e., a schedule) using moves.
@@
 \begin{proposition}[Schedule size]
-Given a pair of filtrations $(K_0, K_1)$ and a one-to-one correspondence $f: K_0 \to K_1$, the coarsest move schedule $S^\ast$ is of size $\Omega(d)$, where:
+Given filtrations $K_0$ and $ K_1$ of size $m$, and an order preserving bijection $f: K_0 \to K_1$, 
+then the smallest move schedule $S^\ast$ is of size $\Omega(d)$, where:
 $$ d = m - \lvert \mathrm{LCS}(K_0, K_1) \rvert $$ 
@@
     \centering
-    \includegraphics[width=0.50\textwidth]{crossings.png}
+    \includegraphics[width=0.60\textwidth]{crossings.png}
     \caption{Example of three possible schedules, $S_1$, $S_2$, and $S_3$.}
@@
 	\centering
-	\includegraphics[width=0.65\textwidth]{boid_sim_results.png}
+	\includegraphics[width=0.75\textwidth]{boid_sim_results.png}
 	\caption{ The cumulative number of $O(m)$ column operations (left, log-scale) of the three approaches tested and the distribution of the normalized Kendall distance between adjacent filtration pairs.

### 2021-09-24 move_schedules.tex
Updates  prior to moving the insertion/deletion composition into place. Included new section to clarify sorting procedure. Just need to finish revising end of Section 3 to include treaps, possibly redo pseudocode for move_left, consider whether the proof involving the constant factor should be removed or reworded, maybe add new results on bulk matrix repairing, add code links and ams classification, possibly edit the entire paper to shorten it dramatically and then it should be good to go

@@
 \end{displaymath}
+%Prior to performing the exchange, observe that that the bold entry in $V_1$ would render $V_1$ non-upper triangular after the exchange.
+This entry is removed by a left-to-right column operation, given by applying the $S_1^{2}$ on the right to $R_1$ and $V_1$.
+After this operation, the permutation may be safely applied to $V_1$.
+In this case, both before and after the permutation, $R_1$ is rendered non-reduced, requiring another column operation to restore the decomposition to a valid state.
 Starting with a valid reduction $R = DV$ and prior to performing the exchange, observe that that the highlighted entry in $V_1$ would render $V_1$ non-upper triangular after the exchange.
@@
  \begin{enumerate}[leftmargin=2\parindent, align=left, labelsep=-5pt, topsep=0pt,itemsep=-0.25ex,parsep=1.2ex]
- 	\item[M1:] Querying the non-zero status of entries in any matrix representation occurs once per move.
- 	\item[M2:] The decomposition $R = D V$ is not maintained during the movement of $\sigma_i \mapsto \sigma_j$, but only when the move is completed.
+ 	\item[M1:] Querying the non-zero status of entries in either $R$ or $V$ occurs once per move.
+ 	\item[M2:] $R = D V$ is not maintained during the movement of $\sigma_i \mapsto \sigma_j$.
  \end{enumerate} 
@@
 First, consider property M1.
-Prior to applying any permutation $P$ to the decomposition, it is necessary to remove non-zero entries in $V$ which render $PVP^T$ non-upper triangular, to maintain invariant I2.
+Prior to applying any permutation $P$ to the decomposition, it is necessary to remove non-zero entries in $V$ which render $P^TVP$ non-upper triangular, to maintain invariant I2.
 If one uses the (augmented) sparse matrix representation described in the original vineyards publication~\cite{cohen2006vines}, checking whether these entries are non-zero takes $O(\log m)$ per column.
@@
 \end{algorithm}
-
 If one has a pair of filtrations $(K_0, K_1)$ each with $m$ simplices, and we assume that each simplex $\sigma_i$ switches its relative ordering with another simplex $\sigma_j$ at most once in some ordered schedule of transpositions, then the number of intermediate persistence diagrams is bounded above by $O(m^2)$ using transpositions, whereas with move operations this bound  is at most $O(m)$.
@@
     	\Ensure Valid $R = D V$ decompositions are computed for each filtration $K_1, K_2, \dots, K_n$
-    	\Procedure{SparseSchedule}{$\mathcal{K} = ( K_1, K_2, \dots, K_n )$} 
+    	\Procedure{CoarseSchedule}{$\mathcal{K} = ( K_1, K_2, \dots, K_n )$} 
     		\State $\mathcal{S} = \emptyset$
@@
     			\State $\mathrm{lis}_q \gets \mathrm{LIS}(q)$ \Comment{$O(m \log \log m)$}
-    			\State $\mathcal{S} \gets \mathcal{S} \cup \mathrm{GreedySchedule}(p, q, \mathrm{lis}_q)$ \Comment{$O(d\log d)$, see section~\ref{sec:proxy_objective}}
+    			\State $\mathcal{S} \gets \mathcal{S} \cup \mathrm{GreedySchedule}(p, q, \mathrm{lis}_q)$ \Comment{$O(d\log m)$, see section~\ref{sec:proxy_objective}}
     		\EndFor
@@
  This bound is exhibited in the grayscale image data example in section~\ref{sec:motivation}: each 9x9 patch contains $(81, 208, 128)$ simplices of dimensions $(0, 1, 2)$ yielding $m=417$ simplices in total which, as the homotopy is simulated, generate $\approx 70,\!000$ transpositions.
-The upper bound size on $\lvert S_F \rvert$ is thus $\binom{417}{2} \approx 87,\!000$, which nearly occurs in this case, since the last filtration is almost the reverse of the filtration at the beginning.
+The upper bound size on $\lvert S_F \rvert$ is thus $\binom{417}{2} \approx 87,\!000$, which nearly occurs in this case, since the last filtration is approximately the reverse of the filtration at the beginning.
  This quadratic scaling can induce a number of issues in practical implementations of vineyards.
@@
 \begin{proof}
-	Consider Algorithms~\ref{alg:tr} and~\ref{alg:mr}, where   Algorithm~\ref{alg:tr} is used to execute a given schedule $S$ and Algorithm~\ref{alg:mr} is used to execute its coarsened version $\widetilde{S}$.
-If maps are created providing $O(1)$ access to the low entries needed by lines (5) and (12), and the non-zero entries in each column are sorted according to some fixed order providing $O(\log(m))$ time for lines (5), (11), (17), and (20), then the dominating cost of each transposition in Algorithm~\ref{alg:tr} are the column operations, each of which take $O(m)$ time.
-There are at most $2$ column operations for any given case.
-For any contiguous sequence of transpositions $(t_i, t_{i{+}1}, \dots, t_j)$, this implies we have at most $2\lvert i - j \rvert$ operations each requiring $O(m)$ time.
-Now consider a single move operation, Move($i$,$j$), as a replacement for the sequence of transpositions above.
-Identifying non-zero entries in the matrix occurs once, which takes $O(\lvert i - j \rvert \log(m))$ time.
-If we maintain a map providing low entries in $O(1)$ time, since permutations take $O(1)$ time and lines (7) and (7) in both Restore* and Move* takes just $O(1)$ via pointer swapping, the dominant cost again are the column operations (line 5).
-Since a move operation requires at most a single column operation for each index in $[i: j]$, the complexity of $\mathrm{Move}(i,j)$ is bounded above by $O(\lvert i - j \rvert m)$, and the claimed inequality follows.
+Consider executing the vineyards algorithm on a given schedule $S$.
+	%Consider Algorithms~\ref{alg:tr} and~\ref{alg:mr}, where Algorithm~\ref{alg:tr} is used to execute a given schedule $S$ and Algorithm~\ref{alg:mr} is used to execute its coarsened version $\widetilde{S}$.
+If maps are created providing $O(1)$ access to the low entries needed by lines (5) and (12), and the non-zero entries in each column are sorted according to some fixed order providing $O(\log(m))$ time for lines (5), (11), (17), and (20), then
+	%If we maintain a map locating the position of the lowest entries in $R$ in $O(1)$ time, 
+	The dominant subcomputation for the vineyards algorithm are the column operations---which take $O(m)$ time---since the permutations can be made to the matrix in $O(1)$ time and all other computations take at most $O(\log m)$ time.
+In the vineyards algorithm, there are at most $2$ column operations for a given adjacent transposition per matrix.
+	Thus, for any contiguous sequence of transpositions $(t_i, t_{i{+}1}, \dots, t_j)$ we require at most $2(\lvert i - j \rvert)$ column operations in both $R$ and $V$, giving a total of $4(\lvert i - j \rvert)$ column operations.
+
+	Now consider using the Move($i$,$j$) in Algorithm~\ref{alg:mr} for a single move operation, as a replacement for the sequence of transpositions above.
+Here again, the dominant cost again are the column operations (line 5), since although permutations take $O(\lvert i - j\rvert)$ time we require only $1$ such permutation per movement.
+	If we allocate an extra $O(m)$ memory space during the movement, notice that lines (7) and (7) in both Restore* and Move* can be completed $O(1)$ via a pointer swapping argument. 
+	Since a move operation requires at most $1$ column operation for each index in $[i: j]$, moving a column from $i$ to $j$ where $i < j$ requires at most $2(\lvert i - j \rvert)$ column operations.
+The claimed inequality follows.
+	%Identifying non-zero entries in the matrix occurs once, which takes $O(\lvert i - j \rvert \log(m))$ time.
 \end{proof}
@@
 \end{enumerate}
-%Restricting the operations given above further to the case where $i = 1$ gives the so-called \emph{prefix} versions of the described distances. 
-%Although each edit distance can be described using common notation, 
-Both the difficulty of obtaining a minimal sorting, and the size of $d_\Sigma(p,q)$, vary dramatically with the choice of $\Sigma$. 
+Both the difficulty of obtaining a minimal sorting and the size of $d_\Sigma(p,q)$ vary dramatically with the choice of $\Sigma$. 
 For example, while sorting by transpositions and reversals is NP-hard, and sorting by prefix  transpositions (i.e., when $i=1$ always) is unknown, there are polynomial time algorithms for sorting by block interchanges, exchanges, and prefix exchanges~\cite{labarre2013lower}.
@@
 Sorting by move operations 
-can be interpreted as finding a minimal sequence of edits  where the only operations allowed are insertions and deletions. 
+can be interpreted as finding a minimal sequence of edits where the only operations allowed are insertions and deletions, where each insertion/deletion operation is coupled. 
 This is exactly the LCS (Longest Common Subsequence) distance. 
@@
 Suppose we are given permutations $p = K_0, q = K_1$, and we seek to compute $\mathrm{LCS}(p, q)$.
-Consider the permutation $r = p^{-1} \circ q$.
-Since the LCS distance is a valid edit distance, if $\lvert \mathrm{LCS}(p, q) \rvert = k$, then $\lvert \mathrm{LCS}(\iota, r) \rvert = k$ as well.
-Notice that $\iota$ is strictly increasing--thus any common subsequence $\iota$ has with $r$ must also be strictly increasing.
-Thus, the problem of computing $\mathrm{LCS}(p, q)$ reduces to the problem of computing the \emph{longest increasing subsequence} (LIS) of $r$, which can done in $O( m \log \log m)$ time~\cite{bespamyatnikh2000enumerating}.
+Consider the permutation $p' = q^{-1} \circ p$.
+Since the LCS distance is a valid edit distance, if $\lvert \mathrm{LCS}(p, q) \rvert = k$, then $\lvert \mathrm{LCS}(p', \iota) \rvert = k$ as well.
+Notice that $\iota$ is strictly increasing and that any common subsequence $\iota$ has with $p'$ must also be strictly increasing.
+Thus, the problem of computing $\mathrm{LCS}(p, q)$ reduces to the problem of computing the \emph{longest increasing subsequence} (LIS) of $p'$, which can done in $O( m \log \log m)$ time using van Emde Boas trees~\cite{bespamyatnikh2000enumerating}.
 The optimality of $d$ follows from the optimality of the well-studied LCS problem~\cite{kumar1987linear}.
@@
 	The essential quantity to show the expected length of $L(p)$ over all permutations: 
-	$$ \ell_m = \mathbb{E} \, L(p) = \frac{1}{m!} \sum\limits_{p \in S_m} L(p)$$
+	$$ \mathbb{E} \, L(p) = \ell_m = \frac{1}{m!} \sum\limits_{p \in S_m} L(p)$$
 	The history of the mathematics of estimating this quantity dates back at least 50 years, and is sometimes called the \emph{Ulam-Hammersley} problem.
@@
 Thus, if $p \in S_m$ denotes a uniformly random permutation in $S_m$, then $L(p)/\sqrt{m} \to 2$ in probability as $m \to \infty$.
-Since the schedule size is defined as $d = m - \lvert \mathrm{LCS}(K_0, K_1) \rvert$, the claimed bound follows.
+Using the reduction from above to show that $\mathrm{LCS}(p,q) \Leftrightarrow \mathrm{LIS}(p')$, the claimed bound follows.
 \end{proof}
 \noindent The upper bound from Corollary~\ref{cor:expectation} captures the size of $S^\ast$ under worst-case conditions.
-It is worth noting that this bound applies to pairs of uniformly sampled permutations, as opposed to uniformly sampled distinct filtrations---intuitively, simplicial filtrations have more structure due to the subcomplex relation that must be respected in the filtration.
+It is worth noting that this bound applies to pairs of uniformly sampled permutations, as opposed to uniformly sampled distinct filtrations---intuitively, simplicial filtrations have more structure due to the subcomplex relation that must be respected in order to qualify as a filtration.
 Thus, while $O(m - \sqrt{m})$ in expectation is an improvement over $O(m)$, it is still not tight.
 However, Boissonnat~\cite{boissonnat2018efficient} has shown that for any $d$-dimensional simplicial complex $K$ with $m$ simplices, the number of distinct filtrations built from $K$ is \emph{at least} $\floor{\frac{t+1}{d+1}}^m$, where $t$ is the number of distinct filtration values associated with the complex $K$.
-In the case where $t \sim O(m)$ and $d << m$ fixed, this lower bound grows similarly to $m!$---thus, $O(m - \sqrt{m})$ is not too pessimistic of an upper bound when the pair of filtrations are truly random.
+In the case where $t \sim O(m)$ and $d << m$ fixed, this lower bound grows similarly to $m!$---thus, $O(m - \sqrt{m})$ is not too pessimistic of an upper bound when the pair of filtrations are no more similar than two random filtrations.
 In practice, when one has a time-varying filtration and the sampling points are relatively close [in time], the LCS between adjacent filtrations is expected to be much larger, subsequently making $d$ much smaller.
-For example, the average size of the LCS across the 10 filtrations from Section~\ref{sec:motivation} each having $m=417$ simplices sampled at evenly spaced time points was $343$, implying $d \approx 70$ permutations needed on average to update the decomposition between adjacent time points.
+For example, the average size of the LCS across the 10 filtrations from Section~\ref{sec:motivation} each having $m=417$ simplices sampled at evenly spaced time points was $343$, implying $d \approx 70$ permutations needed on average to update the decomposition at adjacent time points.
+
+ %% New section
+ \subsection*{Constructing a sorting}
+ In the previous section, we showed how computing the LCS between two permutations $p, q \in S_m$ reduces to computing the LIS of a single permutation $p'$, and we outlined how to do this in $O(m \log \log m)$ time.
+It is not clear, however, that given $\mathcal{L} = \mathrm{LIS}(p')$, one can obtain a sorting $p \mapsto q$ in an efficient way.
+We outline below a simple procedure which constructs such a sorting $\mathcal{S} = (s_1, \dots, s_d)$ in $O(dm\log m)$ time.
+
+ Before introducing the algorithm, we require a few definitions.
+Recall that a \emph{sorting} $\mathcal{S}$ with respect to two permutations $p, q \in S_m$ is an ordered sequence of permutations $(s_1, s_2, \dots, s_d)$ whose product under composition with $p$ yields $q$, i.e.
+ $$ q = s_d \circ \dots s_1 \circ p $$
+By definition, a subsequence $\mathcal{L}$ of symbols common to both $p$ and $q$ satisfies:
+ $$ p^{-1}(\sigma) < p^{-1}(\tau) \implies q^{-1}(\sigma) < q^{-1}(\tau) \quad \forall \sigma, \tau \in \mathcal{L} $$
+ where $p^{-1}(\sigma)$ (resp.
+$q^{-1}(\sigma)$) denotes the position of $\sigma$ in $p$ (resp.
+$q$).
+Thus, obtaining a sorting $p \mapsto q$ of size $d = m - \lvert \mathcal{L} \rvert$ reduces to applying a sequence of edit operations to symbols in the complement of $\mathcal{L}$.
+To ensure only $2d$ such operations are required, insertion operations must increase the size of $\mathcal{L}$ monotonically.
+ Suppose further we require these edit operations be coupled in the sense that for every deletion operation $\mathrm{del}_i$ of the symbol at position $i$, we compose with an insertion operation $\mathrm{ins}_j$ at position $j$.
+In this sense, we call the permutation a \emph{move} permutation if it a composition of insertion and deletion operations:
+$$ m_{ij} = (\mathrm{ins}_j \circ \mathrm{del}_i) $$
+ Similarly, we define a move permutation $s \in S_m$ as a \emph{valid} operation with respect to a fixed pair $p, q \in S_m$ if:
+ \begin{equation}\label{eq:valid_move}
+ 	\lvert \mathrm{LCS}(s \circ p, q) \rvert = \lvert \mathrm{LCS}(p,q) \rvert + 1
+ \end{equation}
+  The problem of constructing a sorting $\mathcal{S}$ of size $d$ thus reduces to the problem of choosing a sequence of $d$ valid moves.
+We call such a sorting a valid sorting.
+To do this efficiently, we need a means of choosing a permutation $s \in S_m$ that satisfies equation~\ref{eq:valid_move}.
+  Let $\mathcal{T}$ denote a ordered set-like data structure that supports the following operations on elements $\sigma \in M$ from the set $M = \{0, 1, \dots, m+1\}$:
+  \begin{enumerate}
+  	\item $\mathcal{T} \cup \sigma$---inserts $\sigma$ into $\mathcal{T}$,
+  	\item $\mathcal{T} \setminus \sigma$---removes $\sigma$ from $\mathcal{T}$,
+  	\item $\mathcal{T}_{\mathrm{succ}}(\sigma)$---obtain the successor of $\sigma$ in $\mathcal{T}$, if it exists, otherwise return $m+1$
+  	\item $\mathcal{T}_{\mathrm{pred}}(\sigma)$---obtain the predecessor of $\sigma$ in $\mathcal{T}$, if it exists, otherwise return $0$
+  \end{enumerate}
+  Given such a structure $\mathcal{T}$, an arbitrary valid sorting can be constructed by repeatedly querying and updating information about the $\mathrm{LCS}$ in $\mathcal{T}$.
+To see this, suppose $\mathcal{T}$ contains all of the symbols in the current $\mathrm{LCS}$ between two permutations $p$ and $q$.
+By definition of the LCS, we have:
+ \begin{equation}\label{eq:valid_move_ineq}
+ 	p^{-1}(\mathcal{T}_{\mathrm{pred}}(\sigma)) < p^{-1}(\sigma) < p^{-1}(\mathcal{T}_{\mathrm{succ}}(\sigma))
+ \end{equation}
+  for every $\sigma \in \mathcal{T}$.
+Now, suppose we choose a symbol $\sigma \notin \mathcal{T}$.
+If $p^{-1}(\sigma) < p^{-1}(\mathcal{T}_{\mathrm{pred}}(\sigma))$, then we must move $\sigma$ to the right in $p$ such that equation~\ref{eq:valid_move_ineq} holds.
+Similarly, if  $p^{-1}(\mathcal{T}_{\mathrm{succ}}(\sigma)) < p^{-1}(\sigma)$, then we must move $\sigma$ left in $p$ to increase the size of the LCS.
+Assuming the structure $\mathcal{T}$ supports all of the above operations in $O(\log m)$ time, we easily deduce a $O(d m\log m)$ algorithm for obtaining a valid sorting, which is shown in pseudocode in Algorithm~\ref{alg:sorting}.
+% one can incrementally build a valid sorting of size $d$ by arbitrarily choosing valid move operations . all elements in $\mathrm{LCS}(p,q)$ into $\mathcal{T}$
+%  incrementally updating the inverse permutation $p^{-1}$ of $p$. Note that   
+ % Further, let $\mathcal{I}_p$ denote an array of length $m = \lvert p \rvert$ whose values $\mathcal{I}(\sigma)$ store the position of the element $\sigma$ in $p$.
+Assume also that $\mathcal{I}_p()$
+ %for every pair of elements $\sigma, \tau \in \mathcal{L}$, if $p_{\sigma} < p_{\tau}$ where $p_{\sigma}$ denotes the the position of $\sigma$ in $p$, then $q_{\sigma} < q_{\tau}$ must also be true. 
+ %under certain insertion and deletion operations.
+
+ %We know that
+
+ \begin{algorithm}[h]
+	\caption{Sorting algorithm}\label{alg:sorting}
+	\setstretch{1.05}
+    \begin{algorithmic}[1]
+    	% \Require Fixed $p' \in S_m$, $\mathcal{L} = \mathrm{LIS}(p')$ 
+    	\Function{LCS-Sort}{$p$, $q$}
+    		\State $\bar{p} \gets q' \circ p$
+    		\State $\mathcal{L} \gets \mathrm{LCS}(p,q) = \mathrm{LIS}(\bar{p})$ \Comment{$O(\log \log m)$}
+    		\State $\mathcal{D}, \mathcal{S} \gets \bar{p} \setminus \mathcal{L}, \emptyset$
+    		\While{$\mathcal{D}$\text{ is not empty }}
+    		    \State $\sigma \gets \text{arbitrary element in }\mathcal{D}$
+    		    %\State $\tau, \rho \gets \mathcal{T}_p(\sigma), \mathcal{T}_n(\sigma)$
+    		    %\State $i, i_p, i_n \gets \mathcal{I}_p(\sigma), \, \mathcal{I}_p(T_{\mathrm{pred}}(\sigma)), \, \mathcal{I}_p(T_{\mathrm{succ}}(\sigma))$
+    		    \State $i, i_p, i_n \gets \bar{p}^{-1}(\sigma), \bar{p}^{-1}(T_{\mathrm{pred}}(\sigma)), \bar{p}^{-1}(T_{\mathrm{succ}}(\sigma))$%\Comment{$O(\log \log m)$}
+%    		    \If{ $p_{\sigma} < p_{\tau}$ }
+%	    		    \State $j \gets \text{arbitrary element in } [\, p_{\tau}, \, p_{\rho})$
+%    		    \Else{ $p_{\rho} < p_{\sigma}$}
+%    		    	\State $j \gets \text{arbitrary element in } (\, p_{\tau}, \, p_{\rho}]$
+%    		    \EndIf
+    		    \If{ $i < i_p$ } 
+    		    	\State $j \gets \text{arbitrary element in } [\, i_p, \, i_n)$
+    		    \Else{ $i_n < i$ }
+	    		    \State $j \gets \text{arbitrary element in } (\, i_p, \, i_n]$ 
+    		    \EndIf
+    		    \State $\mathcal{S}, \mathcal{D}, \mathcal{T} \gets \mathcal{S} \cup (i, \, j), \mathcal{D} \setminus \sigma, \mathcal{T} \cup \sigma$%\Comment{$O(\log \log m)$}
+    		    % \State $p, p^{-1} \gets m_{ij} \circ p, p^{-1} \circ m_{ij}^{-1}$
+    		    \State $\bar{p}^{-1} \gets \bar{p}^{-1} \circ m_{ij}^{-1}$ %\Comment{$O(m)$}
+    		 \EndWhile
+    		 \State \Return{$\mathcal{S}$}
+    	\EndFunction
+	\end{algorithmic}
+\end{algorithm}
 
@@
 Any sequence of moves which increases the LCS on each move is guaranteed to have size $m - \lvert \mathrm{LCS}(K_0, K_1) \rvert$, and the reduction to the permutation edit distance problem ensures this size is optimal.
-However, like in the vineyards algorithm, certain pairs of simplices cost more to exchange depending on whether they are critical pairs in the sense described~\cite{cohen2006vines}, making specific move operations more expensive than others.
-This small variation in cost-per-transposition induces a large variability in the cost of executing any given schedule, which is undesirable computationally.
-Ideally, we would like to generate a schedule which not only is minimal in size, but also cost.
+However, like in the vineyards algorithm, certain pairs of simplices cost more to exchange depending on whether they are critical pairs in the sense described~\cite{cohen2006vines}. 
+This small variation in cost-per-transposition induces a large variability in the cost of randomly generated schedules, which is computationally undesirable in practice.
+Ideally, we would like to generate a schedule which not only is minimal in size, but is also minimal in terms of the column operations it requires.
 Formally, given two permutations $p, q \in S_m$, let $S = ( s_1, s_2, \dots, s_d )$ denote a sorting such that $S \circ p = q$.
@@
 \end{alignat*}
-The cost of each successive move operation for each of these schedules is recorded in Table~\ref{table:move_costs}.
+where the notation $m_{x y}$ represents the move permutation that moves symbol $x$ to the position of symbol $y$.
+The cost of each  move operation and each schedule is recorded in Table~\ref{table:move_costs}.
 \begin{table}[h]
@@
 While the cheapest schedule $S_1$ is in this candidate set, so is $S_6$, the most expensive schedule.
-Moreover, a tie-breaker that keeps track of previous cheapest moves would end up picking either schedule $S_2$ or $S_5$ as the final greedy schedule, both of which are suboptimal cost-wise.
-\\
-\\
-\noindent
+As a result, a greedy strategy which chooses the lowest-cost move may not yield an optimal schedule.
+%Moreover, any reasonable tie-breaker that keeps track of previous cheapest moves would end up picking either schedule $S_2$ or $S_5$ as the final greedy schedule, both of which are suboptimal cost-wise.
+
 Nonetheless, greedy strategies can be effective in practice, even if they are suboptimal.
-In the previous example, since evaluating the cost of move is $\approx O(m)$, a greedy strategy that chooses the least expensive move would require $O(d^2 m)$ time, which is computationally viable when $d << m$.
+For example, in the previous section choosing the lowest-cost move takes $\approx O(dm)$---implying a greedy strategy would require $O(d^2 m)$ time---which is computationally viable as a heuristic when $d << m$.
 
@@
 When $d \sim O(m)$, the quadratic nature of the aforementioned greedy minimization is too expensive.
-In order for Algorithm~\ref{alg:schedule} to be considered as an efficient alternative to vineyards---which requires $O(m \log m + k)$ preprocessing time to compute $k$ transpositions---a solution which is subquadratic in $m$ is desirable.
+In order for Algorithm~\ref{alg:schedule} to be considered as an efficient alternative to vineyards---which requires $O(m \log m + k)$ preprocessing time to compute $k \sim O(m^2)$ transpositions---a solution which is subquadratic in $m$ is desirable.
 %In what follows, we present such an objective, showing its relation to the problem at hand as well its use in other known computational problems.
@@
 %where the set and order of pairs $\hat{s}_\ast$ depend on $S(\sigma)$. 
-% In the vineyards algorithm, permutations take $O(1)$ time if a special sparse matrix structure is used~\cite{cohen2006vines}; others permutations $PRP^T = (PDP^T)(PVP^T)$ are non-trivial operations computationally.
+% In the vineyards algorithm, permutations take $O(1)$ time if a special sparse matrix structure is used~\cite{cohen2006vines}; others permutations $P^TRP = (P^TDP)(P^TVP)$ are non-trivial operations computationally.
 % \begin{displaymath}
@@
 where now $\mathrm{cost}(\cdot)$ is to be replaced by a proxy objective function.
-There are exponentially many ways to choose $\sigma \in S_d$, so we opt for a greedy strategy as before, which takes $O(d \, C(m))$ time where $C(m)$ denotes the cost of evaluating $\mathrm{cost}(\cdot)$.
+There are exponentially many ways to choose $\sigma \in S_d$, so we opt for a greedy strategy as before, which takes $O(d \, C(m))$ time where $C(m)$ denotes the cost of evaluating the greedy step.
 For this minimization to be subquadratic in $m$, we need $C(m) \sim O(\log m)$ at each of the $d$ steps of the greedy optimization, and for the minimization to yield an efficient solution in practice, $\mathrm{cost}(\cdot)$ must be related to the problem of minimizing move costs.
@@
 In practice, due to the sparsity of both $R$ and $V$, this is a very loose upper bound---the number of column operations is generally much smaller.
-Nonetheless, the complexities of the move operations discussed in Section~\ref{sec:moves} are dependent on the magnitude of $\lvert i - j \rvert$, and thus it's worth considering a proxy measure.
+Nonetheless, the complexities of the move operations discussed in Section~\ref{sec:moves} are dependent on the magnitude of $\lvert i - j \rvert$, and thus it's worth considering as a proxy objective.
 If one forms bijections between two given filtrations $p: K_0 \to [m], q: K_1 \to [m]$ to permutations of $[m]$ as before, then an upper bound on the number of column reductions needed to perform any given schedule is given the \emph{Spearman distance}, a common distance for measuring the disarrangement between permutations.
@@
 That is, unlike the Kendall distance, the Spearman distance satisfies:
-$$ F(p, q) \geq F(s_1 \circ p, q) \geq F( s_2 \circ s_1 \circ p, q) \geq F( s_d \circ \dots \circ s_2 \circ s_1 \circ p, q)  = F(q,q) = 0 $$
+$$ F(p, q) \geq F(s_1 \circ p, q) \geq F( s_2 \circ s_1 \circ p, q) \geq F( s_d \circ \dots \circ s_1 \circ p, q)  = F(q,q) = 0 $$
 For any sorting $S$ such that $s_{i}$ that increases the size of longest common subsequence between $s_{i-1} \circ \dots \circ s_1 \circ p$ and $q$.
@@
 \begin{equation}\label{eq:kendall_inequality}
-F(p, q) = F(p, S(\sigma) \circ p) \leq \sum\limits_{i=1}^{d-1} F_\sigma(\hat{s}_i, \hat{s}_{i+1})
+F(p, q) = F(p, S(\sigma) \circ p) \leq \sum\limits_{i=1}^{d-1} F(\hat{s}_i, \hat{s}_{i+1})
 \end{equation}
@@
 \end{equation}
-The minimizer of this optimization can be interpreted as a sorting which minimizes the net displacement amongst all possible sortings $p \mapsto q$, in the $\ell_1$ sense.
-It is related to a constrained version of the crossing minimization problem for $k$ sets of permutations, which is NP-hard for $k \geq 4$~\cite{biedl2009complexity}, see below for an example.
+The minimizer of this optimization can be interpreted as a sorting which minimizes the net displacement amongst all possible $d$-sized sortings $p \mapsto q$, again in the $\ell_1$ sense.
+When one does not need the sortings to be valid, this problem reduces to the $k$-layer crossing minimization problem for $k$ sets of permutations, which is NP-hard for $k \geq 4$~\cite{biedl2009complexity}.
 As a result, we do not expect a greedy solution to yield the optimal solution exactly, but simply a fast approximation.
@@
 It is known that one can apply constant-factor updates to multiple values in a prefix sum in $O(\log m)$ time by any prefix-sum data structure that also supports \emph{range updates}, e.g.
-a Fenwick tree or a Segment Tree that supports lazy propagation~\cite{de1997computational}.
-Single element prefix sum modifications can also be achieved $O(\log m)$ time with these data structures, the time to compute $F(\hat{s}_i, \hat{s}_{i+1})$ is bounded above by $O(\log m)$ and thus the greedy solution to equation~\ref{eq:schedule_opt_succ} can be computed in $O(d \log m)$ time.
+a Fenwick tree, or a Segment Tree that supports lazy propagation~\cite{de1997computational}.
+Since single element prefix sum modifications can also be achieved $O(\log m)$ time with these data structures, the time to compute $F(\hat{s}_i, \hat{s}_{i+1})$ is bounded above by $O(\log m)$, thus the greedy solution to equation~\ref{eq:schedule_opt_succ} can be computed in $O(d \log m)$ time.
 %We give an example and a more detailed discussion for those interested in the appendix\ref{sec:appendix}. 
@@
 %However, unlike the permutation edit distance, the edit operations for the vineyards problem have non-uniform cost and are time-varying---each move $\mathrm{Move}(i,j)$ requires a number of column operations at most proportional to $\lvert i - j\rvert$, and this cost changes as the decomposition is modified.
-Ideally, we want an ordered set of move operations that permutes the given decomposition $R = \partial V$ into the target one $PRP^T=(P \partial P^T)(P V P^T)$ as efficiently as possible.
+Ideally, we want an ordered set of move operations that permutes the given decomposition $R = \partial V$ into the target one $P^TRP=(P \partial P^T)(P^T V P)$ as efficiently as possible.
 In what follows, we outline various roadblocks that prevent us from doing this directly, concluding with a heuristic and greedy procedure that works well in practice.
@@
 %%Let $s$ be some fixed symbol to move.
-A move operation $m_s$ is said to be \emph{free} if its corresponding permutation matrix $P$ induces no column operations in the decomposition $(P\partial P^T)(P V P^T) = PRP^T$.
+A move operation $m_s$ is said to be \emph{free} if its corresponding permutation matrix $P$ induces no column operations in the decomposition $(P\partial P^T)(P^T V P) = P^TRP$.
 When is a move free?

### 2021-09-28 move_schedules.tex
Last updates before overhaul of last part of section 3

@@
   		v \kern\rowsp & \;     &  1 & 1 \; \\
-  		w \kern\rowsp & \; 1 &  1 &    \; \botstrut \\
+  		w \kern\rowsp & \; 1 &  \fbox{1} &    \; \botstrut \\
 		\end{block}
@@
   		u \kern\rowsp  & \; 1 & 1 & 1 \; \topstrut \\
-  		v \kern\rowsp & \;     & 1 & 1 \; \\
+  		v \kern\rowsp & \;     & 1 & \fbox{1} \; \\
   		w \kern\rowsp & \; 1 &    &    \; \botstrut \\
@@
 The original purpose of the vineyards algorithm, as described in~\cite{cohen2006vines}, was to compute a continuous 1-parameter family of persistence diagrams over a time-varying filtration, and to detect homological critical events during the construction.
-Notice that homological critical events can only occur when the filtration order changes, though not all filtration changes result in homological critical events.
-It follows that detecting these homological events reduces to computing matrix decompositions at a finite set of time points interleaved between filtration changes. 
-At the finest scale, changes to the filtration order manifest as adjacent transpositions. 
+Notice that homological critical events can only occur when the filtration order changes, though the converse isn't true.
+Thus, detecting these all such events reduces to computing matrix decompositions at a finite set of time points interleaved between changes in the filtration order. 
+At the finest scale, these changes manifest as adjacent transpositions. 
 As a result, a fixed set of rules for maintaining a valid $R = D V$ decomposition under the transposition of adjacent simplices in the filtration is enough to simulate persistence dynamically. 
@@
   		 & \;  & 1 &  \; \\
-  		 & \; 1 & \textbf{1} & \; \botstrut \\
+  		 & \; 1 & 1 & \; \botstrut \\
 		\end{block}
@@
   		& \; 1 &  &  \; \\
-  		& \; 1 & \textbf{1} &  \; \botstrut \\
+  		& \; 1 & \fbox{1} &  \; \botstrut \\
 		\end{block}
@@
 	\hspace{-0.1em}
-	\begin{blockarray}{cccc}{R_1}
+	\begin{blockarray}{cccc}{}
 	\kern-0.2em & b & a & c  \\
@@
 	\,\,,
-	\begin{blockarray}{cccc}{V_1}
+	\begin{blockarray}{cccc}{}
 	\kern-0.2em & b & a & c  \\
 		\begin{block}{c[ccc]}
-  		b \kern\rowsp  & \; 1 & 1 & 1 \; \topstrut \\
-  		a \kern\rowsp & \;  & 1 & 1 \; \\
-  		c \kern\rowsp & \;  &  & 1 \; \botstrut \\
+  		  & \; 1 & 1 & 1 \; \topstrut \\
+  		 & \; & 1 & 1 \; \\
+  		 & \;  &  & 1 \; \botstrut \\
 		\end{block}
@@
   		  & \;    & 1 &  \; \\
-  		 & \; 1 & 1 &  \; \botstrut \\
+  		 & \; 1 & \fbox{1} &  \; \botstrut \\
 		\end{block}
@@
   		 & \;       &  1  &  1  \; \\
-  		 & \; 1   &   1   &     \; \botstrut \\
+  		 & \; 1   &   \fbox{1}   &     \; \botstrut \\
 		\end{block}
@@
   		 & \; 1 & 1 &  \; \\
-  		 & \; 1 &  & 1 \; \botstrut \\
+  		 & \; 1 &  & \fbox{1} \; \botstrut \\
 		\end{block}
@@
 		\begin{block}{c[ccc]}
-  		 & \; 1 & 1 &  1\; \topstrut \\
+  		 & \; 1 & \fbox{1} &  \fbox{1}\; \topstrut \\
   		& \;  & 1 &  1\; \\
@@
 		\begin{block}{c[ccc]}
-  		 & \; 1 & &  1\; \topstrut \\
+  		 & \; 1 & &  \fbox{1}\; \topstrut \\
   		& \;  & 1 &  1\; \\
@@
 \begin{blockarray}{cc}{}
-	\kern-0.2em & c \\
+	\kern-0.2em & a \\
 		\begin{block}{c[c]}
@@
 
-	Now consider using the Move($i$,$j$) in Algorithm~\ref{alg:mr} for a single move operation, as a replacement for the sequence of transpositions above.
+	Now consider using the MoveRight($i$,$j$) for a single move operation, as a replacement for the sequence of transpositions above.
 Here again, the dominant cost again are the column operations (line 5), since although permutations take $O(\lvert i - j\rvert)$ time we require only $1$ such permutation per movement.
-	If we allocate an extra $O(m)$ memory space during the movement, notice that lines (7) and (7) in both Restore* and Move* can be completed $O(1)$ via a pointer swapping argument. 
+	If we allocate an extra $O(m)$ memory space during the movement, notice that lines (4), and (7) of Restore* (but not line (2)) and line (7) of Move* in Algorithm~\ref{alg:mr} can be completed $O(1)$ via a pointer swapping argument.
 	Since a move operation requires at most $1$ column operation for each index in $[i: j]$, moving a column from $i$ to $j$ where $i < j$ requires at most $2(\lvert i - j \rvert)$ column operations.
@@
 We outline below a simple procedure which constructs such a sorting $\mathcal{S} = (s_1, \dots, s_d)$ in $O(dm\log m)$ time.
+
+ \algnewcommand{\IfThenElse}[3]{% \IfThenElse{<if>}{<then>}{<else>}
+  \State \algorithmicif\ #1\ \algorithmicthen\ #2\ \algorithmicelse\ #3}
+
+  \begin{algorithm}[h]
+	\caption{Sorting algorithm}\label{alg:sorting}
+	\setstretch{1.05}
+    \begin{algorithmic}[1]
+    	% \Require Fixed $p' \in S_m$, $\mathcal{L} = \mathrm{LIS}(p')$ 
+    	\Function{LCS-Sort}{$p$, $q$}
+    		\State $\bar{p} \gets q' \circ p$
+    		\State $\mathcal{L} \gets \mathrm{LCS}(p,q) = \mathrm{LIS}(\bar{p})$ \Comment{$O(\log \log m)$}
+    		\State $(\, \mathcal{S}, \, \mathcal{D}, \, \mathcal{T} \, )\gets (\, \emptyset, \, \bar{p} \setminus \mathcal{L}, \, \mathcal{L} \,)$
+    		\While{$\mathcal{D}$\text{ is not empty }}
+    		    \State $\sigma \gets \text{arbitrary element in }\mathcal{D}$
+    		    %\State $\tau, \rho \gets \mathcal{T}_p(\sigma), \mathcal{T}_n(\sigma)$
+    		    %\State $i, i_p, i_n \gets \mathcal{I}_p(\sigma), \, \mathcal{I}_p(T_{\mathrm{pred}}(\sigma)), \, \mathcal{I}_p(T_{\mathrm{succ}}(\sigma))$
+    		    \State $(\,i, \,i_p, \, i_n \,) \gets ( \, \bar{p}^{-1}(\sigma), \, \bar{p}^{-1}(T_{\mathrm{pred}}(\sigma)), \, \bar{p}^{-1}(T_{\mathrm{succ}}(\sigma)) \, )$%\Comment{$O(\log \log m)$}
+%    		    \If{ $p_{\sigma} < p_{\tau}$ }
+%	    		    \State $j \gets \text{arbitrary element in } [\, p_{\tau}, \, p_{\rho})$
+%    		    \Else{ $p_{\rho} < p_{\sigma}$}
+%    		    	\State $j \gets \text{arbitrary element in } (\, p_{\tau}, \, p_{\rho}]$
+%    		    \EndIf
+				% \IfThenElse{$i < i_p$}{$[\, i_p, \, i_n)$}{$(\, i_p, \, i_n]$}
+				% $\text{Pick } j \in [\, i_p, \, i_n)$ \If{ $i < i_p$ } \Else $(\, i_p, \, i_n]$ \EndIf
+    		    \If{ $i < i_p$ } 
+    		    	\State $j \gets \text{arbitrary element in } [\, i_p, \, i_n)$
+    		    \Else{ $i_n < i$ }
+	    		    \State $j \gets \text{arbitrary element in } (\, i_p, \, i_n]$ 
+    		    \EndIf
+    		    \State $(\, \mathcal{S}, \mathcal{D}, \mathcal{T} \,) \gets (\, \mathcal{S} \cup (i, \, j), \, \mathcal{D} \setminus \sigma, \, \mathcal{T} \cup \sigma \, )$%\Comment{$O(\log \log m)$}
+    		    % \State $p, p^{-1} \gets m_{ij} \circ p, p^{-1} \circ m_{ij}^{-1}$
+    		    \State $\bar{p}^{-1} \gets \bar{p}^{-1} \circ m_{ij}^{-1}$ %\Comment{$O(m)$}
+    		 \EndWhile
+    		 \State \Return{$\mathcal{S}$}
+    	\EndFunction
+	\end{algorithmic}
+\end{algorithm}
 
@@
  %We know that
-
- \begin{algorithm}[h]
-	\caption{Sorting algorithm}\label{alg:sorting}
-	\setstretch{1.05}
-    \begin{algorithmic}[1]
-    	% \Require Fixed $p' \in S_m$, $\mathcal{L} = \mathrm{LIS}(p')$ 
-    	\Function{LCS-Sort}{$p$, $q$}
-    		\State $\bar{p} \gets q' \circ p$
-    		\State $\mathcal{L} \gets \mathrm{LCS}(p,q) = \mathrm{LIS}(\bar{p})$ \Comment{$O(\log \log m)$}
-    		\State $\mathcal{D}, \mathcal{S} \gets \bar{p} \setminus \mathcal{L}, \emptyset$
-    		\While{$\mathcal{D}$\text{ is not empty }}
-    		    \State $\sigma \gets \text{arbitrary element in }\mathcal{D}$
-    		    %\State $\tau, \rho \gets \mathcal{T}_p(\sigma), \mathcal{T}_n(\sigma)$
-    		    %\State $i, i_p, i_n \gets \mathcal{I}_p(\sigma), \, \mathcal{I}_p(T_{\mathrm{pred}}(\sigma)), \, \mathcal{I}_p(T_{\mathrm{succ}}(\sigma))$
-    		    \State $i, i_p, i_n \gets \bar{p}^{-1}(\sigma), \bar{p}^{-1}(T_{\mathrm{pred}}(\sigma)), \bar{p}^{-1}(T_{\mathrm{succ}}(\sigma))$%\Comment{$O(\log \log m)$}
-%    		    \If{ $p_{\sigma} < p_{\tau}$ }
-%	    		    \State $j \gets \text{arbitrary element in } [\, p_{\tau}, \, p_{\rho})$
-%    		    \Else{ $p_{\rho} < p_{\sigma}$}
-%    		    	\State $j \gets \text{arbitrary element in } (\, p_{\tau}, \, p_{\rho}]$
-%    		    \EndIf
-    		    \If{ $i < i_p$ } 
-    		    	\State $j \gets \text{arbitrary element in } [\, i_p, \, i_n)$
-    		    \Else{ $i_n < i$ }
-	    		    \State $j \gets \text{arbitrary element in } (\, i_p, \, i_n]$ 
-    		    \EndIf
-    		    \State $\mathcal{S}, \mathcal{D}, \mathcal{T} \gets \mathcal{S} \cup (i, \, j), \mathcal{D} \setminus \sigma, \mathcal{T} \cup \sigma$%\Comment{$O(\log \log m)$}
-    		    % \State $p, p^{-1} \gets m_{ij} \circ p, p^{-1} \circ m_{ij}^{-1}$
-    		    \State $\bar{p}^{-1} \gets \bar{p}^{-1} \circ m_{ij}^{-1}$ %\Comment{$O(m)$}
-    		 \EndWhile
-    		 \State \Return{$\mathcal{S}$}
-    	\EndFunction
-	\end{algorithmic}
-\end{algorithm}
 

### 2021-09-28 move_schedules.tex
Much better reduction of lcs edit distance -> permutation edit distance

@@
   As one can see from the figure, if one needs to simulate the homotopy exactly, then the vineyards approach is indeed far more efficient than naively applying the reduction algorithm independently at all time points.
-  However, when the discretization of the time domain is sparse enough, the naive approach actually performs less column operations than the vineyards strategy, while still capturing the main events.
+  However, when the discretization of the time domain is coarse enough, the naive approach actually performs less column operations than the vineyards strategy, while still capturing the main events.
 
@@
 \subsubsection{The Discrete Setting} 
-Since moves do not maintain a valid decomposition during intermediate steps, it is worth asking if  the vineyard bounds can be improved by 
-relaxing the goal of simulating persistence across  $F : K \times [0,1] \to \mathbb{R}$. 
+At the finest scale, using move operations simulate persistence across $F : K \times [0,1] \to \mathbb{R}$ reduces to vineyards (i.e. when the moves themselves are adjacent transpositions).
+The benefit of move operations is that one need not maintain a valid decomposition during the movement; thus, it is worth investigating whether the vineyard bounds can be improved by relaxing the assumption that one wants to simulate persistence across $F$ continuously.
 In the case of vineyards, the bound is clear: 
-%whether assuming a homotopy between a pair of filtrations $(K_0, K_1)$ or not, 
 if $\kappa = K_\tau(K_0, K_1)$, then $\Omega(\kappa)$ adjacent transpositions are required to transform $K_0 \mapsto K_1$. 
@@
 There are typically many ways to solve (1), and any solution to (2) necessarily solves (1) and (3).
-When $p, q$ are interpreted as strings, distances defined with respect to a fixed $\Sigma \subseteq S_m$ are commonly referred to as \emph{edit distances}~\cite{labarre2013lower}, which are denoted as $d_\Sigma(\cdot, \cdot)$.
+When $p, q$ are interpreted as strings, distances defined with respect to a fixed $\Sigma \subseteq S_m$ are commonly referred to as \emph{edit distances}~\cite{labarre2013lower}, which are denoted as $d_\Sigma(\cdot, \cdot)$, and the subsequent elementary operations on the strings are called \emph{edit operations}.
 %Note that here we describe distances between permutations (e.g. $p,q \in S_n$) using operations that can themselves be expressed as permutations (e.g. $s \in S \subseteq S_n$), and that 
@@
 
-Here we consider sorting by moves.
-In terms of permutations, 
-moving  $i$ to $j$ in 
-$[m]$, for $i < j$,
-corresponds to 
+Here we consider sorting by moves. In terms of permutations, a permutation $m_{ij}$ that moves $i$ to $j$ in $[m]$, for $i < j$, corresponds to: 
 \[
@@
 \]
-while moving $j$ to $i$ is the inverse permutation.
+In the context of edit operations, observe that a move operation can be interpreted the composition of deletion operation followed by an insertion operation: 
+$$ m_{ij} = (\mathrm{ins}_j \circ \mathrm{del}_i) $$
+where $\mathrm{del}_i$ denotes the edit operation that deletes the character at position $i$ in some string $p$ and $\mathrm{ins}_j$ similarly inserts the same character at position $j$.
+Thus, sorting by move operations can be interpreted as finding a minimal sequence of edits where the only operations allowed are insertions and deletions, where each insertion/deletion operation is coupled.
+This is exactly the LCS (Longest Common Subsequence) distance. It is well known~\cite{bergroth2000survey} that the LCS distance between two strings $p, q$ of sizes $m$ and $n$, respectively, is: 
+\begin{equation}
+	\mathrm{d}_{\mathrm{lcs}}(p,q) = m + n - 2\lvert LCS(p, q)\rvert
+\end{equation}
+Intuitively, in the context of strings, given $\mathcal{L} = \mathrm{LCS}(p,q)$ one may transform $p \mapsto q$ by simply deleting the symbols from $p \setminus \mathcal{L}$ and then inserting symbols from $q \setminus \mathcal{L}$.
+In the context of permutations, these operations are coupled, thus to ensure only $2d$ such edit operations are used to transform $p \mapsto q$, each move operations must increase the size of $\mathcal{L}$ monotonically.
+% while moving $j$ to $i$ is the inverse permutation.
 %The complexity of sorting by moves %is related to the well-known \emph{Levenshtein distance}~\cite{bar2004approximating}, which is the minimum number of character insertions, deletions, and substitutions needed to transform one string into another.
@@
 %Like the \emph{LCS} distance, 
-Sorting by move operations 
-can be interpreted as finding a minimal sequence of edits where the only operations allowed are insertions and deletions, where each insertion/deletion operation is coupled. 
-This is exactly the LCS (Longest Common Subsequence) distance. 
-If these operations each take uniform cost and the universe of strings is restricted to $S_m$, the corresponding edit distance has a special name: the \emph{permutation edit distance}.
-We obtain the following lower bound size of a sorting (i.e., a schedule) using moves.
-
+Thus, by reducing the problem to the \emph{permutation edit distance}, we obtain the following lower bound on the minimum size of a sorting (i.e., a schedule) using moves. 
 \begin{proposition}[Schedule size]
@@
 $q$).
-Thus, obtaining a sorting $p \mapsto q$ of size $d = m - \lvert \mathcal{L} \rvert$ reduces to applying a sequence of edit operations to symbols in the complement of $\mathcal{L}$.
-To ensure only $2d$ such operations are required, insertion operations must increase the size of $\mathcal{L}$ monotonically.
- Suppose further we require these edit operations be coupled in the sense that for every deletion operation $\mathrm{del}_i$ of the symbol at position $i$, we compose with an insertion operation $\mathrm{ins}_j$ at position $j$.
-In this sense, we call the permutation a \emph{move} permutation if it a composition of insertion and deletion operations:
-$$ m_{ij} = (\mathrm{ins}_j \circ \mathrm{del}_i) $$
- Similarly, we define a move permutation $s \in S_m$ as a \emph{valid} operation with respect to a fixed pair $p, q \in S_m$ if:
+Thus, obtaining a sorting $p \mapsto q$ of size $d = m - \lvert \mathcal{L} \rvert$ reduces to applying a sequence of move operations to symbols in the complement of $\mathcal{L}$.
+ Formally, we define a move permutation $s \in S_m$ as a \emph{valid} operation with respect to a fixed pair $p, q \in S_m$ if:
  \begin{equation}\label{eq:valid_move}
@@
 \subsubsection{Proxy objective}\label{sec:proxy_objective}
-When $d \sim O(m)$, the quadratic nature of the aforementioned greedy minimization is too expensive.
+When $d \sim O(m)$, the quadratic nature of the aforementioned greedy minimization can be computationally expensive. 
 In order for Algorithm~\ref{alg:schedule} to be considered as an efficient alternative to vineyards---which requires $O(m \log m + k)$ preprocessing time to compute $k \sim O(m^2)$ transpositions---a solution which is subquadratic in $m$ is desirable.
@@
 \section{Conclusion and Future Work}\label{sec:conclusion}
-In conclusion, we presented a scheduling algorithm for efficiently updating a decomposition in sparse dynamic settings.
+In conclusion, we presented a scheduling algorithm for efficiently updating a decomposition in coarse dynamic settings.
 Our approach is simple, easy to implement, and fully general: it does not depend on the geometry of underlying space nor the choice of triangulation.

### 2021-09-28 move_schedules.tex
First full revision of section 3.3!

@@
 Ideally, we would like to generate a schedule which not only is minimal in size, but is also minimal in terms of the column operations it requires.
-Formally, given two permutations $p, q \in S_m$, let $S = ( s_1, s_2, \dots, s_d )$ denote a sorting such that $S \circ p = q$.
+%Formally, given two permutations $p, q \in S_m$, let $S = ( s_1, s_2, \dots, s_d )$ denote a sorting such that $S \circ p = q$. 
 %The example given in section~\ref{sec:greedy} shows all (minimal size) $d!$ possible choices of $S$ for a fixed pair of permutations and choice of $\mathrm{LCS}(p,q)$. 
-Let $\nu = \mathrm{LCS}(p,q)$ and let $\sigma \in S_d$ denote a choice of ordering for the symbols in  $p \smallsetminus \nu$, such that $S(\sigma)$ denotes a schedule generated from this ordering.
+%Let $\nu = \mathrm{LCS}(p,q)$ and let $\sigma \in S_d$ denote a choice of ordering for the symbols in  $p \smallsetminus \nu$, such that $S(\sigma)$ denotes a schedule generated from this ordering.
 Rephrasing from the description above, we would like to achieve something similar to the following objective:
-\begin{equation}\label{eq:schedule_opt}
-	S_\sigma^\ast = \argmin_{\sigma \in S_d} \mathrm{cost}(S(\sigma) \circ p)
-\end{equation}
-where $\mathrm{cost}(S(\sigma) \circ p)$ measures the number of column operations require to execute $S(\sigma)$ on an existing decomposition. 
+%\begin{equation}\label{eq:schedule_opt}
+%	S_\sigma^\ast = \argmin_{\sigma \in S_d} \mathrm{cost}(S(\sigma) \circ p)
+%\end{equation}
+%where $\mathrm{cost}(S(\sigma) \circ p)$ measures the number of column operations require to execute $S(\sigma)$ on an existing decomposition. 
 %There are exponentially many ways to generate such a sorting $S$. 
 % Dynamic programming discussion
-One advantage of Algorithm~\ref{alg:mr} is that the cost of applying a permutation to some given $R = D V$ decomposition can be determined efficiently prior to performing any column operations. 
-Specifically, the cost of determining $\lvert \mathbb{I}\rvert + \lvert \mathbb{J}\rvert$ in lines (2) and (3) of Algorithm~\ref{alg:mr} cannot take more than $O(m)$ time.
-If we assume a similar complexity to determine the cost of applying a permutation with Algorithm~\ref{alg:ml} and we know to execute $d$ moves, one might expect be to use an optimization method like dynamic programming.
-That is, suppose $O(dm)$ move costs are computed ahead of time and these costs are stored in a table providing $O(1)$ element access and modification.
+
+%If we assume a similar complexity to determine the cost of applying a permutation with Algorithm~\ref{alg:ml} and we know to execute $d$ moves, one might expect be to use an optimization method like dynamic programming.
+% That is, suppose $O(dm)$ move costs are computed ahead of time and these costs are stored in a table providing $O(1)$ element access and modification.
 Using a sparse row representation for $V$, this cannot take more than $O(dm)$ time, which is relatively efficient when $d << m$.
@@
 
-\subsubsection{Greedy approach}\label{sec:greedy} If an optimization procedure like dynamic programming is too expensive, one would hope that a greedy solution which chooses the minimal cost choice at every step could lead be an efficient strategy.
+\subsubsection{Greedy approach}\label{sec:greedy} 
+One advantage of using move operations is that the cost of applying a permutation to some given $R = D V$ decomposition can be determined efficiently prior to performing any column operations. 
+Specifically, the cost of determining $\lvert \mathbb{I}\rvert + \lvert \mathbb{J}\rvert$ in lines (2) and (3) of Algorithm~\ref{alg:mr} cannot take more than $O(m)$ time. 
+Towards minimizing the number of column operations a schedule requires, it is natural to consider whether a greedy-type solution which chooses the minimal cost choice at every step yields an efficient strategy.
 We give a counter-example below demonstrating that a greedy procedure may lead to arbitrarily bad behavior.
@@
 
-Nonetheless, greedy strategies can be effective in practice, even if they are suboptimal.
+\subsubsection{Proxy objective}\label{sec:proxy_objective}
+Although the greedy approach from the last section can lead to arbitrarily high cost schedules, similar greedy-like strategies can yield computationally efficient heuristics in practice, even if they are suboptimal.
 For example, in the previous section choosing the lowest-cost move takes $\approx O(dm)$---implying a greedy strategy would require $O(d^2 m)$ time---which is computationally viable as a heuristic when $d << m$.
-
-\subsubsection{Proxy objective}\label{sec:proxy_objective}
-When $d \sim O(m)$, the quadratic nature of the aforementioned greedy minimization can be computationally expensive. 
-In order for Algorithm~\ref{alg:schedule} to be considered as an efficient alternative to vineyards---which requires $O(m \log m + k)$ preprocessing time to compute $k \sim O(m^2)$ transpositions---a solution which is subquadratic in $m$ is desirable.
+%Since the persistence algorithm requires $\approx O(m^3)$, the schedule generation procedure must at most have $O(m^2)$ complexity in its operation. 
+However, when $d \sim O(m)$, the sorting procedure shown using Algorithm~\ref{alg:sorting} approaches $O(m^2 \log m)$.
+We seek a fast procedure for generating schedules; if the schedule generation procedure exhibits a complexity of $\approx O(m^3)$ or higher, then such a procedure would offers no computational benefit compared to naively reduction algorithm.
+Indeed, observe the vineyards algorithm requires $O(m \log m + k)$ preprocessing time to compute a schedule of size $k \sim O(m^2)$ transpositions, thus for Algorithm~\ref{alg:schedule} to be considered a more efficient alternative to vineyards, a solution which is subquadratic in $m$ is desirable.
 %In what follows, we present such an objective, showing its relation to the problem at hand as well its use in other known computational problems. 
@@
 %Let $p,q \in S^n$ be fixed permutations of $[n]$ and let $S = (s_1, \dots, s_d)$ denote a sorting $p \mapsto q$ such that $q = S \circ p = s_d \circ \cdots \circ s_1 \circ p$, and
-In light of the discussion above, we consider whether there exists an alternative \emph{proxy objective} to optimize that is related to the cost of a schedule but not dependent on the entries in the decomposition.
-Consider again the objective:
-\begin{equation}\label{eq:schedule_opt2}
-	S_\sigma^\ast = \argmin_{\sigma \in S_d} \mathrm{cost}(S(\sigma) \circ p)
-\end{equation}
-where now $\mathrm{cost}(\cdot)$ is to be replaced by a proxy objective function.
+
+%Consider again the objective: 
+%\begin{equation}\label{eq:schedule_opt2}
+%	S_\sigma^\ast = \argmin_{\sigma \in S_d} \mathrm{cost}(S(\sigma) \circ p)
+%\end{equation}
+%where now $\mathrm{cost}(\cdot)$ is to be replaced by a proxy objective function.
 There are exponentially many ways to choose $\sigma \in S_d$, so we opt for a greedy strategy as before, which takes $O(d \, C(m))$ time where $C(m)$ denotes the cost of evaluating the greedy step.
@@
 
-A natural upper bound on the number of column operations needed to move a simplex from position $i$ in the filtration to position $j$ is the displacement between these positions, $\lvert i - j \rvert$.
-In practice, due to the sparsity of both $R$ and $V$, this is a very loose upper bound---the number of column operations is generally much smaller.
-Nonetheless, the complexities of the move operations discussed in Section~\ref{sec:moves} are dependent on the magnitude of $\lvert i - j \rvert$, and thus it's worth considering as a proxy objective.
-If one forms bijections between two given filtrations $p: K_0 \to [m], q: K_1 \to [m]$ to permutations of $[m]$ as before, then an upper bound on the number of column reductions needed to perform any given schedule is given the \emph{Spearman distance}, a common distance for measuring the disarrangement between permutations.
-It is defined as:
+In light of the discussion above, we consider whether there exists an alternative \emph{proxy objective} to optimize that is both related to the cost of a schedule and not dependent on the entries in the decomposition.
+Obviously, the number of column operations needed to move a simplex from position $i$ in the filtration to position $j$ is upper bounded by $\lvert i - j \rvert$.
+In practice, due to the sparsity of both $R$ and $V$, this bound is very loose---the number of column operations is generally much smaller.
+Nonetheless, the complexities of the move operations discussed in Section~\ref{sec:moves} are dependent on the magnitude of $\lvert i - j \rvert$, and thus it's worth considering whether  \emph{minimizing the total displacement} would be a effective proxy objective.
+
+If one forms bijections between two given filtrations $p: K_0 \to [m], q: K_1 \to [m]$ to permutations of $[m]$ as before, then the Kendall-$\tau$ distance yields upper bound on the number of column reductions needed to perform \emph{any} schedule.
+This follows from the fact that any inversion between $p$ and $q$ may induce a column operation, and the number of inversions between $p$ and $q$ is exactly $K_{\tau}(p, q)$.
+Another common distance for measuring the disarrangement between permutations is \emph{Spearman distance}, defined as:
 \begin{equation}\label{eq:spearman_dist}
-F(p, q) = \sum\limits_{i \in [m]} \lvert p(i) - q(i) \rvert
+F(p, q) = \sum\limits_{i \in [m]} \lvert p^{-1}(i) - q^{-1}(i) \rvert
 \end{equation}
@@
 Like the Kendall-$\tau$ distance, it is a metric, it is invariant under relabeling, and the distance between pairs of uniformly random permutations $p, q \sim S_m$ are asymptotically normal.
-Indeed, Diaconis et al.~\cite{diaconis1977spearman} showed the following inequality relating the Spearman distance and the Kendall-$\tau$ distances:
+Indeed, Diaconis et al.~\cite{diaconis1977spearman} showed the the Spearman distance between any two permutaton $p, q \in S_m$ approximates their Kendall-$\tau$ distance within a factor of two:
 \begin{equation}\label{eq:diaconis_inequality}
@@
 \end{equation}
-Although very similar to the Kendall-$\tau$ distance, the Spearman distance offers some convenient computational advantages: it can be computed in $O(m)$ time, the variance of its asymptotic distribution is larger than Kendalls for uniformly random permutations, and it satisfies a certain monotonicity property under LCS edit operations.
-Specifically, given two permutations $p,q \in S_m$ with $\nu = \mathrm{LCS}(p,q)$ between them, the analysis in section~\ref{sec:move_schedules} shows we must move each symbol $x \in \mathcal{D}$ where $\mathcal{D} = p \smallsetminus \nu $ in a way that extends $\nu$ after each move.
-Each symbol $x \in \mathcal{D}$ induces a permutation $s_x$ such that $F(p, q) \geq F(s_x \circ p, q)$.
-That is, unlike the Kendall distance, the Spearman distance satisfies:
-$$ F(p, q) \geq F(s_1 \circ p, q) \geq F( s_2 \circ s_1 \circ p, q) \geq F( s_d \circ \dots \circ s_1 \circ p, q)  = F(q,q) = 0 $$
-For any sorting $S$ such that $s_{i}$ that increases the size of longest common subsequence between $s_{i-1} \circ \dots \circ s_1 \circ p$ and $q$.
-Given $p, q$ and some schedule $S = (s_1, \dots, s_d)$, let $\hat{s}_i =  s_i \circ \cdots \circ s_2 \circ s_1 \circ p$ denote the composition of the first $i$ permutations in $S$, where $i < d$.
-An upper bound on the Spearman distance can be decomposed into a sum of $d-1$ local distances with the following inequality:
-%A possible inequality to utilize is the following: 
-\begin{equation}\label{eq:kendall_inequality}
-F(p, q) = F(p, S(\sigma) \circ p) \leq \sum\limits_{i=1}^{d-1} F(\hat{s}_i, \hat{s}_{i+1})
+Although very similar to the Kendall-$\tau$ distance, the Spearman distance conveys certain computational advantages not inherited by the Kendall distance due to its simplicity.
+
+For example, the Spearman distance can be computed $O(m)$ time, and 
+For example, one of the tasks of the \emph{rank aggregation} problem is to obtain an  aggregation~\cite{} $\sigma \in S_m$ from a set of $k$ rankings $T = \{\tau_1, \tau_2, \dots, \tau_k\}$ that obeys some optimality criteria.
+A Kemeny optimal ranking is a permutation $\sigma^\ast$ that minimizes the sum of Kendall distances to the gievn set of rankings:
+$$ \argmin_{\sigma \in S_m} = \sum\limits_{i = 1}^k K_\tau(\sigma, \tau_i)$$ 
+Computationally, obtaining a Kemeny optimal ranking is NP-hard\cite{}.
+In contrast, a 2-approximation may be obtained by replacing the Kendall distance with the Spearman distance, as in this case the problem reduces to finding a perfect matching in a bipartite graph, which can be accomplished in polynomial time.
+
+We consider the difficulty of obtaining a sorting $p \stackrel{S}{\mapsto} q$ of size $d = m - \mathrm{LCS}(p,q)$ which has minimal total displacement.
+Formally, given $p, q \in S_m$ and some sorting  $S = (s_1, \dots, s_d)$, let $\hat{s}_i =  s_i \circ \cdots \circ s_2 \circ s_1$ denote the composition of the first $i$ permutations in $S$, where $i < d$.
+To incorporate the notion of displacement as an additive minimization problem, we consider minimizing an upper bound on the Spearman distance by decomposing it into a sum of $d-1$ local distances:
+\begin{equation}\label{eq:sorting_additive}
+F(p, q) \leq \sum\limits_{i=1}^{d-1} F(\hat{s}_i \circ p, \hat{s}_{i+1} \circ p)
 \end{equation}
@@
 Note this not guaranteed if $S$ was derived using the strategy from section~\ref{sec:move_schedules}.
+
+The problem of minimizing the right hand side of equation~\ref{eq:sorting_additive} can be interpreted as a crossing minimization problem for a set of $k$-layered bipartite graphs.
+To see this, consider two permutations: $p$ and $m_{ij} \circ p$, where $m_{ij}$ is a move permutation.
+If the $(p, m_{ij} \circ p)$ are drawn as a bipartite graph, with the elements of $p$ in one layer and the elements of $m_{ij} \circ p$ in another, then observe that $F(p, m_{i,j} \circ p)$ is twice the number of edge crossings in the graph.
+We given an example below to show this.
+\\
+\\
+\noindent \textbf{Example:} An example of three possible schedules, $S_1$, $S_2$, and $S_3$ is given in Figure~\ref{fig:crossings}. 
+%Each schedule, from left to right, has $21$, $31$, and $35$ crossings, respectively. 
+Each column represents the successive application of permutations from the three fixed schedules. Black/red vertices correspond to symbols in and outside of $\mathrm{LCS}(p,q)$, respectively. 
+\begin{figure}[ht]
+    \centering
+    \includegraphics[width=0.60\textwidth]{crossings.png}
+    \caption{Three sortings, $S_1$, $S_2$, and $S_3$, with $21$, $31$, and $35$ crossings, resp.}
+    %.
+Each schedule, from left to right, .
+The greedy heuristic proposed to minimize~\ref{eq:schedule_opt_succ} was used to produce $S_1$, which has the minimal number of crossings, since $K_\tau(p,q) = 21$ in this example.
+    \label{fig:crossings}
+\end{figure}
+%The greedy heuristic proposed to minimize equation~\ref{eq:schedule_opt_succ} was used to produce $S_1$, which has 21 edge crossings. 
+\noindent 
+All three schedules were generated from the same $\mathrm{LCS}(p,q) = (4\,7\,8)$ and all three schedules transform $p \mapsto q$ in $d = 6$ moves.
+In this example, this matches the minimal number of crossings amongst all possible schedules in fact, since $K_\tau(p,q) = 21$.
+When one does not need the sortings to be valid, this problem reduces to the $k$-layer crossing minimization problem for $k$ sets of permutations, which is NP-hard for $k \geq 4$~\cite{biedl2009complexity}.
+As a result, we do not expect there to be a polynomial time algorithm for optimally achieving this proxy objective.
+% Since $\mathcal{D}$ is fixed, the complexity of a greedy strategy depends entirely on the complexity of minimizing $F(\hat{s}_i, \hat{s}_{i+1})$ at each greedy step.  
+%For any sorting $S$ such that $s_{i}$ that increases the size of longest common subsequence between $s_{i-1} \circ \dots \circ s_1 \circ p$ and $q$. 
+% and it satisfies a certain monotonicity property under LCS edit operations. 
+%Specifically, given two permutations $p,q \in S_m$ with $\nu = \mathrm{LCS}(p,q)$ between them, the analysis in section~\ref{sec:move_schedules} shows we must move each symbol $x \in \mathcal{D}$ where $\mathcal{D} = p \smallsetminus \nu $ in a way that extends $\nu$ after each move.
+Each symbol $x \in \mathcal{D}$ induces a permutation $s_x$ such that $F(p, q) \geq F(s_x \circ p, q)$.
+That is, unlike the Kendall distance, the Spearman distance satisfies:
+%$$ F(p, q) \geq F(s_1 \circ p, q) \geq F( s_2 \circ s_1 \circ p, q) \geq F( s_d \circ \dots \circ s_1 \circ p, q)  = F(q,q) = 0 $$
 %Since the complexity of the move operations discussed in section~\ref{sec:moves} depend on the size of the movement, it is advantageous to construct a schedule which minimizes this movement. 
 %Since $\lvert \mathcal{D} \rvert$ is fixed, we would like to choose $x \in \mathcal{D}$ such that the corresponding permutation $s_x$ reduces the total displacement of every symbol $x \in \mathcal{D}$ relative to  $q$.
-Substituting objective~\ref{eq:kendall_inequality} for $\mathrm{cost}(\cdot)$ in equation~\ref{eq:schedule_opt2} yields the following global optimization problem over $S_d$: 
-\begin{equation}\label{eq:schedule_opt_succ}
-S_{\sigma}^\ast = \argmin_{\sigma \in S_d} \frac{1}{2} \sum\limits_{i = 1}^{d - 1} F(\hat{s}_i, \hat{s}_{i+1})
-\end{equation}
-The minimizer of this optimization can be interpreted as a sorting which minimizes the net displacement amongst all possible $d$-sized sortings $p \mapsto q$, again in the $\ell_1$ sense.
-When one does not need the sortings to be valid, this problem reduces to the $k$-layer crossing minimization problem for $k$ sets of permutations, which is NP-hard for $k \geq 4$~\cite{biedl2009complexity}.
-As a result, we do not expect a greedy solution to yield the optimal solution exactly, but simply a fast approximation.
+%Substituting objective~\ref{eq:kendall_inequality} for $\mathrm{cost}(\cdot)$ in equation~\ref{eq:schedule_opt2} yields the following global optimization problem over $S_d$: 
+%\begin{equation}\label{eq:schedule_opt_succ}
+%S_{\sigma}^\ast = \argmin_{\sigma \in S_d} \frac{1}{2} \sum\limits_{i = 1}^{d - 1} F(\hat{s}_i, \hat{s}_{i+1})
+%\end{equation}
+%The minimizer of this optimization can be interpreted as a sorting which minimizes the net displacement amongst all possible $d$-sized sortings $p \mapsto q$, again in the $\ell_1$ sense. 
 %A greedy strategy requires $O((d+(d-1)+(d-2)+\dots+1)m) \approx O(d^2 m)$ time to complete, since there are $O(d)$ symbols to choose from and minimizing $F(\hat{s}_i, \hat{s}_{i+1})$ at each step takes $O(md)$.
-Since $\mathcal{D}$ is fixed, the complexity of a greedy strategy depends entirely on the complexity of minimizing $F(\hat{s}_i, \hat{s}_{i+1})$ at each greedy step.  
 %\\
@@
 %% TODO: elaborate on monotonicity, describe segment tree
-\\
-\\
-\noindent \textbf{Example:} An example of three possible schedules, $S_1$, $S_2$, and $S_3$ is given in Figure~\ref{fig:crossings}.
-Each schedule, from left to right, has $21$, $31$, and $35$ crossings, respectively.
-Each column represents the successive application of permutations from the three fixed schedules.
-Black/red vertices correspond to symbols in and outside of $\mathrm{LCS}(p,q)$, respectively.
-\begin{figure}[ht]
-    \centering
-    \includegraphics[width=0.60\textwidth]{crossings.png}
-    \caption{Example of three possible schedules, $S_1$, $S_2$, and $S_3$.}
-    %.
-Each schedule, from left to right, has $21$, $31$, and $35$ crossings, respectively.
-The greedy heuristic proposed to minimize~\ref{eq:schedule_opt_succ} was used to produce $S_1$, which has the minimal number of crossings, since $K_\tau(p,q) = 21$ in this example.
-    \label{fig:crossings}
-\end{figure}
-All three schedules were generated from the same $\mathrm{LCS}(p,q) = (4\,7\,8)$ and all three schedules transform $p \mapsto q$ in $d = 6$ moves.
-The greedy heuristic proposed to minimize equation~\ref{eq:schedule_opt_succ} was used to produce $S_1$, which has the minimal number of crossings amongst all possible schedules in this example, since $K_\tau(p,q) = 21$.
-\\
-\\
-\noindent 
-The monotonic nature of successively applying the Spearman distance enables certain computational advantages.
+
+The simplicity of both the Spearman distance and the types of permutations we consider here enables certain computational advantages.
 Suppose we begin with an array $\mathcal{A}$ of size $m$ which provides $O(1)$ access and modification, initialized with the (signed) \emph{displacement} of every element in $p$ to its corresponding position in $q$.
-Since the Spearman distance is simply the sum of these displacements, if we could replace $\mathcal{A}$ with a data structure that supports $O(\log m)$ access time to aggregate information (i.e.
-the sum) and $O(\log m)$ modification time ability on $\lvert i - j \vert$ entries (simultaneously), we would have an efficient $O(d \log m)$ greedy solution.
+Since the Spearman distance is simply the sum of the absolute value of these displacements, at any point during during the execution of Algorithm~\ref{alg:sorting} we may obtain $F(\bar{p}, q)$ simply by having access to the sum of every entry in $\mathcal{A}$.
+There are two degrees of freedom in the Algorithm~\ref{alg:sorting}: the first is in the choice of $\sigma \in \mathcal{D}$, and the second is in choosing $j$.
+If we fix a heuristic for the latter, then we have a set of possible valid moves $S_\mathcal{D}$ induced by each $\sigma \in \mathcal{D}$ to choose from.
+Observe that each permutation $s_\sigma \in S_\mathcal{D}$ changes the displacement of every symbol in at most three different ways:
+\[
+F(s_\sigma \circ p, q) = 
+\begin{cases} 
+	 \mathcal{A}(\sigma) + \lvert i - j \rvert & p^{-1}(\sigma) = i \\
+	 \mathcal{A}(\sigma) \pm 1 & i < p^{-1}(\sigma) \leq j \\
+	 \mathcal{A}(\sigma) & \text{ otherwise }
+\end{cases}
+\]
+That is, if $s_\sigma$ represents a move permutation that moves $\sigma$ from position $i$ to position $j$, then relative to the previous permutation $p$, the displacement of $\sigma$ in $s_\sigma \circ p$ changes by $\lvert i - j \rvert$ and the displacement of any symbol between $i$ and $j$ changes by $\pm 1$.
+All other symbol displacements are unchanged.
+Thus, if we replace $\mathcal{A}$ with a data structure that supports $O(\log m)$ access time to aggregate information (i.e.
+the sum) and $O(\log m)$ modification time ability on $\lvert i - j \vert$ entries (simultaneously), we could greedily choose the next permutation $s$ to maximize:
+\begin{equation}\label{eq:greedy_step}
+	s_{\text{greedy}} = \argmin_{\text{ valid } s \in S_\mathcal{D}} F(s\circ p, q)
+\end{equation}  
+in $O(d\log m)$ time. 
 The former problem reduces to the problem of efficiently calculating \emph{prefix sums}, which is easily solved.
 It's not immediately clear how to achieve the latter modification complexity in $O(\log m)$ time, since $\lvert i - j \rvert \leq m$ is potentially larger than $O(\log m)$.
-However, observe that each permutation $s_x$ changes the displacement of every symbol in a very predictable way: if $s_x$ represents a move permutation that moves $x$ from position $i$ to position $j$, then relative to the previous permutation $s_{i}$, the displacement of $x$ in $s_{i+1}$ changes by $\lvert i - j \rvert$ and the displacement of any symbol between $i$ and $j$ changes by $\pm 1$.
-All other symbol displacements are unchanged.
 %, and
+% Can actually only achieve O(d^2 log m) it seems 
 It is known that one can apply constant-factor updates to multiple values in a prefix sum in $O(\log m)$ time by any prefix-sum data structure that also supports \emph{range updates}, e.g.
-a Fenwick tree, or a Segment Tree that supports lazy propagation~\cite{de1997computational}.
-Since single element prefix sum modifications can also be achieved $O(\log m)$ time with these data structures, the time to compute $F(\hat{s}_i, \hat{s}_{i+1})$ is bounded above by $O(\log m)$, thus the greedy solution to equation~\ref{eq:schedule_opt_succ} can be computed in $O(d \log m)$ time.
+an \emph{implicit treap}.
+Since single element modifications, removals, and insertions can all be achieved in $O(\log m)$ expected time with this data structure, we conclude that a greedy solution to equation~\ref{eq:greedy_step} can be computed in $O(d^2 \log m)$ time.
+%Since single element prefix sum modifications can also be achieved $O(\log m)$ time with these data structures, the time to compute $F(\hat{s}_i, \hat{s}_{i+1})$ is bounded above by $O(\log m)$, 
+%thus the greedy solution to equation~\ref{eq:schedule_opt_succ} can be computed in $O(d^2 \log m)$ time.
+
+%  e.g. a Fenwick tree, or a Segment Tree that supports lazy propagation~\cite{de1997computational}
+
 %We give an example and a more detailed discussion for those interested in the appendix\ref{sec:appendix}. 

### 2021-10-26 move_schedules.tex
Updates from Overleaf

@@
 \author{
-	Matt Piekenbrock\thanks{Department of Computational Mathematics, Science \& Engineering, Michigan State University.}
+	Matt Piekenbrock\thanks{Khoury College of Computer Sciences, Northeastern University.}
 	\and Jose A. Perea\thanks{
-	Department of Computational Mathematics, Science \& Engineering, 
-	Department of Mathematics, Michigan State University.}
+	Department of Mathematics and Khoury College of Computer Sciences, Northeastern University.}
 }
@@
 \subsection{Overview}\label{sec:overview} 
-Given a triangulable topological space equipped with a sufficiently tame continuous function, persistent homology captures the changes in topology across the sublevel sets of the space, and encodes them in a persistence diagram.
-The stability of persistence contends that small changes in the underlying function produce small changes in the corresponding diagrams:  if the function changes continuously, so too will the points on the persistence diagram~\cite{cohen2007stability, cohen2006vines}.
-This motivates the study and application of persistence to time-varying or dynamic settings, like that of dynamic metric spaces~\cite{kim2020spatiotemporal}. 
+Given a triangulable topological space equipped with a  tame continuous function, persistent homology captures the changes in topology across the sublevel sets of the space, and encodes them in a persistence diagram.
+The stability of persistence contends that   if the function changes continuously, so too will the points on the persistence diagram~\cite{cohen2007stability, cohen2006vines}.
+This motivates the  application of persistence to time-varying  settings, like that of dynamic metric spaces~\cite{kim2020spatiotemporal}. 
 As persistence-related computations tend to exhibit high algorithmic complexity---essentially cubic\footnote{For finite fields, it is known that the persistence computation reduces to the PLU factorization problem, which takes $O(m^\omega)$ where $\omega \approx 2.373$ is the matrix multiplication constant.} in the size of the underlying filtration~\cite{morozov2005persistence}---their adoption to dynamic settings poses a challenging computational problem.
@@
 Indeed, while there is an increasingly rich ecosystem of software packages offering variations of the standard reduction algorithm (e.g.
-Ripser, PHAT, Dionysus, etc...
-see~\cite{otter2017roadmap} for an overview), implementations of the vineyards algorithm are relatively uncommon.\footnote{Dionysus 1 has an implementation of vineyards, but the algorithm was never ported to the 2nd version.}
+Ripser, PHAT, Dionysus, etc.
+See~\cite{otter2017roadmap} for an overview), implementations of the vineyards algorithm are relatively uncommon.\footnote{Dionysus 1 has an implementation of vineyards, but the algorithm was never ported to the 2nd version.}
 The reason for this disparity is perhaps explained by Lesnick and Wright~\cite{lesnick2015interactive}: ``While an update to an $RU$ decomposition involving few transpositions is very fast in practice...
@@
 While vineyards is  designed  for handling   \emph{continuous} time families of diagrams, it is not necessarily efficient when the time parameter is coarsely discretized.
-Our methodology is based on the observation that practitioners often don't need (or want!) all $d$ persistence diagrams generated by a continuous 1-parameter of filtrations; usually just $n << d$ of them are sufficient.
+Our methodology is based on the observation that practitioners often don't need (or want!) all $d$ persistence diagrams generated by a continuous 1-parameter of filtrations; usually just $n << d$ of them  suffice.
 % in many time varying settings, a valid decomposition is often only sought after at a relatively sparse number of time points relative to the size of the (finite) set needed to generated by the family.
@@
  \subsection{A Motivating Example and Contributions}\label{sec:motivation} 
-As a motivating example, we describe a simple experiment illustrating why the vineyards algorithm does not always yield an efficient method for time-varying applications. 
-Consider a series of grayscale images (i.e., a video) depicting a fixed-width annulus expanding about the center of a $9 \times 9$ grid, where the scale of the annulus is parameterized by a single parameter $\tau$.
-Each image contains the same number of pixels whose intensities---these are determined by $\tau$---vary with time.
-For each image, we build a filtered simplicial complex using the \emph{Freudenthal} triangulation of the plane, with the order of inclusions obtained by lower stars.
+As a motivating example, we describe a simple experiment illustrating why the vineyards algorithm does not always yield an efficient strategy for time-varying applications. 
+Consider a series of grayscale images (i.e., a video) depicting a fixed-width annulus expanding about the center of a $9 \times 9$ grid. 
+Each image contains the same number of pixels, whose intensities vary with time.
+For each image, we build a filtered simplicial complex using the \emph{Freudenthal} triangulation of the plane, with the order of inclusions obtained by lower stars and pixel values.
  \begin{figure*}[t]
@@
  \end{figure*}
-Snapshots of the image data and its corresponding time-varying filtration is depicted in Figure~\ref{fig:vineyards} (top right and bottom right, respectively), which we use as a baseline for comparing the standard reduction algorithm $\texttt{pHcol}$ (Algorithm \ref{alg:reduce}) with vineyards.
-Note that the size of the filtration is fixed--no simplices are ever added or removed over time.
-There are two events which dramatically alter the persistence diagrams extracted from these filtrations: the first occurs when the central connected component splits to form a cycle, and the second occurs when the annulus approaches the outer dimensions of the grid, splitting into four components.
+Snapshots of the image data and its corresponding time-varying filtration is depicted in Figure~\ref{fig:vineyards} (top right and bottom right, respectively), which we use as a baseline for comparing vineyards and the standard reduction algorithm $\texttt{pHcol}$ (Algorithm \ref{alg:reduce}).
+Note that the size of the filtration is fixed---no simplices are ever added or removed over time. 
+There are two events which dramatically alter the persistence diagrams extracted from these filtrations: the first occurs when the central connected component splits to form a cycle, and the second  when the annulus   splits into four components.
 As a consequence, in this example, only a few persistence diagrams are needed to capture the major topological changes over time.
@@
   Since it is unknown a priori at which times  the persistent homology will change, one solution is to discretize the time domain at $n$ evenly spaced points and compute the persistence diagrams at each one of them  independently.
-An alternative approach is to construct a homotopy between two contiguous time-points, and then decompose this homotopy into adjacent transpositions--the vineyards approach.
+  An alternative approach is to construct a (e.g., straight-line) homotopy between  contiguous time-points, and then decompose these homotopies into adjacent transpositions in the filtration order---the vineyards approach.
 The former is often used in practice, but it is only an approximation; the latter is guaranteed to capture all homological changes in persistence that occur in simulating the 1-parameter family.
   The graph in Figure~\ref{fig:vineyards} depicts two different size approximations (purple, $n=100$ and green, $n=10$), wherein the reduction algorithm (\texttt{pHcol}) was applied at $n$ time points, and two exact strategies (black and blue), which compute all $\approx 7 \times 10^{4}$ diagrams given by the homotopy.
-  As one can see from the figure, if one needs to simulate the homotopy exactly, then the vineyards approach is indeed far more efficient than naively applying the reduction algorithm independently at all time points.
+  As shown in the figure, if one needs to simulate the homotopy exactly, then the vineyards approach is indeed far more efficient than naively applying the reduction algorithm independently at all time points.
   However, when the discretization of the time domain is coarse enough, the naive approach actually performs less column operations than the vineyards strategy, while still capturing the main events.
 
-  The existence of a time discretization that is more efficient to compute than continually updating the decomposition, indicates that the vineyards algorithm must incur some amount of overhead (in terms of column operations) to perform transpositions, even if the persistence diagrams remain unchanged.
-  Even in situations where one is updating a decomposition to a filtration that is  relatively ``close'', e.g.
-any two adjacent filtrations in the bottom right of Figure~\ref{fig:vineyards}, it is still more efficient to apply \texttt{pHcol} directly than  to iteratively update the decomposition (see the case where $n = 10$).
+  The existence of a time discretization that is more efficient to compute than continually updating the decomposition, indicates that the vineyards algorithm must incur some overhead (in terms of column operations) to perform transpositions, even if the persistence diagrams remain unchanged.
+  Even in situations where one is updating a decomposition to a filtration that is  relatively ``close''---e.g.
+any two adjacent filtrations in the bottom right of Figure~\ref{fig:vineyards}---it is still more efficient to apply \texttt{pHcol} directly than  to iteratively update the decomposition (see the case where $n = 10$).
  Moreover, any optimization to the reduction algorithm (e.g.,  clearing~\cite{chen2011persistent}) would only increase this disparity.
@@
 \textbf{Our approach and contributions are as follows:} First, we leverage the \emph{moves} framework of Busaryev et.
-al.~\cite{busaryev2010tracking} to include  coarser  operations---in terms of valid intermediate decomposition states, as compared to vineyards---for the dynamic setting.
+al.~\cite{busaryev2010tracking} to include  coarser  operations---in terms of the number of valid intermediate decomposition states, as compared to vineyards---for dynamic persistence.
 We give a tight lower bound on the number of moves needed to perform an arbitrary permutation to the $R = D V$ decomposition, and give a proof of optimality by a reduction to the permutation edit distance problem.
@@
 This reduction yields an efficient algorithm for generating sequences of moves $( s_1, s_2, \dots, s_d )$ of minimal size $d$, which we call \emph{schedules}. 
-However, not all valid schedules incur the same permutation costs 
+However, not all minimal size schedules incur the same  cost 
 (i.e., number of column operations).
@@
 A performance comparison with other reduction-based persistent homology computations is given; 
-our move schedules are demonstrated to be an order of magnitude more efficient than existing approaches in applications such as flock simulation and 2D persistence computations. 
+our move schedules are demonstrated to be an order of magnitude more efficient than existing approaches in applications such as flock simulations and 2D persistence computations with RIVET. 
 %We conclude with a discussion of situations where our strategy is most applicable and future directions for research.
@@
 In Section \ref{sec:move_schedules} we introduce our move schedules  and provide efficient algorithms to make this alternative viable. 
-In Section \ref{sec:results} we demonstrate a few applications of the proposed method. 
+In Section \ref{sec:results} we present applications of the proposed method. 
 In particular, to the computation of crocker stacks from flock simulations, and to 2-dimensional persistence on the Mumford data set of image patches from natural images. 
@@
 $\mathbb{F}$-vector space  denoted   $C_p(K)$. 
-The $p$-boundary $\partial_p(\sigma)$  of a $p$-simplex $\sigma\in K$ is the alternating sum of its oriented co-dimension 1 faces; the $p$-boundary of a $p$-chain is defined 
+The $p$-boundary $\partial_p(\sigma)$  of a $p$-simplex $\sigma\in K$ is the alternating sum of its oriented co-dimension 1 faces,
+and the $p$-boundary of a $p$-chain is defined 
 linearly in terms of its constitutive simplices. 
@@
 $p$-th homology of $K$ with coefficients in $\mathbb{F}$. 
-If $\{K_i\}_{i\in [m]}$ is a filtration, then the inclusion maps  $K_i\subset K_j$  for $i < j$, induce linear transformations $f_p^{i,j}: H_p(K_i) \to H_p(K_j)$ at the level of homology:
+If $\{K_i\}_{i\in [m]}$ is a filtration, then the inclusion maps  $K_i\subset K_{i+1}$   induce linear transformations 
+%$f_p^{i,j}: H_p(K_i) \to H_p(K_j)$ 
+at the level of homology:
 \begin{equation}
@@
 \end{equation}
-The $p$-th persistent homology groups are the images of these transformations: $H_{p}^{i,j} = \mathrm{Im}\,f_p^{i,j}$. 
+%The $p$-th persistent homology groups are the images of these transformations: $H_{p}^{i,j} = \mathrm{Im}\,f_p^{i,j}$. 
 %Note that if $i = j$, then $H_{p}^{i,j} = H_{p}(K_i) = H_{p}(K_i)$ is   just the ``standard'' homology. 
@@
 \\
-Given a triangulable topological space $\mathbb{X}$ equipped with a
-  function $f: \mathbb{X} \to \mathbb{R}$, 
-we write $\mathbb{X}_a = f^{-1}(-\infty, a]$ to denote the sublevel sets of $\mathbb{X}$ defined by the value $a$. 
+Let $\mathbb{X}$ be a triangulable topological space. 
+That is, so that there exists an abstract 
+simplicial complex $K$ whose geometric realization  is homeomorphic to 
+$\mathbb{X}$. 
+Let  $f: \mathbb{X} \to \mathbb{R}$ be continuous and  
+write $\mathbb{X}_a = f^{-1}(-\infty, a]$ to denote the sublevel sets of $\mathbb{X}$ defined by the value $a$. 
 A \emph{homological critical value} of $f$ is any value $a \in \mathbb{R}$ such that the homology of the sublevel sets of $f$ changes at $a$, i.e.
-if for some $p$ the inclusion-induced homomorphism  $H_p(\mathbb{X}_{a - \epsilon}) \to H_p(\mathbb{X}_{a+\epsilon})$ is not an isomorphism for  any $\epsilon >0$.
+if for some $p$ the inclusion-induced homomorphism  $H_p(\mathbb{X}_{a - \epsilon}) \to H_p(\mathbb{X}_{a+\epsilon})$ is not an isomorphism for  any small enough $\epsilon >0$.
 If there are only finitely many of these homological critical values, then $f$ is said to be \emph{tame}.
@@
 As $\tau$ varies in $[0,1]$, the points in $\mathrm{dgm}_p(f_\tau)$ trace  curves in $\mathbb{R}^3$, called \emph{vines} and which together form a \emph{vineyard}. 
-The stability of persistence implies that these curves will be continuous if the homotopy is continuous.
+The stability of persistence implies that these curves will be continuous if $F$ is continuous
+and the $f_\tau$'s are tame \cite{cohen2007stability}.
 The vineyard analogy acts as a guidepost for practitioners seeking to understand how subtle changes occurring to the topological structure over time.
@@
 is the row index of its lowest nonzero entry.
-Using $\mathbb{Z}_2$ coefficients and homology in dimension 1 to simplify the presentation, the steps look at follows:
+Using $\mathbb{Z}_2$ coefficients and homology in dimension 1 to simplify the presentation, the reduction proceeds as follows:
 \begin{displaymath}
@@
 \end{displaymath}
-\noindent The column $c$ in $R$ indicates that dimension $1$ homology is born, which in this filtration is never killed since the dimension of the complex is $2$.
+\noindent The column $c$ in $R$ indicates that dimension $1$ homology is born, which in this filtration is never killed since the dimension of the complex is $1$.
 Similarly, the columns at $u, v, w$ in $R_0$ (not shown) are all zero, indicating three $0$-dimensional homology classes  are born, which are then killed by the pivot entries in columns $a$ and $b$ in $R_1$.
@@
 \noindent
-Its clear from Algorithm~\ref{alg:reduce} that a loose upper bound for this reduction is $O(m^3)$, where $m$ is the number of simplices of the filtration; it turns out that this bound is in fact tight, see~\cite{morozov2005persistence} for more details.
+It is clear from Algorithm~\ref{alg:reduce} that a loose upper bound for this reduction is $O(m^3)$, where $m$ is the number of simplices of the filtration; it turns out that this bound is in fact tight, see~\cite{morozov2005persistence} for more details.
 There are actually many variations and optimizations proposed over the past decade to Algorithm~\ref{alg:reduce}; for example, one can   reverse the order of the filtration and compute cohomology~\cite{de2011dualities} instead.
@@
 The original purpose of the vineyards algorithm, as described in~\cite{cohen2006vines}, was to compute a continuous 1-parameter family of persistence diagrams over a time-varying filtration, and to detect homological critical events during the construction.
-Notice that homological critical events can only occur when the filtration order changes, though the converse isn't true.
-Thus, detecting these all such events reduces to computing matrix decompositions at a finite set of time points interleaved between changes in the filtration order. 
-At the finest scale, these changes manifest as adjacent transpositions. 
-As a result, a fixed set of rules for maintaining a valid $R = D V$ decomposition under the transposition of adjacent simplices in the filtration is enough to simulate persistence dynamically. 
+Notice that homological critical events can only occur when the filtration order changes, though the converse is not true.
+Thus, detecting  all such events reduces to computing matrix decompositions at a finite set of time points interleaved between changes in the filtration order. 
+At the finest scale, these changes manifest as transpositions of adjacent simplices. 
+As a result, a fixed set of rules for maintaining a valid $R = D V$ decomposition under adjacent transpositions  is enough to simulate persistence dynamically. 
 These rules prescribe certain column and row operations which must be applied to the matrix decomposition either before, during, or after the transposition, in order to ensure decomposition invariants I1 and I2 are respected.
@@
 
- The time complexity of vineyards is determined entirely by the complexity of performing a single adjacent transposition, and its complexity is often stated in this way.
+ The time complexity of vineyards is determined entirely by the complexity of performing a single adjacent transposition.
  Informally, since column operations are the largest complexity operations needed and each column can have potentially $m$ entries, the complexity of vineyards is $O(m)$ per transposition. 
  However, there are a number of subtleties involved in this statement. 
- The first subtlety is that both the $V$ and $R$ matrices tend to be quite sparse, and thus column and row operations rarely require $m$ operations---in fact, as a rule of thumb, most transpositions require no column operations~\cite{edelsbrunner2000topological}.
+ The first  is that both the $V$ and $R$ matrices tend to be quite sparse, and thus column and row operations rarely require $m$ operations---in fact, as a rule of thumb, most transpositions require no column operations~\cite{edelsbrunner2000topological}.
 A second subtlety, due to the necessity of sparse matrix data structures, is that querying the non-zero status of any particular entry typically takes $O(\log m)$ time.
@@
  	\item[M1:] Querying the non-zero status of entries in either $R$ or $V$ occurs once per move.
- 	\item[M2:] $R = D V$ is not maintained during the movement of $\sigma_i \mapsto \sigma_j$.
+ 	\item[M2:] $R = D V$ is \underline{not} guaranteed to be valid  during the movement of $\sigma_i $ to $ \sigma_j$.
  \end{enumerate} 
@@
 Prior to applying any permutation $P$ to the decomposition, it is necessary to remove non-zero entries in $V$ which render $P^TVP$ non-upper triangular, to maintain invariant I2.
-If one uses the (augmented) sparse matrix representation described in the original vineyards publication~\cite{cohen2006vines}, checking whether these entries are non-zero takes $O(\log m)$ per column.
-Since moving a simplex from $i$ to $j$ using vineyards performs $\lvert i - j \rvert - 1$ transpositions, this query operation is required at most $O(\lvert i - j \rvert)$ times, which is excessive.
-If a row-oriented sparse matrix implementation is used, a move operation can perform these non-zero entry checks in just one $O(m)$ pass, prior to performing any column operations.
+If one uses the (augmented) sparse matrix representation described in~\cite{cohen2006vines}, checking whether these entries are non-zero takes $O(\log m)$ per column.
+Since moving a simplex from $i$ to $j$ using vineyards performs $\lvert i - j \rvert - 1$ transpositions, this query operation is required at most $O(\lvert i - j \rvert)$ times.
+By comparison, if a row-oriented sparse matrix implementation is used, then a move operation can perform these non-zero entry checks in just one $O(m)$ pass, prior to performing any column operations.
 
@@
 Indeed, if one has a pair of filtrations $(K_0, K_1)$ each with $m$ simplices, and we assume that each simplex $\sigma_i$ switches its relative ordering with another simplex $\sigma_j$ at most once in some ordered schedule of transpositions, then the number of intermediate persistence diagrams is bounded above by $O(m^2)$ using transpositions, whereas with move operations this bound clearly is at most $O(m)$.
-\begin{algorithm}
+\\
+
+\noindent \textbf{Example:} We continue the example used in sections~\ref{sec:vineyards} and~\ref{sec:reduction} to illustrate moves.
+Consider moving the edge $a$ to the position of edge $c$ in the filtration.
+The donor columns introduced in~\cite{busaryev2010tracking} are shown on the left side of each matrix.
+Note that using vineyards, the equivalent permutation to the decomposition would require $4$ column operations on both $R_1$ and $V_1$, respectively, whereas a single move operation accomplishes this permutation using only $2$ column operations per matrix.
+\begin{displaymath}
+	\def\rowsp{0.4em}
+	\def\arraystretch{0.95}
+	\makeatletter\setlength\BA@colsep{1.2pt}\makeatother
+	%
+	\begin{blockarray}{cc}{d_R}
+	\kern-0.2em & a \\
+		\begin{block}{c[c]}
+  		u \kern \rowsp & \; 1 \; \topstrut \\
+  		v \kern \rowsp & \;  \; \\
+  		w \kern \rowsp & \; 1 \; \botstrut \\
+		\end{block}
+	\end{blockarray}
+	\; \;
+	\begin{blockarray}{cccc}{R_1}
+	 \kern-0.2em & a & b & c  \\
+		\begin{block}{c[ccc]}
+  		u \kern \rowsp & \; 1 & 1 &  \; \topstrut \\
+  		v \kern \rowsp & \;  & 1 &  \; \\
+  		w \kern \rowsp & \; 1 &  &  \; \botstrut \\
+		\end{block}
+	\end{blockarray}
+	%
+\; \to
+	%
+\begin{blockarray}{cc}{}
+	\kern-0.2em & b \\
+		\begin{block}{c[c]}
+  		  & \; 1 \; \topstrut \\
+  		 & \;  1 \; \\
+  		 & \; \; \botstrut \\
+		\end{block}
+	\end{blockarray}
+	\; \;
+\begin{blockarray}{cccc}{}
+	\kern-0.2em & a & b & c  \\
+		\begin{block}{c[ccc]}
+  		  & \; 1 &    &  \; \topstrut \\
+  		  & \;    & 1 &  \; \\
+  		 & \; 1 & \fbox{1} &  \; \botstrut \\
+		\end{block}
+	\end{blockarray}
+	%
+\; \to
+	%
+	\begin{blockarray}{cc}{}
+	\kern-0.2em & c \\
+		\begin{block}{c[c]}
+  		  & \; \hphantom{1} \; \topstrut \\
+  		 & \;  \hphantom{1} \; \\
+  		 & \; \hphantom{1} \; \botstrut \\
+		\end{block}
+	\end{blockarray}
+	\; \; 
+	\begin{blockarray}{cccc}{}
+	\kern-0.2em & a & b &  c \\
+		\begin{block}{c[ccc]}
+  		  & \; 1   &    &   1  \; \topstrut \\
+  		 & \;       &  1  &  1  \; \\
+  		 & \; 1   &   \fbox{1}   &     \; \botstrut \\
+		\end{block}
+	\end{blockarray}
+%
+\; \xrightarrow{P} 
+\begin{blockarray}{cc}{}
+	\kern-0.2em & c \\
+		\begin{block}{c[c]}
+  		  & \; \hphantom{1} \; \topstrut \\
+  		 & \; \hphantom{1} \; \\
+  		 & \; \hphantom{1} \; \botstrut \\
+		\end{block}
+	\end{blockarray}
+\; \;
+\begin{blockarray}{cccc}{}
+	\kern-0.2em & b & c & a  \\
+		\begin{block}{c[ccc]}
+  		  & \;  & 1 &  1\; \topstrut \\
+  		 & \; 1 & 1 &  \; \\
+  		 & \; 1 &  & \fbox{1} \; \botstrut \\
+		\end{block}
+	\end{blockarray}
+%
+\; \xrightarrow{d_R} 
+\begin{blockarray}{cccc}{R_1}
+\kern-0.2em & b & c & a  \\
+	\begin{block}{c[ccc]}
+		u \kern \rowsp & \;  & 1 &  \; \topstrut \\
+		v \kern \rowsp & \; 1 & 1 &  \; \\
+		w \kern \rowsp & \; 1 &  &  \; \botstrut \\
+	\end{block}
+	% \undermat{B}{ & & }  
+\end{blockarray}
+\end{displaymath}
+
+\vspace{-1.0em}
+
+% Start V example 
+\begin{displaymath}
+	\def\rowsp{0.4em}
+	\def\arraystretch{0.95}
+	\makeatletter\setlength\BA@colsep{1.2pt}\makeatother
+	%
+	\begin{blockarray}{cc}{d_V}
+	\kern-0.2em & a \\
+		\begin{block}{c[c]}
+  		 & \; 1 \; \topstrut \\
+  		& \;  \; \\
+  		& \;  \; \botstrut \\
+		\end{block}
+	\end{blockarray}
+	\; \;
+	\begin{blockarray}{cccc}{V_1}
+	 \kern-0.2em & a & b & c  \\
+		\begin{block}{c[ccc]}
+  		 & \; 1 & \fbox{1} &  \fbox{1}\; \topstrut \\
+  		& \;  & 1 &  1\; \\
+  		& \; &  & 1 \; \botstrut \\
+		\end{block}
+	\end{blockarray}
+	%
+\; \to
+	%
+\begin{blockarray}{cc}{}
+	\kern-0.2em & b \\
+		\begin{block}{c[c]}
+  		 & \; 1 \; \topstrut \\
+  		& \;  1 \; \\
+  		& \; \; \botstrut \\
+		\end{block}
+	\end{blockarray}
+	\; \;
+\begin{blockarray}{cccc}{}
+	 \kern-0.2em & a & b & c  \\
+		\begin{block}{c[ccc]}
+  		 & \; 1 & &  \fbox{1}\; \topstrut \\
+  		& \;  & 1 &  1\; \\
+  		& \; &   & 1 \; \botstrut \\
+		\end{block}
+	\end{blockarray}
+	%
+\; \to
+	%
+	\begin{blockarray}{cc}{}
+	\kern-0.2em & c \\
+		\begin{block}{c[c]}
+  		 & \; 1 \; \topstrut \\
+  		& \;  1 \; \\
+  		& \; 1 \; \botstrut \\
+		\end{block}
+	\end{blockarray}
+	\; \;
+	\begin{blockarray}{cccc}{}
+	\kern-0.2em & a & b &  c \\
+		\begin{block}{c[ccc]}
+  		 & \; 1 & &   \; \topstrut \\
+  		& \;  & 1 &   \; \\
+  		& \;  &   & 1 \; \botstrut \\
+		\end{block}
+	\end{blockarray}
+%
+\; \xrightarrow{P} \\
+% 
+\begin{blockarray}{cc}{}
+	\kern-0.2em & a \\
+		\begin{block}{c[c]}
+  		 & \; 1 \; \topstrut \\
+  		& \;  1 \; \\
+  		& \; 1 \; \botstrut \\
+		\end{block}
+	\end{blockarray}
+\; \;
+\begin{blockarray}{cccc}{}
+	\kern-0.2em & b & c & a  \\
+		\begin{block}{c[ccc]}
+  		 & \; 1 & &\; \topstrut \\
+  		& \;  & 1 &  \; \\
+  		& \; &  & 1 \; \botstrut \\
+		\end{block}
+	\end{blockarray}
+%
+\; \xrightarrow{d_V}
+% 
+	\begin{blockarray}{cccc}{V_1}
+	\kern-0.2em & b & c & a  \\
+		\begin{block}{c[ccc]}
+  		b \kern \rowsp & \; 1 & & 1 \; \topstrut \\
+  		c \kern \rowsp & \; & 1 &  1 \; \\
+  		a \kern \rowsp & \; &  & 1  \; \botstrut \\
+		\end{block}
+	\end{blockarray}
+\end{displaymath}
+
+\begin{algorithm}[!htb]
 	\caption{Move Right Algorithm}\label{alg:mr}
@@
 
-\begin{algorithm}
+\begin{algorithm}[!htb]
 	\caption{Move Left Algorithm}\label{alg:ml}
@@
 \end{algorithm}
-If one has a pair of filtrations $(K_0, K_1)$ each with $m$ simplices, and we assume that each simplex $\sigma_i$ switches its relative ordering with another simplex $\sigma_j$ at most once in some ordered schedule of transpositions, then the number of intermediate persistence diagrams is bounded above by $O(m^2)$ using transpositions, whereas with move operations this bound  is at most $O(m)$.
-\\
-\\
-\noindent \textbf{Example:} We continue the example used in sections~\ref{sec:vineyards} and~\ref{sec:reduction} to illustrate moves.
-Consider moving the edge $a$ to the position of edge $c$ in the filtration.
-The donor columns introduced in~\cite{busaryev2010tracking} are shown on the left side of each matrix.
-Note that using vineyards, the equivalent permutation to the decomposition would require $4$ column operations on both $R_1$ and $V_1$, respectively, whereas a single move operation accomplishes this permutation using only $2$ column operations per matrix.
-\begin{displaymath}
-	\def\rowsp{0.4em}
-	\def\arraystretch{0.95}
-	\makeatletter\setlength\BA@colsep{1.2pt}\makeatother
-	%
-	\begin{blockarray}{cc}{d_R}
-	\kern-0.2em & a \\
-		\begin{block}{c[c]}
-  		u \kern \rowsp & \; 1 \; \topstrut \\
-  		v \kern \rowsp & \;  \; \\
-  		w \kern \rowsp & \; 1 \; \botstrut \\
-		\end{block}
-	\end{blockarray}
-	\; \;
-	\begin{blockarray}{cccc}{R_1}
-	 \kern-0.2em & a & b & c  \\
-		\begin{block}{c[ccc]}
-  		u \kern \rowsp & \; 1 & 1 &  \; \topstrut \\
-  		v \kern \rowsp & \;  & 1 &  \; \\
-  		w \kern \rowsp & \; 1 &  &  \; \botstrut \\
-		\end{block}
-	\end{blockarray}
-	%
-\; \to
-	%
-\begin{blockarray}{cc}{}
-	\kern-0.2em & b \\
-		\begin{block}{c[c]}
-  		  & \; 1 \; \topstrut \\
-  		 & \;  1 \; \\
-  		 & \; \; \botstrut \\
-		\end{block}
-	\end{blockarray}
-	\; \;
-\begin{blockarray}{cccc}{}
-	\kern-0.2em & a & b & c  \\
-		\begin{block}{c[ccc]}
-  		  & \; 1 &    &  \; \topstrut \\
-  		  & \;    & 1 &  \; \\
-  		 & \; 1 & \fbox{1} &  \; \botstrut \\
-		\end{block}
-	\end{blockarray}
-	%
-\; \to
-	%
-	\begin{blockarray}{cc}{}
-	\kern-0.2em & c \\
-		\begin{block}{c[c]}
-  		  & \; \hphantom{1} \; \topstrut \\
-  		 & \;  \hphantom{1} \; \\
-  		 & \; \hphantom{1} \; \botstrut \\
-		\end{block}
-	\end{blockarray}
-	\; \; 
-	\begin{blockarray}{cccc}{}
-	\kern-0.2em & a & b &  c \\
-		\begin{block}{c[ccc]}
-  		  & \; 1   &    &   1  \; \topstrut \\
-  		 & \;       &  1  &  1  \; \\
-  		 & \; 1   &   \fbox{1}   &     \; \botstrut \\
-		\end{block}
-	\end{blockarray}
-%
-\; \xrightarrow{P} 
-\begin{blockarray}{cc}{}
-	\kern-0.2em & c \\
-		\begin{block}{c[c]}
-  		  & \; \hphantom{1} \; \topstrut \\
-  		 & \; \hphantom{1} \; \\
-  		 & \; \hphantom{1} \; \botstrut \\
-		\end{block}
-	\end{blockarray}
-\; \;
-\begin{blockarray}{cccc}{}
-	\kern-0.2em & b & c & a  \\
-		\begin{block}{c[ccc]}
-  		  & \;  & 1 &  1\; \topstrut \\
-  		 & \; 1 & 1 &  \; \\
-  		 & \; 1 &  & \fbox{1} \; \botstrut \\
-		\end{block}
-	\end{blockarray}
-%
-\; \xrightarrow{d_R} 
-\begin{blockarray}{cccc}{R_1}
-\kern-0.2em & b & c & a  \\
-	\begin{block}{c[ccc]}
-		u \kern \rowsp & \;  & 1 &  \; \topstrut \\
-		v \kern \rowsp & \; 1 & 1 &  \; \\
-		w \kern \rowsp & \; 1 &  &  \; \botstrut \\
-	\end{block}
-	% \undermat{B}{ & & }  
-\end{blockarray}
-\end{displaymath}
-
-\vspace{-1.0em}
-
-% Start V example 
-\begin{displaymath}
-	\def\rowsp{0.4em}
-	\def\arraystretch{0.95}
-	\makeatletter\setlength\BA@colsep{1.2pt}\makeatother
-	%
-	\begin{blockarray}{cc}{d_V}
-	\kern-0.2em & a \\
-		\begin{block}{c[c]}
-  		 & \; 1 \; \topstrut \\
-  		& \;  \; \\
-  		& \;  \; \botstrut \\
-		\end{block}
-	\end{blockarray}
-	\; \;
-	\begin{blockarray}{cccc}{V_1}
-	 \kern-0.2em & a & b & c  \\
-		\begin{block}{c[ccc]}
-  		 & \; 1 & \fbox{1} &  \fbox{1}\; \topstrut \\
-  		& \;  & 1 &  1\; \\
-  		& \; &  & 1 \; \botstrut \\
-		\end{block}
-	\end{blockarray}
-	%
-\; \to
-	%
-\begin{blockarray}{cc}{}
-	\kern-0.2em & b \\
-		\begin{block}{c[c]}
-  		 & \; 1 \; \topstrut \\
-  		& \;  1 \; \\
-  		& \; \; \botstrut \\
-		\end{block}
-	\end{blockarray}
-	\; \;
-\begin{blockarray}{cccc}{}
-	 \kern-0.2em & a & b & c  \\
-		\begin{block}{c[ccc]}
-  		 & \; 1 & &  \fbox{1}\; \topstrut \\
-  		& \;  & 1 &  1\; \\
-  		& \; &   & 1 \; \botstrut \\
-		\end{block}
-	\end{blockarray}
-	%
-\; \to
-	%
-	\begin{blockarray}{cc}{}
-	\kern-0.2em & c \\
-		\begin{block}{c[c]}
-  		 & \; 1 \; \topstrut \\
-  		& \;  1 \; \\
-  		& \; 1 \; \botstrut \\
-		\end{block}
-	\end{blockarray}
-	\; \;
-	\begin{blockarray}{cccc}{}
-	\kern-0.2em & a & b &  c \\
-		\begin{block}{c[ccc]}
-  		 & \; 1 & &   \; \topstrut \\
-  		& \;  & 1 &   \; \\
-  		& \;  &   & 1 \; \botstrut \\
-		\end{block}
-	\end{blockarray}
-%
-\; \xrightarrow{P} \\
-% 
-\begin{blockarray}{cc}{}
-	\kern-0.2em & a \\
-		\begin{block}{c[c]}
-  		 & \; 1 \; \topstrut \\
-  		& \;  1 \; \\
-  		& \; 1 \; \botstrut \\
-		\end{block}
-	\end{blockarray}
-\; \;
-\begin{blockarray}{cccc}{}
-	\kern-0.2em & b & c & a  \\
-		\begin{block}{c[ccc]}
-  		 & \; 1 & &\; \topstrut \\
-  		& \;  & 1 &  \; \\
-  		& \; &  & 1 \; \botstrut \\
-		\end{block}
-	\end{blockarray}
-%
-\; \xrightarrow{d_V}
-% 
-	\begin{blockarray}{cccc}{V_1}
-	\kern-0.2em & b & c & a  \\
-		\begin{block}{c[ccc]}
-  		b \kern \rowsp & \; 1 & & 1 \; \topstrut \\
-  		c \kern \rowsp & \; & 1 &  1 \; \\
-  		a \kern \rowsp & \; &  & 1  \; \botstrut \\
-		\end{block}
-	\end{blockarray}
-\end{displaymath}
-\\
+
 \noindent
@@
 \subsection{Proposed approach}
-Let us begin with a briefly overview of the pipeline, which we outline in Algorithm~\ref{alg:schedule}.
-As before, we assume as input  a discrete $1$-parameter family of filtrations $\mathcal{K} = (K_1, K_2, \dots, K_n)$,
+Let us begin with a brief overview of the pipeline, which we also outline in Algorithm~\ref{alg:schedule} below.
+As before, we assume as input  a discrete $1$-parameter family of filtrations $\mathcal{K} = (K_1, K_2, \dots, K_n)$ of an abstract simplicial complex $K$ with $|K|=m$, 
 and the goal being to compute their persistence diagrams.
-
-Without loss of generality, we may assume $\lvert K_i \rvert = m$ for every $i \in [n]$, 
-such that there are  bijections $f_i : K_i \to K_{i+1}$. 
-Otherwise, if $\lvert K_i \rvert \neq \lvert K_{i+1} \rvert$,   
-we may append $K_{i+1} \smallsetminus K_i$ to the end of $K_i$ and likewise append $K_{i} \smallsetminus K_{i+1}$ to the end of $K_{i+1}$ prior to constructing the bijection. 
-At the matrix level, both concatenations can be achieved by appending the boundary chains of the simplices being inserted to $R$, and then running Algorithm~\ref{alg:reduce} only on the appended columns.
+The pipeline can be summarized as follows.
+
+Fix reindexing  bijections $f_i : K_i \to K_{i+1}$. 
 Each $f_i$ induces a bijection $f_i^\ast : [m] \to [m]$, or equivalently, a permutation of the index set $[m]$.
@@
 We compute the longest increasing subsequence (LIS) of $q$ and use this subsequence to recover a longest common subsequence (LCS) $\mathrm{LCS}(p,q)$.
-We pass $p$, $q$, and this LCS to our scheduling algorithm, which returns as output an ordered set of move permutations.
-We collect all of these permutations into a single schedule $\mathcal{S}$, obtain an $R=DV$ decomposition for the first filtration $K_1$, and then execute the permutations stored in $\mathcal{S}$ in sequence.
-During execution, we obtain a valid decomposition for each filtration $K_1, K_2, \dots, K_n$.
-It turns out that not all valid schedules are created equal:
+We pass $p$, $q$, and this LCS to our greedy scheduling algorithm, which returns as output an ordered set of move permutations  $\mathcal{S}$ of minimum size, which we call a \emph{schedule}.
+Not all valid schedules are created equal:
 some incur a much larger number of column operations,
@@
 Section~\ref{sec:schedule_cost}.
+We proceed to computing an $R=DV$ decomposition for the first filtration $K_1$, and then execute the moves stored in $\mathcal{S}$ in sequence, extracting their corresponding persistence diagrams along the way.
+During execution, we obtain a valid decomposition $R = DV$ for each filtration $K_1, K_2, \dots, K_n$.
 
@@
 
-\subsubsection{The Continuous Setting}\label{sec:continuous_setting}
-In the vineyards case, we are given a homotopy $F : K \times [0,1] \to \mathbb{R}$ and a pair of filtrations $K_0, K_1$ of $K$, with simplices  ordered according to $f_0 = F(\cdot, 0)$ and $f_1 = F(\cdot, 1)$, respectively.
-The choice of homotopy $F$ completely determines the number of adjacent transpositions, so to get a meaningful bound we must know something about $F$.
-If we assume that the curves $\tau \mapsto F(\cdot, \tau)$  are in general position in $[0,1]\times \mathbb{R}$, and that each pair of curves cross  at most once, then the set of adjacent transpositions $S_F$ that $F$ decomposes into is given by the pairs of simplices with changes in their relative ordering:
+\subsubsection{The Vineyards Setting}\label{sec:continuous_setting}
+In the vineyards (i.e., continuous time) case, we are given a homotopy $F : K \times [0,1] \to \mathbb{R}$ interpolating between  a pair of filtrations $K_0, K_1$ of $K$, with simplices  ordered according to $f_0 = F(\cdot, 0)$ and $f_1 = F(\cdot, 1)$, respectively.
+The choice of homotopy $F$ completely determines the number of adjacent transpositions that transform $K_0$ into $K_1$ via $F$, so to get a meaningful bound we must know something about $F$. 
+If we assume that the curves $\tau \mapsto F(\cdot, \tau)$  are in general position in $[0,1]\times \mathbb{R}$, and that each pair of curves cross  at most once, then the set of adjacent transpositions $S_F$ that $F$ decomposes into is given by the pairs of simplices in $K$ with changes in their relative ordering:
 \begin{align}\label{eq:sf_schedule}
@@
 %We assume, without loss of generality, that the underlying complexes $K_0$ and $K_1$ have the same cardinality. 
-Since we deal exclusively with essential filtrations, we may think of $K_1$ as simply a reordering of $K_0$ (or vice versa).
-Let $p = [m]$ and let $q = \mathrm{Im}\,f^\ast$, where $f^\ast : [m] \to [m]$ is the one-to-one correspondence induced by the bijection $f: K_0 \to K_1$.
+%Since we deal exclusively with essential filtrations, we may think of $K_1$ as simply a reordering of $K_0$ (or vice versa). 
+Let $p = [m]$ and let $q = \mathrm{Im}\,f^\ast$, where $f^\ast : [m] \to [m]$ is the one-to-one correspondence induced by the reindexing bijection $f: K_0 \to K_1$.
 Then $S_F$ is given by the inversions between $p$ and $q$:
-\[
+\[S_F = 
 \mathrm{Inv}(p, q) = \{ \, (i,j) \, | \, p(i) < p(j) \mbox{ and } q(i) > q(j)  \quad \; i,j \in [m]\, \}
@@
 \end{equation}
-Geometrically, our assumptions on $F$ defines a class of $x$-monotone curves called \emph{pseudo-segments}, i.e.
-parameterized curves whose behavior with respect to a certain restricted set of geometric predicates is invariant.
+Geometrically, our assumptions on $F$ imply that the  functions $\tau\mapsto F(\cdot, \tau)$ belong to a class of $x$-monotone curves called \emph{pseudo-segments}.
+%, i.e. parameterized curves whose behavior with respect to a certain restricted set of geometric predicates is invariant. 
 This family includes 
@@
  This quadratic scaling can induce a number of issues in practical implementations of vineyards.
- Indeed, although many transpositions require $0$ column reductions in practice, detecting the existence of non-zero entries at specific positions in the sparse matrix takes $O(\log(m))$ time.
-Since this check is required after each permutations are applied, it is in some sense unavoidable in the vineyards algorithm.
-
-The ordered set of inversions $(i,j) \in S_F$ can be interpreted as permutations to apply to the decomposition.
-We call $S_F$ a \emph{schedule} with respect to $F$.
+ Indeed, although many transpositions require no column reductions in practice, detecting the existence of non-zero entries at specific positions in the sparse matrix takes $O(\log(m))$ time. 
+ Since this check is required after each permutation is applied, it is in some sense unavoidable in the vineyards algorithm.
+
+The ordered set of inversions $(i,j) \in S_F$ can be interpreted as permutations to apply to the decomposition, and hence 
+can be thought of as a \emph{schedule} with respect to $F$. 
 If our goal is to decrease the size of $\lvert S_F \rvert$, one option is to \emph{coarsen} $S_F$ to a new schedule $\widetilde{S}_F$ by collapsing contiguous sequences of adjacent transpositions to moves, via the map:
@@
 \begin{proposition}\label{prop:factor2}
-	Let $R = D V$ be a valid decomposition of size $(m \times m)$, let $C$ denote the number of $O(m)$ operations required to execute schedule $S$, and similarly let $L$ be the number of $O(m)$ operations required to execute a coarsened schedule $\widetilde{S}$ of $S$.
+	Let $R = D V$ be a valid decomposition of size $(m \times m)$, let $C$ denote the number of $O(m)$ operations required to execute schedule $S_F$, and similarly let $L$ be the number of $O(m)$ operations required to execute a coarsened schedule $\widetilde{S}_F$ of $S_F$.
 We have the following relationship between $C$ and $L$:
@@
 \begin{proof}
-Consider executing the vineyards algorithm on a given schedule $S$.
+Consider executing the vineyards algorithm on a given schedule $S_F$.
 	%Consider Algorithms~\ref{alg:tr} and~\ref{alg:mr}, where Algorithm~\ref{alg:tr} is used to execute a given schedule $S$ and Algorithm~\ref{alg:mr} is used to execute its coarsened version $\widetilde{S}$.
@@
 In the vineyards algorithm, there are at most $2$ column operations for a given adjacent transposition per matrix.
-	Thus, for any contiguous sequence of transpositions $(t_i, t_{i{+}1}, \dots, t_j)$ we require at most $2(\lvert i - j \rvert)$ column operations in both $R$ and $V$, giving a total of $4(\lvert i - j \rvert)$ column operations.
+	Thus, for any contiguous sequence of transpositions $(i,i+1), (i+1,i+2), \ldots (j-1,j)$ we require at most $2(\lvert i - j \rvert)$ column operations in both $R$ and $V$, giving a total of $4(\lvert i - j \rvert)$ column operations.
 
@@
 
-\subsubsection{The Discrete Setting} 
-At the finest scale, using move operations simulate persistence across $F : K \times [0,1] \to \mathbb{R}$ reduces to vineyards (i.e. when the moves themselves are adjacent transpositions).
-The benefit of move operations is that one need not maintain a valid decomposition during the movement; thus, it is worth investigating whether the vineyard bounds can be improved by relaxing the assumption that one wants to simulate persistence across $F$ continuously.
+\subsubsection{The Moves Setting} 
+As we mentioned before, vineyards maintains 
+$j-i$ valid decompositions while 
+performing the sequence 
+$(i,i+1), (i+1,i+2), \ldots, (j-1,j)$
+of adjacent transpositions. 
+The benefit of move operations is that one need not maintain a valid decomposition during the movement $(i,j)$, only at the end; thus, it is worth investigating whether the vineyard bounds can be improved.
 In the case of vineyards, the bound is clear: 
-if $\kappa = K_\tau(K_0, K_1)$, then $\Omega(\kappa)$ adjacent transpositions are required to transform $K_0 \mapsto K_1$. 
+if $\kappa = K_\tau(K_0, K_1)$ is the Kendall-$\tau$ distance, then $\Omega(\kappa)$ adjacent transpositions are required to transform $K_0 \mapsto K_1$. 
 In contrast, if we allow move operations in any order, we have $O(m)$ as a trivial upper bound: simply move each simplex in $K_0$ into its position in the filtration given by $K_1$ in the order given by the latter.
@@
 \]
-In the context of edit operations, observe that a move operation can be interpreted the composition of deletion operation followed by an insertion operation: 
+In the context of edit operations, observe that a move operation can be interpreted as the composition of a deletion operation followed by an insertion operation: 
 $$ m_{ij} = (\mathrm{ins}_j \circ \mathrm{del}_i) $$
-where $\mathrm{del}_i$ denotes the edit operation that deletes the character at position $i$ in some string $p$ and $\mathrm{ins}_j$ similarly inserts the same character at position $j$.
+where $\mathrm{del}_i$ denotes the edit operation that deletes the character at position $i$  and $\mathrm{ins}_j$ similarly inserts the same character at position $j$. 
 Thus, sorting by move operations can be interpreted as finding a minimal sequence of edits where the only operations allowed are insertions and deletions, where each insertion/deletion operation is coupled.
@@
 \begin{proposition}[Schedule size]
-Given filtrations $K_0$ and $ K_1$ of size $m$, and an order preserving bijection $f: K_0 \to K_1$, 
-then the smallest move schedule $S^\ast$ is of size $\Omega(d)$, where:
+Let $K_0$ and $ K_1$ be filtrations of a simplicial complex $K$ with  $m$ simplices.
+Then, a smallest size  move schedule $S^*$ transforming $K_0$ into $K_1$ has cardinality 
 $$ d = m - \lvert \mathrm{LCS}(K_0, K_1) \rvert $$ 
@@
 Recall our definition of  edit distance given above, 
-depending on the choice of $S \subseteq S_m$ of allowable edit operations, and that in order for any edit distance to be symmetric, if $s \in S$ then $s^{-1} \in S$.
-This implies that $d_S(p,q) = d_S(p^{-1}, q)$ for any choice of  $p,q \in S_m$.
+depending on the choice   $\Sigma \subseteq S_m$ of allowable edit operations, and that in order for any edit distance to be symmetric, if $s \in \Sigma$ then $s^{-1} \in \Sigma$.
+This implies that $d_\Sigma(p,q) = d_\Sigma(p^{-1}, q)$ for any choice of  $p,q \in S_m$.
 Moreover, edit distances are \emph{left-invariant}, i.e.
 \[
-d_S(p,q) = d_S(r \circ p, r \circ q) \quad \text{ for all } p,q,r \in S_m
+d_\Sigma(p,q) = d_\Sigma(r \circ p, r \circ q) \quad \text{ for all } p,q,r \in S_m
 \]
@@
 Thus, the following identity always holds:
-$$ d_S(p,q) = d_S(\iota, p^{-1} \circ q) = d_S(q^{-1} \circ p, \iota) $$
+$$ d_\Sigma(p,q) = d_\Sigma(\iota, p^{-1} \circ q) = d_\Sigma(q^{-1} \circ p, \iota) $$
 where $\iota = [m]$, the identity permutation.
@@
 \noindent The upper bound from Corollary~\ref{cor:expectation} captures the size of $S^\ast$ under worst-case conditions. 
-It is worth noting that this bound applies to pairs of uniformly sampled permutations, as opposed to uniformly sampled distinct filtrations---intuitively, simplicial filtrations have more structure due to the subcomplex relation that must be respected in order to qualify as a filtration.
+It is worth noting that this bound applies to pairs of uniformly sampled permutations, as opposed to uniformly sampled  filtrations---simplicial filtrations have more structure due to the subcomplex relation that must be respected.
 Thus, while $O(m - \sqrt{m})$ in expectation is an improvement over $O(m)$, it is still not tight.
 However, Boissonnat~\cite{boissonnat2018efficient} has shown that for any $d$-dimensional simplicial complex $K$ with $m$ simplices, the number of distinct filtrations built from $K$ is \emph{at least} $\floor{\frac{t+1}{d+1}}^m$, where $t$ is the number of distinct filtration values associated with the complex $K$.
-In the case where $t \sim O(m)$ and $d << m$ fixed, this lower bound grows similarly to $m!$---thus, $O(m - \sqrt{m})$ is not too pessimistic of an upper bound when the pair of filtrations are no more similar than two random filtrations.
+In the case where $t \sim O(m)$ and $d << m$ fixed, this lower bound grows similarly to $m!$---thus, $O(m - \sqrt{m})$ is not too pessimistic of an upper bound when the pair of filtrations are no more similar than two random permutations.
 In practice, when one has a time-varying filtration and the sampling points are relatively close [in time], the LCS between adjacent filtrations is expected to be much larger, subsequently making $d$ much smaller.
@@
     	\Function{LCS-Sort}{$p$, $q$}
-    		\State $\bar{p} \gets q' \circ p$
-    		\State $\mathcal{L} \gets \mathrm{LCS}(p,q) = \mathrm{LIS}(\bar{p})$ \Comment{$O(\log \log m)$}
+    		\State $\bar{p} \gets q^{-1} \circ p$
+    		\State $\mathcal{L} \gets \mathrm{LCS}(p,q) = \mathrm{LIS}(\bar{p})$ \Comment{$O(m\log \log m)$}
     		\State $(\, \mathcal{S}, \, \mathcal{D}, \, \mathcal{T} \, )\gets (\, \emptyset, \, \bar{p} \setminus \mathcal{L}, \, \mathcal{L} \,)$
@@
     		    %\State $i, i_p, i_n \gets \mathcal{I}_p(\sigma), \, \mathcal{I}_p(T_{\mathrm{pred}}(\sigma)), \, \mathcal{I}_p(T_{\mathrm{succ}}(\sigma))$
-    		    \State $(\,i, \,i_p, \, i_n \,) \gets ( \, \bar{p}^{-1}(\sigma), \, \bar{p}^{-1}(T_{\mathrm{pred}}(\sigma)), \, \bar{p}^{-1}(T_{\mathrm{succ}}(\sigma)) \, )$%\Comment{$O(\log \log m)$}
+    		    \State $(\,i, \,i_p, \, i_n \,) \gets ( \, \bar{p}^{-1}(\sigma), \, \bar{p}^{-1}(\mathcal{T}_{\mathrm{pred}}(\sigma)), \, \bar{p}^{-1}(\mathcal{T}_{\mathrm{succ}}(\sigma)) \, )$\Comment{$O(\log \log m)$}
 %    		    \If{ $p_{\sigma} < p_{\tau}$ }
@@
     		    \EndIf
-    		    \State $(\, \mathcal{S}, \mathcal{D}, \mathcal{T} \,) \gets (\, \mathcal{S} \cup (i, \, j), \, \mathcal{D} \setminus \sigma, \, \mathcal{T} \cup \sigma \, )$%\Comment{$O(\log \log m)$}
+    		    \State $(\, \mathcal{S}, \mathcal{D}, \mathcal{T} \,) \gets (\, \mathcal{S} \cup (i, \, j), \, \mathcal{D} \setminus \sigma, \, \mathcal{T} \cup \sigma \, )$\Comment{$O(\log \log m)$}
     		    % \State $p, p^{-1} \gets m_{ij} \circ p, p^{-1} \circ m_{ij}^{-1}$
-    		    \State $\bar{p}^{-1} \gets \bar{p}^{-1} \circ m_{ij}^{-1}$ %\Comment{$O(m)$}
+    		    \State $\bar{p}^{-1} \gets \bar{p}^{-1} \circ m_{ij}^{-1}$ \Comment{$O(m)$}
     		 \EndWhile
@@
 Any sequence of moves which increases the LCS on each move is guaranteed to have size $m - \lvert \mathrm{LCS}(K_0, K_1) \rvert$, and the reduction to the permutation edit distance problem ensures this size is optimal.
-However, like in the vineyards algorithm, certain pairs of simplices cost more to exchange depending on whether they are critical pairs in the sense described~\cite{cohen2006vines}. 
+However, like in the vineyards algorithm, certain pairs of simplices cost more to exchange depending on whether they are critical pairs in the sense described in~\cite{cohen2006vines}. 
 This small variation in cost-per-transposition induces a large variability in the cost of randomly generated schedules, which is computationally undesirable in practice.
@@
 $$ K_1 = \{ {\color{red} a \; b \; c \; d } \; x \; y \; z \; {\color{red} u \; v \; w } \}  = { {\color{red} 1 \; 2 \; 3 \; 4 } \;8 \; 9 \; 10 \; {\color{red} 5 \; 6 \; 7 } }$$
-The subset of the filtration which corresponds to the simplices which lie in the LCS between these permutations is colored in red.
-For this example, the permutation edit distance is $d = m - \mathrm{LCS}(K_0, K_1)$ implies exactly $3$ moves are needed to apply the mapping $K_0 \mapsto K_1$.
-There are six possible valid schedules of moves:
+The subset of the filtration which corresponds to the simplices which lie in the LCS between these permutations is colored in red. For this example, the permutation edit distance is 
+$d = m - |\mathrm{LCS}(K_0, K_1)|$ implies that exactly $3$ moves are needed to apply the mapping $K_0 \mapsto K_1$. There are six possible valid schedules of moves: 
 \begin{alignat*}{3}
@@
 However, when $d \sim O(m)$, the sorting procedure shown using Algorithm~\ref{alg:sorting} approaches $O(m^2 \log m)$.
-We seek a fast procedure for generating schedules; if the schedule generation procedure exhibits a complexity of $\approx O(m^3)$ or higher, then such a procedure would offers no computational benefit compared to naively reduction algorithm.
+We seek a fast procedure for generating schedules; if the schedule generation procedure exhibits a complexity of $\approx O(m^3)$ or higher, then such a procedure would offer no computational benefit compared to the reduction algorithm.
 Indeed, observe the vineyards algorithm requires $O(m \log m + k)$ preprocessing time to compute a schedule of size $k \sim O(m^2)$ transpositions, thus for Algorithm~\ref{alg:schedule} to be considered a more efficient alternative to vineyards, a solution which is subquadratic in $m$ is desirable.
@@
 In practice, due to the sparsity of both $R$ and $V$, this bound is very loose---the number of column operations is generally much smaller.
-Nonetheless, the complexities of the move operations discussed in Section~\ref{sec:moves} are dependent on the magnitude of $\lvert i - j \rvert$, and thus it's worth considering whether  \emph{minimizing the total displacement} would be a effective proxy objective.
-
-If one forms bijections between two given filtrations $p: K_0 \to [m], q: K_1 \to [m]$ to permutations of $[m]$ as before, then the Kendall-$\tau$ distance yields upper bound on the number of column reductions needed to perform \emph{any} schedule.
+Nonetheless, the complexities of the move operations discussed in Section~\ref{sec:moves} are dependent on the magnitude of $\lvert i - j \rvert$, and thus it is worth considering whether  \emph{minimizing the total displacement} would be a effective proxy objective.
+
+If one forms bijections between two given filtrations $p: K_0 \to [m], q: K_1 \to [m]$ to permutations of $[m]$ as before, then the Kendall-$\tau$ distance yields an upper bound on the number of column reductions needed to perform \emph{any} schedule.
 This follows from the fact that any inversion between $p$ and $q$ may induce a column operation, and the number of inversions between $p$ and $q$ is exactly $K_{\tau}(p, q)$.
-Another common distance for measuring the disarrangement between permutations is \emph{Spearman distance}, defined as:
+Another common distance for measuring the disarrangement between permutations is the \emph{Spearman distance}, defined as:
 \begin{equation}\label{eq:spearman_dist}
@@
 Although very similar to the Kendall-$\tau$ distance, the Spearman distance conveys certain computational advantages not inherited by the Kendall distance due to its simplicity. 
-
-For example, the Spearman distance can be computed $O(m)$ time, and 
-For example, one of the tasks of the \emph{rank aggregation} problem is to obtain an  aggregation~\cite{} $\sigma \in S_m$ from a set of $k$ rankings $T = \{\tau_1, \tau_2, \dots, \tau_k\}$ that obeys some optimality criteria.
+For example, the Spearman distance can be computed in $O(m)$ time.
+
+%, and 
+%For example, one of the tasks of the \emph{rank aggregation} problem is to obtain an  aggregation~\cite{} $\sigma \in S_m$ from a set of $k$ rankings $T = \{\tau_1, \tau_2, \dots, \tau_k\}$ that obeys some optimality criteria.
 A Kemeny optimal ranking is a permutation $\sigma^\ast$ that minimizes the sum of Kendall distances to the gievn set of rankings:
-$$ \argmin_{\sigma \in S_m} = \sum\limits_{i = 1}^k K_\tau(\sigma, \tau_i)$$ 
-Computationally, obtaining a Kemeny optimal ranking is NP-hard\cite{}.
+%$$ \argmin_{\sigma \in S_m} = \sum\limits_{i = 1}^k K_\tau(\sigma, \tau_i)$$ 
+%Computationally, obtaining a Kemeny optimal ranking is NP-hard\cite{}.
 In contrast, a 2-approximation may be obtained by replacing the Kendall distance with the Spearman distance, as in this case the problem reduces to finding a perfect matching in a bipartite graph, which can be accomplished in polynomial time.
 
-We consider the difficulty of obtaining a sorting $p \stackrel{S}{\mapsto} q$ of size $d = m - \mathrm{LCS}(p,q)$ which has minimal total displacement.
+Next we consider the difficulty of obtaining a sorting $S$ of size $d = m - \mathrm{LCS}(p,q)$ which has minimal total displacement. 
+That is, incorporating the Spearman distance in the proxy objective. 
 Formally, given $p, q \in S_m$ and some sorting  $S = (s_1, \dots, s_d)$, let $\hat{s}_i =  s_i \circ \cdots \circ s_2 \circ s_1$ denote the composition of the first $i$ permutations in $S$, where $i < d$.
@@
 \end{equation}
-where equality is only achieved with schedules $S$ for which each symbols displacement between its initial position in $p$ and its final position in $q$ is non-increasing with every application of $s_i$.
+where equality is only achieved with schedules $S$ for which the displacement of each symbol between its initial position in $p$ and its final position in $q$ is non-increasing with every application of $s_i$.
 Note this not guaranteed if $S$ was derived using the strategy from section~\ref{sec:move_schedules}.
@@
 If the $(p, m_{ij} \circ p)$ are drawn as a bipartite graph, with the elements of $p$ in one layer and the elements of $m_{ij} \circ p$ in another, then observe that $F(p, m_{i,j} \circ p)$ is twice the number of edge crossings in the graph.
-We given an example below to show this.
+We given an example below illustrating this.
 \\
 \\
-\noindent \textbf{Example:} An example of three possible schedules, $S_1$, $S_2$, and $S_3$ is given in Figure~\ref{fig:crossings}. 
+\noindent \textbf{Example:} 
+Let  $p= (1\;2\;3\;4\;5\;6\;7\;8\;9)$ 
+and 
+$q = (9\;4\;2\;7\;1\;8\;6\;3\;5)$.
+An example of three possible schedules, $S_1$, $S_2$, and $S_3$ sorting $p$ into $q$ is given in Figure~\ref{fig:crossings}. 
 %Each schedule, from left to right, has $21$, $31$, and $35$ crossings, respectively. 
-Each column represents the successive application of permutations from the three fixed schedules. Black/red vertices correspond to symbols in and outside of $\mathrm{LCS}(p,q)$, respectively. 
-\begin{figure}[ht]
+\begin{figure}[!htb]
     \centering
-    \includegraphics[width=0.60\textwidth]{crossings.png}
+    \includegraphics[width=0.80\textwidth]{crossings.png}
     \caption{Three sortings, $S_1$, $S_2$, and $S_3$, with $21$, $31$, and $35$ crossings, resp.}
@@
 %The greedy heuristic proposed to minimize equation~\ref{eq:schedule_opt_succ} was used to produce $S_1$, which has 21 edge crossings.
-\noindent 
+
+\noindent  Each column represents the successive application of a move $m_{ij}$ in the schedule, 
+and the edges track how each  symbol has been moved. 
+Black/red vertices correspond to symbols in and outside of $\mathrm{LCS}$, respectively. 
 All three schedules were generated from the same $\mathrm{LCS}(p,q) = (4\,7\,8)$ and all three schedules transform $p \mapsto q$ in $d = 6$ moves. 
-In this example, this matches the minimal number of crossings amongst all possible schedules in fact, since $K_\tau(p,q) = 21$.
-When one does not need the sortings to be valid, this problem reduces to the $k$-layer crossing minimization problem for $k$ sets of permutations, which is NP-hard for $k \geq 4$~\cite{biedl2009complexity}.
+In this example, $S_1$ matches the minimal number of crossings amongst all possible schedules,  since $K_\tau(p,q) = 21$. \\ \\
+In general,  the $k$-layer crossing minimization problem for $k$ sets of permutations  is NP-hard for $k \geq 4$~\cite{biedl2009complexity}.
 As a result, we do not expect there to be a polynomial time algorithm for optimally achieving this proxy objective.
@@
 %% TODO: elaborate on monotonicity, describe segment tree
-
-The simplicity of both the Spearman distance and the types of permutations we consider here enables certain computational advantages.
+However, the simplicity of both the Spearman distance and the types of permutations we consider here enables certain computational advantages.
 Suppose we begin with an array $\mathcal{A}$ of size $m$ which provides $O(1)$ access and modification, initialized with the (signed) \emph{displacement} of every element in $p$ to its corresponding position in $q$.
 Since the Spearman distance is simply the sum of the absolute value of these displacements, at any point during during the execution of Algorithm~\ref{alg:sorting} we may obtain $F(\bar{p}, q)$ simply by having access to the sum of every entry in $\mathcal{A}$.
-There are two degrees of freedom in the Algorithm~\ref{alg:sorting}: the first is in the choice of $\sigma \in \mathcal{D}$, and the second is in choosing $j$.
+There are two degrees of freedom in   Algorithm~\ref{alg:sorting}: the first is in the choice of $\sigma \in \mathcal{D}$, and the second is in choosing $j$.
 If we fix a heuristic for the latter, then we have a set of possible valid moves $S_\mathcal{D}$ induced by each $\sigma \in \mathcal{D}$ to choose from.
@@
 \]
-That is, if $s_\sigma$ represents a move permutation that moves $\sigma$ from position $i$ to position $j$, then relative to the previous permutation $p$, the displacement of $\sigma$ in $s_\sigma \circ p$ changes by $\lvert i - j \rvert$ and the displacement of any symbol between $i$ and $j$ changes by $\pm 1$.
+That is, if $s_\sigma$ represents a move permutation moving $\sigma$ from position $i$ to position $j$, then relative to the previous permutation $p$, the displacement of $\sigma$ in $s_\sigma \circ p$ changes by $\lvert i - j \rvert$ and the displacement of any symbol between $i$ and $j$ changes by $\pm 1$.
 All other symbol displacements are unchanged.
-Thus, if we replace $\mathcal{A}$ with a data structure that supports $O(\log m)$ access time to aggregate information (i.e.
+Thus, if we replace $\mathcal{A}$ with a data structure  supporting $O(\log m)$ access time to aggregate information (i.e.
 the sum) and $O(\log m)$ modification time ability on $\lvert i - j \vert$ entries (simultaneously), we could greedily choose the next permutation $s$ to maximize:
@@
 The former problem reduces to the problem of efficiently calculating \emph{prefix sums}, which is easily solved. 
-It's not immediately clear how to achieve the latter modification complexity in $O(\log m)$ time, since $\lvert i - j \rvert \leq m$ is potentially larger than $O(\log m)$.
-%, and
+It is not immediately clear how to achieve the latter modification complexity in $O(\log m)$ time, since $\lvert i - j \rvert \leq m$ is potentially larger than $O(\log m)$. %, and 
 % Can actually only achieve O(d^2 log m) it seems 
-It is known that one can apply constant-factor updates to multiple values in a prefix sum in $O(\log m)$ time by any prefix-sum data structure that also supports \emph{range updates}, e.g.
+It is known that one can apply constant-factor updates to multiple values in a prefix sum in $O(\log m)$ time by any prefix-sum data structure that also supports \emph{range updates}---e.g.
 an \emph{implicit treap}.
@@
 	\centering
-	\includegraphics[width=0.75\textwidth]{boid_sim_results.png}
+	\includegraphics[width=\textwidth]{boid_sim_results.png}
 	\caption{ The cumulative number of $O(m)$ column operations (left, log-scale) of the three approaches tested and the distribution of the normalized Kendall distance between adjacent filtration pairs.
@@
 \subsection*{Multidimensional persistence}
-A natural question is whether or not there exists a multidimensional generalization of persistence.
-Given a structure which filters the space in multiple dimensions simultaneously, a \emph{multifiltration}, the goal of multidimensional persistence is to identify persistent features by examining the entire multifiltration.
+Given a procedure to filter a space in multiple dimensions simultaneously, a \emph{multifiltration}, the goal of multidimensional persistence is to identify persistent features by examining the entire multifiltration.
 Such a generalization has appeared naturally in many application contexts, showing potential as a tool for exploratory data analysis~\cite{lesnick2012multidimensional}. 
-For example, one of the often noted drawbacks of persistence is its instability with respect to strong outliers.
-Although persistence diagrams are stable with respect to their input and thus are robust to noise, the presence of strong outliers can artificially modify the connectivity of the underlying filtration, obscuring the detection of what would otherwise be significant topological structures~\cite{buchet2015topological}.
-Indeed, since inferring the homology of some underlying topological space around which data are distributed was one of the original purposes of persistence\footnote{One of the first uses of persistence was to infer the topology of attractors in dynamical systems from experimental data\cite{perea2018brief}.}, strong outliers pose a significant barrier to the widespread adoption of persistent homology.
-By filtering the space according to e.g.
-some notion of density, multidimensional persistence enables the practitioner to handle outliers directly, without resorting to sophisticated data preprocessing techniques which may or may not preserve the underlying topology of the data.
-
-Unfortunately, multidimensional persistent homology comes with its own set of computational and theoretical challenges not necessarily inherited by the 1-D case.
-There is not complete discrete invariant for multidimensional persistence, and the computations associated to other incomplete invariants exhibit high algorithmic complexity.
+Indeed, one of the often noted drawbacks of persistence is its instability with respect to strong outliers. 
+Although persistence diagrams are stable with respect to their input---in the Hausdorff sense---and thus are robust to Hausdorff noise, the presence of strong outliers can artificially modify the connectivity of the underlying filtration, obscuring the detection of what would otherwise be significant topological structures~\cite{buchet2015topological}.
+%Indeed, since inferring the homology of some underlying topological space around which data are distributed was one of the original purposes of persistence\footnote{One of the first uses of persistence was to infer the topology of attractors in dynamical systems from experimental data\cite{perea2018brief}.}, strong outliers pose a significant barrier to the widespread adoption of persistent homology.
+By filtering the data according to a notion of density, in tandem with distance predicates, multidimensional persistence can enable the practitioner to handle outliers directly.
+%, without resorting to sophisticated data preprocessing techniques which may or may not preserve the underlying topology.
+
+Unfortunately, multidimensional persistent homology comes with its own set of computational and theoretical challenges. 
+%not necessarily inherited by the 1-D case. 
+There is not complete discrete invariant for multidimensional persistence---like the 1D persistence diagram---and the computations associated to other incomplete invariants exhibit high algorithmic complexity.
 A number of computational strategies have been proposed to counter this, such as homology preserving multifiltration simplification or parallelization schemes using shared-memory computation~\cite{fugacci2019chunk}.
@@
 It is a simple and easy to visualize invariant, but is unstable and yields no information about the persistent features of $M$.
-The multigraded Betti numbers constitute a natural and important class of invariants often studied in algebraic geometry commutative algebra, though both their theory and computation is beyond the scope of this effort---the interested reader is referred to~\cite{lesnick2015interactive, carlsson2009theory} for more details.
-For our purposes, the $i^{\text{th}}$-graded Betti number of $M$ is simply a function $\beta_i(M): \mathbb{R}^n \to \mathbb{N}$ whose values $\beta_i^z(M)$ indicate the the number of elements at grade $z$ in a basis of the $i^{\text{th}}$ module in a free resolution for $M$.
+The multigraded Betti numbers constitute a natural and important class of invariants often studied in algebraic geometry and commutative algebra, though both their theory and computation is beyond the scope of this effort---the interested reader is referred to~\cite{lesnick2015interactive, carlsson2009theory} for more details.
+For our purposes, the $i^{\text{th}}$-graded Betti number of $M$ is simply a function $\beta_i(M): \mathbb{R}^2 \to \mathbb{N}$ whose values indicate the the number of elements at  each degree in a basis of the $i^{\text{th}}$ module in a free resolution for $M$.
 The fibered barcode, denoted as $\mathcal{B}(M)$, is defined as follows: 
@@
 	%\in \overline{\mathcal{L}}
-	The fibered barcode $\mathcal{B}(M)$ of a bipersistence module $M$ is the map which sends each line $L$ with non-negative slope to the barcode $\mathcal{B}_L(M)$: 
+	The fibered barcode $\mathcal{B}(M)$ of a 2D persistence module $M$ is the map which sends each   line  $L \subset \mathbb{R}^https://www.overleaf.com/project/60edd57f71696a2859f4cf602$ with non-negative slope to the barcode $\mathcal{B}_L(M)$:
 %	$$ L \mapsto \mathcal{B}_L(M) $$
@@
 One could compute the 1-D persistence associated to some fixed linear combination of two given filter functions; each linear combination parameterizes a line $L \in \mathbb{R} \times \mathbb{R}^{+}$ which $M$ restricts too.
-However, this an $O(m^3)$ time computation, which is prohibitive for exploratory data analysis purposes.
+However, this is an $O(m^3)$ time computation, which is prohibitive for exploratory data analysis purposes. 
 Ideally, one would want to continuously vary $L \in \mathbb{R} \times \mathbb{R}^{+}$ interactively, inspecting the corresponding changes to the barcodes immediately.
@@
 In practice, the external stability result from~\cite{landi2014rank} justifies the use of a simple procedure which approximates the module $M$ with a smaller module $M'$ using a grid-like coarsening procedure.
-Two binning parameters associated with the coarsening procedure, $x_{bin}$ and $y_{bin}$, together determine the degree of the approximation.
-The coarsening procedure allows the practitioner to restrict the size of $\kappa$ to a relatively small constant, which in-turn dramatically reduces the size of $\mathcal{A}(M)$.
+This coarsening procedure allows the practitioner to restrict the size of $\kappa$ to a relatively small constant, which in-turn dramatically reduces the size of $\mathcal{A}(M)$.
 
@@
 
-%
-%\appendix 
-%\section{}\label{sec:appendix}
-%\subsection*{Vineyard Algorithm}
-%\begin{algorithm}[ht]
-%	\caption{Vineyards: Tranposition Framework}\label{alg:tr}
-%	\setstretch{1.15}
-%	\begin{algorithmic}[1]
-%		\Require Matrices $R, V$ satisfying $R = D V$, $1 \leq i \leq m - 1$
-%		\Ensure Output $(R, V)$ maintains the decomposition invariants
-%        \Function{Transpose}{$R$, $V$, $i$}
-%            \State $\mathrm{pos} \gets $ columns satisfying $\mathrm{col}_R = 0$
-%            \If {$\mathrm{pos}[i]$ \textbf{and} $\mathrm{pos}[i+1]$}
-%	            \If{$V[i,i+1] \neq 0$}
-%            		\State $\mathrm{col}_V(i+1)  \mathrel{+}= \mathrm{col}_V(i)$	
-%            	\EndIf
-%            	\If {$\exists \; k, l$ satisfying $\mathrm{low}_R(k) = i, \mathrm{low}_R(l) = i+1$ and $R[i,l] \neq 0$} \Comment{O(m)}
-%            		\If {$k < l$}
-%            			\State \Return $(\, R, \, V \,) \gets (\, PRPS_k^l, \, PVP S_k^l \,)$
-%            		\Else
-%            			\State \Return $(\, R, \, V \,) \gets  (\, PRPS_l^k, \, PVP S_l^k\,) $
-%            		\EndIf
-%            	\EndIf 
-%            \ElsIf{$!\mathrm{pos}[i]$ \textbf{and} $!\mathrm{pos}[i+1]$}
-%            	\If{$V[i,i+1] \neq 0$}\Comment{O(m)}
-%            		\If{$\mathrm{low}_R(i) < \mathrm{low}_R(i+1)$}
-%            			\State \Return  $(\, R, V \,) \gets (P R S^{i+1}_i P, \, P V S^{i+1}_i  P) $
-%            		\Else 
-%            			\State \Return  $(\, R, V \,) \gets (P R S^{i+1}_i P S^{i+1}_i, \, P V S^{i+1}_i P S^{i+1}_i ) $
-%            		\EndIf 
-%            	\EndIf
-%           	\ElsIf{$!\mathrm{pos}[i]$ \textbf{and} $\mathrm{pos}[i+1]$} 
-%            	\If{$V[i,i+1] \neq 0$}\Comment{O(m)}
-%            		\State \Return  $(\, R, V \,) \gets (P R S^{i+1}_i P S^{i+1}_i, \, P V S^{i+1}_i P S^{i+1}_i ) $
-%            	\EndIf
-%            \ElsIf{$\mathrm{pos}[i]$ \textbf{and} $!\mathrm{pos}[i+1]$}
-%            	\If{$V[i,i+1] \neq 0$}\Comment{O(m)}
-%            		\State $\mathrm{col}_V(i+1) \mathrel{+}= \mathrm{col}_V(i)$
-%            	\EndIf 
-%            \EndIf
-%            \State \Return $(\, R , V \,) \gets (\, P R P , P V P \,)$
-%        \EndFunction
-%    \end{algorithmic}
-%\end{algorithm}
-%
-%
-%\subsection*{Transposition complexity details}
-%Consider Algorithm~\ref{alg:tr}.
+\appendix 
+\section{}\label{sec:appendix}
+\subsection*{Vineyard Algorithm}
+\begin{algorithm}[!htb]
+	\caption{Vineyards: Tranposition Framework}\label{alg:tr}
+	\setstretch{1.15}
+	\begin{algorithmic}[1]
+		\Require Matrices $R, V$ satisfying $R = D V$, $1 \leq i \leq m - 1$
+		\Ensure Output $(R, V)$ maintains the decomposition invariants
+        \Function{Transpose}{$R$, $V$, $i$}
+            \State $\mathrm{pos} \gets $ columns satisfying $\mathrm{col}_R = 0$
+            \If {$\mathrm{pos}[i]$ \textbf{and} $\mathrm{pos}[i+1]$}
+	            \If{$V[i,i+1] \neq 0$}
+            		\State $\mathrm{col}_V(i+1)  \mathrel{+}= \mathrm{col}_V(i)$	
+            	\EndIf
+            	\If {$\exists \; k, l$ satisfying $\mathrm{low}_R(k) = i, \mathrm{low}_R(l) = i+1$ and $R[i,l] \neq 0$} \Comment{O(m)}
+            		\If {$k < l$}
+            			\State \Return $(\, R, \, V \,) \gets (\, PRPS_k^l, \, PVP S_k^l \,)$
+            		\Else
+            			\State \Return $(\, R, \, V \,) \gets  (\, PRPS_l^k, \, PVP S_l^k\,) $
+            		\EndIf
+            	\EndIf 
+            \ElsIf{$!\mathrm{pos}[i]$ \textbf{and} $!\mathrm{pos}[i+1]$}
+            	\If{$V[i,i+1] \neq 0$}\Comment{O(m)}
+            		\If{$\mathrm{low}_R(i) < \mathrm{low}_R(i+1)$}
+            			\State \Return  $(\, R, V \,) \gets (P R S^{i+1}_i P, \, P V S^{i+1}_i  P) $
+            		\Else 
+            			\State \Return  $(\, R, V \,) \gets (P R S^{i+1}_i P S^{i+1}_i, \, P V S^{i+1}_i P S^{i+1}_i ) $
+            		\EndIf 
+            	\EndIf
+          	\ElsIf{$!\mathrm{pos}[i]$ \textbf{and} $\mathrm{pos}[i+1]$} 
+            	\If{$V[i,i+1] \neq 0$}\Comment{O(m)}
+            		\State \Return  $(\, R, V \,) \gets (P R S^{i+1}_i P S^{i+1}_i, \, P V S^{i+1}_i P S^{i+1}_i ) $
+            	\EndIf
+            \ElsIf{$\mathrm{pos}[i]$ \textbf{and} $!\mathrm{pos}[i+1]$}
+            	\If{$V[i,i+1] \neq 0$}\Comment{O(m)}
+            		\State $\mathrm{col}_V(i+1) \mathrel{+}= \mathrm{col}_V(i)$
+            	\EndIf 
+            \EndIf
+            \State \Return $(\, R , V \,) \gets (\, P R P , P V P \,)$
+        \EndFunction
+    \end{algorithmic}
+\end{algorithm}
+
+\subsection*{Transposition complexity details}
+Consider Algorithm~\ref{alg:tr}.
 The matrices $R, V$ are both $(m \times m)$, where $m$ is the number of simplices in the underlying filtration.
@@
 
-%The principle motivation  
-%Upon moving $s$ 
-%
-%Denote this class with $c_s$.
-%Now if the position of $s$ in $P_1$ is $i$ and the position in $P_2$ is $j$, and $i < j$, then the
-
-%% Note: patience sorting gives means of enumerating all LIS's
-%% Note #2: Consider having the identity permutation be the target and going in reverse--at the end, there is only one symbol to move into place
-%% prims-MST like approach possible?
-
-%\begin{definition}[Coarsening]
+% The principle motivation  
+% Upon moving $s$
+
+% Denote this class with $c_s$.
+% Now if the position of $s$ in $P_1$ is $i$ and the position in $P_2$ is $j$, and $i < j$, then the
+
+% % Note: patience sorting gives means of enumerating all LIS's
+% % Note #2: Consider having the identity permutation be the target and going in reverse--at the end, there is only one symbol to move into place
+% % prims-MST like approach possible?
+
+% \begin{definition}[Coarsening]
 % 	Given a schedule $\mathcal{S} = ( s_1, s_2, \dots, s_h)$ of permutations, a \textbf{\emph{coarsening}} of $\mathcal{S}$ is a schedule $\widetilde{S} = ( \tilde{s}_1, \tilde{s}_2, \dots, \tilde{s}_l)$ obtained by collapsing contiguous sequences in $S$ via the map
@@
 % 	where, in cycle notation, $t_i$ is a permutation of the form $(i, i{+}1)$, and $m_{ij}$ is a permutation of the form $(i{+}1, \dots, j, i)$.
-%\end{definition}
-%\begin{lemma}
+% \end{definition}
+% \begin{lemma}
 % 	For some given decomposition $R = \partial V$ where each matrix is $(m \times m)$, let $C$ denote the number of $O(m)$ operations required to execute schedule $\mathcal{S} = (t_1, t_2, \dots, t_h)$.
@@
 % 	$$ \frac{C}{2} \leq K \leq C$$ 
-%\end{lemma}
-%\begin{proof}
+% \end{lemma}
+% \begin{proof}
 % 	Consider Algorithms~\ref{alg:tr} and~\ref{alg:mr}, where it's assumed Algorithms~\ref{alg:tr} is used to execute a given schedule $\mathcal{S}$ and Algorithms~\ref{alg:mr} is used to execute its coarsened schedule $\widetilde{\mathcal{S}}$.
@@
 Since a move operation requires at most a single column operation for each index between $[i, j]$, the complexity of $\mathrm{Move}(i,j)$ is bounded above by $O(\lvert i - j \rvert m)$, and the claimed inequality follows.
-%\end{proof}
-%\noindent \textbf{Free Schedules:} Previously, we assumed one had a homotopy $F(x,\tau) : \mathbb{X} \times [0,1] \to \mathbb{R}$ mapping the functions associated with two fixed filtrations $K_0, K_1$ each of size $m$, and we showed that simulating persistence dynamically reduced to determining a set of ordered ``crossings.'' If these crossings occur at time points $(\tilde{t}_1, \tilde{t}_2, \dots, \tilde{t}_m)$, then obtaining the updating decomposition at all times $t_i$ where $\tilde{t}_i < t_i < \tilde{t}_{i{+}1}$ fully characterizes how the persistent homology changes across the homotopy.
+% \end{proof}
+% \noindent \textbf{Free Schedules:} Previously, we assumed one had a homotopy $F(x,\tau) : \mathbb{X} \times [0,1] \to \mathbb{R}$ mapping the functions associated with two fixed filtrations $K_0, K_1$ each of size $m$, and we showed that simulating persistence dynamically reduced to determining a set of ordered ``crossings.'' If these crossings occur at time points $(\tilde{t}_1, \tilde{t}_2, \dots, \tilde{t}_m)$, then obtaining the updating decomposition at all times $t_i$ where $\tilde{t}_i < t_i < \tilde{t}_{i{+}1}$ fully characterizes how the persistent homology changes across the homotopy.
 If we again assume both general position and that the curves traced out by $F$ intersect at most once, then we obtain an ordered schedule of transpositions $\mathcal{S}_F$.
 The size of this schedule is equal to the number of inversions between $K_0$ and $K_1$, which is at most $O(m^2)$.
-%
-%There are many applications wherein a practitioner is interested some notion of dynamic persistence, but only at specific time points---how the persistence changes between these time points may not be needed.
+
+% There are many applications wherein a practitioner is interested some notion of dynamic persistence, but only at specific time points---how the persistence changes between these time points may not be needed.
 More generally, it's been observed in practice that many applications of persistence homology involve computing persistence across discrete 1-parameter family of (usually related) filtrations.
@@
 One solution is to assume a straight-line homotopy $F(x, \tau)$ between each adjacent filtrations $K_{t_i}, K_{t_{i {+} 1}}$ and proceed using any of the methods described above.
-%
-%If one \emph{only} needs the persistence diagrams at exactly the given time points $\mathcal{T}$, i.e.
+
+% If one \emph{only} needs the persistence diagrams at exactly the given time points $\mathcal{T}$, i.e.
 the intermediate persistence diagrams between each $K_{t_i}, K_{t_{i {+} 1}}$ aren't required, then intuitively we ought to coarsen each schedule as much as possible to reduce the number of diagrams generated.
@@
 We call a schedule $\mathcal{S}$ that can be arbritrarily reordered a \emph{free schedule}.
-%
-%\subsection{Minimizing schedule cost}\label{sec:schedule_cost}
-%\textbf{Motivation:}
-%Using the notation presented in the previous section, we now have two goals to consider: 
-%\begin{enumerate}
+
+% \subsection{Minimizing schedule cost}\label{sec:schedule_cost}
+% \textbf{Motivation:}
+% Using the notation presented in the previous section, we now have two goals to consider: 
+% \begin{enumerate}
 % 	\item Given an ordered schedule, can we produce a more efficient schedule using \emph{move} operations?
 % 	\item If we allow any order of transpositions or moves, can we product an optimally efficient schedule?
-%\end{enumerate}
-%We address each of these in order.
+% \end{enumerate}
+% We address each of these in order.
 % \textbf{Example: Coarsening around homological critical points}
-%( Insert more on detecting homological critical points, coarsening around critical points, etc. )
+% ( Insert more on detecting homological critical points, coarsening around critical points, etc. )
 % Although move operations can be thought of as a generalization of the rules given for the vineyards algorithm, there is a significant departure philosophically between what the two algorithms seek to accomplish.

### 2021-11-02 move_schedules.tex
minor changes
Updates from Overleaf

@@
 \begin{document} 
-
 \maketitle
@@
 We then show a modification of this technique which maintains only a sublinear number of valid states, as opposed to a quadratic number of states, and we provide tight lower bounds for this technique.
-	Finally, we present empirical results suggesting that the decrease in operations needed to compute diagrams across a family of filtrations is proportional to the difference between the expected quadratic number of states, and the proposed sublinear coarsening.
-	Applications to multi-dimenisonal persistence and crocker stacks are also presented. 
+	Finally, we present empirical results showing that the decrease in operations needed by our method to compute diagrams across a family of filtrations is proportional to the difference between the expected quadratic number of states, and the proposed sublinear coarsening.
+	Applications to multi-dimensional persistence and crocker stacks are also presented. 
 \end{abstract}
@@
 \begin{keywords}
-  example, \LaTeX
+  Algorithms, 
 \end{keywords}
@@
 % REQUIRED
+% See: https://mathscinet.ams.org/mathscinet/msc/msc2020.html?t=68T09&btn=Current
 \begin{AMS}
-  68Q25, 68R10, 68U05
+  68T09, 55N31, 62R40
 \end{AMS}
@@
 With the current state of the art tools, there is no recourse to the problem of computing persistence from a time-varying filtration containing millions of simplices at thousands of snapshots in time.
-Efficiently  updating the persistence of a time-varying filtration has far-reaching consequences: methods that vectorize persistence diagrams for machine learning purposes, like adaptive template functions~\cite{polanco2019adaptive} and persistence images~\cite{adams2017persistence}, or summary statistics like $\alpha$-smoothed Betti curves~\cite{ulmer2019topological}, all immediately become computationally viable tools in dynamic  settings.
+Nonetheless, having the capability of efficiently updating the persistence of a time-varying filtration has far-reaching consequences: methods that vectorize persistence diagrams for machine learning purposes, like adaptive template functions~\cite{polanco2019adaptive} and persistence images~\cite{adams2017persistence}, or summary statistics like $\alpha$-smoothed Betti curves~\cite{ulmer2019topological}, all immediately become computationally viable tools in dynamic settings.
 %This analogy provides a strong interpretation at gaining insight on the subtle topological changes that may occur in studying a continuous process.
@@
 Despite its theoretical efficiency, vineyards is often not the method of choice in practical settings. 
-Indeed, while there is an increasingly rich ecosystem of software packages offering variations of the standard reduction algorithm (e.g.
+While there is an increasingly rich ecosystem of software packages offering variations of the standard reduction algorithm (e.g.
 Ripser, PHAT, Dionysus, etc.
@@
 many transpositions can be quite slow...
-it is sometimes much faster to simply recompute the $RU$-decomposition from scratch using the standard persistence algorithm.''
-Whether performing the reductions naively or updating the decomposition dynamically using vineyards, they observe that this is the most computationally demanding aspect of  RIVET \cite{rivet}.
+it is sometimes much faster to simply recompute the $RU$-decomposition from scratch using the standard persistence algorithm.'' Indeed, whether performing the reductions naively or updating the decomposition dynamically using vineyards, they observed maintaining the decomposition is the most computationally demanding aspect of RIVET \cite{rivet}.
 
@@
 % in many time varying settings, a valid decomposition is often only sought after at a relatively sparse number of time points relative to the size of the (finite) set needed to generated by the family.
-By exploiting the ``donor'' concept introduced in in~\cite{busaryev2010tracking}, we are able to make a tradeoff in the number of times the decomposition is restored to a valid state, dramatically reducing the total number of column operations needed to apply an arbitrary permutation to the filtration.
-This trade off, paired with a fast greedy heuristic explained in section~\ref{sec:proxy_objective}, yields an algorithm that can update a $R = DV$ decomposition more efficiently than vineyards in coarse contexts, making dynamic persistence more computationally tractable for a wider class of use-cases.
+By exploiting the ``donor'' concept introduced in in~\cite{busaryev2010tracking}, we are able to make a tradeoff between the number of times the decomposition is restored to a valid state and the granularity of the decomposition repair step, dramatically reducing the total number of column operations needed to apply an arbitrary permutation to the filtration.
+This trade off, paired with a fast greedy heuristic explained in section~\ref{sec:proxy_objective}, yields an algorithm that can update a $R = DV$ decomposition more efficiently than vineyards in coarse time-varying contexts, making dynamic persistence more computationally tractable for a wider class of use-cases.
 
@@
 	%\in \overline{\mathcal{L}}
-	The fibered barcode $\mathcal{B}(M)$ of a 2D persistence module $M$ is the map which sends each   line  $L \subset \mathbb{R}^https://www.overleaf.com/project/60edd57f71696a2859f4cf602$ with non-negative slope to the barcode $\mathcal{B}_L(M)$:
+	The fibered barcode $\mathcal{B}(M)$ of a 2D persistence module $M$ is the map which sends each line  $L \subset \mathbb{R}^2$ with non-negative slope to the barcode $\mathcal{B}_L(M)$: 
 %	$$ L \mapsto \mathcal{B}_L(M) $$

### 2021-11-02 move_schedules.tex
Changed conclusion and updated keywords and AMS subject classifications

@@
 % ------  Begin document ------- 
+% https://github.com/peekxc/dart
 \begin{document} 
@@
 \begin{keywords}
-  Algorithms, 
+  Computational topology, Persistent homology, Topological data analysis
 \end{keywords}
@@
 Ripser, PHAT, Dionysus, etc.
-See~\cite{otter2017roadmap} for an overview), implementations of the vineyards algorithm are relatively uncommon.\footnote{Dionysus 1 has an implementation of vineyards, but the algorithm was never ported to the 2nd version.}
+See~\cite{otter2017roadmap} for an overview), implementations of the vineyards algorithm are relatively uncommon.\footnote{Dionysus 1 does have an implementation of vineyards, however the algorithm was never ported to version 2.
+Other major packages, such as GUDHI and PHAT, do not have vineyards implementations.}
 The reason for this disparity is perhaps explained by Lesnick and Wright~\cite{lesnick2015interactive}: ``While an update to an $RU$ decomposition involving few transpositions is very fast in practice...
@@
 This trade off, paired with a fast greedy heuristic explained in section~\ref{sec:proxy_objective}, yields an algorithm that can update a $R = DV$ decomposition more efficiently than vineyards in coarse time-varying contexts, making dynamic persistence more computationally tractable for a wider class of use-cases.
-
- %TODO fill in section reference above.
+The source code containing both the algorithm we propose and the experiments performed in Section~\ref{sec:results} is available online as an open source \texttt{R} package.\footnote{ Code is available here: \url{https://github.com/peekxc/dart}}
 
@@
 greedy-type approaches can lead to arbitrarily bad behavior. 
-In light of these results, we give an alternative proxy-objective for cost minimization, provide bounds justifying its relevance to the original problem, and give an efficient $O(d \log d)$ algorithm for approximately solving this proxy minimization.
+In light of these results, we give an alternative proxy-objective for cost minimization, provide bounds justifying its relevance to the original problem, and give an efficient $O(d^2 \log m)$ algorithm for approximately solving this proxy minimization.
 A performance comparison with other reduction-based persistent homology computations is given; 
@@
     			\State $\mathrm{lis}_q \gets \mathrm{LIS}(q)$ \Comment{$O(m \log \log m)$}
-    			\State $\mathcal{S} \gets \mathcal{S} \cup \mathrm{GreedySchedule}(p, q, \mathrm{lis}_q)$ \Comment{$O(d\log m)$, see section~\ref{sec:proxy_objective}}
+    			\State $\mathcal{S} \gets \mathcal{S} \cup \mathrm{GreedySchedule}(p, q, \mathrm{lis}_q)$ \Comment{$O(d^2\log m)$, see section~\ref{sec:proxy_objective}}
     		\EndFor
@@
 \begin{equation}
-	\mathrm{d}_{\mathrm{lcs}}(p,q) = m + n - 2\lvert LCS(p, q)\rvert
+	\mathrm{d}_{\mathrm{lcs}}(p,q) = m + n - 2\lvert \mathrm{LCS}(p, q)\rvert
 \end{equation}
@@
 \[
-F(s_\sigma \circ p, q) = 
+\mathcal{A}(s_\sigma \circ p) = 
 \begin{cases} 
-	 \mathcal{A}(\sigma) + \lvert i - j \rvert & p^{-1}(\sigma) = i \\
+	 \mathcal{A}(\sigma) \pm \lvert i - j \rvert & p^{-1}(\sigma) = i \\
 	 \mathcal{A}(\sigma) \pm 1 & i < p^{-1}(\sigma) \leq j \\
@@
 The former problem reduces to the problem of efficiently calculating \emph{prefix sums}, which is easily solved. 
-It is not immediately clear how to achieve the latter modification complexity in $O(\log m)$ time, since $\lvert i - j \rvert \leq m$ is potentially larger than $O(\log m)$. %, and 
+It is not immediately clear how to achieve the latter modification complexity in $O(\log m)$ time, since $\lvert i - j \rvert \leq m$ is potentially larger than $O(\log m)$.  
 % Can actually only achieve O(d^2 log m) it seems 
@@
 an \emph{implicit treap}.
-Since single element modifications, removals, and insertions can all be achieved in $O(\log m)$ expected time with this data structure, we conclude that a greedy solution to equation~\ref{eq:greedy_step} can be computed in $O(d^2 \log m)$ time.
+Since single element modifications, removals, and insertions can all be achieved in $O(\log m)$ expected time with such a data structure, we conclude that a greedy solution to equation~\ref{eq:greedy_step} can be computed in $O(d^2 \log m)$ time.
 %Since single element prefix sum modifications can also be achieved $O(\log m)$ time with these data structures, the time to compute $F(\hat{s}_i, \hat{s}_{i+1})$ is bounded above by $O(\log m)$, 
@@
 In conclusion, we presented a scheduling algorithm for efficiently updating a decomposition in coarse dynamic settings.
-Our approach is simple, easy to implement, and fully general: it does not depend on the geometry of underlying space nor the choice of triangulation.
+Our approach is simple, relatively easy to implement, and fully general: it does not depend on the geometry of underlying space, the choice of triangulation, or the choice of homology dimension.
 Moreover, we supplied efficient algorithms for our scheduling strategy, provided tight bounds where applicable, and demonstrated our algorithms performance with several real world use cases.
@@
 \begin{enumerate}
-	\item Computing persistent images for time-varying systems 
+	\item Accelerating PH featurization methods for time-varying systems 
 	\item Optimization procedures involving persistence diagrams
@@
 
-The simple and combinatorial nature of approach does pose some limitations to its applicability.
-There may be better bounds or algorithms available if stronger assumptions can be made on how the filtration is changing.
-If the filtration $K_0$ shares little similarity to the ``target'' filtration $K_1$, then the overhead of reducing the simplices from $K_1 \setminus K_0$ appended to the decomposition derived from $K_0$ may be large enough to motivate simply computing the decomposition at $K_1$ independently, especially if parallel processors are available.
-This overhead does not exist for the multidimensional persistence use-case: the entire decomposition is needed at each template 2-cell.
+The simple and combinatorial nature of our approach does pose some limitations to its applicability.
+For example, there may be better bounds or algorithms available if stronger assumptions can be made on how the filtration is changing with time.
+Moreover, if the filtration $K_0$ shares little similarity to the ``target'' filtration $K_1$, then the overhead of reducing the simplices from $K_1 \setminus K_0$ appended to the decomposition derived from $K_0$ may be large enough to motivate simply computing the decomposition at $K_1$ independently, especially if parallel processors are available.
+That is, our approach is primarily useful if each adjacent filtration in the ordered family of filtration is ``close.''
+Another limitation to the strategy we have here is that we rely completely on the reduction algorithm for appending new simplices to the filtration.
+If the dynamic filtration is constantly both permanently removing and appending new simplices, then the size of the LIS may very well be on the order $O(m - \sqrt{m})$, rendering the efficiencies gained using our approach as neglible.
+Note that this overhead does not exist for the multidimensional persistence use-case: the entire decomposition is needed at each template 2-cell.
 
@@
 Moving forward, our results suggest there are many aspects of computing persistence in dynamic settings yet to be explored.
-For example, it's not immediately clear whether one could adopt, for example, the twist optimization used in the reduction algorithm to the dynamic setting.
-Another avenue of interest is the setting where one has knowledge a priori of the time-varying...
-particular spaces
+For example, it's not immediately clear whether one could adopt, for example, the twist optimization~\cite{chen2011persistent} used in the reduction algorithm to the dynamic setting.
+Another direction to explore would be the analysis of our approach under the cohomology computation~\cite{de2011dualities}, or the specialization of the move operations to specific types of filtrations such as Rips filtrations.
+Such adaptations may result in even larger reductions in the number of column operations, as have been observed in practice for the standard reduction algorithm~\cite{bauer2021ripser}.
 
@@
             	\EndIf
-            	\If {$\exists \; k, l$ satisfying $\mathrm{low}_R(k) = i, \mathrm{low}_R(l) = i+1$ and $R[i,l] \neq 0$} \Comment{O(m)}
+            	\If {$\exists \; k, l$ s.t. $\mathrm{low}_R(k) = i, \mathrm{low}_R(l) = i+1$ and $R[i,l] \neq 0$} \Comment{O(m)}
             		\If {$k < l$}

### 2021-11-02 move_schedules.tex
Review copy !

@@
-\documentclass{siamart190516}
+\documentclass[review]{siamart190516}
 
@@
 
-% REQUIRED - "Abstracts must be able to stand alone and so cannot contain citations to the paper's references, equations, etc. 
+% REQUIRED - "Abstracts must be able to stand alone and so cannot contain citations to the paper's references, equations, etc. .... not exceeding 250 words"
 \begin{abstract}
@@
 	In practice, the quadratic scaling in the number of  such transpositions often makes this maintenance procedure slower than simply computing the decomposition from scratch,  limiting the application of dynamic persistence to relatively small data sets.
-In this work, we propose a coarser strategy for maintaining the decomposition over a discrete 1-parameter family of filtrations.
+In this work, we propose a coarser strategy for maintaining the decomposition over a 1-parameter family of filtrations.
 Our first result is an analysis of a simple linear-time strategy which reduces the number of column operations needed to simulate persistence across a fixed homotopy by at most a factor of 2.
-We then show a modification of this technique which maintains only a sublinear number of valid states, as opposed to a quadratic number of states, and we provide tight lower bounds for this technique.
-	Finally, we present empirical results showing that the decrease in operations needed by our method to compute diagrams across a family of filtrations is proportional to the difference between the expected quadratic number of states, and the proposed sublinear coarsening.
+We show a modification of this technique which maintains only a sublinear number of valid states, as opposed to a quadratic number of states, and we provide tight lower bounds for this technique.
+	Finally, we present results showing that the decrease in operations to compute diagrams across a family of filtrations is proportional to the difference between the expected quadratic number of states, and the proposed sublinear coarsening.
 	Applications to multi-dimensional persistence and crocker stacks are also presented. 
@@
 
+% Keywords and MSC codes must accompany all articles. A list of the subject classifications can be accessed or searched online in the Annual Index of Mathematical Reviews.
 % REQUIRED
@@
 
+% REQUIRED: siamplain format
 \bibliographystyle{siamplain}

### 2021-11-03 move_schedules.tex
Arxiv version

@@
-\documentclass[review]{siamart190516}
-
-% Journal name: SICOMP
+% \documentclass[review]{siamart190516} % for SICOMP 
+\documentclass{article} % for arxiv 
+\usepackage{arxiv}
+
 %\usepackage[margin=0.75in]{geometry}
@@
 
+\pagestyle{plain} % arxiv only
+
 \newcommand\sbullet[1][.5]{\mathbin{\vcenter{\hbox{\scalebox{#1}{$\bullet$}}}}}
@@
 
-%\newtheorem{theorem}{Theorem}
-%\newtheorem{definition}{Definition}
-%\newtheorem{proposition}{Proposition}
-%\newtheorem{lemma}{Lemma}
-%\newtheorem{corollary}{Corollary}
+% For arxiv; comment out for SICOMP
+\newtheorem{theorem}{Theorem}
+\newtheorem{definition}{Definition}
+\newtheorem{proposition}{Proposition}
+\newtheorem{lemma}{Lemma}
+\newtheorem{corollary}{Corollary}
+\newtheorem{proof}{Proof}
 % \newcommand{\CommentX}[1]{\unskip~\%~#1~\%}
 
-\newsiamremark{remark}{Remark}
+%\newsiamremark{remark}{Remark}
 
@@
 % Sets running headers as well as PDF title and authors
-\headers{Fast Persistence Computations}{M. Piekenbrock and J. Perea}
+% \headers{Fast Persistence Computations}{M. Piekenbrock and J. Perea}
 
@@
 % is automatically inserted before the title.
-\title{Move Schedules: Fast persistence computations in coarse dynamic settings\thanks{\textbf{Funding:} This work was partially supported by the National Science Foundation through grants CCF-2006661 and CAREER award DMS-1943758.}
+\title{
+	Move Schedules: Fast persistence computations in coarse dynamic settings
+	\thanks{
+	\textbf{Funding:} This work was partially supported by the National Science Foundation through grants CCF-2006661 and CAREER award DMS-1943758.
+	}
 }
@@
 \hypersetup{
-  pdftitle={Fast Persistence Computations},
+  pdftitle={Move Schedules},
   pdfauthor={Matt Piekenbrock and Jose A. Perea}
@@
 \begin{document}
+
 \maketitle
@@
 \begin{keywords}
-  Computational topology, Persistent homology, Topological data analysis
+  {Computational topology, Persistent homology, Topological data analysis}
 \end{keywords}
@@
 % See: https://mathscinet.ams.org/mathscinet/msc/msc2020.html?t=68T09&btn=Current
-\begin{AMS}
-  68T09, 55N31, 62R40
-\end{AMS}
+%\begin{AMS}
+%  68T09, 55N31, 62R40
+%\end{AMS}
 
@@
   \vspace*{0.8em}
- \begin{enumerate}[leftmargin=2\parindent, align=left, labelsep=-5pt, topsep=0pt,itemsep=-0.25ex,parsep=1.2ex]
-	\item[\hspace{-1.5em}\textbf{ Decomposition Invariants:}]
+  \\
+  \textbf{ Decomposition Invariants:}
+ \begin{enumerate}[labelsep=2pt, topsep=0pt,itemsep=-0.25ex,parsep=1.2ex]
  	\item[I1] $R = D V$ where $D$ is the boundary matrix of the filtration $K$
- 	\item[I2] $V$ is full-rank upper-triangular, and $R$ is \emph{reduced}: if its $i$-th and $j$-th columns are nonzero, then $\mathrm{low}_R(i) \neq \mathrm{low}_R(j)$ 
+ 	\item[I2] $V$ is full-rank upper-triangular, and $R$ is \emph{reduced}: if its $i$-th and $j$-th columns are nonzero, \\ then $\mathrm{low}_R(i) \neq \mathrm{low}_R(j)$ 
  	\end{enumerate} 
@@
 \\	
- \begin{enumerate}[leftmargin=2\parindent, align=left, labelsep=-5pt, topsep=0pt,itemsep=-0.25ex,parsep=1.2ex]
+ \begin{enumerate}[labelsep=5pt, topsep=0pt,itemsep=-0.25ex,parsep=1.2ex]
  	\item[M1:] Querying the non-zero status of entries in either $R$ or $V$ occurs once per move.
@@
 
-\begin{remark}
-We note that while the pseudocode for \emph{MoveRight} is included in~\cite{busaryev2010tracking}, the one for \emph{MoveLeft} is not, and Algorithm~\ref{alg:ml} is the one we propose and leverage in this paper.
-\end{remark}
+%\begin{remark}
+%We note that while the pseudocode for \emph{MoveRight} is included in~\cite{busaryev2010tracking}, the one for \emph{MoveLeft} is not, and Algorithm~\ref{alg:ml} is the one we propose and leverage in this paper.
+%\end{remark}
 

### 2021-11-03 move_schedules.tex
arxiv version
Updates from Overleaf

--- 2021-11-02 move_schedules.tex
@@
 \author{
-	Matt Piekenbrock\thanks{Khoury College of Computer Sciences, Northeastern University.}
+	Matthew Piekenbrock\thanks{Khoury College of Computer Sciences, Northeastern University.}
 	\and Jose A. Perea\thanks{
@@
 \hypersetup{
-  pdftitle={Fast Persistence Computations},
-  pdfauthor={Matt Piekenbrock and Jose A. Perea}
+  pdftitle={Move Schedules},
+  pdfauthor={Matthew Piekenbrock and Jose A. Perea}
 }

### 2021-11-03 move_schedules.tex

--- 2021-11-03 move_schedules.tex
@@
   pdftitle={Move Schedules},
-  pdfauthor={Matt Piekenbrock and Jose A. Perea}
+  pdfauthor={Matthew Piekenbrock and Jose A. Perea}
 }
@@
 % REQUIRED: siamplain format
-\bibliographystyle{siamplain}
+%\bibliographystyle{siamplain}
+\bibliographystyle{unsrt}
 \bibliography{move_schedules}

### 2022-02-07 move_schedules.tex
arxiv version

--- 2021-11-03 move_schedules.tex
@@
+\pdfoutput=1
 % \documentclass[review]{siamart190516} % for SICOMP 
@@
 We pass $p$, $q$, and this LCS to our greedy scheduling algorithm, which returns as output an ordered set of move permutations  $\mathcal{S}$ of minimum size, which we call a \emph{schedule}.
-Not all valid schedules are created equal:
-some incur a much larger number of column operations,
-even if they have a minimal number of moves. 
-Minimizing schedule costs will be addressed in 
-Section~\ref{sec:schedule_cost}.
-We proceed to computing an $R=DV$ decomposition for the first filtration $K_1$, and then execute the moves stored in $\mathcal{S}$ in sequence, extracting their corresponding persistence diagrams along the way.
-During execution, we obtain a valid decomposition $R = DV$ for each filtration $K_1, K_2, \dots, K_n$.
+Not all valid schedules are created equal: some incur a much larger number of column operations, even if they have a minimal number of moves; we address this issue in Section~\ref{sec:schedule_cost}.
+Given the set of schedules for each pair $(K_i, K_{i+1})$, we proceed to computing an $R=DV$ decomposition for the first filtration $K_1$, and then execute the moves stored in $\mathcal{S}$ in sequence.
+During execution, we obtain a valid decomposition $R = DV$ for each filtration $K_1, K_2, \dots, K_n$ from which we extract persistence diagrams along the way.
 

### 2022-02-07 move_schedules.tex
DC&G prelim reformat

@@
-\pdfoutput=1
 % \documentclass[review]{siamart190516} % for SICOMP 
-\documentclass{article} % for arxiv 
-\usepackage{arxiv}
+%\documentclass{article} % for arxiv 
+% \usepackage{arxiv}
+%\pagestyle{plain} % arxiv only 
+\documentclass[sn-mathphys]{sn-jnl}
 
@@
 \usepackage{algorithm}
+
+\usepackage{array}
 %\usepackage{algorithm2e}
-%\usepackage{algorithmicx}
-\usepackage[noend]{algpseudocode}
-\usepackage{bbm}
+%%\usepackage{algorithmicx}
+%\usepackage[noend]{algpseudocode}
+%\usepackage{bbm}
 \usepackage{graphicx}
 
-\pagestyle{plain} % arxiv only
-
-\newcommand\sbullet[1][.5]{\mathbin{\vcenter{\hbox{\scalebox{#1}{$\bullet$}}}}}
 \usepackage{blkarray, bigstrut}
-\usepackage{tikz}
-\usetikzlibrary{arrows.meta, calc, quotes, tikzmark,shapes}
+\newcommand\topstrut[1][1.2ex]{\setlength\bigstrutjot{#1}{\bigstrut[t]}}
+\newcommand\botstrut[1][0.9ex]{\setlength\bigstrutjot{#1}{\bigstrut[b]}}
 
@@
 \captionsetup[table]{skip=4pt, position=above}
-
-\usepackage{array}
-\newcolumntype{P}[1]{>{\centering\arraybackslash}p{#1}}
-
-\newcommand\topstrut[1][1.1ex]{\setlength\bigstrutjot{#1}{\bigstrut[t]}}
-\newcommand\botstrut[1][0.95ex]{\setlength\bigstrutjot{#1}{\bigstrut[b]}}
 
@@
 \newtheorem{corollary}{Corollary}
-\newtheorem{proof}{Proof}
 % \newcommand{\CommentX}[1]{\unskip~\%~#1~\%}
@@
 	Matthew Piekenbrock\thanks{Khoury College of Computer Sciences, Northeastern University.}
-	\and Jose A. Perea\thanks{
+	\& Jose A. Perea\thanks{
 	Department of Mathematics and Khoury College of Computer Sciences, Northeastern University.}
@@
 
-\maketitle
-
 % REQUIRED - "Abstracts must be able to stand alone and so cannot contain citations to the paper's references, equations, etc. .... not exceeding 250 words"
-\begin{abstract}
+%\begin{abstract}
+\abstract{
 	The standard procedure for computing the persistent homology of a filtered simplicial complex is the matrix reduction algorithm.
@@
 	Applications to multi-dimensional persistence and crocker stacks are also presented.
-\end{abstract}
+}
+%\end{abstract}
 
@@
 %\end{AMS}
+\pacs[MSC Classification]{68T09, 55N31, 62R40}
+
+\maketitle
 
@@
 In Section \ref{sec:conclusion} we conclude the paper by discussing other possible applications and future work.
-We also include an appendix which introduces certain topics in more detail, such as the vineyards algorithm itself, for readers who may be unfamiliar but would like a more in-depth understanding.
 
@@
 The corresponding filtration is called the \emph{Vietoris-Rips} (VR) filtration, indexed by $I = \mathbb{R}$. 
-A VR filtration indexed over $\mathbb{R}$  can be condensed to an essential filtration indexed over the set of pairwise distances $\{\, d(x,x') \; | \; x,x' \in X \, \}$, 
+A VR filtration indexed over $\mathbb{R}$  can be condensed to an essential filtration indexed over the set of pairwise distances $\{\, d(x,x') \; \mid \; x,x' \in X \, \}$, 
 which can then be further reindexed to an essential simplexwise filtration by extending a total order of $X$ to its power set---e.g., 
@@
   		v \kern\rowsp & \;     &  1 & 1 \; \\
-  		w \kern\rowsp & \; 1 &  \fbox{1} &    \; \botstrut \\
+  		w \kern\rowsp & \; 1 &  \underline{1} &    \; \botstrut \\
 		\end{block}
@@
   		u \kern\rowsp  & \; 1 & 1 & 1 \; \topstrut \\
-  		v \kern\rowsp & \;     & 1 & \fbox{1} \; \\
+  		v \kern\rowsp & \;     & 1 & \underline{1} \; \\
   		w \kern\rowsp & \; 1 &    &    \; \botstrut \\
@@
 		\begin{block}{c[ccc]}
-  		u \kern\rowsp  & \; 1 & 1 &  \; \topstrut \\
-  		v \kern\rowsp & \;  & 1 &  \; \\
-  		w \kern\rowsp & \; 1 &  &  \; \botstrut \\  
-		\end{block}
-	\end{blockarray}
-	\,\,, 
+  		u \kern\rowsp  & \; 1 & 1 &  \hphantom{1} \; \topstrut \\
+  		v \kern\rowsp & \;  & 1 &  \hphantom{1} \; \\
+  		w \kern\rowsp & \; 1 &  &  \hphantom{1} \; \botstrut \\  
+		\end{block}
+	\end{blockarray}
+	\xrightarrow{S_{1}^2} 
+	\begin{blockarray}{cccc}{}
+	\kern-0.2em & a & b & c  \\
+		\begin{block}{c[ccc]}
+  		 & \; 1 &  &  \hphantom{1} \; \topstrut \\
+  		 & \;  & 1 &  \hphantom{1} \; \\
+  		 & \; 1 & 1 & \hphantom{1} \; \botstrut \\
+		\end{block}
+	\end{blockarray}
+		\xrightarrow{P}
+	\hspace{-0.2em}
+	\begin{blockarray}{cccc}{}
+	\kern-0.2em & b & a & c  \\
+		\begin{block}{c[ccc]}
+  		& \;  & 1 &  \hphantom{1} \; \topstrut \\
+  		& \; 1 &  &  \hphantom{1} \; \\
+  		& \; 1 & \underline{1} &  \hphantom{1} \; \botstrut \\
+		\end{block}
+	\end{blockarray}
+	\xrightarrow{S_1^2}
+	\hspace{-0.1em}
+	\begin{blockarray}{cccc}{}
+	\kern-0.2em & b & a & c  \\
+		\begin{block}{c[ccc]}
+  		u \kern\rowsp & \;  & 1 &  \hphantom{1} \; \topstrut \\
+  		v \kern\rowsp & \; 1 & 1 &  \hphantom{1} \; \\
+  		w \kern\rowsp & \; 1 &  &  \hphantom{1} \; \botstrut \\
+		\end{block}
+	\end{blockarray}
+\end{displaymath}
+\vspace{-3.0em}
+% Start V 
+\begin{displaymath}	
+	\def\rowsp{0.4em}
+	\def\arraystretch{0.95}
+	\makeatletter\setlength\BA@colsep{1.2pt}\makeatother
+	%
 	\begin{blockarray}{cccc}{V_1}
@@
 		\begin{block}{c[ccc]}
-        a \kern\rowsp  & \; 1 &  \fbox{1} & 1 \; \topstrut \\
+        a \kern\rowsp  & \; 1 &  \underline{1} & 1 \; \topstrut \\
   		b \kern\rowsp & \;  & 1 & 1 \; \\
@@
 	\hspace{-0.1em}
-	\begin{blockarray}{cccc}{}
-	\kern-0.2em & a & b & c  \\
-		\begin{block}{c[ccc]}
-  		 & \; 1 &  &  \; \topstrut \\
-  		 & \;  & 1 &  \; \\
-  		 & \; 1 & 1 & \; \botstrut \\
-		\end{block}
-	\end{blockarray}
-	\,\,, 
+	\, \,
 	\begin{blockarray}{cccc}{}
@@
 	\xrightarrow{P}
-	\hspace{-0.2em}
-	\begin{blockarray}{cccc}{}
-	\kern-0.2em & b & a & c  \\
-		\begin{block}{c[ccc]}
-  		& \;  & 1 &  \; \topstrut \\
-  		& \; 1 &  &  \; \\
-  		& \; 1 & \fbox{1} &  \; \botstrut \\
-		\end{block}
-	\end{blockarray}
-	\,\,, 
 	\begin{blockarray}{cccc}{}
@@
 	\xrightarrow{S_1^2}
-	\hspace{-0.1em}
 	\begin{blockarray}{cccc}{}
@@
 		\begin{block}{c[ccc]}
-  		u \kern\rowsp  & \;  & 1 &  \; \topstrut \\
-  		v \kern\rowsp & \; 1 & 1 &  \; \\
-  		w \kern\rowsp & \; 1 &  &  \; \botstrut \\
-		\end{block}
-	\end{blockarray}
-	\,\,,
-	\begin{blockarray}{cccc}{}
-	\kern-0.2em & b & a & c  \\
-		\begin{block}{c[ccc]}
-  		  & \; 1 & 1 & 1 \; \topstrut \\
-  		 & \; & 1 & 1 \; \\
-  		 & \;  &  & 1 \; \botstrut \\
+  		 b \kern\rowsp & \; 1 & 1 & 1 \; \topstrut \\
+  		 a \kern\rowsp & \; & 1 & 1 \; \\
+  		 c \kern\rowsp & \;  &  & 1 \; \botstrut \\
 		\end{block}
@@
 \end{displaymath}
+%\begin{displaymath}
+%	\def\rowsp{0.4em}
+%	\def\arraystretch{0.95}
+%	\makeatletter\setlength\BA@colsep{1.2pt}\makeatother
+%	%
+%	\begin{blockarray}{cccc}{R_1}
+%	 \kern-0.2em & a & b & c  \\
+%		\begin{block}{c[ccc]}
+%  		u \kern\rowsp  & \; 1 & 1 &  \; \topstrut \\
+%  		v \kern\rowsp & \;  & 1 &  \; \\
+%  		w \kern\rowsp & \; 1 &  &  \; \botstrut \\  
+%		\end{block}
+%	\end{blockarray}
+%	\,\,, 
+%	\begin{blockarray}{cccc}{V_1}
+%	\kern-0.2em & a & b & c  \\
+%		\begin{block}{c[ccc]}
+%        a \kern\rowsp  & \; 1 &  \fbox{1} & 1 \; \topstrut \\
+%  		b \kern\rowsp & \;  & 1 & 1 \; \\
+%  		c \kern\rowsp & \;  &  & 1 \; \botstrut \\
+%		\end{block}
+%	\end{blockarray}
+%	\xrightarrow{S_{1}^2} 
+%	\hspace{-0.1em}
+%	\begin{blockarray}{cccc}{}
+%	\kern-0.2em & a & b & c  \\
+%		\begin{block}{c[ccc]}
+%  		 & \; 1 &  &  \; \topstrut \\
+%  		 & \;  & 1 &  \; \\
+%  		 & \; 1 & 1 & \; \botstrut \\
+%		\end{block}
+%	\end{blockarray}
+%	\,\,, 
+%	\begin{blockarray}{cccc}{}
+%	\kern-0.2em & a & b & c  \\
+%		\begin{block}{c[ccc]}
+%  		 & \; 1 &  & 1 \; \topstrut \\
+%  		 & \;  & 1 & 1 \; \\
+%  		 & \;  &  & 1 \; \botstrut \\
+%		\end{block}
+%	\end{blockarray}
+%	\xrightarrow{P}
+%	\hspace{-0.2em}
+%	\begin{blockarray}{cccc}{}
+%	\kern-0.2em & b & a & c  \\
+%		\begin{block}{c[ccc]}
+%  		& \;  & 1 &  \; \topstrut \\
+%  		& \; 1 &  &  \; \\
+%  		& \; 1 & \fbox{1} &  \; \botstrut \\
+%		\end{block}
+%	\end{blockarray}
+%	\,\,, 
+%	\begin{blockarray}{cccc}{}
+%	\kern-0.2em & b & a & c  \\
+%		\begin{block}{c[ccc]}
+%  		& \; 1 &  & 1 \; \topstrut \\
+%  		& \;  & 1 & 1 \; \\
+%  		& \;  &  & 1 \; \botstrut \\
+%		\end{block}
+%	\end{blockarray}
+%	\xrightarrow{S_1^2}
+%	\hspace{-0.1em}
+%	\begin{blockarray}{cccc}{}
+%	\kern-0.2em & b & a & c  \\
+%		\begin{block}{c[ccc]}
+%  		u \kern\rowsp  & \;  & 1 &  \; \topstrut \\
+%  		v \kern\rowsp & \; 1 & 1 &  \; \\
+%  		w \kern\rowsp & \; 1 &  &  \; \botstrut \\
+%		\end{block}
+%	\end{blockarray}
+%	\,\,,
+%	\begin{blockarray}{cccc}{}
+%	\kern-0.2em & b & a & c  \\
+%		\begin{block}{c[ccc]}
+%  		  & \; 1 & 1 & 1 \; \topstrut \\
+%  		 & \; & 1 & 1 \; \\
+%  		 & \;  &  & 1 \; \botstrut \\
+%		\end{block}
+%	\end{blockarray}
+%\end{displaymath}
 %Prior to performing the exchange, observe that that the bold entry in $V_1$ would render $V_1$ non-upper triangular after the exchange.
@@
 This hidden cost may seem minor, however, the quadratic scaling of the number of transpositions vineyards typically requires makes this cost non-trivial.
-In fact, achieving $O(m)$ complexity per transposition requires a special sparse matrix representation that allows swapping any two rows and columns in $O(1)$---see the appendix for more details and discussion.
-Finally, note the constant factor associated with the stated complexity: inspection of the individual cases of the algorihtm from~\cite{cohen2006vines} shows that there are at most two $O(m)$ operations, on both $R$ and $V$, that may be needed for any single transposition.
+In fact, achieving $O(m)$ complexity per transposition requires a special sparse matrix representation that allows swapping any two rows and columns in $O(1)$, as described in~\cite{cohen2006vines}.
+Finally, note the constant factor associated with the stated complexity: inspection of the individual cases of the algorithm from~\cite{cohen2006vines} shows that there are at most two $O(m)$ operations, on both $R$ and $V$, that may be needed for any single transposition.
 
@@
  \begin{enumerate}[labelsep=5pt, topsep=0pt,itemsep=-0.25ex,parsep=1.2ex]
- 	\item[M1:] Querying the non-zero status of entries in either $R$ or $V$ occurs once per move.
+ 	\item[M1:] Querying the non-zero status of entries in $R$ or $V$ occurs once per move.
  	\item[M2:] $R = D V$ is \underline{not} guaranteed to be valid  during the movement of $\sigma_i $ to $ \sigma_j$.
@@
 	\def\rowsp{0.4em}
-	\def\arraystretch{0.95}
-	\makeatletter\setlength\BA@colsep{1.2pt}\makeatother
+	\def\arraystretch{0.60}
+	\makeatletter\setlength\BA@colsep{0.9pt}\makeatother
 	%
@@
 	\; \;
-	\begin{blockarray}{cccc}{R_1}
+	\begin{blockarray}{cccc}{R}
 	 \kern-0.2em & a & b & c  \\
 		\begin{block}{c[ccc]}
-  		u \kern \rowsp & \; 1 & 1 &  \; \topstrut \\
-  		v \kern \rowsp & \;  & 1 &  \; \\
-  		w \kern \rowsp & \; 1 &  &  \; \botstrut \\
+  		u \kern \rowsp & \; 1 & 1 &  \hphantom{1} \; \topstrut \\
+  		v \kern \rowsp & \;  & 1 &  \hphantom{1} \; \\
+  		w \kern \rowsp & \; 1 &  &  \hphantom{1} \; \botstrut \\
 		\end{block}
@@
 		\begin{block}{c[ccc]}
-  		  & \; 1 &    &  \; \topstrut \\
-  		  & \;    & 1 &  \; \\
-  		 & \; 1 & \fbox{1} &  \; \botstrut \\
+  		  & \; 1 &    &  \hphantom{1} \; \topstrut \\
+  		  & \;    & 1 &  \hphantom{1} \; \\
+  		 & \; 1 & \underline{1} &  \hphantom{1} \; \botstrut \\
 		\end{block}
@@
   		 & \;       &  1  &  1  \; \\
-  		 & \; 1   &   \fbox{1}   &     \; \botstrut \\
+  		 & \; 1   &   \underline{1}   &     \; \botstrut \\
 		\end{block}
@@
   		 & \; 1 & 1 &  \; \\
-  		 & \; 1 &  & \fbox{1} \; \botstrut \\
-		\end{block}
-	\end{blockarray}
-%
-\; \xrightarrow{d_R} 
-\begin{blockarray}{cccc}{R_1}
+  		 & \; 1 &  & \underline{1} \; \botstrut \\
+		\end{block}
+	\end{blockarray}
+%
+\xrightarrow{d_R} 
+\begin{blockarray}{cccc}
 \kern-0.2em & b & c & a  \\
 	\begin{block}{c[ccc]}
-		u \kern \rowsp & \;  & 1 &  \; \topstrut \\
-		v \kern \rowsp & \; 1 & 1 &  \; \\
-		w \kern \rowsp & \; 1 &  &  \; \botstrut \\
+		& \;  & 1 &  \hphantom{1} \; \topstrut \\
+		& \; 1 & 1 &  \hphantom{1} \; \\
+		& \; 1 &  &  \hphantom{1} \; \botstrut \\
 	\end{block}
@@
 
-\vspace{-1.0em}
-
-% Start V example 
+\vspace{-3.0em}
+
+% start V example (1) 
 \begin{displaymath}
 	\def\rowsp{0.4em}
-	\def\arraystretch{0.95}
-	\makeatletter\setlength\BA@colsep{1.2pt}\makeatother
+	\def\arraystretch{0.60}
+	\makeatletter\setlength\BA@colsep{0.9pt}\makeatother
 	%
@@
 		\begin{block}{c[c]}
-  		 & \; 1 \; \topstrut \\
-  		& \;  \; \\
-  		& \;  \; \botstrut \\
+  		a \kern \rowsp & \; 1 \; \topstrut \\
+  		b \kern \rowsp & \;  \; \\
+  		c \kern \rowsp & \;  \; \botstrut \\
 		\end{block}
@@
 	\; \;
-	\begin{blockarray}{cccc}{V_1}
+	\begin{blockarray}{cccc}{V}
 	 \kern-0.2em & a & b & c  \\
 		\begin{block}{c[ccc]}
-  		 & \; 1 & \fbox{1} &  \fbox{1}\; \topstrut \\
-  		& \;  & 1 &  1\; \\
-  		& \; &  & 1 \; \botstrut \\
+  		a \kern \rowsp & \; 1 & \underline{1} & \underline{1} \; \topstrut \\
+  		b \kern \rowsp & \;  & 1 & 1 \; \\
+  		c \kern \rowsp & \;  &  & 1 \; \botstrut \\
 		\end{block}
@@
 		\begin{block}{c[ccc]}
-  		 & \; 1 & &  \fbox{1}\; \topstrut \\
-  		& \;  & 1 &  1\; \\
+  		  & \; 1 &    & \underline{1} \; \topstrut \\
+  		  & \;    & 1 &  1 \; \\
   		 & \;  & &  1 \; \botstrut \\
@@
 %
-\; \xrightarrow{P} \\
-% 
+\; \xrightarrow{P} 
 \begin{blockarray}{cc}{}
@@
 		\begin{block}{c[ccc]}
-  		 & \; 1 & &\; \topstrut \\
+  		  & \; 1 &  &  \; \topstrut \\
   		 & \;  & 1 &  \; \\
@@
 %
-\; \xrightarrow{d_V}
-% 
-	\begin{blockarray}{cccc}{V_1}
+\xrightarrow{d_V} 
+\begin{blockarray}{cccc}
 \kern-0.2em & b & c & a  \\
 	\begin{block}{c[ccc]}
-  		b \kern \rowsp & \; 1 & & 1 \; \topstrut \\
-  		c \kern \rowsp & \; & 1 &  1 \; \\
-  		a \kern \rowsp & \; &  & 1  \; \botstrut \\
-		\end{block}
+		& \;  1 & &  1 \; \topstrut \\
+		& \;  & 1 &  1\; \\
+		& \; &  & 1 \; \botstrut \\
+	\end{block}
+	% \undermat{B}{ & & }  
 \end{blockarray}
@@
 \begin{align}\label{eq:sf_schedule}
-	S_F &= \{ (i,j) \, | \, f_0(\sigma_i) < f_0(\sigma_j) \mbox{ and }  f_1(\sigma_i) > f_1(\sigma_j) \quad  \; i,j \in [m] \} \\
+	S_F &= \{ (i,j) \, \mid \, f_0(\sigma_i) < f_0(\sigma_j) \mbox{ and }  f_1(\sigma_i) > f_1(\sigma_j) \quad  \; i,j \in [m] \} \\
 	&= \{ s_1, s_2, \dots, s_k \}
@@
 \[S_F = 
-\mathrm{Inv}(p, q) = \{ \, (i,j) \, | \, p(i) < p(j) \mbox{ and } q(i) > q(j)  \quad \; i,j \in [m]\, \}
+\mathrm{Inv}(p, q) = \{ \, (i,j) \, \mid \, p(i) < p(j) \mbox{ and } q(i) > q(j)  \quad \; i,j \in [m]\, \}
 \]
@@
 \[
-\setlength\arraycolsep{2.3pt}
-\scalefont{1.0}{
+\setlength\arraycolsep{1.8pt}
+\scalefont{0.90}{
 \setcounter{MaxMatrixCols}{20}
@@
 \[
-\scalefont{1.0}{
+\setlength\arraycolsep{1.8pt}
+\scalefont{0.90}{
 \setcounter{MaxMatrixCols}{20}
@@
 \subsubsection{Motivation}
-The results from the previous section yield a sufficient algorithm for generating move schedules of minimal cardinality: simply compute a $\mathrm{LCS}(K_0, K_1)$ for a pair $(K_0, K_1)$ and fix an order of simplices in the set $\mathrm{LCS}(K_0, K_1) \smallsetminus K_0$ to move.
-Any sequence of moves which increases the LCS on each move is guaranteed to have size $m - \lvert \mathrm{LCS}(K_0, K_1) \rvert$, and the reduction to the permutation edit distance problem ensures this size is optimal.
-However, like in the vineyards algorithm, certain pairs of simplices cost more to exchange depending on whether they are critical pairs in the sense described in~\cite{cohen2006vines}. 
-This small variation in cost-per-transposition induces a large variability in the cost of randomly generated schedules, which is computationally undesirable in practice.
-Ideally, we would like to generate a schedule which not only is minimal in size, but is also minimal in terms of the column operations it requires.
+The results from the previous section yield a sufficient algorithm for generating move schedules of minimal cardinality: any schedule of moves $S$ sorting $K_0 \mapsto K_1$ above is guaranteed to have size $\lvert S \rvert = m - \lvert \mathrm{LCS}(K_0, K_1) \rvert$, and the reduction to the permutation edit distance problem ensures this size is optimal.
+However, as with the vineyards algorithm, certain pairs of simplices cost more to exchange depending on whether they are critical pairs in the sense described in~\cite{cohen2006vines}, resulting in a large variability in the cost of randomly generated schedules.
+This variability is undesirable in practice: we would like to generate a schedule which not only small in size, but is also efficient in terms of its required column operations.
+
 %Formally, given two permutations $p, q \in S_m$, let $S = ( s_1, s_2, \dots, s_d )$ denote a sorting such that $S \circ p = q$. 
@@
 The subset of the filtration which corresponds to the simplices which lie in the LCS between these permutations is colored in red. For this example, the permutation edit distance is 
-$d = m - |\mathrm{LCS}(K_0, K_1)|$ implies that exactly $3$ moves are needed to apply the mapping $K_0 \mapsto K_1$. There are six possible valid schedules of moves: 
+$d = m - \lvert \mathrm{LCS}(K_0, K_1) \rvert$ implies that exactly $3$ moves are needed to apply the mapping $K_0 \mapsto K_1$. There are six possible valid schedules of moves: 
 \begin{alignat*}{3}
@@
 \centering
-\begin{tabular}{ |P{0.4cm}|P{0.8cm}|P{0.8cm}|P{0.8cm}|P{0.8cm}|  }
+\begin{tabular}{ m{0.4cm} m{0.8cm} m{0.8cm} m{0.8cm} m{0.8cm}  }
  \hline
- \multicolumn{5}{|c|}{Cost of each permutation} \\
+ \multicolumn{5}{c}{Cost of each permutation} \\
  \hline
@@
 \centering
-\begin{tabular}{ |P{1.6cm}|P{1.6cm}|P{1.2cm}|P{2.8cm}|P{2.4cm}|}
+\begin{tabular}{ m{1.6cm} m{1.6cm} m{1.2cm} m{2.8cm} m{2.4cm} }
  \hline
@@
  %\hline
-  \multicolumn{1}{|c|}{ $\beta(M)$ } & \multicolumn{1}{|c|}{$\mathcal{A}(M)$ } & \multicolumn{3}{c|}{ Column reductions / Permutations } \\
+  \multicolumn{1}{|c|}{ $\beta(M)$ } & \multicolumn{1}{|c|}{$\mathcal{A}(M)$ } & \multicolumn{3}{c|}{ Col. Reductions / Permutations } \\
  \hline
@@
 %\bibliographystyle{siamplain}
-\bibliographystyle{unsrt}
+%\bibliographystyle{unsrt}
 \bibliography{move_schedules}
 
-\appendix 
-\section{}\label{sec:appendix}
-\subsection*{Vineyard Algorithm}
-\begin{algorithm}[!htb]
-	\caption{Vineyards: Tranposition Framework}\label{alg:tr}
-	\setstretch{1.15}
-	\begin{algorithmic}[1]
-		\Require Matrices $R, V$ satisfying $R = D V$, $1 \leq i \leq m - 1$
-		\Ensure Output $(R, V)$ maintains the decomposition invariants
-        \Function{Transpose}{$R$, $V$, $i$}
-            \State $\mathrm{pos} \gets $ columns satisfying $\mathrm{col}_R = 0$
-            \If {$\mathrm{pos}[i]$ \textbf{and} $\mathrm{pos}[i+1]$}
-	            \If{$V[i,i+1] \neq 0$}
-            		\State $\mathrm{col}_V(i+1)  \mathrel{+}= \mathrm{col}_V(i)$	
-            	\EndIf
-            	\If {$\exists \; k, l$ s.t. $\mathrm{low}_R(k) = i, \mathrm{low}_R(l) = i+1$ and $R[i,l] \neq 0$} \Comment{O(m)}
-            		\If {$k < l$}
-            			\State \Return $(\, R, \, V \,) \gets (\, PRPS_k^l, \, PVP S_k^l \,)$
-            		\Else
-            			\State \Return $(\, R, \, V \,) \gets  (\, PRPS_l^k, \, PVP S_l^k\,) $
-            		\EndIf
-            	\EndIf 
-            \ElsIf{$!\mathrm{pos}[i]$ \textbf{and} $!\mathrm{pos}[i+1]$}
-            	\If{$V[i,i+1] \neq 0$}\Comment{O(m)}
-            		\If{$\mathrm{low}_R(i) < \mathrm{low}_R(i+1)$}
-            			\State \Return  $(\, R, V \,) \gets (P R S^{i+1}_i P, \, P V S^{i+1}_i  P) $
-            		\Else 
-            			\State \Return  $(\, R, V \,) \gets (P R S^{i+1}_i P S^{i+1}_i, \, P V S^{i+1}_i P S^{i+1}_i ) $
-            		\EndIf 
-            	\EndIf
-          	\ElsIf{$!\mathrm{pos}[i]$ \textbf{and} $\mathrm{pos}[i+1]$} 
-            	\If{$V[i,i+1] \neq 0$}\Comment{O(m)}
-            		\State \Return  $(\, R, V \,) \gets (P R S^{i+1}_i P S^{i+1}_i, \, P V S^{i+1}_i P S^{i+1}_i ) $
-            	\EndIf
-            \ElsIf{$\mathrm{pos}[i]$ \textbf{and} $!\mathrm{pos}[i+1]$}
-            	\If{$V[i,i+1] \neq 0$}\Comment{O(m)}
-            		\State $\mathrm{col}_V(i+1) \mathrel{+}= \mathrm{col}_V(i)$
-            	\EndIf 
-            \EndIf
-            \State \Return $(\, R , V \,) \gets (\, P R P , P V P \,)$
-        \EndFunction
-    \end{algorithmic}
-\end{algorithm}
-
-\subsection*{Transposition complexity details}
-Consider Algorithm~\ref{alg:tr}.
+%\appendix 
+%\section{}\label{sec:appendix}
+%\subsection*{Vineyard Algorithm}
+%\begin{algorithm}[!htb]
+%	\caption{Vineyards: Tranposition Framework}\label{alg:tr}
+%	\setstretch{1.15}
+%	\begin{algorithmic}[1]
+%		\Require Matrices $R, V$ satisfying $R = D V$, $1 \leq i \leq m - 1$
+%		\Ensure Output $(R, V)$ maintains the decomposition invariants
+%        \Function{Transpose}{$R$, $V$, $i$}
+%            \State $\mathrm{pos} \gets $ columns satisfying $\mathrm{col}_R = 0$
+%            \If {$\mathrm{pos}[i]$ \textbf{and} $\mathrm{pos}[i+1]$}
+%	            \If{$V[i,i+1] \neq 0$}
+%            		\State $\mathrm{col}_V(i+1)  \mathrel{+}= \mathrm{col}_V(i)$	
+%            	\EndIf
+%            	\If {$\exists \; k, l$ s.t. $\mathrm{low}_R(k) = i, \mathrm{low}_R(l) = i+1$ and $R[i,l] \neq 0$} \Comment{O(m)}
+%            		\If {$k < l$}
+%            			\State \Return $(\, R, \, V \,) \gets (\, PRPS_k^l, \, PVP S_k^l \,)$
+%            		\Else
+%            			\State \Return $(\, R, \, V \,) \gets  (\, PRPS_l^k, \, PVP S_l^k\,) $
+%            		\EndIf
+%            	\EndIf 
+%            \ElsIf{$!\mathrm{pos}[i]$ \textbf{and} $!\mathrm{pos}[i+1]$}
+%            	\If{$V[i,i+1] \neq 0$}\Comment{O(m)}
+%            		\If{$\mathrm{low}_R(i) < \mathrm{low}_R(i+1)$}
+%            			\State \Return  $(\, R, V \,) \gets (P R S^{i+1}_i P, \, P V S^{i+1}_i  P) $
+%            		\Else 
+%            			\State \Return  $(\, R, V \,) \gets (P R S^{i+1}_i P S^{i+1}_i, \, P V S^{i+1}_i P S^{i+1}_i ) $
+%            		\EndIf 
+%            	\EndIf
+%          	\ElsIf{$!\mathrm{pos}[i]$ \textbf{and} $\mathrm{pos}[i+1]$} 
+%            	\If{$V[i,i+1] \neq 0$}\Comment{O(m)}
+%            		\State \Return  $(\, R, V \,) \gets (P R S^{i+1}_i P S^{i+1}_i, \, P V S^{i+1}_i P S^{i+1}_i ) $
+%            	\EndIf
+%            \ElsIf{$\mathrm{pos}[i]$ \textbf{and} $!\mathrm{pos}[i+1]$}
+%            	\If{$V[i,i+1] \neq 0$}\Comment{O(m)}
+%            		\State $\mathrm{col}_V(i+1) \mathrel{+}= \mathrm{col}_V(i)$
+%            	\EndIf 
+%            \EndIf
+%            \State \Return $(\, R , V \,) \gets (\, P R P , P V P \,)$
+%        \EndFunction
+%    \end{algorithmic}
+%\end{algorithm}
+%
+%\subsection*{Transposition complexity details}
+%Consider Algorithm~\ref{alg:tr}.
 The matrices $R, V$ are both $(m \times m)$, where $m$ is the number of simplices in the underlying filtration.
@@
 This data structure requires storing an index using at most $O(m \log m)$ bits with each non-zero entry, as well as several $O(m)$-sized auxiliary arrays to protect the matrices from row exchanges.
-
-% The principle motivation  
-% Upon moving $s$
-
-% Denote this class with $c_s$.
-% Now if the position of $s$ in $P_1$ is $i$ and the position in $P_2$ is $j$, and $i < j$, then the
-
-% % Note: patience sorting gives means of enumerating all LIS's
-% % Note #2: Consider having the identity permutation be the target and going in reverse--at the end, there is only one symbol to move into place
-% % prims-MST like approach possible?
 

### 2023-01-23 move_schedules.tex
Revisions from reviewer, part 1

@@
 %\pagestyle{plain} % arxiv only 
+\PassOptionsToPackage{noend}{algpseudocode}
 \documentclass[sn-mathphys]{sn-jnl}
 
-%\usepackage[margin=0.75in]{geometry}
 \usepackage{amsmath, amssymb}
@@
 \usepackage[english]{babel}
-\usepackage{algorithm}
+% \usepackage{algorithm}
 
@@
 %%\usepackage{algorithmicx}
-%\usepackage[noend]{algpseudocode}
+% \usepackage[noend]{algpseudocode} % keep the noend argument
 %\usepackage{bbm}
-\usepackage{graphicx}
+% \usepackage{graphicx}
 
 \usepackage{blkarray, bigstrut}
-\newcommand\topstrut[1][1.2ex]{\setlength\bigstrutjot{#1}{\bigstrut[t]}}
+\newcommand\topstrut[1][1.0ex]{\setlength\bigstrutjot{#1}{\bigstrut[t]}}
 \newcommand\botstrut[1][0.9ex]{\setlength\bigstrutjot{#1}{\bigstrut[b]}}
@@
 \newtheorem{corollary}{Corollary}
-% \newcommand{\CommentX}[1]{\unskip~\%~#1~\%}
 
@@
 % is automatically inserted before the title.
-\title{
-	Move Schedules: Fast persistence computations in coarse dynamic settings
+\title[Fast Persistence Computations in Dynamic Settings]{
+	\vspace{-2.0em} Move Schedules: Fast persistence computations in coarse dynamic settings
 	\thanks{
@@
 % Authors: full names plus addresses.
-\author{
-	Matthew Piekenbrock\thanks{Khoury College of Computer Sciences, Northeastern University.}
-	\& Jose A. Perea\thanks{
-	Department of Mathematics and Khoury College of Computer Sciences, Northeastern University.}
+%\author{
+%	Matthew Piekenbrock\thanks{Khoury College of Computer Sciences, Northeastern University.} and Jose A. Perea\thanks{
+%	Department of Mathematics and Khoury College of Computer Sciences, Northeastern University.}
+%}
+
+% Spring format
+\author[1]{\fnm{Matthew} \sur{Piekenbrock}}\email{piekenbrock.m@northeastern.edu}
+\affil[1]{
+	\orgdiv{Khoury College of Computer Sciences}, 
+	\orgname{Northeastern University}}
+	%\orgaddress{\street{360 Huntington Ave}, \city{Boston}, \postcode{02115}, \state{MA}%\country{United States}}
+% }
+% , ,
+
+\author[2]{\fnm{Jose} \sur{A. Perea}}\email{j.pereabenitez@northeastern.edu}
+\affil[2]{
+	\orgdiv{Department of Mathematics and Khoury College of Computer Sciences},
+	\orgname{Northeastern University}
+	% \orgaddress{\street{360 Huntington Ave}, \city{Boston}, \postcode{02115}, \state{MA}}
 }
@@
 
+\raggedbottom
+
 % ------  Begin document ------- 
@@
 \abstract{
-	The standard procedure for computing the persistent homology of a filtered simplicial complex is the matrix reduction algorithm.
+	The standard procedure for computing the persistent homology of a filtered simplicial complex with $m$ simplices is the matrix reduction algorithm.
 Its output is a particular decomposition of the total boundary matrix, from which the persistence diagrams and generating cycles can be derived.
-	Persistence diagrams are known to vary continuously with respect to their input, which motivates the algorithmic study of persistence computations for time-varying filtered complexes.
+	Persistence diagrams are known to vary continuously with respect to their input, motivating the  study of their computation for time-varying filtered complexes.
 Computationally, simulating persistence dynamically can be reduced to maintaining a valid decomposition under adjacent transpositions in the filtration order.
-	In practice, the quadratic scaling in the number of  such transpositions often makes this maintenance procedure slower than simply computing the decomposition from scratch,  limiting the application of dynamic persistence to relatively small data sets.
-In this work, we propose a coarser strategy for maintaining the decomposition over a 1-parameter family of filtrations.
-Our first result is an analysis of a simple linear-time strategy which reduces the number of column operations needed to simulate persistence across a fixed homotopy by at most a factor of 2.
-We show a modification of this technique which maintains only a sublinear number of valid states, as opposed to a quadratic number of states, and we provide tight lower bounds for this technique.
-	Finally, we present results showing that the decrease in operations to compute diagrams across a family of filtrations is proportional to the difference between the expected quadratic number of states, and the proposed sublinear coarsening.
+	Since there are $O(m^2)$ such transpositions, this maintenance procedure exhibits limited scalability and often is too fine for many application contexts.
+	As our first result, we propose a coarser strategy for maintaining the decomposition over a 1-parameter family of filtrations that take $O(m^2)$ preprocessing time uses only $O(m)$ space. 
+	We also show that this maintenance procedure yields a factor-2 net reduction in column operations over the state of the art approach. 
+	By reduction to a particular longest common subsequence problem, we give a modification of this strategy that not only can be performed in $O(m \log \log m)$ time, but yields a storage bound that is actually sublinear in expectation.
+	Finally, we show experimentally the decrease in operations to compute diagrams across a family of filtrations is proportional to the difference between the expected quadratic number of states and the proposed sublinear coarsening.
 	Applications to multi-dimensional persistence and crocker stacks are also presented.
@@
 % REQUIRED
-\begin{keywords}
-  {Computational topology, Persistent homology, Topological data analysis}
-\end{keywords}
+%\begin{keywords}
+%  {Computational topology, Persistent homology, Topological data analysis}
+%\end{keywords}
+\keywords{Computational topology, Persistent homology, Topological data analysis}
 
@@
 With the current state of the art tools, there is no recourse to the problem of computing persistence from a time-varying filtration containing millions of simplices at thousands of snapshots in time.
-Nonetheless, having the capability of efficiently updating the persistence of a time-varying filtration has far-reaching consequences: methods that vectorize persistence diagrams for machine learning purposes, like adaptive template functions~\cite{polanco2019adaptive} and persistence images~\cite{adams2017persistence}, or summary statistics like $\alpha$-smoothed Betti curves~\cite{ulmer2019topological}, all immediately become computationally viable tools in dynamic settings.
+Acquiring such a capability has far-reaching consequences: methods that vectorize persistence diagrams for machine learning purposes, like adaptive template functions~\cite{polanco2019adaptive} and persistence images~\cite{adams2017persistence}, or summary statistics like $\alpha$-smoothed Betti curves~\cite{ulmer2019topological}, all immediately become computationally viable tools in dynamic settings.
 %This analogy provides a strong interpretation at gaining insight on the subtle topological changes that may occur in studying a continuous process.
 
-Cohen-Steiner et al.
-refer to a continuous 1-parameter family of persistence diagrams as a \emph{vineyard}, and they give in \cite{cohen2006vines} an efficient algorithm for their computation given a time-varying filtration.
+Cohen-Steiner, Morozov, and Edelsbrunner refer to a continuous 1-parameter family of persistence diagrams as a \emph{vineyard}, and they give in \cite{cohen2006vines} an efficient algorithm for their computation.
 The vineyards approach can be interpreted as an extension of the \emph{matrix reduction} algorithm~\cite{zomorodian2005computing}, which computes the persistence diagram of a fixed filtration $K$ with $m$ simplices in $O(m^3)$ time.
@@
 If one is interested in understanding how the persistent homology of a continuous function changes over time, then this algorithm is sufficient, for  homological critical points can only occur when the filtration order changes.
-This algorithm is also efficient  asymptotically: if there are $d$ time-points where the filtration order changes, then   vineyards   takes $O(m^3 + md)$ time; one $O(m^3)$-time reduction upfront for the filtration at time $t_0$ followed by one $O(m)$ operation to update the decomposition for the remaining time points $(t_1, t_2, \dots, t_d)$.
+The vineyards algorithm is also efficient asymptotically: if there are $d$ time-points where the filtration order changes, then   vineyards   takes $O(m^3 + md)$ time; one $O(m^3)$-time reduction upfront for the filtration at time $t_0$ followed by one $O(m)$ operation to update the decomposition for the remaining time points $(t_1, t_2, \dots, t_d)$.
 When $d >> m$, this $O(md)$ approach is far more efficient than the $O(dm^3)$  ``naive'' strategy of computing the diagrams independently at every time point.
@@
 Ripser, PHAT, Dionysus, etc.
-See~\cite{otter2017roadmap} for an overview), implementations of the vineyards algorithm are relatively uncommon.\footnote{Dionysus 1 does have an implementation of vineyards, however the algorithm was never ported to version 2.
+see~\cite{otter2017roadmap} for an overview), implementations of the vineyards algorithm are relatively uncommon.\footnote{Dionysus 1 does have an implementation of vineyards, however the algorithm was never ported to version 2.
 Other major packages, such as GUDHI and PHAT, do not have vineyards implementations.}
@@
 many transpositions can be quite slow...
-it is sometimes much faster to simply recompute the $RU$-decomposition from scratch using the standard persistence algorithm.'' Indeed, whether performing the reductions naively or updating the decomposition dynamically using vineyards, they observed maintaining the decomposition is the most computationally demanding aspect of RIVET \cite{rivet}.
+it is sometimes much faster to simply recompute the $RU$-decomposition from scratch using the standard persistence algorithm.'' Indeed, they observed maintaining the decomposition along a certain parameterized family is the most computationally demanding aspect of RIVET~\cite{rivet}, software for computing and visualizing	 two-parameter persistent homology.
 
@@
 % in many time varying settings, a valid decomposition is often only sought after at a relatively sparse number of time points relative to the size of the (finite) set needed to generated by the family.
-By exploiting the ``donor'' concept introduced in in~\cite{busaryev2010tracking}, we are able to make a tradeoff between the number of times the decomposition is restored to a valid state and the granularity of the decomposition repair step, dramatically reducing the total number of column operations needed to apply an arbitrary permutation to the filtration.
+By exploiting the ``donor'' concept introduced in~\cite{busaryev2010tracking}, we are able to make a tradeoff between the number of times the decomposition is restored to a valid state and the granularity of the decomposition repair step, reducing the total number of column operations needed to apply an arbitrary permutation to the filtration.
 This trade off, paired with a fast greedy heuristic explained in section~\ref{sec:proxy_objective}, yields an algorithm that can update a $R = DV$ decomposition more efficiently than vineyards in coarse time-varying contexts, making dynamic persistence more computationally tractable for a wider class of use-cases.
-The source code containing both the algorithm we propose and the experiments performed in Section~\ref{sec:results} is available online as an open source \texttt{R} package.\footnote{ Code is available here: \url{https://github.com/peekxc/dart}}
-
- \subsection{A Motivating Example and Contributions}\label{sec:motivation} 
-As a motivating example, we describe a simple experiment illustrating why the vineyards algorithm does not always yield an efficient strategy for time-varying applications. 
-Consider a series of grayscale images (i.e., a video) depicting a fixed-width annulus expanding about the center of a $9 \times 9$ grid. 
+The source code containing both the algorithm we propose and the experiments performed in Section~\ref{sec:results} is open source and available online.\footnote{ Code is available here: \url{https://github.com/peekxc/dart}}
+
+\subsection{Related Work}\label{sec:related_work} 
+To the authors knowledge, work focused on ways of updating a  decomposition $R = DV$, for all homological dimensions, is limited---there is the vineyards algorithm~\cite{cohen2006vines} and the moves algorithm~\cite{busaryev2010tracking}, both of which are discussed extensively in section~\ref{sec:background}.
+
+In contrast, there is extensive work  on improving the efficiency of computing a (static) $R = DV$ decomposition.
+Chen~\cite{chen2011persistent} proposed \emph{persistence with a twist}, also called the \emph{clearing optimization}, which exploits a boundary/cycle relationship to ``kill'' columns early in the reduction rather than reducing them.
+Another popular optimization is to utilize the duality between   homology and cohomology~\cite{de2011dualities}, which dramatically improves the effectiveness of the clearing optimization~\cite{bauer2021ripser}.
+There are many other optimizations on the implementation side: the use of ranking functions defined on the combinatorial number system enables implicit cofacet enumeration, removing the need to store the boundary matrix explicitly; the apparent/emergent pairs optimization identifies columns whose pivot entries are unaffected by the reduction algorithm, reducing the total number of columns which need be reduced; sparse data structures such as bit-trees and lazy heaps allow for efficient column-wise additions with $\mathbb{Z}_2 = \mathbb{Z}/2\mathbb{Z}$ coefficients and effective $O(1)$ pivot entry retrieval, and so on~\cite{bauer2021ripser, bauer2017phat}.
+
+By making stronger assumptions on the underlying topological space, restricting the homological dimension, or targeting  a weaker invariant (e.g.
+Betti numbers), one can usually obtain  faster or alternative approaches  recording information related to persistence.
+For example, Attali et al.~\cite{attali2009persistence} give a linear time algorithm (under the $\texttt{word RAM}$ model of computation) for computing persistence on graphs.
+In the same paper, they describe how to obtain $\epsilon$-simplifications of $1$-dimensional persistence diagrams for filtered $2$-manifolds by using duality and symmetry theorems. 
+Along a similar vein, Edelsbrunner et al.~\cite{edelsbrunner2000topological} give a fast incremental algorithm for computing persistent Betti numbers up to dimension $2$, again by utilizing symmetry, duality, and ``time-reversal''~\cite{delfinado1995incremental}.
+Chen et al.~\cite{chen2013output} give an output-sensitive method for computing persistent homology,  utilizing the property that certain submatrices of  $D$ have the same rank as  $R$, which they exploit through fast sub-cubic rank algorithms specialized for sparse-matrices.
+
+If  zeroth homology is the only dimension of interest, computing and updating both the persistence and rank information  is greatly simplified.
+For example, if the relations are available a-priori, obtaining a tree representation fully characterizing the connectivity of the underlying space (also known as the \emph{incremental connectivity} problem) takes just $O(\alpha(n) n)$ time using the disjoint-set data structure, where $\alpha(n)$ is the extremely slow-growing inverse Ackermann function.
+Adapting this approach to the time-varying setting, Oesterling et al.~\cite{oesterling2015computing} give a $O(e)$-per-update algorithm that allows one to maintain a \emph{merge tree} with $e$ edges,
+associated to the filtration parameter changing continuously over time. 
+If only the zeroth-dimensional Betti numbers are needed for a particular application, this problem reduces even further to tracking the connected components of a dynamic graph---sometimes referred to as the \emph{dynamic connectivity problem}, which can be efficiently  solved in amortized $O(\log n)$ query and update times using either Link-cut trees or multi-level Euler tour trees~\cite{kapron2013dynamic}.
+
+ \subsection{A Motivating Example}\label{sec:motivation} 
+We describe a simple experiment illustrating why the vineyards algorithm does not always yield an efficient strategy for time-varying applications. 
+Consider a series of grayscale images (i.e. a video) depicting a fixed-width annulus expanding about the center of a $9 \times 9$ grid. 
 Each image contains the same number of pixels, whose intensities vary with time.
@@
  	\centering
- 	\includegraphics[width=0.36\textwidth, height = 1.45in]{circle_vineyards.png}
- 	\includegraphics[width=0.63\textwidth, height = 1.45in]{circle_complex.png}
+ 	\includegraphics[width=0.45\textwidth,height=1.63in]{circle_vineyards.png}
+ 	\includegraphics[width=0.54\textwidth,height=1.63in]{circle_complex.png}
  	\caption{On the right, the grayscale image data set and corresponding lower star filtrations.
@@
 The  Betti numbers for each of the shown  steps in the filtration, from left to right, are: $(\beta_0,\beta_1, \beta_2)  = (1,0,0), \; (1,1,0),  \; (1,1,0), \; (1,1,0), \; (4,0,0)$.
-Observe that computing the diagrams independently at 10 evenly spaced time points (the green line) both captures these major topological changes and is the most computationally efficient approach shown.}
+Observe that computing the diagrams independently at 10 evenly spaced time points (the green line) both captures the major topological changes and is the most computationally efficient approach shown.}
  	\label{fig:vineyards}
@@
 
-  The existence of a time discretization that is more efficient to compute than continually updating the decomposition, indicates that the vineyards algorithm must incur some overhead (in terms of column operations) to perform transpositions, even if the persistence diagrams remain unchanged.
+  The existence of a time discretization that is more efficient to compute than continually updating the decomposition indicates that the vineyards algorithm must incur some overhead (in terms of column operations) to perform transpositions, even if the persistence diagrams remain unchanged.
   Even in situations where one is updating a decomposition to a filtration that is  relatively ``close''---e.g.
 any two adjacent filtrations in the bottom right of Figure~\ref{fig:vineyards}---it is still more efficient to apply \texttt{pHcol} directly than  to iteratively update the decomposition (see the case where $n = 10$).
- Moreover, any optimization to the reduction algorithm (e.g.,  clearing~\cite{chen2011persistent}) would only increase this disparity.
- See~\cite{topaz2015topological, xian2020capturing, lesnick2015interactive, kim2020spatiotemporal} 
- for relevant scenarios and applications.
-
-\textbf{Our approach and contributions are as follows:} First, we leverage the \emph{moves} framework of Busaryev et.
-al.~\cite{busaryev2010tracking} to include  coarser  operations---in terms of the number of valid intermediate decomposition states, as compared to vineyards---for dynamic persistence.
+ Moreover, any optimizations to the reduction algorithm (e.g.,  clearing~\cite{chen2011persistent}) would only increase this disparity.
+Since persistence has found many applications in dynamic contexts~\cite{topaz2015topological, xian2020capturing, lesnick2015interactive, kim2020spatiotemporal}, a more efficient alternative to the vineyards is clearly needed.
+
+\textbf{Our approach and contributions are as follows:} First, we leverage the \emph{moves} framework of Busaryev et al.~\cite{busaryev2010tracking} to include  coarser  operations---in terms of the number of valid intermediate decomposition states, as compared to vineyards---for dynamic persistence.
 We give a tight lower bound on the number of moves needed to perform an arbitrary permutation to the $R = D V$ decomposition, and give a proof of optimality by a reduction to the permutation edit distance problem.
@@
 
-\subsection{Related Work}\label{sec:related_work} 
-To the authors knowledge, work focused on ways of updating a  decomposition $R = DV$, for all homological dimensions, is limited---there is the vineyards algorithm~\cite{cohen2006vines} and the moves algorithm~\cite{busaryev2010tracking}, both of which are discussed extensively in section~\ref{sec:background}.
-
-In contrast, there is extensive work  on improving the efficiency of computing a (static) $R = DV$ decomposition.
-Chen~\cite{chen2011persistent} proposed \emph{persistence with a twist}, also called the \emph{clearing optimization}, which exploits a boundary/cycle relationship to ``kill'' columns early in the reduction rather than reducing them.
-Another popular optimization is to utilize the duality between   homology and cohomology~\cite{de2011dualities}, which dramatically improves the effectiveness of the clearing optimization~\cite{bauer2021ripser}.
-There are many other optimizations on the implementation side: the use of ranking functions defined on the combinatorial number system enables implicit cofacet enumeration, removing the need to store the boundary matrix explicitly; the apparent/emergent pairs optimization identifies columns whose pivot entries are unaffected by the reduction algorithm, reducing the total number of columns which need be reduced; sparse data structures such as bit-trees and lazy heaps allow for efficient column-wise additions with $\mathbb{Z}_2 = \mathbb{Z}/2\mathbb{Z}$ coefficients and effective $O(1)$ pivot entry retrieval, and so on~\cite{bauer2021ripser, bauer2017phat}.
-
-By making stronger assumptions on the underlying topological space, restricting the homological dimension, or targeting  a weaker invariant (e.g.
-Betti numbers), one can usually obtain  faster or alternative approaches  recording information related to persistence.
-For example, Attali et al.~\cite{attali2009persistence} give a linear time algorithm (under the $\texttt{word RAM}$ model of computation) for computing persistence on graphs.
-In the same paper, they describe how to obtain $\epsilon$-simplifications of $1$-dimensional persistence diagrams for filtered $2$-manifolds by using duality and symmetry theorems. 
-Along a similar vein, Edelsbrunner et.
-al.~\cite{edelsbrunner2000topological} give a fast incremental algorithm for computing persistent Betti numbers up to dimension $2$, again by utilizing symmetry, duality, and ``time-reversal''~\cite{delfinado1995incremental}.
-Chen et.
-al.~\cite{chen2013output} give an output-sensitive method for computing persistent homology,  utilizing the property that certain submatrices of  $D$ have the same rank as  $R$, which they exploit through fast sub-quadratic sparse-matrix rank algorithms.
-
-If  zeroth homology is the only dimension of interest, computing and updating both the persistence and rank information  is greatly simplified.
-For example, if the relations are available a-priori, obtaining a tree representation fully characterizing the connectivity of the underlying space (also known as the \emph{incremental connectivity} problem) takes just $O(\alpha(n) n)$ time using the disjoint-set data structure, where $\alpha(n)$ is the extremely slow-growing inverse Ackermann function.
-Adapting this approach to the time-varying setting, Oesterling et al.~\cite{oesterling2015computing} give a $O(e)$-per-update algorithm that allows one to maintain a \emph{merge tree} with $e$ edges,
-associated to the filtration parameter changing continuously over time. 
-If only the zeroth-dimensional Betti numbers are needed for a particular application, this problem reduces even further to tracking the connected components of a dynamic graph---sometimes referred to as the \emph{dynamic connectivity problem}, which can be efficiently  solved in amortized $O(\log n)$ query and update times using either Link-cut trees or multi-level Euler tour trees~\cite{kapron2013dynamic}.
+\subsection{Main results}
+Given a filtered simplicial complex $K$, denote by $R = D V$ the decomposition of its corresponding boundary matrix $D$ such that $R$ is reduced and $V$ is upper-triangular (see section~\ref{sec:reduction} for details).
+ If one has a pair of filtrations $K, K'$ of equal size $m = \lvert K \rvert = \lvert K' \rvert$ and $R = D V$ has been computed, then it may be advantageous to use the information stored in ($R$, $V$) to reduce the computation of $R' = D' V'$.
+Given the appropriate permutation $P$ such that $D = P^T D P$, such an update scheme has the form: 
+$$ (\ast P^T \ast R \ast P \ast) = (P^T DP)(\ast P^T \ast V \ast P \ast) $$
+where $\ast$ is substituted with elementary column operations that repair the permuted decomposition.
+It is known how to generate $d \sim O(m^2)$ decompositions requiring on the order of $c \sim O(m^2)$~\cite{cohen2006vines} such column operations.
+Since each column operation take $O(m)$, the complexity of obtaining these $d$ decompositions takes $O(m^3)$.
+This is efficient if all $d$ decompositions are needed; otherwise, updating $R \mapsto R'$ matches the complexity of computing $R = D V$ independently.
+
+We now summarize the main results obtained in this effort (Theorem~\ref{prop:sparsity_result}): suppose one has a sequence of permutations $\mathcal{S} = \left( s_1, s_2, \dots, s_d \right)$ yielding a corresponding sequence of decompositions:
+  \begin{equation}\label{eq:rv_seq}
+   	R = R_0 = D_0 V_0 \overset{s_1}{\to} D_1 V_{1} \overset{s_2}{\to} \dots \overset{s_d}{\to} D_d V_{d} = R_d = R'
+ \end{equation}
+ where each permutation $s_i$ is a particular type of cyclic permutation (see section~\ref{sec:moves_setting}).
+If $i_\ast < j_\ast$, our first result is that this sequence can be computed in output-sensitive $O(d\kappa)$ time:
+\begin{equation}
+	\quad \kappa = \sum\limits_{k = 1}^d \, \lvert \mathbb{I}_{k}\rvert + \lvert \mathbb{J}_{k}\rvert
+\end{equation}
+where the quantities $\lvert \mathbb{I}_{k} \rvert$ and $\lvert \mathbb{J}_{k} \rvert$ depend on the sparsity of the $V_k$ and $R_k$ matrices, respectively, and $d \sim O(m)$ is a constant that depends on how similar $K$ and $K'$ are.
+The advantage of this result is that it depends explicitly on the sparsity pattern of the decomposition itself, which we take advantage of in Section~\ref{sec:schedule_cost}.
+Moreover, if the vineyards algorithm requires $C$ column operations to obtain~\eqref{eq:rv_seq}, we show (Proposition~\ref{prop:factor2}) that our approach obtains the same sequence using at most $L$ column operations, where:
+\begin{equation}
+	\frac{C}{2} \leq L \leq C
+\end{equation}
+Our second result turns towards the computation of $\mathcal{S}$ itself, and towards bounding its size $\lvert \mathcal{S} \rvert = d$. 
+The complexity of determining $\mathcal{S}$ depends on the type of permutations considered; many edit distances exhibit tight quadratic growth rates and tend to require at least quadratic time (or worse) to compute~\cite{labarre2013lower}.
+We consider a special set of cyclic permutations whose sequences scale linearly in $m$ for all permutations in $S_m$: surprisingly, we find that the \emph{smallest} such sequence of these permutations which map $K \mapsto K'$ has size $d$ (Proposition~\ref{prop:schedule_size}), where:
+\begin{equation}
+	d = m - \lvert \mathrm{LCS}(K, K') \rvert 
+\end{equation}
+Moreover, we also show that the information needed to construct any $\mathcal{S}$ with optimal size $\lvert \mathcal{S} \rvert = d$ can be computed in $O(m \log \log m)$ preprocessing time and $O(m)$ memory.
+Although $d$ clearly is $O(m)$ for pathological inputs, we show that $d \sim m - \sqrt{m}$ in expectation for random filtrations (Corollary~\ref{cor:expectation}) is not too pessimistic for truly random filtrations, and we give empirical results suggesting $d$ is tends to be much smaller for time-varying filtrations.
+
+In summary, we obtain an output-sensitive persistence update scheme which depends explicitly on the sparsity pattern of the involved matrices, has sub-cubic complexity in expectation even for random filtrations, and we show performs well in practice on a variety of applications.
+
+% given a decomposition $R = \partial_{K} V$ of a simplicial filtration $K$, algorithm~\ref{alg:schedule} can produce such a sequence $\mathcal{S} = \left( s_1, s_2, \dots, s_d \right)$
+%
+
+ %The actual permutations $s_\ast$ are generated in $O(m \log m)$ time prior to the execution of the column operations.
+
+	% where $\lvert \mathbb{I}_{(k)} \rvert$ and $\lvert \mathbb{J}_{(k)} \rvert$ are the computed quantities in lines (2) and (3) in algorithm~\ref{alg:mr} (here we assume if $i_\ast < j_\ast$).
+%	where $\lvert \mathbb{I}_{(k)} \rvert$ denotes the number of non-zero entries $V_k[i_k, i_k:j_k]$ and $\lvert \mathbb{J}_{(k)}\rvert$ are the number of columns satisfying $\mathrm{low}_{R_k} \in [i_k:j_k]$ and $\mathrm{row}_{R_k}(i_k) \neq 0$.
+
+%We mention in passing that the vineyards algorithm produces a similar sequence $\mathcal{S}'$ of size $\lvert \mathcal{S'} \rvert = d$ where $d = K_\tau(K,K')$ in output sensitive $O(m\log m + d)$ time and $O(d)$ memory, where $d = K_\tau(K,K')$ is the output sensitive terms denoting the Kendall-$\tau$ distance between the two filtrations $K$ and $K'$.
+The corresponding sequence $\partial_{K_1} V_1 \dots \partial_{K_d} V_d$ is then obtained in $O(md)$ time.
+%Since $K_\tau(K,K') \sim O(m^2)$, both the time and storage complexities become significant factors in many practical contexts in the vineyards algorithm, and it is difficult to determine the see how the sparsity of the underlying decomposition affects the runtime.
+In comparison, our proposed approach yields a smaller decomposition sequence using less running time and memory, and conveys more directly how much the sparsity of the corresponding decompositions affect the computation.
+
+%Below, we give a corollary of the main result which conveys more about the worst-case size of $\mathcal{S}$. 
+%\begin{corollary}\label{cor:d_limit}
+%	Let $K$, $K'$ be two uniformly random filtrations of equal sizes $m$.
+Then the expected size of $\lvert \mathcal{S} \rvert$ when $\mathcal{S}$ is generated using Algorithm~\ref{alg:schedule} is bounded above by $O(m - \sqrt{m})$ in expectation, with probability $1$ as $m \to \infty$.
+%\end{corollary}
+%\noindent By upper bounding the size of $\mathcal{S}$, corollary~\ref{cor:d_limit} also bounds the number of decompositions generated in the corresponding sequence.
+This result uses a known bound on the expected size of the longest common subsequence between two uniformly random permutations.
+Due to the face relation requirement of abstract simplicial complexes, one might expect that this bound might be much smaller, however for low-dimensional complexes we give evidence that this tightness gap in tightness may not be so large, see section~\ref{sec:schedule_sizes} for more details.
+
+%\begin{proposition}\label{prop:factor2_main}
+%	Let $R = D V$ be a valid decomposition of size $(m \times m)$, let $C$ denote the number of $O(m)$ operations required to execute schedule $S_F$, and similarly let $L$ be the number of $O(m)$ operations required to execute a coarsened schedule $\widetilde{S}_F$ of $S_F$.
+We have the following relationship between $C$ and $L$:
+%	$$ \frac{C}{2} \leq L \leq C$$ 
+%\end{proposition}
 
@@
 Such a family is called a \emph{filtration},
-which is deemed to be \emph{essential} if $i \neq j$ implies $K_i \neq K_j$.
-Moreover, an essential filtration is said to be  \emph{simplexwise} if $K_j \smallsetminus K_i = \{\sigma_j\}$ 
+which is deemed to be\emph{simplexwise} if $K_j \smallsetminus K_i = \{\sigma_j\}$ 
 whenever $j$ is the immediate successor of $i$ in $I$.
-Any finite filtration may be trivially converted into an essential, simplexwise filtration via a set of \emph{condensing}, \emph{refining}, and \emph{reindexing} maps (see~\cite{bauer2021ripser} for more details).
-As a result, and without loss of generality, here we exclusively consider essential simplexwise filtrations. For brevity, we will simply refer to them as filtrations.
+Any finite filtration may be trivially converted into an simplexwise filtration via a set of \emph{condensing}, \emph{refining}, and \emph{reindexing} maps (see~\cite{bauer2021ripser} for more details).
+Without loss of generality, we exclusively consider simplexwise filtrations but for brevity-sake refer to them simply as filtrations.
 
@@
 The snapshot $f_0$ denotes the initial function at time $\tau = 0$ and $f_1$ denotes the function at the last time step. 
-As $\tau$ varies in $[0,1]$, the points in $\mathrm{dgm}_p(f_\tau)$ trace  curves in $\mathbb{R}^3$, called \emph{vines} and which together form a \emph{vineyard}. 
+As $\tau$ varies in $[0,1]$, the points in $\mathrm{dgm}_p(f_\tau)$ trace curves in $\mathbb{R}^3$, called \emph{vines}, which collectively form a \emph{vineyard}. 
 The stability of persistence implies that these curves will be continuous if $F$ is continuous
@@
 We use $f_\tau$ to denote the function at $F(x, \tau)$, where $\tau \in [0, 1]$.
+
+%\begin{definition}\label{def:schedule}
+%Given a pair of filtrations $K_0, K_1$ of equal size $m$ and a fixed bijection $f : K_0 \to K_{1}$, let $f^\ast : [m] \to [m]$ denote its induced bijection on $S_m$.
+A \emph{schedule}  $\mathcal{S} = \{s_1, s_2, \dots, s_d \}$ is a sequence of permutations mapping $K_0 \mapsto K_1$ that satisfies:
+%	\begin{equation}\label{eq:rv_seq}
+%   		R = D_0 V_0 \overset{s_1}{\to} D_1 V_{1} \overset{s_2}{\to} \dots \overset{s_d}{\to} D_d V_{d} = R'
+%   	\end{equation}
+%   	where, for each $i \in [d]$, $R_i = D_i V_i$ is a valid decomposition. 
+%\end{definition}
+
 \subsection{The Reduction Algorithm}\label{sec:reduction}
@@
   \textbf{ Decomposition Invariants:}
- \begin{enumerate}[labelsep=2pt, topsep=0pt,itemsep=-0.25ex,parsep=1.2ex]
- 	\item[I1] $R = D V$ where $D$ is the boundary matrix of the filtration $K$
- 	\item[I2] $V$ is full-rank upper-triangular, and $R$ is \emph{reduced}: if its $i$-th and $j$-th columns are nonzero, \\ then $\mathrm{low}_R(i) \neq \mathrm{low}_R(j)$ 
+ \begin{enumerate}[labelsep=3pt, topsep=2pt,itemsep=-0.25ex,parsep=1.0ex]
+ 	\item[I1.] $R = D V$ where $D$ is the boundary matrix of the filtration $K$
+ 	\item[I2.] $V$ is full-rank upper-triangular, and $R$ is \emph{reduced}: if its $i$-th and $j$-th columns are nonzero, then $\mathrm{low}_R(i) \neq \mathrm{low}_R(j)$ 
  	\end{enumerate} 
@@
 \begin{displaymath}
-	\def\rowsp{0.4em}
+	\def\rowsp{0.2em}
 	\def\arraystretch{0.95}
@@
 	\begin{blockarray}{cccc}{R_1}
-	 \kern-0.2em & a & b & c  \\
+	 \kern-0.2em  \vspace{-0.3em} & a & b & c  \\
 		\begin{block}{c[ccc]}
@@
 	\begin{blockarray}{cccc}{}
-	\kern-0.2em & a & b & c  \\
+	\kern-0.2em \vspace{-0.3em} & a & b & c  \\
 		\begin{block}{c[ccc]}
@@
 	\begin{blockarray}{cccc}{}
-	\kern-0.2em & b & a & c  \\
+	\kern-0.2em \vspace{-0.3em} & b & a & c  \\
 		\begin{block}{c[ccc]}
@@
 \end{displaymath}
-\vspace{-3.0em}
-% Start V 
+% End R1
+\vspace{-1.8em}
+% Start V1 
 \begin{displaymath}	
-	\def\rowsp{0.4em}
+	\def\rowsp{0.2em}
 	\def\arraystretch{0.95}
@@
 	\begin{blockarray}{cccc}{V_1}
-	\kern-0.2em & a & b & c  \\
+	\kern-0.2em \vspace{-0.3em} & a & b & c  \\
 		\begin{block}{c[ccc]}
@@
 Since moving a simplex from $i$ to $j$ using vineyards performs $\lvert i - j \rvert - 1$ transpositions, this query operation is required at most $O(\lvert i - j \rvert)$ times.
-By comparison, if a row-oriented sparse matrix implementation is used, then a move operation can perform these non-zero entry checks in just one $O(m)$ pass, prior to performing any column operations.
+By comparison, if a row-oriented sparse matrix implementation is used, then a move operation can perform these non-zero entry checks in just one $O(\lvert i - j \rvert)$ pass prior to performing any column operations, saving a logarithmic factor.
 
@@
         \Function{MoveLeft}{$R$, $V$, $i$, $j$} 
-        	\State $\mathbb{J} \gets \{ j \}$
+        	\State $\mathbb{I} \gets \{ j \}$
             % \State $(\mathbb{K}, d_R, d_V) \gets (\emptyset, \mathrm{col}_R(i), \mathrm{col}_V(i))$
@@
             	\State $(\mathrm{col}_R(i), \mathrm{col}_V(i)) \mathrel{+}= (\mathrm{col}_R(k), \mathrm{col}_V(k))$
-            	\State $\mathbb{K} \gets \mathbb{K} \cup k + 1 $
+            	\State $\mathbb{I} \gets \mathbb{I} \cup k + 1 $
             \EndWhile
@@
 
-            \State $(R, V) \gets$ \Call{RestoreLeft}{$R, V, \mathbb{K}$}
+            \State $(R, V) \gets$ \Call{RestoreLeft}{$R, V, \mathbb{I}$}
             \State $(R, V) \gets$ \Call{RestoreLeft}{$R, V, \mathbb{J}$}
@@
 
-\subsection{Proposed approach}
-Let us begin with a brief overview of the pipeline, which we also outline in Algorithm~\ref{alg:schedule} below.
-As before, we assume as input  a discrete $1$-parameter family of filtrations $\mathcal{K} = (K_1, K_2, \dots, K_n)$ of an abstract simplicial complex $K$ with $|K|=m$, 
-and the goal being to compute their persistence diagrams.
-The pipeline can be summarized as follows.
-
+%\subsection{Proposed approach}
+Let us begin with a brief overview of the pipeline, which we outline in pseudocode with Algorithm~\ref{alg:schedule} below.
+As before, we assume as input  a discrete $1$-parameter family of filtrations $\mathcal{K} = (K_1, K_2, \dots, K_n)$ of an abstract simplicial complex $K$ with $|K|=m$, and the goal being to compute their persistence diagrams.
 Fix reindexing  bijections $f_i : K_i \to K_{i+1}$. 
@@
 Denote these permutations by $p$ and $q$.
-We compute the longest increasing subsequence (LIS) of $q$ and use this subsequence to recover a longest common subsequence (LCS) $\mathrm{LCS}(p,q)$.
+We compute the longest increasing subsequence (LIS) of $q$ and use this subsequence to recover a longest common subsequence (LCS) of $p, q$, denoted by $\mathrm{LCS}(p,q)$.
 We pass $p$, $q$, and this LCS to our greedy scheduling algorithm, which returns as output an ordered set of move permutations  $\mathcal{S}$ of minimum size, which we call a \emph{schedule}.
-Not all valid schedules are created equal: some incur a much larger number of column operations, even if they have a minimal number of moves; we address this issue in Section~\ref{sec:schedule_cost}.
-Given the set of schedules for each pair $(K_i, K_{i+1})$, we proceed to computing an $R=DV$ decomposition for the first filtration $K_1$, and then execute the moves stored in $\mathcal{S}$ in sequence.
+Note one need not explicitly store the entire family of filtrations nor the schedules between them---Algorithm~\ref{alg:schedule} may be easily modified to be completely \emph{online}, keeping at most two filtrations and one decomposition in memory at any given time.
+% Not all valid schedules are created equal: some incur a much larger number of column operations, even if they have a minimal number of moves; we address this issue in Section~\ref{sec:schedule_cost}.
+% Given the set of schedules for each pair $(K_i, K_{i+1})$, we proceed to computing an $R=DV$ decomposition for the first filtration $K_1$, and then execute the moves stored in $\mathcal{S}$ in sequence.
 During execution, we obtain a valid decomposition $R = DV$ for each filtration $K_1, K_2, \dots, K_n$ from which we extract persistence diagrams along the way.
@@
     \begin{algorithmic}[1]
-    	\Require Ordered set of filtrations $\mathcal{K}$, each of size $m$, with bijections $f_{i}: K_i \to K_{i+1}$
-    	\Ensure Valid $R = D V$ decompositions are computed for each filtration $K_1, K_2, \dots, K_n$
-    	\Procedure{CoarseSchedule}{$\mathcal{K} = ( K_1, K_2, \dots, K_n )$} 
+    	\Require Ordered set of filtrations $\mathcal{K}$ with bijections $f_{i}: K_i \to K_{i+1}$
+    	\Ensure $R = D V$ is computed for each  $K_1, K_2, \dots, K_n$
+    	\Procedure{MoveSchedule}{$\mathcal{K} = ( K_1, K_2, \dots, K_n )$} 
     		\State $\mathcal{S} = \emptyset$
@@
     			\State $\mathrm{lis}_q \gets \mathrm{LIS}(q)$ \Comment{$O(m \log \log m)$}
-    			\State $\mathcal{S} \gets \mathcal{S} \cup \mathrm{GreedySchedule}(p, q, \mathrm{lis}_q)$ \Comment{$O(d^2\log m)$, see section~\ref{sec:proxy_objective}}
+    			\State $\mathcal{S} \gets \mathcal{S} \cup \mathrm{GreedySchedule}(p, q, \mathrm{lis}_q)$ \Comment{$O(d^2\log m)$}
     		\EndFor
@@
 	    	\For{$(i,j)$\textbf{ in }$\mathcal{S}$}
-	    		\If{$i < j$}
-	    			\State $(R, V) \gets$ \Call{MoveRight}{i, j}
-	    		\Else
-	    			\State $(R, V) \gets$ \Call{MoveLeft}{i, j}
-	    		\EndIf
+	    		\State $(R, V) \gets$ \textbf{if} $i < j$ \Call{MoveRight}{i, j} \textbf{else} \Call{MoveLeft}{i, j}
 	    	\EndFor
@@
 \end{algorithm}
-\noindent Algorithm~\ref{alg:schedule} is purely illustrative at this point, 
-and is meant to serve as a guidepost before discussing its  components in depth. 
+%\noindent Algorithm~\ref{alg:schedule} is purely illustrative at this point, and is meant to serve as a guidepost before discussing its  components in depth.
 Specifically, lines (4-5) are discussed next in section~\ref{sec:schedule_sizes}, and the algorithm for line (6) is given in section~\ref{sec:schedule_cost}.
@@
 the individual schedules $S_i$ may be computed on the fly.
-Due to the combinatorial aspect of our approach, there is actually no need to have access to---or explicitly store---the entire family of filtrations and the schedules between them.
-That is, Algorithm~\ref{alg:schedule} can be modified to be completely online, keeping at most two filtrations and one decomposition in memory at any given time.
 
@@
 Here we describe the reasoning for lines (4-5) in Algorithm~\ref{alg:schedule}.
-Recall in section~\ref{sec:motivation}, an example was given showing the vineyards algorithm efficiency compared to a naïve approach at three varying levels of coarseness.
+Recall in section~\ref{sec:motivation}, an example was given showing the vineyards algorithm efficiency compared to a naïve approach at three varying levels of coarseness, where it was shown that vineyards provides an efficient solution with a fine enough update scheme but otherwise scales poorly if the discretization of time is coarse.
+One may attempt to simply permute the decomposition and then repair it, however its known that applying arbitrary permutations to a given decomposition may require $O(m^3)$ column operations to ensure validity---this was one of the motivations for the creation of the vineyards algorithm.
+
 It was hypothesized that the reason the vineyards algorithm was more expensive than the naïve approach is due to the extra overhead of maintaining the decomposition at each transposition.
 Thus, decreasing the number of times the decomposition is restored to a valid state ought to reduce the total number of column operations needed to update a decomposition.
-This motivates the following question:
-in order to apply an \emph{arbitrary} permutation $P$ to a given $R = DV$ decomposition, 
-what is the minimal number of times the decomposition needs to be restored to a valid state?
+This motivates the question: can one simultaneously minimize the number of times the decomposition is restored to a valid state and retain an efficient update scheme for \emph{arbitrary} permutations?
 
@@
 If we assume that the curves $\tau \mapsto F(\cdot, \tau)$  are in general position in $[0,1]\times \mathbb{R}$, and that each pair of curves cross  at most once, then the set of adjacent transpositions $S_F$ that $F$ decomposes into is given by the pairs of simplices in $K$ with changes in their relative ordering:
-\begin{align}\label{eq:sf_schedule}
-	S_F &= \{ (i,j) \, \mid \, f_0(\sigma_i) < f_0(\sigma_j) \mbox{ and }  f_1(\sigma_i) > f_1(\sigma_j) \quad  \; i,j \in [m] \} \\
+\begin{align*}\label{eq:sf_schedule}
+	S_F &= \{ (i,i+1) \, \mid \, f_0(\sigma_i) < f_0(\sigma_{i+1}), \;  f_1(\sigma_i) > f_1(\sigma_{i+1}), \; i \in [m] \} \\
 	&= \{ s_1, s_2, \dots, s_k \} 
-\end{align}
+\end{align*}
 %We assume, without loss of generality, that the underlying complexes $K_0$ and $K_1$ have the same cardinality. 
@@
 \[S_F = 
-\mathrm{Inv}(p, q) = \{ \, (i,j) \, \mid \, p(i) < p(j) \mbox{ and } q(i) > q(j)  \quad \; i,j \in [m]\, \}
+\mathrm{Inv}(p, q) = \{ \, (i,i+1) \, \mid \, p(i) < p(i+1), \; q(i) > q(i+1), \; i \in [m]\, \}
 \]
-The cardinality of $  S_F $ is exactly the \emph{Kendall}-$\tau$ distance $K_\tau(p,q)$ between $p$ and $q$: 
+The cardinality of $  S_F $ is exactly the \emph{Kendall}-$\tau$ distance~\cite{diaconis1977spearman} between $p$ and $q$: 
 \begin{equation}\label{eq:kendall_dist}
@@
  This quadratic scaling can induce a number of issues in practical implementations of vineyards.
- Indeed, although many transpositions require no column reductions in practice, detecting the existence of non-zero entries at specific positions in the sparse matrix takes $O(\log(m))$ time. 
- Since this check is required after each permutation is applied, it is in some sense unavoidable in the vineyards algorithm.
+ Indeed, although many transpositions require no column reductions in practice, detecting the existence of non-zero entries at specific positions in the sparse matrix takes $O(\log(m))$ time per transposition.
 
@@
 \end{proposition}
-\begin{proof}
+\begin{proof} \normalsize
 Consider executing the vineyards algorithm on a given schedule $S_F$.
@@
 	Now consider using the MoveRight($i$,$j$) for a single move operation, as a replacement for the sequence of transpositions above.
-Here again, the dominant cost again are the column operations (line 5), since although permutations take $O(\lvert i - j\rvert)$ time we require only $1$ such permutation per movement.
-	If we allocate an extra $O(m)$ memory space during the movement, notice that lines (4), and (7) of Restore* (but not line (2)) and line (7) of Move* in Algorithm~\ref{alg:mr} can be completed $O(1)$ via a pointer swapping argument.
-	Since a move operation requires at most $1$ column operation for each index in $[i: j]$, moving a column from $i$ to $j$ where $i < j$ requires at most $2(\lvert i - j \rvert)$ column operations.
+Here, the dominant cost again are the column operations (line 5), since although permutations take $O(\lvert i - j\rvert)$ time we require only $1$ such permutation per movement.
+	If we allocate an extra $O(m)$ memory space for the donor columns $d_\ast$ prior to the movement, notice that the assignment to and from $d_\ast$ (lines (4), and (7) in RestoreRight of MoveRight in Algorithm~\ref{alg:mr}) can be completed in $O(1)$ time by storing the column memory addresses as pointers and swapping their pointers.
+When $d_\mathrm{low}' < d_\mathrm{low}$, instead of copying $\mathrm{col}_\ast(k)$ to $d_\ast'$, which takes $O(m)$ time, we swap their column pointers in $O(1)$ and proceed by modifying the donor column $d_\ast$ with the necessary column operations.
+After the operations, $d_\ast$ contains the newly modified column and $\mathrm{col}_\ast(k)$ contains the unmodified donor $d_\ast'$, so we swap them once more in $O(1)$ time.
+Since we require at most $1$ $O(m)$ column operation for each index in $[i: j]$, moving a column from $i$ to $j$ where $i < j$ requires at most $2(\lvert i - j \rvert)$ column operations for both $R$ and $V$.
 The claimed inequality follows.
@@
 In terms of bounds, we have $\lvert \widetilde{S}_F \rvert \leq \lvert S_F \rvert$, and the associated coarsened schedule $\widetilde{S}_F$ can be computed in $O(m)$ time. 
-However, the coarsening depends entirely on the choice of $F$ and the upper bound from~\ref{eq:sf_m2} remains, as its always possible that there are no contiguous subsequences to collapse.
+However, the coarsening depends entirely on the choice of $F$ and the upper bound from~\eqref{eq:sf_m2} remains, as its always possible that there are no contiguous subsequences to collapse.
 If a straight line homotopy is used, the number of collapsible subsequences may depend on the simplices categorization as critical points; transpositions between pairs of simplices which trigger  changes in homology tend to be more expensive.
@@
 
-\subsubsection{The Moves Setting} 
-As we mentioned before, vineyards maintains 
-$j-i$ valid decompositions while 
-performing the sequence 
-$(i,i+1), (i+1,i+2), \ldots, (j-1,j)$
-of adjacent transpositions. 
+\subsubsection{The Moves Setting}\label{sec:moves_setting}
+As we mentioned before, vineyards maintains $j-i$ valid decompositions while 
+performing the sequence $(i,i+1), (i+1,i+2), \ldots, (j-1,j)$ of adjacent transpositions. 
 The benefit of move operations is that one need not maintain a valid decomposition during the movement $(i,j)$, only at the end; thus, it is worth investigating whether the vineyard bounds can be improved.
-In the case of vineyards, the bound is clear: 
-if $\kappa = K_\tau(K_0, K_1)$ is the Kendall-$\tau$ distance, then $\Omega(\kappa)$ adjacent transpositions are required to transform $K_0 \mapsto K_1$. 
+In the case of vineyards, the bound is clear: if $\kappa = K_\tau(K_0, K_1)$ is the Kendall-$\tau$ distance, then $\Omega(\kappa)$ adjacent transpositions are required to transform $K_0 \mapsto K_1$. 
 In contrast, if we allow move operations in any order, we have $O(m)$ as a trivial upper bound: simply move each simplex in $K_0$ into its position in the filtration given by $K_1$ in the order given by the latter.
@@
 Let $S_m$ denote the symmetric group  of   permutations of  $[m]$ under function composition $\circ$.
-Given two fixed permutations $p, q \in S_m$ and a set $\Sigma \subseteq S_m$,  the following 
-are common problems in the literature:
-\begin{enumerate}
-	\item Find an ordered sequence of permutations $s_1, s_2, \dots, s_d \in \Sigma$ whose composition transforms $p$ into $q$:
-	\begin{displaymath}
+Given two fixed permutations $p, q \in S_m$ and a set $\Sigma \subseteq S_m$,  a common  problem is to find an ordered sequence of permutations $s_1, s_2, \dots, s_d \in \Sigma$ whose composition transforms $p$ into $q$:
+	\begin{equation}\label{eq:sorting}
 		s_d \circ \cdots \circ s_2 \circ s_1 \circ p = q
-		% p \sbullet s_1 \sbullet s_2 \sbullet \cdots \sbullet s_d = q
-	\end{displaymath}
-	\item Find a sequence satisfying (1) of minimal length ($d$)
-	\item Find the minimal length $d$, referred to as the \emph{distance} between $p$ and $q$ with respect to $\Sigma$. 
-\end{enumerate}
+	\end{equation}
+Common variations of this problem include finding such a sequence of minimal length ($d$) and bounding the length $d$ as a function of $m$.
+In the latter case, a lower bound on $d$ is referred to as the \emph{distance} between $p$ and $q$ with respect to $\Sigma$.
 A sequence $S = \{s_1, s_2, \dots, s_d\} \subseteq \Sigma \subseteq S_m$ of operations mapping $p \mapsto q$ is sometimes called a \emph{sorting of $p$}. 
-In the context of performing updates to the $R = D V$ decomposition using moves,  such sequences will form schedules. 
-There are typically many ways to solve (1), and any solution to (2) necessarily solves (1) and (3).
 When $p, q$ are interpreted as strings, distances defined with respect to a fixed $\Sigma \subseteq S_m$ are commonly referred to as \emph{edit distances}~\cite{labarre2013lower}, which are denoted as $d_\Sigma(\cdot, \cdot)$, and the subsequent elementary operations on the strings are called \emph{edit operations}.
-%Note that here we describe distances between permutations (e.g. $p,q \in S_n$) using operations that can themselves be expressed as permutations (e.g. $s \in S \subseteq S_n$), and that 
-The choice of $\Sigma$ defines the universe of operations and thus the type of edit distance being measured---otherwise if $\Sigma = S_m$, then $d_\Sigma(p, q) = 1$ for any $p\neq  q \in S_m$.
-
-Many types of edit distances can be described succinctly. 
-Let $\beta(i,j,k,l)$ with $1 \leq i < j \leq k < l \leq m$ denote the permutation that exchanges the two closed intervals given by $i, j-1$ and $k,l-1$:
-\[
-\setlength\arraycolsep{1.8pt}
-\scalefont{0.90}{
-\setcounter{MaxMatrixCols}{20}
-\left (
-\begin{array}{*3c|*4c|*4c|*4c|*{5}c } 
-\cline{4-7} \cline{12-15}  \rule{0pt}{2ex}
-1 & \cdots & i -1 & i & i + 1 & \cdots & j - 1 & j & j +1 & \cdots & k - 1  & k & k+1 & \cdots & l - 1 & l & l + 1 & \cdots & m \\
-\cline{4-7} \cline{12-15} 
-\cline{4-7} \cline{12-15}  \rule{0pt}{2ex}
-1 & \cdots & i -1 & k & k+1 & \cdots & l - 1 & j & j +1 & \cdots & k - 1  & i & i + 1 & \cdots & j - 1 & l & l + 1 & \cdots & m \\
-\cline{4-7} \cline{12-15} 
-\end{array}
-\right ) 
-}
-\]
-Common edit distances found in the literature expressed as specializations of $\beta$ are given below:
-\begin{enumerate}
-	\item (no restriction on $i,j,k,l$) $\implies$ $\beta$ is called a \emph{block interchange}
-	\item $j = k$ $\implies$ $\beta$ swaps adjacent intervals, called a \emph{transposition}
-	\item $j = i+1$ and $l = k+1$ $\implies$ $\beta$ swaps two not necessarily adjacent elements, called an \emph{exchange}\footnote{In comparative genomics, exchanging two elements is sometimes called a \emph{super short transposition}.}
-\end{enumerate}
+Note the choice of $\Sigma$ defines the universe of operations and thus the type of edit distance being measured---otherwise if $\Sigma = S_m$, then $d_\Sigma(p, q) = 1$ for any $p\neq  q \in S_m$.
+
 Both the difficulty of obtaining a minimal sorting and the size of $d_\Sigma(p,q)$ vary dramatically with the choice of $\Sigma$. 
-For example, while sorting by transpositions and reversals is NP-hard, and sorting by prefix  transpositions (i.e., when $i=1$ always) is unknown, there are polynomial time algorithms for sorting by block interchanges, exchanges, and prefix exchanges~\cite{labarre2013lower}.
+For example, while sorting by transpositions and reversals is NP-hard, and sorting by prefix  transpositions is unknown, there are polynomial time algorithms for sorting by block interchanges, exchanges, and prefix exchanges~\cite{labarre2013lower}.
 Sorting by adjacent transpositions can be achieved in many ways: any sorting algorithm that exchanges two adjacent elements during its execution (e.g.
@@
 
-Here we consider sorting by moves. In terms of permutations, a permutation $m_{ij}$ that moves $i$ to $j$ in $[m]$, for $i < j$, corresponds to: 
-\[
+In the context of filtered simplicial filtration, we refer to sortings that repair the decomposition as \emph{schedules}. 
+\begin{definition}[Schedule]\label{def:schedule}
+Given a pair of filtrations $K_0, K_1$ of equal size $m$ and a fixed bijection $f : K_0 \to K_{1}$, let $f^\ast : [m] \to [m]$ denote its induced bijection on $S_m$.
+A \emph{schedule}  $\mathcal{S} = \{s_1, s_2, \dots, s_d \}$ is a sequence of permutations mapping $K_0 \mapsto K_1$ that satisfies:
+	\begin{equation}\label{eq:rv_seq}
+   		R = D_0 V_0 \overset{s_1}{\to} D_1 V_{1} \overset{s_2}{\to} \dots \overset{s_d}{\to} D_d V_{d} = R'
+   	\end{equation}
+   	where, for each $i \in [d]$, $R_i = D_i V_i$ is a valid decomposition: $R_i$ is \emph{reduced} and $V_i$ is full rank upper-triangular. 
+\end{definition}
+% Note that one cannot arbitrarily permute a given filtration $K$ and its corresponding decomposition: one must respect the face poset.
+% Given a filtered simplicial complex $K$, denote by $R = D V$ the decomposition of its corresponding boundary matrix $\partial_K$ such that $R$ is reduced and $V$ is upper-triangular.
+% If one has a pair of filtrations $K, K'$ of equal size $m = \lvert K \rvert = \lvert K' \rvert$ and $R = \partial_{K} V$ has been computed, then it may be advantageous to use the information stored in ($R$, $V$) to reduce the computation of $R' = \partial_{K'} V'$.
+Our first result makes this more explicit.
+%   \begin{proposition}\label{prop:preprocessing_time}
+%   Given a decomposition $R = \partial_{K} V$ of a simplicial filtration $K$, algorithm~\ref{alg:schedule} can produce such a sequence $\mathcal{S} = \left( s_1, s_2, \dots, s_d \right)$ of permutations with size $\lvert \mathcal{S} \rvert \sim O(m)$ which induces a sequence of decompositions:
+%   	\begin{equation}\label{eq:rv_seq}
+%   		R = D_0 V_0 \overset{s_1}{\to} D_1 V_{1} \overset{s_2}{\to} \dots \overset{s_d}{\to} D_d V_{d} = R'
+%   	\end{equation}
+%	in $O(m \log \log m)$ time and $O(m)$ memory.
+
+%In the context of performing updates to the $R = D V$ decomposition using moves,  such sequences will form schedules. 
+%There are typically many ways to solve (1), and any solution to (2) necessarily solves (1) and (3).
+
+%Many types of edit distances can be described succinctly. 
+%Let $\beta(i,j,k,l)$ with $1 \leq i < j \leq k < l \leq m$ denote the permutation that exchanges the two closed intervals given by $i, j-1$ and $k,l-1$:
+%\[
+%\setlength\arraycolsep{1.5pt}
+%\scalefont{0.90}{
+%\setcounter{MaxMatrixCols}{20}
+%\left (
+%\begin{array}{*3c|*4c|*4c|*4c|*{5}c } 
+%\cline{4-7} \cline{12-15}  \rule{0pt}{2ex}
+%1 & \cdots & i -1 & i & i + 1 & \cdots & j - 1 & j & j +1 & \cdots & k - 1  & k & k+1 & \cdots & l - 1 & l & l + 1 & \cdots & m \\
+%\cline{4-7} \cline{12-15} 
+%\cline{4-7} \cline{12-15}  \rule{0pt}{2ex}
+%1 & \cdots & i -1 & k & k+1 & \cdots & l - 1 & j & j +1 & \cdots & k - 1  & i & i + 1 & \cdots & j - 1 & l & l + 1 & \cdots & m \\
+%\cline{4-7} \cline{12-15} 
+%\end{array}
+%\right ) 
+%}
+%\]
+%Common edit distances found in the literature expressed as specializations of $\beta$ are given below:
+%\begin{enumerate}
+%	\item (no restriction on $i,j,k,l$) $\implies$ $\beta$ is called a \emph{block interchange}
+%	\item $j = k$ $\implies$ $\beta$ swaps adjacent intervals, called a \emph{transposition}
+%	\item $j = i+1$ and $l = k+1$ $\implies$ $\beta$ swaps two not necessarily adjacent elements, called an \emph{exchange}\footnote{In comparative genomics, exchanging two elements is sometimes called a \emph{super short transposition}.}
+%\end{enumerate}
+\noindent Here we consider sorting by moves. Phrased using permutations, a \emph{move operation} $m_{ij}$ that moves $i$ to $j$ in $[m]$, for $i < j$, corresponds to the permutation: 
+\begin{equation}
 \setlength\arraycolsep{1.8pt}
@@
 \begin{array}{*3c | *5c | *{3}c } 
-\cline{4-8} \rule{0pt}{2ex}
+\cline{4-8} \rule{0pt}{1ex}
 1 & \cdots & i -1 & i & i + 1 & \cdots & j - 1 & j & j +1 &  \cdots & m \\
 \cline{4-8}
-\cline{4-8}\rule{0pt}{2ex}
+\cline{4-8}\rule{0pt}{1ex}
 1 & \cdots & i -1 & i + 1 & \cdots & j - 1 & j & i & j +1 &  \cdots & m \\
@@
 }
-\]
-In the context of edit operations, observe that a move operation can be interpreted as the composition of a deletion operation followed by an insertion operation: 
+\end{equation}
+In cycle notation, when $i < j$, this corresponds to the cyclic permutation 
+\begin{equation}\label{eq:move_permutation}
+(\enspace i \enspace j \enspace j\text{-}1 \enspace \dots \enspace i\text{+}2 \enspace i\text{+}1 \enspace )
+\end{equation}
+Similarly, in the context of edit operations, observe that a move operation can be interpreted as a paired delete-and-insert operation: 
 $$ m_{ij} = (\mathrm{ins}_j \circ \mathrm{del}_i) $$
 where $\mathrm{del}_i$ denotes the edit operation that deletes the character at position $i$  and $\mathrm{ins}_j$ similarly inserts the same character at position $j$. 
-Thus, sorting by move operations can be interpreted as finding a minimal sequence of edits where the only operations allowed are insertions and deletions, where each insertion/deletion operation is coupled.
-This is exactly the LCS (Longest Common Subsequence) distance. It is well known~\cite{bergroth2000survey} that the LCS distance between two strings $p, q$ of sizes $m$ and $n$, respectively, is: 
+Thus, sorting by move operations can be interpreted as finding a minimal sequence of edits where the only operations allowed are insertions and deletions, where each insertion/deletion operation is coupled---which is exactly the well known \emph{Longest Common Subsequence} (LCS) distance.
+The LCS distance between two strings $p, q$ of sizes $m$ and $n$, respectively, is given by~\cite{bergroth2000survey}:
 \begin{equation}
@@
 Intuitively, in the context of strings, given $\mathcal{L} = \mathrm{LCS}(p,q)$ one may transform $p \mapsto q$ by simply deleting the symbols from $p \setminus \mathcal{L}$ and then inserting symbols from $q \setminus \mathcal{L}$.
-In the context of permutations, these operations are coupled, thus to ensure only $2d$ such edit operations are used to transform $p \mapsto q$, each move operations must increase the size of $\mathcal{L}$ monotonically.
 % while moving $j$ to $i$ is the inverse permutation.
 %The complexity of sorting by moves %is related to the well-known \emph{Levenshtein distance}~\cite{bar2004approximating}, which is the minimum number of character insertions, deletions, and substitutions needed to transform one string into another.
-%Through a dynamic programming type approach, the Wagner-Fischer algorithm can compute the Levenstein distance in $O(m^2)$, and there are several sub-quadratic approximations~\cite{bar2004approximating}.
+%. 
 %is related to the \emph{LCS} distance, which is the minimum number of character insertions or deletions needed to transform one string into another. 
@@
 %Like the \emph{LCS} distance, 
-Thus, by reducing the problem to the \emph{permutation edit distance}, we obtain the following lower bound on the minimum size of a sorting (i.e., a schedule) using moves. 
-\begin{proposition}[Schedule size]
-Let $K_0$ and $ K_1$ be filtrations of a simplicial complex $K$ with  $m$ simplices.
-Then, a smallest size  move schedule $S^*$ transforming $K_0$ into $K_1$ has cardinality 
+The Wagner-Fischer algorithm---a dynamic programming type approach---can compute the LCS distance in $O(m^2)$~\cite{bar2004approximating}.
+However, we can do better: if the pair of filtrations $K_0, K_1$ are identical up to a permutation, then computing $d$ reduces further to the \emph{permutation edit distance} problem.
+As such, we obtain the following bound on the minimum size of a sorting (i.e., a schedule) using moves and the complexity of computing it. 
+\begin{proposition}[Schedule Size]\label{prop:schedule_size}
+Let $K_0$ and $ K_1$ be filtrations of a fixed simplicial complex $K$ with $\lvert K \rvert = m$ simplices.
+Then, the smallest size any move schedule $S^*$ transforming $K_0$ into $K_1$ can have has cardinality: 
 $$ d = m - \lvert \mathrm{LCS}(K_0, K_1) \rvert $$ 
-Moreover, any size $d$ schedule $S$ can be computed in $O(m \log \log m)$ time. 
+Moreover, any $S^*$ of minimal size can be computed in $O(m \log \log m)$ time. 
 \end{proposition}
-\begin{proof}
+\begin{proof} \normalsize
 Recall our definition of  edit distance given above, 
@@
 \end{proof}
-
 \begin{corollary}\label{cor:expectation}
-	If $K_0, K_1$ are random distinct filtrations each of size $m$, then $d$ is upper bounded by $O(m - \sqrt{m})$ in expectation, with probability $1$ as $m \to \infty$.
+If $K_0, K_1$ are random distinct filtrations each of size $m$, then $d \sim m - \sqrt{m}$ in expectation, with probability $1$ as $m \to \infty$.
 \end{corollary}
-\begin{proof}
+\begin{proof} \normalsize
 	The proof of this result reduces to showing the average length of the LIS for random permutations. Let $L(p) \in [1,m]$ denote the maximal length of a increasing subsequence of $p \in S_m$. 
@@
 \end{proof}
-\noindent The upper bound from Corollary~\ref{cor:expectation} captures the size of $S^\ast$ under worst-case conditions. 
-It is worth noting that this bound applies to pairs of uniformly sampled permutations, as opposed to uniformly sampled  filtrations---simplicial filtrations have more structure due to the subcomplex relation that must be respected.
-Thus, while $O(m - \sqrt{m})$ in expectation is an improvement over $O(m)$, it is still not tight.
-However, Boissonnat~\cite{boissonnat2018efficient} has shown that for any $d$-dimensional simplicial complex $K$ with $m$ simplices, the number of distinct filtrations built from $K$ is \emph{at least} $\floor{\frac{t+1}{d+1}}^m$, where $t$ is the number of distinct filtration values associated with the complex $K$.
-In the case where $t \sim O(m)$ and $d << m$ fixed, this lower bound grows similarly to $m!$---thus, $O(m - \sqrt{m})$ is not too pessimistic of an upper bound when the pair of filtrations are no more similar than two random permutations.
-In practice, when one has a time-varying filtration and the sampling points are relatively close [in time], the LCS between adjacent filtrations is expected to be much larger, subsequently making $d$ much smaller.
+\noindent It's worth noting that the quantity from Corollary~\ref{cor:expectation} captures the size of $S^\ast$ under worst-case conditions, i.e.
+to pairs of uniformly sampled permutations, as opposed to uniformly sampled filtrations.
+Simplicial filtrations have more structure due to the subcomplex relation that must be respected, so one might wonder if this structure reduces the the size of $d$ even further.
+However, Boissonnat~\cite{boissonnat2018efficient} has shown that for any $k$-dimensional simplicial complex $K$ with $m$ simplices, the number of distinct filtrations built from $K$ is \emph{at least} $\floor{\frac{t+1}{k+1}}^m$, where $t$ is the number of distinct filtration values associated with the complex $K$.
+In the case where $t \sim O(m)$ and $k << m$ fixed, this lower bound grows similarly to $m!$---thus,  $d \approx n - \sqrt{n}$ is not too pessimistic a value to expect when the pair of filtrations are no more similar than two random permutations.
+In practice, when one has a time-varying filtration and the sampling points are relatively close [in time], the LCS between adjacent filtrations is expected to be much larger, shrinking $d$ substantially.
 For example, the average size of the LCS across the 10 filtrations from Section~\ref{sec:motivation} each having $m=417$ simplices sampled at evenly spaced time points was $343$, implying $d \approx 70$ permutations needed on average to update the decomposition at adjacent time points.
+
+Before discussing the practice of constructing such a schedule, we conclude by combining the propositions we've given so far and present the main theorem of this effort. 
+  \begin{theorem}\label{prop:sparsity_result}
+   Given a pair of simplicial filtrations $K, K'$ and a decomposition $R = DV$ of $K$, algorithm~\ref{alg:schedule} produces a sequence $\mathcal{S} = \left( s_1, s_2, \dots, s_d \right)$ of permutations $s_k = (i_k, j_k)$ with size $\lvert \mathcal{S} \rvert = m - \lvert \mathrm{LCS}(K, K') \rvert$ satisfying:
+   	\begin{equation}\label{eq:rv_seq}
+   		R = D_0 V_0 \overset{s_1}{\to} D_1 V_{1} \overset{s_2}{\to} \dots \overset{s_d}{\to} D_d V_{d} = R'
+   	\end{equation}
+	in $O(m \log \log m)$ time and $O(m)$ space. Moreover, if $i_k < j_k$ for all $k \in [d]$, then~\eqref{eq:rv_seq} can be computed using $O(d\kappa)$ column operations, where:	  
+	$$  \quad \kappa = \sum\limits_{k = 1}^d \, \lvert \mathbb{I}_{k}\rvert + \lvert \mathbb{J}_{k}\rvert $$
+	% where $\lvert \mathbb{I}_{(k)} \rvert$ and $\lvert \mathbb{J}_{(k)} \rvert$ are the computed quantities in lines (2) and (3) in algorithm~\ref{alg:mr} (here we assume if $i_\ast < j_\ast$).
+	where $\lvert \mathbb{I}_{k} \rvert$ denotes the number of non-zero entries $V_k[i_k, i_k:j_k]$ and $\lvert \mathbb{J}_{k}\rvert$ are the number of columns satisfying $\mathrm{low}_{R_k} \in [i_k:j_k]$ and $\mathrm{row}_{R_k}(i_k) \neq 0$ and $d = m - \lvert \mathrm{LCS}(K, K') \rvert$.
+\end{theorem}
+\begin{proof} \normalsize
+	Proposition~\ref{prop:schedule_size} yields the necessary conditions for constructing $\mathcal{S}$  with optimal size $d$ in $O(m \log \log m)$ time and $O(m)$.
+The definition of $\kappa$ follows directly from Algorithm~\ref{alg:mr}.
+\end{proof}
 
@@
 We outline below a simple procedure which constructs such a sorting $\mathcal{S} = (s_1, \dots, s_d)$ in $O(dm\log m)$ time.
+Equivalently, this algorithm may be performed completely online in $O(m \log m)$ per permutation.
 
@@
 We call such a sorting a valid sorting.
-To do this efficiently, we need a means of choosing a permutation $s \in S_m$ that satisfies equation~\ref{eq:valid_move}.
+To do this efficiently, we need a means of choosing a permutation $s \in S_m$ that satisfies~\eqref{eq:valid_move}.
   Let $\mathcal{T}$ denote a ordered set-like data structure that supports the following operations on elements $\sigma \in M$ from the set $M = \{0, 1, \dots, m+1\}$:
@@
 Now, suppose we choose a symbol $\sigma \notin \mathcal{T}$.
-If $p^{-1}(\sigma) < p^{-1}(\mathcal{T}_{\mathrm{pred}}(\sigma))$, then we must move $\sigma$ to the right in $p$ such that equation~\ref{eq:valid_move_ineq} holds.
+If $p^{-1}(\sigma) < p^{-1}(\mathcal{T}_{\mathrm{pred}}(\sigma))$, then we must move $\sigma$ to the right in $p$ such that~\eqref{eq:valid_move_ineq} holds.
 Similarly, if  $p^{-1}(\mathcal{T}_{\mathrm{succ}}(\sigma)) < p^{-1}(\sigma)$, then we must move $\sigma$ left in $p$ to increase the size of the LCS.
@@
 
-The problem of minimizing the right hand side of equation~\ref{eq:sorting_additive} can be interpreted as a crossing minimization problem for a set of $k$-layered bipartite graphs.
+The problem of minimizing the right hand side of~\eqref{eq:sorting_additive} can be interpreted as a crossing minimization problem for a set of $k$-layered bipartite graphs.
 To see this, consider two permutations: $p$ and $m_{ij} \circ p$, where $m_{ij}$ is a move permutation.
@@
 an \emph{implicit treap}.
-Since single element modifications, removals, and insertions can all be achieved in $O(\log m)$ expected time with such a data structure, we conclude that a greedy solution to equation~\ref{eq:greedy_step} can be computed in $O(d^2 \log m)$ time.
+Since single element modifications, removals, and insertions can all be achieved in $O(\log m)$ expected time with such a data structure, we conclude that a greedy solution to equation~\eqref{eq:greedy_step} can be computed in $O(d^2 \log m)$ time.
 %Since single element prefix sum modifications can also be achieved $O(\log m)$ time with these data structures, the time to compute $F(\hat{s}_i, \hat{s}_{i+1})$ is bounded above by $O(\log m)$, 
@@
 Ideally, one would want to continuously vary $L \in \mathbb{R} \times \mathbb{R}^{+}$ interactively, inspecting the corresponding changes to the barcodes immediately.
-To enable this, ~\cite{lesnick2015interactive} introduced an elegant reparameterization that not only fully characterizes $\mathcal{B}(M)$, but also yields an efficient method for rendering $\mathcal{B}_L(M)$ in real time.
+To enable this,~\cite{lesnick2015interactive} introduced an elegant reparameterization that not only fully characterizes $\mathcal{B}(M)$, but also yields an efficient method for rendering $\mathcal{B}_L(M)$ in real time.
 In the next section, we briefly summarize this reparameterization.
@@
 The optimal path $\Gamma^\ast$ is the one of minimal length with respect to $W$ which visits every vertex of $G$ at least once.
-There is a known $\frac{3}{2}$-approximation that can be computed efficiently which reduces the problem to the traveling salesman problem on a metric graph~\cite{christofides1976worst}.
+There is a known $\frac{3}{2}$-approximation that can be computed efficiently which reduces the problem to the traveling salesman problem on a metric graph~\cite{christofides2022worst}.
 \\
@@
 Moreover, if the filtration $K_0$ shares little similarity to the ``target'' filtration $K_1$, then the overhead of reducing the simplices from $K_1 \setminus K_0$ appended to the decomposition derived from $K_0$ may be large enough to motivate simply computing the decomposition at $K_1$ independently, especially if parallel processors are available.
-That is, our approach is primarily useful if each adjacent filtration in the ordered family of filtration is ``close.''
-Another limitation to the strategy we have here is that we rely completely on the reduction algorithm for appending new simplices to the filtration.
-If the dynamic filtration is constantly both permanently removing and appending new simplices, then the size of the LIS may very well be on the order $O(m - \sqrt{m})$, rendering the efficiencies gained using our approach as neglible.
+Indeed, in this setting the size of the LIS may very well be on the order $\sim m - \sqrt{m}  = O(m)$, rendering the efficiencies gained using our approach as negligible.
+Thus, our approach is primarily useful if filtration in the parameterized family is ``nearby'' in the combinatorial sense. 
 Note that this overhead does not exist for the multidimensional persistence use-case: the entire decomposition is needed at each template 2-cell.

### 2023-01-24 move_schedules.tex
Major edits/simplifications to the 2-parameter section. Moving things to the appendix to shorten the paper

@@
 As persistence-related computations tend to exhibit high algorithmic complexity---essentially cubic\footnote{For finite fields, it is known that the persistence computation reduces to the PLU factorization problem, which takes $O(m^\omega)$ where $\omega \approx 2.373$ is the matrix multiplication constant.} in the size of the underlying filtration~\cite{morozov2005persistence}---their adoption to dynamic settings poses a challenging computational problem.
-With the current state of the art tools, there is no recourse to the problem of computing persistence from a time-varying filtration containing millions of simplices at thousands of snapshots in time.
+% With the current state of the art tools, there is no recourse to the problem of computing persistence from a time-varying filtration containing millions of simplices across thousands of snapshots in time.
 Acquiring such a capability has far-reaching consequences: methods that vectorize persistence diagrams for machine learning purposes, like adaptive template functions~\cite{polanco2019adaptive} and persistence images~\cite{adams2017persistence}, or summary statistics like $\alpha$-smoothed Betti curves~\cite{ulmer2019topological}, all immediately become computationally viable tools in dynamic settings.
@@
 
-Cohen-Steiner, Morozov, and Edelsbrunner refer to a continuous 1-parameter family of persistence diagrams as a \emph{vineyard}, and they give in \cite{cohen2006vines} an efficient algorithm for their computation.
-The vineyards approach can be interpreted as an extension of the \emph{matrix reduction} algorithm~\cite{zomorodian2005computing}, which computes the persistence diagram of a fixed filtration $K$ with $m$ simplices in $O(m^3)$ time.
-The main step in the reduction algorithm is a specific decomposition $R = D V$ of the boundary matrix $D$ of $K$. 
-Since $V$ is always full rank and upper-triangular, an alternative is to compute a $D = RU$ decomposition, where $U = V^{-1}$. 
-The vineyards algorithm transforms a time-varying filtration into an ordered set of permutations to be applied to adjacent rows and columns of the decomposition $D = RU$, each of which takes at most $O(m)$ time to execute.
+Cohen-Steiner et al. refer to a continuous 1-parameter family of persistence diagrams as a \emph{vineyard}, and they give in \cite{cohen2006vines} an efficient algorithm for their computation. 
+The vineyards approach can be interpreted as an extension of the \emph{reduction} algorithm~\cite{zomorodian2005computing}, which computes the persistence diagram of a fixed filtration $K$ with $m$ simplices in $O(m^3)$ time via a decomposition $R = D V$ of the boundary matrix $D$ of $K$.
+%Since $V$ is always full rank and upper-triangular, an alternative is to compute a $D = RU$ decomposition, where $U = V^{-1}$. 
+In particular, the vineyards algorithm transforms a time-varying filtration into a particular set of permutations of the decomposition $R = DV$, each of which takes at most $O(m)$ time to execute.
 If one is interested in understanding how the persistent homology of a continuous function changes over time, then this algorithm is sufficient, for homological critical points can only occur when the filtration order changes.
-The vineyards algorithm is also efficient asymptotically: if there are $d$ time-points where the filtration order changes, then   vineyards   takes $O(m^3 + md)$ time; one $O(m^3)$-time reduction upfront for the filtration at time $t_0$ followed by one $O(m)$ operation to update the decomposition for the remaining time points $(t_1, t_2, \dots, t_d)$.
+The vineyards algorithm is efficient asymptotically: if there are $d$ time-points where the filtration order changes, then  vineyards   takes $O(m^3 + md)$ time; one $O(m^3)$-time reduction upfront for the filtration at time $t_0$ followed by one $O(m)$ operation to update the decomposition for the remaining time points $(t_1, t_2, \dots, t_d)$.
 When $d >> m$, this $O(md)$ approach is far more efficient than the $O(dm^3)$  ``naive'' strategy of computing the diagrams independently at every time point.
@@
 While vineyards is  designed  for handling   \emph{continuous} time families of diagrams, it is not necessarily efficient when the time parameter is coarsely discretized.
-Our methodology is based on the observation that practitioners often don't need (or want!) all $d$ persistence diagrams generated by a continuous 1-parameter of filtrations; usually just $n << d$ of them  suffice.
+Our methodology is based on the observation that practitioners often don't need (or want!) all of the persistence diagrams generated by a continuous 1-parameter of filtrations; usually just $n << d$ of them  suffice.
 % in many time varying settings, a valid decomposition is often only sought after at a relatively sparse number of time points relative to the size of the (finite) set needed to generated by the family.
@@
   Consider the situation where one is concerned with capturing these dramatic changes in homology.
-  On the left side of Figure~\ref{fig:vineyards}, we compare the cumulative cost (in total number of column operations) of various approaches.
-  Since it is unknown a priori at which times  the persistent homology will change, one solution is to discretize the time domain at $n$ evenly spaced points and compute the persistence diagrams at each one of them  independently.
+  On the left side of Figure~\ref{fig:vineyards}, we compare the cumulative cost (in total column operations) of various approaches.
+  Since it is unknown a priori at when the persistent pairing function changes, one solution is to discretize the time domain at $n$ evenly spaced points and compute the diagrams at each time point independently.
   An alternative approach is to construct a (e.g., straight-line) homotopy between  contiguous time-points, and then decompose these homotopies into adjacent transpositions in the filtration order---the vineyards approach.
@@
 \subsection{Main results}
-Given a filtered simplicial complex $K$, denote by $R = D V$ the decomposition of its corresponding boundary matrix $D$ such that $R$ is reduced and $V$ is upper-triangular (see section~\ref{sec:reduction} for details).
- If one has a pair of filtrations $K, K'$ of equal size $m = \lvert K \rvert = \lvert K' \rvert$ and $R = D V$ has been computed, then it may be advantageous to use the information stored in ($R$, $V$) to reduce the computation of $R' = D' V'$.
-Given the appropriate permutation $P$ such that $D = P^T D P$, such an update scheme has the form: 
+Given a filtered simplicial complex $K$ with filtration function $f : K \to [m]$, denote by $R = D V$ the decomposition of its corresponding boundary matrix $D$ such that $R$ is reduced and $V$ is upper-triangular (see section~\ref{sec:reduction} for details).
+ If one has a pair of filtrations $(K, f)$,  $(K', f')$ of equal size $m = \lvert K \rvert = \lvert K' \rvert$ and $R = D V$ has been computed for $K$, then it may be advantageous to use the information stored in ($R$, $V$) to reduce the computation of $R' = D' V'$.
+Given a permutation $P$ such that $D' = P^T D P$, such an update scheme has the form: 
 $$ (\ast P^T \ast R \ast P \ast) = (P^T DP)(\ast P^T \ast V \ast P \ast) $$
 where $\ast$ is substituted with elementary column operations that repair the permuted decomposition.
-It is known how to generate $d \sim O(m^2)$ decompositions requiring on the order of $c \sim O(m^2)$~\cite{cohen2006vines} such column operations.
-Since each column operation take $O(m)$, the complexity of obtaining these $d$ decompositions takes $O(m^3)$.
-This is efficient if all $d$ decompositions are needed; otherwise, updating $R \mapsto R'$ matches the complexity of computing $R = D V$ independently.
+It is known how to linearly interpolate $f \mapsto f'$ using $d \sim O(m^2)$ updates to the decomposition, where each update is comprised of at most two column operations~\cite{cohen2006vines}.
+Since each column operation takes $O(m)$, the complexity performing this update is $O(m^3)$, which is efficient if all $d$ decompositions are needed.
+Otherwise, if only $(R', V')$ is needed (and none of the $d -1$ other decomposition), updating $R \mapsto R'$ using the approach from~\cite{cohen2006vines} matches the complexity of computing $R' = D' V'$ independently.
 
@@
  where each permutation $s_i$ is a particular type of cyclic permutation (see section~\ref{sec:moves_setting}).
-If $i_\ast < j_\ast$, our first result is that this sequence can be computed in output-sensitive $O(d\kappa)$ time:
+If $i_\ast < j_\ast$, our first result is that this sequence can be computed in using $O(d\kappa)$ column operations, where:
 \begin{equation}
@@
 \begin{equation}
-	\frac{C}{2} \leq L \leq C
+	\tfrac{1}{2} \, C \leq L \leq C
 \end{equation}
@@
 
-In summary, we obtain an output-sensitive persistence update scheme which depends explicitly on the sparsity pattern of the involved matrices, has sub-cubic complexity in expectation even for random filtrations, and we show performs well in practice on a variety of applications.
-
 % given a decomposition $R = \partial_{K} V$ of a simplicial filtration $K$, algorithm~\ref{alg:schedule} can produce such a sequence $\mathcal{S} = \left( s_1, s_2, \dots, s_d \right)$
@@
 %\end{proposition}
-
-\subsection{Outline} The remainder of the paper is organized as follows: we review and establish the notations we will use to describe simplicial complexes, persistent homology, and dynamic persistence in Section \ref{sec:background}.
+\noindent 
+\textbf{Outline:} The remainder of the paper is organized as follows: we review and establish the notations we will use to describe simplicial complexes, persistent homology, and dynamic persistence in Section \ref{sec:background}.
 We also cover the reduction algorithm (designated here as \texttt{pHcol}), the vineyards algorithm, and the set of \emph{move}-related algorithms introduced in~\cite{busaryev2010tracking}, which serves as the starting point of this work.
@@
 Such a family is called a \emph{filtration},
-which is deemed to be\emph{simplexwise} if $K_j \smallsetminus K_i = \{\sigma_j\}$ 
+which is deemed \emph{simplexwise} if $K_j \smallsetminus K_i = \{\sigma_j\}$ 
 whenever $j$ is the immediate successor of $i$ in $I$.
@@
 \\
-\noindent \textbf{Example 2.1:} Non-essential filtrations often arise in geometrical contexts.
+\noindent \textbf{Example 2.1:} Non-simplexwise filtrations often arise in geometrical contexts.
 For example, given a finite metric space $(X, d)$, the \emph{Vietoris-Rips} complex at scale $\epsilon \in \mathbb{R}$ is the abstract simplicial complex given by:
@@
 The corresponding filtration is called the \emph{Vietoris-Rips} (VR) filtration, indexed by $I = \mathbb{R}$. 
-A VR filtration indexed over $\mathbb{R}$  can be condensed to an essential filtration indexed over the set of pairwise distances $\{\, d(x,x') \; \mid \; x,x' \in X \, \}$, 
-which can then be further reindexed to an essential simplexwise filtration by extending a total order of $X$ to its power set---e.g., 
-using the lexicographical ordering. 
+A VR filtration indexed over $\mathbb{R}$ can be condensed to an simplexwise filtration indexed over the set of pairwise distances $\{\, d(x,x') \; \mid \; x,x' \in X \, \}$  by extending a total order of $X$ to its power set using any linear extension of the face poset, such as e.g.
+the lexicographical ordering.
 %This is often necessary in the computational setting, where it is easier to work with a filtration with $m$ simplices indexed over $[m]$ instead of $\mathbb{R}$. 
@@
 \\
-Let $\mathbb{X}$ be a triangulable topological space. 
-That is, so that there exists an abstract 
-simplicial complex $K$ whose geometric realization  is homeomorphic to 
-$\mathbb{X}$. 
-Let  $f: \mathbb{X} \to \mathbb{R}$ be continuous and  
-write $\mathbb{X}_a = f^{-1}(-\infty, a]$ to denote the sublevel sets of $\mathbb{X}$ defined by the value $a$. 
+Let $\mathbb{X}$ be a triangulable topological space and $K$ an abstract simplicial complex whose geometric realization is homeomorphic to $\mathbb{X}$. 
+Let  $f: \mathbb{X} \to \mathbb{R}$ be continuous and  write $\mathbb{X}_a = f^{-1}(-\infty, a]$ to denote the sublevel sets of $\mathbb{X}$ defined by the value $a$. 
 A \emph{homological critical value} of $f$ is any value $a \in \mathbb{R}$ such that the homology of the sublevel sets of $f$ changes at $a$, i.e.
@@
  The main output of the reduction algorithm is a matrix decomposition $R = D V$, where the persistence diagram is encoded in $R$, and the generating cycles in the columns of $V$. 
- Given a  filtration $K$ with $m$ simplices and maximal dimension $d$, 
- one assembles the elementary boundary chains $\partial(\sigma)$ as columns ordered according to the filtration, forming the \emph{filtration boundary matrix} $D$. 
- The reduction algorithm can  compute the persistent homology for all dimensions up to $d - 1$.
- In this case,  $D$ is a square matrix of size $m \times m$.
-Alternatively, a single dimension $i \leq d - 1$ can be computed, in which case one reduces a pair of sub-matrices  $D_i$ and $ D_{i+1}$ of $D$,  where $D_i$ has dimension $d_{i-1} \times d_i$ and $D_{i+1}$ is  $d_{i} \times d_{i+1}$.
- If one is interested in a particular dimension of homology, then the latter approach is preferred.
-However, the  description of the algorithms is simpler when considering the full matrix, so $D$ will always be $(m \times m)$ in what follows.
- All algorithms discussed generalize to both situations. We give an example using the matrix pair $(R_1, V_1)$  below. 
+ Given a filtration $K$ with $m$ simplices, one assembles the elementary boundary chains $\partial(\sigma)$ as columns ordered according to the filtration, forming the square $m \times m$  \emph{filtration boundary matrix} $D$.
+ The reduction algorithm computes the persistent homology from $D$ for all dimensions up to $\mathrm{dim}(K) - 1$.
+  Alternatively, a single dimension $p \leq \mathrm{dim}(K) - 1$ can be computed, in which case one reduces a pair of sub-matrices $D_p$ and $ D_{p+1}$ of $D$,  where $D_p$ has dimension $d_{p-1} \times d_p$ and $D_{p+1}$ is $d_{p} \times d_{p+1}$.
+In some situations the description of the ensuing algorithms is simpler when considering all dimensions simultaneously, while in others it is more succinct to consider dimension-specific matrices. 
+In what follows $D$, $R$, and $V$ will always denote square $(m \times m)$ matrices, though we note that all algorithms discussed generalize to both cases. 
+ We give an example of the reduction algorithm using the matrix pair $(R_1, V_1)$ below. 
 \\
@@
 is the row index of its lowest nonzero entry.
-Using $\mathbb{Z}_2$ coefficients and homology in dimension 1 to simplify the presentation, the reduction proceeds as follows:
+Using $\mathbb{Z}_2$ coefficients and homology in dimension 1 for simplicity, the reduction proceeds as follows:
 \begin{displaymath}
@@
 \end{displaymath}
-\noindent The column $c$ in $R$ indicates that dimension $1$ homology is born, which in this filtration is never killed since the dimension of the complex is $1$.
-Similarly, the columns at $u, v, w$ in $R_0$ (not shown) are all zero, indicating three $0$-dimensional homology classes  are born, which are then killed by the pivot entries in columns $a$ and $b$ in $R_1$.
+\noindent Since column $c$ in $R_1$ is $0$, the chain $d_1 + d_2 + d_3$ indicated by the column $c$ in $V_1$ represents a dimension $1$ cycle.
+Similarly, the columns at $u, v, w$ in $R_0$ (not shown) are all zero, indicating three $0$-dimensional homology classes are born, two of which are killed by the pivot entries in columns $a$ and $b$ in $R_1$.
 \\
@@
 \noindent
-It is clear from Algorithm~\ref{alg:reduce} that a loose upper bound for this reduction is $O(m^3)$, where $m$ is the number of simplices of the filtration; it turns out that this bound is in fact tight, see~\cite{morozov2005persistence} for more details.
-There are actually many variations and optimizations proposed over the past decade to Algorithm~\ref{alg:reduce}; for example, one can   reverse the order of the filtration and compute cohomology~\cite{de2011dualities} instead.
-Despite these variations in computation, regardless of how the decomposition is obtained, it must obey the following invariants:
+Inspection of Algorithm~\ref{alg:reduce} suggests that a loose upper bound for the reduction is $O(m^3)$, where $m$ is the number of simplices of the filtration---this bound is in fact tight~\cite{morozov2005persistence}.
+Nonetheless, many variations and optimizations to Algorithm~\ref{alg:reduce} have been proposed over the past decade, see~\cite{bauer2017phat, bauer2021ripser, chen2011persistent} for an overview.
+Despite these variations in computation, the resulting decomposition must satisfy the following invariants:
   \vspace*{0.8em}
@@
  \vspace*{1em}
- If these two invariants are satisfied, then we call the decomposition \emph{valid}. Given any valid $R = D V$ decomposition, then~\cite{zomorodian2005computing}:
-%\begin{proposition}[Persistent Pairing~\cite{zomorodian2005computing}]
-%	Let $R = D V$, where $R$ is in reduced form and $V$ is full-rank upper triangular (invariants (1) and (2)  hold).
+ If these two invariants are satisfied, then we call the decomposition \emph{valid}. 
+%%\begin{proposition}[Persistent Pairing~\cite{zomorodian2005computing}]
+%%	Let $R = D V$, where $R$ is in reduced form and $V$ is full-rank upper triangular (invariants (1) and (2)  hold).
 Then the pairing $(\sigma_i, \sigma_j)$ defined by $i = \mathrm{low}_R(j)$ is a \emph{creator/destroyer (persistent) pair}.
 While the matrices $R,V$ satisfying (1) and (2) are not unique in general,  the persistent pairs $(i,j)$ are unique and define the persistence diagrams of $K$.
-%\end{proposition}
-\noindent
-The collection of persistent pairings is independent of the choice of $R$ and $V$.
-When reducing the boundary matrix, we can therefore perform column operations in any order, as long as columns are added from left to right, and so long as every column operation performed in $R$ is also performed on $V$.
-Once $R$ is reduced, then the low entries in $R$ yield the correct pairings encoding the persistence diagram of the corresponding filtration $K$.
+%%\end{proposition}
+%\noindent
+Though the matrices $R$ and $V$ are not unique, it has been shown that the collection of persistent pairings (derived from $R$) are~\cite{zomorodian2005computing}. 
+%When reducing the boundary matrix, we can therefore perform column operations in any order, as long as columns are added from left to right, and so long as every column operation performed in $R$ is also performed on $V$.
+In particular, once $R$ is in reduced form, the persistence diagram of the corresponding filtration $K$ is given by lowest non-zero ``pivot'' entries in $R$.
 % Important corollaries of the persistence pairing lemma is that the 
@@
 \subsection{Vineyards}\label{sec:vineyards}
-The original purpose of the vineyards algorithm, as described in~\cite{cohen2006vines}, was to compute a continuous 1-parameter family of persistence diagrams over a time-varying filtration, and to detect homological critical events during the construction.
-Notice that homological critical events can only occur when the filtration order changes, though the converse is not true.
-Thus, detecting  all such events reduces to computing matrix decompositions at a finite set of time points interleaved between changes in the filtration order. 
-At the finest scale, these changes manifest as transpositions of adjacent simplices. 
-As a result, a fixed set of rules for maintaining a valid $R = D V$ decomposition under adjacent transpositions  is enough to simulate persistence dynamically. 
-These rules prescribe certain column and row operations which must be applied to the matrix decomposition either before, during, or after the transposition, in order to ensure decomposition invariants I1 and I2 are respected.
-
-Let $S_{i}^j$ represent the upper-triangular matrix such that multiplying by it on the right is equivalent to adding column $i$ to column $j$. 
-Multiplication by $S_{i}^j$ on the left is equivalent to adding row $j$ to row $i$.
-% NOT SURE about this:, and thus $S_{i}^j S_{i}^j = I$. 
-Similarly, let $P$ denote the matrix so that multiplication from the right applies a fixed permutation to the columns of $A$.
+The original purpose of the vineyards algorithm, as described in~\cite{cohen2006vines}, was to compute a continuous 1-parameter family of persistence diagrams over a time-varying filtration, detecting homological critical events along the way.
+As homological critical events can only occur when the filtration order changes, detecting all such critical events can be reduced to computing valid decompositions at all time points interleaving  changes in the filtration order.
+At the finest scale, these changes manifest as transpositions of adjacent simplices for simplexwise filtrations. 
+As a result, any fixed set of rules that maintains a valid $R = D V$ decomposition under adjacent transpositions is sufficient to simulate persistence dynamically. 
+These rules prescribe certain column and row operations to apply to a given matrix decomposition either before, during, or after each individual transposition to ensure it is a valid decomposition.
+
+Let $S_{i}^j$ represent the upper-triangular matrix such that $A S_{i}^j$ is equivalent to the elementary operation that adds column $i$ of $A$ to column $j$ of $A$ and $S_{i}^j A$ is equivalent to adding row $j$ of $A$ to row $i$ of $A$.
+Similarly, let $P$ denote the matrix so that $A P$ applies a fixed permutation to the columns of $A$ and $P^T A$ applied the same permutation to the rows of $A$.
 Since the columns of $P$ are orthonormal, then  $P^{-1} = P^T$, and thus one would write $P^T A P$ to denote the application of the same permutation to both the columns and rows of $A$.
@@
 The goal of the vineyards algorithm can now be described explicitly: to prescribe a set of rules, written as matrices $S_{i}^{j}$, such that if $R = D V$ is a valid decomposition, then $(\ast P \ast R \ast P \ast) = (PDP)(\ast P\ast V \ast P \ast)$ is also a valid decomposition, where $\ast$ is some number (possibly zero) of matrices encoding elementary column or row operations.
-%We refer to the algorithm that maintains the $R = \partial V$ decomposition under adjacent permutations $P$ as the \emph{transposition framework}. 
-% The explicit pseudocode describing these rules is given in the appendix by Algorithm~\ref{alg:tr}. 
-%	\; = 
-%	\begin{blockarray}{cccc}
-%	\tiny{D_1} \kern-0.2em & a & b & c  \\
-%		\begin{block}{c[ccc]}
-%  		u \kern-0.1em & \; 1 & 1 & 0 \; \topstrut \\
-%  		v \kern-0.1em & \; 0 & 1 & 0 \; \\
-%  		w \kern-0.1em & \; 1 & 0 & 0 \; \botstrut \\
-%		\end{block}
-%	\end{blockarray}	
-%	\;\, 
-%	+
 \\
@@
  The time complexity of vineyards is determined entirely by the complexity of performing a single adjacent transposition.
- Informally, since column operations are the largest complexity operations needed and each column can have potentially $m$ entries, the complexity of vineyards is $O(m)$ per transposition. 
- However, there are a number of subtleties involved in this statement. 
- The first  is that both the $V$ and $R$ matrices tend to be quite sparse, and thus column and row operations rarely require $m$ operations---in fact, as a rule of thumb, most transpositions require no column operations~\cite{edelsbrunner2000topological}.
-A second subtlety, due to the necessity of sparse matrix data structures, is that querying the non-zero status of any particular entry typically takes $O(\log m)$ time.
-This hidden cost may seem minor, however, the quadratic scaling of the number of transpositions vineyards typically requires makes this cost non-trivial.
-In fact, achieving $O(m)$ complexity per transposition requires a special sparse matrix representation that allows swapping any two rows and columns in $O(1)$, as described in~\cite{cohen2006vines}.
-Finally, note the constant factor associated with the stated complexity: inspection of the individual cases of the algorithm from~\cite{cohen2006vines} shows that there are at most two $O(m)$ operations, on both $R$ and $V$, that may be needed for any single transposition.
+Formally, since column operations are the largest complexity operations needed and each column can have potentially $O(m)$ entries, the complexity of vineyards is $O(m)$ per transposition. 
+Inspection of the individual cases of the algorithm from~\cite{cohen2006vines} shows that any single transposition requires at most two such operations on both $R$ and $V$.
+However, there are a number of factors that affect the runtime efficiency of the algorithm as a whole. 
+On the positive side, as both the $V$ and $R$ matrices tend to be quite sparse, the cost of a given column operation is proportional the number of non-zero field coefficients in the two columns being modified.
+As a rule of thumb, most transpositions require no column operations~\cite{edelsbrunner2000topological}.
+On the negative side, one needs to constantly query the non-zero status of various entries in $R$ and $V$, which has a non-trivial runtime cost (consider evaluating e.g.
+Case 1.1 in~\cite{cohen2006vines}).
+ Moreover, an efficient implementation of the vineyards algorithm requires a non-standard sparse matrix representation that allows swapping any two rows and columns in $O(1)$, as described in Figure 5 in~\cite{cohen2006vines}.
+ Though these minor operations do not affect the asymptotic complexity of the vineyards algorithm, they collectively accrue a non-trivial runtime cost due to the quadratic scaling in how often they need to be performed.
 
@@
 
-Though they are similar, move operations confer a few computational advantages compared to vineyards. Informally, a move operation has two convenient properties:
+For completeness, we now describe the moves algorithm from~\cite{busaryev2010tracking}.
+
+Though similar, move operations confer a few computational advantages compared to vineyards. We focus on two convenient properties of moves:
 \\	
@@
 Prior to applying any permutation $P$ to the decomposition, it is necessary to remove non-zero entries in $V$ which render $P^TVP$ non-upper triangular, to maintain invariant I2.
-If one uses the (augmented) sparse matrix representation described in~\cite{cohen2006vines}, checking whether these entries are non-zero takes $O(\log m)$ per column.
-Since moving a simplex from $i$ to $j$ using vineyards performs $\lvert i - j \rvert - 1$ transpositions, this query operation is required at most $O(\lvert i - j \rvert)$ times.
-By comparison, if a row-oriented sparse matrix implementation is used, then a move operation can perform these non-zero entry checks in just one $O(\lvert i - j \rvert)$ pass prior to performing any column operations, saving a logarithmic factor.
-
-Property M2 indicates move operations only guarantee the decomposition is valid after the operation completes, implying that the decomposition is not fully maintained during the execution of \textit{RestoreRight} and \textit{RestoreLeft} below.
-This is a significant difference compared to the vineyards algorithm, which was designed specifically to maintain a valid decomposition after each transposition. 
-In this way, we interpret move operations as making a tradeoff in granularity: whereas a sequence of adjacent transpositions $(i, i{+}1), (i{+}1, i{+}2), \dots, (j{-}1, j)$ generates $\lvert i - j \rvert $ valid decompositions in vineyards, a move operation $\mathrm{Move}(i,j)$ generates only one.
-Indeed, if one has a pair of filtrations $(K_0, K_1)$ each with $m$ simplices, and we assume that each simplex $\sigma_i$ switches its relative ordering with another simplex $\sigma_j$ at most once in some ordered schedule of transpositions, then the number of intermediate persistence diagrams is bounded above by $O(m^2)$ using transpositions, whereas with move operations this bound clearly is at most $O(m)$.
+Using vineyards, one must consistently perform $\lvert i - j \rvert - 1$ non-zero status queries interleaved between repairing column operation.
+A move operation, on the other hand, groups these status queries into a single pass prior to performing any modifying operations.
+
+Property M2 implies the decomposition is not fully maintained during the execution of \textit{RestoreRight} and \textit{RestoreLeft} below, which starkly contrasts the vineyards algorithm. 
+In this way, we interpret move operations as making a tradeoff in granularity: whereas a sequence of adjacent transpositions $(i, i{+}1), (i{+}1, i{+}2), \dots, (j{-}1, j)$ generates $\lvert i - j \rvert $ valid decompositions in vineyards, an equivalent move operation $\mathrm{Move}(i,j)$ generates only one.
+Indeed, the number of intermediate persistence diagrams is bounded above by $O(m^2)$ using vineyards if each pair of simplices $\sigma_i, \sigma_j \in K$ switches relative ordering at most once in a given schedule of transpositions, whereas with move operations this bound clearly is at most $O(m)$.
 \\
@@
 \end{displaymath}
-
-\begin{algorithm}[!htb]
-	\caption{Move Right Algorithm}\label{alg:mr}
-	\setstretch{1.15}
-    \begin{algorithmic}[1]
-        \Function{RestoreRight}{$R$, $V$, $\mathbb{I} = \{I_1, I_2, \dots, I_s \}$}
-            \State $(\, d_{low}, \, d_R, \, d_V \,) \gets ( \, \mathrm{low}_R(I_1), \, \mathrm{col}_R(I_1), \, \mathrm{col}_V(I_1) \, )$
-        	\For{$k$ \textbf{in} $I_2, \dots, I_s$} 
-        		\State $( \, d_{low}', \, d_R', \, d_V' \, ) \gets ( \, \mathrm{low}_R(k), \, \mathrm{col}_R(k), \, \mathrm{col}_V(k) \, )$
-        		\State $( \, \mathrm{col}_R(k), \, \mathrm{col}_V(k) \, ) \mathrel{+}= (\, d_R, \, d_V \,) $
-        		\If {$d_{low}' < d_{low}$}
-        			\State $(\, d_{low}, \, d_R, \, d_V \,) \gets (\, d_{low}', \, d_R', \, d_V' \, )$
-        		\EndIf
-        	\EndFor 
-        	\State \Return $(\, R, \, V, \, d_R, \, d_V \, )$
-        \EndFunction
-    \end{algorithmic}
-    \begin{algorithmic}[1]
-        \Function{MoveRight}{$R$, $V$, $i$, $j$} 
-            \State $\mathbb{I} =$ columns satisfying $V[i, i:j] \neq 0$ %\Comment{$O(m)$}
-            \State $\mathbb{J} = $ columns satisfying $\mathrm{low}_R \in [i:j]$ and $\mathrm{row}_R(i) \neq 0$ %\Comment{$O(m)$}
-            \State $(\, R, \, V, d_R, d_V \,) \gets $ \Call{RestoreRight}{$R$, $V$, $\mathbb{I}$} 
-            %\Comment{Restores $p$-simplices} \Comment{$O(\lvert \mathbb{I} \rvert)$}
-            \State $(\, R, \, V \,) \gets $ \Call{RestoreRight}{$R$, $V$, $\mathbb{J}$}
-            	%\Comment{Restores $(p+1)$-simplices}
-            	%\Comment{$O(m \lvert \mathbb{I} \rvert)$}
-            \State $(\, R, \, V\, ) \gets (\, P R P^T, \, P V P^T\,)$  % \Comment{$P$ is a permutation given by (i,j)}
-            \State $(\, \mathrm{col}_R(j), \, \mathrm{col}_V(j) \, ) \gets (\, P d_R, \, P d_V \,)$
-            \State \Return $(\, R, \, V\,)$
-        \EndFunction
-    \end{algorithmic}
-\end{algorithm}
-
-\begin{algorithm}[!htb]
-	\caption{Move Left Algorithm}\label{alg:ml}
-	\setstretch{1.15}
-    \begin{algorithmic}[1]
-        \Function{RestoreLeft}{$R$, $V$, $\mathbb{K} = \{k_1, k_2, \dots, k_s \}$}
-            % \State $\mathrm{low}_K \gets \{ \mathrm{low}_R(k_1), \mathrm{low}_R(k_2), \dots, \mathrm{low}_R(k_s) \}$
-            \State $(l, r) \gets $ indices $l, r \in \mathbb{K}$ satisfying $l < r$, $\mathrm{low}_R(l) = \mathrm{low}_R(r)$ maximal 
-            \While{$\mathrm{low}_R(l) \neq 0$ \textbf{and} $\mathrm{low}_R(r) \neq 0$}
-            	\State $(\mathrm{col}_R(r), \mathrm{col}_V(r)) \mathrel{+}= (\mathrm{col}_R(l), \mathrm{col}_V(l))$
-            	\State $\mathbb{K} \gets \mathbb{K} \setminus l$
-            	\State $(l, r) \gets $ indices $l, r \in \mathbb{K}$ satisfying $l < r$, $\mathrm{low}_R(l) = \mathrm{low}_R(r)$ maximal
-            \EndWhile
-        	\State \Return $(\, R, \, V \, )$
-        \EndFunction
-    \end{algorithmic}
-
-    \begin{algorithmic}[1]
-        \Function{MoveLeft}{$R$, $V$, $i$, $j$} 
-        	\State $\mathbb{I} \gets \{ j \}$
-            % \State $(\mathbb{K}, d_R, d_V) \gets (\emptyset, \mathrm{col}_R(i), \mathrm{col}_V(i))$
-            \While{$V(k, i) \neq 0 $\textbf{ for }$ k = \mathrm{low}_V(i)$\textbf{ where }$j \leq k < i$}
-            	\State $(\mathrm{col}_R(i), \mathrm{col}_V(i)) \mathrel{+}= (\mathrm{col}_R(k), \mathrm{col}_V(k))$
-            	\State $\mathbb{I} \gets \mathbb{I} \cup k + 1 $
-            \EndWhile
-            \State $(R, V) \gets (P R P^T, P V P^T)$
-            \State $\mathbb{J} = $ columns satisfying $\mathrm{low}_R \in [i:j]$ and $\mathrm{row}_R(i) \neq 0$ %\Comment{$O(m)$}
-
-            \State $(R, V) \gets$ \Call{RestoreLeft}{$R, V, \mathbb{I}$}
-            \State $(R, V) \gets$ \Call{RestoreLeft}{$R, V, \mathbb{J}$}
-            % \State $(\mathrm{col}_R(i), \mathrm{col}_V(i)) \gets (P d_R, P d_V )$
-			\State \Return $(\, R, \, V\,)$
-        \EndFunction
-    \end{algorithmic}
-
-\end{algorithm}
 
@@
   \begin{theorem}\label{prop:sparsity_result}
-   Given a pair of simplicial filtrations $K, K'$ and a decomposition $R = DV$ of $K$, algorithm~\ref{alg:schedule} produces a sequence $\mathcal{S} = \left( s_1, s_2, \dots, s_d \right)$ of permutations $s_k = (i_k, j_k)$ with size $\lvert \mathcal{S} \rvert = m - \lvert \mathrm{LCS}(K, K') \rvert$ satisfying:
+   Given a pair of simplicial filtrations $K, K'$ and a decomposition $R = DV$ of $K$, algorithm~\ref{alg:schedule} produces a sequence $\mathcal{S} = \left( s_1, s_2, \dots, s_d \right)$ of cyclic permutations $s_k = (i_k, j_k)$ of size $\lvert \mathcal{S} \rvert = m - \lvert \mathrm{LCS}(K, K') \rvert$, satisfying:
    	\begin{equation}\label{eq:rv_seq}
@@
 \subsection*{Constructing a sorting}
- In the previous section, we showed how computing the LCS between two permutations $p, q \in S_m$ reduces to computing the LIS of a single permutation $p'$, and we outlined how to do this in $O(m \log \log m)$ time.
-It is not clear, however, that given $\mathcal{L} = \mathrm{LIS}(p')$, one can obtain a sorting $p \mapsto q$ in an efficient way.
-We outline below a simple procedure which constructs such a sorting $\mathcal{S} = (s_1, \dots, s_d)$ in $O(dm\log m)$ time.
-Equivalently, this algorithm may be performed completely online in $O(m \log m)$ per permutation.
-
- \algnewcommand{\IfThenElse}[3]{% \IfThenElse{<if>}{<then>}{<else>}
-  \State \algorithmicif\ #1\ \algorithmicthen\ #2\ \algorithmicelse\ #3}
-
-  \begin{algorithm}[h]
-	\caption{Sorting algorithm}\label{alg:sorting}
-	\setstretch{1.05}
-    \begin{algorithmic}[1]
-    	% \Require Fixed $p' \in S_m$, $\mathcal{L} = \mathrm{LIS}(p')$ 
-    	\Function{LCS-Sort}{$p$, $q$}
-    		\State $\bar{p} \gets q^{-1} \circ p$
-    		\State $\mathcal{L} \gets \mathrm{LCS}(p,q) = \mathrm{LIS}(\bar{p})$ \Comment{$O(m\log \log m)$}
-    		\State $(\, \mathcal{S}, \, \mathcal{D}, \, \mathcal{T} \, )\gets (\, \emptyset, \, \bar{p} \setminus \mathcal{L}, \, \mathcal{L} \,)$
-    		\While{$\mathcal{D}$\text{ is not empty }}
-    		    \State $\sigma \gets \text{arbitrary element in }\mathcal{D}$
-    		    %\State $\tau, \rho \gets \mathcal{T}_p(\sigma), \mathcal{T}_n(\sigma)$
-    		    %\State $i, i_p, i_n \gets \mathcal{I}_p(\sigma), \, \mathcal{I}_p(T_{\mathrm{pred}}(\sigma)), \, \mathcal{I}_p(T_{\mathrm{succ}}(\sigma))$
-    		    \State $(\,i, \,i_p, \, i_n \,) \gets ( \, \bar{p}^{-1}(\sigma), \, \bar{p}^{-1}(\mathcal{T}_{\mathrm{pred}}(\sigma)), \, \bar{p}^{-1}(\mathcal{T}_{\mathrm{succ}}(\sigma)) \, )$\Comment{$O(\log \log m)$}
-%    		    \If{ $p_{\sigma} < p_{\tau}$ }
-%	    		    \State $j \gets \text{arbitrary element in } [\, p_{\tau}, \, p_{\rho})$
-%    		    \Else{ $p_{\rho} < p_{\sigma}$}
-%    		    	\State $j \gets \text{arbitrary element in } (\, p_{\tau}, \, p_{\rho}]$
-%    		    \EndIf
-				% \IfThenElse{$i < i_p$}{$[\, i_p, \, i_n)$}{$(\, i_p, \, i_n]$}
-				% $\text{Pick } j \in [\, i_p, \, i_n)$ \If{ $i < i_p$ } \Else $(\, i_p, \, i_n]$ \EndIf
-    		    \If{ $i < i_p$ } 
-    		    	\State $j \gets \text{arbitrary element in } [\, i_p, \, i_n)$
-    		    \Else{ $i_n < i$ }
-	    		    \State $j \gets \text{arbitrary element in } (\, i_p, \, i_n]$ 
-    		    \EndIf
-    		    \State $(\, \mathcal{S}, \mathcal{D}, \mathcal{T} \,) \gets (\, \mathcal{S} \cup (i, \, j), \, \mathcal{D} \setminus \sigma, \, \mathcal{T} \cup \sigma \, )$\Comment{$O(\log \log m)$}
-    		    % \State $p, p^{-1} \gets m_{ij} \circ p, p^{-1} \circ m_{ij}^{-1}$
-    		    \State $\bar{p}^{-1} \gets \bar{p}^{-1} \circ m_{ij}^{-1}$ \Comment{$O(m)$}
-    		 \EndWhile
-    		 \State \Return{$\mathcal{S}$}
-    	\EndFunction
-	\end{algorithmic}
-\end{algorithm}
-
- Before introducing the algorithm, we require a few definitions.
+It is clear from Corollary~\ref{cor:expectation} that one may compute the LCS between two permutations $p, q \in S_m$ via the LIS of a single permutation $p^\ast$, and that computation may be carried out in $O(m \log \log m)$ time.
+It is not immediately clear, however, how to obtain a sorting $p \mapsto q$ from a given $\mathcal{L} = \mathrm{LIS}(p')$ in an efficient way.
+We outline below a simple procedure which constructs such a sorting $\mathcal{S} = (s_1, \dots, s_d)$ in $O(dm\log m)$ time and  and $O(m)$ space.
+%Alternatively, given $\mathcal{L}$, the algorithm may be executed online (up to $d$ times) in $O(m \log m)$ time and $O(m)$ space.
+
+First, we require a few definitions.
 Recall that a \emph{sorting} $\mathcal{S}$ with respect to two permutations $p, q \in S_m$ is an ordered sequence of permutations $(s_1, s_2, \dots, s_d)$ whose product under composition with $p$ yields $q$, i.e.
@@
 $q$).
-Thus, obtaining a sorting $p \mapsto q$ of size $d = m - \lvert \mathcal{L} \rvert$ reduces to applying a sequence of move operations to symbols in the complement of $\mathcal{L}$.
- Formally, we define a move permutation $s \in S_m$ as a \emph{valid} operation with respect to a fixed pair $p, q \in S_m$ if:
+Thus, obtaining a sorting $p \mapsto q$ of size $d = m - \lvert \mathcal{L} \rvert$ reduces to applying a sequence of moves to symbols in the complement of $\mathcal{L}$.
+ Formally, we define a permutation $s \in S_m$ as a \emph{valid} operation with respect to a fixed pair $p, q \in S_m$ if:
  \begin{equation}\label{eq:valid_move}
@@
  \end{equation}
-  The problem of constructing a sorting $\mathcal{S}$ of size $d$ thus reduces to the problem of choosing a sequence of $d$ valid moves.
-We call such a sorting a valid sorting.
-To do this efficiently, we need a means of choosing a permutation $s \in S_m$ that satisfies~\eqref{eq:valid_move}.
-  Let $\mathcal{T}$ denote a ordered set-like data structure that supports the following operations on elements $\sigma \in M$ from the set $M = \{0, 1, \dots, m+1\}$:
+  The problem of constructing a sorting $\mathcal{S}$ of size $d$ thus reduces to the problem of choosing a sequence of $d$ valid moves, which we call a \emph{valid sorting}.
+To do this efficiently, let $\mathcal{T}$ denote a ordered set-like data structure that supports the following operations on elements $\sigma \in M$ from the set $M = \{0, 1, \dots, m+1\}$:
   \begin{enumerate}
@@
   \end{enumerate}
-  Given such a structure $\mathcal{T}$, an arbitrary valid sorting can be constructed by repeatedly querying and updating information about the $\mathrm{LCS}$ in $\mathcal{T}$.
+  Given such a $\mathcal{T}$, an arbitrary valid sorting can be constructed by repeatedly querying and maintaining information about the $\mathrm{LCS}$ in $\mathcal{T}$.
 To see this, suppose $\mathcal{T}$ contains all of the symbols in the current $\mathrm{LCS}$ between two permutations $p$ and $q$.
@@
 Similarly, if  $p^{-1}(\mathcal{T}_{\mathrm{succ}}(\sigma)) < p^{-1}(\sigma)$, then we must move $\sigma$ left in $p$ to increase the size of the LCS.
-Assuming the structure $\mathcal{T}$ supports all of the above operations in $O(\log m)$ time, we easily deduce a $O(d m\log m)$ algorithm for obtaining a valid sorting, which is shown in pseudocode in Algorithm~\ref{alg:sorting}.
+Assuming the structure $\mathcal{T}$ supports all of the above operations in $O(\log m)$ time, we easily deduce a $O(d m\log m)$ algorithm for obtaining a valid sorting, which for completeness is shown via Algorithm~\ref{alg:sorting} in the appendix.
   % one can incrementally build a valid sorting of size $d$ by arbitrarily choosing valid move operations . all elements in $\mathrm{LCS}(p,q)$ into $\mathcal{T}$
@@
 \subsection{Minimizing schedule cost}\label{sec:schedule_cost}
-\subsubsection{Motivation}
 The results from the previous section yield a sufficient algorithm for generating move schedules of minimal cardinality: any schedule of moves $S$ sorting $K_0 \mapsto K_1$ above is guaranteed to have size $\lvert S \rvert = m - \lvert \mathrm{LCS}(K_0, K_1) \rvert$, and the reduction to the permutation edit distance problem ensures this size is optimal.
@@
 Such a generalization has appeared naturally in many application contexts, showing potential as a tool for exploratory data analysis~\cite{lesnick2012multidimensional}.
-Indeed, one of the often noted drawbacks of persistence is its instability with respect to strong outliers. 
-Although persistence diagrams are stable with respect to their input---in the Hausdorff sense---and thus are robust to Hausdorff noise, the presence of strong outliers can artificially modify the connectivity of the underlying filtration, obscuring the detection of what would otherwise be significant topological structures~\cite{buchet2015topological}.
-%Indeed, since inferring the homology of some underlying topological space around which data are distributed was one of the original purposes of persistence\footnote{One of the first uses of persistence was to infer the topology of attractors in dynamical systems from experimental data\cite{perea2018brief}.}, strong outliers pose a significant barrier to the widespread adoption of persistent homology.
-By filtering the data according to a notion of density, in tandem with distance predicates, multidimensional persistence can enable the practitioner to handle outliers directly.
-%, without resorting to sophisticated data preprocessing techniques which may or may not preserve the underlying topology.
-
-Unfortunately, multidimensional persistent homology comes with its own set of computational and theoretical challenges. 
-%not necessarily inherited by the 1-D case. 
-There is not complete discrete invariant for multidimensional persistence---like the 1D persistence diagram---and the computations associated to other incomplete invariants exhibit high algorithmic complexity.
-A number of computational strategies have been proposed to counter this, such as homology preserving multifiltration simplification or parallelization schemes using shared-memory computation~\cite{fugacci2019chunk}.
-Rather than attempting to tackle multiple dimensions simultaneously, another approach is to focus on a fixed dimension, such as 2-D persistence.
-Utilizing the equivalence between the rank and fibered barcode invariants, Lesnick and Wright~\cite{lesnick2015interactive} developed an elegant way of visualizing the former via a reparameterization of the latter using standard point-line duality.
-This clever reparameterization effectively reduces the fibered barcode computation to a sequence of 1-D barcode computations at ``template points'' lying within the 2-cells of particular planar subdivision of the half-plane $[0, \infty) \times \mathbb{R}$.
-Indeed, since algorithm~\ref{alg:schedule} was designed for precisely such a computation, the algorithm proposed by~\cite{lesnick2015interactive} is prototypical of the class of methods that stand to benefit from move scheduling.
-
-The purpose of the following sections is to uncover the intrinsic bottlenecks impeding the computation of the fibered barcode invariant in practical settings.
-Towards this goal, we briefly summarize the theory established by several researchers~\cite{lesnick2015interactive, carlsson2010computing, landi2014rank}, focusing on their computational and algorithmic details.
-Our goal is to show that the highest complexity subcomputation of the fibered barcode invariant algorithm proposed by~\cite{lesnick2015interactive} is precisely the type of computation we focus on in this effort.
-In doing so, we substantiate our claims from section~\ref{sec:motivation} that our scheduling approach can improve the efficiency of a real-world application of multidimensional persistence by an order of magnitude by comparing Algorithm~\ref{alg:schedule} to several alternatives on a real-world data set and problem: the problem of identifying the underlying topological space of a data set extracting from high-contrast natural images.
-% In what follows, we briefly recount an incomplete but nonetheless very useful invariant of multi-dimensional persistence modules, the \emph{rank invariant}, as well as another invariant which in some sense determines the rank invariant, the \emph{fibered barcode}.
-It turns out that computing the latter invariant
-
-% --- For each section of the paper, consider writing a mini-introduction that says what its organization is, what is in each subpart, and how the parts relate to one another.
-
-%To demonstrate how our scheduling approach can accelerate the multidimensional persistence computation, we establish the computational framework we will be considering. 
-% These transforms end up being crucial in decomposing the definition of the fibered barcode (and thus, the rank invariant) into a discrete computation.
-%We begin with a few definitions.
-An $n-D$ \emph{filtration} is a functor $\mathcal{F} : \mathbb{R}^n \to \textbf{Simp}$, the category of simplicial complexes, such that for all $a \leq b$, $\mathcal{F}(a,b) : F_a \to F_b$ is an inclusion.
-We will focus exclusively on the case where $n = 2$, where the object of study is a \emph{bifiltration}.
-Multifiltrations can be broadly categorized into two types: 1-critical and multi-critical filtrations.
-A \emph{1-critical filtration} $\mathcal{F}$ is a finite $n-D$ filtration each simplex $s \in K$ is associated with only one multi-grade (denoting its appearance in each of the filtered dimensions), otherwise $\mathcal{F}$ is said to \emph{multi-critical}.
-%An $n-D$ persistence module is a functor $M : \mathbb{R}^n \to \textbf{Vect}$, where $\textbf{Vect}$ denotes the category of $\mathbf{F}$-vector spaces and linear maps for some suitable field $\mathbb{F}$.
-A module $M$ is said to be \emph{pointwise finite dimensional} (p.f.d) if $\mathrm{dim}(M_a) < \infty$ for all $a \in \mathbb{R}^n$.
-%For $i \geq 0$, let $H_i : \textbf{Simp} \to \textbf{Vect}$ denote the $i^{\text{th}}$ simplicial homology functor with coefficients in $\mathbb{F}$. 
-%Using this notation, given a 1D filtration $\mathcal{F}$ of size $m$, each p.f.d 1D persistence module $M = H_i(\mathcal{F})$ for  may be associated with a multiset $\mathcal{B}(M)$ of intervals in $\mathbb{R}$ which record the isomorphism classes of indecomposable summands of $M$.
-The set $\mathcal{B}(M)$ is called the \emph{barcode}\footnote{Note that \emph{barcodes} and \emph{persistence diagrams} are interchangeable representations of the same multiset.} of $M$.
-%\emph{The rank invariant} proposed by Carlsson et al~\cite{carlsson2009theory} is one such invariant that is complete when $n=1$ but is otherwise incomplete when $n \geq 2$. 
-%% The rank invariant is computable in polynomial time~\cite{carlsson2010computing}, that has been the subject of much interest lately. 
-%It is defined as follows: 
-%\begin{definition}[Rank Invariant]
-%	For $n \geq 1$, the \emph{rank invariant} of a $\mathbb{R}^n$-indexed persistence module $M$ over $\mathcal{H}^{n} \subseteq \mathbb{R}^n \times \mathbb{R}^n$ is the function:
-%	\begin{equation*}
-%		\begin{aligned}
-%		& \mathrm{rank}(M)  \; : & \mathcal{H}^n & \to \mathbb{N}  \\
-%		& & (a,b) & \mapsto \mathrm{rank} \, M(a,b) 
-%		\end{aligned}
-%	\end{equation*}
-%	for $a \leq b \in \mathcal{H}^n$. When $n = 1$, the rank invariant is complete, thus $\mathrm{rank}(M)$ and the persistence barcodes $\mathcal{B}(M)$ determine each other.
-%\end{definition}
-%\noindent Thus, in one dimension, persistent homology provides a complete invariant that fully captures the lifetime of topological features in a given filtration: 1-D persistent homology modules decompose in an essentially unique way into indecomposable summands~\cite{zomorodian2005computing}.
-In contrast, the theory of multidimensional persistence shows that no complete discrete invariant exists---the structure of the target for the invariant depends on the structure of the underlying field, and although multiparameter persistence modules also decompose in a similar way as the 1-D case, the set of isomorphism classes is known to be extremely complicated~\cite{carlsson2009theory, lesnick2015interactive}.
-%Nonetheless, there are invariants related to multi-parameter persistence that can be quite useful in practice, even though they are incomplete. 
-%\noindent For example, even though the rank invariant does not encode the isomorphism type of the underlying module $M$ when $n > 1$, it does captures important ``first-order'' information about the modules structure.
- %The multigraded Betti numbers constitute a natural and important class of invariants often studied in algebraic geometry commutative algebra.
-There are many ways of defining these, with perhaps the simplest definition stemming from minimal resolutions: for a finitely generated $n$-parameter persistence module $M$ and $i \in \{0, 1, \dots, d \}$, the $i^{\text{th}}$ graded Betti number of $M$ at grade $z$, denoted as $\beta_i^M(z)$, is defined as the number of elements at grade $z$ in a basis of the $i^{\text{th}}$ module in a free resolution for $M$.
-
-\subsubsection*{The Fibered Barcode}
-%In the context of persistence, and \emph{invariant} is a function from the collection of persistence modules to some set $S$ with the property that $f(M) = f(M')$ when $M \cong M'$.
-An invariant is said to be \emph{complete} if $M \cong M'$ whenever $f(M) = f(M')$.
-Thus, incomplete invariants are invariants which may have the same value on two non-isomorphic modules.
-Lesnick and Wright~\cite{lesnick2015interactive} associate three simple invariants to a 2-dimensional persistence module $M$: the \emph{dimension function}, the \emph{multigraded Betti numbers}, and the \emph{fibered barcode}.
-The dimension function is simply the the function which maps every point $a \in \mathbb{R}^2$ to $\mathrm{dim}(M_a)$.
-It is a simple and easy to visualize invariant, but is unstable and yields no information about the persistent features of $M$.
-The multigraded Betti numbers constitute a natural and important class of invariants often studied in algebraic geometry and commutative algebra, though both their theory and computation is beyond the scope of this effort---the interested reader is referred to~\cite{lesnick2015interactive, carlsson2009theory} for more details.
-For our purposes, the $i^{\text{th}}$-graded Betti number of $M$ is simply a function $\beta_i(M): \mathbb{R}^2 \to \mathbb{N}$ whose values indicate the the number of elements at  each degree in a basis of the $i^{\text{th}}$ module in a free resolution for $M$.
-The fibered barcode, denoted as $\mathcal{B}(M)$, is defined as follows: 
+
+Unfortunately, unlike one-parameter persistence, there is no complete discrete invariant for multi-parameter persistence.
+Instead, Lesnick et al~\cite{lesnick2015interactive} associate a variety of incomplete invariants to 2-parameter persistence modules: we focus here on the \emph{fibered barcode} invariant, defined as follows:
 \begin{definition}[Fibered barcode]
@@
 $$ \mathcal{B}(M) = \{ \; B_L(M) : L \in \mathbb{R} \times \mathbb{R}^{+} \; \}$$
+Equivalently, $\mathcal{B}(M)$ is the 2-parameter family of barcodes given by restricting $M$ to the of set affine lines with non-negative slope in $\mathbb{R}^2$. 
 \end{definition} 
-\noindent Thus, given a bipersistence module $M$, the fibered barcode of $M$ is the 2-parameter family of barcodes given by restricting $M$ to the of set affine lines with non-negative slope in $\mathbb{R}^2$.
-
-Although the fiber barcode is a very intuitive invariant, it is not clear how one might go about computing it efficiently. 
-One could compute the 1-D persistence associated to some fixed linear combination of two given filter functions; each linear combination parameterizes a line $L \in \mathbb{R} \times \mathbb{R}^{+}$ which $M$ restricts too.
+\noindent Although an intuitive invariant, it is not clear how one might go about computing $\mathcal{B}(M)$ efficiently. 
+One obvious choice is fix $L$ via a linear combination of two filter functions, restrict $M$ to $L$, and compute the associated 1-parameter barcode. 
 However, this is an $O(m^3)$ time computation, which is prohibitive for exploratory data analysis purposes.
-Ideally, one would want to continuously vary $L \in \mathbb{R} \times \mathbb{R}^{+}$ interactively, inspecting the corresponding changes to the barcodes immediately.
-To enable this,~\cite{lesnick2015interactive} introduced an elegant reparameterization that not only fully characterizes $\mathcal{B}(M)$, but also yields an efficient method for rendering $\mathcal{B}_L(M)$ in real time.
-In the next section, we briefly summarize this reparameterization.
-
-%The fibered barcode enjoys a number of stability properties readily expressed using interleaving distances~\cite{landi2014rank}. 
-% More important to this work, the fibered barcode $\mathcal{B}(M)$ and the rank invariant $\mathrm{rank}(M)$ determine each other~\cite{lesnick2015interactive}.
-
-\subsubsection*{Reparameterization using point-line duality}\label{sec:fibered_barcode_reparam}
-Let $\overline{\mathcal{L}}$ denote the collection of all lines in $\mathbb{R}^2$ with non-negative slope, $\mathcal{L} \subset \overline{\mathcal{L}}$ the collection of all lines in $\mathbb{R}^2$ with non-negative, finite slope, and $\mathcal{L}^\circ$ the collection of all affine lines in $\mathbb{R}^2$ with positive, finite slope.
-There is a standard point-line duality that gives a parameterization of $\mathcal{L}$ with the half-plane $[0, \infty) \times \mathbb{R}$ that is convenient to work with here.
-Define the \emph{line} and \emph{point} dual transforms $\mathcal{D}_{\ell}$ and $\mathcal{D}_p$, respectively, as follows:
-\begin{equation}
-	\begin{aligned}
-\mathcal{D}_{\ell}: \mathcal{L} \rightarrow[0, \infty) \times \mathbb{R} & \quad \hfill \quad & \mathcal{D}_{p}:[0, \infty) \times \mathbb{R} \rightarrow \mathcal{L} \\
-y=a x+b \mapsto(a,-b) &\quad \hfill \quad & (c, d) \mapsto y=c x-d
-\end{aligned}
-\end{equation}
-% Use \multicolumn{1}{|r|}{text} to align
-The transforms $\mathcal{D}_{\ell}$ and $\mathcal{D}_p$ are \emph{dual} to each other in the sense that for any point $a \in [0, \infty) \times \mathbb{R}$ and any line $L \in \mathcal{L}$, $a \in L$ if and only if $D_\ell(L) \in D_{p}(a)$.
-Now, for some fixed line $L$, define  the \emph{push map} $\mathrm{push}_L(a):  \mathbb{R}^2 \to L \cup \infty$ as:
-\begin{equation}
-	\mathrm{push}_L(a) \mapsto \mathrm{min}\{ v \in L \mid a \leq v \}
-\end{equation}
-%\begin{equation*}
-%	\begin{aligned}
-%	& \mathrm{push}_L : & \mathbb{R}^2 & \to L \cup \infty \\
-%	& & a \in \mathbb{R}^2 & \mapsto \mathrm{min}\{ v \in L \mid a \leq v \}
-%	\end{aligned}
-%\end{equation*}
-The push map satisfies a number of useful properties. Namely: 
-\begin{enumerate}
-	\item For $r < s \in \mathbb{R}^2$, $\mathrm{push}_L(r) \leq \mathrm{push}_L(s)$
-	\item For each $a \in \mathbb{R}^2$, $\mathrm{push}_L(a)$ is continuous on $\mathcal{L}^\circ$
-	\item For $L \in \mathcal{L}^\circ$ and $S \subset \mathbb{R}^2$, $\mathrm{push}_L$ induces a totally ordered partition $S_L$ on $S$ 
-\end{enumerate}
-Property (1) asserts that for any $L \in \overline{\mathcal{L}}$, the standard partial order on $\mathbb{R}^2$ restricts to a total order on $L$.
-Properties (2) and (3) qualify the following definition:
-\begin{definition}[Critical Lines]
-	For some fixed $S \subset \mathbb{R}^2$, a line $L \in L^\circ$ is defined to be \emph{regular} if there is an open ball $B \in L^\circ$ containing $L$ such that $S_L = S_{L'}$ for all $L' \in B$.
-Otherwise, the line $L$ is defined as \emph{critical}.
-\end{definition}
-\noindent The set of critical lines $\mathrm{crit}(M)$ with respect to some fixed set $S \subset \mathbb{R}^2$ fully characterizes a certain planar subdivision of the half plane $[0, \infty) \times \mathbb{R}$.
-This planar subdivision, denoted by $\mathcal{A}(M)$, is thus entirely determined by $S$ under point line duality.
-%It has been shown~\cite{lesnick2015interactive} that the 1-skeleton $\mathcal{A}^1(M)$ of the line arrangement $\mathcal{A}(M)$ can be characterized b
-%Explicitly, define $\mathcal{A}^1(M)$ to the 1-skeleton of the line arrangement defined by: 
-% \{ \mathcal{D}_p(\alpha) \text{ for } \alpha = \mathrm{LUB}(u,v), u,v \in S \} 
-%\begin{equation}\label{eq:arrangement_crit}
-%	\mathcal{A}^1(M) = \{ \mathcal{D}_{\ell}(\mathrm{crit}(M)) \cup (\{0\} \times \mathbb{R}) \}
-%\end{equation}
-%Expanding the 1-skeleton to a 2-D cell complex $\mathcal{A}(M)$ fully partitions the upper half plane.  
-A corollary from~\cite{lesnick2015interactive} shows that if the duals of two lines $L, L' \in \mathcal{L}$ are contained in the same $2$-cell in $\mathcal{A}(M)$, then $S_L = S_{L'}$, i.e.
-the partitions induced by $\mathrm{push}_L$ are equivalent.
-Indeed, the total order on $S_L$ is simply the pullback of the total order on $L$ with respect to the push map.
-Since $\mathcal{A}(M)$ partitions the entire half-plane, the dual to every line $L \in \mathcal{L}$ is contained within $\mathcal{A}(M)$---the desired reparameterization.
-
-Let $S = \mathrm{supp}\,\beta_0(M) \, \cup \, \mathrm{supp}\,\beta_1(M)$, where the functions $\beta_0(M), \beta_1(M)$ are $0^{\text{th}}$ and $1^{\text{st}}$ bigraded Betti numbers of $M$, respectively.
-The main mathematical result from~\cite{lesnick2015interactive} is a characterization of the barcodes $\mathcal{B}_L(M)$, for any $L \in \mathcal{L}$, in terms of a set of \emph{barcode templates} $\mathcal{T}$ computed at every 2-cell in $\mathcal{A}(M)$.
- More formally, for any line $L \in \overline{\mathcal{L}}$ and $e$ any 2-cell in $\mathcal{A}(M)$ whose closure contains the dual of $L$ under point-line duality, the 1-parameter restriction of the persistence module $M$ induced by $L$ is given by:
- \begin{equation}
- \mathcal{B}_L(M) = \{ [\,\mathrm{push}_L(a), \, \mathrm{push}_L(b)\,) \mid (a,b) \in \mathcal{T}^e, \mathrm{push}_L(a) < \mathrm{push}_L(b)  \} 
-\end{equation}
-Minor additional conditions are needed for handling completely horizontal and vertical lines. 
-The importance of this theorem lies in the fact that the fibered barcodes are completely defined from the precomputed barcode templates $\mathcal{T}$---once every barcode template $\mathcal{T}^e$ has been computed and augmented onto $\mathcal{A}(M)$, $\mathcal{B}(M)$ is completely characterized, and the barcodes $\mathcal{B}_L(M)$ associated to a 1-D filtration induced by \emph{any} choice of $L$ can be efficiently computed via a point-location query on $\mathcal{A}(M)$ and a $O(\lvert \mathcal{B}_L(M) \rvert)$ application of the push map.
-
-%For some fixed set $S \subset \mathbb{R}^2$ and any pair $u, v \in S$ of points which are incomparable with respect to the partial order on $\mathbb{R}^2$
-% The definition of \emph{critical} lines provides an analogue to the
-
-%This is the main mathematical result from~\cite{lesnick2015interactive},
-
-% Indeed, it was shown in~\cite{lesnick2015interactive} that the fibered barcodes $$ of $M$ (i.e.
-the collection of barcodes of 1D affine slices of $M$) and the rank invariant determine each other.
-\subsubsection*{Computing the fibered barcode}
-%Lesnick and Wright introduced the Rank Invariant Visualization and Exploration Tool (RIVET)~\cite{lesnick2015interactive}, which is both a tool and a collection of computational and mathematical theory which enables exploratory analysis of 2D persistence modules built from arbitrary bifiltrations.
-From a computational perspective, the core novelty of the RIVET approach is its ability to compute the fibered barcodes efficiently from precomputed information.
-Given bipersistence module $M$, the three dominant computations needed to compute the fibered barcode using the reparameteratization discussed above are the following: 
-\begin{enumerate}
-	\item Computing the bigraded Betti numbers $\beta(M)$ of $M$
-	\item Constructing a line arrangement $\mathcal{A}(M)$ induced by critical lines from (1) 
-	\item Augmenting $\mathcal{A}(M)$ with a \emph{barcode template} $\mathcal{T}^e$ at every 2-cell $e \in \mathcal{A}(M)$
-	%, yielding the \emph{augmented arrangement} $\mathcal{A}^{\sbullet}(M)$ 
-\end{enumerate}
-Computing (1) takes approximately $\approx O(m^3)$ using a matrix algorithm similar to Algorithm~\ref{alg:reduce}~\cite{lesnick2019computing}.
-Constructing and storing the line arrangement $\mathcal{A}(M)$ with $n$ lines and $k$ vertices is related to the \emph{line segment intersection problem}, which known algorithms in computational geometry can solve in (optimal) output-sensitive $O((n+k) \log n)$ time (see~\cite{boissonnat2000robust} for an overview).
-In terms of memory complexity, the number of 2-cells in $\mathcal{A}(M)$ is upper bounded by $O(\kappa^2)$, where $\kappa$ is a coarseness parameter associated with the computation of $\beta(M)$.
-By combining this bound with a bound on the number of transpositions needed to traverse a particular sequence of 2-cells in $\mathcal{A}(M)$, the analysis from~\cite{lesnick2015interactive} shows that the computation of the barcodes templates $\mathcal{T}$ from step (3) using vineyard updates requires on the order of $O(m^3 \kappa + m \kappa^2 \log \kappa)$ elementary operations and $O(m \kappa^2)$ storage.
-
-Of the three steps above, the barcode template computation is both the highest complexity and most demanding computation in practice.
-The number of 2-cells in $\mathcal{A}(M)$ is on the order $O(\kappa^2)$, and $\kappa$ itself is on the order of $O(m^2)$ in the worst case.
-This suggests the computation of the barcode templates is on the order $O(m^5)$, which is effectively intractable for even moderately-sized filtrations.
-In practice, the external stability result from~\cite{landi2014rank} justifies the use of a simple procedure which approximates the module $M$ with a smaller module $M'$ using a grid-like coarsening procedure.
-This coarsening procedure allows the practitioner to restrict the size of $\kappa$ to a relatively small constant, which in-turn dramatically reduces the size of $\mathcal{A}(M)$.
-
-There are several approaches one can use to compute $\mathcal{T}$, the simplest being to run Algorithm~\ref{alg:reduce} independently on the 1-D filtration induced by the duals of some set of points (e.g.
-the barycenters) lying in the interior of the 2-cells of $\mathcal{A}(M)$.
-The approach taken by~\cite{lesnick2015interactive} is to use the $R = DV$ decomposition computed at some adjacent 2-cell $e \in \mathcal{A}(M)$ to speed up the computation of an adjacent cell $e' \in \mathcal{A}(M)$.
-More explicitly, define the \emph{dual graph} of $\mathcal{A}(M)$ to be the undirected graph $G$ which has a vertex for every 2-cell $e \in \mathcal{A}(M)$ and an edge for each adjacent pair of cells $e, e' \in \mathcal{A}(M)$.
-Each vertex in $G$ is associated with a barcode template $\mathcal{T}^e$, and the computation of $\mathcal{T}$ now reduces to computing a path $\Gamma$ on $G$ which visits each vertex at least once.
-To minimize the computation time, assume the $n$ edges of $G$ are endowed with non-negative weights $W = w_1, w_2, \dots, w_n$ whose values $w_i \in \mathbb{R}_{+}$ represent some notion of distance.
-The optimal path $\Gamma^\ast$ is the one of minimal length with respect to $W$ which visits every vertex of $G$ at least once.
-There is a known $\frac{3}{2}$-approximation that can be computed efficiently which reduces the problem to the traveling salesman problem on a metric graph~\cite{christofides2022worst}.
+
+%Lesnick and Wright~\cite{lesnick2015interactive} introduced an elegant reparameterization that not only fully characterizes $\mathcal{B}(M)$, but also yields an efficient method for rendering $\mathcal{B}_L(M)$ in real time.
+Utilizing the equivalence between the rank and fibered barcode invariants, Lesnick and Wright~\cite{lesnick2015interactive} developed an elegant way of computing $\mathcal{B}(M)$ via a reparameterization using standard point-line duality.
+This clever technique effectively reduces the fibered barcode computation to a sequence of 1-D barcode computations at ``template points'' lying within the 2-cells of particular planar subdivision $\mathcal{A}(M)$ of the half-plane $[0, \infty) \times \mathbb{R}$.
+This subdivision may be obtained via arrangement of ``critical lines'' induced by the bigraded Betti numbers $\beta(M)$ of $M$.
+As the barcode of one template point $\mathcal{T}_e$ at the two cell $e \in \mathcal{A}(M)$ may be computed efficiently by re-using information from an 2-cell-adjacent template point $\mathcal{T}_{e'}$,~\cite{lesnick2015interactive} observed that computing the barcodes of all such template points (and thus, $\mathcal{B}(M)$) may be reduced to ordering the two-cells in $\mathcal{A}(M)$ along a Eulerian path traversing the dual graph of $\mathcal{A}(M)$.
+The full algorithm is out of scope for this effort; we include supplementary details for the curious reader in the appendix~\ref{app:2d_pers}. 
 \\
 \\
-\textbf{Noisy circle example} We include a small example to illustrate many of the concepts and definitions discussed so far. 
-Consider a small set of points distributed noisily around $S^1$ that contains a few strong outliers, shown on the left side of Figure~\ref{fig:bifiltration_ex2}.
-Filtering this data set with respect to the Rips parameter and the complement of a kernel density estimate yields a bifiltration whose dimension function and bigraded Betti numbers are shown in the middle figure.
-The gray areas indicate the region where the dimension function is constant; the lighter gray area where $\mathrm{dim}_1(M) = 1$, indicating where a persistent loop was detected.
-On the right side, the space dual to the space in the middle is shown.
-The black lines represent the lines of $\mathcal{A}(M)$, the blue dashed-lines the edges of the dual graph $\mathcal{G}$, and the orange points the barycenters of the 2-cells in $\mathcal{A}(M)$, depicting where the barcodes templates $\mathcal{T}^e$ are parameterized.
+\noindent 
+\textbf{Example}: Consider a small set of noisy points distributed around $S^1$ containing a few strong outliers, as shown on the left side of Figure~\ref{fig:bifiltration_ex2}.
+Filtering this data set with respect to the Rips parameter and the complement of a kernel density estimate yields a bifiltration whose various invariants are shown in the middle figure.
+The gray areas indicate the region where the dimension function is constant; the lighter gray area where $\mathrm{dim}_1(M) = 1$ indicates a persistent loop was detected.
+On the right side, dual space is shown: the black lines are the critical lines that form $\mathcal{A}(M)$, the blue dashed-lines the edges of the dual graph of $\mathcal{A}(M)$, and the orange barycentric points along the 2-cells of $\mathcal{A}(M)$ represent where the barcodes templates $\mathcal{T}_e$ are parameterized.
 \begin{figure}[t]
@@
 On left, the input data, colored by density.
-In the middle, the bigraded Betti numbers $\beta_0(M)$ and $\beta_1(M)$ (green and red, respectively), the Hilbert function (gray), and a line $L$ emphasizing the persistence of features with high density.
-On the right, the line arrangement $\mathcal{A}(M)$ lying in the dual space $D_p(\alpha)$ derived from the $\beta(M)$.
+In the middle, the bigraded Betti numbers $\beta_0(M)$ and $\beta_1(M)$ (green and red, respectively), the dimension function (gray), and a line $L$ emphasizing the persistence of features with high density.
+On the right, the line arrangement $\mathcal{A}(M)$ lying in the dual space derived from the $\beta(M)$.
 }
@@
 \end{figure}
-The shortest path $\Gamma^\ast$ spanning the vertices of $G$ is shown by the solid rainbow colored lines, with the traversal beginning in the top-left cell (at the red edge) and ending at the bottom-right cell (the indigo edge).
+%The shortest path $\Gamma^\ast$ spanning the vertices of $G$ is shown by the solid rainbow colored lines, with the traversal beginning in the top-left cell (at the red edge) and ending at the bottom-right cell (the indigo edge).
 Traversing $\Gamma^\ast$ can be done via a simple depth-first search, though note that a $R=DV$ decomposition needs to be stored at each degree-3 or higher vertex in $\Gamma^\ast$ to avoid redundant computations as the traversal recurses.
@@
 
-%To an along an Eulerian path spanning the dual graph $G$ of the arrangement $\mathcal{A}(M)$. The
-%As input, one has a set of points $A \subset \mathbb{R}^2$, whose duals induce filtrations ww
-%Since each template point corresponds to a dual 
-%TODO: (Discuss path along dual graph of $\mathcal{A}(M)$)
-% It is worth briefly discussing just how difficult computationally step (3) is.
-For a bifiltration with $m$ simplices, \emph{without any coarsening}, $\kappa$ is on the order of $O(m^2)$ in the worst case.
-Thus, the uncoarsened complexity of the template computation barcode is $\approx O(m^5)$, where again $m$ is the number of simplices in the bifiltration.
-A more common practice in data science is to report the complexity of algorithm in terms of the number of data points.
-Given some input point cloud data (PCD) with $n$ points, the standard rips filtration up to dimension 2 can have upwards of $O(n^3)$ 2-simplices.
-Thus, given a 2-dimensional bifiltration $\mathcal{F}$ with $m = n^3$ simplices derived from a rips filtration on PCD with $n$ data points, the complexity of computing all barcode templates for every $2$-cell in $\mathcal{A}(M)$ for $M = H_1(\mathcal{F})$ is on the order of $O((n^3)^5) \approx O(n^{15})$.
-This is clearly computationally intractable for even relatively small data sets.
-
-%There are a few ways of tackling this computational barrier.
-Recent work has focused on using alternative filtration whose size is much smaller than that of the rips (or cech) filtration, but whose homology is approximately the same, for example by using a sparsified rips filtration~\cite{} or the chunk-based approach to multi-parameter persistence~\cite{fugacci2019chunk}.
-It was shown in~\cite{} that one can alternatively opt to approximate a bipersistence module $M$ by a \emph{coarsened} module $M'$ in a way that bounds the the interleaving distance $d_I(M, M')$ between the two modules.
-The external stability result of Landi~\cite{} implies that if the two modules $M$ and $M'$ are close in their interleaving distance, then the fibered barcodes $\mathcal{B}(M)$ and $\mathcal{B}(M)$ will be close as well.
-In practice, this coarsening procedure is essential to enabling the practical computation of the fibered barcodes.
-
-% However, if a give bipersistence module $M$ is coarsened at all, the 
-%  Specifically, assembling the barcode takes just $O(\log \kappa + \lvert \mathcal{B}_L(M) \rvert ) \approx O(\lvert \mathcal{B}_L(M) \rvert )$ time when $\kappa$, a coarsening parameter discussed below, is treated as a constant.
+Despite its elegance, there are significant computational hurdles to overcome to make the 2-parameter persistence outlined above practical at scale. 
+Using vineyards,~\cite{lesnick2015interactive} give an analysis demonstrating the barcodes template computation requires on the order of $O(m^3 \kappa + m \kappa^2 \log \kappa)$ elementary operations and $O(m \kappa^2)$ storage.
+Since the number of 2-cells in $\mathcal{A}(M)$ is on the order $O(\kappa^2)$, and $\kappa$ itself is on the order of $O(m^2)$ in the worst case, the worse-case complexity of the barcode templates computation $O(m^5)$---this is not only the highest complexity sub-procedure in RIVET~\cite{rivet}, but also the most demanding in practice.
+Despite this significant barrier, there is some room for optimism: in practice, the external stability result from~\cite{landi2014rank} justifies the use of a coarsening procedure which approximates the module $M$ with a smaller module $M'$ via a grid-like reduction, enabling practitioners to restrict the size of $\kappa$ to a relatively small constant.
+This in-turn dramatically reduces the size of $\mathcal{A}(M)$ and thus the number of barcode templates to compute. 
+Moreover, the ordering of barcode templates given by traversing the dual graph of $\mathcal{A}(M)$ along an Eulerian path implies that adjacent barcode templates should be relatively similar, so long as $\kappa$ is not too small.
+%Intuitively, similar barcode templates should be easier to compute in sequence than independently. 
+Indeed, as algorithm~\ref{alg:schedule} was designed for precisely such a computation, 2-parameter persistence is prototypical of the class of methods that stand to benefit from move scheduling.
 
  \subsubsection*{Natural images dataset} 
- In the previous section, we outlined the computational theory of multidimensional persistence, including its relevance to our proposed move scheduling approach.
-In this section, we demonstrate the efficiency of our method on practical use case of multidimensional persistence: detecting the presence of a low-dimensional topological space which well-approximates the distribution of texture patches extracted from real-world image data.
+As we have outlined the computational theory of 2-parameter persistence and elucidated it relevance to our proposed move scheduling approach, we now demonstrate the efficiency of scheduling with a practical application: detecting the presence of a low-dimensional topological space which well-approximates the distribution of texture patches extracted from real-world image data.
 
@@
 
+% REQUIRED: siamplain format
+%\bibliographystyle{siamplain}
+%\bibliographystyle{unsrt}
+\bibliography{move_schedules}
+
+\newpage 
+\appendix
+\section*{Appendix}
+
+\subsection{Algorithms}
+
+\subsubsection*{Move Algorithms}
+\begin{algorithm}[!htb]
+	\caption{Move Right Algorithm}\label{alg:mr}
+	\setstretch{1.15}
+    \begin{algorithmic}[1]
+        \Function{RestoreRight}{$R$, $V$, $\mathbb{I} = \{I_1, I_2, \dots, I_s \}$}
+            \State $(\, d_{low}, \, d_R, \, d_V \,) \gets ( \, \mathrm{low}_R(I_1), \, \mathrm{col}_R(I_1), \, \mathrm{col}_V(I_1) \, )$
+        	\For{$k$ \textbf{in} $I_2, \dots, I_s$} 
+        		\State $( \, d_{low}', \, d_R', \, d_V' \, ) \gets ( \, \mathrm{low}_R(k), \, \mathrm{col}_R(k), \, \mathrm{col}_V(k) \, )$
+        		\State $( \, \mathrm{col}_R(k), \, \mathrm{col}_V(k) \, ) \mathrel{+}= (\, d_R, \, d_V \,) $
+        		\If {$d_{low}' < d_{low}$}
+        			\State $(\, d_{low}, \, d_R, \, d_V \,) \gets (\, d_{low}', \, d_R', \, d_V' \, )$
+        		\EndIf
+        	\EndFor 
+        	\State \Return $(\, R, \, V, \, d_R, \, d_V \, )$
+        \EndFunction
+    \end{algorithmic}
+    \begin{algorithmic}[1]
+        \Function{MoveRight}{$R$, $V$, $i$, $j$} 
+            \State $\mathbb{I} =$ columns satisfying $V[i, i:j] \neq 0$ %\Comment{$O(m)$}
+            \State $\mathbb{J} = $ columns satisfying $\mathrm{low}_R \in [i:j]$ and $\mathrm{row}_R(i) \neq 0$ %\Comment{$O(m)$}
+            \State $(\, R, \, V, d_R, d_V \,) \gets $ \Call{RestoreRight}{$R$, $V$, $\mathbb{I}$} 
+            %\Comment{Restores $p$-simplices} \Comment{$O(\lvert \mathbb{I} \rvert)$}
+            \State $(\, R, \, V \,) \gets $ \Call{RestoreRight}{$R$, $V$, $\mathbb{J}$}
+            	%\Comment{Restores $(p+1)$-simplices}
+            	%\Comment{$O(m \lvert \mathbb{I} \rvert)$}
+            \State $(\, R, \, V\, ) \gets (\, P R P^T, \, P V P^T\,)$  % \Comment{$P$ is a permutation given by (i,j)}
+            \State $(\, \mathrm{col}_R(j), \, \mathrm{col}_V(j) \, ) \gets (\, P d_R, \, P d_V \,)$
+            \State \Return $(\, R, \, V\,)$
+        \EndFunction
+    \end{algorithmic}
+\end{algorithm}
+
+\begin{algorithm}[!htb]
+	\caption{Move Left Algorithm}\label{alg:ml}
+	\setstretch{1.15}
+    \begin{algorithmic}[1]
+        \Function{RestoreLeft}{$R$, $V$, $\mathbb{K} = \{k_1, k_2, \dots, k_s \}$}
+            % \State $\mathrm{low}_K \gets \{ \mathrm{low}_R(k_1), \mathrm{low}_R(k_2), \dots, \mathrm{low}_R(k_s) \}$
+            \State $(l, r) \gets $ indices $l, r \in \mathbb{K}$ satisfying $l < r$, $\mathrm{low}_R(l) = \mathrm{low}_R(r)$ maximal 
+            \While{$\mathrm{low}_R(l) \neq 0$ \textbf{and} $\mathrm{low}_R(r) \neq 0$}
+            	\State $(\mathrm{col}_R(r), \mathrm{col}_V(r)) \mathrel{+}= (\mathrm{col}_R(l), \mathrm{col}_V(l))$
+            	\State $\mathbb{K} \gets \mathbb{K} \setminus l$
+            	\State $(l, r) \gets $ indices $l, r \in \mathbb{K}$ satisfying $l < r$, $\mathrm{low}_R(l) = \mathrm{low}_R(r)$ maximal
+            \EndWhile
+        	\State \Return $(\, R, \, V \, )$
+        \EndFunction
+    \end{algorithmic}
+
+    \begin{algorithmic}[1]
+        \Function{MoveLeft}{$R$, $V$, $i$, $j$} 
+        	\State $\mathbb{I} \gets \{ j \}$
+            % \State $(\mathbb{K}, d_R, d_V) \gets (\emptyset, \mathrm{col}_R(i), \mathrm{col}_V(i))$
+            \While{$V(k, i) \neq 0 $\textbf{ for }$ k = \mathrm{low}_V(i)$\textbf{ where }$j \leq k < i$}
+            	\State $(\mathrm{col}_R(i), \mathrm{col}_V(i)) \mathrel{+}= (\mathrm{col}_R(k), \mathrm{col}_V(k))$
+            	\State $\mathbb{I} \gets \mathbb{I} \cup k + 1 $
+            \EndWhile
+            \State $(R, V) \gets (P R P^T, P V P^T)$
+            \State $\mathbb{J} = $ columns satisfying $\mathrm{low}_R \in [i:j]$ and $\mathrm{row}_R(i) \neq 0$ %\Comment{$O(m)$}
+
+            \State $(R, V) \gets$ \Call{RestoreLeft}{$R, V, \mathbb{I}$}
+            \State $(R, V) \gets$ \Call{RestoreLeft}{$R, V, \mathbb{J}$}
+            % \State $(\mathrm{col}_R(i), \mathrm{col}_V(i)) \gets (P d_R, P d_V )$
+			\State \Return $(\, R, \, V\,)$
+        \EndFunction
+    \end{algorithmic}
+
+\end{algorithm}
+
+\subsubsection{LCS-Sort}
+ \algnewcommand{\IfThenElse}[3]{% \IfThenElse{<if>}{<then>}{<else>}
+  \State \algorithmicif\ #1\ \algorithmicthen\ #2\ \algorithmicelse\ #3}
+  \begin{algorithm}[h]
+	\caption{Sorting algorithm}\label{alg:sorting}
+	\setstretch{1.05}
+    \begin{algorithmic}[1]
+    	% \Require Fixed $p' \in S_m$, $\mathcal{L} = \mathrm{LIS}(p')$ 
+    	\Function{LCS-Sort}{$p$, $q$}
+    		\State $\bar{p} \gets q^{-1} \circ p$
+    		\State $\mathcal{L} \gets \mathrm{LCS}(p,q) = \mathrm{LIS}(\bar{p})$ \Comment{$O(m\log \log m)$}
+    		\State $(\, \mathcal{S}, \, \mathcal{D}, \, \mathcal{T} \, )\gets (\, \emptyset, \, \bar{p} \setminus \mathcal{L}, \, \mathcal{L} \,)$
+    		\While{$\mathcal{D}$\text{ is not empty }}
+    		    \State $\sigma \gets \text{arbitrary element in }\mathcal{D}$
+    		    %\State $\tau, \rho \gets \mathcal{T}_p(\sigma), \mathcal{T}_n(\sigma)$
+    		    %\State $i, i_p, i_n \gets \mathcal{I}_p(\sigma), \, \mathcal{I}_p(T_{\mathrm{pred}}(\sigma)), \, \mathcal{I}_p(T_{\mathrm{succ}}(\sigma))$
+    		    \State $(\,i, \,i_p, \, i_n \,) \gets ( \, \bar{p}^{-1}(\sigma), \, \bar{p}^{-1}(\mathcal{T}_{\mathrm{pred}}(\sigma)), \, \bar{p}^{-1}(\mathcal{T}_{\mathrm{succ}}(\sigma)) \, )$\Comment{$O(\log \log m)$}
+%    		    \If{ $p_{\sigma} < p_{\tau}$ }
+%	    		    \State $j \gets \text{arbitrary element in } [\, p_{\tau}, \, p_{\rho})$
+%    		    \Else{ $p_{\rho} < p_{\sigma}$}
+%    		    	\State $j \gets \text{arbitrary element in } (\, p_{\tau}, \, p_{\rho}]$
+%    		    \EndIf
+				% \IfThenElse{$i < i_p$}{$[\, i_p, \, i_n)$}{$(\, i_p, \, i_n]$}
+				% $\text{Pick } j \in [\, i_p, \, i_n)$ \If{ $i < i_p$ } \Else $(\, i_p, \, i_n]$ \EndIf
+    		    \If{ $i < i_p$ } 
+    		    	\State $j \gets \text{arbitrary element in } [\, i_p, \, i_n)$
+    		    \Else{ $i_n < i$ }
+	    		    \State $j \gets \text{arbitrary element in } (\, i_p, \, i_n]$ 
+    		    \EndIf
+    		    \State $(\, \mathcal{S}, \mathcal{D}, \mathcal{T} \,) \gets (\, \mathcal{S} \cup (i, \, j), \, \mathcal{D} \setminus \sigma, \, \mathcal{T} \cup \sigma \, )$\Comment{$O(\log \log m)$}
+    		    % \State $p, p^{-1} \gets m_{ij} \circ p, p^{-1} \circ m_{ij}^{-1}$
+    		    \State $\bar{p}^{-1} \gets \bar{p}^{-1} \circ m_{ij}^{-1}$ \Comment{$O(m)$}
+    		 \EndWhile
+    		 \State \Return{$\mathcal{S}$}
+    	\EndFunction
+	\end{algorithmic}
+\end{algorithm}
+
+\subsection{2D Persistence}\label{app:2d_pers}
+Indeed, one of the often noted drawbacks of persistence is its instability with respect to strong outliers. 
+Although persistence diagrams are stable with respect to their input---in the Hausdorff sense---and thus are robust to Hausdorff noise, the presence of strong outliers can artificially modify the connectivity of the underlying filtration, obscuring the detection of what would otherwise be significant topological structures~\cite{buchet2015topological}.
+
+The dimension function is simply the the function which maps every point $a \in \mathbb{R}^2$ to $\mathrm{dim}(M_a)$.
+It is a simple and easy to visualize invariant, but is unstable and yields no information about the persistent features of $M$.
+The multigraded Betti numbers constitute a natural and important class of invariants often studied in algebraic geometry and commutative algebra, though both their theory and computation is beyond the scope of this effort---the interested reader is referred to~\cite{lesnick2015interactive, carlsson2009theory} for more details.
+For our purposes, the $i^{\text{th}}$-graded Betti number of $M$ is simply a function $\beta_i(M): \mathbb{R}^2 \to \mathbb{N}$ whose values indicate the the number of elements at  each degree in a basis of the $i^{\text{th}}$ module in a free resolution for $M$.
+
+\subsubsection*{Reparameterization using point-line duality}\label{sec:fibered_barcode_reparam}
+Let $\overline{\mathcal{L}}$ denote the collection of all lines in $\mathbb{R}^2$ with non-negative slope, $\mathcal{L} \subset \overline{\mathcal{L}}$ the collection of all lines in $\mathbb{R}^2$ with non-negative, finite slope, and $\mathcal{L}^\circ$ the collection of all affine lines in $\mathbb{R}^2$ with positive, finite slope.
+There is a standard point-line duality that gives a parameterization of $\mathcal{L}$ with the half-plane $[0, \infty) \times \mathbb{R}$ that is convenient to work with here.
+Define the \emph{line} and \emph{point} dual transforms $\mathcal{D}_{\ell}$ and $\mathcal{D}_p$, respectively, as follows:
+\begin{equation}
+	\begin{aligned}
+\mathcal{D}_{\ell}: \mathcal{L} \rightarrow[0, \infty) \times \mathbb{R} & \quad \hfill \quad & \mathcal{D}_{p}:[0, \infty) \times \mathbb{R} \rightarrow \mathcal{L} \\
+y=a x+b \mapsto(a,-b) &\quad \hfill \quad & (c, d) \mapsto y=c x-d
+\end{aligned}
+\end{equation}
+% Use \multicolumn{1}{|r|}{text} to align
+The transforms $\mathcal{D}_{\ell}$ and $\mathcal{D}_p$ are \emph{dual} to each other in the sense that for any point $a \in [0, \infty) \times \mathbb{R}$ and any line $L \in \mathcal{L}$, $a \in L$ if and only if $D_\ell(L) \in D_{p}(a)$.
+Now, for some fixed line $L$, define  the \emph{push map} $\mathrm{push}_L(a):  \mathbb{R}^2 \to L \cup \infty$ as:
+\begin{equation}
+	\mathrm{push}_L(a) \mapsto \mathrm{min}\{ v \in L \mid a \leq v \}
+\end{equation}
+%\begin{equation*}
+%	\begin{aligned}
+%	& \mathrm{push}_L : & \mathbb{R}^2 & \to L \cup \infty \\
+%	& & a \in \mathbb{R}^2 & \mapsto \mathrm{min}\{ v \in L \mid a \leq v \}
+%	\end{aligned}
+%\end{equation*}
+The push map satisfies a number of useful properties. Namely: 
+\begin{enumerate}
+	\item For $r < s \in \mathbb{R}^2$, $\mathrm{push}_L(r) \leq \mathrm{push}_L(s)$
+	\item For each $a \in \mathbb{R}^2$, $\mathrm{push}_L(a)$ is continuous on $\mathcal{L}^\circ$
+	\item For $L \in \mathcal{L}^\circ$ and $S \subset \mathbb{R}^2$, $\mathrm{push}_L$ induces an ordered partition $S_L$ on $S$ 
+\end{enumerate}
+Property (1) asserts that for any $L \in \overline{\mathcal{L}}$, the standard partial order on $\mathbb{R}^2$ restricts to a total order on $L$.
+Properties (2) and (3) qualify the following definition:
+\begin{definition}[Critical Lines]
+	For some fixed $S \subset \mathbb{R}^2$, a line $L \in L^\circ$ is defined to be \emph{regular} if there is an open ball $B \in L^\circ$ containing $L$ such that $S_L = S_{L'}$ for all $L' \in B$.
+Otherwise, the line $L$ is defined as \emph{critical}.
+\end{definition}
+\noindent The set of critical lines $\mathrm{crit}(M)$ with respect to some fixed set $S \subset \mathbb{R}^2$ fully characterizes a certain planar subdivision of the half plane $[0, \infty) \times \mathbb{R}$.
+This planar subdivision, denoted by $\mathcal{A}(M)$, is thus entirely determined by $S$ under point line duality.
+%It has been shown~\cite{lesnick2015interactive} that the 1-skeleton $\mathcal{A}^1(M)$ of the line arrangement $\mathcal{A}(M)$ can be characterized b
+%Explicitly, define $\mathcal{A}^1(M)$ to the 1-skeleton of the line arrangement defined by: 
+% \{ \mathcal{D}_p(\alpha) \text{ for } \alpha = \mathrm{LUB}(u,v), u,v \in S \} 
+%\begin{equation}\label{eq:arrangement_crit}
+%	\mathcal{A}^1(M) = \{ \mathcal{D}_{\ell}(\mathrm{crit}(M)) \cup (\{0\} \times \mathbb{R}) \}
+%\end{equation}
+%Expanding the 1-skeleton to a 2-D cell complex $\mathcal{A}(M)$ fully partitions the upper half plane.  
+A corollary from~\cite{lesnick2015interactive} shows that if the duals of two lines $L, L' \in \mathcal{L}$ are contained in the same $2$-cell in $\mathcal{A}(M)$, then $S_L = S_{L'}$, i.e.
+the partitions induced by $\mathrm{push}_L$ are equivalent.
+Indeed, the total order on $S_L$ is simply the pullback of the total order on $L$ with respect to the push map.
+Since $\mathcal{A}(M)$ partitions the entire half-plane, the dual to every line $L \in \mathcal{L}$ is contained within $\mathcal{A}(M)$---the desired reparameterization.
+
+Let $S = \mathrm{supp}\,\beta_0(M) \, \cup \, \mathrm{supp}\,\beta_1(M)$, where the functions $\beta_0(M), \beta_1(M)$ are $0^{\text{th}}$ and $1^{\text{st}}$ bigraded Betti numbers of $M$, respectively.
+The main mathematical result from~\cite{lesnick2015interactive} is a characterization of the barcodes $\mathcal{B}_L(M)$, for any $L \in \mathcal{L}$, in terms of a set of \emph{barcode templates} $\mathcal{T}$ computed at every 2-cell in $\mathcal{A}(M)$.
+ More formally, for any line $L \in \overline{\mathcal{L}}$ and $e$ any 2-cell in $\mathcal{A}(M)$ whose closure contains the dual of $L$ under point-line duality, the 1-parameter restriction of the persistence module $M$ induced by $L$ is given by:
+ \begin{equation}
+ \mathcal{B}_L(M) = \{ [\,\mathrm{push}_L(a), \, \mathrm{push}_L(b)\,) \mid (a,b) \in \mathcal{T}^e, \mathrm{push}_L(a) < \mathrm{push}_L(b)  \} 
+\end{equation}
+Minor additional conditions are needed for handling completely horizontal and vertical lines. 
+The importance of this theorem lies in the fact that the fibered barcodes are completely defined from the precomputed barcode templates $\mathcal{T}$---once every barcode template $\mathcal{T}^e$ has been computed and augmented onto $\mathcal{A}(M)$, $\mathcal{B}(M)$ is completely characterized, and the barcodes $\mathcal{B}_L(M)$ associated to a 1-D filtration induced by \emph{any} choice of $L$ can be efficiently computed via a point-location query on $\mathcal{A}(M)$ and a $O(\lvert \mathcal{B}_L(M) \rvert)$ application of the push map.
+
+\subsection{Computation}
+Computationally, the algorithm can be summarized into three core steps: 
+\begin{enumerate}
+	\item Computing the bigraded Betti numbers $\beta(M)$ of $M$
+	\item Constructing a line arrangement $\mathcal{A}(M)$ induced by critical lines from (1) 
+	\item Augmenting $\mathcal{A}(M)$ with a \emph{barcode template} $\mathcal{T}_e$ at every 2-cell $e \in \mathcal{A}(M)$
+	%, yielding the \emph{augmented arrangement} $\mathcal{A}^{\sbullet}(M)$ 
+\end{enumerate}
+Computing (1) takes approximately $\approx O(m^3)$ using a matrix algorithm similar to Algorithm~\ref{alg:reduce}~\cite{lesnick2019computing}.
+Constructing and storing the line arrangement $\mathcal{A}(M)$ with $n$ lines and $k$ vertices is related to the \emph{line segment intersection problem}, which known algorithms in computational geometry can solve in (optimal) output-sensitive $O((n+k) \log n)$ time~\cite{boissonnat2000robust}.
+In terms of space complexity, the number of 2-cells in $\mathcal{A}(M)$ is upper bounded by $O(\kappa^2)$, where $\kappa$ is a coarseness parameter associated with the computation of $\beta(M)$.
+
+There are several approaches one can use to compute $\mathcal{T}$, the simplest being to run Algorithm~\ref{alg:reduce} independently on the 1-D filtration induced by the duals of some set of points (e.g.
+the barycenters) lying in the interior of the 2-cells of $\mathcal{A}(M)$.
+The approach taken by~\cite{lesnick2015interactive} is to use the $R = DV$ decomposition computed at some adjacent 2-cell $e \in \mathcal{A}(M)$ to speed up the computation of an adjacent cell $e' \in \mathcal{A}(M)$.
+More explicitly, define the \emph{dual graph} of $\mathcal{A}(M)$ to be the undirected graph $G$ which has a vertex for every 2-cell $e \in \mathcal{A}(M)$ and an edge for each adjacent pair of cells $e, e' \in \mathcal{A}(M)$.
+Each vertex in $G$ is associated with a barcode template $\mathcal{T}^e$, and the computation of $\mathcal{T}$ now reduces to computing a path $\Gamma$ on $G$ which visits each vertex at least once.
+To minimize the computation time, assume the $n$ edges of $G$ are endowed with non-negative weights $W = w_1, w_2, \dots, w_n$ whose values $w_i \in \mathbb{R}_{+}$ represent some notion of distance.
+The optimal path $\Gamma^\ast$ is the one of minimal length with respect to $W$ which visits every vertex of $G$ at least once.
+There is a known $\frac{3}{2}$-approximation that can be computed efficiently which reduces the problem to the traveling salesman problem on a metric graph~\cite{christofides2022worst}.
+
+%The purpose of the following sections is to uncover the intrinsic bottlenecks impeding the computation of the fibered barcode invariant in practical settings.
+Towards this goal, we briefly summarize the theory established by several researchers~\cite{lesnick2015interactive, carlsson2010computing, landi2014rank}, focusing on their computational and algorithmic details.
+%Our goal is to show that the highest complexity subcomputation of the fibered barcode invariant algorithm proposed by~\cite{lesnick2015interactive} is precisely the type of computation we focus on in this effort.
+%In doing so, we substantiate our claims from section~\ref{sec:motivation} that our scheduling approach can improve the efficiency of a real-world application of multidimensional persistence by an order of magnitude by comparing Algorithm~\ref{alg:schedule} to several alternatives on a real-world data set and problem: the problem of identifying the underlying topological space of a data set extracting from high-contrast natural images.
+% In what follows, we briefly recount an incomplete but nonetheless very useful invariant of multi-dimensional persistence modules, the \emph{rank invariant}, as well as another invariant which in some sense determines the rank invariant, the \emph{fibered barcode}.
+It turns out that computing the latter invariant
+
+% --- For each section of the paper, consider writing a mini-introduction that says what its organization is, what is in each subpart, and how the parts relate to one another.
+
+%To demonstrate how our scheduling approach can accelerate the multidimensional persistence computation, we establish the computational framework we will be considering. 
+% These transforms end up being crucial in decomposing the definition of the fibered barcode (and thus, the rank invariant) into a discrete computation.
+%We begin with a few definitions.
+An $n-D$ \emph{filtration} is a functor $\mathcal{F} : \mathbb{R}^n \to \textbf{Simp}$, the category of simplicial complexes, such that for all $a \leq b$, $\mathcal{F}(a,b) : F_a \to F_b$ is an inclusion.
+We will focus exclusively on the case where $n = 2$, where the object of study is a \emph{bifiltration}.
+Multifiltrations can be broadly categorized into two types: 1-critical and multi-critical filtrations.
+A \emph{1-critical filtration} $\mathcal{F}$ is a finite $n-D$ filtration each simplex $s \in K$ is associated with only one multi-grade (denoting its appearance in each of the filtered dimensions), otherwise $\mathcal{F}$ is said to \emph{multi-critical}.
+%An $n-D$ persistence module is a functor $M : \mathbb{R}^n \to \textbf{Vect}$, where $\textbf{Vect}$ denotes the category of $\mathbf{F}$-vector spaces and linear maps for some suitable field $\mathbb{F}$.
+A module $M$ is said to be \emph{pointwise finite dimensional} (p.f.d) if $\mathrm{dim}(M_a) < \infty$ for all $a \in \mathbb{R}^n$.
+%For $i \geq 0$, let $H_i : \textbf{Simp} \to \textbf{Vect}$ denote the $i^{\text{th}}$ simplicial homology functor with coefficients in $\mathbb{F}$. 
+%Using this notation, given a 1D filtration $\mathcal{F}$ of size $m$, each p.f.d 1D persistence module $M = H_i(\mathcal{F})$ for  may be associated with a multiset $\mathcal{B}(M)$ of intervals in $\mathbb{R}$ which record the isomorphism classes of indecomposable summands of $M$.
+The set $\mathcal{B}(M)$ is called the \emph{barcode}\footnote{Note that \emph{barcodes} and \emph{persistence diagrams} are interchangeable representations of the same multiset.} of $M$.
+%\emph{The rank invariant} proposed by Carlsson et al~\cite{carlsson2009theory} is one such invariant that is complete when $n=1$ but is otherwise incomplete when $n \geq 2$. 
+%% The rank invariant is computable in polynomial time~\cite{carlsson2010computing}, that has been the subject of much interest lately. 
+%It is defined as follows: 
+%\begin{definition}[Rank Invariant]
+%	For $n \geq 1$, the \emph{rank invariant} of a $\mathbb{R}^n$-indexed persistence module $M$ over $\mathcal{H}^{n} \subseteq \mathbb{R}^n \times \mathbb{R}^n$ is the function:
+%	\begin{equation*}
+%		\begin{aligned}
+%		& \mathrm{rank}(M)  \; : & \mathcal{H}^n & \to \mathbb{N}  \\
+%		& & (a,b) & \mapsto \mathrm{rank} \, M(a,b) 
+%		\end{aligned}
+%	\end{equation*}
+%	for $a \leq b \in \mathcal{H}^n$. When $n = 1$, the rank invariant is complete, thus $\mathrm{rank}(M)$ and the persistence barcodes $\mathcal{B}(M)$ determine each other.
+%\end{definition}
+%\noindent Thus, in one dimension, persistent homology provides a complete invariant that fully captures the lifetime of topological features in a given filtration: 1-D persistent homology modules decompose in an essentially unique way into indecomposable summands~\cite{zomorodian2005computing}.
+In contrast, the theory of multidimensional persistence shows that no complete discrete invariant exists---the structure of the target for the invariant depends on the structure of the underlying field, and although multiparameter persistence modules also decompose in a similar way as the 1-D case, the set of isomorphism classes is known to be extremely complicated~\cite{carlsson2009theory, lesnick2015interactive}.
+%Nonetheless, there are invariants related to multi-parameter persistence that can be quite useful in practice, even though they are incomplete. 
+%\noindent For example, even though the rank invariant does not encode the isomorphism type of the underlying module $M$ when $n > 1$, it does captures important ``first-order'' information about the modules structure.
+ %The multigraded Betti numbers constitute a natural and important class of invariants often studied in algebraic geometry commutative algebra.
+There are many ways of defining these, with perhaps the simplest definition stemming from minimal resolutions: for a finitely generated $n$-parameter persistence module $M$ and $i \in \{0, 1, \dots, d \}$, the $i^{\text{th}}$ graded Betti number of $M$ at grade $z$, denoted as $\beta_i^M(z)$, is defined as the number of elements at grade $z$ in a basis of the $i^{\text{th}}$ module in a free resolution for $M$.
+
+%\subsubsection*{The Fibered Barcode}
+%In the context of persistence, and \emph{invariant} is a function from the collection of persistence modules to some set $S$ with the property that $f(M) = f(M')$ when $M \cong M'$.
+An invariant is said to be \emph{complete} if $M \cong M'$ whenever $f(M) = f(M')$.
+Thus, incomplete invariants are invariants which may have the same value on two non-isomorphic modules.
+
+%The fibered barcode enjoys a number of stability properties readily expressed using interleaving distances~\cite{landi2014rank}. 
+% More important to this work, the fibered barcode $\mathcal{B}(M)$ and the rank invariant $\mathrm{rank}(M)$ determine each other~\cite{lesnick2015interactive}.
+
+%For some fixed set $S \subset \mathbb{R}^2$ and any pair $u, v \in S$ of points which are incomparable with respect to the partial order on $\mathbb{R}^2$
+% The definition of \emph{critical} lines provides an analogue to the
+
+%This is the main mathematical result from~\cite{lesnick2015interactive},
+
+% Indeed, it was shown in~\cite{lesnick2015interactive} that the fibered barcodes $$ of $M$ (i.e.
+the collection of barcodes of 1D affine slices of $M$) and the rank invariant determine each other.
+%\subsubsection*{Computing the fibered barcode}
+%Lesnick and Wright introduced the Rank Invariant Visualization and Exploration Tool (RIVET)~\cite{lesnick2015interactive}, which is both a tool and a collection of computational and mathematical theory which enables exploratory analysis of 2D persistence modules built from arbitrary bifiltrations.
+From a computational perspective, the core novelty of the RIVET approach is its ability to compute the fibered barcodes efficiently from precomputed information.
+
+%% --------------- Trash ----------------
 %\[
@@
 
-% REQUIRED: siamplain format
-%\bibliographystyle{siamplain}
-%\bibliographystyle{unsrt}
-\bibliography{move_schedules}
-
 %\appendix 

### 2023-01-24 move_schedules.tex
Crocker stack fixes

@@
 We give a tight lower bound on the number of moves needed to perform an arbitrary permutation to the $R = D V$ decomposition, and give a proof of optimality by a reduction to the permutation edit distance problem.
-We also give worst-case bounds in expectation as well as efficient algorithms for achieving these bounds---both of which are derived from a reduction to the Longest Increasing Subsequence (LIS) problem.
+We also give worst-case sizes of these quantities in expectation as well as efficient algorithms for achieving these bounds---both of which are derived from a reduction to the Longest Increasing Subsequence (LIS) problem.
 This reduction yields an efficient algorithm for generating sequences of moves $( s_1, s_2, \dots, s_d )$ of minimal size $d$, which we call \emph{schedules}. 
@@
 \section{Background}\label{sec:background} 
-\subsection{Notation}\label{sec:notation}
-
+%\subsection{Notation}\label{sec:notation} 
 Suppose one has a family $\{K_i\}_{i\in I}$ of simplicial complexes indexed by a totally ordered   set $I$, and so that for any $i< j \in I$ we have $K_i \subseteq K_j$.
@@
 %For this reason, we'll refer to the difference $\lvert i - j \rvert$ of a homology class $[c]$ as its \emph{index persistence}.
-\\
-\\
-\noindent \textbf{Example 2.1:} Non-simplexwise filtrations often arise in geometrical contexts.
+%\\
+%\\
+%\noindent \textbf{Example 2.1:} Non-simplexwise filtrations often arise in geometrical contexts.
 For example, given a finite metric space $(X, d)$, the \emph{Vietoris-Rips} complex at scale $\epsilon \in \mathbb{R}$ is the abstract simplicial complex given by:
-\begin{equation}\label{eq:rips}
-	\mathrm{Rips_\epsilon}(X) = \{ S \subseteq X : S \neq \emptyset \;\mbox{ and }\;\mathrm{diam}(S) \leq \epsilon \}
-\end{equation} 
-The corresponding filtration is called the \emph{Vietoris-Rips} (VR) filtration, indexed by $I = \mathbb{R}$. 
-A VR filtration indexed over $\mathbb{R}$ can be condensed to an simplexwise filtration indexed over the set of pairwise distances $\{\, d(x,x') \; \mid \; x,x' \in X \, \}$  by extending a total order of $X$ to its power set using any linear extension of the face poset, such as e.g.
+%\begin{equation}\label{eq:rips}
+%	\mathrm{Rips_\epsilon}(X) = \{ S \subseteq X : S \neq \emptyset \;\mbox{ and }\;\mathrm{diam}(S) \leq \epsilon \}
+%\end{equation} 
+%The corresponding filtration is called the \emph{Vietoris-Rips} (VR) filtration, indexed by $I = \mathbb{R}$. 
+%A VR filtration indexed over $\mathbb{R}$ can be condensed to an simplexwise filtration indexed over the set of pairwise distances $\{\, d(x,x') \; \mid \; x,x' \in X \, \}$  by extending a total order of $X$ to its power set using any linear extension of the face poset, such as e.g.
 the lexicographical ordering.
@@
 If there are only finitely many of these homological critical values, then $f$ is said to be \emph{tame}.
-
-Consider a homotopy $F(x,\tau) : \mathbb{X} \times [0,1] \to \mathbb{R}$ and denote its ``snapshot'' at a given time-point $\tau$ by $f_\tau(x) = F(x,\tau)$.
-The snapshot $f_0$ denotes the initial function at time $\tau = 0$ and $f_1$ denotes the function at the last time step. 
-As $\tau$ varies in $[0,1]$, the points in $\mathrm{dgm}_p(f_\tau)$ trace curves in $\mathbb{R}^3$, called \emph{vines}, which collectively form a \emph{vineyard}. 
-The stability of persistence implies that these curves will be continuous if $F$ is continuous
-and the $f_\tau$'s are tame \cite{cohen2007stability}.
-The vineyard analogy acts as a guidepost for practitioners seeking to understand how subtle changes occurring to the topological structure over time.
-A more detailed discussion of the vineyards approach is provided in section~\ref{sec:vineyards}.
+The concept of homological critical points and tameness will be revisited in section~\ref{sec:vineyards}.
+
 %Let $\mathbb{M}$ be a smooth $d$-manifold and $f : \mathbb{M} \to \mathbb{R}$ a smooth function. 
@@
 In this section we briefly recount the original reduction algorithm introduced in~\cite{zomorodian2005computing}, also sometimes called the \emph{standard} algorithm or more explicitly \texttt{pHcol}~\cite{de2011dualities}.
-The pseudocode is outlined in Algorithm~\ref{alg:reduce}.
+The pseudocode is outlined in Algorithm~\ref{alg:reduce} in the appendix.
 Without optimizations, like   clearing   or       implicit matrix reduction, the standard  algorithm is very inefficient.
-Nonetheless, it serves as the foundation of most implementations that compute persistent homology, and its invariants are necessary before introducing both vineyards in section~\ref{sec:vineyards} and our move schedules  in section~\ref{sec:move_schedules}.
-
-\begin{algorithm}[t]
-	\setstretch{1.15}
-	\caption{Reduction Algorithm (\texttt{pHcol}) }
-	\begin{algorithmic}[1]
-		\Require{$D = (m \times m)$ filtration boundary matrix }
-		\Ensure{$R$ is reduced, $V$ is full rank upper triangular, and $R = D V$}
-		\Function{Reduction}{$D$}
-		\State $(R, V) \gets (D, I)$
-		\For{$j = 1$ \textbf{to} $m$} 
-			\While{$\exists \, i < j$ \textbf{such that} $\mathrm{low}_R(i) = \mathrm{low}_R(j)$ }
-				\State $\lambda \gets \mathrm{pivot}_R(j)/\mathrm{pivot}_R(i)$
-				\State $(\mathrm{col}_R(j), \mathrm{col}_V(j)) \mathrel{-}= \left ( \lambda \cdot \mathrm{col}_R(i), \lambda \cdot \mathrm{col}_V(i) \right )$
-			\EndWhile
-		\EndFor 
-		\State \Return $(R, V)$
-		\EndFunction
-	\end{algorithmic}	\label{alg:reduce}
-\end{algorithm}
+Nonetheless, it serves as the foundation of most persistent homology implementations, and its invariants are necessary before introducing both vineyards in section~\ref{sec:vineyards} and our move schedules  in section~\ref{sec:move_schedules}.
 
@@
 \noindent
-Inspection of Algorithm~\ref{alg:reduce} suggests that a loose upper bound for the reduction is $O(m^3)$, where $m$ is the number of simplices of the filtration---this bound is in fact tight~\cite{morozov2005persistence}.
+Inspection of the reduction algorithm from~\cite{edelsbrunner2000topological} suggests that a loose upper bound for the reduction is $O(m^3)$, where $m$ is the number of simplices of the filtration---this bound is in fact tight~\cite{morozov2005persistence}.
 Nonetheless, many variations and optimizations to Algorithm~\ref{alg:reduce} have been proposed over the past decade, see~\cite{bauer2017phat, bauer2021ripser, chen2011persistent} for an overview.
@@
 \subsection{Vineyards}\label{sec:vineyards}
+Consider a homotopy $F(x,\tau) : \mathbb{X} \times [0,1] \to \mathbb{R}$ on a triangulable topological space $\mathbb{X}$, and denote its ``snapshot'' at a given time-point $\tau$ by $f_\tau(x) = F(x,\tau)$.
+The snapshot $f_0$ denotes the initial function at time $\tau = 0$ and $f_1$ denotes the function at the last time step. 
+As $\tau$ varies in $[0,1]$, the points in $\mathrm{dgm}_p(f_\tau)$ trace curves in $\mathbb{R}^3$, which will be continuous if $F$ is continuous and the $f_\tau$'s are tame, due to the stability of persistence.
+Cohen-steiner et al.~\cite{cohen2007stability} referred to these curves as \emph{vines}, as collection of which forms as \emph{vineyard}---the geometric analogy meant to act as a guidepost for practitioners seeking to understand how subtle changes occurring to the topological structure over time.
+
 The original purpose of the vineyards algorithm, as described in~\cite{cohen2006vines}, was to compute a continuous 1-parameter family of persistence diagrams over a time-varying filtration, detecting homological critical events along the way.
@@
 As a result, any fixed set of rules that maintains a valid $R = D V$ decomposition under adjacent transpositions is sufficient to simulate persistence dynamically.
+
 These rules prescribe certain column and row operations to apply to a given matrix decomposition either before, during, or after each individual transposition to ensure it is a valid decomposition.   
-
-Let $S_{i}^j$ represent the upper-triangular matrix such that $A S_{i}^j$ is equivalent to the elementary operation that adds column $i$ of $A$ to column $j$ of $A$ and $S_{i}^j A$ is equivalent to adding row $j$ of $A$ to row $i$ of $A$.
+Formally, let $S_{i}^j$ represent the upper-triangular matrix such that $A S_{i}^j$ is equivalent to the elementary operation that adds column $i$ of $A$ to column $j$ of $A$ and $S_{i}^j A$ is equivalent to adding row $j$ of $A$ to row $i$ of $A$.
 Similarly, let $P$ denote the matrix so that $A P$ applies a fixed permutation to the columns of $A$ and $P^T A$ applied the same permutation to the rows of $A$.
@@
 
-For completeness, we now describe the moves algorithm from~\cite{busaryev2010tracking}.
+%For completeness, we now describe the moves algorithm from~\cite{busaryev2010tracking}.
 
@@
 
-\noindent \textbf{Example:} We continue the example used in sections~\ref{sec:vineyards} and~\ref{sec:reduction} to illustrate moves.
+\noindent \textbf{Example 2.3:} We re-use the running example from sections~\ref{sec:reduction} and~\ref{sec:vineyards} to illustrate moves.
 Consider moving the edge $a$ to the position of edge $c$ in the filtration.
@@
 	\includegraphics[width=0.95\textwidth]{crocker_combo_1.png}
-	\caption{An example of a crocker plot (right) depicting the evolution of dimension $p = 1$ Betti curves over time.
+	\caption{A crocker plot (right) depicts the evolution of dimension $p = 1$ Betti curves over time.
 The green $\mathrm{X}$ marks correspond chronologically to the complexes (left), in row-major order.
@@
 \v{C}ech filtrations).} plot.
-In brief, a crocker plot is a contour plot on a family of Betti curves.
+In brief, a crocker plot is a contour plot of a family of Betti curves.
 Formally, given a filtration $K = K_0 \subseteq K_1 \subseteq \dots \subseteq K_m$, a $p$-dimensional \emph{Betti curve} $\beta_p^{\bullet}$ is defined as the ordered sequence of $p$-th dimensional Betti numbers:
 $$ \beta_p^\bullet = \{ \, \mathrm{rank}(H_p(K_0)), \, \mathrm{rank}(H_p(K_1)), \, \dots, \, \mathrm{rank}(H_p(K_m))\, \}$$
-Given a time-varying filtration $K(\tau)$, a crocker plot can be interpreted as a contour plot on the 1-parameter family of Betti curves $\beta_p^\bullet(\tau)$.
+Given a time-varying filtration $K(\tau)$, a crocker plot displays changes to $\beta_p^\bullet(\tau)$ as a function of $\tau$.
 A example of a crocker plot generated from the simulation described below is given in Figure~\ref{fig:crocker1}.
@@
 
-Acknowledging this, Xian et al.~\cite{xian2020capturing} showed that crocker plots may be \emph{smoothed} to inherit the stability property of persistence diagrams and reduce noise in the visualization.
+Xian et al.~\cite{xian2020capturing} showed that crocker plots may be \emph{smoothed} to inherit the stability property of persistence diagrams and reduce noise in the visualization.
 That is, when applied to a time-varying persistence module $M = \{M_t\}_{t \in [0, T]}$ an $\alpha$-smoothed crocker plot for $\alpha \geq 0$ is the rank of the map $M_t(\epsilon - \alpha) \to M_t(\epsilon + \alpha)$ at time $t$ and scale $\epsilon$.
@@
 	$$ f_M(t,\epsilon, \alpha) = \mathrm{rank}(M_t(\epsilon - \alpha) \to M_t(\epsilon + \alpha)) $$
+	and $f_M$ satisfies $f_M(t,\epsilon,\alpha') \leq f_M(t,\epsilon, \alpha)$ for all $0 \leq \alpha \leq \alpha'$.
 \end{definition}
-\noindent A crocker stack is a sequence of $\alpha$-smoothed crocker plots that vary over $\alpha \geq 0$, satisfying $f_M(t,\epsilon,\alpha) \leq f_M(t,\epsilon, \alpha)$ for all $\alpha \geq \alpha'$.
-Note that, unlike crocker plots, applying this $\alpha$ smoothing efficiently \emph{requires} the persistence pairing.
+%\noindent A crocker stack is a sequence of $\alpha$-smoothed crocker plots that vary over $\alpha \geq 0$, satisfying $f_M(t,\epsilon,\alpha) \leq f_M(t,\epsilon, \alpha)$ for all $\alpha \geq \alpha'$.
+\noindent Note that, unlike crocker plots, applying this $\alpha$ smoothing efficiently \emph{requires} the persistence pairing.
 Indeed, it has been shown that crocker stacks and stacked persistence diagrams (i.e.
@@
 \subsection{Algorithms}
+
+\subsubsection{Reduction Algorithm}
+\begin{algorithm}[t]
+	\setstretch{1.15}
+	\caption{Reduction Algorithm (\texttt{pHcol}) }
+	\begin{algorithmic}[1]
+		\Require{$D = (m \times m)$ filtration boundary matrix }
+		\Ensure{$R$ is reduced, $V$ is full rank upper triangular, and $R = D V$}
+		\Function{Reduction}{$D$}
+		\State $(R, V) \gets (D, I)$
+		\For{$j = 1$ \textbf{to} $m$} 
+			\While{$\exists \, i < j$ \textbf{such that} $\mathrm{low}_R(i) = \mathrm{low}_R(j)$ }
+				\State $\lambda \gets \mathrm{pivot}_R(j)/\mathrm{pivot}_R(i)$
+				\State $(\mathrm{col}_R(j), \mathrm{col}_V(j)) \mathrel{-}= \left ( \lambda \cdot \mathrm{col}_R(i), \lambda \cdot \mathrm{col}_V(i) \right )$
+			\EndWhile
+		\EndFor 
+		\State \Return $(R, V)$
+		\EndFunction
+	\end{algorithmic}	\label{alg:reduce}
+\end{algorithm}
 

### 2023-01-26 move_schedules.tex
Added moves description

@@
 \newtheorem{corollary}{Corollary}
+\newtheorem{remark}{Remark}
 
@@
 In light of these results, we give an alternative proxy-objective for cost minimization, provide bounds justifying its relevance to the original problem, and give an efficient $O(d^2 \log m)$ algorithm for approximately solving this proxy minimization.
-A performance comparison with other reduction-based persistent homology computations is given; 
-our move schedules are demonstrated to be an order of magnitude more efficient than existing approaches in applications such as flock simulations and 2D persistence computations with RIVET. 
+A performance comparison with other reduction-based persistence computations is given, wherein move schedules are demonstrated to be an order of magnitude more efficient than existing approaches at calculating persistence in dynamic settings.
+In particular, we illustrate the effectiveness of efficient scheduling with a variety of real-world applications, including flock analysis in dynamic metric spaces and manifold detection from image data using 2D persistence computations.
 %We conclude with a discussion of situations where our strategy is most applicable and future directions for research.
@@
 Although $d$ clearly is $O(m)$ for pathological inputs, we show that $d \sim m - \sqrt{m}$ in expectation for random filtrations (Corollary~\ref{cor:expectation}) is not too pessimistic for truly random filtrations, and we give empirical results suggesting $d$ is tends to be much smaller for time-varying filtrations.
-
+\\
 % given a decomposition $R = \partial_{K} V$ of a simplicial filtration $K$, algorithm~\ref{alg:schedule} can produce such a sequence $\mathcal{S} = \left( s_1, s_2, \dots, s_d \right)$
@@
 the lexicographical ordering.
-%This is often necessary in the computational setting, where it is easier to work with a filtration with $m$ simplices indexed over $[m]$ instead of $\mathbb{R}$. 
-\\
-\\
+
 Let $\mathbb{X}$ be a triangulable topological space and $K$ an abstract simplicial complex whose geometric realization is homeomorphic to $\mathbb{X}$. 
@@
  The main output of the reduction algorithm is a matrix decomposition $R = D V$, where the persistence diagram is encoded in $R$, and the generating cycles in the columns of $V$. 
- Given a filtration $K$ with $m$ simplices, one assembles the elementary boundary chains $\partial(\sigma)$ as columns ordered according to the filtration, forming the square $m \times m$  \emph{filtration boundary matrix} $D$.
- The reduction algorithm computes the persistent homology from $D$ for all dimensions up to $\mathrm{dim}(K) - 1$.
-  Alternatively, a single dimension $p \leq \mathrm{dim}(K) - 1$ can be computed, in which case one reduces a pair of sub-matrices $D_p$ and $ D_{p+1}$ of $D$,  where $D_p$ has dimension $d_{p-1} \times d_p$ and $D_{p+1}$ is $d_{p} \times d_{p+1}$.
-In some situations the description of the ensuing algorithms is simpler when considering all dimensions simultaneously, while in others it is more succinct to consider dimension-specific matrices. 
-In what follows $D$, $R$, and $V$ will always denote square $(m \times m)$ matrices, though we note that all algorithms discussed generalize to both cases. 
- We give an example of the reduction algorithm using the matrix pair $(R_1, V_1)$ below. 
+ Given a filtration $K$ with $m$ simplices, one assembles the elementary boundary chains $\partial(\sigma)$ as columns ordered according to the filtration into an $m \times m$  \emph{filtration boundary matrix} $D$.
+ It is at times more succinct to restrict to specific sub-matrices $D$ based on the homology dimension $p$, and so we write $D_p$ represent the $d_{p-1} \times d_p$ matrix representing the linear operator $\partial_p(K)$ (the same notation is extended to $R$ and $V$).
+% The reduction algorithm computes the persistent homology from $D$ for all dimensions up to $\mathrm{dim}(K) - 1$.
+%  Alternatively, a single dimension $p \leq \mathrm{dim}(K) - 1$ can be computed, in which case one reduces a pair of sub-matrices $D_p$ and $ D_{p+1}$ of $D$,  where $D_p$ has dimension $d_{p-1} \times d_p$ and $D_{p+1}$ is $d_{p} \times d_{p+1}$.
+%In some situations the description of the ensuing algorithms is simpler when considering all dimensions simultaneously, while in others it is more succinct to consider dimension-specific matrices. 
+%In what follows $D$, $R$, and $V$ will always denote square $(m \times m)$ matrices, though we note that all algorithms discussed generalize to both cases. 
+ We illustrate the reduction algorithm with an example pair $(R_1, V_1)$ below. 
 \\
@@
 	%
-	\begin{blockarray}{cccc}{R_1}
+	\begin{blockarray}{cccc}{D_1}
 	 \kern-0.2em & a & b & c  \\
@@
 	\; , \;
-	\begin{blockarray}{cccc}{V_1}
+	\begin{blockarray}{cccc}{I_1}
 	\kern-0.2em & a & b & c  \\
@@
 	\quad \xrightarrow{} \,
-	\begin{blockarray}{cccc}{R_1}
+	\begin{blockarray}{cccc}{}
 	\kern-0.2em & a & b & c  \\
@@
 	\; , \;
-	\begin{blockarray}{cccc}{V_1}
+	\begin{blockarray}{cccc}{}
 	\kern-0.2em & a & b & c  \\
@@
  	\item[I1.] $R = D V$ where $D$ is the boundary matrix of the filtration $K$
- 	\item[I2.] $V$ is full-rank upper-triangular, and $R$ is \emph{reduced}: if its $i$-th and $j$-th columns are nonzero, then $\mathrm{low}_R(i) \neq \mathrm{low}_R(j)$ 
+ 	\item[I2.] $V$ is full-rank upper-triangular
+ 	\item[l3.] $R$ is \emph{reduced}: if both $\mathrm{col}_i(R) \neq 0$ and $\mathrm{col}_j(R) \neq 0$, then $\mathrm{low}_R(i) \neq \mathrm{low}_R(j)$ 
  	\end{enumerate} 
@@
 The original purpose of the vineyards algorithm, as described in~\cite{cohen2006vines}, was to compute a continuous 1-parameter family of persistence diagrams over a time-varying filtration, detecting homological critical events along the way.
-As homological critical events can only occur when the filtration order changes, detecting all such critical events can be reduced to computing valid decompositions at all time points interleaving  changes in the filtration order.
-At the finest scale, these changes manifest as transpositions of adjacent simplices for simplexwise filtrations. 
-As a result, any fixed set of rules that maintains a valid $R = D V$ decomposition under adjacent transpositions is sufficient to simulate persistence dynamically.
+As homological critical events only occur when the filtration order changes, detecting all such events may be reduced to computing valid decompositions at all time points interleaving  changes in the filtration order.
+For simplexwise filtrations, these changes manifest as transpositions of adjacent simplices, and thus any fixed set of rules that maintains a valid $R = D V$ decomposition under adjacent transpositions is sufficient to simulate persistence dynamically.
 
@@
 
-%For completeness, we now describe the moves algorithm from~\cite{busaryev2010tracking}.
-
-Though similar, move operations confer a few computational advantages compared to vineyards. We focus on two convenient properties of moves:
+For completeness, we recapitulate the moves algorithm from~\cite{busaryev2010tracking}. 
+Let $K$ denote a filtration of size $m = \lvert K \rvert$ and $R = DV = [ r_1, r_2, \dots, r_m]$ its decomposition, where $r_j$ ($v_j$) will denote the $j$-th column of $R$ ($V$).
+By definition, if $r_j = D v_j = 0$, then the inclusion $K_{j-1} \hookrightarrow K_{j}$ introduced a new cycle and homology class of dimension $p = \mathrm{dim}(\sigma_j)$ whose generator is encoded by the non-zero coefficients $(\alpha_i^{(j)}, \alpha_{i+1}^{(j)}, \dots, \alpha_j^{(j)})$ via the combination:
+\begin{equation}
+	r_j = D v_j = \alpha_i^{(j)} \cdot \partial(\sigma_{i+1}) + \alpha_{i+1}^{(j)} \cdot \partial(\sigma_{i+1}) + \dots + \alpha_{j}^{(j)} \cdot \partial(\sigma_{j}) = 0
+\end{equation}
+Now, consider the permutation $P$ that moves a simplex $\sigma_i$ in $K$ to position $j$, shifting all intermediate simplices $\sigma_{i+1}, \dots, \sigma_{j}$ down by one ($i < j$). 
+To perform this shift, these non-zero coefficients $\alpha_i^{(i)}, \alpha_{i}^{(i+1)}, \dots, \alpha_i^{(j)}$ encoded in the boundary chains $r_i, r_{i+1}, \dots, r_j$ must be set to zero, otherwise $\partial(\sigma_i)$ contributes to earlier simplices in the filtration.
+In matrix form, these coefficients are recorded by $v_j$, and thus setting $\alpha_i^k = 0$ for all $k \in [i,j]$ amounts to setting $v_i(i), v_{i+1}(i), \dots, v_{j}(i)$ to $0$.
+
+Computationally, setting $v_i(i) = v_{i+1}(i) = \dots = v_{j}(i) = 0$ with column operations $\ast$ ensures $PV\ast$ is upper-triangular, thus maintaining invariant I2.
+To maintain invariant I1, these operations induce a corresponding $R' = P R \ast P^T$, which may be unreduced (breaking invariant I3).
+We could reducing $R'$ with an additional $k$ operations, however $k$ may scale quadratically in the interval $\lvert i - j\rvert$, rendering the efficiency of this approach at best that of vineyards and at worse that of the standard algorithm.
+To bypass this difficulty, Busaryev et al. observed that since the initial $R$ is reduced, if it contains $s$ pivot entries in the columns $[i,j]$ of $R$, then $R'$ must also have $s$ pivots. 
+In particular, if during the cancellation  of $v_i(i) = v_{i+1}(i) = 0$ the column $r_{i+1} \mapsto r_{i+1}'$ becomes unreduced, then the pivot $\mathrm{low}_R(i+1)$ becomes a \emph{free}, possibly becoming a pivot again in $r_{i+2}', \dots, r_j'$.
+Thus, if we save $r_{i+1}$ prior to modification to a \emph{donor column}, we may re-use or \emph{donate} its pivot entry to one of $r_{i+2}, \dots, r_j$. 
+This process is repeated at most $j - i - 1$ times---by construction, the matrix $R'$ stays reduced during the movement in all except possibly for the $i$-th column, and since these operations are left-to-right $V'$ will be upper-triangular.
+%We may achieve this via left-to-right column operations, as before, however this may leave several chains in $R$ unreduced. 
+%To ameliorate this, observe that since the initial set of chains $r_i, r_{i+1}, \dots, r_j$ are already reduced, their pivots are unique---thus, we should expect these entries. 
+%setting any individual $\alpha_i^(k) = 0$ for some $i < k \leq j$ has the effect of changing its pivot entry
+% participates in a linear combination of another 
+%By definition, the column $r_i$ 
+\\
+\\
+\noindent \textbf{Example 2.3:} We re-use the running example from sections~\ref{sec:reduction} and~\ref{sec:vineyards} to illustrate moves.
+The donor columns of $R$ and $V$ are denoted as $d_R$ and $d_V$, respectively.
+Consider moving edge $a$ to the position of edge $c$ in the filtration.
+\begin{displaymath}
+	\def\rowsp{0.4em}
+	\def\arraystretch{0.60}
+	\makeatletter\setlength\BA@colsep{0.9pt}\makeatother
+	%
+	\begin{blockarray}{cc}{d_R}
+	\kern-0.2em & a \\
+		\begin{block}{c[c]}
+  		u \kern \rowsp & \; 1 \; \topstrut \\
+  		v \kern \rowsp & \;  \; \\
+  		w \kern \rowsp & \; 1 \; \botstrut \\
+		\end{block}
+	\end{blockarray}
+	\; \;
+	\begin{blockarray}{cccc}{R}
+	 \kern-0.2em & a & b & c  \\
+		\begin{block}{c[ccc]}
+  		u \kern \rowsp & \; 1 & 1 &  \hphantom{1} \; \topstrut \\
+  		v \kern \rowsp & \;  & 1 &  \hphantom{1} \; \\
+  		w \kern \rowsp & \; 1 &  &  \hphantom{1} \; \botstrut \\
+		\end{block}
+	\end{blockarray}
+	%
+\; \to
+	%
+\begin{blockarray}{cc}{}
+	\kern-0.2em & b \\
+		\begin{block}{c[c]}
+  		  & \; 1 \; \topstrut \\
+  		 & \;  1 \; \\
+  		 & \; \; \botstrut \\
+		\end{block}
+	\end{blockarray}
+	\; \;
+\begin{blockarray}{cccc}{}
+	\kern-0.2em & a & b & c  \\
+		\begin{block}{c[ccc]}
+  		  & \; 1 &    &  \hphantom{1} \; \topstrut \\
+  		  & \;    & 1 &  \hphantom{1} \; \\
+  		 & \; 1 & \underline{1} &  \hphantom{1} \; \botstrut \\
+		\end{block}
+	\end{blockarray}
+	%
+\; \to
+	%
+	\begin{blockarray}{cc}{}
+	\kern-0.2em & c \\
+		\begin{block}{c[c]}
+  		  & \; \hphantom{1} \; \topstrut \\
+  		 & \;  \hphantom{1} \; \\
+  		 & \; \hphantom{1} \; \botstrut \\
+		\end{block}
+	\end{blockarray}
+	\; \; 
+	\begin{blockarray}{cccc}{}
+	\kern-0.2em & a & b &  c \\
+		\begin{block}{c[ccc]}
+  		  & \; 1   &    &   1  \; \topstrut \\
+  		 & \;       &  1  &  1  \; \\
+  		 & \; 1   &   \underline{1}   &     \; \botstrut \\
+		\end{block}
+	\end{blockarray}
+%
+\; \xrightarrow{P} 
+\begin{blockarray}{cc}{}
+	\kern-0.2em & a \\
+		\begin{block}{c[c]}
+  		  & \; \hphantom{1} \; \topstrut \\
+  		 & \; \hphantom{1} \; \\
+  		 & \; \hphantom{1} \; \botstrut \\
+		\end{block}
+	\end{blockarray}
+\; \;
+\begin{blockarray}{cccc}{}
+	\kern-0.2em & b & c & a  \\
+		\begin{block}{c[ccc]}
+  		  & \;  & 1 &  1\; \topstrut \\
+  		 & \; 1 & 1 &  \; \\
+  		 & \; 1 &  & \underline{1} \; \botstrut \\
+		\end{block}
+	\end{blockarray}
+%
+\xrightarrow{d_R} 
+\begin{blockarray}{cccc}
+\kern-0.2em & b & c & a  \\
+	\begin{block}{c[ccc]}
+		& \;  & 1 &  \hphantom{1} \; \topstrut \\
+		& \; 1 & 1 &  \hphantom{1} \; \\
+		& \; 1 &  &  \hphantom{1} \; \botstrut \\
+	\end{block}
+	% \undermat{B}{ & & }  
+\end{blockarray}
+\end{displaymath}
+
+\vspace{-3.0em}
+
+% start V example (1) 
+\begin{displaymath}
+	\def\rowsp{0.4em}
+	\def\arraystretch{0.60}
+	\makeatletter\setlength\BA@colsep{0.9pt}\makeatother
+	%
+	\begin{blockarray}{cc}{d_V}
+	\kern-0.2em & a \\
+		\begin{block}{c[c]}
+  		a \kern \rowsp & \; 1 \; \topstrut \\
+  		b \kern \rowsp & \;  \; \\
+  		c \kern \rowsp & \;  \; \botstrut \\
+		\end{block}
+	\end{blockarray}
+	\; \;
+	\begin{blockarray}{cccc}{V}
+	 \kern-0.2em & a & b & c  \\
+		\begin{block}{c[ccc]}
+  		a \kern \rowsp & \; 1 & \underline{1} & \underline{1} \; \topstrut \\
+  		b \kern \rowsp & \;  & 1 & 1 \; \\
+  		c \kern \rowsp & \;  &  & 1 \; \botstrut \\
+		\end{block}
+	\end{blockarray}
+	%
+\; \to
+	%
+\begin{blockarray}{cc}{}
+	\kern-0.2em & b \\
+		\begin{block}{c[c]}
+  		  & \; 1 \; \topstrut \\
+  		 & \;  1 \; \\
+  		 & \; \; \botstrut \\
+		\end{block}
+	\end{blockarray}
+	\; \;
+\begin{blockarray}{cccc}{}
+	\kern-0.2em & a & b & c  \\
+		\begin{block}{c[ccc]}
+  		  & \; 1 &    & \underline{1} \; \topstrut \\
+  		  & \;    & 1 &  1 \; \\
+  		 & \;  & &  1 \; \botstrut \\
+		\end{block}
+	\end{blockarray}
+	%
+\; \to
+	%
+	\begin{blockarray}{cc}{}
+	\kern-0.2em & c \\
+		\begin{block}{c[c]}
+  		  & \; 1 \; \topstrut \\
+  		 & \; 1 \; \\
+  		 & \; 1 \; \botstrut \\
+		\end{block}
+	\end{blockarray}
+	\; \; 
+	\begin{blockarray}{cccc}{}
+	\kern-0.2em & a & b &  c \\
+		\begin{block}{c[ccc]}
+  		  & \; 1   &    &    \; \topstrut \\
+  		 & \;   &  1  &   \; \\
+  		 & \;   &   &  1   \; \botstrut \\
+		\end{block}
+	\end{blockarray}
+%
+\; \xrightarrow{P} 
+\begin{blockarray}{cc}{}
+	\kern-0.2em & a \\
+		\begin{block}{c[c]}
+  		  & \; 1 \; \topstrut \\
+  		 & \; 1 \; \\
+  		 & \; 1 \; \botstrut \\
+		\end{block}
+	\end{blockarray}
+\; \;
+\begin{blockarray}{cccc}{}
+	\kern-0.2em & b & c & a  \\
+		\begin{block}{c[ccc]}
+  		  & \; 1 &  &  \; \topstrut \\
+  		 & \;  & 1 &  \; \\
+  		 & \; &  & 1 \; \botstrut \\
+		\end{block}
+	\end{blockarray}
+%
+\xrightarrow{d_V} 
+\begin{blockarray}{cccc}
+\kern-0.2em & b & c & a  \\
+	\begin{block}{c[ccc]}
+		& \;  1 & &  1 \; \topstrut \\
+		& \;  & 1 &  1\; \\
+		& \; &  & 1 \; \botstrut \\
+	\end{block}
+	% \undermat{B}{ & & }  
+\end{blockarray}
+\end{displaymath}
+Note that the equivalent permutation using vineyards requires $4$ column operations on both $R_1$ and $V_1$, respectively, whereas a single move operation accomplishes using only $2$ column operations per matrix.
+The pseudo-code for \emph{MoveRight} is given in Algorithm~\ref{alg:mr} and for \emph{MoveLeft} in Algorithm~\ref{alg:ml}.
+
+%\begin{remark}
+%	Conceptually, there is a conceptual similarity between the cyclic permutation that moves a simplex in a filtration the \emph{cycle sort}~\cite{haddon1990cycle}, a simple sorting algorithm method that has $O(n^2)$ time complexity for general inputs and $O(n)$ complexity for permutations.
+%	Indeed, just as cycle sort rotates elements within a \emph{cycle} (in the permutation sense) using $O(1)$ auxiliary storage, the move right operation  
+%\end{remark}
+Though similar to vineyards, move operations confer a few additional advantages. We note two convenient properties of moves:
 \\	
@@
 Indeed, the number of intermediate persistence diagrams is bounded above by $O(m^2)$ using vineyards if each pair of simplices $\sigma_i, \sigma_j \in K$ switches relative ordering at most once in a given schedule of transpositions, whereas with move operations this bound clearly is at most $O(m)$.
-\\
-
-\noindent \textbf{Example 2.3:} We re-use the running example from sections~\ref{sec:reduction} and~\ref{sec:vineyards} to illustrate moves.
-Consider moving the edge $a$ to the position of edge $c$ in the filtration.
-The donor columns introduced in~\cite{busaryev2010tracking} are shown on the left side of each matrix.
-Note that using vineyards, the equivalent permutation to the decomposition would require $4$ column operations on both $R_1$ and $V_1$, respectively, whereas a single move operation accomplishes this permutation using only $2$ column operations per matrix.
-\begin{displaymath}
-	\def\rowsp{0.4em}
-	\def\arraystretch{0.60}
-	\makeatletter\setlength\BA@colsep{0.9pt}\makeatother
-	%
-	\begin{blockarray}{cc}{d_R}
-	\kern-0.2em & a \\
-		\begin{block}{c[c]}
-  		u \kern \rowsp & \; 1 \; \topstrut \\
-  		v \kern \rowsp & \;  \; \\
-  		w \kern \rowsp & \; 1 \; \botstrut \\
-		\end{block}
-	\end{blockarray}
-	\; \;
-	\begin{blockarray}{cccc}{R}
-	 \kern-0.2em & a & b & c  \\
-		\begin{block}{c[ccc]}
-  		u \kern \rowsp & \; 1 & 1 &  \hphantom{1} \; \topstrut \\
-  		v \kern \rowsp & \;  & 1 &  \hphantom{1} \; \\
-  		w \kern \rowsp & \; 1 &  &  \hphantom{1} \; \botstrut \\
-		\end{block}
-	\end{blockarray}
-	%
-\; \to
-	%
-\begin{blockarray}{cc}{}
-	\kern-0.2em & b \\
-		\begin{block}{c[c]}
-  		  & \; 1 \; \topstrut \\
-  		 & \;  1 \; \\
-  		 & \; \; \botstrut \\
-		\end{block}
-	\end{blockarray}
-	\; \;
-\begin{blockarray}{cccc}{}
-	\kern-0.2em & a & b & c  \\
-		\begin{block}{c[ccc]}
-  		  & \; 1 &    &  \hphantom{1} \; \topstrut \\
-  		  & \;    & 1 &  \hphantom{1} \; \\
-  		 & \; 1 & \underline{1} &  \hphantom{1} \; \botstrut \\
-		\end{block}
-	\end{blockarray}
-	%
-\; \to
-	%
-	\begin{blockarray}{cc}{}
-	\kern-0.2em & c \\
-		\begin{block}{c[c]}
-  		  & \; \hphantom{1} \; \topstrut \\
-  		 & \;  \hphantom{1} \; \\
-  		 & \; \hphantom{1} \; \botstrut \\
-		\end{block}
-	\end{blockarray}
-	\; \; 
-	\begin{blockarray}{cccc}{}
-	\kern-0.2em & a & b &  c \\
-		\begin{block}{c[ccc]}
-  		  & \; 1   &    &   1  \; \topstrut \\
-  		 & \;       &  1  &  1  \; \\
-  		 & \; 1   &   \underline{1}   &     \; \botstrut \\
-		\end{block}
-	\end{blockarray}
-%
-\; \xrightarrow{P} 
-\begin{blockarray}{cc}{}
-	\kern-0.2em & a \\
-		\begin{block}{c[c]}
-  		  & \; \hphantom{1} \; \topstrut \\
-  		 & \; \hphantom{1} \; \\
-  		 & \; \hphantom{1} \; \botstrut \\
-		\end{block}
-	\end{blockarray}
-\; \;
-\begin{blockarray}{cccc}{}
-	\kern-0.2em & b & c & a  \\
-		\begin{block}{c[ccc]}
-  		  & \;  & 1 &  1\; \topstrut \\
-  		 & \; 1 & 1 &  \; \\
-  		 & \; 1 &  & \underline{1} \; \botstrut \\
-		\end{block}
-	\end{blockarray}
-%
-\xrightarrow{d_R} 
-\begin{blockarray}{cccc}
-\kern-0.2em & b & c & a  \\
-	\begin{block}{c[ccc]}
-		& \;  & 1 &  \hphantom{1} \; \topstrut \\
-		& \; 1 & 1 &  \hphantom{1} \; \\
-		& \; 1 &  &  \hphantom{1} \; \botstrut \\
-	\end{block}
-	% \undermat{B}{ & & }  
-\end{blockarray}
-\end{displaymath}
-
-\vspace{-3.0em}
-
-% start V example (1) 
-\begin{displaymath}
-	\def\rowsp{0.4em}
-	\def\arraystretch{0.60}
-	\makeatletter\setlength\BA@colsep{0.9pt}\makeatother
-	%
-	\begin{blockarray}{cc}{d_V}
-	\kern-0.2em & a \\
-		\begin{block}{c[c]}
-  		a \kern \rowsp & \; 1 \; \topstrut \\
-  		b \kern \rowsp & \;  \; \\
-  		c \kern \rowsp & \;  \; \botstrut \\
-		\end{block}
-	\end{blockarray}
-	\; \;
-	\begin{blockarray}{cccc}{V}
-	 \kern-0.2em & a & b & c  \\
-		\begin{block}{c[ccc]}
-  		a \kern \rowsp & \; 1 & \underline{1} & \underline{1} \; \topstrut \\
-  		b \kern \rowsp & \;  & 1 & 1 \; \\
-  		c \kern \rowsp & \;  &  & 1 \; \botstrut \\
-		\end{block}
-	\end{blockarray}
-	%
-\; \to
-	%
-\begin{blockarray}{cc}{}
-	\kern-0.2em & b \\
-		\begin{block}{c[c]}
-  		  & \; 1 \; \topstrut \\
-  		 & \;  1 \; \\
-  		 & \; \; \botstrut \\
-		\end{block}
-	\end{blockarray}
-	\; \;
-\begin{blockarray}{cccc}{}
-	\kern-0.2em & a & b & c  \\
-		\begin{block}{c[ccc]}
-  		  & \; 1 &    & \underline{1} \; \topstrut \\
-  		  & \;    & 1 &  1 \; \\
-  		 & \;  & &  1 \; \botstrut \\
-		\end{block}
-	\end{blockarray}
-	%
-\; \to
-	%
-	\begin{blockarray}{cc}{}
-	\kern-0.2em & c \\
-		\begin{block}{c[c]}
-  		  & \; 1 \; \topstrut \\
-  		 & \; 1 \; \\
-  		 & \; 1 \; \botstrut \\
-		\end{block}
-	\end{blockarray}
-	\; \; 
-	\begin{blockarray}{cccc}{}
-	\kern-0.2em & a & b &  c \\
-		\begin{block}{c[ccc]}
-  		  & \; 1   &    &    \; \topstrut \\
-  		 & \;   &  1  &   \; \\
-  		 & \;   &   &  1   \; \botstrut \\
-		\end{block}
-	\end{blockarray}
-%
-\; \xrightarrow{P} 
-\begin{blockarray}{cc}{}
-	\kern-0.2em & a \\
-		\begin{block}{c[c]}
-  		  & \; 1 \; \topstrut \\
-  		 & \; 1 \; \\
-  		 & \; 1 \; \botstrut \\
-		\end{block}
-	\end{blockarray}
-\; \;
-\begin{blockarray}{cccc}{}
-	\kern-0.2em & b & c & a  \\
-		\begin{block}{c[ccc]}
-  		  & \; 1 &  &  \; \topstrut \\
-  		 & \;  & 1 &  \; \\
-  		 & \; &  & 1 \; \botstrut \\
-		\end{block}
-	\end{blockarray}
-%
-\xrightarrow{d_V} 
-\begin{blockarray}{cccc}
-\kern-0.2em & b & c & a  \\
-	\begin{block}{c[ccc]}
-		& \;  1 & &  1 \; \topstrut \\
-		& \;  & 1 &  1\; \\
-		& \; &  & 1 \; \botstrut \\
-	\end{block}
-	% \undermat{B}{ & & }  
-\end{blockarray}
-\end{displaymath}
-
-\noindent
-We recall an important claim given in~\cite{busaryev2010tracking} on the effect that move operations have on the status of simplices in the pairing.
-Recall from section~\ref{sec:background} that simplices which create new homology classes are called \emph{creators} and simplices that destroy homology classes are called \emph{destroyers}.
-%\begin{proposition}[Locality of moves~\cite{busaryev2010tracking}]
-%	Other than the simplex $\sigma_i$ being moved, $\mathrm{MoveRight}(i,j)$ cannot change a destroyer to a creator, $\mathrm{MoveLeft}(i,j)$ cannot change a creator to destroyer, and neither move operation changes the status of any simplex in $K$ outside of the interval $[i,j]$.
-%\end{proposition}
-\noindent The effect of the movement on intermediate simplices depends on the direction of the movement.
-If $i < j$ (respectively, $j < i$), all simplices at positions $k \in [i+1\;:\;j] $ are shifted down (respectively, up) by $1$.
-The pseudo-code for the \emph{MoveRight} operation is given in Algorithm~\ref{alg:mr}, and the pseudo-code for the \emph{MoveLeft} is given in Algorithm~\ref{alg:ml}.
-
-%\begin{remark}
-%We note that while the pseudocode for \emph{MoveRight} is included in~\cite{busaryev2010tracking}, the one for \emph{MoveLeft} is not, and Algorithm~\ref{alg:ml} is the one we propose and leverage in this paper.
-%\end{remark}
-
-Although conceptually similar, there is an asymmetry between \emph{MoveRight} and \emph{MoveLeft}.
-Moving a simplex upwards in the filtration requires removing non-zero entries along several columns of a particular row in $V$ so that the corresponding permutation does not render $V$ non-upper triangular.
-The key insight of the algorithm presented in~\cite{busaryev2010tracking} is that $R$ can actually be maintained in all but one column during this procedure (by employing the \emph{donor} column).
-In contrast, moving a simplex to an earlier time in the filtration requires removing non-zero entries along several rows of a particular column of $V$.
-As before, $R$ stays reduced during this cancellation procedure in all but one column, however the subsequent permutation to $R$ requires reducing a pair of columns which may cascade into a larger chain of column operations to keep $R$ reduced.
-Since these operations always occur in a left-to-right fashion, its not immediately clear how to apply a donor column kind of concept.
-Fortunately, we can still separate  these two stages.
+
+As a final note, we remark that the combination of \emph{MoveRight} and \emph{MoveLeft} enable the dynamic addition or deletion of simplices to the underlying complex $K$ (in fact this was the motivation for its conception).
+In particular, given $K$ and its decomposition $R = DV$, obtaining the decomposition $R' = D' V'$ of $K' = K \cup \sigma$ can be achieved by appending its requisite elementary chains to $D$ and $V$, reducing them, and then calling \emph{MoveLeft}$(m+1, i)$, where $i = f'(\sigma)$.
+Dually, deleting a simplex $\sigma_i$ may be achieved via \emph{MoveRight} by moving $i$-th to the end and dropping them.
 % with a logarithmic loss in efficiency, although executing the movement has a higher constant factor.
@@
 %\subsection{Proposed approach}
-Let us begin with a brief overview of the pipeline, which we outline in pseudocode with Algorithm~\ref{alg:schedule} below.
-As before, we assume as input  a discrete $1$-parameter family of filtrations $\mathcal{K} = (K_1, K_2, \dots, K_n)$ of an abstract simplicial complex $K$ with $|K|=m$, and the goal being to compute their persistence diagrams.
+Let us begin with a brief overview of the pipeline, outlined in Algorithm~\ref{alg:schedule} below.
+As before, we assume as input  a discrete $1$-parameter family of filtrations $\mathcal{K} = (K_1, K_2, \dots, K_n)$ of an abstract simplicial complex $K$ with $\lvert K \rvert =m$, and the goal being to compute their persistence diagrams.
 Fix reindexing  bijections $f_i : K_i \to K_{i+1}$. 
 Each $f_i$ induces a bijection $f_i^\ast : [m] \to [m]$, or equivalently, a permutation of the index set $[m]$.
-For each pair of filtrations $(K_i, K_{i+1})$, we assign labels to the simplices of $K_i$ using the index set $[m]$ and relabel $K_{i+1}$ accordingly using $f_i$.
-Denote these permutations by $p$ and $q$.
-We compute the longest increasing subsequence (LIS) of $q$ and use this subsequence to recover a longest common subsequence (LCS) of $p, q$, denoted by $\mathrm{LCS}(p,q)$.
-We pass $p$, $q$, and this LCS to our greedy scheduling algorithm, which returns as output an ordered set of move permutations  $\mathcal{S}$ of minimum size, which we call a \emph{schedule}.
+For each pair of filtrations $(K_i, K_{i+1})$, we assign the simplices of $K_i$ labels from index set $[m]$ in filtration order, relabeling $K_{i+1}$ accordingly using $f_i$.
+As $p$ is now a strictly increasing sequence, we may compute the longest increasing subsequence $\mathrm{LIS}(q)$ and use this subsequence to recover a longest common subsequence (LCS) between $(K_i, K_{i+1})$, which we denote later with $\mathrm{LCS}(K_i, K_{i+1})$.
+We pass $q$ and $\mathrm{LIS}(q)$ to our greedy scheduling algorithm, which returns as output an ordered set of move permutations  $\mathcal{S}$ of minimum size, which we call a \emph{schedule}.
 Note one need not explicitly store the entire family of filtrations nor the schedules between them---Algorithm~\ref{alg:schedule} may be easily modified to be completely \emph{online}, keeping at most two filtrations and one decomposition in memory at any given time.
@@
     			\State $\mathrm{lis}_q \gets \mathrm{LIS}(q)$ \Comment{$O(m \log \log m)$}
-    			\State $\mathcal{S} \gets \mathcal{S} \cup \mathrm{GreedySchedule}(p, q, \mathrm{lis}_q)$ \Comment{$O(d^2\log m)$}
+    			\State $\mathcal{S} \gets \mathcal{S} \cup \mathrm{GreedySchedule}(q, \mathrm{lis}_q)$ \Comment{$O(d^2\log m)$}
     		\EndFor
@@
 However, it's not immediately clear whether this bound is tight.
+
 To establish this formally, we require a few definitions.
-
 Let $S_m$ denote the symmetric group  of   permutations of  $[m]$ under function composition $\circ$.
@@
 \end{equation}
-Similarly, in the context of edit operations, observe that a move operation can be interpreted as a paired delete-and-insert operation: 
-$$ m_{ij} = (\mathrm{ins}_j \circ \mathrm{del}_i) $$
-where $\mathrm{del}_i$ denotes the edit operation that deletes the character at position $i$  and $\mathrm{ins}_j$ similarly inserts the same character at position $j$. 
+Similarly, in the context of edit operations, observe that a move operation can be interpreted as a paired delete-and-insert operation, i.e.
+$m_{ij} = (\mathrm{ins}_j \circ \mathrm{del}_i)$, where $\mathrm{del}_i$ denotes the edit operation that deletes the character at position $i$  and $\mathrm{ins}_j$ similarly inserts the same character at position $j$.
 Thus, sorting by move operations can be interpreted as finding a minimal sequence of edits where the only operations allowed are insertions and deletions, where each insertion/deletion operation is coupled---which is exactly the well known \emph{Longest Common Subsequence} (LCS) distance.
@@
 \end{proof}
+\begin{remark}
 \noindent It's worth noting that the quantity from Corollary~\ref{cor:expectation} captures the size of $S^\ast$ under worst-case conditions, i.e.
-to pairs of uniformly sampled permutations, as opposed to uniformly sampled filtrations.
-Simplicial filtrations have more structure due to the subcomplex relation that must be respected, so one might wonder if this structure reduces the the size of $d$ even further.
+between pairs of uniformly sampled permutations, as opposed to uniformly sampled filtrations.
+Simplicial filtrations have more structure due to the subface relation that must be respected, so one might wonder if this structure reduces the the size of $d$ even further.
 However, Boissonnat~\cite{boissonnat2018efficient} has shown that for any $k$-dimensional simplicial complex $K$ with $m$ simplices, the number of distinct filtrations built from $K$ is \emph{at least} $\floor{\frac{t+1}{k+1}}^m$, where $t$ is the number of distinct filtration values associated with the complex $K$.
@@
 For example, the average size of the LCS across the 10 filtrations from Section~\ref{sec:motivation} each having $m=417$ simplices sampled at evenly spaced time points was $343$, implying $d \approx 70$ permutations needed on average to update the decomposition at adjacent time points.
+\end{remark}
 
@@
  %% New section
-\subsection*{Constructing a sorting}
+\subsection*{Constructing schedules}
 It is clear from Corollary~\ref{cor:expectation} that one may compute the LCS between two permutations $p, q \in S_m$ via the LIS of a single permutation $p^\ast$, and that computation may be carried out in $O(m \log \log m)$ time.
 It is not immediately clear, however, how to obtain a sorting $p \mapsto q$ from a given $\mathcal{L} = \mathrm{LIS}(p')$ in an efficient way.
-We outline below a simple procedure which constructs such a sorting $\mathcal{S} = (s_1, \dots, s_d)$ in $O(dm\log m)$ time and  and $O(m)$ space.
+We outline below a simple procedure which constructs such a sorting $\mathcal{S} = (s_1, \dots, s_d)$ in $O(dm\log m)$ time and $O(m)$ space.
 %Alternatively, given $\mathcal{L}$, the algorithm may be executed online (up to $d$ times) in $O(m \log m)$ time and $O(m)$ space.
@@
 First, we require a few definitions.
-Recall that a \emph{sorting} $\mathcal{S}$ with respect to two permutations $p, q \in S_m$ is an ordered sequence of permutations $(s_1, s_2, \dots, s_d)$ whose product under composition with $p$ yields $q$, i.e.
- $$ q = s_d \circ \dots s_1 \circ p $$
-By definition, a subsequence $\mathcal{L}$ of symbols common to both $p$ and $q$ satisfies:
- $$ p^{-1}(\sigma) < p^{-1}(\tau) \implies q^{-1}(\sigma) < q^{-1}(\tau) \quad \forall \sigma, \tau \in \mathcal{L} $$
+Recall that a \emph{sorting} $\mathcal{S}$ with respect to two permutations $p, q \in S_m$ is an ordered sequence of permutations $\mathcal{S} = (s_1, s_2, \dots, s_d)$ satisfying $q = s_d \circ \dots s_1 \circ p$.
+By definition, a subsequence of symbols in $\mathcal{L}$ common to both $p$ and $q$ satisfies:
+\begin{equation}
+	p^{-1}(\sigma) < p^{-1}(\tau) \implies q^{-1}(\sigma) < q^{-1}(\tau) \quad \forall \sigma, \tau \in \mathcal{L}
+\end{equation}
  where $p^{-1}(\sigma)$ (resp.
@@
 
+\begin{figure}[ht]
+	\centering
+	\includegraphics[width=\textwidth, height=16.35em]{boid_sim_results.png}
+	\caption{ The cumulative number of $O(m)$ column operations (left, log-scale) of the three approaches tested and the distribution of the normalized Kendall distance between adjacent filtration pairs.
+The histogram on the right depicts the coarseness of the discretization; most filtration pairs have $\approx 5\%$ of their $\approx O(m^2)$ simplex pairs inverted between adjacent time steps.}\label{fig:boid_sim_results}
+\end{figure}
+
 Xian et al.~\cite{xian2020capturing} showed that crocker plots may be \emph{smoothed} to inherit the stability property of persistence diagrams and reduce noise in the visualization.
@@
 To get a sense of the time domain, we ran the simulation until a vertex made at least 5 rotations around the torus.
-\begin{figure}[ht]
-	\centering
-	\includegraphics[width=\textwidth]{boid_sim_results.png}
-	\caption{ The cumulative number of $O(m)$ column operations (left, log-scale) of the three approaches tested and the distribution of the normalized Kendall distance between adjacent filtration pairs.
-The histogram on the right depicts the coarseness of the discretization; most filtration pairs have $\approx 5\%$ of their $\approx O(m^2)$ simplex pairs inverted between adjacent time steps.}\label{fig:boid_sim_results}
-\end{figure}
 
@@
 Such a generalization has appeared naturally in many application contexts, showing potential as a tool for exploratory data analysis~\cite{lesnick2012multidimensional}.
+Indeed, one of the often noted drawbacks of persistence is its instability with respect to strong outliers, which can obscure the detection of significant topological structures~\cite{buchet2015topological}.
+By filtering the data by density in tandem with the e.g. the ambient metric, multidimensional persistence in theory enables practitioner to handle outliers directly.
 
@@
 This subdivision may be obtained via arrangement of ``critical lines'' induced by the bigraded Betti numbers $\beta(M)$ of $M$.
-As the barcode of one template point $\mathcal{T}_e$ at the two cell $e \in \mathcal{A}(M)$ may be computed efficiently by re-using information from an 2-cell-adjacent template point $\mathcal{T}_{e'}$,~\cite{lesnick2015interactive} observed that computing the barcodes of all such template points (and thus, $\mathcal{B}(M)$) may be reduced to ordering the two-cells in $\mathcal{A}(M)$ along a Eulerian path traversing the dual graph of $\mathcal{A}(M)$.
+As the barcode of one template point $\mathcal{T}_e$ at the 2-cell $e \in \mathcal{A}(M)$ may be computed efficiently by re-using information from an adjacent template point $\mathcal{T}_{e'}$,~\cite{lesnick2015interactive} observed that computing the barcodes of all such template points (and thus, $\mathcal{B}(M)$) may be reduced to ordering the 2-cells in $\mathcal{A}(M)$ along a Eulerian path traversing the dual graph of $\mathcal{A}(M)$.
 The full algorithm is out of scope for this effort; we include supplementary details for the curious reader in the appendix~\ref{app:2d_pers}. 
@@
 Filtering this data set with respect to the Rips parameter and the complement of a kernel density estimate yields a bifiltration whose various invariants are shown in the middle figure.
-The gray areas indicate the region where the dimension function is constant; the lighter gray area where $\mathrm{dim}_1(M) = 1$ indicates a persistent loop was detected.
-On the right side, dual space is shown: the black lines are the critical lines that form $\mathcal{A}(M)$, the blue dashed-lines the edges of the dual graph of $\mathcal{A}(M)$, and the orange barycentric points along the 2-cells of $\mathcal{A}(M)$ represent where the barcodes templates $\mathcal{T}_e$ are parameterized.
+The gray areas indicate homology with positive dimension---the lighter gray area $\mathrm{dim}_1(M) = 1$ indicates a persistent loop was detected.
+On the right side, dual space is shown: the black lines are the critical lines that form $\mathcal{A}(M)$, the blue dashed-lines the edges of the dual graph of $\mathcal{A}(M)$, the rainbow lines overlaying the dashed-lines form the Eulerian path, and the orange barycentric points along the 2-cells of $\mathcal{A}(M)$ represent where the barcodes templates $\mathcal{T}_e$ are parameterized.
 \begin{figure}[t]
@@
 
-Despite its elegance, there are significant computational hurdles to overcome to make the 2-parameter persistence outlined above practical at scale. 
-Using vineyards,~\cite{lesnick2015interactive} give an analysis demonstrating the barcodes template computation requires on the order of $O(m^3 \kappa + m \kappa^2 \log \kappa)$ elementary operations and $O(m \kappa^2)$ storage.
-Since the number of 2-cells in $\mathcal{A}(M)$ is on the order $O(\kappa^2)$, and $\kappa$ itself is on the order of $O(m^2)$ in the worst case, the worse-case complexity of the barcode templates computation $O(m^5)$---this is not only the highest complexity sub-procedure in RIVET~\cite{rivet}, but also the most demanding in practice.
-Despite this significant barrier, there is some room for optimism: in practice, the external stability result from~\cite{landi2014rank} justifies the use of a coarsening procedure which approximates the module $M$ with a smaller module $M'$ via a grid-like reduction, enabling practitioners to restrict the size of $\kappa$ to a relatively small constant.
+Despite its elegance, there are significant computational barriers prohibiting the 2-parameter persistence algorithm outlined from being practical. 
+An analysis from~\cite{lesnick2015interactive} (using vineyards) shows the barcodes template computation requires on the order of $O(m^3 \kappa + m \kappa^2 \log \kappa)$ elementary operations and $O(m \kappa^2)$ storage.
+Since the number of 2-cells in $\mathcal{A}(M)$ is on the order $O(\kappa^2)$, and $\kappa$ itself is on the order of $O(m^2)$ in the worst case, the worse-case complexity of the barcode templates computation $O(m^5)$---this is both the highest complexity and most time-intensive sub-procedure the RIVET software~\cite{rivet} depends on.
+Despite this significant barrier, there is room for optimism: in practice, the external stability result from~\cite{landi2014rank} justifies the use of a coarsening procedure which approximates the module $M$ with a smaller module $M'$ via a grid-like reduction, enabling practitioners to restrict the size of $\kappa$ to a relatively small constant.
 This in-turn dramatically reduces the size of $\mathcal{A}(M)$ and thus the number of barcode templates to compute. 
-Moreover, the ordering of barcode templates given by traversing the dual graph of $\mathcal{A}(M)$ along an Eulerian path implies that adjacent barcode templates should be relatively similar, so long as $\kappa$ is not too small.
+Moreover, the ordering of barcode templates given by the dual graph traversal implies that adjacent  templates should be relatively similar---so long as $\kappa$ is not too small---suggesting barcode computations may be productively shared between adjacent templates.
 %Intuitively, similar barcode templates should be easier to compute in sequence than independently. 
@@
  \subsubsection*{Natural images dataset} 
-As we have outlined the computational theory of 2-parameter persistence and elucidated it relevance to our proposed move scheduling approach, we now demonstrate the efficiency of scheduling with a practical application: detecting the presence of a low-dimensional topological space which well-approximates the distribution of texture patches extracted from real-world image data.
+As we have outlined the computational theory of 2-parameter persistence and elucidated it relevance to our proposed move scheduling approach, we now demonstrate the efficiency of scheduling with a practical application: detecting the presence of a low-dimensional topological space which well-approximates the distribution of patches from high-contrast images.
+
+A common hypothesis is that high dimensional data tend to lie in the vicinity of an embedded, low dimensional manifold or topological space.
+An exemplary demonstration of this is given in~\cite{lee2003nonlinear}, who explored the space of $3 \times 3$ high-contrast patches extracted from Hans van Hateren's still image collection~\cite{hateren_schaaf_1998}, which consists of approximately 4,000 monochrome images\footnote{See \url{http://bethgelab.org/datasets/vanhateren/} for details on the image collection.} depicting various areas outside around Groningen (Holland).
+Motivated by applications to sparse coding applications, Lee et al.~\cite{lee2003nonlinear} was interested in how these natural image patches were distributed, in pixel-space, with respect to predicted spaces and manifolds.
+One of the motivating questions of their research was whether there were any clear qualitative differences the distributions of patches extracted from images of different modalities, such optical and range images.
+A major finding from their work is that the majority of high-contrast image patches are concentrated around a 2-dimensional submanifold embedded in $S^7$.
+Supplemental experiments conducted by Carllsson et al.~\cite{carlsson2008local} using persistent homology confirmed that the distribution of high-contrast $3 \times 3$ patches is indeed well-approximated by a Klein bottle---around 60\% of the high-contrast patches from the still image data set are concentrated within a small neighborhood around a model of a Klein bottle which accounts for $\approx$ 21\% of the volume $S^7$, compared to the $\approx$ 84\% of volume of $S^7$ accounted for by the entire data set.
+Along a similar vein, in the context of sparse coding, Perea et al.~\cite{perea2014klein} introduced a dictionary learning framework for estimating and representing the distribution of patches from texture images.
 
@@
 On the left, a projection the full data set is shown, along with the 15 landmark patches.
-In the middle, the bigraded Betti numbers, the Hilbert function, and a chosen line $L$ are shown for the coarsened grid.
-As before, the $0/1/2$ dimension bigraded Betti numbers are shown in green/red/yellow, respectively.
-The blue regions highlights where $\mathrm{dim}(M) = 5$.
-On the right, the persistence diagram of $M$ restricted to $L$, with larger points indicating a multiplicity of 2.
-Observe five persistent features are revealed in both the middle and right plots, matching $\beta_1$ of the three-circle model.
+(Middle) the bigraded Betti numbers and a fixed line $L$ over parameter space.
+	As before, the $0/1/2$ dimension bigraded Betti numbers are shown in green/red/yellow, respectively, with the blue region highlighting where $\mathrm{dim}(M) = 5$.
+(Right) five persistent features representing $B_L(M)$ are revealed from the middle, matching $\beta_1$ of the three-circle model.
 }
@@
 
-A common hypothesis is that high dimensional data tend to lie in the vicinity of an embedded, low dimensional manifold or topological space.
-An exemplary demonstration of this is given in~\cite{lee2003nonlinear}, who explored the space of $3 \times 3$ high-contrast patches extracted from Hans van Hateren's still image collection~\cite{hateren_schaaf_1998}, which consists of approximately 4,000 monochrome images\footnote{See \url{http://bethgelab.org/datasets/vanhateren/} for details on the image collection.} depicting various areas outside around Groningen (Holland).
-Motivated by applications to sparse coding applications, Lee et al.~\cite{lee2003nonlinear} was interested in how these natural image patches were distributed, in pixel-space, with respect to predicted spaces and manifolds.
-For example, one of the motivating questions of their research was whether there were any clear qualitative differences the distributions of patches extracted from from images of different modalities, e.g.
-optical versus range images.
-A major result from their work is that the majority of high-contrast image patches are concentrated around a 2-dimensional submanifold embedded in $S^7$.
-Supplemental experiments conducted by Carllsson et al.~\cite{carlsson2008local} using persistent homology confirmed that the distribution of high-contrast $3 \times 3$ patches is indeed well-approximated by a Klein bottle---around 60\% of the high-contrast patches from the still image data set are concentrated within a small neighborhood around a model of a Klein bottle which accounts for $\approx$ 21\% of the volume $S^7$, compared to the $\approx$ 84\% of volume of $S^7$ accounted for by the entire data set.
-Along a similar vein, in the context of sparse coding, Perea et al.~\cite{perea2014klein} introduced a dictionary learning framework for estimating and representing the distribution of patches from texture images.
-
 If one was not aware of the analysis done by~\cite{lee2003nonlinear, hateren_schaaf_1998, carlsson2008local, perea2014klein}, it is not immediately clear a priori that the Klein bottle model is indeed a good candidate for capturing non-linearity of image patches.
@@
 
-Consider the (coarsened) fibered barcode computed from a standard Rips / codensity bifiltration on a representative sample of the image data from~\cite{hateren_schaaf_1998} (details to follow), shown in Figure~\ref{fig:patch_data_dgm}.
-Upon inspection of the bigraded Betti number and the dimension function, one finds that a large area of dimension function is constant (highlighted as the blue portion in the middle of Figure~\ref{fig:patch_data_dgm}), wherein the first Betti number is 5.
+Consider the (coarsened) fibered barcode computed from a standard Rips / codensity bifiltration on a representative sample of the image data from~\cite{hateren_schaaf_1998}, shown in Figure~\ref{fig:patch_data_dgm}.
+From the bigraded Betti number and the dimension function, one finds that a large area of dimension function is constant (highlighted as the blue portion in the middle of Figure~\ref{fig:patch_data_dgm}), wherein the first Betti number is 5.
 There are many spaces whose  meets this criterion, though upon further inspection, one candidate space that seems particularly plausible is the three-circle model $C_3$.
@@
 It is advantageous to ``whiten'' the data via a change of coordinates to Euclidean sphere.
-This can be achieved by expressing the patch data as an expansion of a certain convenient basis of 8 non-constant eigenvectors, often called the Discrete Cosine Transform (DCT) basis~\cite{lee2003nonlinear}, which are given below:
+This can be achieved by expressing the patch data as an expansion of a certain convenient basis of 8 non-constant eigenvectors, often called the Discrete Cosine Transform (DCT) basis~\cite{lee2003nonlinear}, shown below:
 %\vspace{-0.2em}
@@
 The results are summarized in Table~\ref{tab:barcode_templates}.
-We also record the number of 2-cells encountered along the path $\Gamma$ permutations that were applied throughout the computation of the barcode templates for both vineyards and moves, denoted in the table as $d_K$ and $d_{\mathrm{LCS}}$, respectively.
+We also record the number of 2-cells in $\mathcal{A}(M)$ and the number of permutations applied throughout the encountered along the traversal of the dual graph for both vineyards and moves, denoted in the table as $d_K$ and $d_{\mathrm{LCS}}$, respectively.
 %Specifically, we begin by extracting a representative subset of the set $X \subset \mathcal{X}$ of patches representing the top 30\% densest patches of the space of all patches $\mathcal{X}$, where the notion of density is approximated with a k-nearest neighbor (knn) density estimator $\rho$.
@@
 \noindent
-As shown on the table, we're able to achieve a significant reduction in the number of total column operations needed to compute $\mathcal{T}$ compared to both vineyards and $\mathtt{pHcol}$, though it's worth noting that permutations are constant time for vineyards while they require at most $O(m)$ elementary operations for moves.
-Observe that initially, when there is a high degree of coarsening, vineyards is particularly inefficient and moves requires only about 3x less column operations that naively computing $\mathcal{T}$ independently.
-As the coarsening becomes more refined and more 2-cells are added to $\mathcal{A}(M)$, however, vineyards quickly becomes a much more viable option compared to $\mathtt{pHcol}$, though even at the highest coarsening we tested the gain in efficient was relatively small.
+As shown on the table, when the coarsening $\kappa$ is small enough, we're able to achieve a significant reduction in the number of total column operations needed to compute $\mathcal{T}$ compared to both vineyards and $\mathtt{pHcol}$.
+%though it's worth noting that permutations are constant time for vineyards while they require at most $O(m)$ elementary operations for moves.
+This is further reinforced by the observation that, when there is a high degree of coarsening, vineyards is particularly inefficient and moves requires only about 3x less column operations that naively computing $\mathcal{T}$ independently.
+As the coarsening becomes more refined and more 2-cells are added to $\mathcal{A}(M)$, however, vineyards quickly becomes a much more viable option compared to $\mathtt{pHcol}$---as predicted---though even at the highest coarsening we tested the gain in efficiency is relatively small.
 In contrast, our proposed moves approach scales quite well with this refinement, requiring about $12\%$  and $\approx 5\%$ of the number of column operations as vineyards and $\mathrm{pHcol}$, respectively.
@@
 \appendix
-\section*{Appendix}
+\section{Appendix}
 
@@
 \subsubsection*{Move Algorithms}
+Although conceptually similar, there is an asymmetry between \emph{MoveRight} and \emph{MoveLeft}.
+Moving a simplex upwards in the filtration requires removing non-zero entries along several columns of a particular row in $V$ so that the corresponding permutation does not render $V$ non-upper triangular.
+The key insight of the algorithm presented in~\cite{busaryev2010tracking} is that $R$ can actually be maintained in all but one column during this procedure (by employing the \emph{donor} column).
+In contrast, moving a simplex to an earlier time in the filtration requires removing non-zero entries along several rows of a particular column of $V$.
+As before, $R$ stays reduced during this cancellation procedure in all but one column, however the subsequent permutation to $R$ requires reducing a pair of columns which may cascade into a larger chain of column operations to keep $R$ reduced.
+Since these operations always occur in a left-to-right fashion, its not immediately clear how to apply a donor column kind of concept.
+Fortunately, we can still separate  these two stages.
+
 \begin{algorithm}[!htb]
@@
 \end{algorithm}
+We recall an important claim given in~\cite{busaryev2010tracking} on the effect that move operations have on the status of simplices in the pairing.
+Recall from section~\ref{sec:background} that simplices which create new homology classes are called \emph{creators} and simplices that destroy homology classes are called \emph{destroyers}.
+%\begin{proposition}[Locality of moves~\cite{busaryev2010tracking}]
+%	Other than the simplex $\sigma_i$ being moved, $\mathrm{MoveRight}(i,j)$ cannot change a destroyer to a creator, $\mathrm{MoveLeft}(i,j)$ cannot change a creator to destroyer, and neither move operation changes the status of any simplex in $K$ outside of the interval $[i,j]$.
+%\end{proposition}
+\noindent The effect of the movement on intermediate simplices depends on the direction of the movement.
+If $i < j$ (respectively, $j < i$), all simplices at positions $k \in [i+1\;:\;j] $ are shifted down (respectively, up) by $1$.
 

### 2023-01-26 move_schedules.tex
Improved Remark on boissonatt bound

@@
 \noindent Here we consider sorting by moves. Phrased using permutations, a \emph{move operation} $m_{ij}$ that moves $i$ to $j$ in $[m]$, for $i < j$, corresponds to the permutation: 
-\begin{equation}
+\begin{equation}\label{eq:move_perm}
 \setlength\arraycolsep{1.8pt}
@@
 Let $K_0$ and $ K_1$ be filtrations of a fixed simplicial complex $K$ with $\lvert K \rvert = m$ simplices.
-Then, the smallest size any move schedule $S^*$ transforming $K_0$ into $K_1$ can have has cardinality: 
+Then, the smallest size any move schedule $S^*$ transforming $K_0$ into $K_1$ is given by: 
 $$ d = m - \lvert \mathrm{LCS}(K_0, K_1) \rvert $$ 
-Moreover, any $S^*$ of minimal size can be computed in $O(m \log \log m)$ time. 
+Moreover, the size of $S^\ast$ can be determined in $O(m \log \log m)$ time. 
 \end{proposition}
@@
 \begin{corollary}\label{cor:expectation}
-If $K_0, K_1$ are random distinct filtrations each of size $m$, then $d \sim m - \sqrt{m}$ in expectation, with probability $1$ as $m \to \infty$.
+If $K, K'$ are random distinct filtrations each of size $m$, then $d \sim m - \sqrt{m}$ in expectation, with probability $1$ as $m \to \infty$.
 \end{corollary}
@@
 \begin{remark}
-\noindent It's worth noting that the quantity from Corollary~\ref{cor:expectation} captures the size of $S^\ast$ under worst-case conditions, i.e.
-between pairs of uniformly sampled permutations, as opposed to uniformly sampled filtrations.
-Simplicial filtrations have more structure due to the subface relation that must be respected, so one might wonder if this structure reduces the the size of $d$ even further.
-However, Boissonnat~\cite{boissonnat2018efficient} has shown that for any $k$-dimensional simplicial complex $K$ with $m$ simplices, the number of distinct filtrations built from $K$ is \emph{at least} $\floor{\frac{t+1}{k+1}}^m$, where $t$ is the number of distinct filtration values associated with the complex $K$.
-In the case where $t \sim O(m)$ and $k << m$ fixed, this lower bound grows similarly to $m!$---thus,  $d \approx n - \sqrt{n}$ is not too pessimistic a value to expect when the pair of filtrations are no more similar than two random permutations.
+\noindent Note the quantity from Corollary~\ref{cor:expectation} captures the size of $S^\ast$ between pairs of uniformly sampled permutations, as opposed to uniformly sampled filtrations, which have more structure.
+%One may wonder if filtrations would exhibit a smaller upper bound due to the subface relation that must be respected.
+However, Boissonnat~\cite{boissonnat2018efficient} prove the number of distinct filtrations built from $k$-dimensional simplicial complex $K$ with $m$ simplices and $t$ distinct filtration values is \emph{at least} $\floor{\frac{t+1}{k+1}}^m$.
+Since this bound grows similarly to $m!$ when $t \sim O(m)$ and $k << m$ fixed, $d \approx n - \sqrt{n}$ is not too pessimistic a bound between random filtrations.     
+\end{remark}
  In practice, when one has a time-varying filtration and the sampling points are relatively close [in time], the LCS between adjacent filtrations is expected to be much larger, shrinking $d$ substantially.
 For example, the average size of the LCS across the 10 filtrations from Section~\ref{sec:motivation} each having $m=417$ simplices sampled at evenly spaced time points was $343$, implying $d \approx 70$ permutations needed on average to update the decomposition at adjacent time points.
-\end{remark}
 
@@
   \begin{theorem}\label{prop:sparsity_result}
-   Given a pair of simplicial filtrations $K, K'$ and a decomposition $R = DV$ of $K$, algorithm~\ref{alg:schedule} produces a sequence $\mathcal{S} = \left( s_1, s_2, \dots, s_d \right)$ of cyclic permutations $s_k = (i_k, j_k)$ of size $\lvert \mathcal{S} \rvert = m - \lvert \mathrm{LCS}(K, K') \rvert$, satisfying:
+   Given a pair of distinct filtrations $K, K'$ of a simplicial complex $S$ and a decomposition $R = DV$ of $K$, the size of a minimal sequence $\mathcal{S} = \left( s_1, s_2, \dots, s_d \right)$ of cyclic permutations $s_k = (i_k, j_k)$ satisfying~\eqref{eq:move_perm} and:
    	\begin{equation}\label{eq:rv_seq}
@@
    	\end{equation}
-	in $O(m \log \log m)$ time and $O(m)$ space. Moreover, if $i_k < j_k$ for all $k \in [d]$, then~\eqref{eq:rv_seq} can be computed using $O(d\kappa)$ column operations, where:	  
+   	can be determined in $O(m \log \log m)$ time and $O(m)$ space. 
+   	Moreover, if $i_k < j_k$ for all $k \in [d]$, then computing~\eqref{eq:rv_seq} requires $O(d\kappa)$ column operations, where:	  
 	$$  \quad \kappa = \sum\limits_{k = 1}^d \, \lvert \mathbb{I}_{k}\rvert + \lvert \mathbb{J}_{k}\rvert $$
 	% where $\lvert \mathbb{I}_{(k)} \rvert$ and $\lvert \mathbb{J}_{(k)} \rvert$ are the computed quantities in lines (2) and (3) in algorithm~\ref{alg:mr} (here we assume if $i_\ast < j_\ast$).
-	where $\lvert \mathbb{I}_{k} \rvert$ denotes the number of non-zero entries $V_k[i_k, i_k:j_k]$ and $\lvert \mathbb{J}_{k}\rvert$ are the number of columns satisfying $\mathrm{low}_{R_k} \in [i_k:j_k]$ and $\mathrm{row}_{R_k}(i_k) \neq 0$ and $d = m - \lvert \mathrm{LCS}(K, K') \rvert$.
+	and where $\lvert \mathbb{I}_{k} \rvert$ denotes the number of non-zero entries $V_k[i_k, i_k:j_k]$, $\lvert \mathbb{J}_{k}\rvert$ the number of columns satisfying $\mathrm{low}_{R_k} \in [i_k:j_k]$ and $\mathrm{row}_{R_k}(i_k) \neq 0$, and $d = m - \lvert \mathrm{LCS}(K, K') \rvert$.
 \end{theorem}
@@
 \subsubsection{LCS-Sort}
+
  \algnewcommand{\IfThenElse}[3]{% \IfThenElse{<if>}{<then>}{<else>}

### 2023-01-27 move_schedules.tex
Added propisition from Busaryev. Revising section 3 and adding new figures is all that needs to be done now

@@
 % \usepackage{algorithm}
-
+\usepackage{bbm}
+\usepackage{dsfont}
 \usepackage{array}
@@
 
-\subsection{Moves}\label{sec:moves} Busaryev et al.~\cite{busaryev2010tracking} introduced an algorithm which maintains an $R = D V$ decomposition under \emph{move operations}.
+\subsection{Moves}\label{sec:moves} Originally developed to accelerate tracking generators with temporal coherence, Busaryev et al.~\cite{busaryev2010tracking} introduced an extension of the vineyards algorithm which maintains a $R = D V$ decomposition under \emph{move operations}.
 A move operation $\mathrm{Move}(i,j)$ is a set of rules for maintaining a valid decomposition under the application of a permutation $P$ that moves a simplex $\sigma_i$ at position $i$ to position $j$.
-If $j = i \pm 1$, this operation is an adjacent transposition, and thus in some sense Busaryev's move algorithm is a generalization of vineyards.
-
-For completeness, we recapitulate the moves algorithm from~\cite{busaryev2010tracking}. 
-Let $K$ denote a filtration of size $m = \lvert K \rvert$ and $R = DV = [ r_1, r_2, \dots, r_m]$ its decomposition, where $r_j$ ($v_j$) will denote the $j$-th column of $R$ ($V$).
-By definition, if $r_j = D v_j = 0$, then the inclusion $K_{j-1} \hookrightarrow K_{j}$ introduced a new cycle and homology class of dimension $p = \mathrm{dim}(\sigma_j)$ whose generator is encoded by the non-zero coefficients $(\alpha_i^{(j)}, \alpha_{i+1}^{(j)}, \dots, \alpha_j^{(j)})$ via the combination:
+If $j = i \pm 1$, this operation is an adjacent transposition, and thus in some sense moves generalizes vineyards.
+However, despite being conceptually equivalent to performing a sequence of contiguous transpositions, the move operation presented by Busaryev exhibits several attractive qualities distinct from the vineyards approach that warrants further study.
+
+For completeness, we recapitulate the motivation of the moves algorithm from~\cite{busaryev2010tracking}. 
+Let $K$ denote a filtration of size $m = \lvert K \rvert$ and $R = DV$ its decomposition, where $R = [r_1, r_2, \dots, r_m]$ and $V = [v_1, v_2, \dots, v_m]$ denote the columns of $R$ and $V$, respectively.
+By definition, if $r_j = D v_j = 0$, then the inclusion $K_{j-1} \hookrightarrow K_{j}$ introduced a new cycle whose generator is given by:
+%the non-zero coefficients $(\alpha_i^{(j)}, \alpha_{i+1}^{(j)}, \dots, \alpha_j^{(j)})$ via the combination: 
 \begin{equation}
@@
 Now, consider the permutation $P$ that moves a simplex $\sigma_i$ in $K$ to position $j$, shifting all intermediate simplices $\sigma_{i+1}, \dots, \sigma_{j}$ down by one ($i < j$). 
-To perform this shift, these non-zero coefficients $\alpha_i^{(i)}, \alpha_{i}^{(i+1)}, \dots, \alpha_i^{(j)}$ encoded in the boundary chains $r_i, r_{i+1}, \dots, r_j$ must be set to zero, otherwise $\partial(\sigma_i)$ contributes to earlier simplices in the filtration.
-In matrix form, these coefficients are recorded by $v_j$, and thus setting $\alpha_i^k = 0$ for all $k \in [i,j]$ amounts to setting $v_i(i), v_{i+1}(i), \dots, v_{j}(i)$ to $0$.
-
-Computationally, setting $v_i(i) = v_{i+1}(i) = \dots = v_{j}(i) = 0$ with column operations $\ast$ ensures $PV\ast$ is upper-triangular, thus maintaining invariant I2.
-To maintain invariant I1, these operations induce a corresponding $R' = P R \ast P^T$, which may be unreduced (breaking invariant I3).
-We could reducing $R'$ with an additional $k$ operations, however $k$ may scale quadratically in the interval $\lvert i - j\rvert$, rendering the efficiency of this approach at best that of vineyards and at worse that of the standard algorithm.
+To perform this shift, the non-zero coefficients $\alpha_{i}^{(i+1)}, \dots, \alpha_i^{(j)}$ encoded in the boundary chains $r_{i+1}, \dots, r_j$ must be set to zero, otherwise $\partial(\sigma_i)$ contributes to earlier simplices in the filtration.
+As these coefficients are recorded by $v_j$, setting $\alpha_i^k = 0$  amounts to setting $v_k(i) = 0$ for all $k \in [i+1,j]$---algorithmically, these manifest as column operations $\ast$ used to ensure $PV\ast$ is upper-triangular, thus maintaining invariant I2.
+Notice that, to maintain invariant I1, these operations induce a potentially unreduced $R' = P R \ast P^T$, breaking invariant I3.
+We could reduce $R'$ with an additional $k$ operations, however $k$ may scale quadratically in the interval $\lvert i - j\rvert$, rendering the efficiency of this approach at best that of vineyards and at worse that of the standard algorithm.
+
 To bypass this difficulty, Busaryev et al. observed that since the initial $R$ is reduced, if it contains $s$ pivot entries in the columns $[i,j]$ of $R$, then $R'$ must also have $s$ pivots. 
-In particular, if during the cancellation  of $v_i(i) = v_{i+1}(i) = 0$ the column $r_{i+1} \mapsto r_{i+1}'$ becomes unreduced, then the pivot $\mathrm{low}_R(i+1)$ becomes a \emph{free}, possibly becoming a pivot again in $r_{i+2}', \dots, r_j'$.
-Thus, if we save $r_{i+1}$ prior to modification to a \emph{donor column}, we may re-use or \emph{donate} its pivot entry to one of $r_{i+2}, \dots, r_j$. 
-This process is repeated at most $j - i - 1$ times---by construction, the matrix $R'$ stays reduced during the movement in all except possibly for the $i$-th column, and since these operations are left-to-right $V'$ will be upper-triangular.
-%We may achieve this via left-to-right column operations, as before, however this may leave several chains in $R$ unreduced. 
-%To ameliorate this, observe that since the initial set of chains $r_i, r_{i+1}, \dots, r_j$ are already reduced, their pivots are unique---thus, we should expect these entries. 
-%setting any individual $\alpha_i^(k) = 0$ for some $i < k \leq j$ has the effect of changing its pivot entry
-% participates in a linear combination of another 
-%By definition, the column $r_i$ 
+In particular, if during the cancellation  of $v_{i+1}(i) = v_{i+2}(i) = 0$ the column $r_{i+2} \mapsto r_{i+2}'$ becomes unreduced, then the pivot $\mathrm{low}_R(i+2)$ becomes a \emph{free}, possibly becoming a pivot again in $r_{i+3}', \dots, r_j'$.
+Thus, if $r_{i+2}$ is copied prior to modification to a \emph{donor column}, it may re-use or \emph{donate} its pivot entry to a later column $r_{i+3}, \dots, r_j$. 
+Repeating this process at most $j - i - 1$ times ensures $R'$ stays reduced in all except possibly for the $i$-th column by the end of the movement---and since the $k$-th such operation simultaneously sets $v_{k}(i) = 0$ without creating non-zeros at indices $v_{k}(j)$ for $j > k$, $V'$ retains its upper-triangularity.
 \\
@@
 %\end{remark}
-Though similar to vineyards, move operations confer a few additional advantages. We note two convenient properties of moves:
+
+As shown by example 2.3, move updates are cheaper than vineyard updates when $j > i+1$. More formally, we recall the proposition shown in~\cite{busaryev2010tracking}: 
+\begin{proposition}[Busaryev et al.~\cite{busaryev2010tracking}]
+	Let $K$ denote a filtration with $n$ simplices of dimension $p-1$, $p$, and $p+1$, respectively, and let $R = DV$ denote its associated decomposition.
+Then the operation $\mathrm{MoveRight}(i,j)$ constructs a valid decomposition $R' = D'V'$ in $O((s+t)n)$ time, where $s,t$ are given by:
+	$$ s = \sum\limits_{l=i+1}^j \mathds{1}\left(\, v_l(i) \neq 0 \, \right), \quad t = \sum\limits_{l=1}^m \mathds{1}\left(\, \mathrm{low}_R(l) \in [i,j] \text{ and } r_l(i) \neq 0 \, \right)$$
+	Moreover, the quantity $k = s + t$ satisfies $k \leq 2(j - i)$. 
+\end{proposition}
+\noindent Though similar to vineyards, move operations confer additional advantages, including the two computationally attractive properties given below:
 \\	
@@
  	\item[M2:] $R = D V$ is \underline{not} guaranteed to be valid during the movement of $\sigma_i $ to $ \sigma_j$.
+ 	\item[M3:] At most $O(m)$ moves are needed to reindex $K \mapsto K'$ 
  \end{enumerate} 
@@
 In this way, we interpret move operations as making a tradeoff in granularity: whereas a sequence of adjacent transpositions $(i, i{+}1), (i{+}1, i{+}2), \dots, (j{-}1, j)$ generates $\lvert i - j \rvert $ valid decompositions in vineyards, an equivalent move operation $\mathrm{Move}(i,j)$ generates only one.
-Indeed, the number of intermediate persistence diagrams is bounded above by $O(m^2)$ using vineyards if each pair of simplices $\sigma_i, \sigma_j \in K$ switches relative ordering at most once in a given schedule of transpositions, whereas with move operations this bound clearly is at most $O(m)$.
-
-As a final note, we remark that the combination of \emph{MoveRight} and \emph{MoveLeft} enable the dynamic addition or deletion of simplices to the underlying complex $K$ (in fact this was the motivation for its conception).
-In particular, given $K$ and its decomposition $R = DV$, obtaining the decomposition $R' = D' V'$ of $K' = K \cup \sigma$ can be achieved by appending its requisite elementary chains to $D$ and $V$, reducing them, and then calling \emph{MoveLeft}$(m+1, i)$, where $i = f'(\sigma)$.
-Dually, deleting a simplex $\sigma_i$ may be achieved via \emph{MoveRight} by moving $i$-th to the end and dropping them.
+Indeed, Property M3 directly follows from this fact, as one may simply move each simplex $\sigma \in K$ into its order in $K'$ via insertion sort.
+Note that the number of valid decomposition produced by vineyards is bounded above by $O(m^2)$, since each pair of simplices $\sigma_i, \sigma_j \in K$ switches its relative ordering at most once is the homotopy assumptions from~\ref{sec:vineyards} hold.
+
+As a final note, we remark the combination of \emph{MoveRight} and \emph{MoveLeft} enable the addition or deletion of simplices to the underlying complex.
+In particular, given $K$ and its decomposition $R = DV$, obtaining a valid decomposition $R' = D' V'$ of $K' = K \cup \sigma$ can be achieved by appending its requisite elementary chains to $D$ and $V$, reducing them, and then executing \emph{MoveLeft}$(m+1, i)$, where $i = f'(\sigma)$.
+Dually, deleting a simplex $\sigma_i$ may be achieved via \emph{MoveRight} by moving $i$-th to the end of the decomposition and dropping the corresponding columns.
 % with a logarithmic loss in efficiency, although executing the movement has a higher constant factor.
@@
 \begin{proposition}[Schedule Size]\label{prop:schedule_size}
-Let $K_0$ and $ K_1$ be filtrations of a fixed simplicial complex $K$ with $\lvert K \rvert = m$ simplices.
-Then, the smallest size any move schedule $S^*$ transforming $K_0$ into $K_1$ is given by: 
-$$ d = m - \lvert \mathrm{LCS}(K_0, K_1) \rvert $$ 
+Let $K, K'$ denote distinct filtrations of a simplicial complex $S$ of size $\lvert S\rvert = m$.
+Then, the smallest size any move schedule $S^*$ transforming $K$ into $K'$ is given by: 
+$$ d = m - \lvert \mathrm{LCS}(K, K') \rvert $$ 
 Moreover, the size of $S^\ast$ can be determined in $O(m \log \log m)$ time. 
@@
 \]
-Conceptually, left-invariance implies that the edit distance between any pair of permutations $p,q$ is invariant under an arbitrary relabeling $p,q$--as long as the relabeling is consistent.
+Conceptually, left-invariance implies that the edit distance between any pair of permutations $p,q$ is invariant under an arbitrary relabeling of $p,q$--as long as the relabeling is consistent.
 Thus, the following identity always holds:
@@
 where $\iota = [m]$, the identity permutation.
-Suppose we are given permutations $p = K_0, q = K_1$, and we seek to compute $\mathrm{LCS}(p, q)$.
+Suppose we are given two permutations $p, q \in S_n$ and we seek to compute $\mathrm{LCS}(p, q)$.
 Consider the permutation $p' = q^{-1} \circ p$.
@@
 \begin{corollary}\label{cor:expectation}
-If $K, K'$ are random distinct filtrations each of size $m$, then $d \sim m - \sqrt{m}$ in expectation, with probability $1$ as $m \to \infty$.
+If $K, K'$ are random filtrations of a common simplicial complex $S$ of size $m$.
+Then the size of longest common subsequence of simplices between $K,K'$ is no larger than $m - \sqrt{m}$ in expectation, with probability $1$ as $m \to \infty$.
 \end{corollary}
@@
 	$$ \mathbb{E} \, L(p) = \ell_m = \frac{1}{m!} \sum\limits_{p \in S_m} L(p)$$
-	The history of the mathematics of estimating this quantity dates back at least 50 years, and is sometimes called the \emph{Ulam-Hammersley} problem.
-The seminal work by Baik et al.~\cite{baik1999distribution} established that as $m \to \infty$:
+	A large body of work dates back at least 50 years has focused on estimating this quantity, which is sometimes called the \emph{Ulam-Hammersley} problem.
+Seminal work by Baik et al.~\cite{baik1999distribution} established that as $m \to \infty$:
 	$$ \displaystyle \ell_m = 2 \sqrt{m} + c m^{1/6} + o(m^{1/6}) $$
-where $c = -1.77108...$. As a result, we have: 
+where $c = -1.77108...$. Moreover, letting $m \to \infty$, we have: 
 $$ \frac{\ell_m}{\sqrt{m}} \to 2\rlap{\quad \text{as } $m \to \infty$} $$	 
@@
 Such a generalization has appeared naturally in many application contexts, showing potential as a tool for exploratory data analysis~\cite{lesnick2012multidimensional}.
-Indeed, one of the often noted drawbacks of persistence is its instability with respect to strong outliers, which can obscure the detection of significant topological structures~\cite{buchet2015topological}.
-By filtering the data by density in tandem with the e.g. the ambient metric, multidimensional persistence in theory enables practitioner to handle outliers directly.
-
-Unfortunately, unlike one-parameter persistence, there is no complete discrete invariant for multi-parameter persistence.
-Instead, Lesnick et al~\cite{lesnick2015interactive} associate a variety of incomplete invariants to 2-parameter persistence modules: we focus here on the \emph{fibered barcode} invariant, defined as follows:
+Indeed, one of the drawbacks of persistence is its instability with respect to strong outliers, which can obscure the detection of significant topological structures~\cite{buchet2015topological}.
+One exemplary use case of multi-parameter persistence is to detect these strong outliers by filtering the data with respect to both density and the associated metric.
+
+Unfortunately, unlike the one-parameter case, there is no complete discrete invariant for multi-parameter persistence.
+Circumventing this, Lesnick et al~\cite{lesnick2015interactive} associate a variety of incomplete invariants to 2-parameter persistence modules; we focus here on the \emph{fibered barcode} invariant, defined as follows:
 \begin{definition}[Fibered barcode]
@@
 Utilizing the equivalence between the rank and fibered barcode invariants, Lesnick and Wright~\cite{lesnick2015interactive} developed an elegant way of computing $\mathcal{B}(M)$ via a reparameterization using standard point-line duality.
-This clever technique effectively reduces the fibered barcode computation to a sequence of 1-D barcode computations at ``template points'' lying within the 2-cells of particular planar subdivision $\mathcal{A}(M)$ of the half-plane $[0, \infty) \times \mathbb{R}$.
-This subdivision may be obtained via arrangement of ``critical lines'' induced by the bigraded Betti numbers $\beta(M)$ of $M$.
+This clever technique effectively reduces the fibered barcode computation to a sequence of 1-D barcode computations at ``template points'' lying within the 2-cells of a particular planar subdivision $\mathcal{A}(M)$ of the half-plane $[0, \infty) \times \mathbb{R}$.
+This particular subdivision is induced by the arrangement of ``critical lines'' derived by the bigraded Betti numbers $\beta(M)$ of $M$.
 As the barcode of one template point $\mathcal{T}_e$ at the 2-cell $e \in \mathcal{A}(M)$ may be computed efficiently by re-using information from an adjacent template point $\mathcal{T}_{e'}$,~\cite{lesnick2015interactive} observed that computing the barcodes of all such template points (and thus, $\mathcal{B}(M)$) may be reduced to ordering the 2-cells in $\mathcal{A}(M)$ along a Eulerian path traversing the dual graph of $\mathcal{A}(M)$.
@@
 Since the number of 2-cells in $\mathcal{A}(M)$ is on the order $O(\kappa^2)$, and $\kappa$ itself is on the order of $O(m^2)$ in the worst case, the worse-case complexity of the barcode templates computation $O(m^5)$---this is both the highest complexity and most time-intensive sub-procedure the RIVET software~\cite{rivet} depends on.
-Despite this significant barrier, there is room for optimism: in practice, the external stability result from~\cite{landi2014rank} justifies the use of a coarsening procedure which approximates the module $M$ with a smaller module $M'$ via a grid-like reduction, enabling practitioners to restrict the size of $\kappa$ to a relatively small constant.
+Despite this significant complexity barrier, there is room for optimism: in practice, the external stability result from~\cite{landi2014rank} justifies the use of a coarsening procedure which approximates the module $M$ with a smaller module $M'$ via a grid-like reduction, enabling practitioners to restrict the size of $\kappa$ to a relatively small constant.
 This in-turn dramatically reduces the size of $\mathcal{A}(M)$ and thus the number of barcode templates to compute. 
-Moreover, the ordering of barcode templates given by the dual graph traversal implies that adjacent  templates should be relatively similar---so long as $\kappa$ is not too small---suggesting barcode computations may be productively shared between adjacent templates.
+Moreover, the ordering of barcode templates given by the dual graph traversal implies that adjacent  template points should be relatively close---so long as $\kappa$ is not too small---suggesting adjacent templates may productively share computations due to the high similarity of their associated filtrations.
 %Intuitively, similar barcode templates should be easier to compute in sequence than independently. 

### 2023-01-27 move_schedules.tex
Cleaning section 3

@@
   \textbf{ Decomposition Invariants:}
- \begin{enumerate}[labelsep=3pt, topsep=2pt,itemsep=-0.25ex,parsep=1.0ex]
+ \begin{enumerate}[labelsep=3pt, topsep=2pt,itemsep=-0.25ex,parsep=1.0ex]\label{inv:decomposition}
  	\item[I1.] $R = D V$ where $D$ is the boundary matrix of the filtration $K$
@@
 Indeed, Property M3 directly follows from this fact, as one may simply move each simplex $\sigma \in K$ into its order in $K'$ via insertion sort.
-Note that the number of valid decomposition produced by vineyards is bounded above by $O(m^2)$, since each pair of simplices $\sigma_i, \sigma_j \in K$ switches its relative ordering at most once is the homotopy assumptions from~\ref{sec:vineyards} hold.
+Note that the number of valid decomposition produced by vineyards is bounded above by $O(m^2)$ if each pair of simplices $\sigma_i, \sigma_j \in K$ switches its relative ordering at most once during the interpolation from $K$ to $K'$.
 
@@
 As before, we assume as input  a discrete $1$-parameter family of filtrations $\mathcal{K} = (K_1, K_2, \dots, K_n)$ of an abstract simplicial complex $K$ with $\lvert K \rvert =m$, and the goal being to compute their persistence diagrams.
-Fix reindexing  bijections $f_i : K_i \to K_{i+1}$. 
-Each $f_i$ induces a bijection $f_i^\ast : [m] \to [m]$, or equivalently, a permutation of the index set $[m]$.
-For each pair of filtrations $(K_i, K_{i+1})$, we assign the simplices of $K_i$ labels from index set $[m]$ in filtration order, relabeling $K_{i+1}$ accordingly using $f_i$.
+Fix reindexing bijections $\rho_i : K_i \to K_{i+1}$. 
+Each $\rho_i$ induces a bijection $\rho_i^\ast : [m] \to [m]$, or equivalently, a permutation of the index set $[m]$.
+For each pair of filtrations $(K_i, K_{i+1})$, we assign the simplices of $K_i$ labels from index set $[m]$ in filtration order, relabeling $K_{i+1}$ accordingly using $\rho_i$.
 As $p$ is now a strictly increasing sequence, we may compute the longest increasing subsequence $\mathrm{LIS}(q)$ and use this subsequence to recover a longest common subsequence (LCS) between $(K_i, K_{i+1})$, which we denote later with $\mathrm{LCS}(K_i, K_{i+1})$.
@@
 
-\subsubsection{The Vineyards Setting}\label{sec:continuous_setting}
-In the vineyards (i.e., continuous time) case, we are given a homotopy $F : K \times [0,1] \to \mathbb{R}$ interpolating between  a pair of filtrations $K_0, K_1$ of $K$, with simplices  ordered according to $f_0 = F(\cdot, 0)$ and $f_1 = F(\cdot, 1)$, respectively.
-The choice of homotopy $F$ completely determines the number of adjacent transpositions that transform $K_0$ into $K_1$ via $F$, so to get a meaningful bound we must know something about $F$. 
-If we assume that the curves $\tau \mapsto F(\cdot, \tau)$  are in general position in $[0,1]\times \mathbb{R}$, and that each pair of curves cross  at most once, then the set of adjacent transpositions $S_F$ that $F$ decomposes into is given by the pairs of simplices in $K$ with changes in their relative ordering:
+\subsubsection{Continuous setting}\label{sec:continuous_setting}
+In the continuous time case (i.e., vineyards), we are given a homotopy $F : K \times [0,1] \to \mathbb{R}$ interpolating between  a pair of filtrations $K, K'$ from some underlying complex $S$, with simplices ordered according to $f = F(\cdot, 0)$ and $f' = F(\cdot, 1)$, respectively.
+As the choice of homotopy $F$ determines the number of transpositions that transform $K$ into $K$ via $F$, bounding this number requires assumptions on $F$. 
+If we assume the curves $\tau \mapsto F(\cdot, \tau)$ are in general position in $[0,1]\times \mathbb{R}$ and that each pair of curves cross  at most once, then the  functions $\tau\mapsto F(\cdot, \tau)$ belong to a class of $x$-monotone curves called \emph{pseudo-segments}.
+%, i.e. parameterized curves whose behavior with respect to a certain restricted set of geometric predicates is invariant. 
+This family includes the straight-line homotopy $F(\sigma, \tau) = (1 - \tau) f(\sigma) + \tau f'(\sigma)$,  studied in the original vineyards paper~\cite{cohen2006vines}. 
+Detecting all $k$ intersections of $m$ pseudo-segments is a well-studied problem in computational geometry that can be optimally solved in output-sensitive $O(m \log m + k)$ time by several algorithms~\cite{boissonnat2000efficient}, where $k$ is the output-sensitive term.
+
+If one has two distinct filtrations $K,K'$ both derived from the same underlying complex $S$ of size $m = \lvert S \rvert$, observe the simplexwise representations of both $K$ and $K'$ can be thought of as elements of the symmetric group $S_m$.
+Moreover, as the application of permutations to $K$ implies additional elementary column operations to the underlying matrix decomposition, we think of the action of any set of such permutations on the initial decomposition $R=DV$ of $K$ as a \emph{schedule} with respect to the pair $K,K'$.
+We formalize this with a definition.
+\begin{definition}[Schedule]\label{def:schedule}
+Given a pair of filtrations $K, K'$ from an underlying complex $S$ and a fixed bijection $\rho : K \to K'$, let $\rho^\ast : [m] \to [m]$ denote its induced bijection on $S_m$.
+Given a decomposition $R = DV$ of $K$, a \emph{schedule}  $\mathcal{S} = (s_1, s_2, \dots, s_d )$ is a sequence of permutations mapping satisfying:
+	\begin{equation}\label{eq:rv_seq}
+   		R = D_0 V_0 \overset{s_1}{\to} D_1 V_{1} \overset{s_2}{\to} \dots \overset{s_d}{\to} D_d V_{d} = R'
+   	\end{equation}
+   	where, for each $i \in [d]$, $R_i = D_i V_i$ is a valid decomposition respecting invariants~\ref{inv:decomposition}, and $R'$ is a valid decomposition of $K'$.
+\end{definition}
+\noindent Let $F : K \times [0,1] \to \mathbb{R}$ denote a homotopy of $x$-monotone curves.
+Then  $F$ decomposes into an ordered set of adjacent transpositions $S_F$ given by pairs of simplices in $K$ that change in their relative ordering between $K$ and $K'$:
 \begin{align*}\label{eq:sf_schedule}
-	S_F &= \{ (i,i+1) \, \mid \, f_0(\sigma_i) < f_0(\sigma_{i+1}), \;  f_1(\sigma_i) > f_1(\sigma_{i+1}), \; i \in [m] \} \\
+	S_F &= \{ (i,i+1) \, \mid \, f(\sigma_i) < f(\sigma_{i+1}), \;  f'(\sigma_i) > f'(\sigma_{i+1}), \; i \in [m] \} \\
 	&= \{ s_1, s_2, \dots, s_k \} 
 \end{align*}
-%We assume, without loss of generality, that the underlying complexes $K_0$ and $K_1$ have the same cardinality. 
-%Since we deal exclusively with essential filtrations, we may think of $K_1$ as simply a reordering of $K_0$ (or vice versa). 
-Let $p = [m]$ and let $q = \mathrm{Im}\,f^\ast$, where $f^\ast : [m] \to [m]$ is the one-to-one correspondence induced by the reindexing bijection $f: K_0 \to K_1$.
+Moreover, let $p = [m]$ and let $q = \mathrm{Im}\,\rho^\ast$, where $\rho^\ast : [m] \to [m]$ is the one-to-one correspondence induced by the reindexing bijection $\rho: K \to K'$.
 Then $S_F$ is given by the inversions between $p$ and $q$:
@@
 \end{equation}
-Geometrically, our assumptions on $F$ imply that the  functions $\tau\mapsto F(\cdot, \tau)$ belong to a class of $x$-monotone curves called \emph{pseudo-segments}.
-%, i.e. parameterized curves whose behavior with respect to a certain restricted set of geometric predicates is invariant. 
-This family includes 
-the straight-line homotopy $F(\sigma, \tau) = (1 - \tau) f_0(\sigma) + \tau f_1(\sigma)$,   studied in the original vineyards paper~\cite{cohen2006vines}.
-Detecting all $k$ intersections of $m$ pseudo-segments is a well-studied problem in computational geometry that can be optimally solved in output-sensitive $O(m \log m + k)$ time by several algorithms~\cite{boissonnat2000efficient}, where $k$ is the output-sensitive term.
-Since $k \sim O(m^2)$ in the worst case, achieved when $f_1 = - f_0$, we have the following upper bound on $\lvert S_F \rvert$
-\begin{equation}\label{eq:sf_m2}
-	\lvert S_F \rvert \approx O(m^2)
-\end{equation}   
- This bound is exhibited in the grayscale image data example in section~\ref{sec:motivation}: each 9x9 patch contains $(81, 208, 128)$ simplices of dimensions $(0, 1, 2)$ yielding $m=417$ simplices in total which, as the homotopy is simulated, generate $\approx 70,\!000$ transpositions.
-The upper bound size on $\lvert S_F \rvert$ is thus $\binom{417}{2} \approx 87,\!000$, which nearly occurs in this case, since the last filtration is approximately the reverse of the filtration at the beginning.
- This quadratic scaling can induce a number of issues in practical implementations of vineyards.
- Indeed, although many transpositions require no column reductions in practice, detecting the existence of non-zero entries at specific positions in the sparse matrix takes $O(\log(m))$ time per transposition.
-
-The ordered set of inversions $(i,j) \in S_F$ can be interpreted as permutations to apply to the decomposition, and hence 
-can be thought of as a \emph{schedule} with respect to $F$. 
+This implies the $\lvert S_F \rvert \sim O(m^2)$ in the worst case, achieved when $f = - f'$, which is exhibited in the grayscale image data example in section~\ref{sec:motivation}: each 9x9 patch contains $(81, 208, 128)$ simplices of dimensions $(0, 1, 2)$ yielding $m=417$ simplices in total.
+As the homotopy is simulated, observe $\approx 70,\!000$ transpositions are generated.
+The upper bound size on $\lvert S_F \rvert$ is thus $\binom{417}{2} \approx 87,\!000$, which nearly occurs in this case, since the last filtration is nearly the reverse of the initial one.
+
 If our goal is to decrease the size of $\lvert S_F \rvert$, one option is to \emph{coarsen} $S_F$ to a new schedule $\widetilde{S}_F$ by collapsing contiguous sequences of adjacent transpositions to moves, via the map:
@@
 \begin{proposition}\label{prop:factor2}
-	Let $R = D V$ be a valid decomposition of size $(m \times m)$, let $C$ denote the number of $O(m)$ operations required to execute schedule $S_F$, and similarly let $L$ be the number of $O(m)$ operations required to execute a coarsened schedule $\widetilde{S}_F$ of $S_F$.
-We have the following relationship between $C$ and $L$:
-	$$ \frac{C}{2} \leq L \leq C$$ 
+	Let $K$ denote a filtration of size $\lvert K \rvert = m$ with valid decomposition $R = D V$, $T$ the maximum number of $O(m)$ operations required to execute schedule $S_F$ using vineyards, and $M$ be the maximum number of $O(m)$ operations required to execute the coarsened schedule $\widetilde{S}_F$ with moves.
+Then:
+	$$ M \leq {\textstyle \frac{1}{2} \, T} $$ 
 \end{proposition}
 \begin{proof} \normalsize
-Consider executing the vineyards algorithm on a given schedule $S_F$.
-	%Consider Algorithms~\ref{alg:tr} and~\ref{alg:mr}, where Algorithm~\ref{alg:tr} is used to execute a given schedule $S$ and Algorithm~\ref{alg:mr} is used to execute its coarsened version $\widetilde{S}$.
-If maps are created providing $O(1)$ access to the low entries needed by lines (5) and (12), and the non-zero entries in each column are sorted according to some fixed order providing $O(\log(m))$ time for lines (5), (11), (17), and (20), then
-	%If we maintain a map locating the position of the lowest entries in $R$ in $O(1)$ time, 
-	The dominant subcomputation for the vineyards algorithm are the column operations---which take $O(m)$ time---since the permutations can be made to the matrix in $O(1)$ time and all other computations take at most $O(\log m)$ time.
-In the vineyards algorithm, there are at most $2$ column operations for a given adjacent transposition per matrix.
-	Thus, for any contiguous sequence of transpositions $(i,i+1), (i+1,i+2), \ldots (j-1,j)$ we require at most $2(\lvert i - j \rvert)$ column operations in both $R$ and $V$, giving a total of $4(\lvert i - j \rvert)$ column operations.
-
-	Now consider using the MoveRight($i$,$j$) for a single move operation, as a replacement for the sequence of transpositions above.
-Here, the dominant cost again are the column operations (line 5), since although permutations take $O(\lvert i - j\rvert)$ time we require only $1$ such permutation per movement.
-	If we allocate an extra $O(m)$ memory space for the donor columns $d_\ast$ prior to the movement, notice that the assignment to and from $d_\ast$ (lines (4), and (7) in RestoreRight of MoveRight in Algorithm~\ref{alg:mr}) can be completed in $O(1)$ time by storing the column memory addresses as pointers and swapping their pointers.
-When $d_\mathrm{low}' < d_\mathrm{low}$, instead of copying $\mathrm{col}_\ast(k)$ to $d_\ast'$, which takes $O(m)$ time, we swap their column pointers in $O(1)$ and proceed by modifying the donor column $d_\ast$ with the necessary column operations.
-After the operations, $d_\ast$ contains the newly modified column and $\mathrm{col}_\ast(k)$ contains the unmodified donor $d_\ast'$, so we swap them once more in $O(1)$ time.
-Since we require at most $1$ $O(m)$ column operation for each index in $[i: j]$, moving a column from $i$ to $j$ where $i < j$ requires at most $2(\lvert i - j \rvert)$ column operations for both $R$ and $V$.
+First, consider executing the vineyards algorithm on a given schedule $S_F$.
+As there are at most $2$ column operations, any contiguous sequence of transpositions $(i,i+1), (i+1,i+2), \ldots (j-1,j)$ induces at most $2(\lvert i - j \rvert)$ column operations in both $R$ and $V$, giving a total of $4(\lvert i - j \rvert)$ column operations.
+
+Now consider a single MoveRight($i$,$j$) outlined in Algorithm~\ref{alg:mr}. Here, the dominant cost again are the column operations (line 5). 
+Though we need an extra $O(m)$ storage allocation for the donor columns $d_\ast$ prior to the movement, notice that the assignment to and from $d_\ast$ (lines (4), and (7) in RestoreRight of MoveRight in Algorithm~\ref{alg:mr}) can be completed in $O(1)$ time via a pointer swapping argument.
+That is, when $d_\mathrm{low}' < d_\mathrm{low}$, instead of copying $\mathrm{col}_\ast(k)$ to $d_\ast'$---which takes $O(m)$ time---we swap their column pointers in $O(1)$ before proceeding with the necessary column operations.
+After the movement, $d_\ast$ contains the newly modified column and $\mathrm{col}_\ast(k)$ contains the unmodified donor $d_\ast'$, so we swap them once more in $O(1)$ time.
+Since we require at most one $O(m)$ column operation for each index in $[i, j]$, moving a column from $i$ to $j$ where $i < j$ requires at most $2(\lvert i - j \rvert)$ column operations for both $R$ and $V$.
 The claimed inequality follows.
+%since although permutations take $O(\lvert i - j\rvert)$ time we require only $1$ such permutation per movement. 
 	%Identifying non-zero entries in the matrix occurs once, which takes $O(\lvert i - j \rvert \log(m))$ time.
 \end{proof}
-In terms of bounds, we have $\lvert \widetilde{S}_F \rvert \leq \lvert S_F \rvert$, and the associated coarsened schedule $\widetilde{S}_F$ can be computed in $O(m)$ time. 
-However, the coarsening depends entirely on the choice of $F$ and the upper bound from~\eqref{eq:sf_m2} remains, as its always possible that there are no contiguous subsequences to collapse.
-If a straight line homotopy is used, the number of collapsible subsequences may depend on the simplices categorization as critical points; transpositions between pairs of simplices which trigger  changes in homology tend to be more expensive.
+\noindent 
+In terms of size bounds, though clearly $\lvert \widetilde{S}_F \rvert \leq \lvert S_F \rvert$ and the associated coarsened $\widetilde{S}_F$ requires just $O(m)$ time to compute, the coarsening depends entirely on the initial choice of $F$---thus the quadratic upper bound remains, as its always possible that there are no contiguous subsequences to collapse.
+As mentioned in~\ref{sec:vineyards}, this quadratic scaling induces a number of issues in the practical implementations of the vineyards algorithm.
+% Indeed, although many transpositions require no column reductions in practice, detecting the existence of non-zero entries at specific positions in the sparse matrix takes $O(\log(m))$ time per transposition.
+
+\subsection{Discrete setting}
+
+%If a straight line homotopy is used, the number of collapsible subsequences may depend on the simplices categorization as critical points; transpositions between pairs of simplices which trigger  changes in homology tend to be more expensive.
 Cohen-Steiner~\cite{cohen2006vines} referred to these as \emph{switches}, see~\cite{edelsbrunner2000topological} for more details.
 
-\subsubsection{The Moves Setting}\label{sec:moves_setting}
-As we mentioned before, vineyards maintains $j-i$ valid decompositions while 
-performing the sequence $(i,i+1), (i+1,i+2), \ldots, (j-1,j)$ of adjacent transpositions. 
-The benefit of move operations is that one need not maintain a valid decomposition during the movement $(i,j)$, only at the end; thus, it is worth investigating whether the vineyard bounds can be improved.
-In the case of vineyards, the bound is clear: if $\kappa = K_\tau(K_0, K_1)$ is the Kendall-$\tau$ distance, then $\Omega(\kappa)$ adjacent transpositions are required to transform $K_0 \mapsto K_1$. 
-In contrast, if we allow move operations in any order, we have $O(m)$ as a trivial upper bound: simply move each simplex in $K_0$ into its position in the filtration given by $K_1$ in the order given by the latter.
-However, it's not immediately clear whether this bound is tight.
+%\subsubsection{The Moves Setting}\label{sec:moves_setting}
+%As we mentioned before, vineyards maintains $j-i$ valid decompositions while 
+%performing the sequence $(i,i+1), (i+1,i+2), \ldots, (j-1,j)$ of adjacent transpositions. 
+%The benefit of move operations is that one need not maintain a valid decomposition during the movement $(i,j)$, only at the end; thus, it is worth investigating whether the vineyard bounds can be improved.
+%In the case of vineyards, the bound is clear: if $\kappa = K_\tau(K_0, K_1)$ is the Kendall-$\tau$ distance, then $\Omega(\kappa)$ adjacent transpositions are required to transform $K_0 \mapsto K_1$.
+
+Contrasting the continuous-time setting, if we discard the use of a homotopy interpolation and allow move operations in any order, we have $\widetilde{S} \sim O(m)$ trivially: simply move each simplex in $K$ into its position in the filtration given by $K'$ in the order given by the latter.
+However, it's not immediately clear whether this bound is tight, or whether this strategy of generating schedules---which we call \emph{naive strategy}---is performant.
+%Indeed, arbitrarily permuting a given filtration $K$: one must respect the face poset.
 
 To establish this formally, we require a few definitions.
-Let $S_m$ denote the symmetric group  of   permutations of  $[m]$ under function composition $\circ$.
-Given two fixed permutations $p, q \in S_m$ and a set $\Sigma \subseteq S_m$,  a common  problem is to find an ordered sequence of permutations $s_1, s_2, \dots, s_d \in \Sigma$ whose composition transforms $p$ into $q$:
+Given two fixed permutations $p, q \in S_m$ and a set $\Sigma \subseteq S_m$,  a common  problem is to find a sequence of permutations $s_1, s_2, \dots, s_d \in \Sigma$ whose composition satisfies:
 	\begin{equation}\label{eq:sorting}
@@
 
-Both the difficulty of obtaining a minimal sorting and the size of $d_\Sigma(p,q)$ vary dramatically with the choice of $\Sigma$. 
-For example, while sorting by transpositions and reversals is NP-hard, and sorting by prefix  transpositions is unknown, there are polynomial time algorithms for sorting by block interchanges, exchanges, and prefix exchanges~\cite{labarre2013lower}.
+Perhaps surprisingly, small changes to set of allowable edit operation $\Sigma$ can dramatically affect both the size of $d_\Sigma(p,q)$ and the difficulty of obtaining a minimal sorting. 
+For example, while sorting by transpositions and reversals is NP-hard and sorting by prefix  transpositions is unknown, there are polynomial time algorithms for sorting by block interchanges, exchanges, and prefix exchanges~\cite{labarre2013lower}.
 Sorting by adjacent transpositions can be achieved in many ways: any sorting algorithm that exchanges two adjacent elements during its execution (e.g.
-bubble sort, insertion sort, merge sort) yields a sorting of size $K_\tau(p, q)$.
-
-In the context of filtered simplicial filtration, we refer to sortings that repair the decomposition as \emph{schedules}. 
-\begin{definition}[Schedule]\label{def:schedule}
-Given a pair of filtrations $K_0, K_1$ of equal size $m$ and a fixed bijection $f : K_0 \to K_{1}$, let $f^\ast : [m] \to [m]$ denote its induced bijection on $S_m$.
-A \emph{schedule}  $\mathcal{S} = \{s_1, s_2, \dots, s_d \}$ is a sequence of permutations mapping $K_0 \mapsto K_1$ that satisfies:
-	\begin{equation}\label{eq:rv_seq}
-   		R = D_0 V_0 \overset{s_1}{\to} D_1 V_{1} \overset{s_2}{\to} \dots \overset{s_d}{\to} D_d V_{d} = R'
-   	\end{equation}
-   	where, for each $i \in [d]$, $R_i = D_i V_i$ is a valid decomposition: $R_i$ is \emph{reduced} and $V_i$ is full rank upper-triangular. 
-\end{definition}
-% Note that one cannot arbitrarily permute a given filtration $K$ and its corresponding decomposition: one must respect the face poset.
-% Given a filtered simplicial complex $K$, denote by $R = D V$ the decomposition of its corresponding boundary matrix $\partial_K$ such that $R$ is reduced and $V$ is upper-triangular.
-% If one has a pair of filtrations $K, K'$ of equal size $m = \lvert K \rvert = \lvert K' \rvert$ and $R = \partial_{K} V$ has been computed, then it may be advantageous to use the information stored in ($R$, $V$) to reduce the computation of $R' = \partial_{K'} V'$.
-Our first result makes this more explicit.
-%   \begin{proposition}\label{prop:preprocessing_time}
-%   Given a decomposition $R = \partial_{K} V$ of a simplicial filtration $K$, algorithm~\ref{alg:schedule} can produce such a sequence $\mathcal{S} = \left( s_1, s_2, \dots, s_d \right)$ of permutations with size $\lvert \mathcal{S} \rvert \sim O(m)$ which induces a sequence of decompositions:
-%   	\begin{equation}\label{eq:rv_seq}
-%   		R = D_0 V_0 \overset{s_1}{\to} D_1 V_{1} \overset{s_2}{\to} \dots \overset{s_d}{\to} D_d V_{d} = R'
-%   	\end{equation}
-%	in $O(m \log \log m)$ time and $O(m)$ memory.
-
-%In the context of performing updates to the $R = D V$ decomposition using moves,  such sequences will form schedules. 
-%There are typically many ways to solve (1), and any solution to (2) necessarily solves (1) and (3).
-
-%Many types of edit distances can be described succinctly. 
-%Let $\beta(i,j,k,l)$ with $1 \leq i < j \leq k < l \leq m$ denote the permutation that exchanges the two closed intervals given by $i, j-1$ and $k,l-1$:
-%\[
-%\setlength\arraycolsep{1.5pt}
-%\scalefont{0.90}{
-%\setcounter{MaxMatrixCols}{20}
-%\left (
-%\begin{array}{*3c|*4c|*4c|*4c|*{5}c } 
-%\cline{4-7} \cline{12-15}  \rule{0pt}{2ex}
-%1 & \cdots & i -1 & i & i + 1 & \cdots & j - 1 & j & j +1 & \cdots & k - 1  & k & k+1 & \cdots & l - 1 & l & l + 1 & \cdots & m \\
-%\cline{4-7} \cline{12-15} 
-%\cline{4-7} \cline{12-15}  \rule{0pt}{2ex}
-%1 & \cdots & i -1 & k & k+1 & \cdots & l - 1 & j & j +1 & \cdots & k - 1  & i & i + 1 & \cdots & j - 1 & l & l + 1 & \cdots & m \\
-%\cline{4-7} \cline{12-15} 
-%\end{array}
-%\right ) 
-%}
-%\]
-%Common edit distances found in the literature expressed as specializations of $\beta$ are given below:
-%\begin{enumerate}
-%	\item (no restriction on $i,j,k,l$) $\implies$ $\beta$ is called a \emph{block interchange}
-%	\item $j = k$ $\implies$ $\beta$ swaps adjacent intervals, called a \emph{transposition}
-%	\item $j = i+1$ and $l = k+1$ $\implies$ $\beta$ swaps two not necessarily adjacent elements, called an \emph{exchange}\footnote{In comparative genomics, exchanging two elements is sometimes called a \emph{super short transposition}.}
-%\end{enumerate}
-\noindent Here we consider sorting by moves. Phrased using permutations, a \emph{move operation} $m_{ij}$ that moves $i$ to $j$ in $[m]$, for $i < j$, corresponds to the permutation: 
+bubble sort, insertion sort) yields a sorting of size $K_\tau(p, q)$.
+
+Here we consider sorting by moves. Phrased using permutations, a \emph{move operation} $m_{ij}$ that moves $i$ to $j$ in $[m]$, for $i < j$, corresponds to the permutation: 
 \begin{equation}\label{eq:move_perm}

### 2023-01-28 move_schedules.tex
Replaced motivating example figure, split them, and editted/streamlined the motivating example section

@@
  \subsection{A Motivating Example}\label{sec:motivation} 
-We describe a simple experiment illustrating why the vineyards algorithm does not always yield an efficient strategy for time-varying applications. 
-Consider a series of grayscale images (i.e. a video) depicting a fixed-width annulus expanding about the center of a $9 \times 9$ grid. 
-Each image contains the same number of pixels, whose intensities vary with time.
-For each image, we build a filtered simplicial complex using the \emph{Freudenthal} triangulation of the plane, with the order of inclusions obtained by lower stars and pixel values.
+We illustrate why the vineyards algorithm does not always yield an efficient strategy for time-varying setting with a simple experiment. 
+Consider a series of grayscale images (i.e. a video) depicting a fixed-width annulus expanding about the center of a $9 \times 9$ grid and its associated sublevel-set filtrations:   
+\begin{center}
+	\vspace{1em}
+	\includegraphics[height=1.15in]{circle_complex.png}
+\end{center}
+Each image contains a fixed number of pixels whose intensities vary with time.
+For each image, we build a fixed-size filtered simplicial complex using the \emph{Freudenthal} triangulation of the plane whose inclusion maps are obtained by lower stars of pixel values.
+%Snapshots of the image data and its corresponding time-varying filtration are shown below Figure~\ref{fig:vineyards}. 
+%\begin{figure*}[h!]
+%\centering	
+%\end{figure*}
+%Note that the size of the filtration is fixed---no simplices are added or removed over time. 
+Two events alter the persistence diagrams over time: the first occurs when the central connected component splits to form a cycle, and the second  when the annulus  splits into four components. 
+From left to right, the Betti numbers of the five evenly spaced `snapshots' of the filtration shown above are: $(\beta_0,\beta_1)  = (1,0), \; (1,1),  \; (1,1), \; (1,1), \; (4,0)$.
+Thus, in this example, only a few persistence diagrams are needed to capture the major changes to the topology. 
 \begin{figure*}[t]
  	\centering
- 	\includegraphics[width=0.45\textwidth,height=1.63in]{circle_vineyards.png}
- 	\includegraphics[width=0.54\textwidth,height=1.63in]{circle_complex.png}
- 	\caption{On the right, the grayscale image data set and corresponding lower star filtrations.
-On the left, the cumulative number of column operations required to compute persistence for this time-varying filtration.
-The  Betti numbers for each of the shown  steps in the filtration, from left to right, are: $(\beta_0,\beta_1, \beta_2)  = (1,0,0), \; (1,1,0),  \; (1,1,0), \; (1,1,0), \; (4,0,0)$.
-Observe that computing the diagrams independently at 10 evenly spaced time points (the green line) both captures the major topological changes and is the most computationally efficient approach shown.}
+ 	\includegraphics[width=0.75\textwidth]{circle_vineyards.png}
+ 	\caption{The cumulative column operations needed to compute persistence across the time-varying filtration of grayscale images.
+Observe 10 independent persistence computations evenly spaced in time (green line) captures the major topological changes and is the most computationally efficient approach shown.}
  	\label{fig:vineyards}
  \end{figure*}
-Snapshots of the image data and its corresponding time-varying filtration is depicted in Figure~\ref{fig:vineyards} (top right and bottom right, respectively), which we use as a baseline for comparing vineyards and the standard reduction algorithm $\texttt{pHcol}$ (Algorithm \ref{alg:reduce}).
-Note that the size of the filtration is fixed---no simplices are ever added or removed over time. 
-There are two events which dramatically alter the persistence diagrams extracted from these filtrations: the first occurs when the central connected component splits to form a cycle, and the second  when the annulus  splits into four components.
-As a consequence, in this example, only a few persistence diagrams are needed to capture the major topological changes over time.
-
-  Consider the situation where one is concerned with capturing these dramatic changes in homology.
-  On the left side of Figure~\ref{fig:vineyards}, we compare the cumulative cost (in total column operations) of various approaches.
-  Since it is unknown a priori at when the persistent pairing function changes, one solution is to discretize the time domain at $n$ evenly spaced points and compute the diagrams at each time point independently.
-  An alternative approach is to construct a (e.g., straight-line) homotopy between  contiguous time-points, and then decompose these homotopies into adjacent transpositions in the filtration order---the vineyards approach.
-The former is often used in practice, but it is only an approximation; the latter is guaranteed to capture all homological changes in persistence that occur in simulating the 1-parameter family.
-  The graph in Figure~\ref{fig:vineyards} depicts two different size approximations (purple, $n=100$ and green, $n=10$), wherein the reduction algorithm (\texttt{pHcol}) was applied at $n$ time points, and two exact strategies (black and blue), which compute all $\approx 7 \times 10^{4}$ diagrams given by the homotopy.
-  As shown in the figure, if one needs to simulate the homotopy exactly, then the vineyards approach is indeed far more efficient than naively applying the reduction algorithm independently at all time points.
-  However, when the discretization of the time domain is coarse enough, the naive approach actually performs less column operations than the vineyards strategy, while still capturing the main events.
-
-  The existence of a time discretization that is more efficient to compute than continually updating the decomposition indicates that the vineyards algorithm must incur some overhead (in terms of column operations) to perform transpositions, even if the persistence diagrams remain unchanged.
-  Even in situations where one is updating a decomposition to a filtration that is  relatively ``close''---e.g.
-any two adjacent filtrations in the bottom right of Figure~\ref{fig:vineyards}---it is still more efficient to apply \texttt{pHcol} directly than  to iteratively update the decomposition (see the case where $n = 10$).
- Moreover, any optimizations to the reduction algorithm (e.g.,  clearing~\cite{chen2011persistent}) would only increase this disparity.
-Since persistence has found many applications in dynamic contexts~\cite{topaz2015topological, xian2020capturing, lesnick2015interactive, kim2020spatiotemporal}, a more efficient alternative to the vineyards is clearly needed.
+
+  We use this data set as a baseline for comparing vineyards and the standard reduction algorithm $\texttt{pHcol}$ (Algorithm \ref{alg:reduce}).
+Suppose a practitioner wanted to know the major homological changes a time-varying filtration encounters over time.
+  Since it is unknown a priori at when the persistent pairing function changes, one solution is to do $n$ independent persistence computations $n$ evenly spaced points in the time domain.
+An alternative approach is to construct a homotopy between contiguous time-points and decompose these homotopies into adjacent transpositions in the filtration order---the vineyards approach.
+ We refer to the former as the \emph{discrete setting}, which is often used in practice, though it is only an approximation; the latter is guaranteed to capture all homological changes in persistence that occur in simulating the 1-parameter family, which we'll refer to as the \emph{continuous setting}.
+
+ The cumulative cost (in total column operations) of these various approaches are shown on the left side of Figure~\ref{fig:vineyards}, wherein the reduction algorithm (\texttt{pHcol}) and the vineyards algorithm are compared.
+Two approximate strategies (green and purple) and two continuous strategies (black and blue) are shown.
+  Note that without knowing where persistence pairing function changes, a continuous strategy must construct all $\approx 7 \times 10^{4}$ diagrams induced by the homotopy. 
+  In this setting, as shown in the figure, the vineyards approach is indeed far more efficient than naively applying the reduction algorithm independently at all time points.
+  However, when the discretization of the time domain is coarse enough, the naive approach actually performs less column operations than vineyards, while still capturing the main events.
+
+  The existence of a time discretization that is more efficient to compute than continually updating the decomposition indicates that the vineyards framework must incur some overhead (in terms of column operations) to maintain the underlying decomposition, even when pairing function is unchanged.
+Indeed, applying \texttt{pHcol} independently between  relatively ``close'' filtrations is more efficient than iteratively update the decomposition, as shown by the case where $n = 10$.
+ Moreover, any optimizations to the reduction algorithm (e.g. clearing~\cite{chen2011persistent}) would only increase this disparity.
+Since persistence has found many applications in dynamic contexts~\cite{topaz2015topological, xian2020capturing, lesnick2015interactive, kim2020spatiotemporal}, a more efficient alternative to vineyards is clearly needed.
 
@@
 \subsection{Main results}
-Given a filtered simplicial complex $K$ with filtration function $f : K \to [m]$, denote by $R = D V$ the decomposition of its corresponding boundary matrix $D$ such that $R$ is reduced and $V$ is upper-triangular (see section~\ref{sec:reduction} for details).
- If one has a pair of filtrations $(K, f)$,  $(K', f')$ of equal size $m = \lvert K \rvert = \lvert K' \rvert$ and $R = D V$ has been computed for $K$, then it may be advantageous to use the information stored in ($R$, $V$) to reduce the computation of $R' = D' V'$.
+Given a simplicial complex $K$ with filtration function $f : K \to [m]$, denote by $R = D V$ the decomposition of its corresponding boundary matrix $D$ such that $R$ is reduced and $V$ is upper-triangular (see section~\ref{sec:reduction} for details).
+ If one has a pair of filtrations $(K, f)$,  $(K, f')$ of size $m = \lvert K \rvert$ and $R = D V$ has been computed for $K$, then it may be advantageous to use the information stored in ($R$, $V$) to reduce the computation of $R' = D' V'$.
 Given a permutation $P$ such that $D' = P^T D P$, such an update scheme has the form: 
@@
 \end{equation}
-where the quantities $\lvert \mathbb{I}_{k} \rvert$ and $\lvert \mathbb{J}_{k} \rvert$ depend on the sparsity of the $V_k$ and $R_k$ matrices, respectively, and $d \sim O(m)$ is a constant that depends on how similar $K$ and $K'$ are.
+where the quantities $\lvert \mathbb{I}_{k} \rvert$ and $\lvert \mathbb{J}_{k} \rvert$ depend on the sparsity of the $V_k$ and $R_k$ matrices, respectively, and $d \sim O(m)$ is a constant that depends on how similar $f$ and $f'$ are.
 The advantage of this result is that it depends explicitly on the sparsity pattern of the decomposition itself, which we take advantage of in Section~\ref{sec:schedule_cost}.
-Moreover, if the vineyards algorithm requires $C$ column operations to obtain~\eqref{eq:rv_seq}, we show (Proposition~\ref{prop:factor2}) that our approach obtains the same sequence using at most $L$ column operations, where:
+Moreover, if the vineyards algorithm requires $C$ column operations to obtain~\eqref{eq:rv_seq}, we show (Proposition~\ref{prop:factor2}) that our approach obtains the same sequence using at most $M$ column operations, where:
 \begin{equation}
-	\tfrac{1}{2} \, C \leq L \leq C
+	M \leq \tfrac{1}{2}  C
 \end{equation}
@@
 The complexity of determining $\mathcal{S}$ depends on the type of permutations considered; many edit distances exhibit tight quadratic growth rates and tend to require at least quadratic time (or worse) to compute~\cite{labarre2013lower}.
-We consider a special set of cyclic permutations whose sequences scale linearly in $m$ for all permutations in $S_m$: surprisingly, we find that the \emph{smallest} such sequence of these permutations which map $K \mapsto K'$ has size $d$ (Proposition~\ref{prop:schedule_size}), where:
+We consider a special set of cyclic permutations whose sequences scale linearly in $m$ for all permutations in $S_m$: surprisingly, we find that the \emph{smallest} such sequence of these permutations which map $f \mapsto f'$ has size $d$ (Proposition~\ref{prop:schedule_size}), where:
 \begin{equation}
-	d = m - \lvert \mathrm{LCS}(K, K') \rvert 
+	d = m - \lvert \mathrm{LCS}(f, f') \rvert 
 \end{equation}
@@
 Suppose one has a family $\{K_i\}_{i\in I}$ of simplicial complexes indexed by a totally ordered   set $I$, and so that for any $i< j \in I$ we have $K_i \subseteq K_j$. 
-There are two index sets of interest in what follows: $\mathbb{R}$ and $[n] = \{ 1, \dots, n\}$.
 Such a family is called a \emph{filtration},
@@
 Any finite filtration may be trivially converted into an simplexwise filtration via a set of \emph{condensing}, \emph{refining}, and \emph{reindexing} maps (see~\cite{bauer2021ripser} for more details).
+Equivalently, we can also define a filtration as a pair $(K, f)$ where $K$ is a simplicial complex and $f : K \to I$ is a \emph{filter function} over a totally ordered index set $I$ satisfying $f(\tau) \leq f(\sigma)$ whenever $\tau \subseteq \sigma$, for any $\tau,\sigma \in K$.
+In this setting, $K_i = \{ \, \sigma \in K :  f(\sigma) \leq i \, \}$.
+Here, we consider two index sets: $[m] = \{ 1, \dots, m\}$ and $\mathbb{R}$.
 Without loss of generality, we exclusively consider simplexwise filtrations but for brevity-sake refer to them simply as filtrations.
@@
  The main output of the reduction algorithm is a matrix decomposition $R = D V$, where the persistence diagram is encoded in $R$, and the generating cycles in the columns of $V$. 
- Given a filtration $K$ with $m$ simplices, one assembles the elementary boundary chains $\partial(\sigma)$ as columns ordered according to the filtration into an $m \times m$  \emph{filtration boundary matrix} $D$.
- It is at times more succinct to restrict to specific sub-matrices $D$ based on the homology dimension $p$, and so we write $D_p$ represent the $d_{p-1} \times d_p$ matrix representing the linear operator $\partial_p(K)$ (the same notation is extended to $R$ and $V$).
+ Given a filtration $(K, f)$ with $m$ simplices, one assembles the elementary boundary chains $\partial(\sigma)$ as columns ordered according to $f$ into a $m \times m$  \emph{filtration boundary matrix} $D$.
+ It is at times more succinct to restrict to specific sub-matrices $D$ based on the homology dimension $p$, and so we write $D_p$ represent the $d_{p-1} \times d_p$ matrix representing the linear operator $\partial_p$ (the same notation is extended to $R$ and $V$).
 % The reduction algorithm computes the persistent homology from $D$ for all dimensions up to $\mathrm{dim}(K) - 1$.
@@
 and whose filtration order, read as inclusions from left to right, is given as $K = (u, v, w, a, b, c)$.
-The reduction algorithm begins by setting $R = D$ and $V = I$, followed by left-to-right column operations until invariant I2 below is satisfied.
+The reduction algorithm begins by setting $R = D$ and $V = I$, followed by left-to-right column operations until invariants I3 is satisfied.
 If the $i$-th column of $R$ is nonzero, then $\mathrm{low}_R(i)$ 
@@
  	\item[I2.] $V$ is full-rank upper-triangular
- 	\item[l3.] $R$ is \emph{reduced}: if both $\mathrm{col}_i(R) \neq 0$ and $\mathrm{col}_j(R) \neq 0$, then $\mathrm{low}_R(i) \neq \mathrm{low}_R(j)$ 
+ 	\item[l3.] $R$ is \emph{reduced}: if $\mathrm{col}_i(R) \neq 0$ and $\mathrm{col}_j(R) \neq 0$, then $\mathrm{low}_R(i) \neq \mathrm{low}_R(j)$ 
  	\end{enumerate} 
  \vspace*{1em}
- If these two invariants are satisfied, then we call the decomposition \emph{valid}. 
+We call the decomposition satisfying these three invariants \emph{valid}. 
 %%\begin{proposition}[Persistent Pairing~\cite{zomorodian2005computing}]
@@
 %\noindent
-Though the matrices $R$ and $V$ are not unique, it has been shown that the collection of persistent pairings (derived from $R$) are~\cite{zomorodian2005computing}. 
-%When reducing the boundary matrix, we can therefore perform column operations in any order, as long as columns are added from left to right, and so long as every column operation performed in $R$ is also performed on $V$.
-In particular, once $R$ is in reduced form, the persistence diagram of the corresponding filtration $K$ is given by lowest non-zero ``pivot'' entries in $R$.
+Note that though the matrices $R$ and $V$ are not unique, the collection of persistent pairings are~\cite{zomorodian2005computing}.
+The persistence diagram of the corresponding filtration $K$ may be read off from lowest entries in $R$, once $R$ is in reduced form.
 % Important corollaries of the persistence pairing lemma is that the 
@@
 %the non-zero coefficients $(\alpha_i^{(j)}, \alpha_{i+1}^{(j)}, \dots, \alpha_j^{(j)})$ via the combination: 
-\begin{equation}
-	r_j = D v_j = \alpha_i^{(j)} \cdot \partial(\sigma_{i+1}) + \alpha_{i+1}^{(j)} \cdot \partial(\sigma_{i+1}) + \dots + \alpha_{j}^{(j)} \cdot \partial(\sigma_{j}) = 0
+\begin{equation}\label{eq:coefficients}
+	r_j = D v_j = \alpha_i^{(i)} \cdot \partial(\sigma_{i}) + \alpha_{i}^{(i+1)} \cdot \partial(\sigma_{i+1}) + \dots + \alpha_{i}^{(j)} \cdot \partial(\sigma_{j}) = 0
 \end{equation}
 Now, consider the permutation $P$ that moves a simplex $\sigma_i$ in $K$ to position $j$, shifting all intermediate simplices $\sigma_{i+1}, \dots, \sigma_{j}$ down by one ($i < j$). 
-To perform this shift, the non-zero coefficients $\alpha_{i}^{(i+1)}, \dots, \alpha_i^{(j)}$ encoded in the boundary chains $r_{i+1}, \dots, r_j$ must be set to zero, otherwise $\partial(\sigma_i)$ contributes to earlier simplices in the filtration.
-As these coefficients are recorded by $v_j$, setting $\alpha_i^k = 0$  amounts to setting $v_k(i) = 0$ for all $k \in [i+1,j]$---algorithmically, these manifest as column operations $\ast$ used to ensure $PV\ast$ is upper-triangular, thus maintaining invariant I2.
-Notice that, to maintain invariant I1, these operations induce a potentially unreduced $R' = P R \ast P^T$, breaking invariant I3.
-We could reduce $R'$ with an additional $k$ operations, however $k$ may scale quadratically in the interval $\lvert i - j\rvert$, rendering the efficiency of this approach at best that of vineyards and at worse that of the standard algorithm.
+To perform this shift, the non-zero coefficients from~\eqref{eq:coefficients} must be set to zero, otherwise $\partial(\sigma_i)$ contributes to boundary chains $r_{i+1}, \dots, r_j$ earlier in the filtration.
+As these coefficients are recorded by $v_j$, setting $\alpha_i^k = 0$  amounts to setting the row entries $v_k(i) = 0$ for all $k \in [i+1,j]$---these cancellations manifest as column operations $(\ast)$ used to ensure $PV(\ast)$ is upper-triangular, thus maintaining invariant I2.
+Notice that these operations may induce an unreduced $R' = P R(\ast) P^T$, breaking invariant I3.
+We could reduce $R'$ with an additional $k$ operations, however its possible that $k \sim O(\lvert i - j\rvert^2)$, rendering the efficiency of this approach at best that of vineyards and at worse that of the standard algorithm.
 
@@
 As shown by example 2.3, move updates are cheaper than vineyard updates when $j > i+1$. More formally, we recall the proposition shown in~\cite{busaryev2010tracking}: 
-\begin{proposition}[Busaryev et al.~\cite{busaryev2010tracking}]
+\begin{proposition}[Busaryev et al.~\cite{busaryev2010tracking}]\label{prop:move_st}
 	Let $K$ denote a filtration with $n$ simplices of dimension $p-1$, $p$, and $p+1$, respectively, and let $R = DV$ denote its associated decomposition.
-Then the operation $\mathrm{MoveRight}(i,j)$ constructs a valid decomposition $R' = D'V'$ in $O((s+t)n)$ time, where $s,t$ are given by:
-	$$ s = \sum\limits_{l=i+1}^j \mathds{1}\left(\, v_l(i) \neq 0 \, \right), \quad t = \sum\limits_{l=1}^m \mathds{1}\left(\, \mathrm{low}_R(l) \in [i,j] \text{ and } r_l(i) \neq 0 \, \right)$$
-	Moreover, the quantity $k = s + t$ satisfies $k \leq 2(j - i)$. 
+Then the operation $\mathrm{MoveRight}(i,j)$ constructs a valid decomposition $R' = D'V'$ in $O((\lvert \mathbb{I} \rvert + \lvert \mathbb{J} \rvert)n)$ time, where $\mathbb{I}, \mathbb{J}$ are given by:
+	$$ \lvert \mathbb{I}\rvert = \sum\limits_{l=i+1}^j \mathds{1}\left(\, v_l(i) \neq 0 \, \right), \quad \mathbb{J} = \sum\limits_{l=1}^m \mathds{1}\left(\, \mathrm{low}_R(l) \in [i,j] \text{ and } r_l(i) \neq 0 \, \right)$$
+	Moreover, the quantity $\lvert \mathbb{I} \rvert + \lvert \mathbb{J} \rvert$ satisfies $\lvert \mathbb{I} \rvert + \lvert \mathbb{J} \rvert \leq 2(j - i)$. 
 \end{proposition}
@@
 As $p$ is now a strictly increasing sequence, we may compute the longest increasing subsequence $\mathrm{LIS}(q)$ and use this subsequence to recover a longest common subsequence (LCS) between $(K_i, K_{i+1})$, which we denote later with $\mathrm{LCS}(K_i, K_{i+1})$.
-We pass $q$ and $\mathrm{LIS}(q)$ to our greedy scheduling algorithm, which returns as output an ordered set of move permutations  $\mathcal{S}$ of minimum size, which we call a \emph{schedule}.
+We pass $q$ and $\mathrm{LIS}(q)$ to our greedy scheduling algorithm, which returns as output an ordered set of move permutations  $\mathcal{S}$ of minimum size, which we call a \emph{schedule} (see Definition~\ref{def:schedule}).
 Note one need not explicitly store the entire family of filtrations nor the schedules between them---Algorithm~\ref{alg:schedule} may be easily modified to be completely \emph{online}, keeping at most two filtrations and one decomposition in memory at any given time.
@@
 
-\subsection{Minimizing schedule size}\label{sec:schedule_sizes}
-Here we describe the reasoning for lines (4-5) in Algorithm~\ref{alg:schedule}.
-Recall in section~\ref{sec:motivation}, an example was given showing the vineyards algorithm efficiency compared to a naïve approach at three varying levels of coarseness, where it was shown that vineyards provides an efficient solution with a fine enough update scheme but otherwise scales poorly if the discretization of time is coarse.
-One may attempt to simply permute the decomposition and then repair it, however its known that applying arbitrary permutations to a given decomposition may require $O(m^3)$ column operations to ensure validity---this was one of the motivations for the creation of the vineyards algorithm.
-
+%\subsection{Minimizing schedule size}\label{sec:schedule_sizes}
+%Recall in section~\ref{sec:motivation}, an example was given showing the vineyards algorithm efficiency compared to a naïve approach at three varying levels of coarseness, where it was shown that vineyards provides an efficient solution with a fine enough update scheme but otherwise scales poorly if the discretization of time is coarse.
+%One may attempt to simply permute the decomposition and then repair it, however its known that applying arbitrary permutations to a given decomposition may require $O(m^3)$ column operations to ensure validity---this was one of the motivations for the creation of the vineyards algorithm.
+
+In the following subsections, we investigate whether we may profitably exploit the increased flexibility Busaryevs move framework yields us.
+% TODO: re-write 
 It was hypothesized that the reason the vineyards algorithm was more expensive than the naïve approach is due to the extra overhead of maintaining the decomposition at each transposition.
-Thus, decreasing the number of times the decomposition is restored to a valid state ought to reduce the total number of column operations needed to update a decomposition.
+Thus, in theory, decreasing the number of times the decomposition is restored to a valid state ought to reduce the total number of column operations needed to update a decomposition.
 This motivates the question: can one simultaneously minimize the number of times the decomposition is restored to a valid state and retain an efficient update scheme for \emph{arbitrary} permutations?
 
-\subsubsection{Continuous setting}\label{sec:continuous_setting}
+\subsection{Continuous setting}\label{sec:continuous_setting}
 In the continuous time case (i.e., vineyards), we are given a homotopy $F : K \times [0,1] \to \mathbb{R}$ interpolating between  a pair of filtrations $K, K'$ from some underlying complex $S$, with simplices ordered according to $f = F(\cdot, 0)$ and $f' = F(\cdot, 1)$, respectively.
@@
 \end{equation}
-This implies the $\lvert S_F \rvert \sim O(m^2)$ in the worst case, achieved when $f = - f'$, which is exhibited in the grayscale image data example in section~\ref{sec:motivation}: each 9x9 patch contains $(81, 208, 128)$ simplices of dimensions $(0, 1, 2)$ yielding $m=417$ simplices in total.
-As the homotopy is simulated, observe $\approx 70,\!000$ transpositions are generated.
-The upper bound size on $\lvert S_F \rvert$ is thus $\binom{417}{2} \approx 87,\!000$, which nearly occurs in this case, since the last filtration is nearly the reverse of the initial one.
+This implies the $\lvert S_F \rvert \sim O(m^2)$ in the worst case, achieved when $f = - f'$, which the grayscale image data example from section~\ref{sec:motivation} exhibits. 
+As the 9x9 patch contains $(81, 208, 128)$ simplices of dimensions $(0, 1, 2)$, the largest number of transpositions a schedule simplices in total.
+As the homotopy is simulated, observe $\approx 70,\!000$ transpositions are generated, approaching the upper bound of , due to the fact that $f'$  is nearly the reverse of $f$.
 
@@
 
-\subsection{Discrete setting}
-
+\subsection{Discrete setting}\label{sec:moves_setting}
 %If a straight line homotopy is used, the number of collapsible subsequences may depend on the simplices categorization as critical points; transpositions between pairs of simplices which trigger  changes in homology tend to be more expensive.
@@
 
-Contrasting the continuous-time setting, if we discard the use of a homotopy interpolation and allow move operations in any order, we have $\widetilde{S} \sim O(m)$ trivially: simply move each simplex in $K$ into its position in the filtration given by $K'$ in the order given by the latter.
-However, it's not immediately clear whether this bound is tight, or whether this strategy of generating schedules---which we call \emph{naive strategy}---is performant.
+Contrasting the continuous-time setting, if we discard the use of a homotopy interpolation and allow move operations in any order, we obtain a trivial upper bound of $O(m)$ on the schedule size:  simply move each simplex in $K$ into its position in the filtration given by $K'$ in the order given by the latter---which we call \emph{naive strategy}.
+However, it's not immediately clear whether this bound is tight.
+Moreover, we expect  investigating the tightness of this bound may lead to exploitable structure.
+% or whether this strategy of generating schedules---which we call \emph{naive strategy}---is performant.  
 %Indeed, arbitrarily permuting a given filtration $K$: one must respect the face poset.
 
-To establish this formally, we require a few definitions.
+Before proceeding, we require a few definitions.
 Given two fixed permutations $p, q \in S_m$ and a set $\Sigma \subseteq S_m$,  a common  problem is to find a sequence of permutations $s_1, s_2, \dots, s_d \in \Sigma$ whose composition satisfies:
@@
 A sequence $S = \{s_1, s_2, \dots, s_d\} \subseteq \Sigma \subseteq S_m$ of operations mapping $p \mapsto q$ is sometimes called a \emph{sorting of $p$}. 
-When $p, q$ are interpreted as strings, distances defined with respect to a fixed $\Sigma \subseteq S_m$ are commonly referred to as \emph{edit distances}~\cite{labarre2013lower}, which are denoted as $d_\Sigma(\cdot, \cdot)$, and the subsequent elementary operations on the strings are called \emph{edit operations}.
+When $p, q$ are interpreted as strings, distances defined with respect to a fixed $\Sigma \subseteq S_m$ are commonly referred to as \emph{edit distances}~\cite{bergroth2000survey}, which are denoted as $d_\Sigma(\cdot, \cdot)$, and the subsequent elementary operations on the strings are called \emph{edit operations}.
 Note the choice of $\Sigma$ defines the universe of operations and thus the type of edit distance being measured---otherwise if $\Sigma = S_m$, then $d_\Sigma(p, q) = 1$ for any $p\neq  q \in S_m$.
@@
 
-Here we consider sorting by moves. Phrased using permutations, a \emph{move operation} $m_{ij}$ that moves $i$ to $j$ in $[m]$, for $i < j$, corresponds to the permutation: 
+Here we consider sorting by moves. Using permutations, a \emph{move operation} $m_{ij}$ that moves $i$ to $j$ in $[m]$, for $i < j$, corresponds to the circular rotation: 
 \begin{equation}\label{eq:move_perm}
@@
 \end{equation}
-In cycle notation, when $i < j$, this corresponds to the cyclic permutation 
+In cycle notation, this corresponds to the cyclic permutation:
 \begin{equation}\label{eq:move_permutation}
@@
 Similarly, in the context of edit operations, observe that a move operation can be interpreted as a paired delete-and-insert operation, i.e.
-$m_{ij} = (\mathrm{ins}_j \circ \mathrm{del}_i)$, where $\mathrm{del}_i$ denotes the edit operation that deletes the character at position $i$  and $\mathrm{ins}_j$ similarly inserts the same character at position $j$.
-Thus, sorting by move operations can be interpreted as finding a minimal sequence of edits where the only operations allowed are insertions and deletions, where each insertion/deletion operation is coupled---which is exactly the well known \emph{Longest Common Subsequence} (LCS) distance.
-The LCS distance between two strings $p, q$ of sizes $m$ and $n$, respectively, is given by~\cite{bergroth2000survey}:
+$m_{ij} = (\mathrm{ins}_j \circ \mathrm{del}_i)$, where $\mathrm{del}_i$ denotes the operation that deletes the character at position $i$  and $\mathrm{ins}_j$ the operation that inserts the same character at position $j$.
+Thus, sorting by move operations can be interpreted as finding a minimal sequence of edits where the only operations allowed are (paired) insertions and deletions---this is exactly the well known \emph{Longest Common Subsequence} (LCS) distance.
+Between strings $p, q$ of sizes $m$ and $n$, the LCS distance is given by~\cite{bergroth2000survey}:
 \begin{equation}
@@
 \end{equation}
-Intuitively, in the context of strings, given $\mathcal{L} = \mathrm{LCS}(p,q)$ one may transform $p \mapsto q$ by simply deleting the symbols from $p \setminus \mathcal{L}$ and then inserting symbols from $q \setminus \mathcal{L}$.
+%Intuitively, in the context of strings, given $\mathcal{L} = \mathrm{LCS}(p,q)$ one may transform $p \mapsto q$ by simply deleting the symbols from $p \setminus \mathcal{L}$ and then inserting symbols from $q \setminus \mathcal{L}$.
 % while moving $j$ to $i$ is the inverse permutation.
@@
 %Like the \emph{LCS} distance, 
-The Wagner-Fischer algorithm---a dynamic programming type approach---can compute the LCS distance in $O(m^2)$~\cite{bar2004approximating}.
-However, we can do better: if the pair of filtrations $K_0, K_1$ are identical up to a permutation, then computing $d$ reduces further to the \emph{permutation edit distance} problem.
-As such, we obtain the following bound on the minimum size of a sorting (i.e., a schedule) using moves and the complexity of computing it. 
+In general, one can compute the LCS itself in $O(mn)$ easily with dynamic programming.
+One might hope computing the size $\mathrm{d}_{\mathrm{lcs}}(p,q)$ alone exhibits a lower complexity, however there is substantial evidence that for general string inputs the complexity cannot be much lower than quadratic~\cite{abboud2015tight}.
+However, if the pair of filtrations $K, K'$ are from the same underlying complex $S$, then they may both be thought of a permutations in $S_m$---the corresponding edit distance $d$ then reduces to the \emph{permutation edit distance} problem.
+With this insight in mind, we obtain the following bound on the minimum size of a sorting (i.e. schedule) using moves and the complexity of computing it. 
 \begin{proposition}[Schedule Size]\label{prop:schedule_size}
@@
 \end{proof}
+\noindent
+Establishing a connection between the permutation edit distance and move scheduling allows use to exploit the combinatorial structure that comes from the developed theory on both LCS's and LIS's, which are both well-studied objects.
+We record a single corollary to demonstrate this fact.
 \begin{corollary}\label{cor:expectation}
-If $K, K'$ are random filtrations of a common simplicial complex $S$ of size $m$.
-Then the size of longest common subsequence of simplices between $K,K'$ is no larger than $m - \sqrt{m}$ in expectation, with probability $1$ as $m \to \infty$.
+If $K, K'$ are random filtrations of a common simplicial complex $S$ of size $m$, then the size of longest common subsequence of simplices between $K,K'$ is no larger than $m - \sqrt{m}$ in expectation, with probability $1$ as $m \to \infty$.
 \end{corollary}
@@
 %One may wonder if filtrations would exhibit a smaller upper bound due to the subface relation that must be respected.
-However, Boissonnat~\cite{boissonnat2018efficient} prove the number of distinct filtrations built from $k$-dimensional simplicial complex $K$ with $m$ simplices and $t$ distinct filtration values is \emph{at least} $\floor{\frac{t+1}{k+1}}^m$.
+However, Boissonnat~\cite{boissonnat2018efficient} prove the number of distinct filtrations built from a $k$-dimensional simplicial complex $K$ with $m$ simplices and $t$ distinct filtration values is \emph{at least} $\floor{\frac{t+1}{k+1}}^m$.
 Since this bound grows similarly to $m!$ when $t \sim O(m)$ and $k << m$ fixed, $d \approx n - \sqrt{n}$ is not too pessimistic a bound between random filtrations.     
 \end{remark}
+\noindent 
 In practice, when one has a time-varying filtration and the sampling points are relatively close [in time], the LCS between adjacent filtrations is expected to be much larger, shrinking $d$ substantially.
-For example, the average size of the LCS across the 10 filtrations from Section~\ref{sec:motivation} each having $m=417$ simplices sampled at evenly spaced time points was $343$, implying $d \approx 70$ permutations needed on average to update the decomposition at adjacent time points.
-
-Before discussing the practice of constructing such a schedule, we conclude by combining the propositions we've given so far and present the main theorem of this effort. 
+For example, for the complex from Section~\ref{sec:motivation} with $m=417$ simplices, the average size of the LCS across the 10 evenly spaced filtrations was $343$, implying $d \approx 70$ permutations needed on average to update the decomposition between adjacent time points.
+
+We conclude this section with the main theorem of this effort: an output-sensitive bound on the simulation of persistence dynamically.
   \begin{theorem}\label{prop:sparsity_result}
-   Given a pair of distinct filtrations $K, K'$ of a simplicial complex $S$ and a decomposition $R = DV$ of $K$, the size of a minimal sequence $\mathcal{S} = \left( s_1, s_2, \dots, s_d \right)$ of cyclic permutations $s_k = (i_k, j_k)$ satisfying~\eqref{eq:move_perm} and:
+   Given a pair of distinct filtrations $K, K'$ of a simplicial complex $S$ and a decomposition $R = DV$ of $K$, the size of a minimal sequence $\mathcal{S} = \left( s_1, s_2, \dots, s_d \right)$ of cyclic `move' permutations $s_k = (i_k, j_k)$ satisfying:
    	\begin{equation}\label{eq:rv_seq}
@@
    	Moreover, if $i_k < j_k$ for all $k \in [d]$, then computing~\eqref{eq:rv_seq} requires $O(d\kappa)$ column operations, where:	  
-	$$  \quad \kappa = \sum\limits_{k = 1}^d \, \lvert \mathbb{I}_{k}\rvert + \lvert \mathbb{J}_{k}\rvert $$
-	% where $\lvert \mathbb{I}_{(k)} \rvert$ and $\lvert \mathbb{J}_{(k)} \rvert$ are the computed quantities in lines (2) and (3) in algorithm~\ref{alg:mr} (here we assume if $i_\ast < j_\ast$).
-	and where $\lvert \mathbb{I}_{k} \rvert$ denotes the number of non-zero entries $V_k[i_k, i_k:j_k]$, $\lvert \mathbb{J}_{k}\rvert$ the number of columns satisfying $\mathrm{low}_{R_k} \in [i_k:j_k]$ and $\mathrm{row}_{R_k}(i_k) \neq 0$, and $d = m - \lvert \mathrm{LCS}(K, K') \rvert$.
+   	$$  \quad \kappa = \sum\limits_{k = 1}^d \, (s_{k} + t_{k}) $$
+%	$$  \quad \kappa = \sum\limits_{k = 1}^d \, \lvert \mathbb{I}_{k}\rvert + \lvert \mathbb{J}_{k}\rvert $$
+	where $s_k, t_k$ are quantities from~\ref{prop:move_st} derived from the non-zeroes of $R_k, V_k$ in the sequence~\ref{eq:rv_seq}.
+Note that since $\kappa$ depends on the sparsity of the intermediate entries $V_1, V_2, \dots, V_d$ and $R_1, R_2, \dots, R_d$, the bound $O(d \kappa)$ is output-sensitive.
+%	and where $\lvert \mathbb{I}_{k} \rvert$ denotes the number of non-zero entries $V_k[i_k, i_k:j_k]$, $\lvert \mathbb{J}_{k}\rvert$ the number of columns satisfying $\mathrm{low}_{R_k} \in [i_k:j_k]$ and $\mathrm{row}_{R_k}(i_k) \neq 0$, and $d = m - \lvert \mathrm{LCS}(K, K') \rvert$.
 \end{theorem}
@@
  %% New section
-\subsection*{Constructing schedules}
+\subsection{Constructing schedules}\label{sec:schedule_construction}
 It is clear from Corollary~\ref{cor:expectation} that one may compute the LCS between two permutations $p, q \in S_m$ via the LIS of a single permutation $p^\ast$, and that computation may be carried out in $O(m \log \log m)$ time.
@@
 \subsection{Minimizing schedule cost}\label{sec:schedule_cost}
-The results from the previous section yield a sufficient algorithm for generating move schedules of minimal cardinality: any schedule of moves $S$ sorting $K_0 \mapsto K_1$ above is guaranteed to have size $\lvert S \rvert = m - \lvert \mathrm{LCS}(K_0, K_1) \rvert$, and the reduction to the permutation edit distance problem ensures this size is optimal.
+The algorithm outlined in section~\ref{sec:schedule_construction} is a sufficient for generating move schedules of minimal cardinality: any schedule of moves $S$ sorting $f \mapsto f'$ above is guaranteed to have size $\lvert S \rvert = m - \lvert \mathrm{LCS}(f, f') \rvert$, and the reduction to the permutation edit distance problem ensures this size is optimal.
 However, as with the vineyards algorithm, certain pairs of simplices cost more to exchange depending on whether they are critical pairs in the sense described in~\cite{cohen2006vines}, resulting in a large variability in the cost of randomly generated schedules.
@@
 \subsubsection{Greedy approach}\label{sec:greedy} 
-One advantage of using move operations is that the cost of applying a permutation to some given $R = D V$ decomposition can be determined efficiently prior to performing any column operations. 
-Specifically, the cost of determining $\lvert \mathbb{I}\rvert + \lvert \mathbb{J}\rvert$ in lines (2) and (3) of Algorithm~\ref{alg:mr} cannot take more than $O(m)$ time. 
+Another flexibility granted by the move framework is that the cost of a move on a given $R = D V$ decomposition can be determined efficiently prior to any column operations. 
+Specifically, the cost of determining $\lvert \mathbb{I}\rvert + \lvert \mathbb{J}\rvert$ in Proposition~\ref{prop:move_st} cannot take more than $O(m)$ time with row-oriented sparse matrices. 
 Towards minimizing the number of column operations a schedule requires, it is natural to consider whether a greedy-type solution which chooses the minimal cost choice at every step yields an efficient strategy.
@@
 \\
-\textbf{Counter-Example} A pair of filtrations is given below, each filtration contains $m = 10$ simplices comprising the $1$-skeletons of a $3$-simplex with a particular choice of ordering.
-Relabeling the simplices of $K_0$ to the index set $[m]$ and modifying the labels of $K_1$ correspondingly yields permutations given below:
+\textbf{Counter-Example:} A pair of filtrations is given below, each comprising the $1$-skeleton of a $3$-simplex.
+Relabeling $(K, f)$ to the index set $f : K \to [m]$ and modifying $(K, f')$ accordingly yields the permutations given below:
 %$$ K_0 = \{ {\color{red} (1) (2) (3) (4) (1,4) (2,4) (3,4) } (1,2) (2,3) (1,3) \} $$ 
 %$$ K_1 = \{ {\color{red} (1) (2) (3) (4) } (1,2) (2,3) (1,3) {\color{red} (1,4) (2,4) (3,4) }\} $$ 
-$$ K_0 = \{ {\color{red} a \; b \; c \; d \; u \; v \; w} \; x \; y \; z  \} = { {\color{red} 1 \; 2 \; 3 \; 4 \; 5 \; 6 \; 7 } \; 8 \; 9 \; 10 } $$
-$$ K_1 = \{ {\color{red} a \; b \; c \; d } \; x \; y \; z \; {\color{red} u \; v \; w } \}  = { {\color{red} 1 \; 2 \; 3 \; 4 } \;8 \; 9 \; 10 \; {\color{red} 5 \; 6 \; 7 } }$$
-The subset of the filtration which corresponds to the simplices which lie in the LCS between these permutations is colored in red. For this example, the permutation edit distance is 
-$d = m - \lvert \mathrm{LCS}(K_0, K_1) \rvert$ implies that exactly $3$ moves are needed to apply the mapping $K_0 \mapsto K_1$. There are six possible valid schedules of moves: 
+$$ (K, f) = \{ {\color{red} a \; b \; c \; d \; u \; v \; w} \; x \; y \; z  \} = { {\color{red} 1 \; 2 \; 3 \; 4 \; 5 \; 6 \; 7 } \; 8 \; 9 \; 10 } $$
+$$ (K, f') = \{ {\color{red} a \; b \; c \; d } \; x \; y \; z \; {\color{red} u \; v \; w } \}  = { {\color{red} 1 \; 2 \; 3 \; 4 } \;8 \; 9 \; 10 \; {\color{red} 5 \; 6 \; 7 } }$$
+The subset of the filtration which corresponds to the simplices which lie in the LCS between these permutations is colored in red. For this example, the edit distance is 
+$d = m - \lvert \mathrm{LCS}(f, f') \rvert$, implying exactly $3$ moves are needed to map $f \mapsto f'$. There are six possible valid schedules of moves: 
 \begin{alignat*}{3}
@@
 \end{table}
-A greedy strategy which always selects the cheapest move in succession would begin by moving $x$ or $z$ first, since these are the cheapest moves available, which implies one of $S_1, S_2, S_5, S_6$ would be picked depending on the choice of a tie-breaker.
+Note the greedy strategy which always selects the cheapest move in succession would begin by moving $x$ or $z$ first, since these are the cheapest moves available, which implies one of $S_1, S_2, S_5, S_6$ would be picked depending on the tie-breaker.
 While the cheapest schedule $S_1$ is in this candidate set, so is $S_6$, the most expensive schedule.
@@
 Although the greedy approach from the last section can lead to arbitrarily high cost schedules, similar greedy-like strategies can yield computationally efficient heuristics in practice, even if they are suboptimal.
-For example, in the previous section choosing the lowest-cost move takes $\approx O(dm)$---implying a greedy strategy would require $O(d^2 m)$ time---which is computationally viable as a heuristic when $d << m$.
-%Since the persistence algorithm requires $\approx O(m^3)$, the schedule generation procedure must at most have $O(m^2)$ complexity in its operation. 
-However, when $d \sim O(m)$, the sorting procedure shown using Algorithm~\ref{alg:sorting} approaches $O(m^2 \log m)$.
-We seek a fast procedure for generating schedules; if the schedule generation procedure exhibits a complexity of $\approx O(m^3)$ or higher, then such a procedure would offer no computational benefit compared to the reduction algorithm.
-Indeed, observe the vineyards algorithm requires $O(m \log m + k)$ preprocessing time to compute a schedule of size $k \sim O(m^2)$ transpositions, thus for Algorithm~\ref{alg:schedule} to be considered a more efficient alternative to vineyards, a solution which is subquadratic in $m$ is desirable.
+Choosing the lowest-cost move takes $\approx O(dm)$---implying a greedy strategy would require $O(d^2 m)$ time---which is computationally viable as a heuristic when $d << m$. 
+We seek a fast procedure for generating schedules that is not only substantially better than a randomly or naive strategy in term of column reductions, but is also low complexity enough that it is practical in a real-world setting.
+For example, the vineyards algorithm requires $O(m \log m + k)$ preprocessing time to compute a schedule of size $k \sim O(m^2)$ transpositions, thus for Algorithm~\ref{alg:schedule} to be considered a more efficient alternative to vineyards, a solution subquadratic in $m$ is desirable.
 %In what follows, we present such an objective, showing its relation to the problem at hand as well its use in other known computational problems. 
@@
 Like the Kendall-$\tau$ distance, it is a metric, it is invariant under relabeling, and the distance between pairs of uniformly random permutations $p, q \sim S_m$ are asymptotically normal.
-Indeed, Diaconis et al.~\cite{diaconis1977spearman} showed the the Spearman distance between any two permutaton $p, q \in S_m$ approximates their Kendall-$\tau$ distance within a factor of two:
+Indeed, Diaconis et al.~\cite{diaconis1977spearman} showed the the Spearman distance between any two permutation $p, q \in S_m$ approximates their Kendall-$\tau$ distance within a factor of two:
 \begin{equation}\label{eq:diaconis_inequality}
@@
 % \section{Applications and Discussion}\label{sec:applications}
-\section{Empirical results}\label{sec:results}
+\section{Applications and Experiments}\label{sec:results}
 \subsection*{Crocker stacks}
-There are many challenges to depicting topological behavior in dynamic settings.
+There are many challenges to characterizing topological behavior in dynamic settings.
 One approach is to trace out the curves constituting a continuous family of persistence diagrams in $\mathbb{R}^3$---the vineyards approach---however this visualization can be cumbersome to work with as there are potentially many such vines tangled together, making topological critical events with low persistence difficult to detect.
@@
 The green $\mathrm{X}$ marks correspond chronologically to the complexes (left), in row-major order.
-The large orange and purple areas depict $1$-cycles persisting in both space (the y-axis) and time (x-axis).
-}\label{fig:crocker1}
+The large orange and purple areas depict $1$-cycles persisting in both space (y-axis) and time (x-axis).} \label{fig:crocker1}
 \end{figure}
@@
 	\includegraphics[width=\textwidth, height=16.35em]{boid_sim_results.png}
-	\caption{ The cumulative number of $O(m)$ column operations (left, log-scale) of the three approaches tested and the distribution of the normalized Kendall distance between adjacent filtration pairs.
-The histogram on the right depicts the coarseness of the discretization; most filtration pairs have $\approx 5\%$ of their $\approx O(m^2)$ simplex pairs inverted between adjacent time steps.}\label{fig:boid_sim_results}
+	\caption{ One the left, the cumulative number of column operations (log-scale) of the three baseline approaches tested.
+On the right, the normalized $K_\tau$ between adjacent filtrations depicts the coarseness of the discretization---about $5\%$ of the $\approx O(m^2)$ simplex pairs between adjacent filtrations are discordant.}
+	\label{fig:boid_sim_results}
 \end{figure}
@@
 The standard algorithm \texttt{pHcol} and our move strategy were computed at 60 evenly spaced time points of the simulations.
-As depicted in Figure~\ref{fig:boid_sim_results}, our move strategy is far more efficient then both vineyards and the naive \texttt{pHcol} strategies.
-
-\subsection*{Multidimensional persistence}
-Given a procedure to filter a space in multiple dimensions simultaneously, a \emph{multifiltration}, the goal of multidimensional persistence is to identify persistent features by examining the entire multifiltration.
+As depicted in Figure~\ref{fig:boid_sim_results}, our move strategy is far more efficient than both vineyards and the naive \texttt{pHcol} strategies.
+
+\subsection*{Multiparameter persistence}
+Given a procedure to filter a space in multiple dimensions simultaneously, a \emph{multifiltration}, the goal of multi-parameter persistence is to identify persistent features by examining the entire multifiltration.
 Such a generalization has appeared naturally in many application contexts, showing potential as a tool for exploratory data analysis~\cite{lesnick2012multidimensional}.
@@
 For example, there may be better bounds or algorithms available if stronger assumptions can be made on how the filtration is changing with time.
-Moreover, if the filtration $K_0$ shares little similarity to the ``target'' filtration $K_1$, then the overhead of reducing the simplices from $K_1 \setminus K_0$ appended to the decomposition derived from $K_0$ may be large enough to motivate simply computing the decomposition at $K_1$ independently, especially if parallel processors are available.
+Moreover, if the filtration $K$ shares little similarity to the ``target'' filtration $K'$, then the overhead of reducing the simplices from $K' \setminus K$ appended to the decomposition derived from $K$ may be large enough to motivate simply computing the decomposition at $K_1$ independently, especially if parallel processors are available.
 Indeed, in this setting the size of the LIS may very well be on the order $\sim m - \sqrt{m}  = O(m)$, rendering the efficiencies gained using our approach as negligible.
-Thus, our approach is primarily useful if filtration in the parameterized family is ``nearby'' in the combinatorial sense. 
-Note that this overhead does not exist for the multidimensional persistence use-case: the entire decomposition is needed at each template 2-cell.
+Thus, our approach is primarily useful if the filtrations in the parameterized family is ``nearby'' in the combinatorial sense.
 
@@
 CSC, CSR, etc.), permuting both the rows and columns generally takes at most $O(Z)$ time, where $Z$ is the number of non-zero entries.
-If the particular filtration has many cycles, permuting these compressed representations can be quite expensive, bottlenecking the computation both in theory and in practice.
+If the particular filtration has many cycles, permuting these compressed representations can be quite expensive.
 As a result, the more complex sparse matrix representation from~\cite{cohen2006vines} is necessary to be efficient in practice.

### 2023-01-28 move_schedules.tex
Greatly simplified natural images section, finishing the simplifications of the multiparameter use case

@@
 
-\subsection*{Multiparameter persistence}
+\subsection{Multiparameter persistence}
 Given a procedure to filter a space in multiple dimensions simultaneously, a \emph{multifiltration}, the goal of multi-parameter persistence is to identify persistent features by examining the entire multifiltration.
@@
 One exemplary use case of multi-parameter persistence is to detect these strong outliers by filtering the data with respect to both density and the associated metric.
-
+In this section, we show the utility of scheduling with a real-world use case: detecting the presence of a low-dimensional topological space which well-approximates the distribution of natural images.
+As a quick outline, in what follows we briefly recall the fibered barcode invariant~\ref{sec:fibered_barcode}, summarize its potential application to a particular data set with known topological structure~\ref{sec:natural_images}, and conclude with experiments of demonstrating how scheduling enables such applications~\ref{sec:empirical_klein}.
+
+\subsubsection{Fibered barcode}\label{sec:fibered_barcode}
 Unfortunately, unlike the one-parameter case, there is no complete discrete invariant for multi-parameter persistence.
@@
 
- \subsubsection*{Natural images dataset} 
-As we have outlined the computational theory of 2-parameter persistence and elucidated it relevance to our proposed move scheduling approach, we now demonstrate the efficiency of scheduling with a practical application: detecting the presence of a low-dimensional topological space which well-approximates the distribution of patches from high-contrast images.
-
+ \subsubsection{Natural images dataset}\label{sec:natural_images}
 A common hypothesis is that high dimensional data tend to lie in the vicinity of an embedded, low dimensional manifold or topological space.
-An exemplary demonstration of this is given in~\cite{lee2003nonlinear}, who explored the space of $3 \times 3$ high-contrast patches extracted from Hans van Hateren's still image collection~\cite{hateren_schaaf_1998}, which consists of approximately 4,000 monochrome images\footnote{See \url{http://bethgelab.org/datasets/vanhateren/} for details on the image collection.} depicting various areas outside around Groningen (Holland).
-Motivated by applications to sparse coding applications, Lee et al.~\cite{lee2003nonlinear} was interested in how these natural image patches were distributed, in pixel-space, with respect to predicted spaces and manifolds.
-One of the motivating questions of their research was whether there were any clear qualitative differences the distributions of patches extracted from images of different modalities, such optical and range images.
-A major finding from their work is that the majority of high-contrast image patches are concentrated around a 2-dimensional submanifold embedded in $S^7$.
-Supplemental experiments conducted by Carllsson et al.~\cite{carlsson2008local} using persistent homology confirmed that the distribution of high-contrast $3 \times 3$ patches is indeed well-approximated by a Klein bottle---around 60\% of the high-contrast patches from the still image data set are concentrated within a small neighborhood around a model of a Klein bottle which accounts for $\approx$ 21\% of the volume $S^7$, compared to the $\approx$ 84\% of volume of $S^7$ accounted for by the entire data set.
-Along a similar vein, in the context of sparse coding, Perea et al.~\cite{perea2014klein} introduced a dictionary learning framework for estimating and representing the distribution of patches from texture images.
-
- \begin{figure}[t]
-	\includegraphics[width=0.98\textwidth]{natural_images}
-	\caption{Bipersistence example of natural images data set on an $12 \times 16$ coarsened grid.
-On the left, a projection the full data set is shown, along with the 15 landmark patches.
-(Middle) the bigraded Betti numbers and a fixed line $L$ over parameter space.
-	As before, the $0/1/2$ dimension bigraded Betti numbers are shown in green/red/yellow, respectively, with the blue region highlighting where $\mathrm{dim}(M) = 5$.
-(Right) five persistent features representing $B_L(M)$ are revealed from the middle, matching $\beta_1$ of the three-circle model.
-}
-	\label{fig:patch_data_dgm}
-\end{figure}
-
-If one was not aware of the analysis done by~\cite{lee2003nonlinear, hateren_schaaf_1998, carlsson2008local, perea2014klein}, it is not immediately clear a priori that the Klein bottle model is indeed a good candidate for capturing non-linearity of image patches.
-Nonetheless, as we have mentioned, among the first uses of persistent homology was the so-called ``homology inference problem''~\cite{perea2018brief}; given a suitable treatment of strong outliers using by e.g.
-filtering both on the geometry and the density of the image patches, the homology of the Klein bottle should become apparent through a persistence computation.
-
-Consider the (coarsened) fibered barcode computed from a standard Rips / codensity bifiltration on a representative sample of the image data from~\cite{hateren_schaaf_1998}, shown in Figure~\ref{fig:patch_data_dgm}.
-From the bigraded Betti number and the dimension function, one finds that a large area of dimension function is constant (highlighted as the blue portion in the middle of Figure~\ref{fig:patch_data_dgm}), wherein the first Betti number is 5.
-There are many spaces whose  meets this criterion, though upon further inspection, one candidate space that seems particularly plausible is the three-circle model $C_3$.
-The $C_3$ model consists of three circles, two of which (say, $S_v$ and $S_h$) intersect the third (say, $S_\mathrm{lin}$) in exactly two points, but themselves do not intersect.
-Upon performing an analysis on the density of the patches analogous to the one done by~\cite{lee2003nonlinear}, one finds that the mean-centered and contrast normalized $3 \times 3$ patches like on a 7-dimensional ellipsoid $\tilde{S}^7 \subset \mathbb{R}^9$.
-Contrast is typically measured using a discrete version of the Dirichlet semi-norm, or the ``$D$-norm'' $\lVert \cdot \rVert_D$ for short, which is a unique scale-invariant norm on images:
+An exemplary demonstration of this is given in the analysis by Lee et al.~\cite{lee2003nonlinear}, who explored the space of high-contrast patches extracted from Hans van Hateren's~\cite{hateren_schaaf_1998} still image collection\footnote{See \url{http://bethgelab.org/datasets/vanhateren/} for details on the image collection.}, which consists of $\approx 4,000$ monochrome images depicting various areas outside Groningen (Holland).
+In particular,~\cite{lee2003nonlinear} were interested in exploring how high-contrast $3 \times 3$ image patches  were distributed, in pixel-space, with respect to predicted spaces and manifolds.
+Formally, they measured contrast using a discrete version of the scale-invariant Dirichlet semi-norm:
 $$ \lVert x \rVert_D = \sqrt{\sum_{i \sim j}(x_i - x_j)^2} = \sqrt{x^T D x}$$
-where $D$ is a fixed matrix which upon application to an image $x \in \mathbb{R}^9$ yields a value proportional to the sum of the differences between the 4-connected neighbors (described by the relation $i \sim j$).
-It is advantageous to ``whiten'' the data via a change of coordinates to Euclidean sphere.
-This can be achieved by expressing the patch data as an expansion of a certain convenient basis of 8 non-constant eigenvectors, often called the Discrete Cosine Transform (DCT) basis~\cite{lee2003nonlinear}, shown below:
+where $D$ is a fixed matrix which upon application to an image $x \in \mathbb{R}^9$ yields a value proportional to the sum of the differences between each pixels 4 connected neighbors (described above by the relation $i \sim j$).
+Their research was primarily motivated by discerning whether there existed clear qualitative differences in the distributions of patches extracted from images of different modalities, such optical and range images.
+By mean-centering, contrast normalizing, and``whitening'' the data via the Discrete Cosine Transform (DCT) basis~\cite{lee2003nonlinear}, they found they could form a convenient basis for $D$ via an expansion of 8 certain non-constant eigenvectors, shown below:
 %\vspace{-0.2em}
@@
 %\vspace{-1.5em}
-\noindent Projecting the image data onto the first two basis vectors of leads to the projection shown in the top left of Figure~\ref{fig:patch_data_dgm}.
-15 landmark points are also shown.
-Observe the data are distributed well around three ``circles''---the outside circle capturing the rotation gradient of the image patches ($S_{\mathrm{lin}}$), and the other two circles capturing the vertical and horizontal gradients ($S_v$ and $S_h$, respectively).
-Since the three circle model is the 1-skeleton of the Klein bottle~\cite{carlsson2008local}, one may conclude that the Klein bottle might be a reasonable candidate upon which the image data are distributed.
-This can be further verified empirically by e.g.
-using the iterative, meanshift-like procedure described in~\cite{carlsson2008local} to fit a parameterization of the model in question to the data.
-
-Using the same high-contrast patch data set studied in~\cite{lee2003nonlinear}, we evaluate the performance of various methods at computing the fibered barcode invariant using the parameterization from~\ref{sec:fibered_barcode_reparam}.
-Due to the aformentioned high complexity of the fibered barcode computation, we begin by working with a subset of the image patch data $\mathcal{X}$.
-Specifically, we combine the use furthest-point sampling and proportionate allocation (stratified) sampling to sample landmarks $X \subset \mathcal{X}$ distributed within $n = 25$ strata.
+Since these images are scale-invariant, the expansion of these basis vectors spans the 7-sphere, $S^7 \subset \mathbb{R}^8$.
+Using a voronoi cell decomposition of the data, their subsequent distribution analysis suggested that the majority of data points concentrated in a few high-density regions.
+
+After Lee et al published their work, Carlsson et al.~\cite{carlsson2008local} subsequently performed extensive experiments conducted using persistent homology, wherein he found that the distribution of high-contrast $3 \times 3$ patches is actually well-approximated by a Klein bottle $\mathcal{M}$---around 60\% of the high-contrast patches from the still image data set lie within a small neighborhood around $\mathcal{M}$ which accounts for only 21\% of the volume $S^7$.
+Along a similar vein, in the context of sparse coding, Perea et al.~\cite{perea2014klein} introduced a dictionary learning framework for estimating and representing the distribution of patches from texture images.
+
+ \begin{figure}[t]
+	\includegraphics[width=0.98\textwidth]{natural_images}
+	\caption{Bipersistence example of natural images data set on an $12 \times 16$ coarsened grid.
+On the left, a projection the full data set is shown, along with the 15 landmark patches.
+(Middle) the bigraded Betti numbers and a fixed line $L$ over parameter space.
+	As before, the $0/1/2$ dimension bigraded Betti numbers are shown in green/red/yellow, respectively, with the blue region highlighting where $\mathrm{dim}(M) = 5$.
+(Right) five persistent features representing $B_L(M)$ are revealed from the middle, matching $\beta_1$ of the three-circle model.
+}
+	\label{fig:patch_data_dgm}
+\end{figure}
+
+If one was not aware of the analysis done by~\cite{lee2003nonlinear, hateren_schaaf_1998, carlsson2008local, perea2014klein}, it is not immediately clear a priori that the Klein bottle model is  a good candidate for capturing non-linearity of image patches.
+Indeed, armed with fine topological intuition, Carlsson still needed to perform extensive sampling, preprocessing, and model fitting techniques in order reveal the underlying the topological space with persistent homology~\cite{carlsson2008local}.
+One reason such preprocessing is needed is due to persistent homology's aforementioned instability with respect to strong outliers. 
+In the ideal setting, a multi-parameter approach that accounts for the local density of points should require far less experimentation. 
+%Nonetheless, given a suitable treatment of strong outliers using by e.g.
+filtering both on the geometry and the density of the image patches, the homology of the Klein bottle should become apparent through a persistence computation.
+%Nonetheless, as we have mentioned, among the first uses of persistent homology was the so-called ``homology inference problem''~\cite{perea2018brief}; given a suitable treatment of strong outliers using by e.g.
+filtering both on the geometry and the density of the image patches, the homology of the Klein bottle should become apparent through a persistence computation.
+
+Consider the (coarsened) fibered barcode computed from a standard Rips / codensity bifiltration on a representative sample of the image data from~\cite{hateren_schaaf_1998}, shown in Figure~\ref{fig:patch_data_dgm}.
+From the bigraded Betti number and the dimension function, one finds that a large area of dimension function is constant (highlighted as the blue portion in the middle of Figure~\ref{fig:patch_data_dgm}), wherein the first Betti number is 5.
+Further inspection suggests one plausible candidate is the three-circle model $C_3$, which consists of three circles, two of which (say, $S_v$ and $S_h$) intersect the third (say, $S_\mathrm{lin}$) in exactly two points, but themselves do not intersect.
+Projecting the image data onto the first two basis vectors of leads to the projection shown in the top left of Figure~\ref{fig:patch_data_dgm}, of which 15 landmark points are also shown.
+Observe the data are distributed well around three ``circles''---the outside circle capturing the rotation gradient of the image patches ($S_{\mathrm{lin}}$), and the other two capturing the vertical and horizontal gradients ($S_v$ and $S_h$, respectively).
+Since the three circle model is the 1-skeleton of the Klein bottle, one may concur with Carlssons analysis~\cite{carlsson2008local} that the Klein bottle may be a reasonable candidate upon which the image data are distributed.
+% This can be further verified empirically by e.g. using the iterative, meanshift-like procedure described in~\cite{carlsson2008local} to fit a parameterization of the model in question to the data. 
+The degree to which multi-parameter persistence simplifies this exploratory phase cannot be understated: we believe multi-parameter persistence has a larger role to play in manifold learning. 
+Unfortunately, as mentioned prior, the compute barriers effective bar its use in practice.
+
+\subsubsection{Accelerating 2D persistence}\label{sec:empirical_klein}
+As we have outlined the computational theory of 2-parameter persistence and elucidated its relevance to our proposed move scheduling approach, we now demonstrate the efficiency of scheduling using the same high-contrast patch data set studied in~\cite{lee2003nonlinear} by evaluating the performance of various methods at computing the fibered barcode invariant via the parameterization from~\ref{sec:fibered_barcode_reparam}.
+
+Due to the aforementioned high complexity of the fibered barcode computation, we begin by working with a subset of the image patch data $\mathcal{X}$. 
+In particular, we combine the use of furthest-point sampling and proportionate allocation (stratified) sampling to sample landmarks $X \subset \mathcal{X}$ distributed within $n = 25$ strata.
 Each strata consists of the $(1/n)$-thick level set given by $\rho_{15}$, the density estimator.
 The use of furthest-point sampling gives us certain coverage guarantees that the geometry is approximately preserved within each level set, whereas the stratification ensures the original density of is approximated preserved as well.
-From this data set, we construct a Rips-(co)density bifiltration using $\rho_{15}$ equipped with the geodesic metric computed over the same $k = 15$ knn graph on $X$.
+From this data set, we construct a Rips-(co)density bifiltration using $\rho_{15}$ equipped with the geodesic metric computed over the same $k = 15$ k-nearest neighbor graph on $X$. 
 Finally, we record the number of column reductions needed to compute the fibered barcode at a variety of levels of coarsening using $\mathtt{pHcol}$, vineyards, and our moves approach.
@@
 
-\noindent
 As shown on the table, when the coarsening $\kappa$ is small enough, we're able to achieve a significant reduction in the number of total column operations needed to compute $\mathcal{T}$ compared to both vineyards and $\mathtt{pHcol}$.

### 2023-01-28 move_schedules.tex
Cleaned up reduction section

@@
 
- The main output of the reduction algorithm is a matrix decomposition $R = D V$, where the persistence diagram is encoded in $R$, and the generating cycles in the columns of $V$. 
- Given a filtration $(K, f)$ with $m$ simplices, one assembles the elementary boundary chains $\partial(\sigma)$ as columns ordered according to $f$ into a $m \times m$  \emph{filtration boundary matrix} $D$.
+ Given a filtration $(K, f)$ with $m$ simplices,  the main output of the reduction algorithm is a matrix decomposition $R = D V$, where the persistence diagram is encoded in $R$ and the generating cycles in the columns of $V$.
+ To begin the reduction, one starts by assembling the elementary boundary chains $\partial(\sigma)$ as columns ordered according to $f$ into a $m \times m$  \emph{filtration boundary matrix} $D$.
+Setting $V = I$ and $R = D$, one proceeds by performing elementary left-to-right column operations on $V$ and $R$ until the following invariants are satisfied:
+\textbf{ Decomposition Invariants:}
+ \begin{enumerate}[labelsep=3pt, topsep=2pt,itemsep=-0.25ex,parsep=1.0ex]\label{inv:decomposition}
+ 	\item[I1.] $R = D V$ where $D$ is the boundary matrix of the filtration $K$
+ 	\item[I2.] $V$ is full-rank upper-triangular
+ 	\item[l3.] $R$ is \emph{reduced}: if $\mathrm{col}_i(R) \neq 0$ and $\mathrm{col}_j(R) \neq 0$, then $\mathrm{low}_R(i) \neq \mathrm{low}_R(j)$ 
+ 	\end{enumerate} 
+ \vspace*{1em}
+ where $\mathrm{low}_R(i)$ denotes the largest row index of a non-zero entry in column $i$ of $R$. 
+We call the decomposition satisfying these three invariants \emph{valid}.
+Note that though the matrices $R$ and $V$ are not unique, the collection of persistent pairings are~\cite{zomorodian2005computing}.
+The persistence diagram of the corresponding filtration $K$ may be read off from lowest entries in $R$, once $R$ is in reduced form.
+
  It is at times more succinct to restrict to specific sub-matrices $D$ based on the homology dimension $p$, and so we write $D_p$ represent the $d_{p-1} \times d_p$ matrix representing the linear operator $\partial_p$ (the same notation is extended to $R$ and $V$).
-% The reduction algorithm computes the persistent homology from $D$ for all dimensions up to $\mathrm{dim}(K) - 1$.
-%  Alternatively, a single dimension $p \leq \mathrm{dim}(K) - 1$ can be computed, in which case one reduces a pair of sub-matrices $D_p$ and $ D_{p+1}$ of $D$,  where $D_p$ has dimension $d_{p-1} \times d_p$ and $D_{p+1}$ is $d_{p} \times d_{p+1}$.
-%In some situations the description of the ensuing algorithms is simpler when considering all dimensions simultaneously, while in others it is more succinct to consider dimension-specific matrices. 
-%In what follows $D$, $R$, and $V$ will always denote square $(m \times m)$ matrices, though we note that all algorithms discussed generalize to both cases. 
 We illustrate the reduction algorithm with an example pair $(R_1, V_1)$ below. 
@@
 \textbf{Example 2.1: Reduction} Consider a triangle with vertices $u,v,w$, edges $a = (u,w)$, $b = (v,w)$, $c = (u, v)$,
-and whose filtration order, read as inclusions from left to right, is given as $K = (u, v, w, a, b, c)$.
-The reduction algorithm begins by setting $R = D$ and $V = I$, followed by left-to-right column operations until invariants I3 is satisfied.
-If the $i$-th column of $R$ is nonzero, then $\mathrm{low}_R(i)$ 
-is the row index of its lowest nonzero entry.
-Using $\mathbb{Z}_2$ coefficients and homology in dimension 1 for simplicity, the reduction proceeds as follows:
+and whose filtration order is given as $K = (u, v, w, a, b, c)$. 
+%The reduction algorithm begins by setting $R = D$ and $V = I$, followed by left-to-right column operations until invariants I3 is satisfied.
+%If the $i$-th column of $R$ is nonzero, then $\mathrm{low}_R(i)$  is the row index of its lowest nonzero entry.
+Using $\mathbb{Z}_2$ coefficients for simplicity, the reduction proceeds to compute $H_1(K)$ as follows:
 \begin{displaymath}
@@
 Inspection of the reduction algorithm from~\cite{edelsbrunner2000topological} suggests that a loose upper bound for the reduction is $O(m^3)$, where $m$ is the number of simplices of the filtration---this bound is in fact tight~\cite{morozov2005persistence}.
-Nonetheless, many variations and optimizations to Algorithm~\ref{alg:reduce} have been proposed over the past decade, see~\cite{bauer2017phat, bauer2021ripser, chen2011persistent} for an overview.
-Despite these variations in computation, the resulting decomposition must satisfy the following invariants:
-  \vspace*{0.8em}
-  \\
-  \textbf{ Decomposition Invariants:}
- \begin{enumerate}[labelsep=3pt, topsep=2pt,itemsep=-0.25ex,parsep=1.0ex]\label{inv:decomposition}
- 	\item[I1.] $R = D V$ where $D$ is the boundary matrix of the filtration $K$
- 	\item[I2.] $V$ is full-rank upper-triangular
- 	\item[l3.] $R$ is \emph{reduced}: if $\mathrm{col}_i(R) \neq 0$ and $\mathrm{col}_j(R) \neq 0$, then $\mathrm{low}_R(i) \neq \mathrm{low}_R(j)$ 
- 	\end{enumerate} 
- \vspace*{1em}
-We call the decomposition satisfying these three invariants \emph{valid}. 
-%%\begin{proposition}[Persistent Pairing~\cite{zomorodian2005computing}]
-%%	Let $R = D V$, where $R$ is in reduced form and $V$ is full-rank upper triangular (invariants (1) and (2)  hold).
-Then the pairing $(\sigma_i, \sigma_j)$ defined by $i = \mathrm{low}_R(j)$ is a \emph{creator/destroyer (persistent) pair}.
-While the matrices $R,V$ satisfying (1) and (2) are not unique in general,  the persistent pairs $(i,j)$ are unique and define the persistence diagrams of $K$.
-%%\end{proposition}
-%\noindent
-Note that though the matrices $R$ and $V$ are not unique, the collection of persistent pairings are~\cite{zomorodian2005computing}.
-The persistence diagram of the corresponding filtration $K$ may be read off from lowest entries in $R$, once $R$ is in reduced form.
+Despite this high algorithmic complexity, the number of column operations has been observed to be super-linear in practice, due impart to the high sparsity and structure of $D$. 
+Moreover, many variations and optimizations to Algorithm~\ref{alg:reduce} have been proposed over the past decade, see~\cite{bauer2017phat, bauer2021ripser, chen2011persistent} for an overview. 
 % Important corollaries of the persistence pairing lemma is that the 
@@
 
-These rules prescribe certain column and row operations to apply to a given matrix decomposition either before, during, or after each individual transposition to ensure it is a valid decomposition.   
+These rules prescribe certain column and row operations to apply to a given matrix decomposition either before, during, or after each transposition to ensure the decomposition is valid.   
 Formally, let $S_{i}^j$ represent the upper-triangular matrix such that $A S_{i}^j$ is equivalent to the elementary operation that adds column $i$ of $A$ to column $j$ of $A$ and $S_{i}^j A$ is equivalent to adding row $j$ of $A$ to row $i$ of $A$.
@@
 \noindent 
-\textbf{Example}: Consider a small set of noisy points distributed around $S^1$ containing a few strong outliers, as shown on the left side of Figure~\ref{fig:bifiltration_ex2}.
+\textbf{Example 4.1}: Consider a small set of noisy points distributed around $S^1$ containing a few strong outliers, as shown on the left side of Figure~\ref{fig:bifiltration_ex2}.
 Filtering this data set with respect to the Rips parameter and the complement of a kernel density estimate yields a bifiltration whose various invariants are shown in the middle figure.

### 2023-01-29 move_schedules.tex
Refactored section 3.4.2 on the proxy minimization to streamline it.

@@
 \usepackage{array}
+\usepackage[inkscape={/usr/local/bin/inkscape -z -C}]{svg}
 %\usepackage{algorithm2e}
@@
 	\vspace{1em}
-	\includegraphics[height=1.15in]{circle_complex.png}
+	\includegraphics[height=1.25in]{circle_complex.png}
 \end{center}
@@
  	\centering
- 	\includegraphics[width=0.75\textwidth]{circle_vineyards.png}
+ 	\includegraphics[width=0.75\textwidth]{circle_vineyards_vector.pdf}
+%	\includesvg{circle_vineyards}
  	\caption{The cumulative column operations needed to compute persistence across the time-varying filtration of grayscale images.
@@
 Before proceeding, we require a few definitions.
-Given two fixed permutations $p, q \in S_m$ and a set $\Sigma \subseteq S_m$,  a common  problem is to find a sequence of permutations $s_1, s_2, \dots, s_d \in \Sigma$ whose composition satisfies:
+Given two fixed permutations $p, q \in S_m$ and a set allowable permutations $\Sigma \subseteq S_m$,  a common  problem is to find a sequence of permutations $s_1, s_2, \dots, s_d \in \Sigma$ whose composition satisfies:
 	\begin{equation}\label{eq:sorting}
@@
 In the latter case, a lower bound on $d$ is referred to as the \emph{distance} between $p$ and $q$ with respect to $\Sigma$.
-A sequence $S = \{s_1, s_2, \dots, s_d\} \subseteq \Sigma \subseteq S_m$ of operations mapping $p \mapsto q$ is sometimes called a \emph{sorting of $p$}. 
+A sequence $S = (s_1, s_2, \dots, s_d)$ of operations $s \in \Sigma \subseteq S_m$ mapping $p \mapsto q$ is sometimes called a \emph{sorting of $p$}. 
 When $p, q$ are interpreted as strings, distances defined with respect to a fixed $\Sigma \subseteq S_m$ are commonly referred to as \emph{edit distances}~\cite{bergroth2000survey}, which are denoted as $d_\Sigma(\cdot, \cdot)$, and the subsequent elementary operations on the strings are called \emph{edit operations}.
-Note the choice of $\Sigma$ defines the universe of operations and thus the type of edit distance being measured---otherwise if $\Sigma = S_m$, then $d_\Sigma(p, q) = 1$ for any $p\neq  q \in S_m$.
+We denote the space of sequences transforming $p \mapsto q$ using $d$ permutations in $\Sigma \subseteq S_m$ with $\Phi_\Sigma(p,q,d)$.
+Note the choice of $\Sigma$ defines the type of distance being measured---otherwise if $\Sigma = S_m$, then $d_\Sigma(p, q) = 1$ trivially for any $p\neq  q \in S_m$.
 
@@
 \begin{proposition}[Schedule Size]\label{prop:schedule_size}
-Let $K, K'$ denote distinct filtrations of a simplicial complex $S$ of size $\lvert S\rvert = m$.
-Then, the smallest size any move schedule $S^*$ transforming $K$ into $K'$ is given by: 
-$$ d = m - \lvert \mathrm{LCS}(K, K') \rvert $$ 
-Moreover, the size of $S^\ast$ can be determined in $O(m \log \log m)$ time. 
+Let $(K, f), (K, f')$ denote two filtrations of size $\lvert K \rvert = m$. Then, the smallest move schedule $S^*$ reindexing $f \mapsto f'$ has size: 
+$$ \lvert S^\ast \rvert = d = m - \lvert \mathrm{LCS}(f, f') \rvert $$ 
+where we use $\mathrm{LCS}(f, f')$ to denote the LCS of the permutations of $K$ induced by $f$ and $f'$.
+Moreover, $\lvert S^\ast \rvert = d$ can be determined in $O(m \log \log m)$ time. 
 \end{proposition}
@@
   \begin{theorem}\label{prop:sparsity_result}
-   Given a pair of distinct filtrations $K, K'$ of a simplicial complex $S$ and a decomposition $R = DV$ of $K$, the size of a minimal sequence $\mathcal{S} = \left( s_1, s_2, \dots, s_d \right)$ of cyclic `move' permutations $s_k = (i_k, j_k)$ satisfying:
+   Given a pair of filtrations $(K,f), (K,f')$ and a decomposition $R = DV$ of $K$, the size of a minimal sequence $\mathcal{S} = \left( s_1, s_2, \dots, s_d \right)$ of cyclic `move' permutations $s_k = (i_k, j_k)$ satisfying:
    	\begin{equation}\label{eq:rv_seq}
@@
    	\end{equation}
-   	can be determined in $O(m \log \log m)$ time and $O(m)$ space. 
+   	can be determined in $O(m \log \log m)$ time and $O(m)$ space, where $R' = D_d V_d$ denotes a valid decomposition of $(K, f')$. 
    	Moreover, if $i_k < j_k$ for all $k \in [d]$, then computing~\eqref{eq:rv_seq} requires $O(d\kappa)$ column operations, where:	  
-   	$$  \quad \kappa = \sum\limits_{k = 1}^d \, (s_{k} + t_{k}) $$
+   	$$  \quad \kappa = \sum\limits_{k = 1}^d \, (\lvert \mathbb{I}_{k} \rvert + \lvert \mathbb{J}_{k} \rvert) $$
 %	$$  \quad \kappa = \sum\limits_{k = 1}^d \, \lvert \mathbb{I}_{k}\rvert + \lvert \mathbb{J}_{k}\rvert $$
-	where $s_k, t_k$ are quantities from~\ref{prop:move_st} derived from the non-zeroes of $R_k, V_k$ in the sequence~\ref{eq:rv_seq}.
+	where the quantities $\lvert \mathbb{I}_{k} \rvert, \lvert \mathbb{J}_{k} \rvert$ of the intermediate $R_k, V_k$ are given in Proposition~\ref{prop:move_st}.
 Note that since $\kappa$ depends on the sparsity of the intermediate entries $V_1, V_2, \dots, V_d$ and $R_1, R_2, \dots, R_d$, the bound $O(d \kappa)$ is output-sensitive.
@@
 \subsubsection{Greedy approach}\label{sec:greedy} 
-Another flexibility granted by the move framework is that the cost of a move on a given $R = D V$ decomposition can be determined efficiently prior to any column operations. 
-Specifically, the cost of determining $\lvert \mathbb{I}\rvert + \lvert \mathbb{J}\rvert$ in Proposition~\ref{prop:move_st} cannot take more than $O(m)$ time with row-oriented sparse matrices. 
-Towards minimizing the number of column operations a schedule requires, it is natural to consider whether a greedy-type solution which chooses the minimal cost choice at every step yields an efficient strategy.
+Ideally, we would like to minimize the cost of a schedule $\mathcal{S} \in \Phi_\Sigma(p,q,d)$ directly, which recall is given by the number of non-zeros at certain entries in $R$ and $V$: 
+\begin{equation}\label{eq:schedule_cost}
+\mathrm{cost}(\mathcal{S}) = \sum\limits_{k=1}^d \lvert \mathbb{I}_k \rvert + \lvert \mathbb{J}_k\rvert
+\end{equation}
+where $\lvert \mathbb{I}\rvert + \lvert \mathbb{J}\rvert$ are the quantities from Proposition~\ref{prop:move_st}.
+One advantages to the moves framework is that the cost of a single move on a given $R = D V$ decomposition can be determined efficiently prior to any column operations---no more than $O(m)$ time with row-oriented sparse matrices.
+It is natural to consider whether a greedy-type solution which chooses the minimal cost choice at every step yields an efficient strategy.
 We give a counter-example below demonstrating that a greedy procedure may lead to arbitrarily bad behavior.
@@
 \subsubsection{Proxy objective}\label{sec:proxy_objective}
-Although the greedy approach from the last section can lead to arbitrarily high cost schedules, similar greedy-like strategies can yield computationally efficient heuristics in practice, even if they are suboptimal.
+Although the greedy approach from the last section can lead to arbitrarily high cost schedules, we find similar greedy-like strategies can yield computationally efficient heuristics in practice, even if they are suboptimal.
 Choosing the lowest-cost move takes $\approx O(dm)$---implying a greedy strategy would require $O(d^2 m)$ time---which is computationally viable as a heuristic when $d << m$. 
@@
 
-In light of the discussion above, we consider whether there exists an alternative \emph{proxy objective} to optimize that is both related to the cost of a schedule and not dependent on the entries in the decomposition.
-Obviously, the number of column operations needed to move a simplex from position $i$ in the filtration to position $j$ is upper bounded by $\lvert i - j \rvert$.
-In practice, due to the sparsity of both $R$ and $V$, this bound is very loose---the number of column operations is generally much smaller.
-Nonetheless, the complexities of the move operations discussed in Section~\ref{sec:moves} are dependent on the magnitude of $\lvert i - j \rvert$, and thus it is worth considering whether  \emph{minimizing the total displacement} would be a effective proxy objective.
-
-If one forms bijections between two given filtrations $p: K_0 \to [m], q: K_1 \to [m]$ to permutations of $[m]$ as before, then the Kendall-$\tau$ distance yields an upper bound on the number of column reductions needed to perform \emph{any} schedule.
-This follows from the fact that any inversion between $p$ and $q$ may induce a column operation, and the number of inversions between $p$ and $q$ is exactly $K_{\tau}(p, q)$.
-Another common distance for measuring the disarrangement between permutations is the \emph{Spearman distance}, defined as:
+We seek a \emph{proxy objective} that correlates with~\eqref{eq:schedule_cost} and does not explicitly depend on the entries in the decomposition, as minimizing these terms directly is difficult due to  the changing sparsity of the intermediate matrices $R_k, V_k$,
+Consider a strategy $\mathcal{S} \in \Phi(p,q,d)$ that minimizes the upper bound:
+\begin{equation}\label{eq:upper_bound_move_cost}
+\mathrm{cost}(\mathcal{S}) = \sum\limits_{k=1}^d ( \lvert \mathbb{I}_k \rvert + \lvert \mathbb{J}_k\rvert ) \leq  \sum\limits_{k=1}^d 2 \lvert i_k - j_k \rvert 
+\end{equation}
+In practice, this bound is very loose due to the sparsity of both $R$ and $V$.
+Nonetheless, the complexities of the move operations discussed in Section~\ref{sec:moves} depend on $\lvert i - j \rvert$, and minimizing~\eqref{eq:upper_bound_move_cost} has the intuitive interpretation as minimizing \emph{net displacement}.
+%The Kendall-$\tau$ distance yields an upper bound on the number of column reductions needed to perform any schedule. 
+A similar $\ell_1$-type distance for measuring the disarrangement between permutations is the \emph{Spearman distance}, defined as:
 \begin{equation}\label{eq:spearman_dist}
-F(p, q) = \sum\limits_{i \in [m]} \lvert p^{-1}(i) - q^{-1}(i) \rvert
+F(p, q) = \sum\limits_{i =1}^m \lvert p(i) - q(i) \rvert =  \sum\limits_{i =1}^m \lvert i - (q^{-1} \circ p)(i) \rvert
 \end{equation}
-Intuitively, the Spearman distance can be interpreted as $\ell_1$-type distance between two permutations.
-Like the Kendall-$\tau$ distance, it is a metric, it is invariant under relabeling, and the distance between pairs of uniformly random permutations $p, q \sim S_m$ are asymptotically normal.
-Indeed, Diaconis et al.~\cite{diaconis1977spearman} showed the the Spearman distance between any two permutation $p, q \in S_m$ approximates their Kendall-$\tau$ distance within a factor of two:
+The Spearman distance shares any similarities with $K_\tau$: it is a metric on $S_m$ that is invariant under consistent relabeling. 
+%and the distance between pairs of uniformly random permutations $p, q \sim S_m$ are asymptotically normal. 
+Indeed, Diaconis et al.~\cite{diaconis1977spearman} showed the the Spearman distance approximates $K_\tau$ within a factor of two: 
 \begin{equation}\label{eq:diaconis_inequality}
@@
 \end{equation}
-Although very similar to the Kendall-$\tau$ distance, the Spearman distance conveys certain computational advantages not inherited by the Kendall distance due to its simplicity. 
-For example, the Spearman distance can be computed in $O(m)$ time.
-
+In contrast to $K_\tau$, the Spearman can be computed in $O(m)$ time, making it an attractive alternative quantity in the well-known \emph{rank aggregation problem}.
+% Indeed, it is used...
 %, and 
@@
 In contrast, a 2-approximation may be obtained by replacing the Kendall distance with the Spearman distance, as in this case the problem reduces to finding a perfect matching in a bipartite graph, which can be accomplished in polynomial time.
-
-Next we consider the difficulty of obtaining a sorting $S$ of size $d = m - \mathrm{LCS}(p,q)$ which has minimal total displacement. 
-That is, incorporating the Spearman distance in the proxy objective. 
-Formally, given $p, q \in S_m$ and some sorting  $S = (s_1, \dots, s_d)$, let $\hat{s}_i =  s_i \circ \cdots \circ s_2 \circ s_1$ denote the composition of the first $i$ permutations in $S$, where $i < d$.
-To incorporate the notion of displacement as an additive minimization problem, we consider minimizing an upper bound on the Spearman distance by decomposing it into a sum of $d-1$ local distances:
+To adapt $F(\cdot, \cdot)$ to sortings, we consider the additive minimization problem which minimizes the following upper bound on $F(p, q)$:
 \begin{equation}\label{eq:sorting_additive}
-F(p, q) \leq \sum\limits_{i=1}^{d-1} F(\hat{s}_i \circ p, \hat{s}_{i+1} \circ p)
+\hat{F}_{\mathcal{S}}(p, \iota) = \sum\limits_{i=1}^{d-1} F(\hat{s}_i \circ p, \hat{s}_{i+1} \circ p) \geq F(p, \iota)
 \end{equation}
-where equality is only achieved with schedules $S$ for which the displacement of each symbol between its initial position in $p$ and its final position in $q$ is non-increasing with every application of $s_i$.
-Note this not guaranteed if $S$ was derived using the strategy from section~\ref{sec:move_schedules}.
+where $\hat{s}_i =  s_i \circ \cdots \circ s_2 \circ s_1$ denote the composition of the first $i$ permutations of a sorting  $S = (s_1, \dots, s_d)$ that maps $p \mapsto \iota$.
+Note that equality in~\eqref{eq:sorting_additive} is achieved when the displacement of each symbol between its initial position in $p$ to its value is non-increasing with every application of $s_i$, which in general not guaranteed using the schedule construction method derived in section~\ref{sec:move_schedules}.
 
@@
 All three schedules were generated from the same $\mathrm{LCS}(p,q) = (4\,7\,8)$ and all three schedules transform $p \mapsto q$ in $d = 6$ moves. 
-In this example, $S_1$ matches the minimal number of crossings amongst all possible schedules,  since $K_\tau(p,q) = 21$. \\ \\
-In general,  the $k$-layer crossing minimization problem for $k$ sets of permutations  is NP-hard for $k \geq 4$~\cite{biedl2009complexity}.
-As a result, we do not expect there to be a polynomial time algorithm for optimally achieving this proxy objective.
+In this example, $S_1$ matches the minimal number of crossings amongst all possible schedules,  since $K_\tau(p,q) = 21$. 
+\\
+\\
+The example above suggests minimizing~\eqref{eq:sorting_additive} is reminiscent of the $k$-layer crossing minimization problem for $k$ sets of permutations, which is NP-hard for $k \geq 4$~\cite{biedl2009complexity}.
+%As a result, we do not expect there to be a polynomial time algorithm for optimally achieving this proxy objective.  
 % Since $\mathcal{D}$ is fixed, the complexity of a greedy strategy depends entirely on the complexity of minimizing $F(\hat{s}_i, \hat{s}_{i+1})$ at each greedy step.  
@@
 %\\
+
 %% TODO: elaborate on monotonicity, describe segment tree
-However, the simplicity of both the Spearman distance and the types of permutations we consider here enables certain computational advantages.
+The simplicity of both the Spearman distance and the types of permutations we consider here enables certain computational advantages.
 Suppose we begin with an array $\mathcal{A}$ of size $m$ which provides $O(1)$ access and modification, initialized with the (signed) \emph{displacement} of every element in $p$ to its corresponding position in $q$.

### 2023-01-30 move_schedules.tex
Added performance images

@@
 %% TODO: elaborate on monotonicity, describe segment tree
-The simplicity of both the Spearman distance and the types of permutations we consider here enables certain computational advantages.
+We propose a heuristic strategy to minimize~\eqref{eq:sorting_additive} which we observed is both efficient to compute quite effective in practice.
+The heuristic is inspired by the simplicity of computing the Spearman distance between cyclic permutations at the schedule construction phase.
 Suppose we begin with an array $\mathcal{A}$ of size $m$ which provides $O(1)$ access and modification, initialized with the (signed) \emph{displacement} of every element in $p$ to its corresponding position in $q$.
@@
 \begin{equation}\label{eq:greedy_step}
-	s_{\text{greedy}} = \argmin_{\text{ valid } s \in S_\mathcal{D}} F(s\circ p, q)
+	s_{\text{greedy}} = \argmin_{\text{ valid } s \in \Sigma} F(s\circ p, \iota)
 \end{equation}  
@@
 \section{Applications and Experiments}\label{sec:results}
+\subsection*{Movies}
+Move data description
+\begin{figure}
+	\centering
+	\includegraphics[width=0.48\textwidth]{varying_d.pdf}
+	\includegraphics[width=0.48\textwidth]{varying_nd.pdf}
+	\caption{Performance}
+\end{figure}
+
 \subsection*{Crocker stacks}

### 2023-01-30 move_schedules.tex
full paper modulo the appendix and final edits from jose

@@
 	The standard procedure for computing the persistent homology of a filtered simplicial complex with $m$ simplices is the matrix reduction algorithm.
-Its output is a particular decomposition of the total boundary matrix, from which the persistence diagrams and generating cycles can be derived.
+Its output is a particular decomposition of the total boundary matrix, from which the persistence diagrams and generating cycles are derived.
 	Persistence diagrams are known to vary continuously with respect to their input, motivating the  study of their computation for time-varying filtered complexes.
 Computationally, simulating persistence dynamically can be reduced to maintaining a valid decomposition under adjacent transpositions in the filtration order.
-	Since there are $O(m^2)$ such transpositions, this maintenance procedure exhibits limited scalability and often is too fine for many application contexts.
-	As our first result, we propose a coarser strategy for maintaining the decomposition over a 1-parameter family of filtrations that take $O(m^2)$ preprocessing time uses only $O(m)$ space. 
-	We also show that this maintenance procedure yields a factor-2 net reduction in column operations over the state of the art approach. 
-	By reduction to a particular longest common subsequence problem, we give a modification of this strategy that not only can be performed in $O(m \log \log m)$ time, but yields a storage bound that is actually sublinear in expectation.
-	Finally, we show experimentally the decrease in operations to compute diagrams across a family of filtrations is proportional to the difference between the expected quadratic number of states and the proposed sublinear coarsening.
-	Applications to multi-dimensional persistence and crocker stacks are also presented.
+	Since there are $O(m^2)$ such transpositions, this maintenance procedure exhibits limited scalability and often is too fine for many applications.
+%	As our first result, we propose a coarser strategy for maintaining the decomposition over a 1-parameter family of filtrations that takes $O(m^2)$ preprocessing time uses only $O(m)$ space. 
+%	We also show that this maintenance procedure yields a factor-2 net reduction in column operations over the state of the art approach. 
+We propose a coarser strategy for maintaining the decomposition over a 1-parameter family of filtrations that requires only $O(m \log \log m)$ time and $O(m)$ space. 
+By reduction to a particular longest common subsequence problem, we show the storage needed to employ this strategy is actually sublinear in expectation. 
+	Exploiting this connection, we show experimentally the decrease in operations to compute diagrams across a family of filtrations is proportional to the difference between the expected quadratic number of states and the proposed sublinear coarsening.
+	Applications to movie data, dynamic metric spaces, and multi-parameter persistence are also presented.
 }
@@
 where $\ast$ is substituted with elementary column operations that repair the permuted decomposition.
-It is known how to linearly interpolate $f \mapsto f'$ using $d \sim O(m^2)$ updates to the decomposition, where each update is comprised of at most two column operations~\cite{cohen2006vines}.
-Since each column operation takes $O(m)$, the complexity performing this update is $O(m^3)$, which is efficient if all $d$ decompositions are needed.
-Otherwise, if only $(R', V')$ is needed (and none of the $d -1$ other decomposition), updating $R \mapsto R'$ using the approach from~\cite{cohen2006vines} matches the complexity of computing $R' = D' V'$ independently.
+It is known how to linearly interpolate $f \mapsto f'$ using $d \sim O(m^2)$ updates to the decomposition, where each update  requires at most two column operations~\cite{cohen2006vines}.
+Since each column operation takes $O(m)$, the complexity of reindexing $f \mapsto f'$ is $O(m^3)$, which is efficient if all $d$ decompositions are needed.
+Otherwise, if only $(R', V')$ is needed (and none of the intermediate $d -1$ decompositions), updating $R \mapsto R'$ using the approach from~\cite{cohen2006vines} matches the complexity of computing $R' = D' V'$ independently.
 
@@
  where each permutation $s_i$ is a particular type of cyclic permutation (see section~\ref{sec:moves_setting}).
-If $i_\ast < j_\ast$, our first result is that this sequence can be computed in using $O(d\kappa)$ column operations, where:
+If $i_\ast < j_\ast$, our first result extends~\cite{busaryev2010tracking} by showing that~\eqref{eq:rv_seq} can be computed in using $O(\kappa)$ column operations, where:
 \begin{equation}
@@
 where the quantities $\lvert \mathbb{I}_{k} \rvert$ and $\lvert \mathbb{J}_{k} \rvert$ depend on the sparsity of the $V_k$ and $R_k$ matrices, respectively, and $d \sim O(m)$ is a constant that depends on how similar $f$ and $f'$ are.
-The advantage of this result is that it depends explicitly on the sparsity pattern of the decomposition itself, which we take advantage of in Section~\ref{sec:schedule_cost}.
-Moreover, if the vineyards algorithm requires $C$ column operations to obtain~\eqref{eq:rv_seq}, we show (Proposition~\ref{prop:factor2}) that our approach obtains the same sequence using at most $M$ column operations, where:
-\begin{equation}
-	M \leq \tfrac{1}{2}  C
-\end{equation}
-Our second result turns towards the computation of $\mathcal{S}$ itself, and towards bounding its size $\lvert \mathcal{S} \rvert = d$. 
-The complexity of determining $\mathcal{S}$ depends on the type of permutations considered; many edit distances exhibit tight quadratic growth rates and tend to require at least quadratic time (or worse) to compute~\cite{labarre2013lower}.
-We consider a special set of cyclic permutations whose sequences scale linearly in $m$ for all permutations in $S_m$: surprisingly, we find that the \emph{smallest} such sequence of these permutations which map $f \mapsto f'$ has size $d$ (Proposition~\ref{prop:schedule_size}), where:
+The advantage of this result is that it depends explicitly on the sparsity pattern of the decomposition itself and is thus output sensitive, which we take advantage of in Section~\ref{sec:schedule_cost}.
+%Surprisingly, due to a refined analysis of restore subprocedure from~\cite{busaryev2010tracking} we give in section~\ref{sec:continuous_setting}, we find that move operations are actually cheaper than vineyards, see Proposition~\ref{prop:factor2}.
+%if the vineyards algorithm requires $T$ column operations to obtain~\eqref{eq:rv_seq}, we show (Proposition~\ref{prop:factor2}) that obtains the same sequence using at most $M$ column operations, where:
+%\begin{equation}
+%	M \leq \tfrac{1}{2}  T
+%\end{equation}
+
+Our second result turns towards upper bounding $d = \lvert \mathcal{S} \rvert$ and the complexity of constructing $\mathcal{S}$ itself. 
+%The complexity of determining $\mathcal{S}$ depends on the type of permutations considered; many edit distances exhibit tight quadratic growth rates and tend to require at least quadratic time (or worse) to compute~\cite{labarre2013lower}.
+By reinterpreting a special set of cyclic permutations as edit operations on strings, we find that the \emph{smallest} such sequence that maps $f \mapsto f'$ has size $d$ (Proposition~\ref{prop:schedule_size}), where:
 \begin{equation}
@@
 \end{equation}
-Moreover, we also show that the information needed to construct any $\mathcal{S}$ with optimal size $\lvert \mathcal{S} \rvert = d$ can be computed in $O(m \log \log m)$ preprocessing time and $O(m)$ memory.
-Although $d$ clearly is $O(m)$ for pathological inputs, we show that $d \sim m - \sqrt{m}$ in expectation for random filtrations (Corollary~\ref{cor:expectation}) is not too pessimistic for truly random filtrations, and we give empirical results suggesting $d$ is tends to be much smaller for time-varying filtrations.
-\\
+We also show that the information needed to construct any $\mathcal{S}$ with optimal size can be computed in $O(m \log \log m)$ preprocessing time and $O(m)$ memory.
+Although $d$ clearly is $O(m)$ for pathological inputs, we give evidence  that $d \sim m - \sqrt{m}$ in expectation for random filtrations is not too pessimistic for truly random filtrations (Corollary~\ref{cor:expectation}), and we give empirical results suggesting $d$ is tends to be much smaller for time-varying filtrations.
+
+%Finally, we perform extensive experiments 
 % given a decomposition $R = \partial_{K} V$ of a simplicial filtration $K$, algorithm~\ref{alg:schedule} can produce such a sequence $\mathcal{S} = \left( s_1, s_2, \dots, s_d \right)$
@@
 \noindent 
-\textbf{Outline:} The remainder of the paper is organized as follows: we review and establish the notations we will use to describe simplicial complexes, persistent homology, and dynamic persistence in Section \ref{sec:background}.
+\textbf{Outline:} The paper is organized as follows: we review and establish the notations we will use to describe simplicial complexes, persistent homology, and dynamic persistence in Section~\ref{sec:background}.
 We also cover the reduction algorithm (designated here as \texttt{pHcol}), the vineyards algorithm, and the set of \emph{move}-related algorithms introduced in~\cite{busaryev2010tracking}, which serves as the starting point of this work.
 In Section \ref{sec:move_schedules} we introduce our move schedules  and provide efficient algorithms to make this alternative viable. 
-In Section \ref{sec:results} we present applications of the proposed method. 
-In particular, to the computation of crocker stacks from flock simulations, and to 2-dimensional persistence on the Mumford data set of image patches from natural images. 
+In Section \ref{sec:results} we present applications of the proposed method, including the computation of crocker stacks from flock simulations and of a 2-dimensional persistence invariant on a data set of image patches derived from natural images.
 In Section \ref{sec:conclusion} we conclude the paper by discussing other possible applications and future work.
@@
 Given a decomposition $R = DV$ of $K$, a \emph{schedule}  $\mathcal{S} = (s_1, s_2, \dots, s_d )$ is a sequence of permutations mapping satisfying:
-	\begin{equation}\label{eq:rv_seq}
+	\begin{equation}\label{eq:rv_seq_schedule}
    		R = D_0 V_0 \overset{s_1}{\to} D_1 V_{1} \overset{s_2}{\to} \dots \overset{s_d}{\to} D_d V_{d} = R'
@@
 \begin{proposition}\label{prop:factor2}
-	Let $K$ denote a filtration of size $\lvert K \rvert = m$ with valid decomposition $R = D V$, $T$ the maximum number of $O(m)$ operations required to execute schedule $S_F$ using vineyards, and $M$ be the maximum number of $O(m)$ operations required to execute the coarsened schedule $\widetilde{S}_F$ with moves.
-Then:
-	$$ M \leq {\textstyle \frac{1}{2} \, T} $$ 
+	Let $K$ denote a filtration of size $\lvert K \rvert = m$ with decomposition $R = D V$, $T$ the maximum number of $O(m)$ operations required to execute $S_F$ via vineyards, and $M$ be the maximum number of $O(m)$ operations required to execute $\widetilde{S}_F$ with moves.
+Then, the inequality $ M \leq {\textstyle \frac{1}{2} \, T}$ holds.
 \end{proposition}
@@
 However, it's not immediately clear whether this bound is tight.
-Moreover, we expect  investigating the tightness of this bound may lead to exploitable structure.
+In this section, we investigate the tightness of this bound by revisiting the problem from a combinatorial perspective.
 % or whether this strategy of generating schedules---which we call \emph{naive strategy}---is performant.  
@@
 
-Before proceeding, we require a few definitions.
+Let $S_m$ denote the symmetric group.
 Given two fixed permutations $p, q \in S_m$ and a set allowable permutations $\Sigma \subseteq S_m$,  a common  problem is to find a sequence of permutations $s_1, s_2, \dots, s_d \in \Sigma$ whose composition satisfies:
@@
 A sequence $S = (s_1, s_2, \dots, s_d)$ of operations $s \in \Sigma \subseteq S_m$ mapping $p \mapsto q$ is sometimes called a \emph{sorting of $p$}. 
-When $p, q$ are interpreted as strings, distances defined with respect to a fixed $\Sigma \subseteq S_m$ are commonly referred to as \emph{edit distances}~\cite{bergroth2000survey}, which are denoted as $d_\Sigma(\cdot, \cdot)$, and the subsequent elementary operations on the strings are called \emph{edit operations}.
+When $p, q$ are interpreted as strings, these operations $s \in \Sigma$ are called \emph{edit operations}.
+The minimal number of edit operations $d_\Sigma(p, q)$ needed to sort $p \mapsto q$ with respect to $\Sigma$ is referred to as the \emph{edit distance}~\cite{bergroth2000survey} between $p$ and $q$.
 We denote the space of sequences transforming $p \mapsto q$ using $d$ permutations in $\Sigma \subseteq S_m$ with $\Phi_\Sigma(p,q,d)$.
@@
 
-Perhaps surprisingly, small changes to set of allowable edit operation $\Sigma$ can dramatically affect both the size of $d_\Sigma(p,q)$ and the difficulty of obtaining a minimal sorting. 
+Perhaps surprisingly, small changes to set of allowable edit operations $\Sigma$ dramatically affects both the size of $d_\Sigma(p,q)$ and the difficulty of obtaining a minimal sorting. 
 For example, while sorting by transpositions and reversals is NP-hard and sorting by prefix  transpositions is unknown, there are polynomial time algorithms for sorting by block interchanges, exchanges, and prefix exchanges~\cite{labarre2013lower}.
@@
 \begin{corollary}\label{cor:expectation}
-If $K, K'$ are random filtrations of a common simplicial complex $S$ of size $m$, then the size of longest common subsequence of simplices between $K,K'$ is no larger than $m - \sqrt{m}$ in expectation, with probability $1$ as $m \to \infty$.
+If $(K, f), (K,f')$ are random filtrations of a common complex $K$ of size $m$, then the size of longest common subsequence of simplices between $f,f'$ is no larger than $m - \sqrt{m}$ in expectation, with probability $1$ as $m \to \infty$.
 \end{corollary}
@@
    Given a pair of filtrations $(K,f), (K,f')$ and a decomposition $R = DV$ of $K$, the size of a minimal sequence $\mathcal{S} = \left( s_1, s_2, \dots, s_d \right)$ of cyclic `move' permutations $s_k = (i_k, j_k)$ satisfying:
-   	\begin{equation}\label{eq:rv_seq}
+   	\begin{equation}\label{eq:rv_seq_theorem}
    		R = D_0 V_0 \overset{s_1}{\to} D_1 V_{1} \overset{s_2}{\to} \dots \overset{s_d}{\to} D_d V_{d} = R'
@@
    	can be determined in $O(m \log \log m)$ time and $O(m)$ space, where $R' = D_d V_d$ denotes a valid decomposition of $(K, f')$. 
-   	Moreover, if $i_k < j_k$ for all $k \in [d]$, then computing~\eqref{eq:rv_seq} requires $O(d\kappa)$ column operations, where:	  
+   	Moreover, if $i_k < j_k$ for all $k \in [d]$, then computing~\eqref{eq:rv_seq_theorem} requires $O(d\kappa)$ column operations, where:	  
    	$$  \quad \kappa = \sum\limits_{k = 1}^d \, (\lvert \mathbb{I}_{k} \rvert + \lvert \mathbb{J}_{k} \rvert) $$
@@
 In contrast, a 2-approximation may be obtained by replacing the Kendall distance with the Spearman distance, as in this case the problem reduces to finding a perfect matching in a bipartite graph, which can be accomplished in polynomial time.
-To adapt $F(\cdot, \cdot)$ to sortings, we consider the additive minimization problem which minimizes the following upper bound on $F(p, q)$:
+To adapt $F(\cdot, \cdot)$ to sortings, we consider the additive minimization problem which minimizes the following upper bound on $F(p, \iota)$:
 \begin{equation}\label{eq:sorting_additive}
@@
 That is, if $s_\sigma$ represents a move permutation moving $\sigma$ from position $i$ to position $j$, then relative to the previous permutation $p$, the displacement of $\sigma$ in $s_\sigma \circ p$ changes by $\lvert i - j \rvert$ and the displacement of any symbol between $i$ and $j$ changes by $\pm 1$.
-All other symbol displacements are unchanged.
+All other symbols are unchanged.
 Thus, if we replace $\mathcal{A}$ with a data structure  supporting $O(\log m)$ access time to aggregate information (i.e.
@@
 \section{Applications and Experiments}\label{sec:results}
-\subsection*{Movies}
-Move data description
+\subsection{Movies}
+A common application of persistence is to characterizing topological structure in image data.
+Since a  set of ``snapshot'' frame of a movie can be equivalently thought of as discrete 1-parameter family, our framework provides a natural extension of the typical image analysis to video data.
+To demonstrate the benefit of using minimal sized scheduled and the scalability of the greedy approach proposed in section~\ref{sec:proxy_objective}, we re-use the movie data from~\ref{sec:motivation} as a baseline benchmark.
+We perform two performance tests: one to  one to test the impact of shrinking the number of permutations a given reindexing operation needs and one to test the asymptotic behavior of the greedy approach and.
+
+ In the first test, we fix a grid size of $9 \times 9$ and record the cumulative number of column operations needed to simulate persistence dynamically across 25 evenly-spaced time points using a variety of scheduling strategies.
+ The primary three strategies we test are the greedy approach from section~\ref{sec:proxy_objective}, the ``naive'' approach which uses upwards of $O(m)$ move permutations via selection sort, and a third strategy which interpolates between the two.
+ In particular, if $d = m - \lvert \mathrm{LCS}(f, f') \rvert$ we use a parameter $\alpha \in [0,1]$ to choose $m - \alpha \cdot d$ random symbols to move using the same construction method outlined in section~\ref{sec:schedule_construction}.
+When $\alpha = 0$, the strategy reduces to using selection sort to reindex $f \mapsto f'$ using a random ordering of simplices; otherwise, $\alpha = 1$ reduces to using a minimal sized schedule (with a random ordering of simplices).
+The results are summarized in the left graph on Figure~\ref{fig:movie_perf}.
+To get an idea of the variation of the random sampling, we run 10 independent iterations and shade the upper and lower bounds of the schedule costs, which is shown around the mean schedule cost depicted by the solid lines in Figure~\ref{fig:movie_perf}.
+ As one can see from the figure, while using less move operations (lower $\alpha$) does progressively reduce column operations, constructing random schedules of minimal size is no more competitive than the naive selection sort strategy.
+This suggests that efficient schedule construction needs to account for the structure of performing several permutations in sequence, like the greedy heuristic we introduced, to yield a performance boost.
+
 \begin{figure}
@@
 	\includegraphics[width=0.48\textwidth]{varying_nd.pdf}
-	\caption{Performance}
+	\caption{Performance data }
+	\label{fig:movie_perf}
 \end{figure}
 
-\subsection*{Crocker stacks}
+In the second test, we aim to measure the asymptotics of our greedy LCS-based approach.
+To do this, we generated 8 movie data sets again of the expanding annulus outlined in section~\ref{sec:motivation}, each of increasing grid sizes of $5 \times 5$, $6 \times 6$, $\dots$, $12 \times 12$.
+For each data set, we simulate persistence over the duration of the movie, again testing five evenly spaced settings of $\alpha \in [0,1]$---the results are shown in the right plot of Figure~\ref{fig:movie_perf}.
+On the vertical axis, we plot the total number of column operations needed to simulate the movie again across 25 evenly-spaced time points \emph{as a ratio} of the data set size ($m$); we also show the regression curves one obtains from considering each schedule size.
+As one can see from the Figure, the cost of using the greedy heuristic tends to increase sub-linearly as a function of the data set size, suggesting the move scheduling approach is indeed quite scalable.
+Moreover, using schedules with minimal size tended to be cheaper than otherwise, confirming our initial hypothesis that repairing the decomposition less can lead to substantial reductions at runtime.
+
+\subsection{Crocker stacks}
 There are many challenges to characterizing topological behavior in dynamic settings.
@@
 This dependence only on the Betti numbers makes crocker plots easier to compute than standard persistence, however what one gains in efficiency one loses in stability; it is known that Betti curves are inherently unstable with respect to small fluctuations about the diagonal of the persistence diagram.
-
-\begin{figure}[ht]
-	\centering
-	\includegraphics[width=\textwidth, height=16.35em]{boid_sim_results.png}
-	\caption{ One the left, the cumulative number of column operations (log-scale) of the three baseline approaches tested.
-On the right, the normalized $K_\tau$ between adjacent filtrations depicts the coarseness of the discretization---about $5\%$ of the $\approx O(m^2)$ simplex pairs between adjacent filtrations are discordant.}
-	\label{fig:boid_sim_results}
-\end{figure}
 
@@
 As depicted in Figure~\ref{fig:boid_sim_results}, our move strategy is far more efficient than both vineyards and the naive \texttt{pHcol} strategies.
+
+\begin{figure}[ht]
+	\centering
+	\includegraphics[width=\textwidth, height=16.35em]{boid_sim_results.png}
+	\caption{ One the left, the cumulative number of column operations (log-scale) of the three baseline approaches tested.
+On the right, the normalized $K_\tau$ between adjacent filtrations depicts the coarseness of the discretization---about $5\%$ of the $\approx O(m^2)$ simplex pairs between adjacent filtrations are discordant.}
+	\label{fig:boid_sim_results}
+\end{figure}
 
@@
 
-There are many possible applications of our work beyond the example application of generating crocker stacks and simulating multidimensional persistence.
-As discussed in sections~\ref{sec:related_work}, examples include:
+There are many possible applications of our work beyond the ones discussed in section~\ref{sec:results}. As mentioned in section~\ref{sec:related_work}, examples include: 
 \begin{enumerate}
@@
 \end{enumerate}
-Indeed, we envision our approach as potentially useful to any situation where the structure of interest is a parameterized family of filtrations and the goal involves computing the persistent homology for each filtration in the family.
+Indeed, we see our approach as potentially useful to any situation where the structure of interest is a parameterized family of persistence diagrams.
 Areas of particular interest include time-series analysis and dynamic metric spaces~\cite{kim2020spatiotemporal}.
@@
 The simple and combinatorial nature of our approach does pose some limitations to its applicability.
-For example, there may be better bounds or algorithms available if stronger assumptions can be made on how the filtration is changing with time.
-Moreover, if the filtration $K$ shares little similarity to the ``target'' filtration $K'$, then the overhead of reducing the simplices from $K' \setminus K$ appended to the decomposition derived from $K$ may be large enough to motivate simply computing the decomposition at $K_1$ independently, especially if parallel processors are available.
-Indeed, in this setting the size of the LIS may very well be on the order $\sim m - \sqrt{m}  = O(m)$, rendering the efficiencies gained using our approach as negligible.
-Thus, our approach is primarily useful if the filtrations in the parameterized family is ``nearby'' in the combinatorial sense.
+For example, better bounds or algorithms may be obtainable if stronger assumptions can be made on how the filtration is changing with time.
+Moreover, if the filtration $(K, f)$ shares little similarity to the ``target'' filtration $(L,f')$, then the overhead of reducing the simplices from $L \setminus K$ appended to the decomposition derived from $K$ may be large enough to motivate simply computing the decomposition at $L$ independently, especially if parallel processors are available.
+Our approach is primarily useful if the filtrations in the parameterized family is ``nearby'' in the combinatorial sense.
 
 From an implementation perspective, one non-trivial complication of our approach is its heavy dependence on a particular sparse matrix data structure which permits permuting both the row and columns of a given matrix in at most $O(m)$ time.
-Our use of this data structure is motivated by the fact that, like vineyards and unlike the standard persistence algorithm, the $R = DV$ decomposition is permuted with every $O(m)$ operation.
-Indeed, as shown with the natural images example in section~\ref{sec:results}, there are often more permutation operations being applied than there are column reductions.
-In the more standard \emph{compressed} sparse matrix representations (i.e.
-CSC, CSR, etc.), permuting both the rows and columns generally takes at most $O(Z)$ time, where $Z$ is the number of non-zero entries.
-If the particular filtration has many cycles, permuting these compressed representations can be quite expensive.
+As shown with the natural images example in section~\ref{sec:results}, there are often more permutation operations being applied than there are column reductions.
+In the more standard \emph{compressed} sparse matrix representations\footnote{
+By ``standard,'' we mean any of the common sparse representations used in scientific computing packages, like SciPy's sparse module (\href{https://docs.scipy.org/doc/scipy/reference/sparse.html} 
+))
+}, permuting both the rows and columns generally takes at most $O(Z)$ time, where $Z$ is the number of non-zero entries, which can be quite expensive if the particular filtration has many cycles.
 As a result, the more complex sparse matrix representation from~\cite{cohen2006vines} is necessary to be efficient in practice.
@@
 \subsubsection{Reduction Algorithm}
-\begin{algorithm}[t]
+\begin{algorithm}[h!]
 	\setstretch{1.15}

### 2023-01-31 move_schedules.tex
Organized appendix, small typos changes

@@
 \subsection{Continuous setting}\label{sec:continuous_setting}
-In the continuous time case (i.e., vineyards), we are given a homotopy $F : K \times [0,1] \to \mathbb{R}$ interpolating between  a pair of filtrations $K, K'$ from some underlying complex $S$, with simplices ordered according to $f = F(\cdot, 0)$ and $f' = F(\cdot, 1)$, respectively.
-As the choice of homotopy $F$ determines the number of transpositions that transform $K$ into $K$ via $F$, bounding this number requires assumptions on $F$. 
+In the continuous time case, we are given a homotopy $F : K \times [0,1] \to \mathbb{R}$ interpolating between  a pair of filtrations $(K, f), (K,f')$ satisfying $f = F(\cdot, 0)$ and $f' = F(\cdot, 1)$, respectively.
+As the choice of homotopy $F$ determines the number of transpositions that transform $f$ into $f'$ via $F$, bounding this number requires assumptions on $F$. 
 If we assume the curves $\tau \mapsto F(\cdot, \tau)$ are in general position in $[0,1]\times \mathbb{R}$ and that each pair of curves cross  at most once, then the  functions $\tau\mapsto F(\cdot, \tau)$ belong to a class of $x$-monotone curves called \emph{pseudo-segments}.
@@
 
-If one has two distinct filtrations $K,K'$ both derived from the same underlying complex $S$ of size $m = \lvert S \rvert$, observe the simplexwise representations of both $K$ and $K'$ can be thought of as elements of the symmetric group $S_m$.
-Moreover, as the application of permutations to $K$ implies additional elementary column operations to the underlying matrix decomposition, we think of the action of any set of such permutations on the initial decomposition $R=DV$ of $K$ as a \emph{schedule} with respect to the pair $K,K'$.
+Given two filtrations $(K,f), (K,f')$, observe their simplexwise representations induced by $f$ and $f'$ can be thought of as elements of the symmetric group $S_m$.
+As applying permutations to $K$ implies additional column operations on the underlying matrix decomposition, we think of any set of such permutations acting on the initial decomposition $R=DV$ of $(K,f)$ as a \emph{schedule} with respect to the pair $f,f'$.
 We formalize this with a definition.
 \begin{definition}[Schedule]\label{def:schedule}
-Given a pair of filtrations $K, K'$ from an underlying complex $S$ and a fixed bijection $\rho : K \to K'$, let $\rho^\ast : [m] \to [m]$ denote its induced bijection on $S_m$.
-Given a decomposition $R = DV$ of $K$, a \emph{schedule}  $\mathcal{S} = (s_1, s_2, \dots, s_d )$ is a sequence of permutations mapping satisfying:
+Given a pair of filtrations $(K, f), (K,f)$ 
+%and a fixed bijection $\rho : f \to f'$, let $\rho^\ast : [m] \to [m]$ denote its induced bijection on $S_m$ 
+and $R = DV$ the initial decomposition of $(K, f)$, a \emph{schedule}  $\mathcal{S} = (s_1, s_2, \dots, s_d )$ is a sequence of permutations  satisfying: 
 	\begin{equation}\label{eq:rv_seq_schedule}
@@
 	\centering
-	\includegraphics[width=\textwidth, height=16.35em]{boid_sim_results.png}
+	\includegraphics[width=\textwidth, height=16.15em]{boid_sim_results.png}
 	\caption{ One the left, the cumulative number of column operations (log-scale) of the three baseline approaches tested.
@@
 The degree to which multi-parameter persistence simplifies this exploratory phase cannot be understated: we believe multi-parameter persistence has a larger role to play in manifold learning. 
-Unfortunately, as mentioned prior, the compute barriers effective bar its use in practice.
+Unfortunately, as mentioned prior, the compute barriers effectively bar its use in practice.
 
@@
 In particular, we combine the use of furthest-point sampling and proportionate allocation (stratified) sampling to sample landmarks $X \subset \mathcal{X}$ distributed within $n = 25$ strata.
-Each strata consists of the $(1/n)$-thick level set given by $\rho_{15}$, the density estimator.
+Each strata consists of the $(1/n)$-thick level set given the $k$-nearest neighbor density estimator $\rho_{15}$ with $k = 15$.
 The use of furthest-point sampling gives us certain coverage guarantees that the geometry is approximately preserved within each level set, whereas the stratification ensures the original density of is approximated preserved as well.
-From this data set, we construct a Rips-(co)density bifiltration using $\rho_{15}$ equipped with the geodesic metric computed over the same $k = 15$ k-nearest neighbor graph on $X$. 
+From this data set, we construct a Rips-(co)density bifiltration using $\rho_{15}$ equipped with the geodesic metric computed over the same k-nearest neighbor graph on $X$. 
 Finally, we record the number of column reductions needed to compute the fibered barcode at a variety of levels of coarsening using $\mathtt{pHcol}$, vineyards, and our moves approach.
@@
 \subsubsection{Reduction Algorithm}
+The reduction algorithm, also called the ``standard algorithm,'' is the most often used modality  for computing persistence.
+While there exists other algorithms for computing persistence, they are typically not competitive with the reduction algorithm in practice.
+We outline the reduction algorithm below in Algorithm~\ref{alg:reduce}.
 \begin{algorithm}[h!]
@@
 \end{algorithm}
+The algorithm begins by copying $D$ to a new matrix $R$, to be subsequently modified in-place.
+After setting $V$ is set to the identity, the algorithm proceeds with column operations on both $R$ and $V$, left to right, until the decomposition invariants are satisfied.
+Since each column operation takes $O(m)$ and there are potentially $O(k)$ columns in $D$ with identical low entries (line 4 in ~\ref{alg:reduce}, observe the reduction algorithm below clearly takes $O(m^2 k)$ time.
+Since there exists complexes where $k \sim O(m)$, one concludes the bound of $O(m^3)$ is tight~\cite{morozov2005persistence}, though this seems to only be true on pathological inputs.
+Indeed, a more refined analysis by Edelsbrunner et al.~\cite{edelsbrunner2000topological} shows the reduction algorithm scales by the sum of squares of the cycle persistences, which is an output-sensitive bound.
 
 \subsubsection*{Move Algorithms}
-Although conceptually similar, there is an asymmetry between \emph{MoveRight} and \emph{MoveLeft}.
-Moving a simplex upwards in the filtration requires removing non-zero entries along several columns of a particular row in $V$ so that the corresponding permutation does not render $V$ non-upper triangular.
+As we've covered the moves algorithm extensively in section~\ref{sec:moves}, we now record the algorithmic components of both \emph{MoveRight} and \emph{MoveLeft}.
+Though conceptually similar, note that there is an asymmetry between \emph{MoveRight} and \emph{MoveLeft}: moving a simplex upwards in the filtration requires removing non-zero entries along several columns of a particular row in $V$ so that the corresponding permutation does not render $V$ non-upper triangular.
 The key insight of the algorithm presented in~\cite{busaryev2010tracking} is that $R$ can actually be maintained in all but one column during this procedure (by employing the \emph{donor} column).
 In contrast, moving a simplex to an earlier time in the filtration requires removing non-zero entries along several rows of a particular column of $V$.
-As before, $R$ stays reduced during this cancellation procedure in all but one column, however the subsequent permutation to $R$ requires reducing a pair of columns which may cascade into a larger chain of column operations to keep $R$ reduced.
+As before, though $R$ stays reduced during this cancellation procedure in all but one column, the subsequent permutation to $R$ requires reducing a pair of columns which may cascade into a larger chain of column operations to keep $R$ reduced.
+This is due to the fact that higher entries in columns in $R$ (above the pivot entry) may very well introduce additional non-reduced columns after $R$ is permuted. 
 Since these operations always occur in a left-to-right fashion, its not immediately clear how to apply a donor column kind of concept.
-Fortunately, we can still separate  these two stages.
-
+Fortunately, like move right, we can still separate the algorithm into a reduction and restoration phase---see Algorithm~\ref{alg:ml}.
+Moreover, since $R$ is reduced in all but one column by line 6 in~Algorithm~\ref{alg:ml}, we can still guarantee the number of column operations in $R$ will scale with $\lvert i - j \rvert$.
+For a supplementary description of the move algorithm, see~\cite{busaryev2010tracking}.
 \begin{algorithm}[!htb]
@@
         \Function{MoveLeft}{$R$, $V$, $i$, $j$} 
-        	\State $\mathbb{I} \gets \{ j \}$
+        	\State $\mathbb{I} \gets \emptyset$
             % \State $(\mathbb{K}, d_R, d_V) \gets (\emptyset, \mathrm{col}_R(i), \mathrm{col}_V(i))$
@@
 \subsubsection{LCS-Sort}
+Here we record explicitly the simple schedule construction algorithm outlined in section~\ref{sec:schedule_construction}.
+The algorithm is simple enough to derive using the rules discussed in section~\ref{sec:schedule_construction} (namely, equation~\eqref{eq:valid_move}, but nonetheless for posterity sake we record it here for the curious reader; it is given in Algorithm~\ref{alg:sorting}.
+The high level idea of the algorithm is to first construct the LCS between two permutations $p,q \in S_m$.
+To do this efficiently, one re-labels $q \mapsto \iota$ to the identity permutation $\iota = [m]$ and applies a consistent re-labeling $p \mapsto \bar{p}$.
+This relabeling preserves the LCS distance and has the additional advantage that $\bar{q} = \iota = [m]$ is a strictly increasing subsequence, and thus computing the LCS between $p,q \in S_m$ reduces to computing the LIS $\mathcal{L}$ of $\bar{p}$.
+By sorting $\bar{p} \mapsto p$ via operations which (strictly) increase the size of $\mathcal{L}$, we ensure that the size of the set of corresponding permutations is exactly $m - \lvert \mathcal{L} \rvert$.
 
@@
   \State \algorithmicif\ #1\ \algorithmicthen\ #2\ \algorithmicelse\ #3}
-  \begin{algorithm}[h]
+  \begin{algorithm}[h!]
 	\caption{Sorting algorithm}\label{alg:sorting}
@@
 \end{algorithm}
-
-\subsection{2D Persistence}\label{app:2d_pers}
-Indeed, one of the often noted drawbacks of persistence is its instability with respect to strong outliers. 
-Although persistence diagrams are stable with respect to their input---in the Hausdorff sense---and thus are robust to Hausdorff noise, the presence of strong outliers can artificially modify the connectivity of the underlying filtration, obscuring the detection of what would otherwise be significant topological structures~\cite{buchet2015topological}.
-
-The dimension function is simply the the function which maps every point $a \in \mathbb{R}^2$ to $\mathrm{dim}(M_a)$.
+Suppose $\mathcal{L}$ has been computed from $\bar{p}$.
+Since $\mathcal{L}$ is strictly increasing, the only symbols left to permute are in $\mathcal{L} \setminus \bar{p}$, which we denote with $\mathcal{D}$.
+After choosing any symbol $\sigma \in \mathcal{D}$, one then applies a cyclic permutation to $\bar{p}$ that moves $\sigma$ into any position that increases the size of $\mathcal{L}$.
+To do this efficiently, we maintain a data structure $\mathcal{T}$ which enables us to query the successor and predecessor of any given symbol $s \in \bar{p}$ \emph{in} $\mathcal{L}$.
+We also require a data structure to query the position of a given element $\sigma \in \bar{p}$, which for now we simply use the inverse permutation $\bar{p}$ (though a more efficient representation can be used based off of symbol \emph{displacements}, see section~\ref{sec:proxy_objective}.
+After the symbol is inserted into $\mathcal{L}$, we update $\bar{p}$, its inverse permutations $\bar{p}^{-1}$, $\mathcal{D}$ and $\mathcal{T}$ prior to the next move.
+The final set of permutations which sort $p \mapsto q$ (or equivalently, $\bar{p} \to \iota$ are stored in an array $\mathcal{S}$, which is then returned for further use.
+
+\subsection{2-parameter persistence}\label{app:2d_pers}\label{sec:fibered_barcode_reparam}
+We now describe the reparameterization between the bigraded Betti numbers and the set of ``critical lines'' Lesnick and Wright~\cite{lesnick2012multidimensional} used to create their interactive 2d persistence algorithm, beginning with point-line duality.
+%One of the often noted drawbacks of persistence is its instability with respect to strong outliers. 
+%Although persistence diagrams are stable with respect to their input---in the Hausdorff sense---and thus are robust to Hausdorff noise, the presence of strong outliers can artificially modify the connectivity of the underlying filtration, obscuring the detection of what would otherwise be significant topological structures~\cite{buchet2015topological}.
+%The dimension function is simply the the function which maps every point $a \in \mathbb{R}^2$ to $\mathrm{dim}(M_a)$.
 It is a simple and easy to visualize invariant, but is unstable and yields no information about the persistent features of $M$.
-The multigraded Betti numbers constitute a natural and important class of invariants often studied in algebraic geometry and commutative algebra, though both their theory and computation is beyond the scope of this effort---the interested reader is referred to~\cite{lesnick2015interactive, carlsson2009theory} for more details.
-For our purposes, the $i^{\text{th}}$-graded Betti number of $M$ is simply a function $\beta_i(M): \mathbb{R}^2 \to \mathbb{N}$ whose values indicate the the number of elements at  each degree in a basis of the $i^{\text{th}}$ module in a free resolution for $M$.
-
-\subsubsection*{Reparameterization using point-line duality}\label{sec:fibered_barcode_reparam}
-Let $\overline{\mathcal{L}}$ denote the collection of all lines in $\mathbb{R}^2$ with non-negative slope, $\mathcal{L} \subset \overline{\mathcal{L}}$ the collection of all lines in $\mathbb{R}^2$ with non-negative, finite slope, and $\mathcal{L}^\circ$ the collection of all affine lines in $\mathbb{R}^2$ with positive, finite slope.
-There is a standard point-line duality that gives a parameterization of $\mathcal{L}$ with the half-plane $[0, \infty) \times \mathbb{R}$ that is convenient to work with here.
+Let $\overline{\mathcal{L}}$ denote the collection of all lines in $\mathbb{R}^2$ with non-negative slope, $\mathcal{L} \subset \overline{\mathcal{L}}$ the collection of all lines with non-negative finite slope, and $\mathcal{L}^\circ$ the collection of all affine lines with positive finite slope.
+%There is a standard point-line duality that gives a parameterization of $\mathcal{L}$ with the half-plane $[0, \infty) \times \mathbb{R}$ that is convenient to work with here. 
 Define the \emph{line} and \emph{point} dual transforms $\mathcal{D}_{\ell}$ and $\mathcal{D}_p$, respectively, as follows: 
@@
 \end{enumerate}
-Property (1) asserts that for any $L \in \overline{\mathcal{L}}$, the standard partial order on $\mathbb{R}^2$ restricts to a total order on $L$.
-Properties (2) and (3) qualify the following definition:
+Property (1) elucidates how the standard partial order on $\mathbb{R}^2$ restricts to a total order on $L$ for any $L \in \overline{\mathcal{L}}$, whereas Properties (2) and (3) qualify the following definition:
 \begin{definition}[Critical Lines]
@@
 
+%To connect this construction back to persistence, one requires the definition of bigraded Betti numbers. 
+%In general, the multigraded Betti numbers constitute a natural and important class of invariants often studied in algebraic geometry and commutative algebra, though both their theory and computation is beyond the scope of this effort---the interested reader is referred to~\cite{lesnick2015interactive, carlsson2009theory} for more details.
+
+For our purposes, the $i^{\text{th}}$-graded Betti number of $M$ is simply a function $\beta_i(M): \mathbb{R}^2 \to \mathbb{N}$ whose values indicate the the number of elements at  each degree in a basis of the $i^{\text{th}}$ module in a free resolution for $M$.
 Let $S = \mathrm{supp}\,\beta_0(M) \, \cup \, \mathrm{supp}\,\beta_1(M)$, where the functions $\beta_0(M), \beta_1(M)$ are $0^{\text{th}}$ and $1^{\text{st}}$ bigraded Betti numbers of $M$, respectively.
@@
 
-\subsection{Computation}
-Computationally, the algorithm can be summarized into three core steps: 
+\subsubsection{Invariant computation}
+Computationally, the algorithm from~\cite{lesnick2019computing} can be summarized into three steps: 
 \begin{enumerate}
-	\item Computing the bigraded Betti numbers $\beta(M)$ of $M$
-	\item Constructing a line arrangement $\mathcal{A}(M)$ induced by critical lines from (1) 
-	\item Augmenting $\mathcal{A}(M)$ with a \emph{barcode template} $\mathcal{T}_e$ at every 2-cell $e \in \mathcal{A}(M)$
+	\item Compute the bigraded Betti numbers $\beta(M)$ of $M$
+	\item Construct a line arrangement $\mathcal{A}(M)$ induced by critical lines from (1) 
+	\item Augment $\mathcal{A}(M)$ with \emph{barcode templates} $\mathcal{T}_e$ at every 2-cell $e \in \mathcal{A}(M)$
 	%, yielding the \emph{augmented arrangement} $\mathcal{A}^{\sbullet}(M)$ 

### 2023-01-31 move_schedules.tex
new plots, new revision to 3.4.3, and small typo fixes

@@
 \abstract{
-	The standard procedure for computing the persistent homology of a filtered simplicial complex with $m$ simplices is the matrix reduction algorithm.
+	The standard procedure for computing the persistent homology of a filtered simplicial complex with $m$ simplices is the reduction algorithm.
 Its output is a particular decomposition of the total boundary matrix, from which the persistence diagrams and generating cycles are derived.
@@
 %	We also show that this maintenance procedure yields a factor-2 net reduction in column operations over the state of the art approach. 
-We propose a coarser strategy for maintaining the decomposition over a 1-parameter family of filtrations that requires only $O(m \log \log m)$ time and $O(m)$ space. 
+We propose a coarser strategy for maintaining the decomposition over a 1-parameter family of filtrations that requires only $O(m \log \log m)$ time and $O(m)$ space to construct. 
 By reduction to a particular longest common subsequence problem, we show the storage needed to employ this strategy is actually sublinear in expectation. 
 Exploiting this connection, we show experimentally the decrease in operations to compute diagrams across a family of filtrations is proportional to the difference between the expected quadratic number of states and the proposed sublinear coarsening.
-	Applications to movie data, dynamic metric spaces, and multi-parameter persistence are also presented.
+Applications to video data, dynamic metric space data, and multi-parameter persistence are also presented.
 }
@@
 As persistence-related computations tend to exhibit high algorithmic complexity---essentially cubic\footnote{For finite fields, it is known that the persistence computation reduces to the PLU factorization problem, which takes $O(m^\omega)$ where $\omega \approx 2.373$ is the matrix multiplication constant.} in the size of the underlying filtration~\cite{morozov2005persistence}---their adoption to dynamic settings poses a challenging computational problem.
-% With the current state of the art tools, there is no recourse to the problem of computing persistence from a time-varying filtration containing millions of simplices across thousands of snapshots in time.
-Acquiring such a capability has far-reaching consequences: methods that vectorize persistence diagrams for machine learning purposes, like adaptive template functions~\cite{polanco2019adaptive} and persistence images~\cite{adams2017persistence}, or summary statistics like $\alpha$-smoothed Betti curves~\cite{ulmer2019topological}, all immediately become computationally viable tools in dynamic settings.
+With state of the art tools, there is no recourse when faced with a time-varying complex containing millions of simplices across thousands of snapshots in time.
+Moreover, acquiring such a capability has far-reaching consequences: methods that vectorize persistence diagrams for machine learning purposes all immediately become computationally viable tools in dynamic settings.
+Persistence-based examples include adaptive template functions~\cite{polanco2019adaptive}, persistence images~\cite{adams2017persistence}, and $\alpha$-smoothed Betti curves~\cite{ulmer2019topological}.
 %This analogy provides a strong interpretation at gaining insight on the subtle topological changes that may occur in studying a continuous process.
@@
 Cohen-Steiner et al. refer to a continuous 1-parameter family of persistence diagrams as a \emph{vineyard}, and they give in \cite{cohen2006vines} an efficient algorithm for their computation. 
-The vineyards approach can be interpreted as an extension of the \emph{reduction} algorithm~\cite{zomorodian2005computing}, which computes the persistence diagram of a fixed filtration $K$ with $m$ simplices in $O(m^3)$ time via a decomposition $R = D V$ of the boundary matrix $D$ of $K$.
+The vineyards approach can be interpreted as an extension of the \emph{reduction} algorithm~\cite{zomorodian2005computing}, which computes the persistence diagram of a fixed filtration $K$ with $m$ simplices in $O(m^3)$ time via a decomposition $R = D V$ (or $RU = D$) of the boundary matrix $D$ of $K$.
 %Since $V$ is always full rank and upper-triangular, an alternative is to compute a $D = RU$ decomposition, where $U = V^{-1}$. 
-In particular, the vineyards algorithm transforms a time-varying filtration into a particular set of permutations of the decomposition $R = DV$, each of which takes at most $O(m)$ time to execute.
+In particular, the vineyards algorithm transforms a time-varying filtration into a certain set of permutations of the decomposition $R = DV$, each of which takes at most $O(m)$ time to execute.
 If one is interested in understanding how the persistent homology of a continuous function changes over time, then this algorithm is sufficient, for homological critical points can only occur when the filtration order changes.
-The vineyards algorithm is efficient asymptotically: if there are $d$ time-points where the filtration order changes, then  vineyards   takes $O(m^3 + md)$ time; one $O(m^3)$-time reduction upfront for the filtration at time $t_0$ followed by one $O(m)$ operation to update the decomposition for the remaining time points $(t_1, t_2, \dots, t_d)$.
-When $d >> m$, this $O(md)$ approach is far more efficient than the $O(dm^3)$  ``naive'' strategy of computing the diagrams independently at every time point.
+The vineyards algorithm is efficient asymptotically: if there are $d$ time-points where the filtration order changes, then  vineyards   takes $O(m^3 + md)$ time; one initial $O(m^3)$-time reduction at time $t_0$ followed by one $O(m)$ operation to update the decomposition at the remaining time points $(t_1, t_2, \dots, t_d)$.
+When $d >> m$, this $O(md)$ approach is far more efficient than the $O(dm^3)$  ``naive'' strategy of computing the diagrams at every time point independently.
 
@@
 This work seeks to further understand and remedy this discrepancy: building on the work presented in~\cite{busaryev2010tracking}, we introduce a coarser approach to the vineyards algorithm.
-While vineyards is  designed  for handling   \emph{continuous} time families of diagrams, it is not necessarily efficient when the time parameter is coarsely discretized.
+While vineyards is  designed  for handling a \emph{continuous} 1-parameter family of diagrams, it is not necessarily efficient when the parameter is coarsely discretized.
 Our methodology is based on the observation that practitioners often don't need (or want!) all of the persistence diagrams generated by a continuous 1-parameter of filtrations; usually just $n << d$ of them  suffice.
@@
 \subsection{Related Work}\label{sec:related_work} 
-To the authors knowledge, work focused on ways of updating a  decomposition $R = DV$, for all homological dimensions, is limited---there is the vineyards algorithm~\cite{cohen2006vines} and the moves algorithm~\cite{busaryev2010tracking}, both of which are discussed extensively in section~\ref{sec:background}.
-
-In contrast, there is extensive work  on improving the efficiency of computing a (static) $R = DV$ decomposition.
+To the authors knowledge, work focused on ways of updating a  decomposition $R = DV$, for all homological dimensions, is limited: there is the vineyards algorithm~\cite{cohen2006vines} and the moves algorithm~\cite{busaryev2010tracking}, both of which are discussed extensively in section~\ref{sec:background}.
+At the time of writing, we were made aware of very recent work~\cite{luo2021accelerating} that iteratively repairs a permuted decomposition via a column swapping and reduce strategy, which they call ``warm starts.'' Though similar in spirit, their approach relies on the reduction algorithm as a subprocedure, which is quite different from the strategy we employ here.
+
+Contrasting the dynamic setting, there is extensive work  on improving the efficiency of computing a single (static) $R = DV$ decomposition.
 Chen~\cite{chen2011persistent} proposed \emph{persistence with a twist}, also called the \emph{clearing optimization}, which exploits a boundary/cycle relationship to ``kill'' columns early in the reduction rather than reducing them.
@@
 By making stronger assumptions on the underlying topological space, restricting the homological dimension, or targeting  a weaker invariant (e.g.
-Betti numbers), one can usually obtain  faster or alternative approaches  recording information related to persistence.
-For example, Attali et al.~\cite{attali2009persistence} give a linear time algorithm (under the $\texttt{word RAM}$ model of computation) for computing persistence on graphs.
+Betti numbers), one can usually obtain faster algorithms.
+For example, Attali et al.~\cite{attali2009persistence} give a linear time algorithm for computing persistence on graphs.
 In the same paper, they describe how to obtain $\epsilon$-simplifications of $1$-dimensional persistence diagrams for filtered $2$-manifolds by using duality and symmetry theorems. 
@@
 
-If  zeroth homology is the only dimension of interest, computing and updating both the persistence and rank information  is greatly simplified.
+If  zeroth homology ($p = 0$) is the only dimension of interest, computing and updating both the persistence and rank information  is greatly simplified.
 For example, if the relations are available a-priori, obtaining a tree representation fully characterizing the connectivity of the underlying space (also known as the \emph{incremental connectivity} problem) takes just $O(\alpha(n) n)$ time using the disjoint-set data structure, where $\alpha(n)$ is the extremely slow-growing inverse Ackermann function.
-Adapting this approach to the time-varying setting, Oesterling et al.~\cite{oesterling2015computing} give a $O(e)$-per-update algorithm that allows one to maintain a \emph{merge tree} with $e$ edges,
-associated to the filtration parameter changing continuously over time. 
-If only the zeroth-dimensional Betti numbers are needed for a particular application, this problem reduces even further to tracking the connected components of a dynamic graph---sometimes referred to as the \emph{dynamic connectivity problem}, which can be efficiently  solved in amortized $O(\log n)$ query and update times using either Link-cut trees or multi-level Euler tour trees~\cite{kapron2013dynamic}.
+Adapting this approach to the time-varying setting, Oesterling et al.~\cite{oesterling2015computing} give an algorithm that maintains a \emph{merge tree} with $e$ edges in $O(e)$ time per-update.
+If only Betti numbers are needed, the zeroth-dimension problem reduces even further to the \emph{dynamic connectivity problem}, which can be be efficiently solved in amortized $O(\log n)$ query and update times using either Link-cut trees or multi-level Euler tour trees~\cite{kapron2013dynamic}.
 
@@
 %Note that the size of the filtration is fixed---no simplices are added or removed over time. 
-Two events alter the persistence diagrams over time: the first occurs when the central connected component splits to form a cycle, and the second  when the annulus  splits into four components. 
+Two events critically change the pairing function: the first occurs when the central connected component splits to form a cycle, and the second  when the annulus  splits into four components. 
 From left to right, the Betti numbers of the five evenly spaced `snapshots' of the filtration shown above are: $(\beta_0,\beta_1)  = (1,0), \; (1,1),  \; (1,1), \; (1,1), \; (4,0)$.
@@
 Suppose a practitioner wanted to know the major homological changes a time-varying filtration encounters over time.
-  Since it is unknown a priori at when the persistent pairing function changes, one solution is to do $n$ independent persistence computations $n$ evenly spaced points in the time domain.
-An alternative approach is to construct a homotopy between contiguous time-points and decompose these homotopies into adjacent transpositions in the filtration order---the vineyards approach.
- We refer to the former as the \emph{discrete setting}, which is often used in practice, though it is only an approximation; the latter is guaranteed to capture all homological changes in persistence that occur in simulating the 1-parameter family, which we'll refer to as the \emph{continuous setting}.
+  Since it is unknown a priori at when the persistent pairing function changes, one solution is to do $n$ independent persistence computations at $n$ evenly spaced points in the time domain.
+An alternative approach is to construct a homotopy between a pair of filtrations $(K, f)$, $(K,f')$ and then decompose this homotopy into adjacent transpositions of the filtration order---the vineyards approach.
+ We refer to the former as the \emph{discrete setting}, which is often used in practice, and the  latter as the \emph{continuous setting}.
+Note that though the discrete setting is often more practical, it is only not guaranteed to capture all homological changes in persistence that occur in simulating a continuous 1-parameter family of diagrams.
 
  The cumulative cost (in total column operations) of these various approaches are shown on the left side of Figure~\ref{fig:vineyards}, wherein the reduction algorithm (\texttt{pHcol}) and the vineyards algorithm are compared.
-Two approximate strategies (green and purple) and two continuous strategies (black and blue) are shown.
+Two discrete strategies (green and purple) and two continuous strategies (black and blue) are shown.
   Note that without knowing where persistence pairing function changes, a continuous strategy must construct all $\approx 7 \times 10^{4}$ diagrams induced by the homotopy. 
@@
   The existence of a time discretization that is more efficient to compute than continually updating the decomposition indicates that the vineyards framework must incur some overhead (in terms of column operations) to maintain the underlying decomposition, even when pairing function is unchanged.
-Indeed, applying \texttt{pHcol} independently between  relatively ``close'' filtrations is more efficient than iteratively update the decomposition, as shown by the case where $n = 10$.
+Indeed, as shown by the case where $n = 10$, applying \texttt{pHcol} independently between  relatively ``close'' filtrations is substantially more efficient than iteratively updating the decomposition.
  Moreover, any optimizations to the reduction algorithm (e.g. clearing~\cite{chen2011persistent}) would only increase this disparity.
 Since persistence has found many applications in dynamic contexts~\cite{topaz2015topological, xian2020capturing, lesnick2015interactive, kim2020spatiotemporal}, a more efficient alternative to vineyards is clearly needed.
-
+ \\
+ \\
+ \noindent
 \textbf{Our approach and contributions are as follows:} First, we leverage the \emph{moves} framework of Busaryev et al.~\cite{busaryev2010tracking} to include  coarser  operations---in terms of the number of valid intermediate decomposition states, as compared to vineyards---for dynamic persistence.
@@
 Although the greedy approach from the last section can lead to arbitrarily high cost schedules, we find similar greedy-like strategies can yield computationally efficient heuristics in practice, even if they are suboptimal.
-Choosing the lowest-cost move takes $\approx O(dm)$---implying a greedy strategy would require $O(d^2 m)$ time---which is computationally viable as a heuristic when $d << m$. 
-We seek a fast procedure for generating schedules that is not only substantially better than a randomly or naive strategy in term of column reductions, but is also low complexity enough that it is practical in a real-world setting.
-For example, the vineyards algorithm requires $O(m \log m + k)$ preprocessing time to compute a schedule of size $k \sim O(m^2)$ transpositions, thus for Algorithm~\ref{alg:schedule} to be considered a more efficient alternative to vineyards, a solution subquadratic in $m$ is desirable.
+%Choosing the lowest-cost move takes $\approx O(dm)$---implying a greedy strategy would require $O(d^2 m)$ time---which is computationally viable as a heuristic when $d << m$. 
+We seek a fast procedure for generating schedules that is not only substantially better than a random or simple schedule strategy in term of column reductions, but has a low enough time and storage complexity to be practical for larger data sets.
+%For example, the vineyards algorithm requires $O(m \log m + k)$ preprocessing time to compute a schedule of size $k \sim O(m^2)$ transpositions, thus for Algorithm~\ref{alg:schedule} to be considered a more efficient alternative to vineyards, a solution subquadratic in $m$ is desirable.
 %In what follows, we present such an objective, showing its relation to the problem at hand as well its use in other known computational problems. 
@@
 We seek a \emph{proxy objective} that correlates with~\eqref{eq:schedule_cost} and does not explicitly depend on the entries in the decomposition, as minimizing these terms directly is difficult due to  the changing sparsity of the intermediate matrices $R_k, V_k$,
-Consider a strategy $\mathcal{S} \in \Phi(p,q,d)$ that minimizes the upper bound:
+Given a pair of filtrations $(K, f), (K,f')$, consider a schedule $\mathcal{S} \in \Phi(f,f',d)$ of cyclic permutations $((i_1, j_1), (i_2, j_2), \dots, (i_d, j_d))$ minimizing:
 \begin{equation}\label{eq:upper_bound_move_cost}
-\mathrm{cost}(\mathcal{S}) = \sum\limits_{k=1}^d ( \lvert \mathbb{I}_k \rvert + \lvert \mathbb{J}_k\rvert ) \leq  \sum\limits_{k=1}^d 2 \lvert i_k - j_k \rvert 
+\mathrm{\widetilde{cost}}(\mathcal{S}) = \sum\limits_{k=1}^d 2 \lvert i_k - j_k \rvert \geq \sum\limits_{k=1}^d ( \lvert \mathbb{I}_k \rvert + \lvert \mathbb{J}_k\rvert )
 \end{equation}
@@
 %and the distance between pairs of uniformly random permutations $p, q \sim S_m$ are asymptotically normal. 
-Indeed, Diaconis et al.~\cite{diaconis1977spearman} showed the the Spearman distance approximates $K_\tau$ within a factor of two: 
+Indeed, Diaconis et al.~\cite{diaconis1977spearman} showed the Spearman distance approximates $K_\tau$ within a factor of two: 
 \begin{equation}\label{eq:diaconis_inequality}
@@
 \end{equation}
-In contrast to $K_\tau$, the Spearman can be computed in $O(m)$ time, making it an attractive alternative quantity in the well-known \emph{rank aggregation problem}.
+Moreover, in contrast to $K_\tau$, the Spearman distance can be computed in $O(m)$ time, has a $\ell_1$ interpretation between ranking, and is often used in the well-known \emph{rank aggregation problem}: whereas obtaining a Kemeni optimal aggregation with respect to $K_\tau$ is NP-hard, the optimal such aggregation with respect to $F$ is obtainable in polynomial time~\cite{dinu2006efficient}.
 % Indeed, it is used...
@@
 In contrast, a 2-approximation may be obtained by replacing the Kendall distance with the Spearman distance, as in this case the problem reduces to finding a perfect matching in a bipartite graph, which can be accomplished in polynomial time.
-To adapt $F(\cdot, \cdot)$ to sortings, we consider the additive minimization problem which minimizes the following upper bound on $F(p, \iota)$:
+To adapt $F(\cdot, \cdot)$ to sortings, we decompose the footrule distance additively via the bound:
 \begin{equation}\label{eq:sorting_additive}
@@
 where $\hat{s}_i =  s_i \circ \cdots \circ s_2 \circ s_1$ denote the composition of the first $i$ permutations of a sorting  $S = (s_1, \dots, s_d)$ that maps $p \mapsto \iota$.
-Note that equality in~\eqref{eq:sorting_additive} is achieved when the displacement of each symbol between its initial position in $p$ to its value is non-increasing with every application of $s_i$, which in general not guaranteed using the schedule construction method derived in section~\ref{sec:move_schedules}.
-
 The problem of minimizing the right hand side of~\eqref{eq:sorting_additive} can be interpreted as a crossing minimization problem for a set of $k$-layered bipartite graphs.
 To see this, consider two permutations: $p$ and $m_{ij} \circ p$, where $m_{ij}$ is a move permutation.
-If the $(p, m_{ij} \circ p)$ are drawn as a bipartite graph, with the elements of $p$ in one layer and the elements of $m_{ij} \circ p$ in another, then observe that $F(p, m_{i,j} \circ p)$ is twice the number of edge crossings in the graph.
-We given an example below illustrating this.
+Drawing $(p, m_{ij} \circ p)$ as a bipartite graph,  observe that $F(p, m_{i,j} \circ p)$ is twice the number of edge crossings in the graph, and that equality in~\eqref{eq:sorting_additive} is achieved when the displacement of each symbol between its initial position in $p$ to its value is non-increasing with every application of $s_i$, which in general not guaranteed using the schedule construction method derived in section~\ref{sec:move_schedules}.
+Unfortunately, the $k$-layer crossing minimization problem is NP-hard for $k$ sets of permutations, when $k \geq 4$~\cite{biedl2009complexity}.
 \\
@@
 \noindent \textbf{Example:} 
-Let  $p= (1\;2\;3\;4\;5\;6\;7\;8\;9)$ 
-and 
-$q = (9\;4\;2\;7\;1\;8\;6\;3\;5)$.
-An example of three possible schedules, $S_1$, $S_2$, and $S_3$ sorting $p$ into $q$ is given in Figure~\ref{fig:crossings}. 
+Let  $p= (1\;2\;3\;4\;5\;6\;7\;8\;9)$ and $q = (9\;4\;2\;7\;1\;8\;6\;3\;5)$.
+An example of three possible schedules, $S_1$, $S_2$, and $S_3$ sorting $p$ into $q$ is given in the figure below. 
 %Each schedule, from left to right, has $21$, $31$, and $35$ crossings, respectively. 
@@
     \includegraphics[width=0.80\textwidth]{crossings.png}
-    \caption{Three sortings, $S_1$, $S_2$, and $S_3$, with $21$, $31$, and $35$ crossings, resp.}
+    %\caption{Three sortings, $S_1$, $S_2$, and $S_3$, with $21$, $31$, and $35$ crossings, resp.}
     %.
@@
 \end{figure}
-%The greedy heuristic proposed to minimize equation~\ref{eq:schedule_opt_succ} was used to produce $S_1$, which has 21 edge crossings.
-
-\noindent  Each column represents the successive application of a move $m_{ij}$ in the schedule, 
+Each column represents the successive application of a move $m_{ij}$ in the schedule, 
 and the edges track how each  symbol has been moved. 
 Black/red vertices correspond to symbols in and outside of $\mathrm{LCS}$, respectively. 
-All three schedules were generated from the same $\mathrm{LCS}(p,q) = (4\,7\,8)$ and all three schedules transform $p \mapsto q$ in $d = 6$ moves. 
+All three schedules were generated from the same $\mathrm{LCS}(p,q) = (4\,7\,8)$ and each schedule transforms $p \mapsto q$ in $d = 6$ moves. 
 In this example, $S_1$ matches the minimal number of crossings amongst all possible schedules,  since $K_\tau(p,q) = 21$. 
@@
 \\
-The example above suggests minimizing~\eqref{eq:sorting_additive} is reminiscent of the $k$-layer crossing minimization problem for $k$ sets of permutations, which is NP-hard for $k \geq 4$~\cite{biedl2009complexity}.
+%The greedy heuristic proposed to minimize equation~\ref{eq:schedule_opt_succ} was used to produce $S_1$, which has 21 edge crossings. 
 %As a result, we do not expect there to be a polynomial time algorithm for optimally achieving this proxy objective.  
@@
 %% TODO: elaborate on monotonicity, describe segment tree
-We propose a heuristic strategy to minimize~\eqref{eq:sorting_additive} which we observed is both efficient to compute quite effective in practice.
+In light of the discussion above, we propose a heuristic strategy to minimize~\eqref{eq:sorting_additive} which we observed is both efficient to compute and effective in practice.
 The heuristic is inspired by the simplicity of computing the Spearman distance between cyclic permutations at the schedule construction phase.
@@
 \]
-That is, if $s_\sigma$ represents a move permutation moving $\sigma$ from position $i$ to position $j$, then relative to the previous permutation $p$, the displacement of $\sigma$ in $s_\sigma \circ p$ changes by $\lvert i - j \rvert$ and the displacement of any symbol between $i$ and $j$ changes by $\pm 1$.
-All other symbols are unchanged.
-Thus, if we replace $\mathcal{A}$ with a data structure  supporting $O(\log m)$ access time to aggregate information (i.e.
-the sum) and $O(\log m)$ modification time ability on $\lvert i - j \vert$ entries (simultaneously), we could greedily choose the next permutation $s$ to maximize:
+%That is, if $s_\sigma$ represents a move permutation moving $\sigma$ from position $i$ to position $j$, then relative to the previous permutation $p$, the displacement of $\sigma$ in $s_\sigma \circ p$ changes by $\lvert i - j \rvert$ and the displacement of any symbol between $i$ and $j$ changes by $\pm 1$.
+%All other symbols are unchanged. 
+Thus, if we replace $\mathcal{A}$ with a data structure  supporting $O(\log m)$ access time to aggregate information and $O(\log m)$ modification time ability on $\lvert i - j \vert$ entries simultaneously, we could greedily choose the next permutation $s$ to minimize:
 \begin{equation}\label{eq:greedy_step}
-	s_{\text{greedy}} = \argmin_{\text{ valid } s \in \Sigma} F(s\circ p, \iota)
+	s_{\text{greedy}} = \argmin_{\text{ valid } s_{\sigma \in S_{\mathcal{D}}}} F(s\circ p, \iota)
 \end{equation}  
@@
 The former problem reduces to the problem of efficiently calculating \emph{prefix sums}, which is easily solved. 
-It is not immediately clear how to achieve the latter modification complexity in $O(\log m)$ time, since $\lvert i - j \rvert \leq m$ is potentially larger than $O(\log m)$.  
+It is not immediately clear how to achieve the latter modification complexity in $O(\log m)$ time, since $\lvert i - j \rvert \leq m$ is potentially larger than $O(\log m)$---to handle this, note that one can apply constant-factor updates to multiple values in a \emph{implicit treap} data structure in $O(\log m)$ time via \emph{range updates}.
  % Can actually only achieve O(d^2 log m) it seems
-It is known that one can apply constant-factor updates to multiple values in a prefix sum in $O(\log m)$ time by any prefix-sum data structure that also supports \emph{range updates}---e.g.
-an \emph{implicit treap}.
-Since single element modifications, removals, and insertions can all be achieved in $O(\log m)$ expected time with such a data structure, we conclude that a greedy solution to equation~\eqref{eq:greedy_step} can be computed in $O(d^2 \log m)$ time.
+Since single element modifications, removals, and insertions can all be achieved in $O(\log m)$ expected time with such a data structure, we conclude that the greedy approach solving equation~\eqref{eq:greedy_step} requires just $O(d^2 \log m)$ time.
 %Since single element prefix sum modifications can also be achieved $O(\log m)$ time with these data structures, the time to compute $F(\hat{s}_i, \hat{s}_{i+1})$ is bounded above by $O(\log m)$, 
@@
 \section{Applications and Experiments}\label{sec:results}
-\subsection{Movies}
+\subsection{Video data}
 A common application of persistence is to characterizing topological structure in image data.
-Since a  set of ``snapshot'' frame of a movie can be equivalently thought of as discrete 1-parameter family, our framework provides a natural extension of the typical image analysis to video data.
-To demonstrate the benefit of using minimal sized scheduled and the scalability of the greedy approach proposed in section~\ref{sec:proxy_objective}, we re-use the movie data from~\ref{sec:motivation} as a baseline benchmark.
+Since a  set of ``snapshot'' frames of a video can be equivalently thought of as discrete 1-parameter family, our framework provides a natural extension of the typical image analysis to video data.
+To demonstrate the benefit of using minimal sized scheduled and the scalability of the greedy approach proposed in section~\ref{sec:proxy_objective}, we re-use the video data from~\ref{sec:motivation} as a baseline benchmark.
 We perform two performance tests: one to  one to test the impact of shrinking the number of permutations a given reindexing operation needs and one to test the asymptotic behavior of the greedy approach and.
@@
  In the first test, we fix a grid size of $9 \times 9$ and record the cumulative number of column operations needed to simulate persistence dynamically across 25 evenly-spaced time points using a variety of scheduling strategies.
- The primary three strategies we test are the greedy approach from section~\ref{sec:proxy_objective}, the ``naive'' approach which uses upwards of $O(m)$ move permutations via selection sort, and a third strategy which interpolates between the two.
+ The primary three strategies we test are the greedy approach from section~\ref{sec:proxy_objective}, the ``simple'' approach which uses upwards of $O(m)$ move permutations via selection sort, and a third strategy which interpolates between the two.
  In particular, if $d = m - \lvert \mathrm{LCS}(f, f') \rvert$ we use a parameter $\alpha \in [0,1]$ to choose $m - \alpha \cdot d$ random symbols to move using the same construction method outlined in section~\ref{sec:schedule_construction}.
@@
 To get an idea of the variation of the random sampling, we run 10 independent iterations and shade the upper and lower bounds of the schedule costs, which is shown around the mean schedule cost depicted by the solid lines in Figure~\ref{fig:movie_perf}.
- As one can see from the figure, while using less move operations (lower $\alpha$) does progressively reduce column operations, constructing random schedules of minimal size is no more competitive than the naive selection sort strategy.
+ As one can see from the figure, while using less move operations (lower $\alpha$) does progressively reduce column operations, constructing random schedules of minimal size is no more competitive than the selection sort strategy.
 This suggests that efficient schedule construction needs to account for the structure of performing several permutations in sequence, like the greedy heuristic we introduced, to yield a performance boost.
@@
 In the second test, we aim to measure the asymptotics of our greedy LCS-based approach.
-To do this, we generated 8 movie data sets again of the expanding annulus outlined in section~\ref{sec:motivation}, each of increasing grid sizes of $5 \times 5$, $6 \times 6$, $\dots$, $12 \times 12$.
-For each data set, we simulate persistence over the duration of the movie, again testing five evenly spaced settings of $\alpha \in [0,1]$---the results are shown in the right plot of Figure~\ref{fig:movie_perf}.
-On the vertical axis, we plot the total number of column operations needed to simulate the movie again across 25 evenly-spaced time points \emph{as a ratio} of the data set size ($m$); we also show the regression curves one obtains from considering each schedule size.
+To do this, we generated 8 video data sets again of the expanding annulus outlined in section~\ref{sec:motivation}, each of increasing grid sizes of $5 \times 5$, $6 \times 6$, $\dots$, $12 \times 12$.
+For each data set, we simulate persistence over the duration of the video, again testing five evenly spaced settings of $\alpha \in [0,1]$---the results are shown in the right plot of Figure~\ref{fig:movie_perf}.
+On the vertical axis, we plot the total number of column operations needed to simulate the video again across 25 evenly-spaced time points \emph{as a ratio} of the data set size ($m$); we also show the regression curves one obtains from considering each schedule size.
 As one can see from the Figure, the cost of using the greedy heuristic tends to increase sub-linearly as a function of the data set size, suggesting the move scheduling approach is indeed quite scalable.
@@
 $$ \lVert x \rVert_D = \sqrt{\sum_{i \sim j}(x_i - x_j)^2} = \sqrt{x^T D x}$$
-where $D$ is a fixed matrix which upon application to an image $x \in \mathbb{R}^9$ yields a value proportional to the sum of the differences between each pixels 4 connected neighbors (described above by the relation $i \sim j$).
+where $D$ is a fixed matrix which upon application $x^T D x$ to an image $x \in \mathbb{R}^9$ yields a value proportional to the sum of the differences between each pixels 4 connected neighbors (given above by the relation $i \sim j$).
 Their research was primarily motivated by discerning whether there existed clear qualitative differences in the distributions of patches extracted from images of different modalities, such optical and range images.
-By mean-centering, contrast normalizing, and``whitening'' the data via the Discrete Cosine Transform (DCT) basis~\cite{lee2003nonlinear}, they found they could form a convenient basis for $D$ via an expansion of 8 certain non-constant eigenvectors, shown below:
+By mean-centering, contrast normalizing, and``whitening'' the data via the Discrete Cosine Transform (DCT) basis, they a convenient basis for $D$ may be obtained via an expansion of 8 certain non-constant eigenvectors, shown below:
 %\vspace{-0.2em}
@@
 Since these images are scale-invariant, the expansion of these basis vectors spans the 7-sphere, $S^7 \subset \mathbb{R}^8$.
-Using a voronoi cell decomposition of the data, their subsequent distribution analysis suggested that the majority of data points concentrated in a few high-density regions.
-
-After Lee et al published their work, Carlsson et al.~\cite{carlsson2008local} subsequently performed extensive experiments conducted using persistent homology, wherein he found that the distribution of high-contrast $3 \times 3$ patches is actually well-approximated by a Klein bottle $\mathcal{M}$---around 60\% of the high-contrast patches from the still image data set lie within a small neighborhood around $\mathcal{M}$ which accounts for only 21\% of the volume $S^7$.
-Along a similar vein, in the context of sparse coding, Perea et al.~\cite{perea2014klein} introduced a dictionary learning framework for estimating and representing the distribution of patches from texture images.
+Using a voronoi cell decomposition of the data, their distribution analysis suggested that the majority of data points concentrated in a few high-density regions.
+
+After Lee et al published their work, Carlsson et al.~\cite{carlsson2008local} subsequently performed extensive experiments using persistent homology, wherein he found that the distribution of high-contrast $3 \times 3$ patches is actually well-approximated by a Klein bottle $\mathcal{M}$---around 60\% of the high-contrast patches from the still image data set lie within a small neighborhood around $\mathcal{M}$ which accounts for only 21\% of the 7-spheres volume.
+Along a similar vein in the sparse coding context, Perea et al.~\cite{perea2014klein} introduced a dictionary learning framework for estimating the distribution of patches from texture images.
 
@@
 
-If one was not aware of the analysis done by~\cite{lee2003nonlinear, hateren_schaaf_1998, carlsson2008local, perea2014klein}, it is not immediately clear a priori that the Klein bottle model is  a good candidate for capturing non-linearity of image patches.
-Indeed, armed with fine topological intuition, Carlsson still needed to perform extensive sampling, preprocessing, and model fitting techniques in order reveal the underlying the topological space with persistent homology~\cite{carlsson2008local}.
+If one was not aware of the analysis done by~\cite{lee2003nonlinear, hateren_schaaf_1998, carlsson2008local, perea2014klein}, it is not immediately clear a priori that the Klein bottle model is  a good candidate for capturing the non-linearity of image patches.
+Indeed, armed with a refined topological intuition, Carlsson still needed to perform extensive sampling, preprocessing, and model fitting techniques in order reveal the underlying the topological space with persistent homology~\cite{carlsson2008local}.
 One reason such preprocessing is needed is due to persistent homology's aforementioned instability with respect to strong outliers. 
@@
 Such adaptations may result in even larger reductions in the number of column operations, as have been observed in practice for the standard reduction algorithm~\cite{bauer2021ripser}.
+Moreover, though we have carefully constructed an efficient greedy heuristic in section~\ref{sec:proxy_objective} and illustrated a different perspective with which to view our heuristic (via crossing minimization), it is an open question whether there exists a more structured reduction of~\eqref{eq:schedule_cost} or~\eqref{eq:sorting_additive} to a better-known problem.
 
@@
 
-%To connect this construction back to persistence, one requires the definition of bigraded Betti numbers. 
-%In general, the multigraded Betti numbers constitute a natural and important class of invariants often studied in algebraic geometry and commutative algebra, though both their theory and computation is beyond the scope of this effort---the interested reader is referred to~\cite{lesnick2015interactive, carlsson2009theory} for more details.
-
-For our purposes, the $i^{\text{th}}$-graded Betti number of $M$ is simply a function $\beta_i(M): \mathbb{R}^2 \to \mathbb{N}$ whose values indicate the the number of elements at  each degree in a basis of the $i^{\text{th}}$ module in a free resolution for $M$.
+%The multigraded Betti numbers constitute a natural and important class of invariants often studied in algebraic geometry and commutative algebra, though both their theory and computation is beyond the scope of this effort---
+
+To connect this construction back to persistence, one requires the definition of bigraded Betti numbers.
+For our purposes, the $i^{\text{th}}$-graded Betti number of $M$ is simply a function $\beta_i(M): \mathbb{R}^2 \to \mathbb{N}$ whose values indicate the the number of elements at each degree in a basis of the $i^{\text{th}}$ module in a free resolution for $M$---the interested reader is referred to~\cite{lesnick2015interactive, carlsson2009theory} for a more precise algebraic definition.
 Let $S = \mathrm{supp}\,\beta_0(M) \, \cup \, \mathrm{supp}\,\beta_1(M)$, where the functions $\beta_0(M), \beta_1(M)$ are $0^{\text{th}}$ and $1^{\text{st}}$ bigraded Betti numbers of $M$, respectively.
@@
 Each vertex in $G$ is associated with a barcode template $\mathcal{T}^e$, and the computation of $\mathcal{T}$ now reduces to computing a path $\Gamma$ on $G$ which visits each vertex at least once.
-To minimize the computation time, assume the $n$ edges of $G$ are endowed with non-negative weights $W = w_1, w_2, \dots, w_n$ whose values $w_i \in \mathbb{R}_{+}$ represent some notion of distance.
-The optimal path $\Gamma^\ast$ is the one of minimal length with respect to $W$ which visits every vertex of $G$ at least once.
-There is a known $\frac{3}{2}$-approximation that can be computed efficiently which reduces the problem to the traveling salesman problem on a metric graph~\cite{christofides2022worst}.
+To minimize the computation time, assume the $n$ edges of $G$ are endowed with non-negative weights $W = w_1, w_2, \dots, w_n$ whose values $w_i \in \mathbb{R}_{+}$ represent some notion of distance which is proportional to the computational disparity between adjacent template computations.
+The optimal path $\Gamma^\ast$ that minimizes the computation time is then the minimal length path with respect to $W$ which visits every vertex of $G$ at least once.
+There is a known $\frac{3}{2}$-approximation that can be computed efficiently which reduces the problem to the traveling salesman problem on a metric graph~\cite{christofides2022worst}, and thus can be used so long as the distance function between templates is a valid metrics.
+\cite{lesnick2012multidimensional} use the kendall distance between the push-map induced filtrations, but other options are available---for example, any of the combinatorial metrics we studied in Section~\ref{sec:schedule_cost}.
 

### 2023-01-31 move_schedules.tex
cleaning up

@@
 Inspection of the individual cases of the algorithm from~\cite{cohen2006vines} shows that any single transposition requires at most two such operations on both $R$ and $V$.
-However, there are a number of factors that affect the runtime efficiency of the algorithm as a whole. 
+However, a host of other factors affect the runtime efficiency of the vineyards algorithm. 
 On the positive side, as both the $V$ and $R$ matrices tend to be quite sparse, the cost of a given column operation is proportional the number of non-zero field coefficients in the two columns being modified.
 As a rule of thumb, most transpositions require no column operations~\cite{edelsbrunner2000topological}.
-On the negative side, one needs to constantly query the non-zero status of various entries in $R$ and $V$, which has a non-trivial runtime cost (consider evaluating e.g.
-Case 1.1 in~\cite{cohen2006vines}).
- Moreover, an efficient implementation of the vineyards algorithm requires a non-standard sparse matrix representation that allows swapping any two rows and columns in $O(1)$, as described in Figure 5 in~\cite{cohen2006vines}.
- Though these minor operations do not affect the asymptotic complexity of the vineyards algorithm, they collectively accrue a non-trivial runtime cost due to the quadratic scaling in how often they need to be performed.
+On the negative side, one needs to frequently query the non-zero status of various entries in $R$ and $V$ (consider evaluating e.g.
+Case 1.1 in~\cite{cohen2006vines}), which accrues a non-trivial runtime cost due to the quadratic frequency with which they are required.
+% Moreover, an efficient implementation of the vineyards algorithm requires a non-standard sparse matrix representation that allows swapping any two rows and columns in $O(1)$, as described in Figure 5 in~\cite{cohen2006vines}.
+% Though these minor operations do not affect the asymptotic complexity of the vineyards algorithm, they collectively accrue a non-trivial runtime cost
 

### 2023-01-31 move_schedules.tex
updates to get overleaf working

@@
 % \documentclass[review]{siamart190516} % for SICOMP 
-%\documentclass{article} % for arxiv 
-% \usepackage{arxiv}
+%\documentclass{article} % for arxiv only
+%\usepackage{arxiv} % for arxiv only 
 %\pagestyle{plain} % arxiv only 
@@
 \usepackage[english]{babel}
-% \usepackage{algorithm}
 \usepackage{bbm}
@@
 \usepackage{array}
-\usepackage[inkscape={/usr/local/bin/inkscape -z -C}]{svg}
-%\usepackage{algorithm2e}
-%%\usepackage{algorithmicx}
-% \usepackage[noend]{algpseudocode} % keep the noend argument
-%\usepackage{bbm}
-% \usepackage{graphicx}
-
 \usepackage{blkarray, bigstrut}
@@
 \newcommand\botstrut[1][0.9ex]{\setlength\bigstrutjot{#1}{\bigstrut[b]}}
+
+\usepackage{epstopdf}
+%\epstopdfDeclareGraphicsRule{.pdf}{png}{.png}{convert #1 \OutputFile}
+%\DeclareGraphicsExtensions{.png,.pdf}
+\DeclareGraphicsExtensions{.pdf,.png}
 
@@
 % Optional PDF information
-\ifpdf
-\hypersetup{
-  pdftitle={Move Schedules},
-  pdfauthor={Matthew Piekenbrock and Jose A. Perea}
-}
-\fi
-
-\raggedbottom
+% \ifpdf
+% \hypersetup{
+%   pdftitle={Move Schedules},
+%   pdfauthor={Matthew Piekenbrock and Jose A. Perea}
+% }
+% \fi
+
+% \raggedbottom
 
@@
  	\includegraphics[width=0.75\textwidth]{circle_vineyards_vector.pdf}
-%	\includesvg{circle_vineyards}
  	\caption{The cumulative column operations needed to compute persistence across the time-varying filtration of grayscale images.
@@
 Using $\mathbb{Z}_2$ coefficients for simplicity, the reduction proceeds to compute $H_1(K)$ as follows:
-\begin{displaymath}
+$$
 	\def\rowsp{0.4em}
@@
 	\end{blockarray}
-	\vspace*{-1em}
-\end{displaymath}
+$$
 \noindent Since column $c$ in $R_1$ is $0$, the chain $d_1 + d_2 + d_3$ indicated by the column $c$ in $V_1$ represents a dimension $1$ cycle.
@@
 Below, we illustrate the case of exchanging simplices $a$ and $b$ in the filtration order, and restoring $RV$ to a valid decomposition.
-\begin{displaymath}
+$$
 	\def\rowsp{0.2em}
@@
 	\end{blockarray}
-\end{displaymath}
+$$
 % End R1
@@
 % Start V1 
-\begin{displaymath}	
+$$
 	\def\rowsp{0.2em}
@@
 	\end{blockarray}
-\end{displaymath}
+$$
 %\begin{displaymath}
@@
 Consider moving edge $a$ to the position of edge $c$ in the filtration.
-\begin{displaymath}
+$$
 	\def\rowsp{0.4em}
@@
 \end{blockarray}
-\end{displaymath}
+$$
 
@@
 % start V example (1) 
-\begin{displaymath}
+$$
 	\def\rowsp{0.4em}
@@
 \end{blockarray}
-\end{displaymath}
+$$
+
 Note that the equivalent permutation using vineyards requires $4$ column operations on both $R_1$ and $V_1$, respectively, whereas a single move operation accomplishes using only $2$ column operations per matrix.
@@
  %under certain insertion and deletion operations.
-
- %We know that
 
@@
 
- \algnewcommand{\IfThenElse}[3]{% \IfThenElse{<if>}{<then>}{<else>}
-  \State \algorithmicif\ #1\ \algorithmicthen\ #2\ \algorithmicelse\ #3}
   \begin{algorithm}[h!]

### 2023-01-31 move_schedules.tex
amending prop 2 slightly to be more transparent
typo fixes

@@
 \end{equation}
-We expect $\widetilde{S}_F$ to be cheaper to execute than $S_F$, with the tradeoff  that less diagrams are produced.
+We expect $\widetilde{S}_F$ to be cheaper to execute using move operations than $S_F$, with the tradeoff  that less diagrams are produced.
 We make this precise next:
@@
 	Let $K$ denote a filtration of size $\lvert K \rvert = m$ with decomposition $R = D V$, $T$ the maximum number of $O(m)$ operations required to execute $S_F$ via vineyards, and $M$ be the maximum number of $O(m)$ operations required to execute $\widetilde{S}_F$ with moves.
-Then, the inequality $ M \leq {\textstyle \frac{1}{2} \, T}$ holds.
+Then, if all pairs $(i, j) \in \tilde{S}_F$ satisfy $i < j$ and $\lvert i - j \rvert > 1$, the inequality $ M \leq {\textstyle \frac{1}{2} \, T}$ holds.
 \end{proposition}

### 2023-02-02 move_schedules.tex
overleaf changes from jose

@@
 	\vspace{-2.0em} Move Schedules: Fast persistence computations in coarse dynamic settings
-	\thanks{
+	\footnote{
 	\textbf{Funding:} This work was partially supported by the National Science Foundation through grants CCF-2006661 and CAREER award DMS-1943758.
@@
 Moreover, acquiring such a capability has far-reaching consequences: methods that vectorize persistence diagrams for machine learning purposes all immediately become computationally viable tools in dynamic settings.
-Persistence-based examples include adaptive template functions~\cite{polanco2019adaptive}, persistence images~\cite{adams2017persistence}, and $\alpha$-smoothed Betti curves~\cite{ulmer2019topological}.
+Such persistence summaries include adaptive template functions~\cite{polanco2019adaptive}, persistence images~\cite{adams2017persistence}, and $\alpha$-smoothed Betti curves~\cite{ulmer2019topological}.
 %This analogy provides a strong interpretation at gaining insight on the subtle topological changes that may occur in studying a continuous process.
@@
 Cohen-Steiner et al. refer to a continuous 1-parameter family of persistence diagrams as a \emph{vineyard}, and they give in \cite{cohen2006vines} an efficient algorithm for their computation. 
-The vineyards approach can be interpreted as an extension of the \emph{reduction} algorithm~\cite{zomorodian2005computing}, which computes the persistence diagram of a fixed filtration $K$ with $m$ simplices in $O(m^3)$ time via a decomposition $R = D V$ (or $RU = D$) of the boundary matrix $D$ of $K$.
+The vineyards approach can be interpreted as an extension of the \emph{reduction} algorithm~\cite{zomorodian2005computing}, which computes the persistence diagram of a filtered simplicial complex $K$ with $m$ simplices in $O(m^3)$ time, via a particular decomposition $R = D V$ (or $RU = D$) of the boundary matrix $D$ of $K$.
 %Since $V$ is always full rank and upper-triangular, an alternative is to compute a $D = RU$ decomposition, where $U = V^{-1}$. 
-In particular, the vineyards algorithm transforms a time-varying filtration into a certain set of permutations of the decomposition $R = DV$, each of which takes at most $O(m)$ time to execute.
+The vineyards algorithm, in turn,  transforms a time-varying filtration into a certain set of permutations of the decomposition $R = DV$, each of which takes at most $O(m)$ time to execute.
 If one is interested in understanding how the persistent homology of a continuous function changes over time, then this algorithm is sufficient, for homological critical points can only occur when the filtration order changes.
@@
 many transpositions can be quite slow...
-it is sometimes much faster to simply recompute the $RU$-decomposition from scratch using the standard persistence algorithm.'' Indeed, they observed maintaining the decomposition along a certain parameterized family is the most computationally demanding aspect of RIVET~\cite{rivet}, software for computing and visualizing	 two-parameter persistent homology.
-
-This work seeks to further understand and remedy this discrepancy: building on the work presented in~\cite{busaryev2010tracking}, we introduce a coarser approach to the vineyards algorithm.
+it is sometimes much faster to simply recompute the $RU$-decomposition from scratch using the standard persistence algorithm.'' Indeed, they observe that maintaining the decomposition along a certain parameterized family is the most computationally demanding aspect of RIVET~\cite{rivet}, a software for computing and visualizing	 two-parameter persistent homology.
+
+The work presented here seeks to further understand and remedy this discrepancy: building on the work presented in~\cite{busaryev2010tracking}, we introduce a coarser approach to the vineyards algorithm.
 While vineyards is  designed  for handling a \emph{continuous} 1-parameter family of diagrams, it is not necessarily efficient when the parameter is coarsely discretized.
@@
 Another popular optimization is to utilize the duality between   homology and cohomology~\cite{de2011dualities}, which dramatically improves the effectiveness of the clearing optimization~\cite{bauer2021ripser}.
-There are many other optimizations on the implementation side: the use of ranking functions defined on the combinatorial number system enables implicit cofacet enumeration, removing the need to store the boundary matrix explicitly; the apparent/emergent pairs optimization identifies columns whose pivot entries are unaffected by the reduction algorithm, reducing the total number of columns which need be reduced; sparse data structures such as bit-trees and lazy heaps allow for efficient column-wise additions with $\mathbb{Z}_2 = \mathbb{Z}/2\mathbb{Z}$ coefficients and effective $O(1)$ pivot entry retrieval, and so on~\cite{bauer2021ripser, bauer2017phat}.
+There are many other optimizations on the implementation side: the use of ranking functions defined on the combinatorial number system enables implicit cofacet enumeration, removing the need to store the boundary matrix explicitly; the apparent/emergent pairs optimization identifies columns whose pivot entries are unaffected by the reduction algorithm, reducing the total number of columns which need to be reduced; sparse data structures such as bit-trees and lazy heaps allow for efficient column-wise additions with $\mathbb{Z}_2 = \mathbb{Z}/2\mathbb{Z}$ coefficients and effective $O(1)$ pivot entry retrieval, and so on~\cite{bauer2021ripser, bauer2017phat}.
 
@@
 
-If  zeroth homology ($p = 0$) is the only dimension of interest, computing and updating both the persistence and rank information  is greatly simplified.
+If  zeroth homology   is the only dimension of interest, computing and updating both the persistence and rank information  is greatly simplified.
 For example, if the relations are available a-priori, obtaining a tree representation fully characterizing the connectivity of the underlying space (also known as the \emph{incremental connectivity} problem) takes just $O(\alpha(n) n)$ time using the disjoint-set data structure, where $\alpha(n)$ is the extremely slow-growing inverse Ackermann function.
@@
  \subsection{A Motivating Example}\label{sec:motivation} 
-We illustrate why the vineyards algorithm does not always yield an efficient strategy for time-varying setting with a simple experiment. 
-Consider a series of grayscale images (i.e. a video) depicting a fixed-width annulus expanding about the center of a $9 \times 9$ grid and its associated sublevel-set filtrations:   
-\begin{center}
-	\vspace{1em}
+We  will use a simple experiment to illustrate why the vineyards algorithm does not always yield an efficient strategy for time-varying settings. 
+Consider a series of grayscale images (i.e.
+a video) depicting a fixed-width annulus expanding about the center of a $9 \times 9$ grid, and its associated sublevel-set filtrations, as shown in Figure~\ref{fig:vidExample}.
+
+\begin{figure}[!htb]
+    \centering
     \includegraphics[height=1.25in]{circle_complex.png}
-\end{center}
+    \caption{Top: A video of an expanding annulus. Bottom: Sublevel-set filtrations, via pixel intensity, of a Freudenthal triangulation of the plane.}
+    \label{fig:vidExample}
+\end{figure}
+
 Each image contains a fixed number of pixels whose intensities vary with time.
@@
 %Note that the size of the filtration is fixed---no simplices are added or removed over time. 
-Two events critically change the pairing function: the first occurs when the central connected component splits to form a cycle, and the second  when the annulus  splits into four components. 
+Two events critically change the persistence diagrams: the first occurs when the central connected component splits to form a cycle, and the second  when the annulus  splits into four components. 
 From left to right, the Betti numbers of the five evenly spaced `snapshots' of the filtration shown above are: $(\beta_0,\beta_1)  = (1,0), \; (1,1),  \; (1,1), \; (1,1), \; (4,0)$.
 Thus, in this example, only a few persistence diagrams are needed to capture the major changes to the topology.
-\begin{figure*}[t]
+
+We use this data set as a baseline for comparing vineyards and the standard reduction algorithm $\texttt{pHcol}$ (Algorithm \ref{alg:reduce}).
+Suppose a practitioner wanted to know the major homological changes a time-varying filtration encounters over time.
+Since it is unknown a priori when the persistent pairing function changes, one solution is to do $n$ independent persistence computations at $n$ evenly spaced points in the time domain.
+An alternative approach is to construct a homotopy between a pair of filtrations $(K, f)$, $(K,f')$ and then decompose this homotopy into adjacent transpositions based on the filtration order---the vineyards approach.
+ We refer to the former as the \emph{discrete setting}, which is often used in practice, and the  latter as the \emph{continuous setting}.
+Note that though the discrete setting is often more practical, it is not guaranteed to capture all homological changes in persistence that occur in simulating a continuous 1-parameter family of diagrams.
+
+The cumulative cost (in total column operations) of these various approaches are shown in Figure~\ref{fig:vineyards}, wherein the reduction (\texttt{pHcol}) and vineyard algorithms are compared.
+Two discrete strategies (green and purple) and two continuous strategies (black and blue) are shown.
+
+\begin{figure*}[!htb]
  	\centering
@@
 
-  We use this data set as a baseline for comparing vineyards and the standard reduction algorithm $\texttt{pHcol}$ (Algorithm \ref{alg:reduce}).
-Suppose a practitioner wanted to know the major homological changes a time-varying filtration encounters over time.
-  Since it is unknown a priori at when the persistent pairing function changes, one solution is to do $n$ independent persistence computations at $n$ evenly spaced points in the time domain.
-An alternative approach is to construct a homotopy between a pair of filtrations $(K, f)$, $(K,f')$ and then decompose this homotopy into adjacent transpositions of the filtration order---the vineyards approach.
- We refer to the former as the \emph{discrete setting}, which is often used in practice, and the  latter as the \emph{continuous setting}.
-Note that though the discrete setting is often more practical, it is only not guaranteed to capture all homological changes in persistence that occur in simulating a continuous 1-parameter family of diagrams.
-
- The cumulative cost (in total column operations) of these various approaches are shown on the left side of Figure~\ref{fig:vineyards}, wherein the reduction algorithm (\texttt{pHcol}) and the vineyards algorithm are compared.
-Two discrete strategies (green and purple) and two continuous strategies (black and blue) are shown.
 Note that without knowing where persistence pairing function changes, a continuous strategy must construct all $\approx 7 \times 10^{4}$ diagrams induced by the homotopy. 
@@
 
-  The existence of a time discretization that is more efficient to compute than continually updating the decomposition indicates that the vineyards framework must incur some overhead (in terms of column operations) to maintain the underlying decomposition, even when pairing function is unchanged.
+The existence of a time discretization that is more efficient to compute than continually updating the decomposition indicates that the vineyards framework must incur some overhead (in terms of column operations) to maintain the underlying decomposition, even when the pairing function determining the persistence diagram is unchanged.
 Indeed, as shown by the case where $n = 10$, applying \texttt{pHcol} independently between  relatively ``close'' filtrations is substantially more efficient than iteratively updating the decomposition.
@@
  \noindent
-\textbf{Our approach and contributions are as follows:} First, we leverage the \emph{moves} framework of Busaryev et al.~\cite{busaryev2010tracking} to include  coarser  operations---in terms of the number of valid intermediate decomposition states, as compared to vineyards---for dynamic persistence.
-We give a tight lower bound on the number of moves needed to perform an arbitrary permutation to the $R = D V$ decomposition, and give a proof of optimality by a reduction to the permutation edit distance problem.
-We also give worst-case sizes of these quantities in expectation as well as efficient algorithms for achieving these bounds---both of which are derived from a reduction to the Longest Increasing Subsequence (LIS) problem.
-This reduction yields an efficient algorithm for generating sequences of moves $( s_1, s_2, \dots, s_d )$ of minimal size $d$, which we call \emph{schedules}. 
+\textbf{Our approach and contributions are as follows:} First, we leverage the \emph{moves} framework of Busaryev et al.~\cite{busaryev2010tracking} to include  coarser  operations for dynamic persistence settings.
+By a reduction to an edit distance problem, we give a   lower bound on the number of moves needed to perform an arbitrary permutation to the $R = D V$ decomposition,  along with a proof of its optimality.
+We also give worst-case sizes of these quantities in expectation as well as efficient algorithms for constructing these operations---both of which are derived from a reduction to the Longest Increasing Subsequence (LIS) problem.
+These operations parameterize sequences of cyclic permutations $( s_1, s_2, \dots, s_d )$ of minimal size $d$, which we call \emph{schedules}. 
 However, not all minimal size schedules incur the same  cost 
@@
 greedy-type approaches can lead to arbitrarily bad behavior. 
-In light of these results, we give an alternative proxy-objective for cost minimization, provide bounds justifying its relevance to the original problem, and give an efficient $O(d^2 \log m)$ algorithm for approximately solving this proxy minimization.
+In light of these results, we give an alternative proxy-objective for cost minimization, provide bounds justifying its relevance to the original problem, and give an efficient $O(d^2 \log m)$ algorithm for heuristically solving this proxy minimization.
 A performance comparison with other reduction-based persistence computations is given, wherein move schedules are demonstrated to be an order of magnitude more efficient than existing approaches at calculating persistence in dynamic settings.
@@
 \subsection{Main results}
-Given a simplicial complex $K$ with filtration function $f : K \to [m]$, denote by $R = D V$ the decomposition of its corresponding boundary matrix $D$ such that $R$ is reduced and $V$ is upper-triangular (see section~\ref{sec:reduction} for details).
- If one has a pair of filtrations $(K, f)$,  $(K, f')$ of size $m = \lvert K \rvert$ and $R = D V$ has been computed for $K$, then it may be advantageous to use the information stored in ($R$, $V$) to reduce the computation of $R' = D' V'$.
-Given a permutation $P$ such that $D' = P^T D P$, such an update scheme has the form: 
-$$ (\ast P^T \ast R \ast P \ast) = (P^T DP)(\ast P^T \ast V \ast P \ast) $$
+Given a simplicial complex $K$ with filtration function $f$, denote by $R = D V$ the decomposition of its corresponding boundary matrix $D$ such that $R$ is reduced and $V$ is upper-triangular (see section~\ref{sec:reduction} for details).
+ If one has a pair of filtrations $(K, f)$,  $(K, f')$ of size $m = \lvert K \rvert$ and $R = D V$ has been computed for $(K,f)$, then it may be advantageous to use the information stored in ($R$, $V$) to reduce the computation of $R' = D' V'$.
+Given a permutation $P$ such that $D' = P D P^T$, such an update scheme has the form: 
+$$ (\ast P \ast R \ast P^T \ast) = (P DP^T)(\ast P \ast V \ast P^T \ast) $$
 where $\ast$ is substituted with elementary column operations that repair the permuted decomposition.
@@
 Since each column operation takes $O(m)$, the complexity of reindexing $f \mapsto f'$ is $O(m^3)$, which is efficient if all $d$ decompositions are needed.
-Otherwise, if only $(R', V')$ is needed (and none of the intermediate $d -1$ decompositions), updating $R \mapsto R'$ using the approach from~\cite{cohen2006vines} matches the complexity of computing $R' = D' V'$ independently.
-
-We now summarize the main results obtained in this effort (Theorem~\ref{prop:sparsity_result}): suppose one has a sequence of permutations $\mathcal{S} = \left( s_1, s_2, \dots, s_d \right)$ yielding a corresponding sequence of decompositions:
+Otherwise, if only $(R', V')$ is needed, updating $R \mapsto R'$ using the approach from~\cite{cohen2006vines} matches the complexity of computing $R' = D' V'$ independently.
+
+We now summarize the main results obtained in this effort (Theorem~\ref{prop:sparsity_result}): suppose one has a sequence of permutations, i.e.
+a schedule,  $\mathcal{S} = \left( s_1, s_2, \dots, s_d \right)$ yielding a corresponding sequence of decompositions:
   \begin{equation}\label{eq:rv_seq}
@@
  \end{equation}
- where each permutation $s_i$ is a particular type of cyclic permutation (see section~\ref{sec:moves_setting}).
-If $i_\ast < j_\ast$, our first result extends~\cite{busaryev2010tracking} by showing that~\eqref{eq:rv_seq} can be computed in using $O(\kappa)$ column operations, where:
+ where each permutation $s_k = (i_k, j_k)$ is a particular type of cyclic permutation (see section~\ref{sec:moves_setting}).
+If $i_\ast < j_\ast$ for all $s_\ast \in \mathcal{S}$, our first result extends~\cite{busaryev2010tracking} by showing that~\eqref{eq:rv_seq} can be computed  using $O(\kappa)$ column operations, where:
 \begin{equation}
@@
 \end{equation}
-where the quantities $\lvert \mathbb{I}_{k} \rvert$ and $\lvert \mathbb{J}_{k} \rvert$ depend on the sparsity of the $V_k$ and $R_k$ matrices, respectively, and $d \sim O(m)$ is a constant that depends on how similar $f$ and $f'$ are.
+The quantities $\lvert \mathbb{I}_{k} \rvert$ and $\lvert \mathbb{J}_{k} \rvert$ depend on the sparsity of the $V_k$ and $R_k$ matrices, respectively, and $d \sim O(m)$ is a constant that depends on how similar $f$ and $f'$ are.
 The advantage of this result is that it depends explicitly on the sparsity pattern of the decomposition itself and is thus output sensitive, which we take advantage of in Section~\ref{sec:schedule_cost}.
@@
 
-Our second result turns towards upper bounding $d = \lvert \mathcal{S} \rvert$ and the complexity of constructing $\mathcal{S}$ itself. 
+Our second result turns towards lower bounding $d = \lvert \mathcal{S} \rvert$ and the complexity of constructing $\mathcal{S}$ itself. 
 %The complexity of determining $\mathcal{S}$ depends on the type of permutations considered; many edit distances exhibit tight quadratic growth rates and tend to require at least quadratic time (or worse) to compute~\cite{labarre2013lower}.
-By reinterpreting a special set of cyclic permutations as edit operations on strings, we find that the \emph{smallest} such sequence that maps $f \mapsto f'$ has size $d$ (Proposition~\ref{prop:schedule_size}), where:
+By reinterpreting a special set of cyclic permutations as edit operations on strings, we find that the \emph{smallest} such sequence that maps $f \mapsto f'$ has size  (Proposition~\ref{prop:schedule_size}) :
 \begin{equation}
@@
 Although $d$ clearly is $O(m)$ for pathological inputs, we give evidence  that $d \sim m - \sqrt{m}$ in expectation for random filtrations is not too pessimistic for truly random filtrations (Corollary~\ref{cor:expectation}), and we give empirical results suggesting $d$ is tends to be much smaller for time-varying filtrations.
-
 %Finally, we perform extensive experiments 
@@
 % 
-
  %The actual permutations $s_\ast$ are generated in $O(m \log m)$ time prior to the execution of the column operations.  
-
 	% where $\lvert \mathbb{I}_{(k)} \rvert$ and $\lvert \mathbb{J}_{(k)} \rvert$ are the computed quantities in lines (2) and (3) in algorithm~\ref{alg:mr} (here we assume if $i_\ast < j_\ast$).
 %	where $\lvert \mathbb{I}_{(k)} \rvert$ denotes the number of non-zero entries $V_k[i_k, i_k:j_k]$ and $\lvert \mathbb{J}_{(k)}\rvert$ are the number of columns satisfying $\mathrm{low}_{R_k} \in [i_k:j_k]$ and $\mathrm{row}_{R_k}(i_k) \neq 0$.
-
 %We mention in passing that the vineyards algorithm produces a similar sequence $\mathcal{S}'$ of size $\lvert \mathcal{S'} \rvert = d$ where $d = K_\tau(K,K')$ in output sensitive $O(m\log m + d)$ time and $O(d)$ memory, where $d = K_\tau(K,K')$ is the output sensitive terms denoting the Kendall-$\tau$ distance between the two filtrations $K$ and $K'$.
@@
 In comparison, our proposed approach yields a smaller decomposition sequence using less running time and memory, and conveys more directly how much the sparsity of the corresponding decompositions affect the computation.
-
 %Below, we give a corollary of the main result which conveys more about the worst-case size of $\mathcal{S}$. 
@@
 Due to the face relation requirement of abstract simplicial complexes, one might expect that this bound might be much smaller, however for low-dimensional complexes we give evidence that this tightness gap in tightness may not be so large, see section~\ref{sec:schedule_sizes} for more details.
-
 %\begin{proposition}\label{prop:factor2_main}
@@
 %\end{proposition}
+\\
+\\
 \noindent 
+% Goal: Interesting+ hard + solved + novel.
 \textbf{Outline:} The paper is organized as follows: we review and establish the notations we will use to describe simplicial complexes, persistent homology, and dynamic persistence in Section~\ref{sec:background}.
@@
 In Section \ref{sec:conclusion} we conclude the paper by discussing other possible applications and future work.
-
-%\section*{Outline}
-%\begin{outline}[enumerate]
-%   \1 Background
-%      \2 Point A1
-%      \2 Point A2
-%         \3 ...
-%            \4 ...
-%   \1 Point B
-%      \2 Point B1
-%         \3 Point B1a
-%            \4 ...
-%\end{outline}
-% Interesting+ hard + solved + novel.
 
@@
 $p$-th homology of $K$ with coefficients in $\mathbb{F}$. 
-If $\{K_i\}_{i\in [m]}$ is a filtration, then the inclusion maps  $K_i\subset K_{i+1}$   induce linear transformations 
+If $ f: K \rightarrow  [m]$ is a filtration, then the inclusion maps  $K_i\subseteq K_{i+1}$   induce linear transformations 
 %$f_p^{i,j}: H_p(K_i) \to H_p(K_j)$ 
@@
 The collection of birth/death  pairs 
-$(i,j)$ is denoted $\mathrm{dgm}_p(K)$, 
-and referred to as the $p$-th \emph{persistence diagram} of $K$.
+$(i,j)$ is denoted $\mathrm{dgm}_p(K ,f )$, 
+and referred to as the $p$-th \emph{persistence diagram} of $(K,f)$.
 If a homology class is born at time $i$ and dies entering time $j$, the difference $\lvert i - j \rvert$ is called the \emph{persistence} of that class.
@@
 \noindent
-\textbf{Example 2.1: Reduction} Consider a triangle with vertices $u,v,w$, edges $a = (u,w)$, $b = (v,w)$, $c = (u, v)$,
+\textbf{Example 2.1:} Consider a triangle with vertices $u,v,w$, edges $a = (u,w)$, $b = (v,w)$, $c = (u, v)$,
 and whose filtration order is given as $K = (u, v, w, a, b, c)$. 
@@
 %If the $i$-th column of $R$ is nonzero, then $\mathrm{low}_R(i)$  is the row index of its lowest nonzero entry.
-Using $\mathbb{Z}_2$ coefficients for simplicity, the reduction proceeds to compute $H_1(K)$ as follows:
+Using $\mathbb{Z}_2$ coefficients, the reduction proceeds to compute $H_1(K)$ as follows:
 $$
@@
 As $\tau$ varies in $[0,1]$, the points in $\mathrm{dgm}_p(f_\tau)$ trace curves in $\mathbb{R}^3$, which will be continuous if $F$ is continuous and the $f_\tau$'s are tame, due to the stability of persistence.
-Cohen-steiner et al.~\cite{cohen2007stability} referred to these curves as \emph{vines}, as collection of which forms as \emph{vineyard}---the geometric analogy meant to act as a guidepost for practitioners seeking to understand how subtle changes occurring to the topological structure over time.
+Cohen-steiner et al.~\cite{cohen2007stability} referred to these curves as \emph{vines}, a collection of which forms as \emph{vineyard}---the geometric analogy meant to act as a guidepost for practitioners seeking to understand the evolution of topological structure over time.
 
@@
 
-These rules prescribe certain column and row operations to apply to a given matrix decomposition either before, during, or after each transposition to ensure the decomposition is valid.   
-Formally, let $S_{i}^j$ represent the upper-triangular matrix such that $A S_{i}^j$ is equivalent to the elementary operation that adds column $i$ of $A$ to column $j$ of $A$ and $S_{i}^j A$ is equivalent to adding row $j$ of $A$ to row $i$ of $A$.
-Similarly, let $P$ denote the matrix so that $A P$ applies a fixed permutation to the columns of $A$ and $P^T A$ applied the same permutation to the rows of $A$.
-Since the columns of $P$ are orthonormal, then  $P^{-1} = P^T$, and thus one would write $P^T A P$ to denote the application of the same permutation to both the columns and rows of $A$.
+To ensure a decomposition is valid, these rules prescribe certain column and row operations to apply to a given matrix decomposition either before, during, or after each transposition.   
+Formally, let $S_{i}^j$ represent the upper-triangular matrix such that $A S_{i}^j$ is equivalent to the elementary operation that adds column $i$ of $A$ to column $j$ of $A$ and $S_{i}^j A$ the same operation on rows $i$ and $j$.
+Similarly, let $P$ denote the matrix so that $A P^T$ permutes the columns of $A$ and $P A$ permutes the rows.
+Since the columns of $P$ are orthonormal, $P^{-1} = P^T$, and thus writes $P A P^T$ to denote the same permutation of both the columns and rows of $A$.
 In the special case where $P$ represents a transposition, we have $P = P^T$ and may instead simply write $P A P$.
@@
 However, a host of other factors affect the runtime efficiency of the vineyards algorithm. 
-On the positive side, as both the $V$ and $R$ matrices tend to be quite sparse, the cost of a given column operation is proportional the number of non-zero field coefficients in the two columns being modified.
+On the positive side, as both the $V$ and $R$ matrices tend to be quite sparse, the cost of a given column operation is proportional to the number of non-zero  entries in the two columns being modified.
 As a rule of thumb, most transpositions require no column operations~\cite{edelsbrunner2000topological}.
@@
 \subsection{Moves}\label{sec:moves} Originally developed to accelerate tracking generators with temporal coherence, Busaryev et al.~\cite{busaryev2010tracking} introduced an extension of the vineyards algorithm which maintains a $R = D V$ decomposition under \emph{move operations}.
-A move operation $\mathrm{Move}(i,j)$ is a set of rules for maintaining a valid decomposition under the application of a permutation $P$ that moves a simplex $\sigma_i$ at position $i$ to position $j$.
-If $j = i \pm 1$, this operation is an adjacent transposition, and thus in some sense moves generalizes vineyards.
-However, despite being conceptually equivalent to performing a sequence of contiguous transpositions, the move operation presented by Busaryev exhibits several attractive qualities distinct from the vineyards approach that warrants further study.
+A move operation $\mathrm{Move}(i,j)$ is a set of rules for maintaining a valid decomposition under the permutation $P$ that moves a simplex $\sigma_i$ at position $i$ to position $j$.
+If $j = i \pm 1$, this operation is an adjacent transposition, and in this sense moves generalizes vineyards.
+However, the move framework presented by Busaryev is actually distinct in that it exhibits several attractive qualities not inherited by the vineyards approach that warrants further study.
 
 For completeness, we recapitulate the motivation of the moves algorithm from~\cite{busaryev2010tracking}. 
-Let $K$ denote a filtration of size $m = \lvert K \rvert$ and $R = DV$ its decomposition, where $R = [r_1, r_2, \dots, r_m]$ and $V = [v_1, v_2, \dots, v_m]$ denote the columns of $R$ and $V$, respectively.
+Let $f: K \rightarrow [m]$ denote a filtration of size $m = \lvert K \rvert$ and $R = DV$ its decomposition, where $R = [r_1, r_2, \dots, r_m]$ and $V = [v_1, v_2, \dots, v_m]$ denote the columns of $R$ and $V$, respectively.
 By definition, if $r_j = D v_j = 0$, then the inclusion $K_{j-1} \hookrightarrow K_{j}$ introduced a new cycle whose generator is given by:
@@
 To perform this shift, the non-zero coefficients from~\eqref{eq:coefficients} must be set to zero, otherwise $\partial(\sigma_i)$ contributes to boundary chains $r_{i+1}, \dots, r_j$ earlier in the filtration.
-As these coefficients are recorded by $v_j$, setting $\alpha_i^k = 0$  amounts to setting the row entries $v_k(i) = 0$ for all $k \in [i+1,j]$---these cancellations manifest as column operations $(\ast)$ used to ensure $PV(\ast)$ is upper-triangular, thus maintaining invariant I2.
-Notice that these operations may induce an unreduced $R' = P R(\ast) P^T$, breaking invariant I3.
-We could reduce $R'$ with an additional $k$ operations, however its possible that $k \sim O(\lvert i - j\rvert^2)$, rendering the efficiency of this approach at best that of vineyards and at worse that of the standard algorithm.
+As these coefficients are recorded by $v_j$, setting $\alpha_i^{(k)} = 0$  amounts to setting the row entries $v_k(i) = 0$ for all $k \in [i+1,j]$---these cancellations manifest as column operations $(\ast)$ used to ensure $P V(\ast)$ is upper-triangular, thus maintaining invariant I2.
+Notice that these operations may yield an unreduced $R' = P R(\ast) P^T$, breaking invariant I3.
+We could reduce $R'$ with an additional $k$ operations, however its possible that $k \sim O(\lvert i - j\rvert^2)$, which is no more efficient than re-running the reduction algorithm on columns $[i,j]$ in $R$.
 
 To bypass this difficulty, Busaryev et al. observed that since the initial $R$ is reduced, if it contains $s$ pivot entries in the columns $[i,j]$ of $R$, then $R'$ must also have $s$ pivots. 
-In particular, if during the cancellation  of $v_{i+1}(i) = v_{i+2}(i) = 0$ the column $r_{i+2} \mapsto r_{i+2}'$ becomes unreduced, then the pivot $\mathrm{low}_R(i+2)$ becomes a \emph{free}, possibly becoming a pivot again in $r_{i+3}', \dots, r_j'$.
-Thus, if $r_{i+2}$ is copied prior to modification to a \emph{donor column}, it may re-use or \emph{donate} its pivot entry to a later column $r_{i+3}, \dots, r_j$. 
+Thus, if during the cancellation  of $v_{i+1}(i) = v_{i+2}(i) = 0$ the column $r_{i+2} \mapsto r_{i+2}'$ becomes unreduced, then the pivot $\mathrm{low}_R(i+2)$ becomes \emph{free}, possibly becoming a pivot later in $r_{i+3}', \dots, r_j'$.
+If $r_{i+2}$ is copied prior to modification to a \emph{donor column}, it may re-use or \emph{donate} its pivot entry to a later column $r_{i+3}, \dots, r_j$. 
 Repeating this process at most $j - i - 1$ times ensures $R'$ stays reduced in all except possibly for the $i$-th column by the end of the movement---and since the $k$-th such operation simultaneously sets $v_{k}(i) = 0$ without creating non-zeros at indices $v_{k}(j)$ for $j > k$, $V'$ retains its upper-triangularity.
@@
 $$
-
+\noindent 
 Note that the equivalent permutation using vineyards requires $4$ column operations on both $R_1$ and $V_1$, respectively, whereas a single move operation accomplishes using only $2$ column operations per matrix.
@@
 \begin{proposition}[Busaryev et al.~\cite{busaryev2010tracking}]\label{prop:move_st}
-	Let $K$ denote a filtration with $n$ simplices of dimension $p-1$, $p$, and $p+1$, respectively, and let $R = DV$ denote its associated decomposition.
-Then the operation $\mathrm{MoveRight}(i,j)$ constructs a valid decomposition $R' = D'V'$ in $O((\lvert \mathbb{I} \rvert + \lvert \mathbb{J} \rvert)n)$ time, where $\mathbb{I}, \mathbb{J}$ are given by:
+	Given a filtration with $n$ simplices of dimensions $p-1$, $p$, and $p+1$,    let $R = DV$ denote its associated decomposition.
+Then, the operation $\mathrm{MoveRight}(i,j)$ constructs a valid decomposition $R' = D'V'$ in $O((\lvert \mathbb{I} \rvert + \lvert \mathbb{J} \rvert)n)$ time, where $\mathbb{I}, \mathbb{J}$ are given by:
 	$$ \lvert \mathbb{I}\rvert = \sum\limits_{l=i+1}^j \mathds{1}\left(\, v_l(i) \neq 0 \, \right), \quad \mathbb{J} = \sum\limits_{l=1}^m \mathds{1}\left(\, \mathrm{low}_R(l) \in [i,j] \text{ and } r_l(i) \neq 0 \, \right)$$
@@
  	\item[M2:] $R = D V$ is \underline{not} guaranteed to be valid during the movement of $\sigma_i $ to $ \sigma_j$.
- 	\item[M3:] At most $O(m)$ moves are needed to reindex $K \mapsto K'$ 
+ 	\item[M3:] At most $O(m)$ moves are needed to reindex $f \mapsto f'$ 
  \end{enumerate} 
@@
 In this way, we interpret move operations as making a tradeoff in granularity: whereas a sequence of adjacent transpositions $(i, i{+}1), (i{+}1, i{+}2), \dots, (j{-}1, j)$ generates $\lvert i - j \rvert $ valid decompositions in vineyards, an equivalent move operation $\mathrm{Move}(i,j)$ generates only one.
-Indeed, Property M3 directly follows from this fact, as one may simply move each simplex $\sigma \in K$ into its order in $K'$ via insertion sort.
-Note that the number of valid decomposition produced by vineyards is bounded above by $O(m^2)$ if each pair of simplices $\sigma_i, \sigma_j \in K$ switches its relative ordering at most once during the interpolation from $K$ to $K'$.
-
-As a final note, we remark the combination of \emph{MoveRight} and \emph{MoveLeft} enable the addition or deletion of simplices to the underlying complex.
-In particular, given $K$ and its decomposition $R = DV$, obtaining a valid decomposition $R' = D' V'$ of $K' = K \cup \sigma$ can be achieved by appending its requisite elementary chains to $D$ and $V$, reducing them, and then executing \emph{MoveLeft}$(m+1, i)$, where $i = f'(\sigma)$.
+Indeed, Property M3 directly follows from this fact, as one may simply move each simplex $\sigma \in K$ into its new order   $f'(\sigma)$ via insertion sort.
+Note that the number of valid decomposition produced by vineyards is bounded above by $O(m^2)$ if each pair of simplices $\sigma_i, \sigma_j \in K$ switches its relative ordering at most once during the interpolation from $f$ to $f'$.
+
+As a final remark, we note the combination of \emph{MoveRight} and \emph{MoveLeft} enable efficient simplex additions or deletions to the underlying complex.
+In particular, given $K$ and a decomposition $R = DV$, obtaining a valid decomposition $R' = D' V'$ of $K' = K \cup \{\sigma\}$ can be achieved by appending its requisite elementary chains to $D$ and $V$, reducing them, and then executing \emph{MoveLeft}$(m+1, i)$, where $i = f'(\sigma)$.
 Dually, deleting a simplex $\sigma_i$ may be achieved via \emph{MoveRight} by moving $i$-th to the end of the decomposition and dropping the corresponding columns.
@@
 %\subsection{Proposed approach}
-Let us begin with a brief overview of the pipeline, outlined in Algorithm~\ref{alg:schedule} below.
-As before, we assume as input  a discrete $1$-parameter family of filtrations $\mathcal{K} = (K_1, K_2, \dots, K_n)$ of an abstract simplicial complex $K$ with $\lvert K \rvert =m$, and the goal being to compute their persistence diagrams.
-Fix reindexing bijections $\rho_i : K_i \to K_{i+1}$. 
-Each $\rho_i$ induces a bijection $\rho_i^\ast : [m] \to [m]$, or equivalently, a permutation of the index set $[m]$.
-For each pair of filtrations $(K_i, K_{i+1})$, we assign the simplices of $K_i$ labels from index set $[m]$ in filtration order, relabeling $K_{i+1}$ accordingly using $\rho_i$.
-As $p$ is now a strictly increasing sequence, we may compute the longest increasing subsequence $\mathrm{LIS}(q)$ and use this subsequence to recover a longest common subsequence (LCS) between $(K_i, K_{i+1})$, which we denote later with $\mathrm{LCS}(K_i, K_{i+1})$.
+Let us begin with a brief overview of the pipeline, which we outline in Algorithm~\ref{alg:schedule} below.
+As before, we assume as input  a discrete $1$-parameter family $\mathcal{F} = (f_1, f_2, \dots, f_n)$ 
+of filtrations $f_i : K \rightarrow [m]$ of a simplicial complex $K$ with $\lvert K \rvert =m$, and the goal being to compute the  persistence diagrams of each $(K,f_i)$.
+Fix   bijections $\rho_i : [m] \to [m]$ so that 
+$f_{i+1} = \rho_i \circ f_i$, or equivalently,  permutations of the index set $[m]$. For each $\rho_i$ we let 
+$q = (\rho_i(1), \rho_i(2), \ldots, \rho_i(m) )$
+and compute the longest increasing subsequence $\mathrm{LIS}(q)$;  this subsequence  is used to recover a longest common subsequence (LCS) between $f_i$ and $ f_{i+1} $, which we denote later with $\mathrm{LCS}(f_i, f_{i+1})$.
 We pass $q$ and $\mathrm{LIS}(q)$ to our greedy scheduling algorithm, which returns as output an ordered set of move permutations  $\mathcal{S}$ of minimum size, which we call a \emph{schedule} (see Definition~\ref{def:schedule}).
-Note one need not explicitly store the entire family of filtrations nor the schedules between them---Algorithm~\ref{alg:schedule} may be easily modified to be completely \emph{online}, keeping at most two filtrations and one decomposition in memory at any given time.
+Note one need not explicitly store the family of filtrations nor the schedules between them---Algorithm~\ref{alg:schedule} may be easily modified to be completely \emph{online}, keeping at most two filtrations and one decomposition in memory at any given time.
 % Not all valid schedules are created equal: some incur a much larger number of column operations, even if they have a minimal number of moves; we address this issue in Section~\ref{sec:schedule_cost}.
@@
     \begin{algorithmic}[1]
-    	\Require Ordered set of filtrations $\mathcal{K}$ with bijections $f_{i}: K_i \to K_{i+1}$
-    	\Ensure $R = D V$ is computed for each  $K_1, K_2, \dots, K_n$
-    	\Procedure{MoveSchedule}{$\mathcal{K} = ( K_1, K_2, \dots, K_n )$} 
+    	\Require Ordered set of filtrations $\mathcal{F}$ with permutations $\rho_{i}: [m] \to [m]$
+    	\Ensure $R = D V$ is computed for each  $(K,f_1),(K,f_2), \dots, (K,f_n)$
+    	\Procedure{MoveSchedule}{$\mathcal{F} = ( f_1, f_2, \dots, f_n )$} 
     		\State $\mathcal{S} = \emptyset$
     		\For{$i = 1$\textbf{ to }$n-1$}
-    		    \State $(p, q) \gets ([m], \, \mathrm{Im}\, f^\ast_{i})$ 
+    		    \State $q \gets  (\rho_i(1), \rho_i(2), \ldots, \rho_i(m))$ 
     			\State $\mathrm{lis}_q \gets \mathrm{LIS}(q)$ \Comment{$O(m \log \log m)$}
@@
     		\EndFor
-	    	\State $(R, V) \gets$ \Call{Reduction}{$D = \partial K_1$}
+	    	\State $(R, V) \gets$ \Call{Reduction}{$D = \partial (K,f_1)$}
 	    	\For{$(i,j)$\textbf{ in }$\mathcal{S}$}
@@
 
-In the following subsections, we investigate whether we may profitably exploit the increased flexibility Busaryevs move framework yields us.
-% TODO: re-write 
-It was hypothesized that the reason the vineyards algorithm was more expensive than the naïve approach is due to the extra overhead of maintaining the decomposition at each transposition.
-Thus, in theory, decreasing the number of times the decomposition is restored to a valid state ought to reduce the total number of column operations needed to update a decomposition.
-This motivates the question: can one simultaneously minimize the number of times the decomposition is restored to a valid state and retain an efficient update scheme for \emph{arbitrary} permutations?
+In the following subsections, we investigate how to profitably exploit the increased flexibility Busaryev's move framework yields us, beginning with the hypothesis formed experimentally from section~\ref{sec:motivation} that vineyards exhibits extraneous overhead maintaining the decomposition.
+In theory, decreasing the number of times the decomposition is restored to a valid state ought to reduce this overhead, motivating the question: can one simultaneously minimize the number of times the decomposition is restored to a valid state while retaining an efficient update scheme?
 
@@
 \begin{definition}[Schedule]\label{def:schedule}
-Given a pair of filtrations $(K, f), (K,f)$ 
+Given a pair of filtrations $(K, f), (K,f')$ 
 %and a fixed bijection $\rho : f \to f'$, let $\rho^\ast : [m] \to [m]$ denote its induced bijection on $S_m$ 
@@
    	\end{equation}
-   	where, for each $i \in [d]$, $R_i = D_i V_i$ is a valid decomposition respecting invariants~\ref{inv:decomposition}, and $R'$ is a valid decomposition of $K'$.
+   	where, for each $i \in [d]$, $R_i = D_i V_i$ is a valid decomposition respecting invariants~\ref{inv:decomposition}, and $R'$ is a valid decomposition for $(K,f')$.
 \end{definition}
@@
 One might hope computing the size $\mathrm{d}_{\mathrm{lcs}}(p,q)$ alone exhibits a lower complexity, however there is substantial evidence that for general string inputs the complexity cannot be much lower than quadratic~\cite{abboud2015tight}.
-However, if the pair of filtrations $K, K'$ are from the same underlying complex $S$, then they may both be thought of a permutations in $S_m$---the corresponding edit distance $d$ then reduces to the \emph{permutation edit distance} problem.
+However, if the pair of filtrations $(K, f), (K, f')$ come from the same underlying complex $K$, then their filter function $f, f'$ may both be thought of a permutations in $S_m$---the corresponding edit distance $d$ then reduces to the \emph{permutation edit distance} problem.
 With this insight in mind, we obtain the following bound on the minimum size of a sorting (i.e. schedule) using moves and the complexity of computing it. 
@@
 \begin{corollary}\label{cor:expectation}
-If $(K, f), (K,f')$ are random filtrations of a common complex $K$ of size $m$, then the size of longest common subsequence of simplices between $f,f'$ is no larger than $m - \sqrt{m}$ in expectation, with probability $1$ as $m \to \infty$.
+If $(K, f), (K,f')$ are random filtrations of a common complex $K$ of size $m$, then the expected size of longest common subsequence of simplices between $f,f'$ is no larger than $m - \sqrt{m}$, with probability $1$ as $m \to \infty$.
 \end{corollary}
@@
 	\includegraphics[width=0.48\textwidth]{varying_nd.pdf}
-	\caption{Performance data }
+	\caption{Performance comparison }
 	\label{fig:movie_perf}

### 2023-02-02 move_schedules.tex
refined propisition 2 to make it clearer, fixed typos in 4.1, added figure caption, and eased one transition paragraph

@@
 If $r_{i+2}$ is copied prior to modification to a \emph{donor column}, it may re-use or \emph{donate} its pivot entry to a later column $r_{i+3}, \dots, r_j$. 
-Repeating this process at most $j - i - 1$ times ensures $R'$ stays reduced in all except possibly for the $i$-th column by the end of the movement---and since the $k$-th such operation simultaneously sets $v_{k}(i) = 0$ without creating non-zeros at indices $v_{k}(j)$ for $j > k$, $V'$ retains its upper-triangularity.
+Repeating this process at most $j - i - 1$ times ensures $R'$ stays reduced in all except possibly for the $i$-th column---and since the $k$-th such operation simultaneously sets $v_{k}(i) = 0$ without creating non-zeros at indices $v_{k}(j)$ for $j > k$, $V'$ retains its upper-triangularity.
 \\
@@
 %\end{remark}
-
-As shown by example 2.3, move updates are cheaper than vineyard updates when $j > i+1$. More formally, we recall the proposition shown in~\cite{busaryev2010tracking}: 
+Regarding the complexity of move operations, which clearly depend on the sparsity of $R$ and $V$, we recall the proposition shown in~\cite{busaryev2010tracking}: 
 \begin{proposition}[Busaryev et al.~\cite{busaryev2010tracking}]\label{prop:move_st}
@@
 Note that the number of valid decomposition produced by vineyards is bounded above by $O(m^2)$ if each pair of simplices $\sigma_i, \sigma_j \in K$ switches its relative ordering at most once during the interpolation from $f$ to $f'$.
+
+There is another aspect of the move algorithm that has a distinct benefit compared to vineyards.
+As shown by example 2.3, move updates can be cheaper than vineyards in terms of column operations.
+However, it is not immediately clear that this is always case upon inspection of~\ref{alg:mr}, as the usage of a donor column seemingly implies many $O(m)$ copy operations need be performed.
+It turns out that we may handle all such operations except the first in $O(1)$ time, which we formalize below.
+\begin{proposition}\label{prop:factor2}
+	Let $K$ denote a filtration of size $\lvert K \rvert = m$ with decomposition $R = D V$ and let $T$ denote the number of column operations needed by the vineyards algorithm to perform the sequence of transpositions:
+	$$ R = R_1 \overset{s_1}{\to} R_2 \overset{s_2}{\to} \dots \overset{s_k}{\to} R_{k-1} = R' $$
+	%$$(i,i+1), (i+1, i+2), \dots, (j-1, j)$$.
+	where $s_i$ denotes the transposition $(i,i+1)$, $i < j$, and $k = \lvert i - j\rvert$.
+Moreover, let $M$ denote the number column operations to perform the same update $R \mapsto R'$ with $\mathrm{Move}(i,j)$.
+Then the inequality $M \leq T$ holds.
+\end{proposition}
+\begin{proof} \normalsize
+First, consider executing the vineyards algorithm with a given pair $(i, j)$.
+As there are at most $2$ column operations, any contiguous sequence of transpositions $(i,i+1), (i+1,i+2), \ldots (j-1,j)$ induces at most $2(\lvert i - j \rvert)$ column operations in both $R$ and $V$, giving a total of $4(\lvert i - j \rvert)$ column operations.
+
+Now consider a single MoveRight($i$,$j$) outlined in Algorithm~\ref{alg:mr}. Here, the dominant cost again are the column operations (line 5). 
+Though we need an extra $O(m)$ storage allocation for the donor columns $d_\ast$ prior to the movement, notice that assignment to and from $d_\ast$ (lines (4), and (7) in RestoreRight of MoveRight, respectively) requires just $O(1)$ time via a pointer swapping argument.
+That is, when $d_\mathrm{low}' < d_\mathrm{low}$, instead of copying $\mathrm{col}_\ast(k)$ to $d_\ast'$---which takes $O(m)$ time---we instead their column pointers in $O(1)$ before prior to column operations.
+After the movement, $d_\ast$ contains the newly modified column and $\mathrm{col}_\ast(k)$ contains the unmodified donor $d_\ast'$, so the final donor swap also requires $O(1)$ time.
+Since at most one $O(m)$ column operation is required for each index in $[i, j]$, moving a column from $i$ to $j$ where $i < j$ requires at most $2(\lvert i - j \rvert)$ column operations for both $R$ and $V$.
+The claimed inequality follows.
+%since although permutations take $O(\lvert i - j\rvert)$ time we require only $1$ such permutation per movement. 
+	%Identifying non-zero entries in the matrix occurs once, which takes $O(\lvert i - j \rvert \log(m))$ time.
+\end{proof}
 
@@
 \end{equation}
-We expect $\widetilde{S}_F$ to be cheaper to execute using move operations than $S_F$, with the tradeoff  that less diagrams are produced.
-We make this precise next:
-\begin{proposition}\label{prop:factor2}
-	Let $K$ denote a filtration of size $\lvert K \rvert = m$ with decomposition $R = D V$, $T$ the maximum number of $O(m)$ operations required to execute $S_F$ via vineyards, and $M$ be the maximum number of $O(m)$ operations required to execute $\widetilde{S}_F$ with moves.
-Then, if all pairs $(i, j) \in \tilde{S}_F$ satisfy $i < j$ and $\lvert i - j \rvert > 1$, the inequality $ M \leq {\textstyle \frac{1}{2} \, T}$ holds.
-\end{proposition}
-\begin{proof} \normalsize
-First, consider executing the vineyards algorithm on a given schedule $S_F$.
-As there are at most $2$ column operations, any contiguous sequence of transpositions $(i,i+1), (i+1,i+2), \ldots (j-1,j)$ induces at most $2(\lvert i - j \rvert)$ column operations in both $R$ and $V$, giving a total of $4(\lvert i - j \rvert)$ column operations.
-
-Now consider a single MoveRight($i$,$j$) outlined in Algorithm~\ref{alg:mr}. Here, the dominant cost again are the column operations (line 5). 
-Though we need an extra $O(m)$ storage allocation for the donor columns $d_\ast$ prior to the movement, notice that the assignment to and from $d_\ast$ (lines (4), and (7) in RestoreRight of MoveRight in Algorithm~\ref{alg:mr}) can be completed in $O(1)$ time via a pointer swapping argument.
-That is, when $d_\mathrm{low}' < d_\mathrm{low}$, instead of copying $\mathrm{col}_\ast(k)$ to $d_\ast'$---which takes $O(m)$ time---we swap their column pointers in $O(1)$ before proceeding with the necessary column operations.
-After the movement, $d_\ast$ contains the newly modified column and $\mathrm{col}_\ast(k)$ contains the unmodified donor $d_\ast'$, so we swap them once more in $O(1)$ time.
-Since we require at most one $O(m)$ column operation for each index in $[i, j]$, moving a column from $i$ to $j$ where $i < j$ requires at most $2(\lvert i - j \rvert)$ column operations for both $R$ and $V$.
-The claimed inequality follows.
-%since although permutations take $O(\lvert i - j\rvert)$ time we require only $1$ such permutation per movement. 
-	%Identifying non-zero entries in the matrix occurs once, which takes $O(\lvert i - j \rvert \log(m))$ time.
-\end{proof}
-\noindent 
-In terms of size bounds, though clearly $\lvert \widetilde{S}_F \rvert \leq \lvert S_F \rvert$ and the associated coarsened $\widetilde{S}_F$ requires just $O(m)$ time to compute, the coarsening depends entirely on the initial choice of $F$---thus the quadratic upper bound remains, as its always possible that there are no contiguous subsequences to collapse.
+Clearly $\lvert \widetilde{S}_F \rvert \leq \lvert S_F \rvert$ and the associated coarsened $\widetilde{S}_F$ requires just $O(m)$ time to compute.
+However, the coarsening depends entirely on the initial choice of $F$ and the quadratic upper bound remains---it is always possible that there are no contiguous subsequences to collapse.
 As mentioned in~\ref{sec:vineyards}, this quadratic scaling induces a number of issues in the practical implementations of the vineyards algorithm.
@@
 \subsection{Video data}
-A common application of persistence is to characterizing topological structure in image data.
+A common application of persistence is characterizing topological structure in image data.
 Since a  set of ``snapshot'' frames of a video can be equivalently thought of as discrete 1-parameter family, our framework provides a natural extension of the typical image analysis to video data.
-To demonstrate the benefit of using minimal sized scheduled and the scalability of the greedy approach proposed in section~\ref{sec:proxy_objective}, we re-use the video data from~\ref{sec:motivation} as a baseline benchmark.
-We perform two performance tests: one to  one to test the impact of shrinking the number of permutations a given reindexing operation needs and one to test the asymptotic behavior of the greedy approach and.
+To demonstrate the benefit of using minimal size schedules and the scalability of the greedy approach proposed in section~\ref{sec:proxy_objective}, we re-use the video data from~\ref{sec:motivation} as a baseline benchmark.
+We perform two performance tests: one to test the impact of shrinking the number of permutations a given reindexing operation needs and one to test the asymptotic behavior of the greedy approach.
 
@@
  The primary three strategies we test are the greedy approach from section~\ref{sec:proxy_objective}, the ``simple'' approach which uses upwards of $O(m)$ move permutations via selection sort, and a third strategy which interpolates between the two.
- In particular, if $d = m - \lvert \mathrm{LCS}(f, f') \rvert$ we use a parameter $\alpha \in [0,1]$ to choose $m - \alpha \cdot d$ random symbols to move using the same construction method outlined in section~\ref{sec:schedule_construction}.
+To perform this interpolation, we use a parameter $\alpha \in [0,1]$ to choose $m - \alpha \cdot d$ random symbols to move using the same construction method outlined in section~\ref{sec:schedule_construction}.
 When $\alpha = 0$, the strategy reduces to using selection sort to reindex $f \mapsto f'$ using a random ordering of simplices; otherwise, $\alpha = 1$ reduces to using a minimal sized schedule (with a random ordering of simplices).
-The results are summarized in the left graph on Figure~\ref{fig:movie_perf}.
-To get an idea of the variation of the random sampling, we run 10 independent iterations and shade the upper and lower bounds of the schedule costs, which is shown around the mean schedule cost depicted by the solid lines in Figure~\ref{fig:movie_perf}.
- As one can see from the figure, while using less move operations (lower $\alpha$) does progressively reduce column operations, constructing random schedules of minimal size is no more competitive than the selection sort strategy.
-This suggests that efficient schedule construction needs to account for the structure of performing several permutations in sequence, like the greedy heuristic we introduced, to yield a performance boost.
+The results are summarized in the left graph on Figure~\ref{fig:movie_perf}, wherein the mean schedule cost of the random strategies are depicted by solid lines.
+To capture the variation in performance for the random sampling approach, we run 10 independent iterations and shade the upper and lower bounds of the schedule costs.
+ As one can see from the figure, while using less move operations (lower $\alpha$) does progressively reduce column operations, constructing random schedules of minimal size is no more competitive than the simple selection sort strategy.
+This suggests that efficient schedule construction needs to account for the structure of performing several permutations in sequence, like the greedy heuristic we introduced, to yield an adequate performance boost.
 
@@
 	\includegraphics[width=0.48\textwidth]{varying_nd.pdf}
-	\caption{Performance comparison }
+	\caption{Performance comparison between various scheduling strategies.
+On the left, the cumulative column operations required to simulate the 1-parameter family is shown for varying schedule sizes ($d$) and strategies.
+On the right, both the size of the schedule and the data set $(m)$ are varied.
+}
 	\label{fig:movie_perf}
@@
 For each data set, we simulate persistence over the duration of the video, again testing five evenly spaced settings of $\alpha \in [0,1]$---the results are shown in the right plot of Figure~\ref{fig:movie_perf}.
-On the vertical axis, we plot the total number of column operations needed to simulate the video again across 25 evenly-spaced time points \emph{as a ratio} of the data set size ($m$); we also show the regression curves one obtains from considering each schedule size.
+On the vertical axis, we plot the total number of column operations needed to simulate the video again across 25 evenly-spaced time points \emph{as a ratio} of the data set size ($m$); we also show the regression curves one obtains for each setting of $\alpha$.
 As one can see from the Figure, the cost of using the greedy heuristic tends to increase sub-linearly as a function of the data set size, suggesting the move scheduling approach is indeed quite scalable.
-Moreover, using schedules with minimal size tended to be cheaper than otherwise, confirming our initial hypothesis that repairing the decomposition less can lead to substantial reductions at runtime.
+Moreover, schedules with minimal size tended to be cheaper than otherwise, confirming our initial hypothesis that repairing the decomposition less can lead to substantial reductions at runtime.
 

### 2023-02-12 move_schedules.tex
Updates from Overleaf

@@
 \abstract{
-	The standard procedure for computing the persistent homology of a filtered simplicial complex with $m$ simplices is the reduction algorithm.
+Matrix reduction is the standard procedure for computing the persistent homology of a filtered simplicial complex with $m$ simplices.
 Its output is a particular decomposition of the total boundary matrix, from which the persistence diagrams and generating cycles are derived.
@@
 By reduction to a particular longest common subsequence problem, we show the storage needed to employ this strategy is actually sublinear in expectation. 
-Exploiting this connection, we show experimentally the decrease in operations to compute diagrams across a family of filtrations is proportional to the difference between the expected quadratic number of states and the proposed sublinear coarsening.
+Exploiting this connection, we show experimentally that the decrease in operations to compute diagrams across a family of filtrations, is proportional to the difference between the expected quadratic number of states and the proposed sublinear coarsening.
 Applications to video data, dynamic metric space data, and multi-parameter persistence are also presented.
@@
 Cohen-Steiner et al. refer to a continuous 1-parameter family of persistence diagrams as a \emph{vineyard}, and they give in \cite{cohen2006vines} an efficient algorithm for their computation. 
-The vineyards approach can be interpreted as an extension of the \emph{reduction} algorithm~\cite{zomorodian2005computing}, which computes the persistence diagram of a filtered simplicial complex $K$ with $m$ simplices in $O(m^3)$ time, via a particular decomposition $R = D V$ (or $RU = D$) of the boundary matrix $D$ of $K$.
+The vineyards approach can be interpreted as an extension of the \emph{reduction} algorithm~\cite{zomorodian2005computing}, which computes the persistence diagrams of a filtered simplicial complex $K$ with $m$ simplices in $O(m^3)$ time, via a particular decomposition $R = D V$ (or $RU = D$) of the boundary matrix $D$ of $K$.
 %Since $V$ is always full rank and upper-triangular, an alternative is to compute a $D = RU$ decomposition, where $U = V^{-1}$. 
@@
 To the authors knowledge, work focused on ways of updating a  decomposition $R = DV$, for all homological dimensions, is limited: there is the vineyards algorithm~\cite{cohen2006vines} and the moves algorithm~\cite{busaryev2010tracking}, both of which are discussed extensively in section~\ref{sec:background}.
-At the time of writing, we were made aware of very recent work~\cite{luo2021accelerating} that iteratively repairs a permuted decomposition via a column swapping and reduce strategy, which they call ``warm starts.'' Though similar in spirit, their approach relies on the reduction algorithm as a subprocedure, which is quite different from the strategy we employ here.
+At the time of writing, we were made aware of very recent work~\cite{luo2021accelerating} that iteratively repairs a permuted decomposition via a column swapping and reduce strategy, which they call ``warm starts.'' Though their motivation is similar to our own, their approach relies on the reduction algorithm as a subprocedure, which is quite different from the strategy we employ here.
 
@@
 
-Each image contains a fixed number of pixels whose intensities vary with time.
-For each image, we build a fixed-size filtered simplicial complex using the \emph{Freudenthal} triangulation of the plane whose inclusion maps are obtained by lower stars of pixel values.
+Each image in the series is comprised of pixels whose intensities vary with time, upon which we build a simplicial complex using the \emph{Freudenthal} triangulation of the plane.
+For each complex, we create a filtration of simplices whose order is determined by the lower stars of pixel values.
 %Snapshots of the image data and its corresponding time-varying filtration are shown below Figure~\ref{fig:vineyards}. 
@@
 
-Note that without knowing where persistence pairing function changes, a continuous strategy must construct all $\approx 7 \times 10^{4}$ diagrams induced by the homotopy. 
+Note that without knowing where the persistence pairing function changes, a continuous strategy must construct all $\approx 7 \times 10^{4}$ diagrams induced by the homotopy. 
 In this setting, as shown in the figure, the vineyards approach is indeed far more efficient than naively applying the reduction algorithm independently at all time points.
@@
 
-The existence of a time discretization that is more efficient to compute than continually updating the decomposition indicates that the vineyards framework must incur some overhead (in terms of column operations) to maintain the underlying decomposition, even when the pairing function determining the persistence diagram is unchanged.
+The existence of a time discretization that is more efficient to compute than continually updating the decomposition, indicates that the vineyards framework must incur some overhead (in terms of column operations) to maintain the underlying decomposition, even when the pairing function determining the persistence diagram is unchanged.
 Indeed, as shown by the case where $n = 10$, applying \texttt{pHcol} independently between  relatively ``close'' filtrations is substantially more efficient than iteratively updating the decomposition.
@@
 \textbf{Our approach and contributions are as follows:} First, we leverage the \emph{moves} framework of Busaryev et al.~\cite{busaryev2010tracking} to include  coarser  operations for dynamic persistence settings.
-By a reduction to an edit distance problem, we give a   lower bound on the number of moves needed to perform an arbitrary permutation to the $R = D V$ decomposition,  along with a proof of its optimality.
+By a reduction to an edit distance problem, we give a   lower bound on the minimal number of moves needed to perform an arbitrary permutation to the $R = D V$ decomposition,  along with a proof of its optimality.
 We also give worst-case sizes of these quantities in expectation as well as efficient algorithms for constructing these operations---both of which are derived from a reduction to the Longest Increasing Subsequence (LIS) problem.
-These operations parameterize sequences of cyclic permutations $( s_1, s_2, \dots, s_d )$ of minimal size $d$, which we call \emph{schedules}. 
+These operations parameterize sequences of permutations $\mathcal{S}= ( s_1, s_2, \dots, s_d )$ of minimal size $d$, which we call \emph{schedules}. 
 However, not all minimal size schedules incur the same  cost 
@@
 
-We now summarize the main results obtained in this effort (Theorem~\ref{prop:sparsity_result}): suppose one has a sequence of permutations, i.e.
-a schedule,  $\mathcal{S} = \left( s_1, s_2, \dots, s_d \right)$ yielding a corresponding sequence of decompositions:
+We now summarize our main results  (Theorem~\ref{prop:sparsity_result}): suppose one has a  schedule  $\mathcal{S} = \left( s_1, s_2, \dots, s_d \right)$ yielding a corresponding sequence of decompositions:
   \begin{equation}\label{eq:rv_seq}
@@
  \end{equation}
- where each permutation $s_k = (i_k, j_k)$ is a particular type of cyclic permutation (see section~\ref{sec:moves_setting}).
-If $i_\ast < j_\ast$ for all $s_\ast \in \mathcal{S}$, our first result extends~\cite{busaryev2010tracking} by showing that~\eqref{eq:rv_seq} can be computed  using $O(\kappa)$ column operations, where:
+ where  $s_k = (i_k, j_k)$ for $k=1,\ldots, d$,  denotes a particular type of cyclic permutation (see section~\ref{sec:moves_setting}).
+If $i_k < j_k$ for all $s_k \in \mathcal{S}$, our first result extends~\cite{busaryev2010tracking} by showing that~\eqref{eq:rv_seq} can be computed  using $O(\kappa)$ column operations, where:
 \begin{equation}
@@
 The quantities $\lvert \mathbb{I}_{k} \rvert$ and $\lvert \mathbb{J}_{k} \rvert$ depend on the sparsity of the $V_k$ and $R_k$ matrices, respectively, and $d \sim O(m)$ is a constant that depends on how similar $f$ and $f'$ are.
-The advantage of this result is that it depends explicitly on the sparsity pattern of the decomposition itself and is thus output sensitive, which we take advantage of in Section~\ref{sec:schedule_cost}.
+The advantage of this result is that it depends explicitly on the sparsity pattern of the decomposition itself and is thus output sensitive, which we leverage in Section~\ref{sec:schedule_cost}.
 %Surprisingly, due to a refined analysis of restore subprocedure from~\cite{busaryev2010tracking} we give in section~\ref{sec:continuous_setting}, we find that move operations are actually cheaper than vineyards, see Proposition~\ref{prop:factor2}.
@@
 %The complexity of determining $\mathcal{S}$ depends on the type of permutations considered; many edit distances exhibit tight quadratic growth rates and tend to require at least quadratic time (or worse) to compute~\cite{labarre2013lower}.
-By reinterpreting a special set of cyclic permutations as edit operations on strings, we find that the \emph{smallest} such sequence that maps $f \mapsto f'$ has size  (Proposition~\ref{prop:schedule_size}) :
+By reinterpreting a special set of cyclic permutations as edit operations on strings, we find that a  \emph{smallest} such sequence mapping $f $ to $  f'$, has size  (Proposition~\ref{prop:schedule_size}) :
 \begin{equation}
@@
 We also show that the information needed to construct any $\mathcal{S}$ with optimal size can be computed in $O(m \log \log m)$ preprocessing time and $O(m)$ memory.
-Although $d$ clearly is $O(m)$ for pathological inputs, we give evidence  that $d \sim m - \sqrt{m}$ in expectation for random filtrations is not too pessimistic for truly random filtrations (Corollary~\ref{cor:expectation}), and we give empirical results suggesting $d$ is tends to be much smaller for time-varying filtrations.
+Although $d$  can be $O(m)$ for pathological inputs, we provide evidence  that $d \sim m - \sqrt{m}$ in expectation for random filtrations   (Corollary~\ref{cor:expectation}), and we give empirical results suggesting $d$  can be much smaller for time-varying filtrations.
 %Finally, we perform extensive experiments 
@@
 We also cover the reduction algorithm (designated here as \texttt{pHcol}), the vineyards algorithm, and the set of \emph{move}-related algorithms introduced in~\cite{busaryev2010tracking}, which serves as the starting point of this work.
-In Section \ref{sec:move_schedules} we introduce our move schedules  and provide efficient algorithms to make this alternative viable. 
+In Section \ref{sec:move_schedules} we introduce move schedules and provide efficient algorithms to construct them. 
 In Section \ref{sec:results} we present applications of the proposed method, including the computation of crocker stacks from flock simulations and of a 2-dimensional persistence invariant on a data set of image patches derived from natural images.
@@
 Any finite filtration may be trivially converted into an simplexwise filtration via a set of \emph{condensing}, \emph{refining}, and \emph{reindexing} maps (see~\cite{bauer2021ripser} for more details).
-Equivalently, we can also define a filtration as a pair $(K, f)$ where $K$ is a simplicial complex and $f : K \to I$ is a \emph{filter function} over a totally ordered index set $I$ satisfying $f(\tau) \leq f(\sigma)$ whenever $\tau \subseteq \sigma$, for any $\tau,\sigma \in K$.
+Equivalently, a filtration can be also defined as a pair $(K, f)$ where $K$ is a simplicial complex and $f : K \to I$ is a \emph{filter function}  satisfying $f(\tau) \leq f(\sigma)$ in $I$, whenever $\tau \subseteq \sigma$ in  $  K$.
 In this setting, $K_i = \{ \, \sigma \in K :  f(\sigma) \leq i \, \}$.
 Here, we consider two index sets: $[m] = \{ 1, \dots, m\}$ and $\mathbb{R}$.
-Without loss of generality, we exclusively consider simplexwise filtrations but for brevity-sake refer to them simply as filtrations.
+Without loss of generality, we exclusively consider simplexwise filtrations, but for brevity-sake refer to them simply as filtrations.
 
@@
 and referred to as the $p$-th \emph{persistence diagram} of $(K,f)$.
-If a homology class is born at time $i$ and dies entering time $j$, the difference $\lvert i - j \rvert$ is called the \emph{persistence} of that class.
+If a homology class is born at   $K_i$ and dies entering   $K_j$, the difference $\lvert i - j \rvert$ is called the \emph{persistence} of that class.
 In practice, filtrations often arise from triangulations parameterized by geometric scaling parameters, and the ``persistence'' of a homology class actually refers to its lifetime with respect to the scaling parameter.
@@
 
-Let $\mathbb{X}$ be a triangulable topological space and $K$ an abstract simplicial complex whose geometric realization is homeomorphic to $\mathbb{X}$. 
+Let $\mathbb{X}$ be a triangulable topological space; that is, so that there exists   an abstract simplicial complex $K$ whose geometric realization is homeomorphic to $\mathbb{X}$. 
 Let  $f: \mathbb{X} \to \mathbb{R}$ be continuous and  write $\mathbb{X}_a = f^{-1}(-\infty, a]$ to denote the sublevel sets of $\mathbb{X}$ defined by the value $a$. 
@@
 
- Given a filtration $(K, f)$ with $m$ simplices,  the main output of the reduction algorithm is a matrix decomposition $R = D V$, where the persistence diagram is encoded in $R$ and the generating cycles in the columns of $V$.
- To begin the reduction, one starts by assembling the elementary boundary chains $\partial(\sigma)$ as columns ordered according to $f$ into a $m \times m$  \emph{filtration boundary matrix} $D$.
+ Given a filtration $(K, f)$ with $m$ simplices,  the main output of the reduction algorithm is a matrix decomposition $R = D V$, where the persistence diagrams are encoded in $R$ and the generating cycles in the columns of $V$.
+ To begin the reduction, one first   assembles the elementary boundary chains $\partial(\sigma)$ as columns ordered according to $f$ into a $m \times m$  \emph{filtration boundary matrix} $D$.
 Setting $V = I$ and $R = D$, one proceeds by performing elementary left-to-right column operations on $V$ and $R$ until the following invariants are satisfied:
@@
  \begin{enumerate}[labelsep=3pt, topsep=2pt,itemsep=-0.25ex,parsep=1.0ex]\label{inv:decomposition}
- 	\item[I1.] $R = D V$ where $D$ is the boundary matrix of the filtration $K$
+ 	\item[I1.] $R = D V$ where $D$ is the boundary matrix of the filtration $(K,f)$
  	\item[I2.] $V$ is full-rank upper-triangular
@@
 Note that though the matrices $R$ and $V$ are not unique, the collection of persistent pairings are~\cite{zomorodian2005computing}.
-The persistence diagram of the corresponding filtration $K$ may be read off from lowest entries in $R$, once $R$ is in reduced form.
-
- It is at times more succinct to restrict to specific sub-matrices $D$ based on the homology dimension $p$, and so we write $D_p$ represent the $d_{p-1} \times d_p$ matrix representing the linear operator $\partial_p$ (the same notation is extended to $R$ and $V$).
-We illustrate the reduction algorithm with an example pair $(R_1, V_1)$ below. 
+The persistence diagrams of the corresponding filtration  can be determined from the lowest entries in $R$, once it has been reduced.
+
+ It is at times more succinct to restrict to specific sub-matrices of $D$ based on the homology dimension $p$, and so we write $D_p$ to represent the $d_{p-1} \times d_p$ matrix representing $\partial_p$ (the same notation is extended to $R$ and $V$).
+We illustrate the reduction algorithm with an example   below. 
 \\
@@
 \textbf{Example 2.1:} Consider a triangle with vertices $u,v,w$, edges $a = (u,w)$, $b = (v,w)$, $c = (u, v)$,
-and whose filtration order is given as $K = (u, v, w, a, b, c)$. 
+and whose filtration order is given as $ (u, v, w, a, b, c)$. 
 %The reduction algorithm begins by setting $R = D$ and $V = I$, followed by left-to-right column operations until invariants I3 is satisfied.
 %If the $i$-th column of $R$ is nonzero, then $\mathrm{low}_R(i)$  is the row index of its lowest nonzero entry.
-Using $\mathbb{Z}_2$ coefficients, the reduction proceeds to compute $H_1(K)$ as follows:
+Using $\mathbb{Z}_2$ coefficients, the reduction proceeds to compute $(R_1,V_1)$ as follows:
 $$
@@
 $$
-\noindent Since column $c$ in $R_1$ is $0$, the chain $d_1 + d_2 + d_3$ indicated by the column $c$ in $V_1$ represents a dimension $1$ cycle.
+\noindent Since column $c$ in $R_1$ is $0$, the 1-chain  indicated by the column $c$ in $V_1$ represents a dimension $1$ cycle.
 Similarly, the columns at $u, v, w$ in $R_0$ (not shown) are all zero, indicating three $0$-dimensional homology classes are born, two of which are killed by the pivot entries in columns $a$ and $b$ in $R_1$.
@@
 Inspection of the reduction algorithm from~\cite{edelsbrunner2000topological} suggests that a loose upper bound for the reduction is $O(m^3)$, where $m$ is the number of simplices of the filtration---this bound is in fact tight~\cite{morozov2005persistence}.
-Despite this high algorithmic complexity, the number of column operations has been observed to be super-linear in practice, due impart to the high sparsity and structure of $D$. 
+Despite this high algorithmic complexity, the number of column operations has been observed to be super-linear in practice, due in part to the high sparsity and structure of $D$. 
 Moreover, many variations and optimizations to Algorithm~\ref{alg:reduce} have been proposed over the past decade, see~\cite{bauer2017phat, bauer2021ripser, chen2011persistent} for an overview. 
@@
 \subsection{Vineyards}\label{sec:vineyards}
-Consider a homotopy $F(x,\tau) : \mathbb{X} \times [0,1] \to \mathbb{R}$ on a triangulable topological space $\mathbb{X}$, and denote its ``snapshot'' at a given time-point $\tau$ by $f_\tau(x) = F(x,\tau)$.
-The snapshot $f_0$ denotes the initial function at time $\tau = 0$ and $f_1$ denotes the function at the last time step. 
-As $\tau$ varies in $[0,1]$, the points in $\mathrm{dgm}_p(f_\tau)$ trace curves in $\mathbb{R}^3$, which will be continuous if $F$ is continuous and the $f_\tau$'s are tame, due to the stability of persistence.
+Consider a homotopy $F(x,t) : \mathbb{X} \times [0,1] \to \mathbb{R}$ on a triangulable topological space $\mathbb{X}$, and denote its ``snapshot'' at a given time-point $t$ by $f_t(x) = F(x,t)$.
+The snapshot $f_0$ denotes the initial function at time $t = 0$ and $f_1$ denotes the function at the last time step. 
+As $t$ varies in $[0,1]$, the points in $\mathrm{dgm}_p(f_t)$ trace curves in $\mathbb{R}^3$ which, by the stability of persistence, will be continuous if $F$ is continuous and the $f_t$'s are tame.
 Cohen-steiner et al.~\cite{cohen2007stability} referred to these curves as \emph{vines}, a collection of which forms as \emph{vineyard}---the geometric analogy meant to act as a guidepost for practitioners seeking to understand the evolution of topological structure over time.
@@
 The original purpose of the vineyards algorithm, as described in~\cite{cohen2006vines}, was to compute a continuous 1-parameter family of persistence diagrams over a time-varying filtration, detecting homological critical events along the way.
-As homological critical events only occur when the filtration order changes, detecting all such events may be reduced to computing valid decompositions at all time points interleaving  changes in the filtration order.
-For simplexwise filtrations, these changes manifest as transpositions of adjacent simplices, and thus any fixed set of rules that maintains a valid $R = D V$ decomposition under adjacent transpositions is sufficient to simulate persistence dynamically.
+As homological critical events only occur when the filtration order changes, detecting all such events may be reduced to computing valid decompositions at   time points interleaving all changes in the filtration order.
+For simplexwise filtrations, these changes manifest as transpositions of adjacent simplices, and thus any fixed set of rules that maintains a valid $R = D V$ decomposition under adjacent column transpositions, is sufficient to simulate persistence dynamically.
 
 To ensure a decomposition is valid, these rules prescribe certain column and row operations to apply to a given matrix decomposition either before, during, or after each transposition.   
-Formally, let $S_{i}^j$ represent the upper-triangular matrix such that $A S_{i}^j$ is equivalent to the elementary operation that adds column $i$ of $A$ to column $j$ of $A$ and $S_{i}^j A$ the same operation on rows $i$ and $j$.
+Formally, let $S_{i}^j$ represent the upper-triangular matrix such that $A S_{i}^j$ results in adding column $i$ of $A$ to column $j$ of $A$, and let $S_{i}^j A$ be the same operation on rows $i$ and $j$.
 Similarly, let $P$ denote the matrix so that $A P^T$ permutes the columns of $A$ and $P A$ permutes the rows.
-Since the columns of $P$ are orthonormal, $P^{-1} = P^T$, and thus writes $P A P^T$ to denote the same permutation of both the columns and rows of $A$.
+Since the columns of $P$ are orthonormal, $P^{-1} = P^T$, then $P A P^T$ performs the same permutation to both the columns and rows of $A$.
 In the special case where $P$ represents a transposition, we have $P = P^T$ and may instead simply write $P A P$.
@@
 \noindent 
-\textbf{Example 2.2} To illustrate the basic principles on which vineyards works, we re-use the running example introduced in the previous section.
+\textbf{Example 2.2} To illustrate the basic principles of vineyards, we re-use the running example introduced in the previous section.
 Below, we illustrate the case of exchanging simplices $a$ and $b$ in the filtration order, and restoring $RV$ to a valid decomposition.
@@
 Starting with a valid reduction $R = DV$ and prior to performing the exchange, observe that that the highlighted entry in $V_1$ would render $V_1$ non-upper triangular after the exchange.
-This entry is removed by a left-to-right column operation, given by applying the $S_1^{2}$ on the right to $R_1$ and $V_1$.
+This entry is removed by a left-to-right column operation, given by applying  $S_1^{2}$ on the right to $R_1$ and $V_1$.
 After this operation, the permutation may be safely applied to $V_1$.
@@
 Inspection of the individual cases of the algorithm from~\cite{cohen2006vines} shows that any single transposition requires at most two such operations on both $R$ and $V$.
-However, a host of other factors affect the runtime efficiency of the vineyards algorithm. 
+However, several factors can affect the runtime efficiency of the vineyards algorithm. 
 On the positive side, as both the $V$ and $R$ matrices tend to be quite sparse, the cost of a given column operation is proportional to the number of non-zero  entries in the two columns being modified.
@@
 Notice that these operations may yield an unreduced $R' = P R(\ast) P^T$, breaking invariant I3.
-We could reduce $R'$ with an additional $k$ operations, however its possible that $k \sim O(\lvert i - j\rvert^2)$, which is no more efficient than re-running the reduction algorithm on columns $[i,j]$ in $R$.
+We could reduce $R'$ with an additional $k$ operations, with the possibility that $k \sim O(\lvert i - j\rvert^2)$, which is no more efficient than re-running the reduction algorithm on columns $[i,j]$ in $R$.
 
@@
 \end{proposition}
-\noindent Though similar to vineyards, move operations confer additional advantages, including the two computationally attractive properties given below:
+\noindent Though similar to vineyards, move operations confer additional advantages:
 \\	
@@
 
-Property M2 implies the decomposition is not fully maintained during the execution of \textit{RestoreRight} and \textit{RestoreLeft} below, which starkly contrasts the vineyards algorithm. 
+Property M2 implies that the decomposition is not fully maintained during the execution of \textit{RestoreRight} and \textit{RestoreLeft} below, which starkly contrasts the vineyards algorithm. 
 In this way, we interpret move operations as making a tradeoff in granularity: whereas a sequence of adjacent transpositions $(i, i{+}1), (i{+}1, i{+}2), \dots, (j{-}1, j)$ generates $\lvert i - j \rvert $ valid decompositions in vineyards, an equivalent move operation $\mathrm{Move}(i,j)$ generates only one.
@@
 As shown by example 2.3, move updates can be cheaper than vineyards in terms of column operations.
-However, it is not immediately clear that this is always case upon inspection of~\ref{alg:mr}, as the usage of a donor column seemingly implies many $O(m)$ copy operations need be performed.
+However, it is not immediately clear that this is always the case upon inspection of~\ref{alg:mr}, as the usage of a donor column seemingly implies that many $O(m)$ copy operations need to be performed.
 It turns out that we may handle all such operations except the first in $O(1)$ time, which we formalize below.
 \begin{proposition}\label{prop:factor2}
-	Let $K$ denote a filtration of size $\lvert K \rvert = m$ with decomposition $R = D V$ and let $T$ denote the number of column operations needed by the vineyards algorithm to perform the sequence of transpositions:
+	Let $(K,f)$ denote a filtration of size $\lvert K \rvert = m$ with decomposition $R = D V$ and let $T$ denote the number of column operations needed by the vineyards algorithm to perform the sequence of transpositions:
 	$$ R = R_1 \overset{s_1}{\to} R_2 \overset{s_2}{\to} \dots \overset{s_k}{\to} R_{k-1} = R' $$
-	%$$(i,i+1), (i+1, i+2), \dots, (j-1, j)$$.
 	where $s_i$ denotes the transposition $(i,i+1)$, $i < j$, and $k = \lvert i - j\rvert$.
-Moreover, let $M$ denote the number column operations to perform the same update $R \mapsto R'$ with $\mathrm{Move}(i,j)$.
+Moreover, let $M$ denote the number of column operations to perform the same update $R \mapsto R'$ with $\mathrm{Move}(i,j)$.
 Then the inequality $M \leq T$ holds.
@@
 Though we need an extra $O(m)$ storage allocation for the donor columns $d_\ast$ prior to the movement, notice that assignment to and from $d_\ast$ (lines (4), and (7) in RestoreRight of MoveRight, respectively) requires just $O(1)$ time via a pointer swapping argument.
-That is, when $d_\mathrm{low}' < d_\mathrm{low}$, instead of copying $\mathrm{col}_\ast(k)$ to $d_\ast'$---which takes $O(m)$ time---we instead their column pointers in $O(1)$ before prior to column operations.
+That is, when $d_\mathrm{low}' < d_\mathrm{low}$, instead of copying $\mathrm{col}_\ast(k)$ to $d_\ast'$---which takes $O(m)$ time---we instead swap their column pointers in $O(1)$   prior to column operations.
 After the movement, $d_\ast$ contains the newly modified column and $\mathrm{col}_\ast(k)$ contains the unmodified donor $d_\ast'$, so the final donor swap also requires $O(1)$ time.
@@
 
-As a final remark, we note the combination of \emph{MoveRight} and \emph{MoveLeft} enable efficient simplex additions or deletions to the underlying complex.
+As a final remark, we note that the combination of \emph{MoveRight} and \emph{MoveLeft} enable efficient simplex additions or deletions to the underlying complex.
 In particular, given $K$ and a decomposition $R = DV$, obtaining a valid decomposition $R' = D' V'$ of $K' = K \cup \{\sigma\}$ can be achieved by appending its requisite elementary chains to $D$ and $V$, reducing them, and then executing \emph{MoveLeft}$(m+1, i)$, where $i = f'(\sigma)$.
@@
 $q = (\rho_i(1), \rho_i(2), \ldots, \rho_i(m) )$
-and compute the longest increasing subsequence $\mathrm{LIS}(q)$;  this subsequence  is used to recover a longest common subsequence (LCS) between $f_i$ and $ f_{i+1} $, which we denote later with $\mathrm{LCS}(f_i, f_{i+1})$.
+and compute a longest increasing subsequence $\mathrm{LIS}(q)$;  this subsequence  is used to recover a longest common subsequence (LCS) between $f_i$ and $ f_{i+1} $, which we denote later with $\mathrm{LCS}(f_i, f_{i+1})$.
 We pass $q$ and $\mathrm{LIS}(q)$ to our greedy scheduling algorithm, which returns as output an ordered set of move permutations  $\mathcal{S}$ of minimum size, which we call a \emph{schedule} (see Definition~\ref{def:schedule}).
@@
 
-In the following subsections, we investigate how to profitably exploit the increased flexibility Busaryev's move framework yields us, beginning with the hypothesis formed experimentally from section~\ref{sec:motivation} that vineyards exhibits extraneous overhead maintaining the decomposition.
+In the following subsections, we investigate how to  leverage the increased flexibility of the move framework, beginning with the hypothesis formed experimentally from section~\ref{sec:motivation} that vineyards exhibits extraneous overhead maintaining the decomposition.
 In theory, decreasing the number of times the decomposition is restored to a valid state ought to reduce this overhead, motivating the question: can one simultaneously minimize the number of times the decomposition is restored to a valid state while retaining an efficient update scheme?
+Next we will show how the analysis of the 
+continuous time setting leads to the idea of a schedule, 
+and then how a schedule comprised of move operations 
+can reduce the overhead from the Vineyards algorithm.
 
 \subsection{Continuous setting}\label{sec:continuous_setting}
-In the continuous time case, we are given a homotopy $F : K \times [0,1] \to \mathbb{R}$ interpolating between  a pair of filtrations $(K, f), (K,f')$ satisfying $f = F(\cdot, 0)$ and $f' = F(\cdot, 1)$, respectively.
-As the choice of homotopy $F$ determines the number of transpositions that transform $f$ into $f'$ via $F$, bounding this number requires assumptions on $F$. 
-If we assume the curves $\tau \mapsto F(\cdot, \tau)$ are in general position in $[0,1]\times \mathbb{R}$ and that each pair of curves cross  at most once, then the  functions $\tau\mapsto F(\cdot, \tau)$ belong to a class of $x$-monotone curves called \emph{pseudo-segments}.
+In the continuous time case, we are given a homotopy $F : K \times [0,1] \to \mathbb{R}$ interpolating between  a pair of filtrations $f,f' : K \rightarrow [m]$.
+That is,  the function $t\mapsto F(\sigma, t)$ is continuous,  $f(\sigma) = F(\sigma, 0)$ and $f'(\sigma) = F(\sigma, 1)$ for every  $\sigma \in K$.
+As the choice of homotopy $F$ determines the number of distinct filtrations while deforming $f$ into $f'$ via $F$, bounding this number requires assumptions on $F$. 
+Indeed, if we assume that each pair of curves $\big(t,  F(\cdot, t)\big) \subset [0,1]\times \mathbb{R}$  intersect in at most one point, at which they cross, then the  functions $t\mapsto F(\cdot, t)$ belong to a class of $x$-monotone curves called \emph{pseudo-segments}.
 %, i.e. parameterized curves whose behavior with respect to a certain restricted set of geometric predicates is invariant. 
-This family includes the straight-line homotopy $F(\sigma, \tau) = (1 - \tau) f(\sigma) + \tau f'(\sigma)$,  studied in the original vineyards paper~\cite{cohen2006vines}. 
+This family includes the straight-line homotopy $F(\sigma, t) = (1 - t) f(\sigma) + t f'(\sigma)$,  studied in the original vineyards paper~\cite{cohen2006vines}. 
+If there are no crossings in the interval $(a,b)$,
+then the filtration order given by $\sigma \mapsto F(\sigma, t)$
+is the same for every $t\in (a,b)$.
+Thus, changes to the filtration occur only after a crossing event. 
 Detecting all $k$ intersections of $m$ pseudo-segments is a well-studied problem in computational geometry that can be optimally solved in output-sensitive $O(m \log m + k)$ time by several algorithms~\cite{boissonnat2000efficient}, where $k$ is the output-sensitive term.
 
-Given two filtrations $(K,f), (K,f')$, observe their simplexwise representations induced by $f$ and $f'$ can be thought of as elements of the symmetric group $S_m$.
-As applying permutations to $K$ implies additional column operations on the underlying matrix decomposition, we think of any set of such permutations acting on the initial decomposition $R=DV$ of $(K,f)$ as a \emph{schedule} with respect to the pair $f,f'$.
-We formalize this with a definition.
+Let $F : K \times [0,1] \to \mathbb{R}$ be a homotopy of pseudo-segments between the  filtrations 
+$f,f' : K \rightarrow [m]$.
+The continuity and genericity assumptions on $F$ imply that for $ \sigma,\mu \in K$ distinct, 
+ the curves   $t \mapsto F(\sigma ,t)$ and 
+$t \mapsto F(\mu,t)$ intersect if and only if
+ $f(\sigma) > f(\mu)$ and $f'(\sigma)  < f'(\mu)$, or $f(\sigma) < f(\mu)$ and $f'(\sigma)  > f'(\mu)$. 
+ In other words, the number of crossings in $F$ is 
+ exactly the  \emph{Kendall}-$\tau$ distance~\cite{diaconis1977spearman} between $f$ and $f'$: 
+\begin{equation}\label{eq:kendall_dist}
+	K_\tau(f, f') = 
+ \frac{1}{2} \left\lvert \, \left\{(\sigma, \mu) \mid \mathsf{sign}\big(f(\sigma) - f(\mu) \big) \neq \mathsf{sign}\big(f'(\sigma) - f'(\mu)\big) \right\} \, \right\rvert
+\end{equation}
+After slightly perturbing $F$ if necessary,  we can further assume that its crossings occur at  $k = K_\tau(f,f')$ distinct time points $0< t_1 < \cdots < t_k < 1$. 
+Let $t_0 = 0$, $t_{k+1}= 1$ and fix 
+$a_i \in (t_{i}, t_{i + 1})$ for $i=0, \ldots, k$. 
+Then, the order in $K$ induced by $\sigma \mapsto F(\sigma, a_i)$
+defines a filtration $f_i : K \rightarrow [m]$
+so  that $f_0 = f$, $f_k = f'$ and 
+$\mathcal{F} = (f_0, f_1,\ldots, f_k)$ is the ordered sequence of all distinct filtrations in the interpolation from $f$ to $f'$ via $F$.
+
+The continuity of the curves $t\mapsto F(\cdot, t)$ and the fact that $t_i$ is the sole crossing time  in the interval $(t_{i-1}, t_{i+1})$, imply that the
+permutation $\rho_i $   
+transforming $f_{i-1}$ into $f_i$, i.e. so that  $f_i = \rho_i \circ f_{i-1}$, must be (in cycle notation) of the form 
+$\rho_i = (\ell_i \;\; \ell_i +1 )$ for $1 \leq \ell_i < m$. In other words, $\rho_i$ is an adjacent transposition for each $i =1,\ldots, k$.
+The ordered sequence of adjacent transpositions  
+$S_F = (\rho_1, \rho_2, \ldots, \rho_k)$ can thus be thought of as a \emph{schedule} of permutations to be applied to the initial decomposition $R = DV$ of $(K,f)$, in order to   interpolate from $f$ to $f'$ via $F$.
+Going beyond transpositions, the idea of a schedule can be generalized as follows:
 \begin{definition}[Schedule]\label{def:schedule}
@@
 \end{definition}
-\noindent Let $F : K \times [0,1] \to \mathbb{R}$ denote a homotopy of $x$-monotone curves.
-Then  $F$ decomposes into an ordered set of adjacent transpositions $S_F$ given by pairs of simplices in $K$ that change in their relative ordering between $K$ and $K'$:
-\begin{align*}\label{eq:sf_schedule}
-	S_F &= \{ (i,i+1) \, \mid \, f(\sigma_i) < f(\sigma_{i+1}), \;  f'(\sigma_i) > f'(\sigma_{i+1}), \; i \in [m] \} \\
-	&= \{ s_1, s_2, \dots, s_k \} 
-\end{align*}
-Moreover, let $p = [m]$ and let $q = \mathrm{Im}\,\rho^\ast$, where $\rho^\ast : [m] \to [m]$ is the one-to-one correspondence induced by the reindexing bijection $\rho: K \to K'$.
-Then $S_F$ is given by the inversions between $p$ and $q$:
-\[S_F = 
-\mathrm{Inv}(p, q) = \{ \, (i,i+1) \, \mid \, p(i) < p(i+1), \; q(i) > q(i+1), \; i \in [m]\, \}
-\]
-The cardinality of $  S_F $ is exactly the \emph{Kendall}-$\tau$ distance~\cite{diaconis1977spearman} between $p$ and $q$: 
-\begin{equation}\label{eq:kendall_dist}
-	K_\tau(p, q) = \lvert \, \mathrm{Inv}(p, q) \, \rvert
-\end{equation}
-This implies the $\lvert S_F \rvert \sim O(m^2)$ in the worst case, achieved when $f = - f'$, which the grayscale image data example from section~\ref{sec:motivation} exhibits. 
-As the 9x9 patch contains $(81, 208, 128)$ simplices of dimensions $(0, 1, 2)$, the largest number of transpositions a schedule simplices in total.
-As the homotopy is simulated, observe $\approx 70,\!000$ transpositions are generated, approaching the upper bound of , due to the fact that $f'$  is nearly the reverse of $f$.
-
-If our goal is to decrease the size of $\lvert S_F \rvert$, one option is to \emph{coarsen} $S_F$ to a new schedule $\widetilde{S}_F$ by collapsing contiguous sequences of adjacent transpositions to moves, via the map:
+
+The size of the schedule $S_F$ defined from the homotopy $F$ above 
+is exactly $K_\tau(f,f')$. In the worst case this
+is $\sim O(m^2)$, achieved when $f' = - f$, which the grayscale image data example from section~\ref{sec:motivation} exhibits. 
+Indeed,  the Freudenthal triangulation of the $9\times 9$ grid contains $(81, 208, 128)$ simplices of dimensions $(0, 1, 2)$, respectively. 
+Therefore, $m = 417$ and $\lvert S_F \rvert \leq \frac{1}{2}m(m-1) = 86,736$.
+As the homotopy given by the video is simulated,  $\approx 70,\!000$ transpositions are generated, approaching the worst case upper bound due to the fact that $f'$  is nearly the reverse of $f$.
+
+If our goal is to decrease  $\lvert S_F \rvert$, one option is to \emph{coarsen} $S_F$ to a new schedule $\widetilde{S}_F$ by collapsing contiguous sequences of adjacent transpositions to moves, via the map:
 \begin{equation}\label{eq:tr_to_mv}
@@
 
-Contrasting the continuous-time setting, if we discard the use of a homotopy interpolation and allow move operations in any order, we obtain a trivial upper bound of $O(m)$ on the schedule size:  simply move each simplex in $K$ into its position in the filtration given by $K'$ in the order given by the latter---which we call \emph{naive strategy}.
-However, it's not immediately clear whether this bound is tight.
+Contrasting the continuous-time setting, if we discard the use of a homotopy interpolation and allow move operations in any order, we obtain a trivial upper bound of $O(m)$ on the schedule size:  simply move each simplex in $K$ from  its position in the filtration given $f$ to the position given by $f'$---which we call the \emph{naive strategy}.
+However, it is not immediately clear whether this bound is tight.
 In this section, we investigate the tightness of this bound by revisiting the problem from a combinatorial perspective.
@@
 Common variations of this problem include finding such a sequence of minimal length ($d$) and bounding the length $d$ as a function of $m$.
-In the latter case, a lower bound on $d$ is referred to as the \emph{distance} between $p$ and $q$ with respect to $\Sigma$.
+In the latter case, the largest  lower bound on $d$ is referred to as the \emph{distance} between $p$ and $q$ with respect to $\Sigma$.
 A sequence $S = (s_1, s_2, \dots, s_d)$ of operations $s \in \Sigma \subseteq S_m$ mapping $p \mapsto q$ is sometimes called a \emph{sorting of $p$}. 
@@
 %Like the \emph{LCS} distance, 
-In general, one can compute the LCS itself in $O(mn)$ easily with dynamic programming.
+In general, one can compute the LCS itself in $O(mn)$  with dynamic programming.
 One might hope computing the size $\mathrm{d}_{\mathrm{lcs}}(p,q)$ alone exhibits a lower complexity, however there is substantial evidence that for general string inputs the complexity cannot be much lower than quadratic~\cite{abboud2015tight}.
@@
 We seek a \emph{proxy objective} that correlates with~\eqref{eq:schedule_cost} and does not explicitly depend on the entries in the decomposition, as minimizing these terms directly is difficult due to  the changing sparsity of the intermediate matrices $R_k, V_k$,
-Given a pair of filtrations $(K, f), (K,f')$, consider a schedule $\mathcal{S} \in \Phi(f,f',d)$ of cyclic permutations $((i_1, j_1), (i_2, j_2), \dots, (i_d, j_d))$ minimizing:
+Given a pair of filtrations $(K, f), (K,f')$, consider a schedule $\mathcal{S} \in \Phi(f,f',d)$ of cyclic permutations $(i_1, j_1), (i_2, j_2), \dots, (i_d, j_d)$ minimizing:
 \begin{equation}\label{eq:upper_bound_move_cost}

### 2023-02-15 move_schedules.tex
removed lineno since journal does it anyways

@@
    	can be determined in $O(m \log \log m)$ time and $O(m)$ space, where $R' = D_d V_d$ denotes a valid decomposition of $(K, f')$. 
-   	Moreover, if $i_k < j_k$ for all $k \in [d]$, then computing~\eqref{eq:rv_seq_theorem} requires $O(d\kappa)$ column operations, where:	  
+   	Moreover, if $i_k < j_k$ for all $k \in [d]$, then computing~\eqref{eq:rv_seq_theorem} requires $O(\kappa)$ column operations, where:	  
    	$$  \quad \kappa = \sum\limits_{k = 1}^d \, (\lvert \mathbb{I}_{k} \rvert + \lvert \mathbb{J}_{k} \rvert) $$
@@
 	where the quantities $\lvert \mathbb{I}_{k} \rvert, \lvert \mathbb{J}_{k} \rvert$ of the intermediate $R_k, V_k$ are given in Proposition~\ref{prop:move_st}.
-Note that since $\kappa$ depends on the sparsity of the intermediate entries $V_1, V_2, \dots, V_d$ and $R_1, R_2, \dots, R_d$, the bound $O(d \kappa)$ is output-sensitive.
+Note that since $\kappa$ depends on the sparsity of the intermediate entries $V_1, V_2, \dots, V_d$ and $R_1, R_2, \dots, R_d$, the bound $O(\kappa)$ is output-sensitive.
 %	and where $\lvert \mathbb{I}_{k} \rvert$ denotes the number of non-zero entries $V_k[i_k, i_k:j_k]$, $\lvert \mathbb{J}_{k}\rvert$ the number of columns satisfying $\mathrm{low}_{R_k} \in [i_k:j_k]$ and $\mathrm{row}_{R_k}(i_k) \neq 0$, and $d = m - \lvert \mathrm{LCS}(K, K') \rvert$.
@@
 It is not immediately clear, however, how to obtain a sorting $p \mapsto q$ from a given $\mathcal{L} = \mathrm{LIS}(p')$ in an efficient way.
-We outline below a simple procedure which constructs such a sorting $\mathcal{S} = (s_1, \dots, s_d)$ in $O(dm\log m)$ time and $O(m)$ space.
-%Alternatively, given $\mathcal{L}$, the algorithm may be executed online (up to $d$ times) in $O(m \log m)$ time and $O(m)$ space.
+We outline below a simple procedure which constructs such a sorting $\mathcal{S} = (s_1, \dots, s_d)$ in $O(dm\log m)$ time and $O(m)$ space, or $O(m \log m)$ time and $O(m)$ space per update in the online setting.
 
@@
 \end{equation}
-Moreover, in contrast to $K_\tau$, the Spearman distance can be computed in $O(m)$ time, has a $\ell_1$ interpretation between ranking, and is often used in the well-known \emph{rank aggregation problem}: whereas obtaining a Kemeni optimal aggregation with respect to $K_\tau$ is NP-hard, the optimal such aggregation with respect to $F$ is obtainable in polynomial time~\cite{dinu2006efficient}.
+Moreover, in contrast to $K_\tau$, the Spearman distance can be computed in $O(m)$ time and is often used in the well-known \emph{rank aggregation problem}.
+Indeed, whereas obtaining a Kemeni optimal aggregation with respect to $K_\tau$ is NP-hard, the optimal such aggregation with respect to $F$ is obtainable in polynomial time~\cite{dinu2006efficient}.
 % Indeed, it is used...
@@
 In contrast, a 2-approximation may be obtained by replacing the Kendall distance with the Spearman distance, as in this case the problem reduces to finding a perfect matching in a bipartite graph, which can be accomplished in polynomial time.
-To adapt $F(\cdot, \cdot)$ to sortings, we decompose the footrule distance additively via the bound:
+To adapt $F(\cdot, \cdot)$ to sortings, we decompose the Spearman distance additively via the bound:
 \begin{equation}\label{eq:sorting_additive}
@@
 Suppose we begin with an array $\mathcal{A}$ of size $m$ which provides $O(1)$ access and modification, initialized with the (signed) \emph{displacement} of every element in $p$ to its corresponding position in $q$.
-Since the Spearman distance is simply the sum of the absolute value of these displacements, at any point during during the execution of Algorithm~\ref{alg:sorting} we may obtain $F(\bar{p}, q)$ simply by having access to the sum of every entry in $\mathcal{A}$.
-There are two degrees of freedom in   Algorithm~\ref{alg:sorting}: the first is in the choice of $\sigma \in \mathcal{D}$, and the second is in choosing $j$.
-If we fix a heuristic for the latter, then we have a set of possible valid moves $S_\mathcal{D}$ induced by each $\sigma \in \mathcal{D}$ to choose from.
-Observe that each permutation $s_\sigma \in S_\mathcal{D}$ changes the displacement of every symbol in at most three different ways:
+Since the Spearman distance is simply the sum of the absolute value of these displacements, at any point during during the execution of Algorithm~\ref{alg:sorting} we may obtain $F(\cdot, q)$ simply by having access to the sum of every entry in $\mathcal{A}$.
+At every step, there are only two degrees of freedom in Algorithm~\ref{alg:sorting}: the choice of $\sigma \in \mathcal{D}$ to move, and choice of the target index $j$ to move $\sigma$ to.
+If we fix a heuristic for the latter, then each $\sigma \in \mathcal{D}$ induces a set of possible valid moves satisfying~\eqref{eq:valid_move}, which we denote as $S_\mathcal{D}(\sigma)$.
+Now, observe that each permutation $s \in S_\mathcal{D}(\sigma)$ changes the displacement of every symbol in at most three different ways:
 \[
-\mathcal{A}(s_\sigma \circ p) = 
+\mathcal{A}(s \circ p) = 
 \begin{cases} 
@@
 \begin{equation}\label{eq:greedy_step}
-	s_{\text{greedy}} = \argmin_{\text{ valid } s_{\sigma \in S_{\mathcal{D}}}} F(s\circ p, \iota)
+	s_{\text{greedy}} = \argmin_{s \in S_{\mathcal{D}}(\sigma)} F(s\circ p, \iota)
 \end{equation}  
 in $O(d\log m)$ time. 
-The former problem reduces to the problem of efficiently calculating \emph{prefix sums}, which is easily solved. 
-It is not immediately clear how to achieve the latter modification complexity in $O(\log m)$ time, since $\lvert i - j \rvert \leq m$ is potentially larger than $O(\log m)$---to handle this, note that one can apply constant-factor updates to multiple values in a \emph{implicit treap} data structure in $O(\log m)$ time via \emph{range updates}.
+The former problem of accessing aggregate information reduces to the problem of efficiently calculating \emph{prefix sums}, which is easily solved. 
+It is not immediately clear how to achieve the latter modification complexity in $O(\log m)$ time, since $\lvert i - j \rvert \leq m$ is potentially larger than $O(\log m)$---however, because we are working with displacements, note the additive modification to each entry in $\lvert i - j \rvert$ is a constant.
+It is known that one can apply constant-factor updates to multiple values in a \emph{implicit treap} data structure in $O(\log m)$ time via \emph{range updates}~\cite{blelloch1998fast}.
  % Can actually only achieve O(d^2 log m) it seems
-Since single element modifications, removals, and insertions can all be achieved in $O(\log m)$ expected time with such a data structure, we conclude that the greedy approach solving equation~\eqref{eq:greedy_step} requires just $O(d^2 \log m)$ time.
+Since single element modifications, removals, and insertions can all be achieved in $O(\log m)$ expected time with such a data structure, we conclude that equation~\eqref{eq:greedy_step} may solved in just $O(d^2 \log m)$ time.
 %Since single element prefix sum modifications can also be achieved $O(\log m)$ time with these data structures, the time to compute $F(\hat{s}_i, \hat{s}_{i+1})$ is bounded above by $O(\log m)$, 

### 2023-05-25 move_schedules.tex
Updated declarations in prep for APCT

@@
 
+%\usepackage[nodisplayskipstretch]{setspace}
+%\setstretch{1.5}
+
 \usepackage{caption} 
@@
 \setenumerate[4]{label*=.\arabic*}
+
+\usepackage{setspace}
+\setstretch{1}
+%\setlength{\abovedisplayskip}{0pt}
+%\setlength{\belowdisplayskip}{0pt}
+%\setlength{\abovedisplayshortskip}{0pt}
+%\setlength{\belowdisplayshortskip}{0pt}
 
@@
 The vineyards algorithm is efficient asymptotically: if there are $d$ time-points where the filtration order changes, then  vineyards   takes $O(m^3 + md)$ time; one initial $O(m^3)$-time reduction at time $t_0$ followed by one $O(m)$ operation to update the decomposition at the remaining time points $(t_1, t_2, \dots, t_d)$.
-When $d >> m$, this $O(md)$ approach is far more efficient than the $O(dm^3)$  ``naive'' strategy of computing the diagrams at every time point independently.
+When $d >> m$, the initial reduction cost is amortized by the cost of maintaining the decomposition, implying each diagram produced by vineyards takes just linear time per time point.
+% approach is far more efficient than the $O(dm^3)$  ``naive'' strategy of computing the diagrams at every time point independently.
 
@@
 The work presented here seeks to further understand and remedy this discrepancy: building on the work presented in~\cite{busaryev2010tracking}, we introduce a coarser approach to the vineyards algorithm.
-While vineyards is  designed  for handling a \emph{continuous} 1-parameter family of diagrams, it is not necessarily efficient when the parameter is coarsely discretized.
+Though the vineyards algorithm is efficient at constructing a \emph{continuous} 1-parameter family of diagrams, it is not necessarily efficient when the parameter is coarsely discretized.
 Our methodology is based on the observation that practitioners often don't need (or want!) all of the persistence diagrams generated by a continuous 1-parameter of filtrations; usually just $n << d$ of them  suffice.
@@
 This trade off, paired with a fast greedy heuristic explained in section~\ref{sec:proxy_objective}, yields an algorithm that can update a $R = DV$ decomposition more efficiently than vineyards in coarse time-varying contexts, making dynamic persistence more computationally tractable for a wider class of use-cases.
-The source code containing both the algorithm we propose and the experiments performed in Section~\ref{sec:results} is open source and available online.\footnote{ Code is available here: \url{https://github.com/peekxc/dart}}
+The source code containing both the algorithm we propose and the experiments performed in Section~\ref{sec:results} is open source and available online.\footnote{ For all accompanying software and materials, see: \url{https://github.com/peekxc/move_schedules}}
 
@@
 Along a similar vein, Edelsbrunner et al.~\cite{edelsbrunner2000topological} give a fast incremental algorithm for computing persistent Betti numbers up to dimension $2$, again by utilizing symmetry, duality, and ``time-reversal''~\cite{delfinado1995incremental}.
-Chen et al.~\cite{chen2013output} give an output-sensitive method for computing persistent homology,  utilizing the property that certain submatrices of  $D$ have the same rank as  $R$, which they exploit through fast sub-cubic rank algorithms specialized for sparse-matrices.
+Chen \& Kerber~\cite{chen2013output} give an output-sensitive method for computing persistent homology,  utilizing the property that certain submatrices of  $D$ have the same rank as  $R$, which they exploit through fast sub-cubic rank algorithms specialized for sparse-matrices.
 
@@
 \end{equation}
+where $\mathrm{LCS}(f,f')$ refers to the size of the longest common subsequence between the simplexwise filtrations $(K, f)$ and $(K,f')$ (see section~\ref{sec:moves_setting} for more details).
 We also show that the information needed to construct any $\mathcal{S}$ with optimal size can be computed in $O(m \log \log m)$ preprocessing time and $O(m)$ memory.
@@
 As $t$ varies in $[0,1]$, the points in $\mathrm{dgm}_p(f_t)$ trace curves in $\mathbb{R}^3$ which, by the stability of persistence, will be continuous if $F$ is continuous and the $f_t$'s are tame.
-Cohen-steiner et al.~\cite{cohen2007stability} referred to these curves as \emph{vines}, a collection of which forms as \emph{vineyard}---the geometric analogy meant to act as a guidepost for practitioners seeking to understand the evolution of topological structure over time.
+Cohen-Steiner et al.~\cite{cohen2007stability} referred to these curves as \emph{vines}, a collection of which forms as \emph{vineyard}---the geometric analogy meant to act as a guidepost for practitioners seeking to understand the evolution of topological structure over time.
 
@@
 \begin{equation}\label{eq:coefficients}
-	r_j = D v_j = \alpha_i^{(i)} \cdot \partial(\sigma_{i}) + \alpha_{i}^{(i+1)} \cdot \partial(\sigma_{i+1}) + \dots + \alpha_{i}^{(j)} \cdot \partial(\sigma_{j}) = 0
+	r_j = D v_j = \alpha_j^{(1)} \cdot \partial(\sigma_{1}) + \alpha_{j}^{(2)} \cdot \partial(\sigma_{2}) + \dots + \alpha_{j}^{(j)} \cdot \partial(\sigma_{j}) = 0
 \end{equation}
 Now, consider the permutation $P$ that moves a simplex $\sigma_i$ in $K$ to position $j$, shifting all intermediate simplices $\sigma_{i+1}, \dots, \sigma_{j}$ down by one ($i < j$). 
-To perform this shift, the non-zero coefficients from~\eqref{eq:coefficients} must be set to zero, otherwise $\partial(\sigma_i)$ contributes to boundary chains $r_{i+1}, \dots, r_j$ earlier in the filtration.
-As these coefficients are recorded by $v_j$, setting $\alpha_i^{(k)} = 0$  amounts to setting the row entries $v_k(i) = 0$ for all $k \in [i+1,j]$---these cancellations manifest as column operations $(\ast)$ used to ensure $P V(\ast)$ is upper-triangular, thus maintaining invariant I2.
+To perform this shift, all of the non-zero coefficients $\alpha_k^{(i)}$ contributing to $Dv_k$ for $k \in [i+1, j]$ must be set to zero, as otherwise $\partial(\sigma_i)$ may participate in boundary chains earlier than its own birth in the permuted filtration.
+Setting these coefficients to zero amounts to setting row entries $v_k(i) = 0$ for all $k \in [i+1,j]$---these cancellations manifest as column operations $(\ast)$ used to ensure $P V(\ast)$ is upper-triangular, thus maintaining invariant I2.
 Notice that these operations may yield an unreduced $R' = P R(\ast) P^T$, breaking invariant I3.
-We could reduce $R'$ with an additional $k$ operations, with the possibility that $k \sim O(\lvert i - j\rvert^2)$, which is no more efficient than re-running the reduction algorithm on columns $[i,j]$ in $R$.
+We could reduce $R'$ with an additional $k$ operations, with the possibility that $k \sim O(\lvert i - j\rvert^2)$, but then this is no more efficient than re-running the reduction algorithm on columns $[i,j]$ in $R$.
 
@@
 In theory, decreasing the number of times the decomposition is restored to a valid state ought to reduce this overhead, motivating the question: can one simultaneously minimize the number of times the decomposition is restored to a valid state while retaining an efficient update scheme?
-Next we will show how the analysis of the 
-continuous time setting leads to the idea of a schedule, 
-and then how a schedule comprised of move operations 
-can reduce the overhead from the Vineyards algorithm.
+%Next we will show how the analysis of the 
+%continuous time setting leads to the idea of a schedule, 
+%and then how a schedule comprised of move operations 
+%can reduce the overhead from the Vineyards algorithm.
 
 \subsection{Continuous setting}\label{sec:continuous_setting}
-In the continuous time case, we are given a homotopy $F : K \times [0,1] \to \mathbb{R}$ interpolating between  a pair of filtrations $f,f' : K \rightarrow [m]$.
-That is,  the function $t\mapsto F(\sigma, t)$ is continuous,  $f(\sigma) = F(\sigma, 0)$ and $f'(\sigma) = F(\sigma, 1)$ for every  $\sigma \in K$.
-As the choice of homotopy $F$ determines the number of distinct filtrations while deforming $f$ into $f'$ via $F$, bounding this number requires assumptions on $F$. 
-Indeed, if we assume that each pair of curves $\big(t,  F(\cdot, t)\big) \subset [0,1]\times \mathbb{R}$  intersect in at most one point, at which they cross, then the  functions $t\mapsto F(\cdot, t)$ belong to a class of $x$-monotone curves called \emph{pseudo-segments}.
-%, i.e. parameterized curves whose behavior with respect to a certain restricted set of geometric predicates is invariant. 
-This family includes the straight-line homotopy $F(\sigma, t) = (1 - t) f(\sigma) + t f'(\sigma)$,  studied in the original vineyards paper~\cite{cohen2006vines}. 
-If there are no crossings in the interval $(a,b)$,
-then the filtration order given by $\sigma \mapsto F(\sigma, t)$
-is the same for every $t\in (a,b)$.
-Thus, changes to the filtration occur only after a crossing event. 
-Detecting all $k$ intersections of $m$ pseudo-segments is a well-studied problem in computational geometry that can be optimally solved in output-sensitive $O(m \log m + k)$ time by several algorithms~\cite{boissonnat2000efficient}, where $k$ is the output-sensitive term.
-
-Let $F : K \times [0,1] \to \mathbb{R}$ be a homotopy of pseudo-segments between the  filtrations 
-$f,f' : K \rightarrow [m]$.
-The continuity and genericity assumptions on $F$ imply that for $ \sigma,\mu \in K$ distinct, 
- the curves   $t \mapsto F(\sigma ,t)$ and 
-$t \mapsto F(\mu,t)$ intersect if and only if
- $f(\sigma) > f(\mu)$ and $f'(\sigma)  < f'(\mu)$, or $f(\sigma) < f(\mu)$ and $f'(\sigma)  > f'(\mu)$. 
+In the original vineyards setting, a given homotopy  $F : K \times [0,1] \to \mathbb{R}$ continuously interpolating between $(K, f)$ and $(K, f')$ is discretized into a set of critical events that alter the filtration order.
+As $F$ determines the number of distinct filtrations encountered during the deformation from $f$ to $f'$, a natural question is whether such an interpolation can be modified so as to minimize the number of times the decomposition is restored to a valid state.
+Towards explaining the phenomenon exhibited in Figure~\ref{fig:vineyards}, in what follows we analyze a particular class of interpolation schemes in order to establish an upper bound on this quantity.
+
+%a homotopy $F : K \times [0,1] \to \mathbb{R}$  whose function $t\mapsto F(\sigma, t)$ is continuous and satisfies $f(\sigma) = F(\sigma, 0)$ and $f'(\sigma) = F(\sigma, 1)$ for every  $\sigma \in K$.
+%Indeed, if we assume that each pair of curves $\big(t,  F(\cdot, t)\big) \subset [0,1]\times \mathbb{R}$  intersect in at most one point, at which they cross, then the  functions $t\mapsto F(\cdot, t)$ belong to a class of $x$-monotone curves called \emph{pseudo-segments}.
+%%, i.e. parameterized curves whose behavior with respect to a certain restricted set of geometric predicates is invariant. 
+%This family includes the straight-line homotopy $F(\sigma, t) = (1 - t) f(\sigma) + t f'(\sigma)$,  studied in the original vineyards paper~\cite{cohen2006vines}. 
+%If there are no crossings in the interval $(a,b)$,
+%then the filtration order given by $\sigma \mapsto F(\sigma, t)$
+%is the same for every $t\in (a,b)$.
+%Thus, changes to the filtration occur only after a crossing event. 
+%Detecting all $k$ intersections of $m$ pseudo-segments is a well-studied problem in computational geometry that can be optimally solved in output-sensitive $O(m \log m + k)$ time by several algorithms~\cite{boissonnat2000efficient}, where $k$ is the output-sensitive term.
+
+%Necessarily, bounding this quantity requires assumptions on $F$. 
+Let $F : K \times [0,1] \to \mathbb{R}$ be a homotopy of $x$-monotone curves\footnote{This term has been used in reference to parameterized curves whose behavior with respect to a certain restricted set of geometric predicates is invariant, see~\cite{boissonnat2000efficient}.
+} between the  filtrations
+$f,f' : K \rightarrow [m]$ whose function $t\mapsto F(\sigma, t)$ is continuous and satisfies $f(\sigma) = F(\sigma, 0)$ and $f'(\sigma) = F(\sigma, 1)$ for every  $\sigma \in K$.
+Note this family includes the straight-line homotopy $F(\sigma, t) = (1 - t) f(\sigma) + t f'(\sigma)$,  studied in the original vineyards paper~\cite{cohen2006vines}. 
+If we assume that each pair of curves $\big(t,  F(\cdot, t)\big) \subset [0,1]\times \mathbb{R}$  intersect in at most one point---at which they cross---the continuity and genericity assumptions on $F$ imply that for $ \sigma,\mu \in K$ distinct,  the curves   $t \mapsto F(\sigma ,t)$ and $t \mapsto F(\mu,t)$ intersect if and only if $f(\sigma) > f(\mu)$ and $f'(\sigma)  < f'(\mu)$, or $f(\sigma) < f(\mu)$ and $f'(\sigma)  > f'(\mu)$.
  In other words, the number of crossings in $F$ is 
@@
 \end{definition}
-
-The size of the schedule $S_F$ defined from the homotopy $F$ above 
-is exactly $K_\tau(f,f')$. In the worst case this
-is $\sim O(m^2)$, achieved when $f' = - f$, which the grayscale image data example from section~\ref{sec:motivation} exhibits. 
+\noindent Observe the size of the schedule $S_F$ defined from the homotopy $F$ above is exactly $K_\tau(f,f')$.
+On the positive side, the reduction of schedule planning to crossing detection implies the former can be  that can be solved optimally in output-sensitive $O(m \log m + k)$ time by several algorithms~\cite{boissonnat2000efficient}, where $k$ is the output-sensitive term and $m$ is the number of simplices in the filtration(s).
+On  the negative side, $k = K_\tau(f,f')$ scales in size to $\sim O(m^2)$ in the worst case, achieved when $f' = - f$.
+As mentioned in~\ref{sec:vineyards}, this quadratic scaling induces a number of issues in the practical implementations of the vineyards algorithm
+
+%detecting all $k$ intersections of $m$ such curves is a well-studied problem in computational geometry
+%If there are no crossings in the interval $(a,b)$,
+%then the filtration order given by $\sigma \mapsto F(\sigma, t)$
+%is the same for every $t\in (a,b)$.
+%Thus, changes to the filtration occur only after a crossing event. 
+\begin{remark}
+	The grayscale image data example from section~\ref{sec:motivation} exhibits this quadratic scaling.
 Indeed, the Freudenthal triangulation of the $9\times 9$ grid contains $(81, 208, 128)$ simplices of dimensions $(0, 1, 2)$, respectively. 
@@
 As the homotopy given by the video is simulated,  $\approx 70,\!000$ transpositions are generated, approaching the worst case upper bound due to the fact that $f'$  is nearly the reverse of $f$.  
-
-If our goal is to decrease  $\lvert S_F \rvert$, one option is to \emph{coarsen} $S_F$ to a new schedule $\widetilde{S}_F$ by collapsing contiguous sequences of adjacent transpositions to moves, via the map:
+\end{remark}
+
+If our goal is to decrease  $\lvert S_F \rvert$, one option is to \emph{coarsen} $S_F$ to a new schedule $\widetilde{S}_F$ by e.g.
+collapsing contiguous sequences of adjacent transpositions to moves, via the map:
 \begin{equation}\label{eq:tr_to_mv}
@@
 However, the coarsening depends entirely on the initial choice of $F$ and the quadratic upper bound remains---it is always possible that there are no contiguous subsequences to collapse.
-As mentioned in~\ref{sec:vineyards}, this quadratic scaling induces a number of issues in the practical implementations of the vineyards algorithm.
+Suggesting one must either abandon the continuous  setting or make stronger assumptions on $F$ to have any hope of keeping $\lvert S_F \rvert \sim O(m)$ in size.
 % Indeed, although many transpositions require no column reductions in practice, detecting the existence of non-zero entries at specific positions in the sparse matrix takes $O(\log(m))$ time per transposition.
@@
 
-Contrasting the continuous-time setting, if we discard the use of a homotopy interpolation and allow move operations in any order, we obtain a trivial upper bound of $O(m)$ on the schedule size:  simply move each simplex in $K$ from  its position in the filtration given $f$ to the position given by $f'$---which we call the \emph{naive strategy}.
-However, it is not immediately clear whether this bound is tight.
-In this section, we investigate the tightness of this bound by revisiting the problem from a combinatorial perspective.
+Contrasting the continuous-time setting, if we discard the use of a homotopy interpolation and allow move operations in any order, we obtain a trivial upper bound of $O(m)$ on the schedule size:  simply move each simplex in $K$ from  its position in the filtration given by $f$ to the position given by $f'$---which we call the \emph{naive strategy}.
+However, in losing the interpolation interpretation it is no longer clear the $O(m)$ bound is tight, and the ``intermediate'' filtrations need no longer respect the face poset of the underlying complex $K$.
+In this section, we investigate these issues from a combinatorial perspective.
 % or whether this strategy of generating schedules---which we call \emph{naive strategy}---is performant.  
@@
 %and the distance between pairs of uniformly random permutations $p, q \sim S_m$ are asymptotically normal. 
-Indeed, Diaconis et al.~\cite{diaconis1977spearman} showed the Spearman distance approximates $K_\tau$ within a factor of two: 
+Indeed, Diaconis~\cite{diaconis1977spearman} showed the Spearman distance approximates $K_\tau$ within a factor of two: 
 \begin{equation}\label{eq:diaconis_inequality}
@@
 One obvious choice is fix $L$ via a linear combination of two filter functions, restrict $M$ to $L$, and compute the associated 1-parameter barcode. 
-However, this is an $O(m^3)$ time computation, which is prohibitive for exploratory data analysis purposes.
+However, this is an $O(m^3)$ time computation, which is prohibitive for interactive data analysis purposes.
 
@@
 After Lee et al published their work, Carlsson et al.~\cite{carlsson2008local} subsequently performed extensive experiments using persistent homology, wherein he found that the distribution of high-contrast $3 \times 3$ patches is actually well-approximated by a Klein bottle $\mathcal{M}$---around 60\% of the high-contrast patches from the still image data set lie within a small neighborhood around $\mathcal{M}$ which accounts for only 21\% of the 7-spheres volume.
-Along a similar vein in the sparse coding context, Perea et al.~\cite{perea2014klein} introduced a dictionary learning framework for estimating the distribution of patches from texture images.
+Along a similar vein in the sparse coding context, Perea~\cite{perea2014klein} introduced a dictionary learning framework for estimating the distribution of patches from texture images.
 
@@
 Moreover, though we have carefully constructed an efficient greedy heuristic in section~\ref{sec:proxy_objective} and illustrated a different perspective with which to view our heuristic (via crossing minimization), it is an open question whether there exists a more structured reduction of~\eqref{eq:schedule_cost} or~\eqref{eq:sorting_additive} to a better-known problem.
+
+\section{Declarations}
+
+\textbf{Ethical Approval:} Not applicable.
+\\
+\\
+\noindent \textbf{Competing interests:} 
+The authors declare that they have no competing interests that could influence the interpretation or presentation of the research findings.
+There are no financial or personal relationships with individuals or organizations that could bias the outcomes of this work.
+\\
+\\
+\noindent
+\textbf{Authors' contributions:}
+The contributions of each author to this article were as follows:
+\begin{enumerate}
+	\item Matt Piekenbrock: conceptualization, methodology design, algorithm development, experimental design \& analysis, software development, drafting of the article. 
+	\item Jose Perea: Literature review, critical revision of the manuscript, final approval of the version to be published.
+\end{enumerate}
+All authors have reviewed and approved the final version of the manuscript and have agreed to be accountable for all aspects of the work.
+\\
+\\
+\noindent 
+\textbf{Funding:}
+The research presented in this work is partially supported by the National Science Foundation through grants CCF-2006661 and CAREER award DMS-1943758.
+The funding source had no role in the study design, data collection and analysis, decision to publish, or preparation of the manuscript.
+\\
+\\
+\noindent \textbf{Availability of data and materials:} Links to the software and data used for the experiments can be found at: \url{https://github.com/peekxc/move_schedules}.
 

### 2023-05-26 move_schedules.tex
Cleaning up the document

@@
-% \documentclass[review]{siamart190516} % for SICOMP 
 %\documentclass{article} % for arxiv only
@@
 %\pagestyle{plain} % arxiv only
-\PassOptionsToPackage{noend}{algpseudocode}
+
 \documentclass[sn-mathphys]{sn-jnl}
 
-\usepackage{amsmath, amssymb}
+% --- Packages ---
+\usepackage{amsmath, amssymb, amsthm}
 \usepackage{mathtools}
@@
 \usepackage{blkarray, bigstrut}
-\newcommand\topstrut[1][1.0ex]{\setlength\bigstrutjot{#1}{\bigstrut[t]}}
-\newcommand\botstrut[1][0.9ex]{\setlength\bigstrutjot{#1}{\bigstrut[b]}}
-
-\usepackage{epstopdf}
-%\epstopdfDeclareGraphicsRule{.pdf}{png}{.png}{convert #1 \OutputFile}
-%\DeclareGraphicsExtensions{.png,.pdf}
-\DeclareGraphicsExtensions{.pdf,.png}
-
-%\usepackage[nodisplayskipstretch]{setspace}
-%\setstretch{1.5}
-
+\usepackage{epstopdf} % \DeclareGraphicsExtensions{.pdf,.png} % for pdf 
 \usepackage{caption} 
-\captionsetup[table]{skip=4pt, position=above}
-
-\makeatletter
-\newcommand{\customlabel}[2]{%
-   \protected@write \@auxout {}{\string \newlabel {#1}{{#2}{\thepage}{#2}{#1}{}} }%
-   \hypertarget{#1}{#2}
-}
-\makeatother
-
-\newcommand\undermat[2]{%
-  	\makebox[0pt][l]{$\smash{\underbrace{\phantom{%
-    \begin{matrix}#2\end{matrix}}}_{\text{$#1$}}}$}#2}
-
+\include{commands.tex}
+\usepackage{outlines}
+\usepackage{enumitem}
 \usepackage{hyperref} % load hyperref last
 
-% For arxiv; comment out for SICOMP
+% --- arxiv only ---- 
+%\usepackage{xcolor}
+%\usepackage{algorithm}% http://ctan.org/pkg/algorithms
+%\PassOptionsToPackage{noend}{algpseudocode}
+%\usepackage{algpseudocode}% http://ctan.org/pkg/algorithmicx
+
+% --- Customizations --- 
 \newtheorem{theorem}{Theorem}
@@
 
-%\newsiamremark{remark}{Remark}
+\captionsetup[table]{skip=4pt, position=above}
 
@@
 
-\usepackage{outlines}
-\usepackage{enumitem}
 \setenumerate[1]{label=\arabic*}
@@
 \setstretch{1}
-%\setlength{\abovedisplayskip}{0pt}
-%\setlength{\belowdisplayskip}{0pt}
-%\setlength{\abovedisplayshortskip}{0pt}
-%\setlength{\belowdisplayshortskip}{0pt}
-
-% Sets running headers as well as PDF title and authors
-% \headers{Fast Persistence Computations}{M. Piekenbrock and J. Perea}
-
-% Title. If the supplement option is on, then "Supplementary Material"
-% is automatically inserted before the title.
+
+% --- Springer only ---
 \title[Fast Persistence Computations in Dynamic Settings]{
@@
 
-% Authors: full names plus addresses.
-%\author{
-%	Matthew Piekenbrock\thanks{Khoury College of Computer Sciences, Northeastern University.} and Jose A. Perea\thanks{
-%	Department of Mathematics and Khoury College of Computer Sciences, Northeastern University.}
-%}
-
-% Spring format
+% --- Arxiv only --- 
+%\title{\vspace{-0.0em} Move Schedules: Fast persistence \\ computations  in coarse dynamic settings\thanks{This work was partially supported by the National Science Foundation through grants CCF-2006661 and CAREER award DMS-1943758.}\vspace{-0.5em}}
+%\author{Matt Piekenbrock\thanks{Khoury College of Computer Sciences, Northeastern University.}\;\; and Jose A.
+Perea$^\dagger$\thanks{Department of Mathematics and Khoury College of Computer Sciences, Northeastern University.}}
+%\date{}
+
+% --- Springer only ---
 \author[1]{\fnm{Matthew} \sur{Piekenbrock}}\email{piekenbrock.m@northeastern.edu}
@@
 	\orgdiv{Khoury College of Computer Sciences}, 
-	\orgname{Northeastern University}}
-	%\orgaddress{\street{360 Huntington Ave}, \city{Boston}, \postcode{02115}, \state{MA}%\country{United States}}
-% }
-% , ,
-
+	\orgname{Northeastern University}
+}
+
+% --- Springer only --- 
 \author[2]{\fnm{Jose} \sur{A. Perea}}\email{j.pereabenitez@northeastern.edu}
@@
 
-%\title{\vspace{-2.0em} Move Schedules: \\ Fast persistence computations in sparse dynamic settings\thanks{This work was partially supported by the National Science Foundation through grants CCF-2006661 and CAREER award DMS-1943758.}\vspace{-0.5em}}
-%\author{Matt Piekenbrock\thanks{Department of Computational Mathematics, Science \& Engineering, Michigan State University}\;\; and Jose A.
-Perea$^\dagger$\thanks{Department of Mathematics, Michigan State University}}
-%\date{}
-
-% Optional PDF information
-% \ifpdf
-% \hypersetup{
-%   pdftitle={Move Schedules},
-%   pdfauthor={Matthew Piekenbrock and Jose A. Perea}
-% }
-% \fi
-
-% \raggedbottom
-
 % ------  Begin document ------- 
-% https://github.com/peekxc/dart
 \begin{document}
 
+%\maketitle  % arxiv only, ensures abstract follows title
+
 % REQUIRED - "Abstracts must be able to stand alone and so cannot contain citations to the paper's references, equations, etc. .... not exceeding 250 words"
-%\begin{abstract}
-\abstract{
+%\begin{abstract} % arxiv only 
+\abstract{ % springer only
 Matrix reduction is the standard procedure for computing the persistent homology of a filtered simplicial complex with $m$ simplices.
@@
 	Since there are $O(m^2)$ such transpositions, this maintenance procedure exhibits limited scalability and often is too fine for many applications. 
-%	As our first result, we propose a coarser strategy for maintaining the decomposition over a 1-parameter family of filtrations that takes $O(m^2)$ preprocessing time uses only $O(m)$ space. 
-%	We also show that this maintenance procedure yields a factor-2 net reduction in column operations over the state of the art approach. 
 We propose a coarser strategy for maintaining the decomposition over a 1-parameter family of filtrations that requires only $O(m \log \log m)$ time and $O(m)$ space to construct. 
@@
 Applications to video data, dynamic metric space data, and multi-parameter persistence are also presented.
-}
-%\end{abstract}
+} % springer only
+%\end{abstract} % arxiv only
+
+\maketitle %  Springer only, ensures abstract follows title
 
 % Keywords and MSC codes must accompany all articles. A list of the subject classifications can be accessed or searched online in the Annual Index of Mathematical Reviews.
-% REQUIRED
-%\begin{keywords}
-%  {Computational topology, Persistent homology, Topological data analysis}
-%\end{keywords}
 \keywords{Computational topology, Persistent homology, Topological data analysis}
 
-% REQUIRED
-% See: https://mathscinet.ams.org/mathscinet/msc/msc2020.html?t=68T09&btn=Current
-%\begin{AMS}
-%  68T09, 55N31, 62R40
-%\end{AMS}
+% --- Springer only (REQUIRED) ---
 \pacs[MSC Classification]{68T09, 55N31, 62R40}
-
-\maketitle
 
@@
 \subsection{Vineyards}\label{sec:vineyards}
+\vspace{-0.5em}
 Consider a homotopy $F(x,t) : \mathbb{X} \times [0,1] \to \mathbb{R}$ on a triangulable topological space $\mathbb{X}$, and denote its ``snapshot'' at a given time-point $t$ by $f_t(x) = F(x,t)$.

### 2023-10-06 reviewer2_jact_response.txt
Addressing most of reviewer 2's comments

Reviewer 1:

Reviewer 2
I believe the paper contains very good and interesting results, but it needs a revision before being published.

Pros: I really like the main idea of the paper: using the machinery of moves that allows one to efficiently recompute the R=DV decomposition after one cyclic permutation in the input filtration, the authors find a heuristic to decompose an arbitrary permutation of the filtration into a sequence of moves of provably minimal length and small cost (this is the heuristic part). The authors show experimentally that this idea works. The paper is as self-contained as possible and the authors work out some concrete examples which are helpful to the reader.

Cons: The detailed list of the remarks follows. Sorry, it’s not ordered by importance, so there are trivial things like grammar/typos and more serious issues.
 In general, I feel that the paper is a bit too verbose - it can be made shorter without losing virtually any content.
The most important issues that are listed below are: the comment about ‘MoveLeft’ and some missing (or hard to find) details about the algorithm (GreedySchedule in Alg. 1).
Also, I would appreciate if the authors formulated  exactly what is the implementation they used in their experiments (e.g., did they really use treaps?).

----

Point by point responses:

Page 1, abstract: I don’t understand the $O(m \log \log m)$ estimate here. According to Thm. 1, this is the time it takes just to find $d$, the length of the optimal schedule. Computing the schedule itself costs $O(dm \log m)$ or $O(m \log m)$ per update. Shouldn’t this be in the abstract? Also, why ‘simulation of persistence’ (this expression will appear below, too)?

* - Yes, the $O(m \log \log m)$ complexity refers to the complexity of finding the length of the schedule and the LIS (only). We leave out the complexity of constructing the schedule itself in the abstract as we envision there are situations where the algorithm used to actually construct the schedule may vary---depending on the particular data structure of choice supporting the operations outlined in 3.3 (e.g. implicit treaps, segment tree, vEB tree), the deterministic and expected time complexity may change.

Page 4, 'if the relations are available a-priori' – do the authors mean ‘edges’?

- Yes.

Page 7, ‘we summarize our main results (Theorem 1)’. The main claim in theorem 1 is that one can efficiently compute only the length of the optimal schedule. Will the authors consider reorganizing?

- see above.

Page 8: Although $d$ can be  $O(m)$, we provide evidence that $d ~ m - sqrt(m)$. Well, $m - \sqrt(m)  = O(m)$, why ‘although’? Should it be ‘Although d can be O(m) and we provide evidence that d ~ m - sqrt(m), d is much smaller in practice’?

- Yes, as the previous reviewers also pointed out, we aware that $m - \sqrt(m)  = O(m)$, which is why we omitted the big-O in the notation $d ~ m - sqrt(m)$. To remedy this, we amended the statement to clarify.

Page 8: ‘an simplexwise’ -> a simplexwise

- fixed.

Page 9: I am more used to the ‘positive/negative simplex’ terminology than ‘creator/destroyer’.
This is a matter of personal preference, of course.

-

Page 10: ‘Despite this high algorithmic complexity, the number of column operations has been observed to be super-linear in practice’ – this has to be rephrased, because there is no contradiction (order of m^2 column operations is super-linear, too).

- TODO: cite ripser

Sec. 2.3, Moves. Sorry, it might be my fault but I had difficulty understanding the procedure from the description. I believe having the pseudocode right here will be helpful (after I read it in the appendix, I was happy with the algorithm). Especially since the algorithm is needed in the proof of Prop. 2, too.

- We previously moved much of pseudocode to the appendix in order to streamline the presentation of the concepts, as the pseudocode for both algorithms takes up an entire page (and the pseudocode for MoveRight is also available in Busaryev's paper). We have moved them back. 
(todo: fix?)

Page 16, Alg. 1. 
a) The authors say ‘Algorithm 1 may be easily modified to be completely online, keeping at most two filtrations and one decomposition in memory’. I agree, but I believe it would be equally easy and more natural to state it in this form immediately. The change is in modifying the meaning of \rho_i: while now it is the permutation taking f_1 to f_i, it should be replaced with the permutation taking f_{i-1} to f_i.

- The algorithm is now presented in the much simpler "online form", which only requires as input a pair of filtrations and a single D = RV decomposition. The original batch-processing presentation of pseudocode reflected more closely the implementation we used (we found it more efficient to compute a batch of schedules, concatenate them, and then give them to the moving algorithm to process sequentially, saving the diagrams between each schedule execution).

b) Is it correct that the authors use both MoveRight and MoveLeft in their implementation? I believe it is necessary, and Alg. 1 reflects that; however, it raises some questions about the sequel. I’ll get back to this point.

- Yes!

c) Line 6 calls the function GreedySchedule. There is no such algorithm in the paper. There is Alg. 5 in the appendix, LCS-sort, but there is no greedy heuristic in it. Also, it computes the LIS internally, while GreedySchedule takes the longest increasing sequence as a second argument. Please fix this.

- The GreedySchedule algorithm wasn't included in part because it corresponds to simply modifying line 6 in Alg. 5 (choosing the next 'symbol'/simplex to move) with equation (22).
(todo: mention this or actually build out the pseudocode for 3.4.2)   (TODO... make it explicit)

Page 17, the paragraph after Def. 1: ‘the former can be that can be solved optimally’ - this sentence should be fixed.

- Fixed.

Page 19: ‘small changes … affects’ -> affect.

- Fixed

Note that $m_{i,j}$ is defined here and used in Alg. 5 at the very end of the paper without explanation.

- The usage in Alg. 5 was intentional as it defined in equations (10) and (11); a reference to those equations has been added in Alg. 5.

Page 20: ‘allows use to exploit’ -> allows us to exploit

- Fixed.

Page 21: Do I understand correctly that this is what is going on in the proof of Corollary 1: $E(m - |LCS(p, q)|) = m - 2 \sqrt(m) - cm^{⅙} + o(m^{⅙})$ by the known result of Baik for $E|LIS(p)|$, therefore it is less than $m - \sqrt(m)$ ? So any function which is asymptotically less than $ 2 \sqrt(m) - cm^{⅙} + o(m^{⅙} $ would do? Another question : Corollary 1 speaks about two random filtrations, which the authors interpret as ‘two permutations chosen uniformly at random’ (explaining why this is okay in light of Boissonat’s result). But the longest increasing sequence result by Baik is for a single permutation drawn uniformly at random. Is it obvious that, if p and q are drawn from S_m uniformly at random, then $p \circ q^{-1}$ (the permutation we should consider for LIS) is also uniformly distributed? Well, for symmetry reasons it seems to be true and obvious, but, given the level of details in the other parts of the paper, the authors may want to include this as well.

- We agree that $p \circ q^{-1}$ is uniform, and we agree the proof strategy for Corollary 1 outlined by the reviewer would match the level of detail of the rest of the paper. That said, in our struggle to keep the length of the paper manageable we chose to present the proof in a succinct and informal way.

Page 22, ‘It is clear from Corollary 1’ -  wrong reference. Probably, from the proof of proposition 3 (Prop. 3 itself does not say anything about LIS).

- Yes, this used to be its own corollary. The proof of prop 3 is indeed the reduction we're referring too. We've amended the statement to explicitly point to the proof.

Let T denote a ordered set-like -> an ordered set-like

- Fixed.

Page 27: equation (22) may solved  -> equation (22) can be solved

- Fixed.

General comment on the greedy heuristic: does most of the reasoning in section 3.4 apply to the case of MoveRight only? The starting point is the term \sum |I_k| + |J_k|, which only applies to moving simplex i to simplex j if i < j. What about the MoveLeft? Can it be analyzed theoretically? Does it occur frequently in the authors’ experiments? Can the freedom that the authors have in choosing the schedule of the optimal length be used to minimize the number of left moves and, if I want to implement this approach, should I do that or would it only harm me?

- We actually included a discussion and small analysis of MoveLeft, including some empirical comparisons, in the original draft of the page. To answer your questions successively:

* "does most of the reasoning ... apply to the case of MoveRight only?"

	- Yes, due to theoretical limitations, the statement only holds for MoveRight.

* "What about the MoveLeft? Can it be analyzed theoretically?"

	- We have a small discussion of MoveLeft in the appendix. We left out 
	Our internal

Sec. 4.1: I find the phrase ‘simulate persistence’ unusual. Does it just mean ‘compute persistent homology’? As for the second test (page 28): was the Algorithm 1 used verbatim, or was the scheduling performed to move from one snapshot to the next one?

Page 32, 4.3.1:  Is  \kappa from [10] the same as defined in Theorem 1? I suspect not, [10] says it’s coarseness. Please clarify.

Page 34: ‘they a convenient basis for D may be obtained’ - is this supposed to be ‘they prove that a convenient basis …’ ?
voronoi  -> Voronoi
7-spheres volume -> 7-sphere’s volume
In order reveal -> In order to reveal

Page 35: ‘Projecting the image data onto the first two basis vectors of leads’ - basis vectors of what?

Page 36: ‘the number of permutations applied throughout the encountered along the traversal of the dual graph’ - please fix this sentence.
vineyards is, moves requires -> either ‘vineyards/moves algorithm/approach is/requires’, or ‘vineyards are/moves require’. Or make vineyards and moves boldface or italics, then they are visually perceived as names.
that naively computing -> than naively computing

Page 37: if the filtrations in the parameterization family is nearby -> are nearby

Page 43 , A1.1.1 'After setting V is set' -> After setting V
'Since there exists complexes' -> Since there exists a complex / there exist complexes

Page 45. We recall an important claim .. - What is the claim? Seems to be missing, it should say something about positive/negative simplices.

Page 48. Well, in A.2.1 the authors mention \kappa as coarseness parameter! Also, the naming clash between the structure T used to compute the sorting and the barcode template T is unfortunate.

kendall -> Kendall

-----

readability - We switched the notation from symbol to permutations / indices in section 3.4, and we've reworked the pseudocode ; symbols 
on the overhead of moves - we agree that the figures may seem misleading; future work, output / geometry sensitive . Talk about balance the fraction  of kendall distances
assessment - talk abotu vineyards and non-trivial implementation details; it woudl portray vineyards inaccuractely

Figure 1 - fixed with upper level set filtration or equiv lower filtatratin with 1 - intensity

### 2023-10-06 move_schedules.tex
Addressing most of reviewer 2's comments

@@
 %\pagestyle{plain} % arxiv only 
-
+%%\usepackage[margin=1in]{geometry} % arxiv only
+
+\PassOptionsToPackage{noend}{algorithmic}
+\PassOptionsToPackage{noend}{algpseudocode}
 \documentclass[sn-mathphys]{sn-jnl}
@@
 
+%\usepackage[noend]{algorithmic}
+
 % --- arxiv only ---- 
@@
 	Persistence diagrams are known to vary continuously with respect to their input, motivating the  study of their computation for time-varying filtered complexes.
-Computationally, simulating persistence dynamically can be reduced to maintaining a valid decomposition under adjacent transpositions in the filtration order.
+Computing persistence dynamically can be reduced to maintaining a valid decomposition under adjacent transpositions in the filtration order.
 	Since there are $O(m^2)$ such transpositions, this maintenance procedure exhibits limited scalability and often is too fine for many applications. 
-We propose a coarser strategy for maintaining the decomposition over a 1-parameter family of filtrations that requires only $O(m \log \log m)$ time and $O(m)$ space to construct. 
-By reduction to a particular longest common subsequence problem, we show the storage needed to employ this strategy is actually sublinear in expectation. 
+We propose a coarser strategy for maintaining the decomposition over a 1-parameter family of filtrations.
+By reduction to a particular longest common subsequence problem, we show that the minimal number $d$ of updates to the decomposition can be found in $O(m \log \log m)$ time and $O(m)$ space, and that the corresponding sequence of permutations---which we call a \emph{schedule}---can be constructed in $O(d m \log m)$ time.
+We also show that, in expectation, the storage needed to employ this strategy is actually sublinear in $m$. 
 Exploiting this connection, we show experimentally that the decrease in operations to compute diagrams across a family of filtrations, is proportional to the difference between the expected quadratic number of states and the proposed sublinear coarsening.
@@
 If  zeroth homology  is the only dimension of interest, computing and updating both the persistence and rank information  is greatly simplified.
-For example, if the relations are available a-priori, obtaining a tree representation fully characterizing the connectivity of the underlying space (also known as the \emph{incremental connectivity} problem) takes just $O(\alpha(n) n)$ time using the disjoint-set data structure, where $\alpha(n)$ is the extremely slow-growing inverse Ackermann function.
+For example, if the edge are available a-priori, obtaining a tree representation fully characterizing the connectivity of the underlying space (also known as the \emph{incremental connectivity} problem) takes just $O(\alpha(n) n)$ time using the disjoint-set data structure, where $\alpha(n)$ is the extremely slow-growing inverse Ackermann function.
 Adapting this approach to the time-varying setting, Oesterling et al.~\cite{oesterling2015computing} give an algorithm that maintains a \emph{merge tree} with $e$ edges in $O(e)$ time per-update.
@@
 \subsection{A Motivating Example}\label{sec:motivation} 
-We  will use a simple experiment to illustrate why the vineyards algorithm does not always yield an efficient strategy for time-varying settings. 
+To motivate this effort, we begin with an illustrative example of why the vineyards algorithm does not always yield an efficient strategy for time-varying settings. 
 Consider a series of grayscale images (i.e.
@@
  We refer to the former as the \emph{discrete setting}, which is often used in practice, and the  latter as the \emph{continuous setting}.
-Note that though the discrete setting is often more practical, it is not guaranteed to capture all homological changes in persistence that occur in simulating a continuous 1-parameter family of diagrams.
+Note that though the discrete setting is often more practical, it is not guaranteed to capture all homological changes in persistence that occur in the continuous 1-parameter family of diagrams.
 
@@
 We also show that the information needed to construct any $\mathcal{S}$ with optimal size can be computed in $O(m \log \log m)$ preprocessing time and $O(m)$ memory.
-Although $d$  can be $O(m)$ for pathological inputs, we provide evidence  that $d \sim m - \sqrt{m}$ in expectation for random filtrations   (Corollary~\ref{cor:expectation}), and we give empirical results suggesting $d$  can be much smaller for time-varying filtrations.
+We provide evidence  that $d \sim m - \sqrt{m}$ in expectation for random filtrations (Corollary~\ref{cor:expectation}).
+Although this implies $d$ can be $O(m)$ for pathological inputs, we give empirical results suggesting $d$  can be much smaller in practice.
 %Finally, we perform extensive experiments 
@@
 whenever $j$ is the immediate successor of $i$ in $I$.
-Any finite filtration may be trivially converted into an simplexwise filtration via a set of \emph{condensing}, \emph{refining}, and \emph{reindexing} maps (see~\cite{bauer2021ripser} for more details).
+Any finite filtration may be trivially converted into a simplexwise filtration via a set of \emph{condensing}, \emph{refining}, and \emph{reindexing} maps (see~\cite{bauer2021ripser} for more details).
 Equivalently, a filtration can be also defined as a pair $(K, f)$ where $K$ is a simplicial complex and $f : K \to I$ is a \emph{filter function}  satisfying $f(\tau) \leq f(\sigma)$ in $I$, whenever $\tau \subseteq \sigma$ in  $  K$.
@@
 
- Given a filtration $(K, f)$ with $m$ simplices,  the main output of the reduction algorithm is a matrix decomposition $R = D V$, where the persistence diagrams are encoded in $R$ and the generating cycles in the columns of $V$.
+ Given a filtration $(K, f)$ with $m$ simplices,  the output of the reduction algorithm is a matrix decomposition $R = D V$, where the persistence diagrams are encoded in $R$ and the generating cycles in the columns of $V$.
  To begin the reduction, one first  assembles the elementary boundary chains $\partial(\sigma)$ as columns ordered according to $f$ into a $m \times m$  \emph{filtration boundary matrix} $D$.
 Setting $V = I$ and $R = D$, one proceeds by performing elementary left-to-right column operations on $V$ and $R$ until the following invariants are satisfied:
+\\
+\\
 \textbf{ Decomposition Invariants:}
@@
 We call the decomposition satisfying these three invariants \emph{valid}.
+The persistence diagrams of the corresponding filtration  can be determined from the lowest entries in $R$, once it has been reduced.
 Note that though the matrices $R$ and $V$ are not unique, the collection of persistent pairings are~\cite{zomorodian2005computing}.
-The persistence diagrams of the corresponding filtration  can be determined from the lowest entries in $R$, once it has been reduced.
 
@@
 Inspection of the reduction algorithm from~\cite{edelsbrunner2000topological} suggests that a loose upper bound for the reduction is $O(m^3)$, where $m$ is the number of simplices of the filtration---this bound is in fact tight~\cite{morozov2005persistence}.
-Despite this high algorithmic complexity, the number of column operations has been observed to be super-linear in practice, due in part to the high sparsity and structure of $D$. 
-Moreover, many variations and optimizations to Algorithm~\ref{alg:reduce} have been proposed over the past decade, see~\cite{bauer2017phat, bauer2021ripser, chen2011persistent} for an overview. 
+% the number of column operations has been observed to scale linearly in the size of the filtration in practice~\cite{bauer2021ripser}, due in part to the high sparsity and structure of $D$. 
+Despite this high algorithmic complexity, many variations and optimizations to Algorithm~\ref{alg:reduce} have been proposed over the past decade, see~\cite{bauer2021ripser, bauer2017phat, chen2011persistent} for an overview.
 % Important corollaries of the persistence pairing lemma is that the 
@@
 As homological critical events only occur when the filtration order changes, detecting all such events may be reduced to computing valid decompositions at   time points interleaving all changes in the filtration order.
-For simplexwise filtrations, these changes manifest as transpositions of adjacent simplices, and thus any fixed set of rules that maintains a valid $R = D V$ decomposition under adjacent column transpositions, is sufficient to simulate persistence dynamically.
+For simplexwise filtrations, these changes manifest as transpositions of adjacent simplices, and thus any fixed set of rules that maintains a valid $R = D V$ decomposition under adjacent column transpositions is sufficient to compute persistence dynamically.
 
@@
 However, several factors can affect the runtime efficiency of the vineyards algorithm. 
-On the positive side, as both the $V$ and $R$ matrices tend to be quite sparse, the cost of a given column operation is proportional to the number of non-zero  entries in the two columns being modified.
-As a rule of thumb, most transpositions require no column operations~\cite{edelsbrunner2000topological}.
+On the positive side, as both the $V$ and $R$ matrices are often sparse, the cost of a given column operation is proportional to the number of non-zero  entries in the two columns being modified.
+Moreover, as a rule of thumb, its been observed that most transpositions require no column operations~\cite{edelsbrunner2000topological}.
 On the negative side, one needs to frequently query the non-zero status of various entries in $R$ and $V$ (consider evaluating e.g.
@@
 Notice that these operations may yield an unreduced $R' = P R(\ast) P^T$, breaking invariant I3.
-We could reduce $R'$ with an additional $k$ operations, with the possibility that $k \sim O(\lvert i - j\rvert^2)$, but then this is no more efficient than re-running the reduction algorithm on columns $[i,j]$ in $R$.
+We could reduce $R'$ with an additional column operations, but the number of such operation scales $O(\lvert i - j\rvert^2)$, which no more efficient than simply performing the permutation and re-reducing the columns $[i,j]$ in $R$.
 
@@
 
-\vspace{-3.0em}
+\vspace{-2.2em}
 
@@
 The pseudo-code for \emph{MoveRight} is given in Algorithm~\ref{alg:mr} and for \emph{MoveLeft} in Algorithm~\ref{alg:ml}.
+
+\begin{algorithm}[!htb]
+	\caption{Move Right Algorithm}\label{alg:mr}
+	\setstretch{1.15}
+    \begin{algorithmic}[1]
+        \Function{RestoreRight}{$R$, $V$, $\mathbb{I} = \{I_1, I_2, \dots, I_s \}$, donors = $\mathrm{true}$}
+            \State $(\, d_{low}, \, d_R, \, d_V \,) \gets ( \, \mathrm{low}_R(I_1), \, \mathrm{col}_R(I_1), \, \mathrm{col}_V(I_1) \, )$
+        	\For{$k$ \textbf{in} $I_2, \dots, I_s$} 
+        		\State $( \, d_{low}', \, d_R', \, d_V' \, ) \gets ( \, \mathrm{low}_R(k), \, \mathrm{col}_R(k), \, \mathrm{col}_V(k) \, )$
+        		\State $( \, \mathrm{col}_R(k), \, \mathrm{col}_V(k) \, ) \mathrel{+}= (\, d_R, \, d_V \,) $
+        		\If {$d_{low}' < d_{low}$}
+        			\State $(\, d_{low}, \, d_R, \, d_V \,) \gets (\, d_{low}', \, d_R', \, d_V' \, )$
+        		\EndIf
+        	\EndFor 
+        	\State \Return $(\, R, \, V, \, d_R, \, d_V \, )$ \textbf{if} $\mathrm{donors}=\mathrm{true}$ \textbf{else} $(\, R, \, V \, )$
+        \EndFunction
+    \end{algorithmic}
+    \begin{algorithmic}[1]
+        \Function{MoveRight}{$R$, $V$, $i$, $j$} 
+            \State $\mathbb{I} =$ columns satisfying $V[i, i:j] \neq 0$ %\Comment{$O(m)$}
+            \State $\mathbb{J} = $ columns satisfying $\mathrm{low}_R \in [i:j]$ and $\mathrm{row}_R(i) \neq 0$ %\Comment{$O(m)$}
+            \State $(\, R, \, V, d_R, d_V \,) \gets $ \Call{RestoreRight}{$R$, $V$, $\mathbb{I}$} 
+            %\Comment{Restores $p$-simplices} \Comment{$O(\lvert \mathbb{I} \rvert)$}
+            \State $(\, R, \, V \,) \gets $ \Call{RestoreRight}{$R$, $V$, $\mathbb{J}$, $\mathrm{false}$}
+            	%\Comment{Restores $(p+1)$-simplices}
+            	%\Comment{$O(m \lvert \mathbb{I} \rvert)$}
+            \State $(\, R, \, V\, ) \gets (\, P R P^T, \, P V P^T\,)$  % \Comment{$P$ is a permutation given by (i,j)}
+            \State $(\, \mathrm{col}_R(j), \, \mathrm{col}_V(j) \, ) \gets (\, P d_R, \, P d_V \,)$
+            \State \Return $(\, R, \, V\,)$
+        \EndFunction
+    \end{algorithmic}
+\end{algorithm}
+
+\begin{algorithm}[!htb]
+	\caption{Move Left Algorithm}\label{alg:ml}
+	\setstretch{1.15}
+    \begin{algorithmic}[1]
+        \Function{RestoreLeft}{$R$, $V$, $\mathbb{K} = \{k_1, k_2, \dots, k_s \}$}
+            % \State $\mathrm{low}_K \gets \{ \mathrm{low}_R(k_1), \mathrm{low}_R(k_2), \dots, \mathrm{low}_R(k_s) \}$
+            \State $(l, r) \gets $ indices $l, r \in \mathbb{K}$ satisfying $l < r$, $\mathrm{low}_R(l) = \mathrm{low}_R(r)$ maximal 
+            \While{$\mathrm{low}_R(l) \neq 0$ \textbf{and} $\mathrm{low}_R(r) \neq 0$}
+            	\State $(\mathrm{col}_R(r), \mathrm{col}_V(r)) \mathrel{+}= (\mathrm{col}_R(l), \mathrm{col}_V(l))$
+            	\State $\mathbb{K} \gets \mathbb{K} \setminus l$
+            	\State $(l, r) \gets $ indices $l, r \in \mathbb{K}$ satisfying $l < r$, $\mathrm{low}_R(l) = \mathrm{low}_R(r)$ maximal
+            \EndWhile
+        	\State \Return $(\, R, \, V \, )$
+        \EndFunction
+    \end{algorithmic}
+
+    \begin{algorithmic}[1]
+        \Function{MoveLeft}{$R$, $V$, $i$, $j$} 
+        	\State $\mathbb{I} \gets \emptyset$
+            % \State $(\mathbb{K}, d_R, d_V) \gets (\emptyset, \mathrm{col}_R(i), \mathrm{col}_V(i))$
+            \While{$V(k, i) \neq 0 $\textbf{ for }$ k = \mathrm{low}_V(i)$\textbf{ where }$j \leq k < i$}
+            	\State $(\mathrm{col}_R(i), \mathrm{col}_V(i)) \mathrel{+}= (\mathrm{col}_R(k), \mathrm{col}_V(k))$
+            	\State $\mathbb{I} \gets \mathbb{I} \cup k + 1 $
+            \EndWhile
+            \State $(R, V) \gets (P R P^T, P V P^T)$
+            \State $\mathbb{J} = $ columns satisfying $\mathrm{low}_R \in [i:j]$ and $\mathrm{row}_R(i) \neq 0$ %\Comment{$O(m)$}
+
+            \State $(R, V) \gets$ \Call{RestoreLeft}{$R, V, \mathbb{I}$}
+            \State $(R, V) \gets$ \Call{RestoreLeft}{$R, V, \mathbb{J}$}
+            % \State $(\mathrm{col}_R(i), \mathrm{col}_V(i)) \gets (P d_R, P d_V )$
+			\State \Return $(\, R, \, V\,)$
+        \EndFunction
+    \end{algorithmic}
+\end{algorithm}
 
@@
 Prior to applying any permutation $P$ to the decomposition, it is necessary to remove non-zero entries in $V$ which render $P^TVP$ non-upper triangular, to maintain invariant I2.
-Using vineyards, one must consistently perform $\lvert i - j \rvert - 1$ non-zero status queries interleaved between repairing column operation.
+Using vineyards, one must consistently perform $\lvert i - j \rvert - 1$ non-zero status queries interleaved between repairing column operations.
 A move operation, on the other hand, groups these status queries into a single pass prior to performing any modifying operations.
@@
 Indeed, Property M3 directly follows from this fact, as one may simply move each simplex $\sigma \in K$ into its new order $f'(\sigma)$ via insertion sort.
-Note that the number of valid decomposition produced by vineyards is bounded above by $O(m^2)$ if each pair of simplices $\sigma_i, \sigma_j \in K$ switches its relative ordering at most once during the interpolation from $f$ to $f'$.
+Note that the number of valid decompositions produced by vineyards is bounded above by $O(m^2)$, as each pair of simplices $\sigma_i, \sigma_j \in K$ may switch its relative ordering at most once during the interpolation from $f$ to $f'$.
 
@@
     \begin{algorithmic}[1]
-    	\Require Ordered set of filtrations $\mathcal{F}$ with permutations $\rho_{i}: [m] \to [m]$
-    	\Ensure $R = D V$ is computed for each  $(K,f_1),(K,f_2), \dots, (K,f_n)$
-    	\Procedure{MoveSchedule}{$\mathcal{F} = ( f_1, f_2, \dots, f_n )$} 
-    		\State $\mathcal{S} = \emptyset$
-    		\For{$i = 1$\textbf{ to }$n-1$}
-    		    \State $q \gets  (\rho_i(1), \rho_i(2), \ldots, \rho_i(m))$ 
-    			\State $\mathrm{lis}_q \gets \mathrm{LIS}(q)$ \Comment{$O(m \log \log m)$}
-    			\State $\mathcal{S} \gets \mathcal{S} \cup \mathrm{GreedySchedule}(q, \mathrm{lis}_q)$ \Comment{$O(d^2\log m)$}
-    		\EndFor
-	    	\State $(R, V) \gets$ \Call{Reduction}{$D = \partial (K,f_1)$}
-	    	\For{$(i,j)$\textbf{ in }$\mathcal{S}$}
+    	\Require Filtration pair $\mathcal{F} = \{f_0, f_1\}$ and valid $R = DV$ for $(K,f_0)$
+    	\Ensure Valid $R = D V$ is computed for $(K, f_1)$
+    	\Procedure{MoveSchedule}{$\mathcal{F} = ( f_0, f_1 )$, $R$, $V$} 
+%    		\State Fix $\rho \in S_m$ such that $f_1 = \rho \circ f_0$ \Comment{$O(m)$} 
+%    		\State $\mathrm{lis}_{\rho} \gets \mathrm{LIS}(\rho)$ \Comment{$O(m \log \log m)$}
+			\State $h \gets \texttt{greedy}$ \Comment{Heuristic for picking moves}
+    		\State $\mathcal{S} \gets \mathrm{Schedule}(f_0, f_1, h)$ \Comment{$O(d^2\log m)$}
+%	    	\State $(R, V) \gets$ \Call{Reduction}{$D = \partial (K,f_1)$}
+	    	\For{$(i,j)$\textbf{ in }$\mathcal{S}$} \Comment{$\lvert \mathcal{S} \rvert = d$}
 	    		\State $(R, V) \gets$ \textbf{if} $i < j$ \Call{MoveRight}{i, j} \textbf{else} \Call{MoveLeft}{i, j}
@@
 \noindent Observe the size of the schedule $S_F$ defined from the homotopy $F$ above is exactly $K_\tau(f,f')$.
-On the positive side, the reduction of schedule planning to crossing detection implies the former can be  that can be solved optimally in output-sensitive $O(m \log m + k)$ time by several algorithms~\cite{boissonnat2000efficient}, where $k$ is the output-sensitive term and $m$ is the number of simplices in the filtration(s).
+On the positive side, the reduction of schedule planning to crossing detection implies the former can be   solved optimally in output-sensitive $O(m \log m + k)$ time by several algorithms~\cite{boissonnat2000efficient}, where $k$ is the output-sensitive term and $m$ is the number of simplices in the filtration(s).
 On  the negative side, $k = K_\tau(f,f')$ scales in size to $\sim O(m^2)$ in the worst case, achieved when $f' = - f$.
@@
 Therefore, $m = 417$ and $\lvert S_F \rvert \leq \frac{1}{2}m(m-1) = 86,736$.
-As the homotopy given by the video is simulated,  $\approx 70,\!000$ transpositions are generated, approaching the worst case upper bound due to the fact that $f'$  is nearly the reverse of $f$.  
+As the homotopy given by the video is varied,  $\approx 70,\!000$ transpositions are generated, approaching the worst case upper bound due to the fact that $f'$  is nearly the reverse of $f$.  
 \end{remark}
@@
 
-Perhaps surprisingly, small changes to set of allowable edit operations $\Sigma$ dramatically affects both the size of $d_\Sigma(p,q)$ and the difficulty of obtaining a minimal sorting. 
+Perhaps surprisingly, small changes to set of allowable edit operations $\Sigma$ dramatically affect both the size of $d_\Sigma(p,q)$ and the difficulty of obtaining a minimal sorting. 
 For example, while sorting by transpositions and reversals is NP-hard and sorting by prefix  transpositions is unknown, there are polynomial time algorithms for sorting by block interchanges, exchanges, and prefix exchanges~\cite{labarre2013lower}.
@@
 \begin{equation}\label{eq:move_perm}
-\setlength\arraycolsep{1.8pt}
+m_{ij} = \setlength\arraycolsep{1.8pt}
 \scalefont{0.90}{
@@
 \begin{equation}\label{eq:move_permutation}
-(\enspace i \enspace j \enspace j\text{-}1 \enspace \dots \enspace i\text{+}2 \enspace i\text{+}1 \enspace )
+m_{ij} = (\enspace i \enspace j \enspace j\text{-}1 \enspace \dots \enspace i\text{+}2 \enspace i\text{+}1 \enspace )
 \end{equation}
@@
 One might hope computing the size $\mathrm{d}_{\mathrm{lcs}}(p,q)$ alone exhibits a lower complexity, however there is substantial evidence that for general string inputs the complexity cannot be much lower than quadratic~\cite{abboud2015tight}.
-However, if the pair of filtrations $(K, f), (K, f')$ come from the same underlying complex $K$, then their filter function $f, f'$ may both be thought of a permutations in $S_m$---the corresponding edit distance $d$ then reduces to the \emph{permutation edit distance} problem.
+However, in our setting, the total orders of any pair of simplexwise filtrations $(K, f), (K, f')$ of the same underlying complex $K$ may be thought of a permutations in $S_m$---the corresponding edit distance $d$ then reduces to the \emph{permutation edit distance} problem.
 With this insight in mind, we obtain the following bound on the minimum size of a sorting (i.e. schedule) using moves and the complexity of computing it. 
@@
 Since the LCS distance is a valid edit distance, if $\lvert \mathrm{LCS}(p, q) \rvert = k$, then $\lvert \mathrm{LCS}(p', \iota) \rvert = k$ as well.
-Notice that $\iota$ is strictly increasing and that any common subsequence $\iota$ has with $p'$ must also be strictly increasing.
+Notice that $\iota$ is strictly increasing and that any common subsequence $p'$ has with $\iota$ must also be strictly increasing.
 Thus, the problem of computing $\mathrm{LCS}(p, q)$ reduces to the problem of computing the \emph{longest increasing subsequence} (LIS) of $p'$, which can done in $O( m \log \log m)$ time using van Emde Boas trees~\cite{bespamyatnikh2000enumerating}.
@@
 \noindent
-Establishing a connection between the permutation edit distance and move scheduling allows use to exploit the combinatorial structure that comes from the developed theory on both LCS's and LIS's, which are both well-studied objects.
+Establishing a connection between the permutation edit distance and move scheduling allows us to exploit the combinatorial structure that comes from the developed theory on both LCS's and LIS's, which are both well-studied objects.
 We record a single corollary to demonstrate this fact.
@@
 \begin{remark}
-\noindent Note the quantity from Corollary~\ref{cor:expectation} captures the size of $S^\ast$ between pairs of uniformly sampled permutations, as opposed to uniformly sampled filtrations, which have more structure.
+\noindent Note the quantity from Corollary~\ref{cor:expectation} captures the size of $S^\ast$ between pairs of uniformly sampled permutations, as opposed to uniformly sampled filtrations, which have more structure due to the face poset.
 %One may wonder if filtrations would exhibit a smaller upper bound due to the subface relation that must be respected.
@@
 
-We conclude this section with the main theorem of this effort: an output-sensitive bound on the simulation of persistence dynamically.
+We conclude this section with the main theorem of this effort: an output-sensitive bound on the computation of persistence dynamically.
   \begin{theorem}\label{prop:sparsity_result}
@@
 \subsection{Constructing schedules}\label{sec:schedule_construction}
-It is clear from Corollary~\ref{cor:expectation} that one may compute the LCS between two permutations $p, q \in S_m$ via the LIS of a single permutation $p^\ast$, and that computation may be carried out in $O(m \log \log m)$ time.
+It is clear from the proof in Proposition~\ref{prop:schedule_size} that one may compute the LCS between two permutations $p, q \in S_m$ via the LIS of a single permutation $p^\ast$, and that computation may be carried out in $O(m \log \log m)$ time.
 It is not immediately clear, however, how to obtain a sorting $p \mapsto q$ from a given $\mathcal{L} = \mathrm{LIS}(p')$ in an efficient way.
@@
 Recall that a \emph{sorting} $\mathcal{S}$ with respect to two permutations $p, q \in S_m$ is an ordered sequence of permutations $\mathcal{S} = (s_1, s_2, \dots, s_d)$ satisfying $q = s_d \circ \dots s_1 \circ p$.
-By definition, a subsequence of symbols in $\mathcal{L}$ common to both $p$ and $q$ satisfies:
+By definition, a subsequence in $\mathcal{L}$ common to both $p$ and $q$ satisfies:
 \begin{equation}
@@
 $q$).
-Thus, obtaining a sorting $p \mapsto q$ of size $d = m - \lvert \mathcal{L} \rvert$ reduces to applying a sequence of moves to symbols in the complement of $\mathcal{L}$.
+Thus, obtaining a sorting $p \mapsto q$ of size $d = m - \lvert \mathcal{L} \rvert$ reduces to applying a sequence of moves in the complement of $\mathcal{L}$.
  Formally, we define a permutation $s \in S_m$ as a \emph{valid} operation with respect to a fixed pair $p, q \in S_m$ if:
@@
   The problem of constructing a sorting $\mathcal{S}$ of size $d$ thus reduces to the problem of choosing a sequence of $d$ valid moves, which we call a \emph{valid sorting}.
-To do this efficiently, let $\mathcal{T}$ denote a ordered set-like data structure that supports the following operations on elements $\sigma \in M$ from the set $M = \{0, 1, \dots, m+1\}$:
+To do this efficiently, let $\mathcal{T}$ denote an ordered set-like data structure that supports the following operations on elements $\sigma \in M$ from the set $M = \{0, 1, \dots, m+1\}$:
   \begin{enumerate}
@@
   Given such a $\mathcal{T}$, an arbitrary valid sorting can be constructed by repeatedly querying and maintaining information about the $\mathrm{LCS}$ in $\mathcal{T}$.
-To see this, suppose $\mathcal{T}$ contains all of the symbols in the current $\mathrm{LCS}$ between two permutations $p$ and $q$.
+To see this, suppose $\mathcal{T}$ contains the current $\mathrm{LCS}$ between two permutations $p$ and $q$.
 By definition of the LCS, we have:
@@
   for every $\sigma \in \mathcal{T}$.
-Now, suppose we choose a symbol $\sigma \notin \mathcal{T}$.
-If $p^{-1}(\sigma) < p^{-1}(\mathcal{T}_{\mathrm{pred}}(\sigma))$, then we must move $\sigma$ to the right in $p$ such that~\eqref{eq:valid_move_ineq} holds.
-Similarly, if  $p^{-1}(\mathcal{T}_{\mathrm{succ}}(\sigma)) < p^{-1}(\sigma)$, then we must move $\sigma$ left in $p$ to increase the size of the LCS.
-Assuming the structure $\mathcal{T}$ supports all of the above operations in $O(\log m)$ time, we easily deduce a $O(d m\log m)$ algorithm for obtaining a valid sorting, which for completeness is shown via Algorithm~\ref{alg:sorting} in the appendix.
+Now, suppose we choose some element $\sigma \notin \mathcal{T}$ which we would like to add to the LCS.
+If $p^{-1}(\sigma) < p^{-1}(\mathcal{T}_{\mathrm{pred}}(\sigma))$, then we must move $\sigma$ to the right of its predecessor in $p$ such that~\eqref{eq:valid_move_ineq} holds.
+Similarly, if  $p^{-1}(\mathcal{T}_{\mathrm{succ}}(\sigma)) < p^{-1}(\sigma)$, then we must move $\sigma$ left of its successor in $p$.
+Assuming the structure $\mathcal{T}$ supports all of the above operations in $O(\log m)$ time, we easily deduce a $O(d m\log m)$ algorithm for obtaining a valid sorting.
+%  which for completeness is shown via Algorithm~\ref{alg:sorting} in the appendix. 
   % one can incrementally build a valid sorting of size $d$ by arbitrarily choosing valid move operations . all elements in $\mathrm{LCS}(p,q)$ into $\mathcal{T}$
@@
 \end{alignat*}
-where the notation $m_{x y}$ represents the move permutation that moves symbol $x$ to the position of symbol $y$.
-The cost of each  move operation and each schedule is recorded in Table~\ref{table:move_costs}.
+where the notation $m_{x y}$ represents the move permutation that moves $x$ to the position of $y$. The cost of each  move operation and each schedule is recorded in Table~\ref{table:move_costs}.
 \begin{table}[h]
@@
 \end{table}
-Note the greedy strategy which always selects the cheapest move in succession would begin by moving $x$ or $z$ first, since these are the cheapest moves available, which implies one of $S_1, S_2, S_5, S_6$ would be picked depending on the tie-breaker.
-While the cheapest schedule $S_1$ is in this candidate set, so is $S_6$, the most expensive schedule.
-As a result, a greedy strategy which chooses the lowest-cost move may not yield an optimal schedule.
+Note the greedy strategy which always selects the cheapest move in succession would begin by moving $x$ or $z$ first, since these are the cheapest moves available, which implies one of $S_1, S_2, S_5, S_6$ would be picked on the first iteration.
+While the cheapest schedule $S_1$ is in this candidate set, an iterative greedy procedure would pick either $S_2$ or $S_5$, depending on the tie-breaker---thus, a greedy approach picking the lowest-cost move may not yield an optimal schedule.
+	Indeed, as the most expensive schedule $S_6$ is in initial iterations candidate set, we see that a greedy-procedure with an arbitrary tie-breaker could potentially yield a \emph{maximal-cost} schedule.
 %Moreover, any reasonable tie-breaker that keeps track of previous cheapest moves would end up picking either schedule $S_2$ or $S_5$ as the final greedy schedule, both of which are suboptimal cost-wise.
@@
 \end{equation}
-The Spearman distance shares any similarities with $K_\tau$: it is a metric on $S_m$ that is invariant under consistent relabeling. 
+The Spearman distance shares any similarities with $K_\tau$; like $K_\tau$, it is also metric on $S_m$ that is invariant under consistent relabeling. 
 %and the distance between pairs of uniformly random permutations $p, q \sim S_m$ are asymptotically normal. 
@@
 To see this, consider two permutations: $p$ and $m_{ij} \circ p$, where $m_{ij}$ is a move permutation.
-Drawing $(p, m_{ij} \circ p)$ as a bipartite graph,  observe that $F(p, m_{i,j} \circ p)$ is twice the number of edge crossings in the graph, and that equality in~\eqref{eq:sorting_additive} is achieved when the displacement of each symbol between its initial position in $p$ to its value is non-increasing with every application of $s_i$, which in general not guaranteed using the schedule construction method derived in section~\ref{sec:move_schedules}.
+Drawing $(p, m_{ij} \circ p)$ as a bipartite graph,  observe that $F(p, m_{ij} \circ p)$ is twice the number of edge crossings in the graph, and that equality in~\eqref{eq:sorting_additive} is achieved when the displacement of each $\sigma \in p$ between its initial position in $p$ to its value is non-increasing with every application of $s_i$, which in general not guaranteed using the schedule construction method derived in section~\ref{sec:move_schedules}.
 Unfortunately, the $k$-layer crossing minimization problem is NP-hard for $k$ sets of permutations, when $k \geq 4$~\cite{biedl2009complexity}.
@@
     \centering
-    \includegraphics[width=0.80\textwidth]{crossings.png}
+%    \includegraphics[width=0.60\textwidth]{crossings.png} % arxiv only 
+    \includegraphics[width=0.75\textwidth]{crossings.png} % springer only
     %\caption{Three sortings, $S_1$, $S_2$, and $S_3$, with $21$, $31$, and $35$ crossings, resp.}
@@
 Each column represents the successive application of a move $m_{ij}$ in the schedule, 
-and the edges track how each  symbol has been moved. 
-Black/red vertices correspond to symbols in and outside of $\mathrm{LCS}$, respectively. 
+and the edges track the movement of each element of the permutation. 
+Black/red vertices correspond to elements in and outside of $\mathrm{LCS}$, respectively. 
 All three schedules were generated from the same $\mathrm{LCS}(p,q) = (4\,7\,8)$ and each schedule transforms $p \mapsto q$ in $d = 6$ moves. 
@@
 Since the Spearman distance is simply the sum of the absolute value of these displacements, at any point during during the execution of Algorithm~\ref{alg:sorting} we may obtain $F(\cdot, q)$ simply by having access to the sum of every entry in $\mathcal{A}$.
-At every step, there are only two degrees of freedom in Algorithm~\ref{alg:sorting}: the choice of $\sigma \in \mathcal{D}$ to move, and choice of the target index $j$ to move $\sigma$ to.
-If we fix a heuristic for the latter, then each $\sigma \in \mathcal{D}$ induces a set of possible valid moves satisfying~\eqref{eq:valid_move}, which we denote as $S_\mathcal{D}(\sigma)$.
-Now, observe that each permutation $s \in S_\mathcal{D}(\sigma)$ changes the displacement of every symbol in at most three different ways:
+At every step, there are only two degrees of freedom in Algorithm~\ref{alg:sorting}: the choice of index $k \in \mathcal{D}$ to move, and the choice of the target index $j$ to move $k$ to.
+If we fix a heuristic for the latter (for instance, $j$ chosen arbitrarily from an appropriate set, as in Algorithm~\ref{alg:sorting}), then each $k \in \mathcal{D}$ yields a corresponding cyclic permutation $m_{kj} \in S_m$~\eqref{eq:move_permutation} satisfying~\eqref{eq:valid_move}.
+Moreover, such a permutation changes the displacement of every index in at most three different ways:
 \[
-\mathcal{A}(s \circ p) = 
+\mathcal{A}(m_{kj} \circ p) = 
 \begin{cases} 
-	 \mathcal{A}(\sigma) \pm \lvert i - j \rvert & p^{-1}(\sigma) = i \\
-	 \mathcal{A}(\sigma) \pm 1 & i < p^{-1}(\sigma) \leq j \\
-	 \mathcal{A}(\sigma) & \text{ otherwise }
+	 \mathcal{A}(k) \pm \lvert i - j \rvert & p^{-1}(k) = i \\
+	 \mathcal{A}(k) \pm 1 & i < p^{-1}(k) \leq j \\
+	 \mathcal{A}(k) & \text{ otherwise }
 \end{cases}
@@
 %All other symbols are unchanged. 
-Thus, if we replace $\mathcal{A}$ with a data structure  supporting $O(\log m)$ access time to aggregate information and $O(\log m)$ modification time ability on $\lvert i - j \vert$ entries simultaneously, we could greedily choose the next permutation $s$ to minimize:
+Thus, if we replace $\mathcal{A}$ with a data structure  supporting $O(\log m)$ access time to aggregate information and $O(\log m)$ modification time over $\lvert i - j \vert$ entries simultaneously, we could greedily choose the next index $k$ to minimize:
 \begin{equation}\label{eq:greedy_step}
-	s_{\text{greedy}} = \argmin_{s \in S_{\mathcal{D}}(\sigma)} F(s\circ p, \iota)
+	k_{\text{greedy}} = \argmin_{k \in \mathcal{D}} F(m_{kj} \circ p, \iota)
 \end{equation}  
@@
  % Can actually only achieve O(d^2 log m) it seems
-Since single element modifications, removals, and insertions can all be achieved in $O(\log m)$ expected time with such a data structure, we conclude that equation~\eqref{eq:greedy_step} may solved in just $O(d^2 \log m)$ time.
+Since single element modifications, removals, and insertions can all be achieved in $O(\log m)$ expected time with such a data structure, we conclude that equation~\eqref{eq:greedy_step} may be solved in just $O(d^2 \log m)$ time.
 %Since single element prefix sum modifications can also be achieved $O(\log m)$ time with these data structures, the time to compute $F(\hat{s}_i, \hat{s}_{i+1})$ is bounded above by $O(\log m)$, 
@@
 
- In the first test, we fix a grid size of $9 \times 9$ and record the cumulative number of column operations needed to simulate persistence dynamically across 25 evenly-spaced time points using a variety of scheduling strategies.
+ In the first test, we fix a grid size of $9 \times 9$ and record the cumulative number of column operations needed to compute persistence dynamically across 25 evenly-spaced time points using a variety of scheduling strategies.
  The primary three strategies we test are the greedy approach from section~\ref{sec:proxy_objective}, the ``simple'' approach which uses upwards of $O(m)$ move permutations via selection sort, and a third strategy which interpolates between the two.
-To perform this interpolation, we use a parameter $\alpha \in [0,1]$ to choose $m - \alpha \cdot d$ random symbols to move using the same construction method outlined in section~\ref{sec:schedule_construction}.
+To perform this interpolation, we use a parameter $\alpha \in [0,1]$ to choose $m - \alpha \cdot d$ random simplices to move using the same construction method outlined in section~\ref{sec:schedule_construction}.
 When $\alpha = 0$, the strategy reduces to using selection sort to reindex $f \mapsto f'$ using a random ordering of simplices; otherwise, $\alpha = 1$ reduces to using a minimal sized schedule (with a random ordering of simplices).
@@
 	\caption{Performance comparison between various scheduling strategies.
-On the left, the cumulative column operations required to simulate the 1-parameter family is shown for varying schedule sizes ($d$) and strategies.
+On the left, the cumulative column operations required to compute the 1-parameter family is shown for varying schedule sizes ($d$) and strategies.
 On the right, both the size of the schedule and the data set $(m)$ are varied.
@@
 To do this, we generated 8 video data sets again of the expanding annulus outlined in section~\ref{sec:motivation}, each of increasing grid sizes of $5 \times 5$, $6 \times 6$, $\dots$, $12 \times 12$.
-For each data set, we simulate persistence over the duration of the video, again testing five evenly spaced settings of $\alpha \in [0,1]$---the results are shown in the right plot of Figure~\ref{fig:movie_perf}.
-On the vertical axis, we plot the total number of column operations needed to simulate the video again across 25 evenly-spaced time points \emph{as a ratio} of the data set size ($m$); we also show the regression curves one obtains for each setting of $\alpha$.
+For each data set, we compute persistence over the duration of the video, again testing five evenly spaced settings of $\alpha \in [0,1]$---the results are shown in the right plot of Figure~\ref{fig:movie_perf}.
+On the vertical axis, we plot the total number of column operations needed to compute persistence across 25 evenly-spaced time points \emph{as a ratio} of the data set size ($m$); we also show the regression curves one obtains for each setting of $\alpha$.
 As one can see from the Figure, the cost of using the greedy heuristic tends to increase sub-linearly as a function of the data set size, suggesting the move scheduling approach is indeed quite scalable.
@@
 We test the efficiency of computing the necessary information to generate these crocker stacks using a spatiotemporal data set to illustrate the applicability of our method.
-Specifically, we ran a \emph{flocking} simulation similar to the simulation run in~\cite{topaz2015topological} with $m = 20$ vertices moving around on the unit square equipped with periodic boundary conditions (i.e.
+Specifically, we ran a \emph{flocking} simulation similar to the one run in~\cite{topaz2015topological} with $m = 20$ vertices moving around on the unit square equipped with periodic boundary conditions (i.e.
 $S^1 \times S^1$).
@@
 Note again that vineyards requires generating many decompositions by design (in this case, $\approx 1.8M$).
-The standard algorithm \texttt{pHcol} and our move strategy were computed at 60 evenly spaced time points of the simulations.
+The standard algorithm \texttt{pHcol} and our move strategy were computed at 60 evenly spaced time points.
 As depicted in Figure~\ref{fig:boid_sim_results}, our move strategy is far more efficient than both vineyards and the naive \texttt{pHcol} strategies.
@@
 	\centering
-	\includegraphics[width=\textwidth, height=16.15em]{boid_sim_results.png}
+	\includegraphics[width=0.7\textwidth, height=12.15em]{boid_sim_results.png} % arxiv only
+%	\includegraphics[width=\textwidth, height=16.15em]{boid_sim_results.png}
 	\caption{ One the left, the cumulative number of column operations (log-scale) of the three baseline approaches tested.
@@
 Their research was primarily motivated by discerning whether there existed clear qualitative differences in the distributions of patches extracted from images of different modalities, such optical and range images.
-By mean-centering, contrast normalizing, and``whitening'' the data via the Discrete Cosine Transform (DCT) basis, they a convenient basis for $D$ may be obtained via an expansion of 8 certain non-constant eigenvectors, shown below:
+By mean-centering, contrast normalizing, and``whitening'' the data via the Discrete Cosine Transform (DCT), they show a convenient basis for $D$ may be obtained via an expansion of 8 certain non-constant eigenvectors, shown below:
 %\vspace{-0.2em}
@@
 Since these images are scale-invariant, the expansion of these basis vectors spans the 7-sphere, $S^7 \subset \mathbb{R}^8$.
-Using a voronoi cell decomposition of the data, their distribution analysis suggested that the majority of data points concentrated in a few high-density regions.
-
-After Lee et al published their work, Carlsson et al.~\cite{carlsson2008local} subsequently performed extensive experiments using persistent homology, wherein he found that the distribution of high-contrast $3 \times 3$ patches is actually well-approximated by a Klein bottle $\mathcal{M}$---around 60\% of the high-contrast patches from the still image data set lie within a small neighborhood around $\mathcal{M}$ which accounts for only 21\% of the 7-spheres volume.
+Using a Voronoi cell decomposition of the data, their distribution analysis suggested that the majority of data points concentrated in a few high-density regions.
+
+After Lee et al published their work, Carlsson et al.~\cite{carlsson2008local} subsequently performed extensive experiments using persistent homology, wherein he found that the distribution of high-contrast $3 \times 3$ patches is actually well-approximated by a Klein bottle $\mathcal{M}$---around 60\% of the high-contrast patches from the still image data set lie within a small neighborhood around $\mathcal{M}$ which accounts for only 21\% of the 7-sphere's volume.
 Along a similar vein in the sparse coding context, Perea~\cite{perea2014klein} introduced a dictionary learning framework for estimating the distribution of patches from texture images.
@@
 If one was not aware of the analysis done by~\cite{lee2003nonlinear, hateren_schaaf_1998, carlsson2008local, perea2014klein}, it is not immediately clear a priori that the Klein bottle model is  a good candidate for capturing the non-linearity of image patches.
-Indeed, armed with a refined topological intuition, Carlsson still needed to perform extensive sampling, preprocessing, and model fitting techniques in order reveal the underlying the topological space with persistent homology~\cite{carlsson2008local}.
+Indeed, armed with a refined topological intuition, Carlsson still needed to perform extensive sampling, preprocessing, and model fitting techniques in order to reveal the underlying the topological space with persistent homology~\cite{carlsson2008local}.
 One reason such preprocessing is needed is due to persistent homology's aforementioned instability with respect to strong outliers. 
@@
 Further inspection suggests one plausible candidate is the three-circle model $C_3$, which consists of three circles, two of which (say, $S_v$ and $S_h$) intersect the third (say, $S_\mathrm{lin}$) in exactly two points, but themselves do not intersect.
-Projecting the image data onto the first two basis vectors of leads to the projection shown in the top left of Figure~\ref{fig:patch_data_dgm}, of which 15 landmark points are also shown.
+Projecting the image data onto the first two basis vectors from the DCT shown above leads to the projection shown in the top left of Figure~\ref{fig:patch_data_dgm}, of which 15 landmark points are also shown.
 Observe the data are distributed well around three ``circles''---the outside circle capturing the rotation gradient of the image patches ($S_{\mathrm{lin}}$), and the other two capturing the vertical and horizontal gradients ($S_v$ and $S_h$, respectively).
@@
 In the more standard \emph{compressed} sparse matrix representations\footnote{
-By ``standard,'' we mean any of the common sparse representations used in scientific computing packages, like SciPy's sparse module (\href{https://docs.scipy.org/doc/scipy/reference/sparse.html} 
-))
+By ``standard,'' we mean any of the common sparse representations used in scientific computing packages, like SciPy's sparse module (\url{https://docs.scipy.org/doc/scipy/reference/sparse.html})
 }, permuting both the rows and columns generally takes at most $O(Z)$ time, where $Z$ is the number of non-zero entries, which can be quite expensive if the particular filtration has many cycles.
@@
 %\bibliographystyle{unsrt}
+\bibliographystyle{plain} % arxiv only
 \bibliography{move_schedules}
@@
 Fortunately, like move right, we can still separate the algorithm into a reduction and restoration phase---see Algorithm~\ref{alg:ml}.
-Moreover, since $R$ is reduced in all but one column by line 6 in~Algorithm~\ref{alg:ml}, we can still guarantee the number of column operations in $R$ will scale with $\lvert i - j \rvert$.
+Moreover, since $R$ is reduced in all but one column by line 6 in~Algorithm~\ref{alg:ml}, we can still guarantee the number of low entries to reduce in $R$ will be at most $\lvert i - j \rvert$.
 For a supplementary description of the move algorithm, see~\cite{busaryev2010tracking}.
-\begin{algorithm}[!htb]
-	\caption{Move Right Algorithm}\label{alg:mr}
-	\setstretch{1.15}
-    \begin{algorithmic}[1]
-        \Function{RestoreRight}{$R$, $V$, $\mathbb{I} = \{I_1, I_2, \dots, I_s \}$}
-            \State $(\, d_{low}, \, d_R, \, d_V \,) \gets ( \, \mathrm{low}_R(I_1), \, \mathrm{col}_R(I_1), \, \mathrm{col}_V(I_1) \, )$
-        	\For{$k$ \textbf{in} $I_2, \dots, I_s$} 
-        		\State $( \, d_{low}', \, d_R', \, d_V' \, ) \gets ( \, \mathrm{low}_R(k), \, \mathrm{col}_R(k), \, \mathrm{col}_V(k) \, )$
-        		\State $( \, \mathrm{col}_R(k), \, \mathrm{col}_V(k) \, ) \mathrel{+}= (\, d_R, \, d_V \,) $
-        		\If {$d_{low}' < d_{low}$}
-        			\State $(\, d_{low}, \, d_R, \, d_V \,) \gets (\, d_{low}', \, d_R', \, d_V' \, )$
-        		\EndIf
-        	\EndFor 
-        	\State \Return $(\, R, \, V, \, d_R, \, d_V \, )$
-        \EndFunction
-    \end{algorithmic}
-    \begin{algorithmic}[1]
-        \Function{MoveRight}{$R$, $V$, $i$, $j$} 
-            \State $\mathbb{I} =$ columns satisfying $V[i, i:j] \neq 0$ %\Comment{$O(m)$}
-            \State $\mathbb{J} = $ columns satisfying $\mathrm{low}_R \in [i:j]$ and $\mathrm{row}_R(i) \neq 0$ %\Comment{$O(m)$}
-            \State $(\, R, \, V, d_R, d_V \,) \gets $ \Call{RestoreRight}{$R$, $V$, $\mathbb{I}$} 
-            %\Comment{Restores $p$-simplices} \Comment{$O(\lvert \mathbb{I} \rvert)$}
-            \State $(\, R, \, V \,) \gets $ \Call{RestoreRight}{$R$, $V$, $\mathbb{J}$}
-            	%\Comment{Restores $(p+1)$-simplices}
-            	%\Comment{$O(m \lvert \mathbb{I} \rvert)$}
-            \State $(\, R, \, V\, ) \gets (\, P R P^T, \, P V P^T\,)$  % \Comment{$P$ is a permutation given by (i,j)}
-            \State $(\, \mathrm{col}_R(j), \, \mathrm{col}_V(j) \, ) \gets (\, P d_R, \, P d_V \,)$
-            \State \Return $(\, R, \, V\,)$
-        \EndFunction
-    \end{algorithmic}
-\end{algorithm}
-
-\begin{algorithm}[!htb]
-	\caption{Move Left Algorithm}\label{alg:ml}
-	\setstretch{1.15}
-    \begin{algorithmic}[1]
-        \Function{RestoreLeft}{$R$, $V$, $\mathbb{K} = \{k_1, k_2, \dots, k_s \}$}
-            % \State $\mathrm{low}_K \gets \{ \mathrm{low}_R(k_1), \mathrm{low}_R(k_2), \dots, \mathrm{low}_R(k_s) \}$
-            \State $(l, r) \gets $ indices $l, r \in \mathbb{K}$ satisfying $l < r$, $\mathrm{low}_R(l) = \mathrm{low}_R(r)$ maximal 
-            \While{$\mathrm{low}_R(l) \neq 0$ \textbf{and} $\mathrm{low}_R(r) \neq 0$}
-            	\State $(\mathrm{col}_R(r), \mathrm{col}_V(r)) \mathrel{+}= (\mathrm{col}_R(l), \mathrm{col}_V(l))$
-            	\State $\mathbb{K} \gets \mathbb{K} \setminus l$
-            	\State $(l, r) \gets $ indices $l, r \in \mathbb{K}$ satisfying $l < r$, $\mathrm{low}_R(l) = \mathrm{low}_R(r)$ maximal
-            \EndWhile
-        	\State \Return $(\, R, \, V \, )$
-        \EndFunction
-    \end{algorithmic}
-
-    \begin{algorithmic}[1]
-        \Function{MoveLeft}{$R$, $V$, $i$, $j$} 
-        	\State $\mathbb{I} \gets \emptyset$
-            % \State $(\mathbb{K}, d_R, d_V) \gets (\emptyset, \mathrm{col}_R(i), \mathrm{col}_V(i))$
-            \While{$V(k, i) \neq 0 $\textbf{ for }$ k = \mathrm{low}_V(i)$\textbf{ where }$j \leq k < i$}
-            	\State $(\mathrm{col}_R(i), \mathrm{col}_V(i)) \mathrel{+}= (\mathrm{col}_R(k), \mathrm{col}_V(k))$
-            	\State $\mathbb{I} \gets \mathbb{I} \cup k + 1 $
-            \EndWhile
-            \State $(R, V) \gets (P R P^T, P V P^T)$
-            \State $\mathbb{J} = $ columns satisfying $\mathrm{low}_R \in [i:j]$ and $\mathrm{row}_R(i) \neq 0$ %\Comment{$O(m)$}
-
-            \State $(R, V) \gets$ \Call{RestoreLeft}{$R, V, \mathbb{I}$}
-            \State $(R, V) \gets$ \Call{RestoreLeft}{$R, V, \mathbb{J}$}
-            % \State $(\mathrm{col}_R(i), \mathrm{col}_V(i)) \gets (P d_R, P d_V )$
-			\State \Return $(\, R, \, V\,)$
-        \EndFunction
-    \end{algorithmic}
-
-\end{algorithm}
+
 We recall an important claim given in~\cite{busaryev2010tracking} on the effect that move operations have on the status of simplices in the pairing.
@@
   \begin{algorithm}[h!]
-	\caption{Sorting algorithm}\label{alg:sorting}
+	\caption{Schedule construction algorithm}\label{alg:sorting}
 	\setstretch{1.05}
     \begin{algorithmic}[1]
-    	% \Require Fixed $p' \in S_m$, $\mathcal{L} = \mathrm{LIS}(p')$ 
-    	\Function{LCS-Sort}{$p$, $q$}
+    	\Require Fixed $p,q \in S_m$, heuristic $h$, set data structure $\mathcal{T}$ from section~\ref{sec:schedule_construction}
+    	\Function{Schedule}{$p$, $q$, $h = \texttt{greedy}$} \Comment{See~\ref{sec:greedy} for heuristic discussion}
     		\State $\bar{p} \gets q^{-1} \circ p$
     		\State $\mathcal{L} \gets \mathrm{LCS}(p,q) = \mathrm{LIS}(\bar{p})$ \Comment{$O(m\log \log m)$}
-    		\State $(\, \mathcal{S}, \, \mathcal{D}, \, \mathcal{T} \, )\gets (\, \emptyset, \, \bar{p} \setminus \mathcal{L}, \, \mathcal{L} \,)$
+    		\State $(\, \mathcal{S}, \, \mathcal{D} \,  )\gets (\, \emptyset, \, [m] \setminus \mathcal{L} \,)$
     		\While{$\mathcal{D}$\text{ is not empty }}
-    		    \State $\sigma \gets \text{arbitrary element in }\mathcal{D}$
+    		    \State Select an element $k \in \mathcal{D}$ using heuristic $h$ \Comment{e.g. equation~\eqref{eq:greedy_step}}   		        	
+%    		    \State $\mathcal{T}_{\mathrm{pred}}(k) \gets \max \{ \ell \in \mathcal{L} \mid \ell \leq k \}$ 
+				\State $\mathrm{k}_{\text{pred}} \gets \max \{ \ell \in \mathcal{L} \mid \ell \leq k \}$\Comment{$O(\log \log m)$ using $\mathcal{T}$}
+				\State $\mathrm{k}_{\text{succ}} \gets \min \{ \ell \in \mathcal{L} \mid \ell \geq k \}$  \Comment{$O(\log \log m)$ using $\mathcal{T}$}
+    		    %\text{arbitrary element in }
     		    %\State $\tau, \rho \gets \mathcal{T}_p(\sigma), \mathcal{T}_n(\sigma)$
     		    %\State $i, i_p, i_n \gets \mathcal{I}_p(\sigma), \, \mathcal{I}_p(T_{\mathrm{pred}}(\sigma)), \, \mathcal{I}_p(T_{\mathrm{succ}}(\sigma))$
-    		    \State $(\,i, \,i_p, \, i_n \,) \gets ( \, \bar{p}^{-1}(\sigma), \, \bar{p}^{-1}(\mathcal{T}_{\mathrm{pred}}(\sigma)), \, \bar{p}^{-1}(\mathcal{T}_{\mathrm{succ}}(\sigma)) \, )$\Comment{$O(\log \log m)$}
+    		    \State $(\,i, \,i_p, \, i_n \,) \gets ( \, \bar{p}^{-1}(k), \, \bar{p}^{-1}(\mathrm{k}_{\text{pred}}), \, \bar{p}^{-1}(\mathrm{k}_{\text{succ}} ) \, )$ \Comment{$O(1)$}
     		    %    		    \If{ $p_{\sigma} < p_{\tau}$ }
@@
 %				\IfThenElse{$i < i_p$}{$[\, i_p, \, i_n)$}{$(\, i_p, \, i_n]$}
+				\State $j \gets \text{arbitrary } j \in [\, i_p, \, i_n)$ \textbf{if} $i < i_p$ \textbf{else} $j \in (\, i_p, \, i_n]$ \Comment{$O(1)$}
 				% $\text{Pick } j \in [\, i_p, \, i_n)$ \If{ $i < i_p$ } \Else $(\, i_p, \, i_n]$ \EndIf
-    		    \If{ $i < i_p$ } 
-    		    	\State $j \gets \text{arbitrary element in } [\, i_p, \, i_n)$
-    		    \Else{ $i_n < i$ }
-	    		    \State $j \gets \text{arbitrary element in } (\, i_p, \, i_n]$ 
-    		    \EndIf
-    		    \State $(\, \mathcal{S}, \mathcal{D}, \mathcal{T} \,) \gets (\, \mathcal{S} \cup (i, \, j), \, \mathcal{D} \setminus \sigma, \, \mathcal{T} \cup \sigma \, )$\Comment{$O(\log \log m)$}
+%    		    \If{ $i < i_p$ } 
+%    		    	\State $j \gets \text{arbitrary element in } [\, i_p, \, i_n)$
+%    		    \Else{ $i_n < i$ }
+%	    		    \State $j \gets \text{arbitrary element in } (\, i_p, \, i_n]$ 
+%    		    \EndIf
+    		    \State $(\, \mathcal{S}, \mathcal{D}, \mathcal{L} \,) \gets (\, \mathcal{S} \cup (i, \, j), \, \mathcal{D} \setminus k, \, \mathcal{L} \cup k \, )$\Comment{$O(\log \log m)$}
     		    % \State $p, p^{-1} \gets m_{ij} \circ p, p^{-1} \circ m_{ij}^{-1}$
-    		    \State $\bar{p}^{-1} \gets \bar{p}^{-1} \circ m_{ij}^{-1}$ \Comment{$O(m)$}
+    		    \State $\bar{p}^{-1} \gets \bar{p}^{-1} \circ m_{ij}^{-1}$ where $m_{ij}$ is given by~\eqref{eq:move_permutation}\Comment{$O(m)$}
     		 \EndWhile
@@
 Suppose $\mathcal{L}$ has been computed from $\bar{p}$.
-Since $\mathcal{L}$ is strictly increasing, the only symbols left to permute are in $\mathcal{L} \setminus \bar{p}$, which we denote with $\mathcal{D}$.
-After choosing any symbol $\sigma \in \mathcal{D}$, one then applies a cyclic permutation to $\bar{p}$ that moves $\sigma$ into any position that increases the size of $\mathcal{L}$.
-To do this efficiently, we maintain a data structure $\mathcal{T}$ which enables us to query the successor and predecessor of any given symbol $s \in \bar{p}$ \emph{in} $\mathcal{L}$.
-We also require a data structure to query the position of a given element $\sigma \in \bar{p}$, which for now we simply use the inverse permutation $\bar{p}$ (though a more efficient representation can be used based off of symbol \emph{displacements}, see section~\ref{sec:proxy_objective}.
-After the symbol is inserted into $\mathcal{L}$, we update $\bar{p}$, its inverse permutations $\bar{p}^{-1}$, $\mathcal{D}$ and $\mathcal{T}$ prior to the next move.
+Since $\mathcal{L}$ is strictly increasing, the only elements left to permute are in $\mathcal{L} \setminus \bar{p}$, which we denote with $\mathcal{D}$.
+After choosing any $\sigma \in \mathcal{D}$, one then applies a cyclic permutation to $\bar{p}$ that moves $\sigma$ into any position that increases the size of $\mathcal{L}$.
+To do this efficiently, we maintain a data structure $\mathcal{T}$ which enables us to query the successor and predecessor of any given $s \in \bar{p}$ \emph{in} $\mathcal{L}$.
+We also require a data structure to query the position of a given element $\sigma \in \bar{p}$, which for now we simply use the inverse permutation $\bar{p}$ (though a more efficient representation can be used based off of \emph{displacements}, see section~\ref{sec:proxy_objective}.
+After $\sigma$ is inserted into $\mathcal{L}$, we update $\bar{p}$, its inverse permutations $\bar{p}^{-1}$, $\mathcal{D}$ and $\mathcal{T}$ prior to the next move.
 The final set of permutations which sort $p \mapsto q$ (or equivalently, $\bar{p} \to \iota$ are stored in an array $\mathcal{S}$, which is then returned for further use.

### 2023-10-10 move_schedules.tex
Edits to streamline section 3. Added reviewer point-by-point responses to most of the reviewer feedbakc so far

@@
 	\orgdiv{Khoury College of Computer Sciences}, 
-	\orgname{Northeastern University}
-}
+	\orgname{Northeastern University}}
 
@@
 	\orgdiv{Department of Mathematics and Khoury College of Computer Sciences},
-	\orgname{Northeastern University}
+	\orgname{Northeastern University}}
 % \orgaddress{\street{360 Huntington Ave}, \city{Boston}, \postcode{02115}, \state{MA}}
-}
+%}
 
@@
 We propose a coarser strategy for maintaining the decomposition over a 1-parameter family of filtrations.
-By reduction to a particular longest common subsequence problem, we show that the minimal number $d$ of updates to the decomposition can be found in $O(m \log \log m)$ time and $O(m)$ space, and that the corresponding sequence of permutations---which we call a \emph{schedule}---can be constructed in $O(d m \log m)$ time.
+By reduction to a particular longest common subsequence problem, we show that the minimal number of decomposition updates $d$ can be found in $O(m \log \log m)$ time and $O(m)$ space, and that the corresponding sequence of permutations---which we call a \emph{schedule}---can be constructed in $O(d m \log m)$ time.
 We also show that, in expectation, the storage needed to employ this strategy is actually sublinear in $m$. 
@@
 \section{Introduction} 
-\subsection{Overview}\label{sec:overview} 
+%\subsection{Overview}\label{sec:overview} 
 Given a triangulable topological space equipped with a tame continuous function, persistent homology captures the changes in topology across the sublevel sets of the space, and encodes them in a persistence diagram.
@@
 As persistence-related computations tend to exhibit high algorithmic complexity---essentially cubic\footnote{For finite fields, it is known that the persistence computation reduces to the PLU factorization problem, which takes $O(m^\omega)$ where $\omega \approx 2.373$ is the matrix multiplication constant.} in the size of the underlying filtration~\cite{morozov2005persistence}---their adoption to dynamic settings poses a challenging computational problem.
-With state of the art tools, there is no recourse when faced with a time-varying complex containing millions of simplices across thousands of snapshots in time.
-Moreover, acquiring such a capability has far-reaching consequences: methods that vectorize persistence diagrams for machine learning purposes all immediately become computationally viable tools in dynamic settings.
+Currently, there is no recourse when faced with a time-varying complex containing millions of simplices across thousands of snapshots in time.
+%---the state of the art is just to compute the persistence at each time point independently. % ask jose what he thinks
+Acquiring such a capability has far-reaching consequences: methods that vectorize persistence diagrams for machine learning purposes all immediately become computationally viable tools in the dynamic setting.
 Such persistence summaries include adaptive template functions~\cite{polanco2019adaptive}, persistence images~\cite{adams2017persistence}, and $\alpha$-smoothed Betti curves~\cite{ulmer2019topological}.
@@
 If one is interested in understanding how the persistent homology of a continuous function changes over time, then this algorithm is sufficient, for homological critical points can only occur when the filtration order changes.
-The vineyards algorithm is efficient asymptotically: if there are $d$ time-points where the filtration order changes, then  vineyards   takes $O(m^3 + md)$ time; one initial $O(m^3)$-time reduction at time $t_0$ followed by one $O(m)$ operation to update the decomposition at the remaining time points $(t_1, t_2, \dots, t_d)$.
-When $d >> m$, the initial reduction cost is amortized by the cost of maintaining the decomposition, implying each diagram produced by vineyards takes just linear time per time point.
+Moreover, the vineyards algorithm is efficient asymptotically: if there are $d$ time-points where the filtration order changes, then  \emph{vineyards}  takes $O(m^3 + md)$ time; one initial $O(m^3)$-time reduction at time $t_0$ followed by one $O(m)$ operation to update the decomposition at the remaining time points $(t_1, t_2, \dots, t_d)$.
+When $d >> m$, the initial reduction cost is amortized by the cost of maintaining the decomposition, implying each diagram produced takes just linear time per time point to obtain.
 % approach is far more efficient than the $O(dm^3)$  ``naive'' strategy of computing the diagrams at every time point independently.
 
-Despite its theoretical efficiency, vineyards is often not the method of choice in practical settings. 
+Despite its theoretical efficiency, \emph{vineyards} is often not the method of choice in practical settings. 
 While there is an increasingly rich ecosystem of software packages offering variations of the standard reduction algorithm (e.g.
 Ripser, PHAT, Dionysus, etc.
-see~\cite{otter2017roadmap} for an overview), implementations of the vineyards algorithm are relatively uncommon.\footnote{Dionysus 1 does have an implementation of vineyards, however the algorithm was never ported to version 2.
-Other major packages, such as GUDHI and PHAT, do not have vineyards implementations.}
+see~\cite{otter2017roadmap} for an overview), implementations of the vineyards algorithm are relatively uncommon.\footnote{Dionysus 1 does have an implementation of \emph{vineyards}, however the algorithm was never ported to version 2.
+Other major packages, such as GUDHI and PHAT, do not have \emph{vineyards} implementations.}
 The reason for this disparity is perhaps explained by Lesnick and Wright~\cite{lesnick2015interactive}: ``While an update to an $RU$ decomposition involving few transpositions is very fast in practice...
@@
 By exploiting the ``donor'' concept introduced in~\cite{busaryev2010tracking}, we are able to make a tradeoff between the number of times the decomposition is restored to a valid state and the granularity of the decomposition repair step, reducing the total number of column operations needed to apply an arbitrary permutation to the filtration.
-This trade off, paired with a fast greedy heuristic explained in section~\ref{sec:proxy_objective}, yields an algorithm that can update a $R = DV$ decomposition more efficiently than vineyards in coarse time-varying contexts, making dynamic persistence more computationally tractable for a wider class of use-cases.
+This trade off, paired with a fast greedy heuristic explained in section~\ref{sec:proxy_objective}, yields an algorithm that can update a $R = DV$ decomposition more efficiently than \emph{vineyards} in coarse time-varying contexts, making dynamic persistence more computationally tractable for a wider class of use-cases.
 The source code containing both the algorithm we propose and the experiments performed in Section~\ref{sec:results} is open source and available online.\footnote{ For all accompanying software and materials, see: \url{https://github.com/peekxc/move_schedules}}
@@
 If  zeroth homology  is the only dimension of interest, computing and updating both the persistence and rank information  is greatly simplified.
-For example, if the edge are available a-priori, obtaining a tree representation fully characterizing the connectivity of the underlying space (also known as the \emph{incremental connectivity} problem) takes just $O(\alpha(n) n)$ time using the disjoint-set data structure, where $\alpha(n)$ is the extremely slow-growing inverse Ackermann function.
+For example, if the edges of the graph are in filtered order a-priori, obtaining a tree representation fully characterizing the connectivity of the underlying space (also known as the \emph{incremental connectivity} problem) takes just $O(\alpha(n) n)$ time using the disjoint-set data structure, where $\alpha(n)$ is the extremely slow-growing inverse Ackermann function.
 Adapting this approach to the time-varying setting, Oesterling et al.~\cite{oesterling2015computing} give an algorithm that maintains a \emph{merge tree} with $e$ edges in $O(e)$ time per-update.
@@
 
-We use this data set as a baseline for comparing vineyards and the standard reduction algorithm $\texttt{pHcol}$ (Algorithm \ref{alg:reduce}).
+We use this data set as a baseline for comparing \emph{vineyards} and the standard reduction algorithm $\texttt{pHcol}$ (Algorithm \ref{alg:reduce}).
 Suppose a practitioner wanted to know the major homological changes a time-varying filtration encounters over time.
@@
 In this setting, as shown in the figure, the vineyards approach is indeed far more efficient than naively applying the reduction algorithm independently at all time points.
-However, when the discretization of the time domain is coarse enough, the naive approach actually performs less column operations than vineyards, while still capturing the main events.
-
-The existence of a time discretization that is more efficient to compute than continually updating the decomposition, indicates that the vineyards framework must incur some overhead (in terms of column operations) to maintain the underlying decomposition, even when the pairing function determining the persistence diagram is unchanged.
+However, when the discretization of the time domain is coarse enough, the naive approach actually performs less column operations than \emph{vineyards}, while still capturing the main events.
+
+The existence of a time discretization that is more efficient to compute than continually updating the decomposition indicates that the vineyards framework must incur some overhead (in terms of column operations) to maintain the underlying decomposition, even when the pairing function determining the persistence diagram is unchanged.
 Indeed, as shown by the case where $n = 10$, applying \texttt{pHcol} independently between  relatively ``close'' filtrations is substantially more efficient than iteratively updating the decomposition.
  Moreover, any optimizations to the reduction algorithm (e.g. clearing~\cite{chen2011persistent}) would only increase this disparity.
-Since persistence has found many applications in dynamic contexts~\cite{topaz2015topological, xian2020capturing, lesnick2015interactive, kim2020spatiotemporal}, a more efficient alternative to vineyards is clearly needed.
+Since persistence has found many applications in dynamic contexts~\cite{topaz2015topological, xian2020capturing, lesnick2015interactive, kim2020spatiotemporal}, a more efficient alternative to \emph{vineyards} is clearly needed.
  \\
@@
  where  $s_k = (i_k, j_k)$ for $k=1,\ldots, d$,  denotes a particular type of cyclic permutation (see section~\ref{sec:moves_setting}).
-If $i_k < j_k$ for all $s_k \in \mathcal{S}$, our first result extends~\cite{busaryev2010tracking} by showing that~\eqref{eq:rv_seq} can be computed  using $O(\kappa)$ column operations, where:
+If $i_k < j_k$ for all $s_k \in \mathcal{S}$, our first result extends~\cite{busaryev2010tracking} by showing that~\eqref{eq:rv_seq} can be computed  using $O(\nu)$ column operations, where:
 \begin{equation}
-	\quad \kappa = \sum\limits_{k = 1}^d \, \lvert \mathbb{I}_{k}\rvert + \lvert \mathbb{J}_{k}\rvert 
+	\quad \nu = \sum\limits_{k = 1}^d \, \lvert \mathbb{I}_{k}\rvert + \lvert \mathbb{J}_{k}\rvert 
 \end{equation}
@@
 Without optimizations, like   clearing   or implicit matrix reduction, the standard  algorithm is very inefficient.
-Nonetheless, it serves as the foundation of most persistent homology implementations, and its invariants are necessary before introducing both vineyards in section~\ref{sec:vineyards} and our move schedules  in section~\ref{sec:move_schedules}.
+Nonetheless, it serves as the foundation of most persistent homology implementations, and its invariants are necessary before introducing both \emph{vineyards} in section~\ref{sec:vineyards} and our move schedules in section~\ref{sec:move_schedules}.
 
@@
 As $t$ varies in $[0,1]$, the points in $\mathrm{dgm}_p(f_t)$ trace curves in $\mathbb{R}^3$ which, by the stability of persistence, will be continuous if $F$ is continuous and the $f_t$'s are tame.
-Cohen-Steiner et al.~\cite{cohen2007stability} referred to these curves as \emph{vines}, a collection of which forms as \emph{vineyard}---the geometric analogy meant to act as a guidepost for practitioners seeking to understand the evolution of topological structure over time.
-
-The original purpose of the vineyards algorithm, as described in~\cite{cohen2006vines}, was to compute a continuous 1-parameter family of persistence diagrams over a time-varying filtration, detecting homological critical events along the way.
+Cohen-Steiner et al.~\cite{cohen2007stability} referred to these curves as \emph{vines}, a collection of which forms as \emph{vineyard}---the geometric analogy is meant to act as a guidepost for practitioners seeking to understand the evolution of topological structure over time.
+
+The original purpose of \emph{vineyards}, as described in~\cite{cohen2006vines}, was to compute a continuous 1-parameter family of persistence diagrams over a time-varying filtration, detecting homological critical events along the way.
 As homological critical events only occur when the filtration order changes, detecting all such events may be reduced to computing valid decompositions at   time points interleaving all changes in the filtration order.
@@
 \noindent 
-\textbf{Example 2.2} To illustrate the basic principles of vineyards, we re-use the running example introduced in the previous section.
+\textbf{Example 2.2} To illustrate the basic principles of \emph{vineyards}, we re-use the running example introduced in the previous section.
 Below, we illustrate the case of exchanging simplices $a$ and $b$ in the filtration order, and restoring $RV$ to a valid decomposition.
@@
 
- The time complexity of vineyards is determined entirely by the complexity of performing a single adjacent transposition.
-Formally, since column operations are the largest complexity operations needed and each column can have potentially $O(m)$ entries, the complexity of vineyards is $O(m)$ per transposition. 
+ The time complexity of \emph{vineyards} is determined entirely by the complexity of performing a single adjacent transposition.
+Formally, since column operations are the largest complexity operations needed and each column can have potentially $O(m)$ entries, the complexity of \emph{vineyards} is $O(m)$ per transposition. 
 Inspection of the individual cases of the algorithm from~\cite{cohen2006vines} shows that any single transposition requires at most two such operations on both $R$ and $V$.
@@
 A move operation $\mathrm{Move}(i,j)$ is a set of rules for maintaining a valid decomposition under the permutation $P$ that moves a simplex $\sigma_i$ at position $i$ to position $j$.
-If $j = i \pm 1$, this operation is an adjacent transposition, and in this sense moves generalizes vineyards.
-However, the move framework presented by Busaryev is actually distinct in that it exhibits several attractive qualities not inherited by the vineyards approach that warrants further study.
+If $j = i \pm 1$, this operation is an adjacent transposition, and in this sense \emph{moves} generalizes \emph{vineyards}.
+However, the move framework presented by Busaryev is actually distinct in that it exhibits several attractive qualities not inherited by the \emph{vineyards} approach that warrants further study.
 
@@
 \\
-\noindent \textbf{Example 2.3:} We re-use the running example from sections~\ref{sec:reduction} and~\ref{sec:vineyards} to illustrate moves.
+\noindent \textbf{Example 2.3:} We re-use the running example from sections~\ref{sec:reduction} and~\ref{sec:vineyards} to illustrate \emph{moves}.
 The donor columns of $R$ and $V$ are denoted as $d_R$ and $d_V$, respectively.
@@
 \noindent 
-Note that the equivalent permutation using vineyards requires $4$ column operations on both $R_1$ and $V_1$, respectively, whereas a single move operation accomplishes using only $2$ column operations per matrix.
+Note that the equivalent permutation using \emph{vineyards} requires $4$ column operations on both $R_1$ and $V_1$, respectively, whereas a single move operation accomplishes using only $2$ column operations per matrix.
 The pseudo-code for \emph{MoveRight} is given in Algorithm~\ref{alg:mr} and for \emph{MoveLeft} in Algorithm~\ref{alg:ml}.
@@
 \end{proposition}
-\noindent Though similar to vineyards, move operations confer additional advantages:
+\noindent Though similar to \emph{vineyards}, move operations confer additional advantages:
 \\	
@@
 Prior to applying any permutation $P$ to the decomposition, it is necessary to remove non-zero entries in $V$ which render $P^TVP$ non-upper triangular, to maintain invariant I2.
-Using vineyards, one must consistently perform $\lvert i - j \rvert - 1$ non-zero status queries interleaved between repairing column operations.
+Using \emph{vineyards}, one must consistently perform $\lvert i - j \rvert - 1$ non-zero status queries interleaved between repairing column operations.
 A move operation, on the other hand, groups these status queries into a single pass prior to performing any modifying operations.
 
-Property M2 implies that the decomposition is not fully maintained during the execution of \textit{RestoreRight} and \textit{RestoreLeft} below, which starkly contrasts the vineyards algorithm. 
-In this way, we interpret move operations as making a tradeoff in granularity: whereas a sequence of adjacent transpositions $(i, i{+}1), (i{+}1, i{+}2), \dots, (j{-}1, j)$ generates $\lvert i - j \rvert $ valid decompositions in vineyards, an equivalent move operation $\mathrm{Move}(i,j)$ generates only one.
+Property M2 implies that the decomposition is not fully maintained during the execution of \textit{RestoreRight} and \textit{RestoreLeft} below, which starkly contrasts the \emph{vineyards} algorithm.
+In this way, we interpret move operations as making a tradeoff in granularity: whereas a sequence of adjacent transpositions $(i, i{+}1), (i{+}1, i{+}2), \dots, (j{-}1, j)$ generates $\lvert i - j \rvert $ valid decompositions in \emph{vineyards}, an equivalent move operation $\mathrm{Move}(i,j)$ generates only one.
 Indeed, Property M3 directly follows from this fact, as one may simply move each simplex $\sigma \in K$ into its new order $f'(\sigma)$ via insertion sort.
-Note that the number of valid decompositions produced by vineyards is bounded above by $O(m^2)$, as each pair of simplices $\sigma_i, \sigma_j \in K$ may switch its relative ordering at most once during the interpolation from $f$ to $f'$.
-
-There is another aspect of the move algorithm that has a distinct benefit compared to vineyards.
-As shown by example 2.3, move updates can be cheaper than vineyards in terms of column operations.
+Note that the number of valid decompositions produced by \emph{vineyards} is bounded above by $O(m^2)$, as each pair of simplices $\sigma_i, \sigma_j \in K$ may switch its relative ordering at most once during the interpolation from $f$ to $f'$.
+
+As shown by example 2.3, \emph{moves} can be cheaper than \emph{vineyards} in terms of column operations.
 However, it is not immediately clear that this is always the case upon inspection of~\ref{alg:mr}, as the usage of a donor column seemingly implies that many $O(m)$ copy operations need to be performed.
@@
 \begin{proof} \normalsize
-First, consider executing the vineyards algorithm with a given pair $(i, j)$.
+First, consider executing the \emph{vineyards} algorithm with a given pair $(i, j)$.
 As there are at most $2$ column operations, any contiguous sequence of transpositions $(i,i+1), (i+1,i+2), \ldots (j-1,j)$ induces at most $2(\lvert i - j \rvert)$ column operations in both $R$ and $V$, giving a total of $4(\lvert i - j \rvert)$ column operations.
@@
 %\subsection{Proposed approach}
-Let us begin with a brief overview of the pipeline, which we outline in Algorithm~\ref{alg:schedule} below.
-As before, we assume as input  a discrete $1$-parameter family $\mathcal{F} = (f_1, f_2, \dots, f_n)$ 
-of filtrations $f_i : K \rightarrow [m]$ of a simplicial complex $K$ with $\lvert K \rvert =m$, and the goal being to compute the  persistence diagrams of each $(K,f_i)$.
-Fix   bijections $\rho_i : [m] \to [m]$ so that 
-$f_{i+1} = \rho_i \circ f_i$, or equivalently,  permutations of the index set $[m]$. For each $\rho_i$ we let 
-$q = (\rho_i(1), \rho_i(2), \ldots, \rho_i(m) )$
-and compute a longest increasing subsequence $\mathrm{LIS}(q)$;  this subsequence  is used to recover a longest common subsequence (LCS) between $f_i$ and $ f_{i+1} $, which we denote later with $\mathrm{LCS}(f_i, f_{i+1})$.
-We pass $q$ and $\mathrm{LIS}(q)$ to our greedy scheduling algorithm, which returns as output an ordered set of move permutations  $\mathcal{S}$ of minimum size, which we call a \emph{schedule} (see Definition~\ref{def:schedule}).
-Note one need not explicitly store the family of filtrations nor the schedules between them---Algorithm~\ref{alg:schedule} may be easily modified to be completely \emph{online}, keeping at most two filtrations and one decomposition in memory at any given time.
+We begin with a brief overview of the pipeline to compute the persistence diagrams of a discrete $1$-parameter family $(f_1, f_2, \dots, f_n)$ of filtrations. 
+%$f_i : K \rightarrow [m]$ of a simplicial complex $K$. 
+Without loss of generality, assume each filtration $f_i : K \rightarrow [m]$ is a simplexwise filtration of a fixed simplicial complex $K$ with $\lvert K \rvert =m$, and let $(f, f')$ denote any pair of such filtrations.
+%or equivalently, permutations of the index set $[m]$. For each $\rho_i$ we let  .
+%a discrete $1$-parameter family $\mathcal{F} = (f_1, f_2, \dots, f_n)$ of filtrations $f_i : K \rightarrow [m]$ of a simplicial complex $K$ with $\lvert K \rvert =m$, and the goal being to compute the  persistence diagrams of each $(K,f_i)$.
+Our strategy to efficiently obtain a valid $RV$ decomposition of the filtration $(K,f')$ from a given decomposition of $(K, f)$ is to decompose a fixed bijection $\rho : f \to f'$ satisfying $f' = \rho \circ f$ into a \emph{schedule} of updates:
+\begin{definition}[Schedule]\label{def:schedule}
+Given a pair of filtrations $(K, f), (K,f')$ 
+%and a fixed bijection $\rho : f \to f'$, let $\rho^\ast : [m] \to [m]$ denote its induced bijection on $S_m$ 
+and $R = DV$ the initial decomposition of $(K, f)$, a \emph{schedule}  $\mathcal{S} = (s_1, s_2, \dots, s_d )$ is a sequence of permutations  satisfying: 
+	\begin{equation}\label{eq:rv_seq_schedule}
+   		R = D_0 V_0 \overset{s_1}{\to} D_1 V_{1} \overset{s_2}{\to} \dots \overset{s_d}{\to} D_d V_{d} = R'
+   	\end{equation}
+   	where, for each $i \in [d]$, $R_i = D_i V_i$ is a valid decomposition respecting invariants~\ref{inv:decomposition}, and $R'$ is a valid decomposition for $(K,f')$.
+\end{definition}
+\noindent
+To produce this sequence of permutations $\mathcal{S} = (s_1, \dots, s_d)$ from $\rho$, our approach is as follows: let $\rho^\ast : [m] \to [m]$ denote its \emph{induced}
+%\footnote{Recall that an induced representation...} 
+bijection on $S_m$, and define $q = (\rho^\ast(1), \rho^\ast(2), \ldots, \rho^\ast(m) )$. 
+We compute a longest increasing subsequence $\mathrm{LIS}(q)$, and then use this subsequence to recover a longest common subsequence (LCS) between $f$ and $f'$, which we denote later with $\mathrm{LCS}(f, f')$.
+We pass $q$, $\mathrm{LIS}(q)$, and a ``greedy'' heuristic for picking moves $h$ to our scheduling algorithm (to be defined later), which returns as output an ordered set of move permutations  $\mathcal{S}$ of minimum size satisfying~\eqref{eq:rv_seq_schedule}---a \emph{schedule} for the pair $(f, f')$.
+These sequence of steps is outlined in Algorithm~\ref{alg:schedule} below.
+%Note one need not explicitly store the family of filtrations nor the schedules between them---Algorithm~\ref{alg:schedule} may be easily modified to be completely \emph{online}, keeping at most two filtrations and one decomposition in memory at any given time.
 % Not all valid schedules are created equal: some incur a much larger number of column operations, even if they have a minimal number of moves; we address this issue in Section~\ref{sec:schedule_cost}.
@@
     	\Procedure{MoveSchedule}{$\mathcal{F} = ( f_0, f_1 )$, $R$, $V$} 
-%    		\State Fix $\rho \in S_m$ such that $f_1 = \rho \circ f_0$ \Comment{$O(m)$} 
-%    		\State $\mathrm{lis}_{\rho} \gets \mathrm{LIS}(\rho)$ \Comment{$O(m \log \log m)$}
+    		\State Fix $\rho \in S_m$ such that $f_1 = \rho \circ f_0$ \Comment{$O(m)$} 
+    		\State $q \gets  (\rho^\ast(1), \rho^\ast(2), \ldots, \rho^\ast(m) )$ \Comment{$O(m)$}
+    		\State $\mathrm{lis}_{q} \gets \mathrm{LIS}(q)$ \Comment{$O(m \log \log m)$}
+%    		\State $\bar{p} \gets q^{-1} \circ p$
+%    		\State $\mathcal{L} \gets \mathrm{LCS}(p,q) = \mathrm{LIS}(\bar{p})$ \Comment{$O(m\log \log m)$}
 			\State $h \gets \texttt{greedy}$ \Comment{Heuristic for picking moves}
-    		\State $\mathcal{S} \gets \mathrm{Schedule}(f_0, f_1, h)$ \Comment{$O(d^2\log m)$}
+    		\State $\mathcal{S} \gets \mathrm{Schedule}(q, \mathrm{lis}_{q}, h)$ \Comment{$O(d^2\log m)$}
 %	    	\State $(R, V) \gets$ \Call{Reduction}{$D = \partial (K,f_1)$}
@@
 
-In the following subsections, we investigate how to  leverage the increased flexibility of the move framework, beginning with the hypothesis formed experimentally from section~\ref{sec:motivation} that vineyards exhibits extraneous overhead maintaining the decomposition.
-In theory, decreasing the number of times the decomposition is restored to a valid state ought to reduce this overhead, motivating the question: can one simultaneously minimize the number of times the decomposition is restored to a valid state while retaining an efficient update scheme?
+% REMOVING
+%In the following subsections, we investigate how to  leverage the increased flexibility of the move framework, beginning with the hypothesis formed experimentally from section~\ref{sec:motivation} that \emph{vineyards} exhibits extraneous overhead maintaining the decomposition.
+%In theory, decreasing the number of times the decomposition is restored to a valid state ought to reduce this overhead, motivating the question: can one simultaneously minimize the number of times the decomposition is restored to a valid state while retaining an efficient update scheme?
+Though using the LCS between a pair of permutations induced by simplexwise filtrations is a relatively intuitive way of producing a schedule of move-type decomposition updates, it is not immediately clear whether such an algorithm has any computational benefits compared to \emph{vineyards}.
+Indeed, since \emph{moves} is a generalization of \emph{vineyards}, several important questions arise when considering practical aspects of how to implement Algorithm~\ref{alg:schedule}, such as e.g.
+how to pick a ``good'' heuristic $h$, how expensive can the $\text{Schedule}$ algorithm be, or how  $\lvert \mathcal{S}\rvert$ scales with respect to the size of the complex $K$.
+We address these issues in the following sections.
+
 %Next we will show how the analysis of the 
@@
 \subsection{Continuous setting}\label{sec:continuous_setting}
-In the original vineyards setting, a given homotopy  $F : K \times [0,1] \to \mathbb{R}$ continuously interpolating between $(K, f)$ and $(K, f')$ is discretized into a set of critical events that alter the filtration order.
-As $F$ determines the number of distinct filtrations encountered during the deformation from $f$ to $f'$, a natural question is whether such an interpolation can be modified so as to minimize the number of times the decomposition is restored to a valid state.
-Towards explaining the phenomenon exhibited in Figure~\ref{fig:vineyards}, in what follows we analyze a particular class of interpolation schemes in order to establish an upper bound on this quantity.
+In the \emph{vineyards} setting, a given homotopy  $F : K \times [0,1] \to \mathbb{R}$ continuously interpolating between $(K, f)$ and $(K, f')$ is discretized into a set of critical events that alter the filtration order.
+As $F$ determines the number of distinct filtrations encountered during the deformation from $f$ to $f'$, a natural question is whether such an interpolation can be modified so as to minimize $\lvert \mathcal{S}\rvert$---the number of times the decomposition is restored to a valid state.
+Towards explaining the phenomenon exhibited in Figure~\ref{fig:vineyards}, we begin by analyzing a particular class of interpolation schemes in order to establish an upper bound on this quantity.
 
@@
 $f,f' : K \rightarrow [m]$ whose function $t\mapsto F(\sigma, t)$ is continuous and satisfies $f(\sigma) = F(\sigma, 0)$ and $f'(\sigma) = F(\sigma, 1)$ for every  $\sigma \in K$.
-Note this family includes the straight-line homotopy $F(\sigma, t) = (1 - t) f(\sigma) + t f'(\sigma)$,  studied in the original vineyards paper~\cite{cohen2006vines}. 
+Note this family includes the straight-line homotopy $F(\sigma, t) = (1 - t) f(\sigma) + t f'(\sigma)$,  studied in the original \emph{vineyards} paper~\cite{cohen2006vines}. 
 If we assume that each pair of curves $\big(t,  F(\cdot, t)\big) \subset [0,1]\times \mathbb{R}$  intersect in at most one point---at which they cross---the continuity and genericity assumptions on $F$ imply that for $ \sigma,\mu \in K$ distinct,  the curves   $t \mapsto F(\sigma ,t)$ and $t \mapsto F(\mu,t)$ intersect if and only if $f(\sigma) > f(\mu)$ and $f'(\sigma)  < f'(\mu)$, or $f(\sigma) < f(\mu)$ and $f'(\sigma)  > f'(\mu)$.
@@
 $\rho_i = (\ell_i \;\; \ell_i +1 )$ for $1 \leq \ell_i < m$. In other words, $\rho_i$ is an adjacent transposition for each $i =1,\ldots, k$.
-The ordered sequence of adjacent transpositions  
-$S_F = (\rho_1, \rho_2, \ldots, \rho_k)$ can thus be thought of as a \emph{schedule} of permutations to be applied to the initial decomposition $R = DV$ of $(K,f)$, in order to   interpolate from $f$ to $f'$ via $F$.
-Going beyond transpositions, the idea of a schedule can be generalized as follows:
-\begin{definition}[Schedule]\label{def:schedule}
-Given a pair of filtrations $(K, f), (K,f')$ 
-%and a fixed bijection $\rho : f \to f'$, let $\rho^\ast : [m] \to [m]$ denote its induced bijection on $S_m$ 
-and $R = DV$ the initial decomposition of $(K, f)$, a \emph{schedule}  $\mathcal{S} = (s_1, s_2, \dots, s_d )$ is a sequence of permutations  satisfying: 
-	\begin{equation}\label{eq:rv_seq_schedule}
-   		R = D_0 V_0 \overset{s_1}{\to} D_1 V_{1} \overset{s_2}{\to} \dots \overset{s_d}{\to} D_d V_{d} = R'
-   	\end{equation}
-   	where, for each $i \in [d]$, $R_i = D_i V_i$ is a valid decomposition respecting invariants~\ref{inv:decomposition}, and $R'$ is a valid decomposition for $(K,f')$.
-\end{definition}
-\noindent Observe the size of the schedule $S_F$ defined from the homotopy $F$ above is exactly $K_\tau(f,f')$.
+%The ordered sequence of adjacent transpositions  $S_F = (\rho_1, \rho_2, \ldots, \rho_k)$ can thus be thought of as a \emph{schedule} of permutations to be applied to the initial decomposition $R = DV$ of $(K,f)$, in order to   interpolate from $f$ to $f'$ via $F$.
+%Going beyond transpositions, the idea of a schedule can be generalized as follows:
+Observe the size of the ordered sequence of adjacent transpositions  $S_F = (\rho_1, \rho_2, \ldots, \rho_k)$ defined from the homotopy $F$ above is exactly $K_\tau(f,f')$.
 On the positive side, the reduction of schedule planning to crossing detection implies the former can be   solved optimally in output-sensitive $O(m \log m + k)$ time by several algorithms~\cite{boissonnat2000efficient}, where $k$ is the output-sensitive term and $m$ is the number of simplices in the filtration(s).
 On  the negative side, $k = K_\tau(f,f')$ scales in size to $\sim O(m^2)$ in the worst case, achieved when $f' = - f$.
-As mentioned in~\ref{sec:vineyards}, this quadratic scaling induces a number of issues in the practical implementations of the vineyards algorithm
+As mentioned in~\ref{sec:vineyards}, this quadratic scaling induces a number of issues in the practical implementations of the \emph{vineyards} algorithm
 
@@
 Contrasting the continuous-time setting, if we discard the use of a homotopy interpolation and allow move operations in any order, we obtain a trivial upper bound of $O(m)$ on the schedule size:  simply move each simplex in $K$ from  its position in the filtration given by $f$ to the position given by $f'$---which we call the \emph{naive strategy}.
-However, in losing the interpolation interpretation it is no longer clear the $O(m)$ bound is tight, and the ``intermediate'' filtrations need no longer respect the face poset of the underlying complex $K$.
+In particular, in losing the interpolation interpretation it is no longer clear the $O(m)$ bound is tight.
+Indeed, the ``intermediate'' filtrations need no longer even respect the face poset of the underlying complex $K$.
 In this section, we investigate these issues from a combinatorial perspective. 
@@
 \end{equation}
-Similarly, in the context of edit operations, observe that a move operation can be interpreted as a paired delete-and-insert operation, i.e.
+Observe that a move operation can be interpreted as a paired delete-and-insert operation, i.e.
 $m_{ij} = (\mathrm{ins}_j \circ \mathrm{del}_i)$, where $\mathrm{del}_i$ denotes the operation that deletes the character at position $i$  and $\mathrm{ins}_j$ the operation that inserts the same character at position $j$.
@@
 %Like the \emph{LCS} distance, 
-In general, one can compute the LCS itself in $O(mn)$  with dynamic programming.
-One might hope computing the size $\mathrm{d}_{\mathrm{lcs}}(p,q)$ alone exhibits a lower complexity, however there is substantial evidence that for general string inputs the complexity cannot be much lower than quadratic~\cite{abboud2015tight}.
-However, in our setting, the total orders of any pair of simplexwise filtrations $(K, f), (K, f')$ of the same underlying complex $K$ may be thought of a permutations in $S_m$---the corresponding edit distance $d$ then reduces to the \emph{permutation edit distance} problem.
 With this insight in mind, we obtain the following bound on the minimum size of a sorting (i.e. schedule) using moves and the complexity of computing it. 
@@
 where we use $\mathrm{LCS}(f, f')$ to denote the LCS of the permutations of $K$ induced by $f$ and $f'$.
-Moreover, $\lvert S^\ast \rvert = d$ can be determined in $O(m \log \log m)$ time. 
 \end{proposition}
@@
 Notice that $\iota$ is strictly increasing and that any common subsequence $p'$ has with $\iota$ must also be strictly increasing.
-Thus, the problem of computing $\mathrm{LCS}(p, q)$ reduces to the problem of computing the \emph{longest increasing subsequence} (LIS) of $p'$, which can done in $O( m \log \log m)$ time using van Emde Boas trees~\cite{bespamyatnikh2000enumerating}.
 The optimality of $d$ follows from the optimality of the well-studied LCS problem~\cite{kumar1987linear}.
@@
 \noindent
-Establishing a connection between the permutation edit distance and move scheduling allows us to exploit the combinatorial structure that comes from the developed theory on both LCS's and LIS's, which are both well-studied objects.
-We record a single corollary to demonstrate this fact.
+For any pair of general string inputs of sizes $n$ and $m$, respectively, the LCS between them is computable in $O(mn)$ with dynamic programming, and there is substantial evidence that the complexity cannot be much lower than this~\cite{abboud2015tight}.
+In our setting, however, observe the total orders of any pair of simplexwise filtrations $(K, f), (K, f')$ of the same underlying complex $K$ may be thought of a permutations in $S_m$, in which case $d_{\mathrm{LCS}}$ reduces further to the \emph{permutation edit distance} problem.
+%Establishing a connection between the permutation edit distance and move scheduling allows us to exploit the combinatorial structure that comes from the developed theory on both LCS's and LIS's, which are both well-studied objects.
+This special type of edit distance has additional structure to it, which we demonstrate below.  
+\begin{corollary}\label{cor:lis_schedule_complexity}
+Let $(K, f), (K, f')$ denote two filtrations of size $\lvert K \rvert = m$, and let $S^*$ denote schedule of minimal size reindexing $f \mapsto f'$.
+Then $\lvert S^\ast \rvert = d$ can be determined in $O(m \log \log m)$ time.
+\end{corollary}
+\begin{proof}
+	By the same reduction from Proposition~\ref{prop:schedule_size}, the problem of computing $\mathrm{LCS}(f, f')$ reduces to the problem of computing the \emph{longest increasing subsequence} (LIS) of a particular permutation $p' \in S_m$, which can done in $O( m \log \log m)$ time using van Emde Boas trees~\cite{bespamyatnikh2000enumerating}
+\end{proof}
+\noindent 
+Reduced complexity is not the only immediate benefit from Proposition~\ref{prop:schedule_size}; by the same reduction to the LIS problem, we obtain the worst-case bounds on $\mathcal{S}$ in expectation.
 \begin{corollary}\label{cor:expectation}
-If $(K, f), (K,f')$ are random filtrations of a common complex $K$ of size $m$, then the expected size of longest common subsequence of simplices between $f,f'$ is no larger than $m - \sqrt{m}$, with probability $1$ as $m \to \infty$.
+If $(K, f), (K,f')$ are random filtrations of a common complex $K$ of size $m$, then the expected size of longest common subsequence $\mathrm{LCS}(f, f')$ between $f,f'$ is no larger than $m - \sqrt{m}$, with probability $1$ as $m \to \infty$.
 \end{corollary}
@@
    	can be determined in $O(m \log \log m)$ time and $O(m)$ space, where $R' = D_d V_d$ denotes a valid decomposition of $(K, f')$. 
-   	Moreover, if $i_k < j_k$ for all $k \in [d]$, then computing~\eqref{eq:rv_seq_theorem} requires $O(\kappa)$ column operations, where:	  
-   	$$  \quad \kappa = \sum\limits_{k = 1}^d \, (\lvert \mathbb{I}_{k} \rvert + \lvert \mathbb{J}_{k} \rvert) $$
+   	Moreover, if $i_k < j_k$ for all $k \in [d]$, then computing~\eqref{eq:rv_seq_theorem} requires $O(\\nu)$ column operations, where:	  
+   	$$  \quad \nu = \sum\limits_{k = 1}^d \, (\lvert \mathbb{I}_{k} \rvert + \lvert \mathbb{J}_{k} \rvert) $$
 %	$$  \quad \kappa = \sum\limits_{k = 1}^d \, \lvert \mathbb{I}_{k}\rvert + \lvert \mathbb{J}_{k}\rvert $$
 	where the quantities $\lvert \mathbb{I}_{k} \rvert, \lvert \mathbb{J}_{k} \rvert$ of the intermediate $R_k, V_k$ are given in Proposition~\ref{prop:move_st}.
-Note that since $\kappa$ depends on the sparsity of the intermediate entries $V_1, V_2, \dots, V_d$ and $R_1, R_2, \dots, R_d$, the bound $O(\kappa)$ is output-sensitive.
+Note that since $\nu$ depends on the sparsity of the intermediate entries $V_1, V_2, \dots, V_d$ and $R_1, R_2, \dots, R_d$, the bound $O(\nu)$ is output-sensitive.
 %	and where $\lvert \mathbb{I}_{k} \rvert$ denotes the number of non-zero entries $V_k[i_k, i_k:j_k]$, $\lvert \mathbb{J}_{k}\rvert$ the number of columns satisfying $\mathrm{low}_{R_k} \in [i_k:j_k]$ and $\mathrm{row}_{R_k}(i_k) \neq 0$, and $d = m - \lvert \mathrm{LCS}(K, K') \rvert$.
@@
 	Proposition~\ref{prop:schedule_size} yields the necessary conditions for constructing $\mathcal{S}$  with optimal size $d$ in $O(m \log \log m)$ time and $O(m)$.
-The definition of $\kappa$ follows directly from Algorithm~\ref{alg:mr}.
+The definition of $\nu$ follows directly from Algorithm~\ref{alg:mr}.
 \end{proof}
@@
 \subsection{Constructing schedules}\label{sec:schedule_construction}
-It is clear from the proof in Proposition~\ref{prop:schedule_size} that one may compute the LCS between two permutations $p, q \in S_m$ via the LIS of a single permutation $p^\ast$, and that computation may be carried out in $O(m \log \log m)$ time.
-It is not immediately clear, however, how to obtain a sorting $p \mapsto q$ from a given $\mathcal{L} = \mathrm{LIS}(p')$ in an efficient way.
+While it is clear from the proof in Proposition~\ref{prop:schedule_size} that one may compute the LCS between two permutations $p, q \in S_m$ in $O(m \log \log m)$ time, it is not immediately clear how to obtain a sorting $p \mapsto q$ from a given $\mathcal{L} = \mathrm{LCS}(p,q)$ in an efficient way.
 We outline below a simple procedure which constructs such a sorting $\mathcal{S} = (s_1, \dots, s_d)$ in $O(dm\log m)$ time and $O(m)$ space, or $O(m \log m)$ time and $O(m)$ space per update in the online setting.
@@
 The algorithm outlined in section~\ref{sec:schedule_construction} is a sufficient for generating move schedules of minimal cardinality: any schedule of moves $S$ sorting $f \mapsto f'$ above is guaranteed to have size $\lvert S \rvert = m - \lvert \mathrm{LCS}(f, f') \rvert$, and the reduction to the permutation edit distance problem ensures this size is optimal.
-However, as with the vineyards algorithm, certain pairs of simplices cost more to exchange depending on whether they are critical pairs in the sense described in~\cite{cohen2006vines}, resulting in a large variability in the cost of randomly generated schedules.
+However, as with the \emph{vineyards} algorithm, certain pairs of simplices cost more to exchange depending on whether they are critical pairs in the sense described in~\cite{cohen2006vines}, resulting in a large variability in the cost of randomly generated schedules.
 This variability is undesirable in practice: we would like to generate a schedule which not only small in size, but is also efficient in terms of its required column operations.
@@
 where $\lvert \mathbb{I}\rvert + \lvert \mathbb{J}\rvert$ are the quantities from Proposition~\ref{prop:move_st}.
-One advantages to the moves framework is that the cost of a single move on a given $R = D V$ decomposition can be determined efficiently prior to any column operations---no more than $O(m)$ time with row-oriented sparse matrices.
+One advantage of the \emph{moves} framework is that the cost of a single move on a given $R = D V$ decomposition can be determined efficiently prior to any column operations---no more than $O(m)$ time with row-oriented sparse matrices.
 It is natural to consider whether a greedy-type solution which chooses the minimal cost choice at every step yields an efficient strategy.
@@
 %The Kendall-$\tau$ distance yields an upper bound on the number of column reductions needed to perform any schedule. 
-A similar $\ell_1$-type distance for measuring the disarrangement between permutations is the \emph{Spearman distance}, defined as:
+A similar $\ell_1$-type distance for measuring the disarrangement between permutations is the \emph{Spearman footrule distance}, defined as:
 \begin{equation}\label{eq:spearman_dist}
@@
 The results are summarized in the left graph on Figure~\ref{fig:movie_perf}, wherein the mean schedule cost of the random strategies are depicted by solid lines.
-To capture the variation in performance for the random sampling approach, we run 10 independent iterations and shade the upper and lower bounds of the schedule costs.
+To capture the variation in performance, we run 10 independent iterations and shade the upper and lower bounds of the schedule costs.
  As one can see from the figure, while using less move operations (lower $\alpha$) does progressively reduce column operations, constructing random schedules of minimal size is no more competitive than the simple selection sort strategy.
@@
 There are many challenges to characterizing topological behavior in dynamic settings.
-One approach is to trace out the curves constituting a continuous family of persistence diagrams in $\mathbb{R}^3$---the vineyards approach---however this visualization can be cumbersome to work with as there are potentially many such vines tangled together, making topological critical events with low persistence difficult to detect.
-Moreover, the vineyards visualization does not admit a natural simplification utilizing the stability properties of persistence, as individual vines are not stable: if two vines move near each other and then pull apart without touching, then a pairing in their corresponding persistence diagrams may cross under a small perturbation, signaling the presence of an erroneous topological critical event~\cite{topaz2015topological, xian2020capturing}.
+One approach is to trace out the curves constituting a continuous family of persistence diagrams in $\mathbb{R}^3$---the \emph{vineyards} approach---however this visualization can be cumbersome to work with as there are potentially many such vines tangled together, making topological critical events with low persistence difficult to detect.
+Moreover, the \emph{vineyards} visualization does not admit a natural simplification utilizing the stability properties of persistence, as individual vines are not stable: if two vines move near each other and then pull apart without touching, then a pairing in their corresponding persistence diagrams may cross under a small perturbation, signaling the presence of an erroneous topological critical event~\cite{topaz2015topological, xian2020capturing}.
 
@@
 Indeed, it has been shown that crocker stacks and stacked persistence diagrams (i.e.
-vineyards) are equivalent to each other in the sense that either one contains the information needed to reconstruct the other~\cite{xian2020capturing}.
+\emph{vineyards}) are equivalent to each other in the sense that either one contains the information needed to reconstruct the other~\cite{xian2020capturing}.
 Thus, computing crocker stacks reduces to computing the persistence of a (time-varying) family of filtrations.
@@
 
-Given this time-evolving data set, we computed the persistence diagram of the Rips filtration up to $\epsilon = 0.30$ at 60 evenly spaced time points using three approaches: the standard algorithm \texttt{pHcol} applied naïvely at each of the 60 time steps, the vineyards algorithm applied to (linear) homotopy connecting filtrations adjacent in time, and our approach using moves.
+Given this time-evolving data set, we computed the persistence diagram of the Rips filtration up to $\epsilon = 0.30$ at 60 evenly spaced time points using three approaches: the standard algorithm \texttt{pHcol} applied naïvely at each of the 60 time steps, the \emph{vineyards} algorithm applied to (linear) homotopy connecting filtrations adjacent in time, and our approach using \emph{moves}.
 The cumulative number of $O(m)$ column operations executed by three different approaches.
-Note again that vineyards requires generating many decompositions by design (in this case, $\approx 1.8M$).
+Note again that \emph{vineyards} requires generating many decompositions by design (in this case, $\approx 1.8M$).
 The standard algorithm \texttt{pHcol} and our move strategy were computed at 60 evenly spaced time points.
-As depicted in Figure~\ref{fig:boid_sim_results}, our move strategy is far more efficient than both vineyards and the naive \texttt{pHcol} strategies.
+As depicted in Figure~\ref{fig:boid_sim_results}, our \emph{moves} strategy is far more efficient than both \emph{vineyards} and the naive \texttt{pHcol} strategies.
 
@@
 
-Despite its elegance, there are significant computational barriers prohibiting the 2-parameter persistence algorithm outlined from being practical. 
-An analysis from~\cite{lesnick2015interactive} (using vineyards) shows the barcodes template computation requires on the order of $O(m^3 \kappa + m \kappa^2 \log \kappa)$ elementary operations and $O(m \kappa^2)$ storage.
-Since the number of 2-cells in $\mathcal{A}(M)$ is on the order $O(\kappa^2)$, and $\kappa$ itself is on the order of $O(m^2)$ in the worst case, the worse-case complexity of the barcode templates computation $O(m^5)$---this is both the highest complexity and most time-intensive sub-procedure the RIVET software~\cite{rivet} depends on.
-Despite this significant complexity barrier, there is room for optimism: in practice, the external stability result from~\cite{landi2014rank} justifies the use of a coarsening procedure which approximates the module $M$ with a smaller module $M'$ via a grid-like reduction, enabling practitioners to restrict the size of $\kappa$ to a relatively small constant.
+Despite its elegance, there are significant computational barriers prohibiting the 2-parameter persistence algorithm from being practical. 
+An analysis from~\cite{lesnick2015interactive} (using \emph{vineyards}) shows the barcodes template computation requires on the order of $O(m^3 \kappa + m \kappa^2 \log \kappa)$ elementary operations and $O(m \kappa^2)$ storage, where $\kappa$ is a coarseness parameter.
+Since the number of 2-cells in $\mathcal{A}(M)$ is on the order $O(\kappa^2)$, and $\kappa$ itself is on the order of $O(m^2)$ in the worst case, the scaling of the barcode template computation may approach $\approx O(m^5)$---this is both the highest complexity and most time-intensive sub-procedure the RIVET software~\cite{rivet} depends on.
+Despite this significant complexity barrier, in practice the external stability result from~\cite{landi2014rank} justifies the use of a grid-like reduction procedure which approximates the module $M$ with a smaller module $M'$, enabling practitioners to restrict the size of $\kappa$ to a relatively small constant.
 This in-turn dramatically reduces the size of $\mathcal{A}(M)$ and thus the number of barcode templates to compute. 
@@
 %Intuitively, similar barcode templates should be easier to compute in sequence than independently. 
-Indeed, as algorithm~\ref{alg:schedule} was designed for precisely such a computation, 2-parameter persistence is prototypical of the class of methods that stand to benefit from move scheduling.
+Indeed, as algorithm~\ref{alg:schedule} was designed for precisely such a computation, 2-parameter persistence is prototypical of the class of methods that stand to benefit from \emph{moves}.
 
@@
 
-After Lee et al published their work, Carlsson et al.~\cite{carlsson2008local} subsequently performed extensive experiments using persistent homology, wherein he found that the distribution of high-contrast $3 \times 3$ patches is actually well-approximated by a Klein bottle $\mathcal{M}$---around 60\% of the high-contrast patches from the still image data set lie within a small neighborhood around $\mathcal{M}$ which accounts for only 21\% of the 7-sphere's volume.
-Along a similar vein in the sparse coding context, Perea~\cite{perea2014klein} introduced a dictionary learning framework for estimating the distribution of patches from texture images.
+In follow-up work, Carlsson et al.~\cite{carlsson2008local} found---using persistent homology---that the distribution of high-contrast $3 \times 3$ patches is actually well-approximated by a Klein bottle $\mathcal{M}$---around 60\% of the high-contrast patches from the still image data set lie within a small neighborhood around $\mathcal{M}$ accounting for only 21\% of the 7-sphere's volume.
+Along a similar vein, Perea~\cite{perea2014klein} established a dictionary learning framework for efficiently estimating the distribution of patches from texture images, prompting applications for persistent homology in sparse coding contexts.
 
@@
 
-Consider the (coarsened) fibered barcode computed from a standard Rips / codensity bifiltration on a representative sample of the image data from~\cite{hateren_schaaf_1998}, shown in Figure~\ref{fig:patch_data_dgm}.
+To demonstrate the benefit of 2-parameter persistence on the patch data, consider the (coarsened) fibered barcode computed from a standard Rips / codensity bifiltration on a representative sample of the image data from~\cite{hateren_schaaf_1998}, shown in Figure~\ref{fig:patch_data_dgm}.
 From the bigraded Betti number and the dimension function, one finds that a large area of dimension function is constant (highlighted as the blue portion in the middle of Figure~\ref{fig:patch_data_dgm}), wherein the first Betti number is 5.
@@
 % This can be further verified empirically by e.g. using the iterative, meanshift-like procedure described in~\cite{carlsson2008local} to fit a parameterization of the model in question to the data.
+
 The degree to which multi-parameter persistence simplifies this exploratory phase cannot be understated: we believe multi-parameter persistence has a larger role to play in manifold learning. 
@@
 \subsubsection{Accelerating 2D persistence}\label{sec:empirical_klein}
-As we have outlined the computational theory of 2-parameter persistence and elucidated its relevance to our proposed move scheduling approach, we now demonstrate the efficiency of scheduling using the same high-contrast patch data set studied in~\cite{lee2003nonlinear} by evaluating the performance of various methods at computing the fibered barcode invariant via the parameterization from~\ref{sec:fibered_barcode_reparam}.
+%As we have outlined the computational theory of 2-parameter persistence and elucidated its relevance to our proposed move scheduling approach, 
+Having outlined the computational theory of 2-parameter persistence, we now demonstrate the efficiency of \emph{moves} using the same high-contrast patch data set studied in~\cite{lee2003nonlinear} by evaluating the performance of various methods at computing the fibered barcode invariant via the parameterization from~\ref{sec:fibered_barcode_reparam}.
 
 Due to the aforementioned high complexity of the fibered barcode computation, we begin by working with a subset of the image patch data $\mathcal{X}$. 
-In particular, we combine the use of furthest-point sampling and proportionate allocation (stratified) sampling to sample landmarks $X \subset \mathcal{X}$ distributed within $n = 25$ strata.
+We combine the use of furthest-point sampling and proportionate allocation (stratified) sampling to sample landmarks $X \subset \mathcal{X}$ distributed within $n = 25$ strata.
 Each strata consists of the $(1/n)$-thick level set given the $k$-nearest neighbor density estimator $\rho_{15}$ with $k = 15$.
@@
 From this data set, we construct a Rips-(co)density bifiltration using $\rho_{15}$ equipped with the geodesic metric computed over the same k-nearest neighbor graph on $X$. 
-Finally, we record the number of column reductions needed to compute the fibered barcode at a variety of levels of coarsening using $\mathtt{pHcol}$, vineyards, and our moves approach.
+Finally, we record the number of column reductions needed to compute the fibered barcode at a variety of levels of coarsening using $\mathtt{pHcol}$, \emph{vineyards}, and our \emph{moves} approach.
 The results are summarized in Table~\ref{tab:barcode_templates}.
-We also record the number of 2-cells in $\mathcal{A}(M)$ and the number of permutations applied throughout the encountered along the traversal of the dual graph for both vineyards and moves, denoted in the table as $d_K$ and $d_{\mathrm{LCS}}$, respectively.
+We also record the number of 2-cells in $\mathcal{A}(M)$ and the number of permutations applied encountered along the traversal of the dual graph for both \emph{vineyards} and \emph{moves}, denoted in the table as $d_K$ and $d_{\mathrm{LCS}}$, respectively.
 %Specifically, we begin by extracting a representative subset of the set $X \subset \mathcal{X}$ of patches representing the top 30\% densest patches of the space of all patches $\mathcal{X}$, where the notion of density is approximated with a k-nearest neighbor (knn) density estimator $\rho$.
@@
 
-As shown on the table, when the coarsening $\kappa$ is small enough, we're able to achieve a significant reduction in the number of total column operations needed to compute $\mathcal{T}$ compared to both vineyards and $\mathtt{pHcol}$.
+As shown on the table, when the coarsening $\kappa$ is small enough, we're able to achieve a significant reduction in the number of total column operations needed to compute $\mathcal{T}$ compared to both \emph{vineyards} and $\mathtt{pHcol}$.
 %though it's worth noting that permutations are constant time for vineyards while they require at most $O(m)$ elementary operations for moves.
-This is further reinforced by the observation that, when there is a high degree of coarsening, vineyards is particularly inefficient and moves requires only about 3x less column operations that naively computing $\mathcal{T}$ independently.
-As the coarsening becomes more refined and more 2-cells are added to $\mathcal{A}(M)$, however, vineyards quickly becomes a much more viable option compared to $\mathtt{pHcol}$---as predicted---though even at the highest coarsening we tested the gain in efficiency is relatively small.
-In contrast, our proposed moves approach scales quite well with this refinement, requiring about $12\%$  and $\approx 5\%$ of the number of column operations as vineyards and $\mathrm{pHcol}$, respectively.
+This is further reinforced by the observation that \emph{vineyards} is particularly inefficient when then 1-parameter family is coarse.
+Indeed, \emph{moves} requires about 3x less column operations than naively computing $\mathcal{T}$ independently.
+However, note that as the coarsening becomes more refined and more 2-cells are added to $\mathcal{A}(M)$, \emph{vineyards} becomes a more viable option compared to $\mathtt{pHcol}$---as the asymptotics suggests---though even at the highest coarsening we tested the gain in efficiency is relatively small.
+In contrast, \emph{moves} scales quite well with this refinement, requiring about $12\%$  and $\approx 5\%$ of the number of column operations as \emph{vineyards} and $\mathtt{pHcol}$, respectively.
 
@@
 
-There are many possible applications of our work beyond the ones discussed in section~\ref{sec:results}. As mentioned in section~\ref{sec:related_work}, examples include: 
-\begin{enumerate}
-	\item Accelerating PH featurization methods for time-varying systems 
-	\item Optimization procedures involving persistence diagrams
-	\item Detecting homological critical points in time-varying filtrations
-\end{enumerate}
-Indeed, we see our approach as potentially useful to any situation where the structure of interest is a parameterized family of persistence diagrams.
+There are many possible applications of our work beyond the ones discussed in section~\ref{sec:results}, such as e.g.
+accelerating PH featurization methods or detecting homological critical points in dynamic settings.
+%\begin{enumerate}
+%	\item Accelerating PH featurization methods for time-varying systems 
+%	\item Optimization procedures involving persistence diagrams
+%	\item Detecting homological critical points in time-varying filtrations
+%\end{enumerate}
+Indeed, we see our approach as potentially useful to any situation where the structure of interest can be cast as a parameterized family of persistence diagrams.
 Areas of particular interest include time-series analysis and dynamic metric spaces~\cite{kim2020spatiotemporal}.
@@
 Moreover, if the filtration $(K, f)$ shares little similarity to the ``target'' filtration $(L,f')$, then the overhead of reducing the simplices from $L \setminus K$ appended to the decomposition derived from $K$ may be large enough to motivate simply computing the decomposition at $L$ independently, especially if parallel processors are available.
-Our approach is primarily useful if the filtrations in the parameterized family is ``nearby'' in the combinatorial sense.
+Our approach is primarily useful if the filtrations in the parameterized family are ``nearby'' in the combinatorial sense.
 
@@
 
-We recall an important claim given in~\cite{busaryev2010tracking} on the effect that move operations have on the status of simplices in the pairing.
+%We recall an important claim given in~\cite{busaryev2010tracking} on the effect that move operations have on the status of simplices in the pairing.
 Recall from section~\ref{sec:background} that simplices which create new homology classes are called \emph{creators} and simplices that destroy homology classes are called \emph{destroyers}.
-%\begin{proposition}[Locality of moves~\cite{busaryev2010tracking}]
-%	Other than the simplex $\sigma_i$ being moved, $\mathrm{MoveRight}(i,j)$ cannot change a destroyer to a creator, $\mathrm{MoveLeft}(i,j)$ cannot change a creator to destroyer, and neither move operation changes the status of any simplex in $K$ outside of the interval $[i,j]$.
-%\end{proposition}
-\noindent The effect of the movement on intermediate simplices depends on the direction of the movement.
+%%\begin{proposition}[Locality of moves~\cite{busaryev2010tracking}]
+%%	Other than the simplex $\sigma_i$ being moved, $\mathrm{MoveRight}(i,j)$ cannot change a destroyer to a creator, $\mathrm{MoveLeft}(i,j)$ cannot change a creator to destroyer, and neither move operation changes the status of any simplex in $K$ outside of the interval $[i,j]$.
+%%\end{proposition}
+%\noindent The effect of the movement on intermediate simplices depends on the direction of the movement.
 If $i < j$ (respectively, $j < i$), all simplices at positions $k \in [i+1\;:\;j] $ are shifted down (respectively, up) by $1$.
@@
     \begin{algorithmic}[1]
-    	\Require Fixed $p,q \in S_m$, heuristic $h$, set data structure $\mathcal{T}$ from section~\ref{sec:schedule_construction}
-    	\Function{Schedule}{$p$, $q$, $h = \texttt{greedy}$} \Comment{See~\ref{sec:greedy} for heuristic discussion}
-    		\State $\bar{p} \gets q^{-1} \circ p$
-    		\State $\mathcal{L} \gets \mathrm{LCS}(p,q) = \mathrm{LIS}(\bar{p})$ \Comment{$O(m\log \log m)$}
+    	\Require Fixed $\bar{p} \in S_m$, LIS $\mathcal{L}$ of $\bar{p}$, and heuristic $h$
+    	\Ensure $[m] = s_d \circ s_{d-1} \circ \dots \circ s_1 \circ \bar{p}$ for output sequence $\mathcal{S} = (s_i)_{i=1}^d$,
+    	%set data structure $\mathcal{T}$ from section~\ref{sec:schedule_construction}
+    	\Function{Schedule}{$\bar{p}$, $\mathcal{L}$, $h = \texttt{greedy}$} \Comment{See~\ref{sec:greedy} for heuristic discussion}
+%    		\State $\bar{p} \gets q^{-1} \circ p$
+%    		\State $\mathcal{L} \gets \mathrm{LCS}(p,q) = \mathrm{LIS}(\bar{p})$ \Comment{$O(m\log \log m)$}
     		\State $(\, \mathcal{S}, \, \mathcal{D} \,  )\gets (\, \emptyset, \, [m] \setminus \mathcal{L} \,)$
@@
 There is a known $\frac{3}{2}$-approximation that can be computed efficiently which reduces the problem to the traveling salesman problem on a metric graph~\cite{christofides2022worst}, and thus can be used so long as the distance function between templates is a valid metrics.
-\cite{lesnick2012multidimensional} use the kendall distance between the push-map induced filtrations, but other options are available---for example, any of the combinatorial metrics we studied in Section~\ref{sec:schedule_cost}.
+\cite{lesnick2012multidimensional} use the Kendall distance between the push-map induced filtrations, but other options are available---for example, any of the combinatorial metrics we studied in Section~\ref{sec:schedule_cost}.
 

### 2023-10-13 move_schedules.tex
Finalizing the next revision prior to final couple checks for grammar and typos

@@
     \includegraphics[height=1.25in]{circle_complex.png}
-    \caption{Top: A video of an expanding annulus. Bottom: Sublevel-set filtrations, via pixel intensity, of a Freudenthal triangulation of the plane.}
+    \caption{Top: A video of an expanding annulus. Bottom: Sublevel-set filtrations, via (negative) pixel intensity, of a Freudenthal triangulation of the plane.}
     \label{fig:vidExample}
@@
 Using \emph{vineyards}, one must consistently perform $\lvert i - j \rvert - 1$ non-zero status queries interleaved between repairing column operations.
-A move operation, on the other hand, groups these status queries into a single pass prior to performing any modifying operations.
+A move operation groups these status queries into a single pass prior to performing any modifying operations.
 
@@
 As shown by example 2.3, \emph{moves} can be cheaper than \emph{vineyards} in terms of column operations.
-However, it is not immediately clear that this is always the case upon inspection of~\ref{alg:mr}, as the usage of a donor column seemingly implies that many $O(m)$ copy operations need to be performed.
+However, it is not clear that this is always the case upon inspection of~\ref{alg:mr}, as the usage of a donor column seemingly implies that many $O(m)$ copy operations need to be performed.
 It turns out that we may handle all such operations except the first in $O(1)$ time, which we formalize below.
@@
 \end{proof}
-
 As a final remark, we note that the combination of \emph{MoveRight} and \emph{MoveLeft} enable efficient simplex additions or deletions to the underlying complex.
-In particular, given $K$ and a decomposition $R = DV$, obtaining a valid decomposition $R' = D' V'$ of $K' = K \cup \{\sigma\}$ can be achieved by appending its requisite elementary chains to $D$ and $V$, reducing them, and then executing \emph{MoveLeft}$(m+1, i)$, where $i = f'(\sigma)$.
+In particular, given $K$ and a decomposition $R = DV$, obtaining a valid decomposition $R' = D' V'$ of $K' = K \cup \{\sigma\}$ can be achieved by appending its requisite elementary chains to $D$ and $V$, reducing them, and then executing \emph{MoveLeft}$(m+1, i)$ with $i = f'(\sigma)$.
 Dually, deleting a simplex $\sigma_i$ may be achieved via \emph{MoveRight} by moving $i$-th to the end of the decomposition and dropping the corresponding columns.
@@
 %a discrete $1$-parameter family $\mathcal{F} = (f_1, f_2, \dots, f_n)$ of filtrations $f_i : K \rightarrow [m]$ of a simplicial complex $K$ with $\lvert K \rvert =m$, and the goal being to compute the  persistence diagrams of each $(K,f_i)$.
-Our strategy to efficiently obtain a valid $RV$ decomposition of the filtration $(K,f')$ from a given decomposition of $(K, f)$ is to decompose a fixed bijection $\rho : f \to f'$ satisfying $f' = \rho \circ f$ into a \emph{schedule} of updates:
+Our strategy to efficiently obtain a valid $RV$ decomposition of the filtration $(K,f')$ from a given decomposition of $(K, f)$ is to decompose a fixed bijection $\rho : [m] \to [m]$ satisfying $f' = \rho \circ f$ into a \emph{schedule} of updates:
 \begin{definition}[Schedule]\label{def:schedule}
@@
 \noindent
-To produce this sequence of permutations $\mathcal{S} = (s_1, \dots, s_d)$ from $\rho$, our approach is as follows: let $\rho^\ast : [m] \to [m]$ denote its \emph{induced}
+To produce this sequence of permutations $\mathcal{S} = (s_1, \dots, s_d)$ from $\rho$, our approach is as follows:
 %\footnote{Recall that an induced representation...} 
-bijection on $S_m$, and define $q = (\rho^\ast(1), \rho^\ast(2), \ldots, \rho^\ast(m) )$. 
+define $q = (\rho(1), \rho(2), \ldots, \rho(m) )$. 
 We compute a longest increasing subsequence $\mathrm{LIS}(q)$, and then use this subsequence to recover a longest common subsequence (LCS) between $f$ and $f'$, which we denote later with $\mathrm{LCS}(f, f')$.
@@
     		\State Fix $\rho \in S_m$ such that $f_1 = \rho \circ f_0$ \Comment{$O(m)$} 
-    		\State $q \gets  (\rho^\ast(1), \rho^\ast(2), \ldots, \rho^\ast(m) )$ \Comment{$O(m)$}
+    		\State $q \gets  (\rho(1), \rho(2), \ldots, \rho(m) )$ \Comment{$O(m)$}
     		\State $\mathrm{lis}_{q} \gets \mathrm{LIS}(q)$ \Comment{$O(m \log \log m)$}
@@
 \end{corollary}
-\begin{proof}
+\begin{proof} \normalsize
 	By the same reduction from Proposition~\ref{prop:schedule_size}, the problem of computing $\mathrm{LCS}(f, f')$ reduces to the problem of computing the \emph{longest increasing subsequence} (LIS) of a particular permutation $p' \in S_m$, which can done in $O( m \log \log m)$ time using van Emde Boas trees~\cite{bespamyatnikh2000enumerating}
@@
   \begin{theorem}\label{prop:sparsity_result}
-   Given a pair of filtrations $(K,f), (K,f')$ and a decomposition $R = DV$ of $K$, the size of a minimal sequence $\mathcal{S} = \left( s_1, s_2, \dots, s_d \right)$ of cyclic `move' permutations $s_k = (i_k, j_k)$ satisfying:
+   Given a pair of filtrations $(K,f), (K,f')$, a decomposition $R = DV$ of $K$, and a sequence $\mathcal{S} = \left( s_1, s_2, \dots, s_d \right)$ of cyclic `move' permutations $s_k = (i_k, j_k)$ satisfying $i_k < j_k$ for all $k \in [d]$, computing the updates:
    \begin{equation}\label{eq:rv_seq_theorem}
@@
    	\end{equation}
-   	can be determined in $O(m \log \log m)$ time and $O(m)$ space, where $R' = D_d V_d$ denotes a valid decomposition of $(K, f')$. 
-   	Moreover, if $i_k < j_k$ for all $k \in [d]$, then computing~\eqref{eq:rv_seq_theorem} requires $O(\\nu)$ column operations, where:	  
-   	$$  \quad \nu = \sum\limits_{k = 1}^d \, (\lvert \mathbb{I}_{k} \rvert + \lvert \mathbb{J}_{k} \rvert) $$
+   	where $R' = D_d V_d$ denotes a valid decomposition of $(K, f')$ requires $O(\nu)$ column operations, where $\nu$ depends on the sparsity of the intermediate entries $V_1, V_2, \dots, V_d$ and $R_1, R_2, \dots, R_d$:
+   	$$  \quad \nu = \sum\limits_{k = 1}^d \, (\lvert \mathbb{I}_{k} \rvert + \lvert \mathbb{J}_{k} \rvert), \quad \text{ where }  \lvert \mathbb{I}_{k} \rvert, \lvert \mathbb{J}_{k} \rvert \text{are given in Proposition~\ref{prop:move_st} } $$
+   	Moreover, the size of a minimal such $\mathcal{S}$ can be determined in $O(m \log \log m)$ time and $O(m)$ space. 
 %	$$  \quad \kappa = \sum\limits_{k = 1}^d \, \lvert \mathbb{I}_{k}\rvert + \lvert \mathbb{J}_{k}\rvert $$
-	where the quantities $\lvert \mathbb{I}_{k} \rvert, \lvert \mathbb{J}_{k} \rvert$ of the intermediate $R_k, V_k$ are given in Proposition~\ref{prop:move_st}.
-Note that since $\nu$ depends on the sparsity of the intermediate entries $V_1, V_2, \dots, V_d$ and $R_1, R_2, \dots, R_d$, the bound $O(\nu)$ is output-sensitive.
 	%	and where $\lvert \mathbb{I}_{k} \rvert$ denotes the number of non-zero entries $V_k[i_k, i_k:j_k]$, $\lvert \mathbb{J}_{k}\rvert$ the number of columns satisfying $\mathrm{low}_{R_k} \in [i_k:j_k]$ and $\mathrm{row}_{R_k}(i_k) \neq 0$, and $d = m - \lvert \mathrm{LCS}(K, K') \rvert$.
@@
   The problem of constructing a sorting $\mathcal{S}$ of size $d$ thus reduces to the problem of choosing a sequence of $d$ valid moves, which we call a \emph{valid sorting}. 
-To do this efficiently, let $\mathcal{T}$ denote an ordered set-like data structure that supports the following operations on elements $\sigma \in M$ from the set $M = \{0, 1, \dots, m+1\}$:
+  To do this efficiently, let $\mathcal{U}$ denote an ordered set-like data structure that supports the following operations on elements $\sigma \in M$ from the set $M = \{0, 1, \dots, m+1\}$:
   \begin{enumerate}
-  	\item $\mathcal{T} \cup \sigma$---inserts $\sigma$ into $\mathcal{T}$,
-  	\item $\mathcal{T} \setminus \sigma$---removes $\sigma$ from $\mathcal{T}$,
-  	\item $\mathcal{T}_{\mathrm{succ}}(\sigma)$---obtain the successor of $\sigma$ in $\mathcal{T}$, if it exists, otherwise return $m+1$
-  	\item $\mathcal{T}_{\mathrm{pred}}(\sigma)$---obtain the predecessor of $\sigma$ in $\mathcal{T}$, if it exists, otherwise return $0$
+  	\item $\mathcal{U} \cup \sigma$---inserts $\sigma$ into $\mathcal{U}$,
+  	\item $\mathcal{U} \setminus \sigma$---removes $\sigma$ from $\mathcal{U}$,
+  	\item $\mathcal{U}_{\mathrm{succ}}(\sigma)$---obtain the successor of $\sigma$ in $\mathcal{U}$, if it exists, otherwise return $m+1$
+  	\item $\mathcal{U}_{\mathrm{pred}}(\sigma)$---obtain the predecessor of $\sigma$ in $\mathcal{U}$, if it exists, otherwise return $0$
   \end{enumerate}
-  Given such a $\mathcal{T}$, an arbitrary valid sorting can be constructed by repeatedly querying and maintaining information about the $\mathrm{LCS}$ in $\mathcal{T}$.
-To see this, suppose $\mathcal{T}$ contains the current $\mathrm{LCS}$ between two permutations $p$ and $q$.
+  Given $\mathcal{U}$, a valid sorting can be constructed by querying and maintaining information about the $\mathrm{LCS}$ in $\mathcal{U}$.
+To see this, suppose $\mathcal{U}$ contains the current $\mathrm{LCS}$ between two permutations $p$ and $q$.
 By definition of the LCS, we have:
  \begin{equation}\label{eq:valid_move_ineq}
- 	p^{-1}(\mathcal{T}_{\mathrm{pred}}(\sigma)) < p^{-1}(\sigma) < p^{-1}(\mathcal{T}_{\mathrm{succ}}(\sigma))
+ 	p^{-1}(\mathcal{U}_{\mathrm{pred}}(\sigma)) < p^{-1}(\sigma) < p^{-1}(\mathcal{U}_{\mathrm{succ}}(\sigma))
  \end{equation}
-  for every $\sigma \in \mathcal{T}$.
-Now, suppose we choose some element $\sigma \notin \mathcal{T}$ which we would like to add to the LCS.
-If $p^{-1}(\sigma) < p^{-1}(\mathcal{T}_{\mathrm{pred}}(\sigma))$, then we must move $\sigma$ to the right of its predecessor in $p$ such that~\eqref{eq:valid_move_ineq} holds.
-Similarly, if  $p^{-1}(\mathcal{T}_{\mathrm{succ}}(\sigma)) < p^{-1}(\sigma)$, then we must move $\sigma$ left of its successor in $p$.
-Assuming the structure $\mathcal{T}$ supports all of the above operations in $O(\log m)$ time, we easily deduce a $O(d m\log m)$ algorithm for obtaining a valid sorting.
+  for every $\sigma \in \mathcal{U}$.
+Now, suppose we choose some element $\sigma \notin \mathcal{U}$ which we would like to add to the LCS.
+If $p^{-1}(\sigma) < p^{-1}(\mathcal{U}_{\mathrm{pred}}(\sigma))$, then we must move $\sigma$ to the right of its predecessor in $p$ such that~\eqref{eq:valid_move_ineq} holds.
+Similarly, if  $p^{-1}(\mathcal{U}_{\mathrm{succ}}(\sigma)) < p^{-1}(\sigma)$, then we must move $\sigma$ left of its successor in $p$.
+Assuming the structure $\mathcal{U}$ supports all of the above operations in $O(\log m)$ time, we easily deduce a $O(d m\log m)$ algorithm for obtaining a valid sorting.
 %  which for completeness is shown via Algorithm~\ref{alg:sorting} in the appendix. 
@@
 where $\lvert \mathbb{I}\rvert + \lvert \mathbb{J}\rvert$ are the quantities from Proposition~\ref{prop:move_st}. 
-One advantage of the \emph{moves} framework is that the cost of a single move on a given $R = D V$ decomposition can be determined efficiently prior to any column operations---no more than $O(m)$ time with row-oriented sparse matrices.
-It is natural to consider whether a greedy-type solution which chooses the minimal cost choice at every step yields an efficient strategy.
-We give a counter-example below demonstrating that a greedy procedure may lead to arbitrarily bad behavior.
-\\
-\\
-\textbf{Counter-Example:} A pair of filtrations is given below, each comprising the $1$-skeleton of a $3$-simplex.
-Relabeling $(K, f)$ to the index set $f : K \to [m]$ and modifying $(K, f')$ accordingly yields the permutations given below:
-%$$ K_0 = \{ {\color{red} (1) (2) (3) (4) (1,4) (2,4) (3,4) } (1,2) (2,3) (1,3) \} $$ 
-%$$ K_1 = \{ {\color{red} (1) (2) (3) (4) } (1,2) (2,3) (1,3) {\color{red} (1,4) (2,4) (3,4) }\} $$ 
-$$ (K, f) = \{ {\color{red} a \; b \; c \; d \; u \; v \; w} \; x \; y \; z  \} = { {\color{red} 1 \; 2 \; 3 \; 4 \; 5 \; 6 \; 7 } \; 8 \; 9 \; 10 } $$
-$$ (K, f') = \{ {\color{red} a \; b \; c \; d } \; x \; y \; z \; {\color{red} u \; v \; w } \}  = { {\color{red} 1 \; 2 \; 3 \; 4 } \;8 \; 9 \; 10 \; {\color{red} 5 \; 6 \; 7 } }$$
-The subset of the filtration which corresponds to the simplices which lie in the LCS between these permutations is colored in red. For this example, the edit distance is 
-$d = m - \lvert \mathrm{LCS}(f, f') \rvert$, implying exactly $3$ moves are needed to map $f \mapsto f'$. There are six possible valid schedules of moves: 
-\begin{alignat*}{3}
-	S_1 = m_{x u}, m_{y u}, m_{z u} \quad & S_3 = m_{y u}, m_{x y}, m_{z u} \quad  & S_5 = m_{z u}, m_{x z}, m_{y z} \\
-	S_2 = m_{x u}, m_{z u}, m_{y z} \quad  & S_4 = m_{y u}, m_{z u}, m_{x y} \quad  & S_6 = m_{z u}, m_{y z}, m_{x z} 
-\end{alignat*}
-where the notation $m_{x y}$ represents the move permutation that moves $x$ to the position of $y$. The cost of each  move operation and each schedule is recorded in Table~\ref{table:move_costs}.
-\begin{table}[h]
-\caption{Move schedule costs}
-\centering
-\begin{tabular}{ m{0.4cm} m{0.8cm} m{0.8cm} m{0.8cm} m{0.8cm}  }
- \hline
- \multicolumn{5}{c}{Cost of each permutation} \\
- \hline
- & 1st & 2nd & 3rd & Total\\
- \hline
- $S_1$ & 2 & 3 & 1 & 6 \\
- \hline 
- $S_2$ & 2 & 2 & 4 & 8 \\
-  \hline 
- $S_3$ & 4 & 2 & 2 & 8 \\
-  \hline 
- $S_4$ & 4 & 3 & 3 & 10 \\
-  \hline 
- $S_5$ & 2 & 2 & 4 & 8  \\
-  \hline 
- $S_6$ & 2 & 5 & 3 & 10\\
- \hline
-\end{tabular}
-\label{table:move_costs}
-\end{table}
-Note the greedy strategy which always selects the cheapest move in succession would begin by moving $x$ or $z$ first, since these are the cheapest moves available, which implies one of $S_1, S_2, S_5, S_6$ would be picked on the first iteration.
-While the cheapest schedule $S_1$ is in this candidate set, an iterative greedy procedure would pick either $S_2$ or $S_5$, depending on the tie-breaker---thus, a greedy approach picking the lowest-cost move may not yield an optimal schedule.
-	Indeed, as the most expensive schedule $S_6$ is in initial iterations candidate set, we see that a greedy-procedure with an arbitrary tie-breaker could potentially yield a \emph{maximal-cost} schedule.
+Globally minimizing the objective~\eqref{eq:schedule_cost} directly is difficult due to the changing sparsity of the intermediate matrices $R_k, V_k$. 
+One advantage of the \emph{moves} framework is that the cost of a single move on a given $R = D V$ decomposition can be determined efficiently prior to any column operations.
+Thus, it is natural to consider whether one could minimize~\eqref{eq:schedule_cost} by greedily choosing the lowest cost move each step.
+Unfortunately, not only does this approach does not yield a minimal cost solution, we give a counter-example in the appendix (\ref{sec:greedy_counter_ex}) demonstrating such a greedy procedure may lead to arbitrarily bad behavior.
+%---no more than $O(m)$ time with row-oriented sparse matrices. 
 %Moreover, any reasonable tie-breaker that keeps track of previous cheapest moves would end up picking either schedule $S_2$ or $S_5$ as the final greedy schedule, both of which are suboptimal cost-wise.
@@
 \subsubsection{Proxy objective}\label{sec:proxy_objective}
-Although the greedy approach from the last section can lead to arbitrarily high cost schedules, we find similar greedy-like strategies can yield computationally efficient heuristics in practice, even if they are suboptimal.
+%Although the greedy approach from the last section can lead to arbitrarily high cost schedules, we find similar greedy-like strategies can yield computationally efficient heuristics in practice, even if they are suboptimal.
+%We seek a fast procedure for generating schedules that is not only substantially better than a random or simple schedule strategy in term of column reductions, but has a low enough time and storage complexity to be practical for larger data sets.
+
 %Choosing the lowest-cost move takes $\approx O(dm)$---implying a greedy strategy would require $O(d^2 m)$ time---which is computationally viable as a heuristic when $d << m$. 
-We seek a fast procedure for generating schedules that is not only substantially better than a random or simple schedule strategy in term of column reductions, but has a low enough time and storage complexity to be practical for larger data sets.
 %For example, the vineyards algorithm requires $O(m \log m + k)$ preprocessing time to compute a schedule of size $k \sim O(m^2)$ transpositions, thus for Algorithm~\ref{alg:schedule} to be considered a more efficient alternative to vineyards, a solution subquadratic in $m$ is desirable.
@@
 
-We seek a \emph{proxy objective} that correlates with~\eqref{eq:schedule_cost} and does not explicitly depend on the entries in the decomposition, as minimizing these terms directly is difficult due to  the changing sparsity of the intermediate matrices $R_k, V_k$,
-Given a pair of filtrations $(K, f), (K,f')$, consider a schedule $\mathcal{S} \in \Phi(f,f',d)$ of cyclic permutations $(i_1, j_1), (i_2, j_2), \dots, (i_d, j_d)$ minimizing:
+In light of section~\ref{sec:greedy}, we seek an alternative objective that correlates with~\eqref{eq:schedule_cost} and does not depend on the entries in the decomposition.
+Given a pair of filtrations $(K, f), (K,f')$, a natural schedule $\mathcal{S} \in \Phi(f,f',d)$ of cyclic permutations $(i_1, j_1), (i_2, j_2), \dots, (i_d, j_d)$ is one minimizing the upper bound:
 \begin{equation}\label{eq:upper_bound_move_cost}
@@
 \end{equation}
-In practice, this bound is very loose due to the sparsity of both $R$ and $V$.
-Nonetheless, the complexities of the move operations discussed in Section~\ref{sec:moves} depend on $\lvert i - j \rvert$, and minimizing~\eqref{eq:upper_bound_move_cost} has the intuitive interpretation as minimizing \emph{net displacement}.
-%The Kendall-$\tau$ distance yields an upper bound on the number of column reductions needed to perform any schedule. 
-A similar $\ell_1$-type distance for measuring the disarrangement between permutations is the \emph{Spearman footrule distance}, defined as:
+%In practice, this bound is very loose due to the sparsity of both $R$ and $V$.
+Nonetheless, the objective is intuitive, and the complexities of the move operations discussed in Section~\ref{sec:moves} depend on $\lvert i - j \rvert$.
+%and minimizing~\eqref{eq:upper_bound_move_cost} has the intuitive interpretation as minimizing \emph{net displacement}. 
+Unfortunately, even obtaining an optimal schedule $\mathcal{S}^\ast \in \Phi(f,f',d)$ minimizing~\eqref{eq:upper_bound_move_cost}
+% $\mathcal{S}^\ast = \argmin_{\mathcal{S} \in \Phi(f,f',d)} \mathrm{\widetilde{cost}}(\mathcal{S})$
+does not appear tractable due to its similarity with the $k$-layer crossing minimization problem, which is NP-hard for $k$ sets of permutations when $k \geq 4$~\cite{biedl2009complexity}.
+For additional discussion on the relationship between these two problems, see section~\ref{sec:cross_minimization} in the appendix.
+
+In light of the discussion above, we devise a \emph{proxy} objective function based on the Spearman distance~\cite{diaconis1977spearman} which we observed is both efficient to optimize and effective in practice.
+The \emph{Spearman footrule distance} $F(p,q)$ between two $p,q \in S_m$ is an $\ell_1$-type distance for measuring permutation disarrangement; it is classically defined as:
 \begin{equation}\label{eq:spearman_dist}
-F(p, q) = \sum\limits_{i =1}^m \lvert p(i) - q(i) \rvert =  \sum\limits_{i =1}^m \lvert i - (q^{-1} \circ p)(i) \rvert
+F(p, q) = \sum\limits_{i =1}^m \lvert \, p(i) - q(i) \, \rvert =  \sum\limits_{i =1}^m \lvert \,i - (q^{-1} \circ p)(i) \, \rvert
 \end{equation}
-The Spearman distance shares any similarities with $K_\tau$; like $K_\tau$, it is also metric on $S_m$ that is invariant under consistent relabeling. 
-%and the distance between pairs of uniformly random permutations $p, q \sim S_m$ are asymptotically normal. 
-Indeed, Diaconis~\cite{diaconis1977spearman} showed the Spearman distance approximates $K_\tau$ within a factor of two: 
-\begin{equation}\label{eq:diaconis_inequality}
-K_\tau(p, q) \leq F(p, q) \leq 2 K_\tau(p, q)
-\end{equation}
-Moreover, in contrast to $K_\tau$, the Spearman distance can be computed in $O(m)$ time and is often used in the well-known \emph{rank aggregation problem}.
-Indeed, whereas obtaining a Kemeni optimal aggregation with respect to $K_\tau$ is NP-hard, the optimal such aggregation with respect to $F$ is obtainable in polynomial time~\cite{dinu2006efficient}.
-% Indeed, it is used...
-%, and 
-%For example, one of the tasks of the \emph{rank aggregation} problem is to obtain an  aggregation~\cite{} $\sigma \in S_m$ from a set of $k$ rankings $T = \{\tau_1, \tau_2, \dots, \tau_k\}$ that obeys some optimality criteria.
-A Kemeny optimal ranking is a permutation $\sigma^\ast$ that minimizes the sum of Kendall distances to the gievn set of rankings:
-%$$ \argmin_{\sigma \in S_m} = \sum\limits_{i = 1}^k K_\tau(\sigma, \tau_i)$$ 
-%Computationally, obtaining a Kemeny optimal ranking is NP-hard\cite{}.
-In contrast, a 2-approximation may be obtained by replacing the Kendall distance with the Spearman distance, as in this case the problem reduces to finding a perfect matching in a bipartite graph, which can be accomplished in polynomial time.
-To adapt $F(\cdot, \cdot)$ to sortings, we decompose the Spearman distance additively via the bound:
+Like $K_\tau$, $F$ forms a metric on $S_m$ and is invariant under relabeling.
+Our motivation for considering the footrule distance is motivated by the fact that $F$ recovers $\mathrm{\widetilde{cost}}(\mathcal{S})$ when $(p,q)$ differ by a cyclic permutation (i.e.
+$F(p, m_{ij} \circ p) = 2 \lvert i - j \rvert$) and by its usage on similar\footnote{
+$F$ is often used to approximate $K_\tau$ in \emph{rank aggregation} problems due to the fact that computing the [Kemini] optimal rank aggregation is NP-hard with respect to $K_\tau$, but only polynomial time with respect to $F$~\cite{dinu2006efficient}.
+Note $F$ bounds $K_\tau$ via the inequalities $K_\tau(p, q) \leq F(p, q) \leq 2 K_\tau(p, q)$~\cite{diaconis1977spearman}.
+} combinatorial optimization problems. 
+To adapt $F$ to sortings, we decompose $F$ additively via the bound:
 \begin{equation}\label{eq:sorting_additive}
@@
 where $\hat{s}_i =  s_i \circ \cdots \circ s_2 \circ s_1$ denote the composition of the first $i$ permutations of a sorting  $S = (s_1, \dots, s_d)$ that maps $p \mapsto \iota$.
-The problem of minimizing the right hand side of~\eqref{eq:sorting_additive} can be interpreted as a crossing minimization problem for a set of $k$-layered bipartite graphs.
-To see this, consider two permutations: $p$ and $m_{ij} \circ p$, where $m_{ij}$ is a move permutation.
-Drawing $(p, m_{ij} \circ p)$ as a bipartite graph,  observe that $F(p, m_{ij} \circ p)$ is twice the number of edge crossings in the graph, and that equality in~\eqref{eq:sorting_additive} is achieved when the displacement of each $\sigma \in p$ between its initial position in $p$ to its value is non-increasing with every application of $s_i$, which in general not guaranteed using the schedule construction method derived in section~\ref{sec:move_schedules}.
-Unfortunately, the $k$-layer crossing minimization problem is NP-hard for $k$ sets of permutations, when $k \geq 4$~\cite{biedl2009complexity}.
-\\
-\\
-\noindent \textbf{Example:} 
-Let  $p= (1\;2\;3\;4\;5\;6\;7\;8\;9)$ and $q = (9\;4\;2\;7\;1\;8\;6\;3\;5)$.
-An example of three possible schedules, $S_1$, $S_2$, and $S_3$ sorting $p$ into $q$ is given in the figure below. 
-%Each schedule, from left to right, has $21$, $31$, and $35$ crossings, respectively. 
-\begin{figure}[!htb]
-    \centering
-%    \includegraphics[width=0.60\textwidth]{crossings.png} % arxiv only 
-    \includegraphics[width=0.75\textwidth]{crossings.png} % springer only
-    %\caption{Three sortings, $S_1$, $S_2$, and $S_3$, with $21$, $31$, and $35$ crossings, resp.}
-    %.
-Each schedule, from left to right, .
-The greedy heuristic proposed to minimize~\ref{eq:schedule_opt_succ} was used to produce $S_1$, which has the minimal number of crossings, since $K_\tau(p,q) = 21$ in this example.
-    \label{fig:crossings}
+As a heuristic to minimize~\eqref{eq:sorting_additive}, we greedily select the optimal $k \in \mathcal{L}$ at each step which minimizes the Spearman distance to the identity permutation: 
+\begin{equation}\label{eq:greedy_step}
+	k_{\text{greedy}} = \argmin_{k \in \mathcal{D}} F(m_{ij} \circ p, \iota), \quad i = p^{-1}(k)
+\end{equation}  
+Note that equality between $F$ and $\hat{F}_{\mathcal{S}}$~\eqref{eq:sorting_additive} is achieved when the displacement of each $\sigma \in p$ between its initial position in $p$ to its value is non-increasing with every application of $s_i$, which in general not guaranteed using the schedule construction method in section~\ref{sec:move_schedules}.
+To build intuition for how this heuristic interacts with Algorithm~\ref{alg:sorting}, we show a purely combinatorial example below.
+\\
+\\  
+\noindent \textbf{Example:} The permutation $p \in S_m$ to sort to the identity $p \mapsto \iota$ is given as a sequence $(4, 2, 7, 1, 8, 6, 3, 5,0)$ and a precomputed LIS $\mathcal{L} = (1, 3, 5)$, which is highlighted in red.
+\begin{figure}[htb!]\label{fig:schedule_construct}
+	\center
+	\includegraphics[width=0.90\textwidth]{greedy_schedule_construct}
+%	\caption{Greedy schedule construction example.  }
 \end{figure}
-Each column represents the successive application of a move $m_{ij}$ in the schedule, 
-and the edges track the movement of each element of the permutation. 
-Black/red vertices correspond to elements in and outside of $\mathrm{LCS}$, respectively. 
-All three schedules were generated from the same $\mathrm{LCS}(p,q) = (4\,7\,8)$ and each schedule transforms $p \mapsto q$ in $d = 6$ moves. 
-In this example, $S_1$ matches the minimal number of crossings amongst all possible schedules,  since $K_\tau(p,q) = 21$. 
-\\
-\\
-%The greedy heuristic proposed to minimize equation~\ref{eq:schedule_opt_succ} was used to produce $S_1$, which has 21 edge crossings. 
-%As a result, we do not expect there to be a polynomial time algorithm for optimally achieving this proxy objective.  
-% Since $\mathcal{D}$ is fixed, the complexity of a greedy strategy depends entirely on the complexity of minimizing $F(\hat{s}_i, \hat{s}_{i+1})$ at each greedy step.  
-%For any sorting $S$ such that $s_{i}$ that increases the size of longest common subsequence between $s_{i-1} \circ \dots \circ s_1 \circ p$ and $q$. 
-% and it satisfies a certain monotonicity property under LCS edit operations. 
-%Specifically, given two permutations $p,q \in S_m$ with $\nu = \mathrm{LCS}(p,q)$ between them, the analysis in section~\ref{sec:move_schedules} shows we must move each symbol $x \in \mathcal{D}$ where $\mathcal{D} = p \smallsetminus \nu $ in a way that extends $\nu$ after each move.
-Each symbol $x \in \mathcal{D}$ induces a permutation $s_x$ such that $F(p, q) \geq F(s_x \circ p, q)$.
-That is, unlike the Kendall distance, the Spearman distance satisfies:
-%$$ F(p, q) \geq F(s_1 \circ p, q) \geq F( s_2 \circ s_1 \circ p, q) \geq F( s_d \circ \dots \circ s_1 \circ p, q)  = F(q,q) = 0 $$
-%Since the complexity of the move operations discussed in section~\ref{sec:moves} depend on the size of the movement, it is advantageous to construct a schedule which minimizes this movement. 
-%Since $\lvert \mathcal{D} \rvert$ is fixed, we would like to choose $x \in \mathcal{D}$ such that the corresponding permutation $s_x$ reduces the total displacement of every symbol $x \in \mathcal{D}$ relative to  $q$.
-%Substituting objective~\ref{eq:kendall_inequality} for $\mathrm{cost}(\cdot)$ in equation~\ref{eq:schedule_opt2} yields the following global optimization problem over $S_d$: 
-%\begin{equation}\label{eq:schedule_opt_succ}
-%S_{\sigma}^\ast = \argmin_{\sigma \in S_d} \frac{1}{2} \sum\limits_{i = 1}^{d - 1} F(\hat{s}_i, \hat{s}_{i+1})
-%\end{equation}
-%The minimizer of this optimization can be interpreted as a sorting which minimizes the net displacement amongst all possible $d$-sized sortings $p \mapsto q$, again in the $\ell_1$ sense. 
-%A greedy strategy requires $O((d+(d-1)+(d-2)+\dots+1)m) \approx O(d^2 m)$ time to complete, since there are $O(d)$ symbols to choose from and minimizing $F(\hat{s}_i, \hat{s}_{i+1})$ at each step takes $O(md)$.
-%\\
-%\\
-%\noindent
-%\textbf{Example:} We reconsider the greedy optimization example from section~\ref{sec:greedy}. 
-%Table~\ref{table:greedy_costs} lists the \emph{successive} Kendall-$\tau$ and Spearman distances, respectively, between adjacent LCS edit (move) operations.
-Observe that, in this example, both greedy approaches lead to the schedule with minimal cost ($S_1$) (though this is not guaranteed in the general case).
-%\begin{table}[h]
-%\caption{Move schedule costs (kendall and spearman)}
-%\centering
-%\begin{tabular}{ |P{0.4cm}|P{0.8cm}|P{0.8cm}|P{0.8cm}|P{0.8cm}|  }
-% \hline
-% \multicolumn{5}{|c|}{ Successive $K_\tau$ cost} \\
-% \hline
-% & 1st & 2nd & 3rd & Total\\
-% \hline
-% $S_1$ & 3 & 5 & 7 & 15 \\
-% \hline 
-% $S_2$ & 3 & 6 & 6 & 15 \\
-%  \hline 
-% $S_3$ & 4 & 4 & 7 & 15 \\
-%  \hline 
-% $S_4$ & 4 & 6 & 5 & 15 \\
-%  \hline 
-% $S_5$ & 5 & 6 & 6 & 17 \\
-%  \hline 
-% $S_6$ & 5 & 5 & 5 & 15 \\
-% \hline
-%\end{tabular}
-%\quad 
-%\begin{tabular}{ |P{0.4cm}|P{0.8cm}|P{0.8cm}|P{0.8cm}|P{0.8cm}|  }
-% \hline
-% \multicolumn{5}{|c|}{ Successive $F$ cost} \\
-% \hline
-% & 1st & 2nd & 3rd & Total\\
-% \hline
-% $S_1$ & 3 & 3 & 3 & 9 \\
-% \hline 
-% $S_2$ & 3 & 4 & 4 & 11 \\
-%  \hline 
-% $S_3$ & 4 & 4 & 3 & 11 \\
-%  \hline 
-% $S_4$ & 4 & 4 & 5 & 13 \\
-%  \hline 
-% $S_5$ & 5 & 4 & 4 & 13 \\
-%  \hline 
-% $S_6$ & 5 & 5 & 5 & 15 \\
-% \hline
-%\end{tabular}
-%\label{table:greedy_costs}
-%\end{table}  
-%\\
-
+On the left, a table records permuted sequences $\{ \hat{s}_i \circ p \}_{i=0}^d$, given on rows $i \in \{0, 1, \dots, 6\}$, for the moved elements $(7, 0, 8, 4, 6, 2)$ (highlighted in yellow). 
+ In the middle, we show the Spearman distance cost~\eqref{eq:greedy_step} associated with both each permuted sequence (left column) and with each candidate permutation $m_{ij} \circ p$ (rows).
+Note that elements $\sigma \in \mathcal{L}$ (red) induce identity permutations that do not modify the cost, and thus the minimization is restricted to elements $\sigma \in p \setminus \mathcal{L}$ (black).
+ The right plot shows the connection to the bipartite crossing minimization problem discussed in section~\ref{sec:greedy}.
+
+\subsubsection{Heuristic Computation}
+%The Kendall-$\tau$ distance yields an upper bound on the number of column reductions needed to perform any schedule.
+We now introduce an efficient $O(d^2 \log m)$ algorithm for constructing a move schedule that greedily minimizes~\eqref{eq:sorting_additive}.
+The algorithm is purely combinatorial, and is motivated by the simplicity of updating the Spearman distance between sequences which differ by a cyclic permutation.
+
+%%% TODO: elaborate on monotonicity, describe segment tree
+%Consider the schedule construction procedure outlined in Algorithm~\ref{alg:sorting}.
+At every step of the sorting, there are only two degrees of freedom: the choice of $k \in \mathcal{D}$ which yields a source position $i = \bar{p}^{-1}(k)$ to move, and the choice of target position $j$ to move to.
+The LIS $\mathcal{L}(p)$ of $p$ determines the valid target positions to move $k$ to---any target position $j$ satisfying~\eqref{eq:valid_move_ineq} is sufficient.
+If we fix a heuristic for the latter (for instance, choosing $j$ such that $\lvert i - j \lvert$ is minimal), then each $k \in \mathcal{D}$ yields a corresponding cyclic permutation $m_{ij} \in S_m$~\eqref{eq:move_permutation} that is valid in the sense of~\eqref{eq:valid_move}.
+Thus, the problem of constructing a schedule reduces to choosing  the order elements in $\mathcal{D} = [m] \setminus \mathcal{L}(p)$ to move.
+%
 %% TODO: elaborate on monotonicity, describe segment tree
-In light of the discussion above, we propose a heuristic strategy to minimize~\eqref{eq:sorting_additive} which we observed is both efficient to compute and effective in practice.
-The heuristic is inspired by the simplicity of computing the Spearman distance between cyclic permutations at the schedule construction phase.
-Suppose we begin with an array $\mathcal{A}$ of size $m$ which provides $O(1)$ access and modification, initialized with the (signed) \emph{displacement} of every element in $p$ to its corresponding position in $q$.
-Since the Spearman distance is simply the sum of the absolute value of these displacements, at any point during during the execution of Algorithm~\ref{alg:sorting} we may obtain $F(\cdot, q)$ simply by having access to the sum of every entry in $\mathcal{A}$.
-At every step, there are only two degrees of freedom in Algorithm~\ref{alg:sorting}: the choice of index $k \in \mathcal{D}$ to move, and the choice of the target index $j$ to move $k$ to.
-If we fix a heuristic for the latter (for instance, $j$ chosen arbitrarily from an appropriate set, as in Algorithm~\ref{alg:sorting}), then each $k \in \mathcal{D}$ yields a corresponding cyclic permutation $m_{kj} \in S_m$~\eqref{eq:move_permutation} satisfying~\eqref{eq:valid_move}.
-Moreover, such a permutation changes the displacement of every index in at most three different ways:
+% --- I guess this is unneeded --- 
+%Consider the schedule construction procedure outlined in Algorithm~\ref{alg:sorting}.
+At every step of the sorting, there are only two degrees of freedom: the choice of $k \in \mathcal{D}$ which yields a source position $i = \bar{p}^{-1}(k)$ to move, and the choice of target position $j$ to move to, which depends on a fixed LIS $\mathcal{L}(p)$ of $p$---any $j$ satisfying~\eqref{eq:valid_move_ineq} will suffice.
+If we fix a heuristic for the latter (for instance, choosing $j$ such that $\lvert i - j \lvert$ is minimal), then each $k \in \mathcal{D}$ yields a corresponding cyclic permutation $m_{ij} \in S_m$~\eqref{eq:move_permutation} that is valid in the sense of~\eqref{eq:valid_move}.
+Thus, the problem of constructing a schedule reduces to choosing  an order of elements in $\mathcal{D} = [m] \setminus \mathcal{L}(p)$ to move.
+First, consider an array $\mathcal{A}$ of size $m$ which provides $O(1)$ access and modification, initialized with the \emph{signed displacement} of every element in $p$ to its corresponding position in $q$.
+In the case where $q = \iota$, note $\mathcal{A}(i) = i - p(i)$, thus $F(p, \iota)$ is given by the sum of the absolute values of the elements in $\mathcal{A}$.
+%$\lVert \mathcal{A} \rVert_1$. 
+%Akin to how cyclic move permutations may be interpreted as a paired insertion-deletion operations,
+In other words, computing $F(m_{ij} \circ p, \iota)$ in $O(\log m)$ time corresponds to updating an arrays \emph{prefix sum} under element insertions and deletions, which is easily solved by many data structures (e.g.
+segment trees).
+
+Because the values of $\mathcal{A}$ are signed, the reduction to prefix sums is not exact: it is not immediately clear how to modify $\lvert i - j \rvert$ elements in $O(\log m)$ time. 
+To address this, observe that at any point during the execution of Algorithm~\ref{alg:sorting}, a cyclic permutation changes each value of $\mathcal{A}$ in at most three different ways: 
 \[
-\mathcal{A}(m_{kj} \circ p) = 
+\mathcal{A}(m_{ij} \circ p) = 
 \begin{cases} 
@@
 %All other symbols are unchanged. 
-Thus, if we replace $\mathcal{A}$ with a data structure  supporting $O(\log m)$ access time to aggregate information and $O(\log m)$ modification time over $\lvert i - j \vert$ entries simultaneously, we could greedily choose the next index $k$ to minimize:
-\begin{equation}\label{eq:greedy_step}
-	k_{\text{greedy}} = \argmin_{k \in \mathcal{D}} F(m_{kj} \circ p, \iota)
-\end{equation}  
-in $O(d\log m)$ time. 
-The former problem of accessing aggregate information reduces to the problem of efficiently calculating \emph{prefix sums}, which is easily solved. 
-It is not immediately clear how to achieve the latter modification complexity in $O(\log m)$ time, since $\lvert i - j \rvert \leq m$ is potentially larger than $O(\log m)$---however, because we are working with displacements, note the additive modification to each entry in $\lvert i - j \rvert$ is a constant.
-It is known that one can apply constant-factor updates to multiple values in a \emph{implicit treap} data structure in $O(\log m)$ time via \emph{range updates}~\cite{blelloch1998fast}.
+%Thus, if we replace $\mathcal{A}$ with a data structure  supporting $O(\log m)$ access time to aggregate information and $O(\log m)$ modification time over $\lvert i - j \vert$ entries simultaneously, we could greedily choose the next index $k$ to minimize~\eqref{eq:greedy_step} in $O(\log m)$ time.
+Thus, $\mathcal{A}$ may be partitioned into at most four contiguous intervals each upon which the update is constant and we need only apply a constant number of \emph{range updates} to  $\mathcal{A}$.
+Such updates are known to require $O(\log m)$ time using e.g.
+an \emph{implicit treap} data structure~\cite{blelloch1998fast}.
  % Can actually only achieve O(d^2 log m) it seems
-Since single element modifications, removals, and insertions can all be achieved in $O(\log m)$ expected time with such a data structure, we conclude that equation~\eqref{eq:greedy_step} may be solved in just $O(d^2 \log m)$ time.
+Since single element modifications, deletions, insertions, and constant range updates can all be achieved in $O(\log m)$ expected time with such a data structure, we conclude that equation~\eqref{eq:greedy_step} may be solved in just $O(d^2 \log m)$ time.
 %Since single element prefix sum modifications can also be achieved $O(\log m)$ time with these data structures, the time to compute $F(\hat{s}_i, \hat{s}_{i+1})$ is bounded above by $O(\log m)$, 
 %thus the greedy solution to equation~\ref{eq:schedule_opt_succ} can be computed in $O(d^2 \log m)$ time. 
+\\
+\\
+\noindent \textbf{Example:} 
+%valid moves in a corresponding $RV$ decomposition (not shown).    into new positions (given by the following row) 
+We re-use the previous example of sorting the sequence $p = (4, 2, 7, 1, 8, 6, 3, 5,0)$ to the identity $\iota = [m]$ using a precomputed LIS $\mathcal{L} = (1, 3, 5)$.
+The left table we show the same sorting as before, with elements chosen to move highlighted in yellow; on the right, a table showing the entries of $\mathcal{A}$ are recorded.
+\begin{figure}[htb!]\label{fig:schedule_construct}
+	\center
+	\includegraphics[width=0.80\textwidth]{displacement_proof}
+%	\caption{Greedy schedule construction example.  }
+\end{figure}
+The colors of each entry in the right table indicate its partition: green values are unchanged between permutations, blue values are modified by $\pm 1$, and orange values are modified arbitrarily.
+As before, the $i$th entry of the column on the left-side shows $F(\hat{s}_i \circ p, \iota)$.
+Small black lines are used to show the movement of the entries in $\mathcal{A}$ which change by $\pm 1$.
 
@@
 Since a  set of ``snapshot'' frames of a video can be equivalently thought of as discrete 1-parameter family, our framework provides a natural extension of the typical image analysis to video data.
-To demonstrate the benefit of using minimal size schedules and the scalability of the greedy approach proposed in section~\ref{sec:proxy_objective}, we re-use the video data from~\ref{sec:motivation} as a baseline benchmark.
-We perform two performance tests: one to test the impact of shrinking the number of permutations a given reindexing operation needs and one to test the asymptotic behavior of the greedy approach.
+To demonstrate the benefit of scheduling and the scalability of the greedy heuristic, we perform two performance tests on the video data from section~\ref{sec:motivation}: one to test the impact of repairing the decomposition less and one to measure the asymptotic behavior of the greedy approach.
+%we re-as a baseline benchmark.
 
  In the first test, we fix a grid size of $9 \times 9$ and record the cumulative number of column operations needed to compute persistence dynamically across 25 evenly-spaced time points using a variety of scheduling strategies.
- The primary three strategies we test are the greedy approach from section~\ref{sec:proxy_objective}, the ``simple'' approach which uses upwards of $O(m)$ move permutations via selection sort, and a third strategy which interpolates between the two.
+ The three strategies we test are the greedy approach from section~\ref{sec:proxy_objective}, the ``simple'' approach which uses upwards of $O(m)$ move permutations via selection sort, and a third strategy which interpolates between the two.
 To perform this interpolation, we use a parameter $\alpha \in [0,1]$ to choose $m - \alpha \cdot d$ random simplices to move using the same construction method outlined in section~\ref{sec:schedule_construction}.
-When $\alpha = 0$, the strategy reduces to using selection sort to reindex $f \mapsto f'$ using a random ordering of simplices; otherwise, $\alpha = 1$ reduces to using a minimal sized schedule (with a random ordering of simplices).
+%When $\alpha = 0$, the strategy reduces to using selection sort to reindex $f \mapsto f'$ using a random ordering of simplices; otherwise, $\alpha = 1$ reduces to using a minimal sized schedule (with a random ordering of simplices).
 The results are summarized in the left graph on Figure~\ref{fig:movie_perf}, wherein the mean schedule cost of the random strategies are depicted by solid lines.
 To capture the variation in performance, we run 10 independent iterations and shade the upper and lower bounds of the schedule costs.
- As one can see from the figure, while using less move operations (lower $\alpha$) does progressively reduce column operations, constructing random schedules of minimal size is no more competitive than the simple selection sort strategy.
+ As seen in Figure~\ref{fig:movie_perf}, while using less move operations (lower $\alpha$) tends to reduce column operations, constructing \emph{random} schedules of minimal size is no more competitive than the selection sort strategy.
 This suggests that efficient schedule construction needs to account for the structure of performing several permutations in sequence, like the greedy heuristic we introduced, to yield an adequate performance boost.
@@
 
-We test the efficiency of computing the necessary information to generate these crocker stacks using a spatiotemporal data set to illustrate the applicability of our method.
+To illustrate the applicability of our method, we test the efficiency of computing these crocker stacks using a spatiotemporal data set.
 Specifically, we ran a \emph{flocking} simulation similar to the one run in~\cite{topaz2015topological} with $m = 20$ vertices moving around on the unit square equipped with periodic boundary conditions (i.e.
@@
 Indeed, one of the drawbacks of persistence is its instability with respect to strong outliers, which can obscure the detection of significant topological structures~\cite{buchet2015topological}.
-One exemplary use case of multi-parameter persistence is to detect these strong outliers by filtering the data with respect to both density and the associated metric.
+One exemplary use case of multi-parameter persistence is to detect these strong outliers by filtering the data with respect to both the original filter function \emph{and} density.
 In this section, we show the utility of scheduling with a real-world use case: detecting the presence of a low-dimensional topological space which well-approximates the distribution of natural images.
@@
 
-From an implementation perspective, one non-trivial complication of our approach is its heavy dependence on a particular sparse matrix data structure which permits permuting both the row and columns of a given matrix in at most $O(m)$ time.
+From an implementation perspective, one non-trivial complication of our approach is its heavy dependence on a particular sparse matrix data structure which permits permuting both the row and columns of a given matrix in at most $O(m)$ time~\cite{cohen2006vines}.
 As shown with the natural images example in section~\ref{sec:results}, there are often more permutation operations being applied than there are column reductions.
@@
 \begin{enumerate}
-	\item Matt Piekenbrock: conceptualization, methodology design, algorithm development, experimental design \& analysis, software development, drafting of the article. 
-	\item Jose Perea: Literature review, critical revision of the manuscript, final approval of the version to be published.
+	\item Matt Piekenbrock: conceptualization, methodology design, algorithm development, experimental design \& analysis, software development. 
+	\item Jose Perea: Conceptualization, methodology design, literature review, critical revision of the manuscript, final approval of the version to be published.
 \end{enumerate}
@@
 \subsubsection{LCS-Sort}
-Here we record explicitly the simple schedule construction algorithm outlined in section~\ref{sec:schedule_construction}.
-The algorithm is simple enough to derive using the rules discussed in section~\ref{sec:schedule_construction} (namely, equation~\eqref{eq:valid_move}, but nonetheless for posterity sake we record it here for the curious reader; it is given in Algorithm~\ref{alg:sorting}.
+Here we record explicitly the schedule construction algorithm outlined in section~\ref{sec:schedule_construction}.
+The algorithm is simple enough to derive using the rules discussed in section~\ref{sec:schedule_construction} (namely, equation~\eqref{eq:valid_move}, but nonetheless for posterity sake we record it here for the curious reader; the pseudocode is given in Algorithm~\ref{alg:sorting}.
+
 The high level idea of the algorithm is to first construct the LCS between two permutations $p,q \in S_m$.
@@
 This relabeling preserves the LCS distance and has the additional advantage that $\bar{q} = \iota = [m]$ is a strictly increasing subsequence, and thus computing the LCS between $p,q \in S_m$ reduces to computing the LIS $\mathcal{L}$ of $\bar{p}$.
-By sorting $\bar{p} \mapsto p$ via operations which (strictly) increase the size of $\mathcal{L}$, we ensure that the size of the set of corresponding permutations is exactly $m - \lvert \mathcal{L} \rvert$.
+By sorting $\bar{p} \mapsto \iota$ via operations which (strictly) increase the size of $\mathcal{L}$, we ensure that the size of the set of corresponding schedule is exactly $m - \lvert \mathcal{L} \rvert$.
 
@@
 %    		    \State $\mathcal{T}_{\mathrm{pred}}(k) \gets \max \{ \ell \in \mathcal{L} \mid \ell \leq k \}$ 
-				\State $\mathrm{k}_{\text{pred}} \gets \max \{ \ell \in \mathcal{L} \mid \ell \leq k \}$\Comment{$O(\log \log m)$ using $\mathcal{T}$}
-				\State $\mathrm{k}_{\text{succ}} \gets \min \{ \ell \in \mathcal{L} \mid \ell \geq k \}$  \Comment{$O(\log \log m)$ using $\mathcal{T}$}
+				\State $\mathrm{k}_{\text{pred}} \gets \max \{ \ell \in \mathcal{L} \mid \ell \leq k \}$\Comment{$O(\log \log m)$ using $\mathcal{U}$}
+				\State $\mathrm{k}_{\text{succ}} \gets \min \{ \ell \in \mathcal{L} \mid \ell \geq k \}$  \Comment{$O(\log \log m)$ using $\mathcal{U}$}
     		    %\text{arbitrary element in }
@@
 \end{algorithm}
-Suppose $\mathcal{L}$ has been computed from $\bar{p}$.
-Since $\mathcal{L}$ is strictly increasing, the only elements left to permute are in $\mathcal{L} \setminus \bar{p}$, which we denote with $\mathcal{D}$.
+The algorithmic steps are as follows: given a LIS $\mathcal{L}$ has been computed from $\bar{p}$, since $\mathcal{L}$ is strictly increasing, the only elements left to permute are in $\mathcal{L} \setminus \bar{p}$, which we denote with $\mathcal{D}$.
 After choosing any $\sigma \in \mathcal{D}$, one then applies a cyclic permutation to $\bar{p}$ that moves $\sigma$ into any position that increases the size of $\mathcal{L}$.
-To do this efficiently, we maintain a data structure $\mathcal{T}$ which enables us to query the successor and predecessor of any given $s \in \bar{p}$ \emph{in} $\mathcal{L}$.
-We also require a data structure to query the position of a given element $\sigma \in \bar{p}$, which for now we simply use the inverse permutation $\bar{p}$ (though a more efficient representation can be used based off of \emph{displacements}, see section~\ref{sec:proxy_objective}.
+To do this efficiently, we use a set-like data structure $\mathcal{U}$ that supports efficient querying the successor and predecessor of any given $s \in \bar{p}$ with respect to $\mathcal{L}$ (such as a vEB tree).
+The simplified pseudocode also uses the inverse permutation $\bar{p}$ to query the position of a given element $\sigma \in \bar{p}$, though a more efficient representation can be used based off of an implicit treap of the \emph{displacements} array could also be used, see section~\ref{sec:proxy_objective}.
 After $\sigma$ is inserted into $\mathcal{L}$, we update $\bar{p}$, its inverse permutations $\bar{p}^{-1}$, $\mathcal{D}$ and $\mathcal{T}$ prior to the next move.
-The final set of permutations which sort $p \mapsto q$ (or equivalently, $\bar{p} \to \iota$ are stored in an array $\mathcal{S}$, which is then returned for further use.
+The final set of permutations which sort $\bar{p} \to \iota$ (or equivalently, $p \mapsto q$) are stored in an array $\mathcal{S}$, which is then returned for further use.
+
+\subsection{Greedy Counter Example}\label{sec:greedy_counter_ex}
+In this section we give a simple counter-example showing that the strategy that greedily chooses valid move permutations $\{ m_{ij} \}$ minimizing the quantity
+\begin{equation*}
+%\lvert \mathbb{I}\rvert  + \lvert \mathbb{J} \rvert
+\mathrm{cost}_{RV}(m_{ij}) = \sum\limits_{l=i+1}^j \mathds{1}\left(\, v_l(i) \neq 0 \, \right) + \sum\limits_{l=1}^m \mathds{1}\left(\, \mathrm{low}_R(l) \in [i,j] \text{ and } r_l(i) \neq 0 \, \right)
+\end{equation*}
+can lead to arbitrarily bad behavior.
+A pair of filtrations is given below, each comprising the $1$-skeleton of a $3$-simplex.
+Relabeling $(K, f)$ to the index set $f : K \to [m]$ and modifying $(K, f')$ accordingly yields the permutations:
+%$$ K_0 = \{ {\color{red} (1) (2) (3) (4) (1,4) (2,4) (3,4) } (1,2) (2,3) (1,3) \} $$ 
+%$$ K_1 = \{ {\color{red} (1) (2) (3) (4) } (1,2) (2,3) (1,3) {\color{red} (1,4) (2,4) (3,4) }\} $$ 
+$$ (K, f) = \{ {\color{red} a \; b \; c \; d \; u \; v \; w} \; x \; y \; z  \} = { {\color{red} 1 \; 2 \; 3 \; 4 \; 5 \; 6 \; 7 } \; 8 \; 9 \; 10 } $$
+$$ (K, f') = \{ {\color{red} a \; b \; c \; d } \; x \; y \; z \; {\color{red} u \; v \; w } \}  = { {\color{red} 1 \; 2 \; 3 \; 4 } \;8 \; 9 \; 10 \; {\color{red} 5 \; 6 \; 7 } }$$
+The values colored in red corresponds to $\mathrm{LCS}(f, f')$. For this example, the edit distance is 
+$d = m - \lvert \mathrm{LCS}(f, f') \rvert$ implies exactly $3$ moves are needed to map $f \mapsto f'$. There are six possible valid schedules of moves: 
+\begin{alignat*}{3}
+	S_1 = m_{x u}, m_{y u}, m_{z u} \quad & S_3 = m_{y u}, m_{x y}, m_{z u} \quad  & S_5 = m_{z u}, m_{x z}, m_{y z} \\
+	S_2 = m_{x u}, m_{z u}, m_{y z} \quad  & S_4 = m_{y u}, m_{z u}, m_{x y} \quad  & S_6 = m_{z u}, m_{y z}, m_{x z} 
+\end{alignat*}
+where the notation $m_{x y}$ represents the move permutation that moves $x$ to the position of $y$. The cost of each  move operation and each schedule is recorded in Table~\ref{table:move_costs}.
+\begin{table}[h]
+\caption{Move schedule costs}
+\centering
+\begin{tabular}{ m{0.4cm} m{0.8cm} m{0.8cm} m{0.8cm} m{0.8cm}  }
+ \hline
+ \multicolumn{5}{c}{Cost of each permutation} \\
+ \hline
+ & 1st & 2nd & 3rd & Total\\
+ \hline
+ $S_1$ & 2 & 3 & 1 & 6 \\
+ \hline 
+ $S_2$ & 2 & 2 & 4 & 8 \\
+  \hline 
+ $S_3$ & 4 & 2 & 2 & 8 \\
+  \hline 
+ $S_4$ & 4 & 3 & 3 & 10 \\
+  \hline 
+ $S_5$ & 2 & 2 & 4 & 8  \\
+  \hline 
+ $S_6$ & 2 & 5 & 3 & 10\\
+ \hline
+\end{tabular}
+\label{table:move_costs}
+\end{table}
+Note the greedy strategy which always selects the cheapest move in succession would begin by moving $x$ or $z$ first, since these are the cheapest moves available, which implies one of $S_1, S_2, S_5, S_6$ would be picked on the first iteration.
+While the cheapest schedule $S_1$ is in this candidate set, an iterative greedy procedure would pick either $S_2$ or $S_5$, depending on the tie-breaker---thus, a greedy approach picking the lowest-cost move may not yield an optimal schedule.
+	Indeed, as the most expensive schedule $S_6$ is in initial iterations candidate set, we see that a greedy-procedure with an arbitrary tie-breaker could potentially yield a \emph{maximal-cost} schedule.
+
+\subsection{Crossing minimization}\label{sec:cross_minimization}
+Conceptually, one way to view~\eqref{eq:upper_bound_move_cost} is as a crossing minimization problem over a set of $k-1$ bipartite graphs.
+To see this, consider two permutations: $p$ and $m_{ij} \circ p$, where $m_{ij}$ is a move permutation.
+Drawing $(p, m_{ij} \circ p)$ as a bipartite graph $(U, V, E)$, observe that there are $\lvert i - j \rvert$ edge crossings in the graph, and thus minimizing ~\eqref{eq:upper_bound_move_cost} is akin to a  structured variation of the $k$-layered crossing minimization problem.
+We give an example visually relating the problem of minimizing \emph{net displacement} to the problem of minimizing crossing in a $k$-layered bipartite graph. 
+Let  $p= (1\;2\;3\;4\;5\;6\;7\;8\;9)$ and $q = (9\;4\;2\;7\;1\;8\;6\;3\;5)$.
+An example of three possible schedules, $S_1$, $S_2$, and $S_3$ sorting $p$ into $q$ is given in the figure below. 
+%Each schedule, from left to right, has $21$, $31$, and $35$ crossings, respectively. 
+\begin{figure}[!htb]
+    \centering
+%    \includegraphics[width=0.60\textwidth]{crossings.png} % arxiv only 
+    \includegraphics[width=0.75\textwidth]{crossings.png} % springer only
+    %\caption{Three sortings, $S_1$, $S_2$, and $S_3$, with $21$, $31$, and $35$ crossings, resp.}
+    %.
+Each schedule, from left to right, .
+The greedy heuristic proposed to minimize~\ref{eq:schedule_opt_succ} was used to produce $S_1$, which has the minimal number of crossings, since $K_\tau(p,q) = 21$ in this example.
+    \label{fig:crossings}
+\end{figure}
+Each column represents the successive application of a move $m_{ij}$ in the schedule, 
+and the edges track the movement of each element of the permutation. 
+Black/red vertices correspond to elements inside and outside the $\mathrm{LCS}$, respectively. 
+All three schedules were generated from the same $\mathrm{LCS}(p,q) = (4\,7\,8)$ and each schedule transforms $p \mapsto q$ in $d = 6$ moves. 
+In this example, $S_1$ matches the minimal number of crossings amongst all possible schedules,  since $K_\tau(p,q) = 21$.
 

### 2023-10-13 move_schedules.tex
All reviewer points have not been addressed

@@
 For completeness, we recapitulate the motivation of the moves algorithm from~\cite{busaryev2010tracking}. 
-Let $f: K \rightarrow [m]$ denote a filtration of size $m = \lvert K \rvert$ and $R = DV$ its decomposition, where $R = [r_1, r_2, \dots, r_m]$ and $V = [v_1, v_2, \dots, v_m]$ denote the columns of $R$ and $V$, respectively.
-By definition, if $r_j = D v_j = 0$, then the inclusion $K_{j-1} \hookrightarrow K_{j}$ introduced a new cycle whose generator is given by:
-%the non-zero coefficients $(\alpha_i^{(j)}, \alpha_{i+1}^{(j)}, \dots, \alpha_j^{(j)})$ via the combination: 
-\begin{equation}\label{eq:coefficients}
-	r_j = D v_j = \alpha_j^{(1)} \cdot \partial(\sigma_{1}) + \alpha_{j}^{(2)} \cdot \partial(\sigma_{2}) + \dots + \alpha_{j}^{(j)} \cdot \partial(\sigma_{j}) = 0
-\end{equation}
-Now, consider the permutation $P$ that moves a simplex $\sigma_i$ in $K$ to position $j$, shifting all intermediate simplices $\sigma_{i+1}, \dots, \sigma_{j}$ down by one ($i < j$). 
-To perform this shift, all of the non-zero coefficients $\alpha_k^{(i)}$ contributing to $Dv_k$ for $k \in [i+1, j]$ must be set to zero, as otherwise $\partial(\sigma_i)$ may participate in boundary chains earlier than its own birth in the permuted filtration.
-Setting these coefficients to zero amounts to setting row entries $v_k(i) = 0$ for all $k \in [i+1,j]$---these cancellations manifest as column operations $(\ast)$ used to ensure $P V(\ast)$ is upper-triangular, thus maintaining invariant I2.
-Notice that these operations may yield an unreduced $R' = P R(\ast) P^T$, breaking invariant I3.
-We could reduce $R'$ with an additional column operations, but the number of such operation scales $O(\lvert i - j\rvert^2)$, which no more efficient than simply performing the permutation and re-reducing the columns $[i,j]$ in $R$.
-
-To bypass this difficulty, Busaryev et al. observed that since the initial $R$ is reduced, if it contains $s$ pivot entries in the columns $[i,j]$ of $R$, then $R'$ must also have $s$ pivots. 
-Thus, if during the cancellation  of $v_{i+1}(i) = v_{i+2}(i) = 0$ the column $r_{i+2} \mapsto r_{i+2}'$ becomes unreduced, then the pivot $\mathrm{low}_R(i+2)$ becomes \emph{free}, possibly becoming a pivot later in $r_{i+3}', \dots, r_j'$.
-If $r_{i+2}$ is copied prior to modification to a \emph{donor column}, it may re-use or \emph{donate} its pivot entry to a later column $r_{i+3}, \dots, r_j$. 
-Repeating this process at most $j - i - 1$ times ensures $R'$ stays reduced in all except possibly for the $i$-th column---and since the $k$-th such operation simultaneously sets $v_{k}(i) = 0$ without creating non-zeros at indices $v_{k}(j)$ for $j > k$, $V'$ retains its upper-triangularity.
+Let $f: K \rightarrow [m]$ denote a filtration of size $m = \lvert K \rvert$ and $R = DV$ its decomposition. 
+%For any given matrix $A$, we denote its $i$th row (column) with $\mathrm{row}_A(i)$ ($\mathrm{col}_A(i)$, respectively).
+Consider the permutation $P$ that moves a simplex $\sigma_i$ in $K$ to position $j$, shifting all intermediate simplices $\sigma_{i+1}, \dots, \sigma_{j}$ down by one ($i < j$). 
+To perform this shift, all entries $V_{ik} \neq 0$ with column positions $k \in [i+1,j]$ need to set to zero, otherwise $PV$ is not upper-triangular. 
+We may zero these entries in $V$ using column operations $V(\ast)$, ensuring invariant I2~\eqref{inv:decomposition} is maintained, however these operations may render $R' = P R(\ast) P^T$ unreduced, breaking invariant I3.
+Of course, we could then reduce $R'$ with additional column operations, but the number of such operations scales $O(\lvert i - j\rvert^2)$, which no more efficient than simply performing the permutation and applying the reduction algorithm to columns $[i,j]$ in $R$.
+
+% in $\mathrm{row}_V(i)$.
+
+%where $R = [r_1, r_2, \dots, r_m]$ and $V = [v_1, v_2, \dots, v_m]$ denote the columns of $R$ and $V$, respectively, and $D = [\partial(\sigma_i)]_{i=1}^m$.
+By definition, if $r_j = D v_j = 0$, then the inclusion $K_{j-1} \hookrightarrow K_{j}$ introduced a new cycle whose generator is given by a linear combination of boundary chains:
+%%the non-zero coefficients $(\alpha_i^{(j)}, \alpha_{i+1}^{(j)}, \dots, \alpha_j^{(j)})$ via the combination: 
+%\begin{equation}\label{eq:coefficients}
+%	r_j = D v_j = \alpha_{(1)} \cdot \partial(\sigma_{1}) + \alpha_{(2)} \cdot \partial(\sigma_{2}) + \dots + \alpha_{(j)} \cdot \partial(\sigma_{j}) = 0
+%\end{equation}
+%Now, consider the permutation $P$ that moves a simplex $\sigma_i$ in $K$ to position $j$, shifting all intermediate simplices $\sigma_{i+1}, \dots, \sigma_{j}$ down by one ($i < j$). 
+%To perform this shift, all of the non-zero coefficients $\alpha_{(i)}$ for $k \in [i+1, j]$ must be set to zero, as otherwise $\partial(\sigma_i)$ may participate in boundary chains earlier than its own birth in the permuted filtration.
+%Observe these coefficients ${\alpha_{i}}_{i=1}^j$ corresponds exactly to non-zero entries in...Setting these coefficients to zero amounts to setting row entries $v_k(i) = 0$ for all $k \in [i+1,j]$---these cancellations manifest as column operations $(\ast)$ used to ensure $P V(\ast)$ is upper-triangular, thus maintaining invariant I2.
+%Notice that these operations may yield an unreduced $R' = P R(\ast) P^T$, breaking invariant I3.
+%We could reduce $R'$ with an additional column operations, but the number of such operation scales $O(\lvert i - j\rvert^2)$, which no more efficient than simply performing the permutation and re-reducing the columns $[i,j]$ in $R$.
+
+To bypass this difficulty, Busaryev et al. observed that since $R$ is reduced, if it contains $s$ pivot entries in the columns $[i,j]$ of $R$, then $R'$ must also have $s$ pivots. 
+Thus, if column operations render some pivot-column $r_k$ of $R$ unreduced, then its pivot entry $\mathrm{low}_R(k)$ becomes \emph{free}\footnote{The process of donating pivot columns using $O(1)$ auxiliary storage is similar in spirit to the in-place sorting algorithm \emph{cycle sort}, which is often used to sort permutations in $O(n)$ time.}---if $r_{k}$ is copied prior to its modification, we may re-use or \emph{donate} its pivot entry to a later column $r_{k+1}, \dots, r_j$.
+%possibly appearing as a pivot entry in some later column $r_{k+1}$.
+%one adds column $v_i$ to $v_{k}$ and $r_i$ to $r_{k}$, for some $i < k$, and $r_{k}$ becomes unreduced
+%during the cancellation of $v_{i, i+1} = v_{i, i+2} = 0$ the column $r_{i+2} \mapsto r_{i+2}'$ becomes unreduced, then the pivot $\mathrm{low}_R(i+2)$ becomes \emph{free}, possibly becoming a pivot later in $r_{i+3}', \dots, r_j'$.
+Repeating this process at most $j - i - 1$ times ensures $R'$ stays reduced in all except possibly at its  $i$-th column.
+Moreover, since the $k$-th such operation simultaneously sets $v_{ik} = 0$, $V'$ retains its upper-triangularity.
 \\

### 2023-10-13 move_schedules.tex
Ran the language checker; fixed many typos and grammar issues

@@
 \usepackage{caption} 
-\include{commands.tex}
 \usepackage{outlines}
@@
 \begin{document}
+\include{commands.tex}
 
@@
 Computing persistence dynamically can be reduced to maintaining a valid decomposition under adjacent transpositions in the filtration order.
-	Since there are $O(m^2)$ such transpositions, this maintenance procedure exhibits limited scalability and often is too fine for many applications. 
+	Since there are $O(m^2)$ such transpositions, this maintenance procedure exhibits limited scalability and is often too fine for many applications. 
 We propose a coarser strategy for maintaining the decomposition over a 1-parameter family of filtrations.
@@
 Exploiting this connection, we show experimentally that the decrease in operations to compute diagrams across a family of filtrations, is proportional to the difference between the expected quadratic number of states and the proposed sublinear coarsening.
-Applications to video data, dynamic metric space data, and multi-parameter persistence are also presented.
+Applications to video data, dynamic metric space data, and multiparameter persistence are also presented.
 } % springer only
@@
 Though the vineyards algorithm is efficient at constructing a \emph{continuous} 1-parameter family of diagrams, it is not necessarily efficient when the parameter is coarsely discretized.
-Our methodology is based on the observation that practitioners often don't need (or want!) all of the persistence diagrams generated by a continuous 1-parameter of filtrations; usually just $n << d$ of them  suffice.
+Our methodology is based on the observation that practitioners often don't need (or want!) \emph{all} the persistence diagrams generated by a continuous 1-parameter of filtrations; usually just $n << d$ of them  suffice.
 % in many time varying settings, a valid decomposition is often only sought after at a relatively sparse number of time points relative to the size of the (finite) set needed to generated by the family.
-By exploiting the ``donor'' concept introduced in~\cite{busaryev2010tracking}, we are able to make a tradeoff between the number of times the decomposition is restored to a valid state and the granularity of the decomposition repair step, reducing the total number of column operations needed to apply an arbitrary permutation to the filtration.
-This trade off, paired with a fast greedy heuristic explained in section~\ref{sec:proxy_objective}, yields an algorithm that can update a $R = DV$ decomposition more efficiently than \emph{vineyards} in coarse time-varying contexts, making dynamic persistence more computationally tractable for a wider class of use-cases.
-The source code containing both the algorithm we propose and the experiments performed in Section~\ref{sec:results} is open source and available online.\footnote{ For all accompanying software and materials, see: \url{https://github.com/peekxc/move_schedules}}
+By exploiting the ``donor'' concept introduced in~\cite{busaryev2010tracking}, we are able to make a trade-off between the number of times the decomposition is restored to a valid state and the granularity of the decomposition repair step, reducing the total number of column operations needed to apply an arbitrary permutation to the filtration.
+This trade-off, paired with a fast greedy heuristic explained in section~\ref{sec:proxy_objective}, yields an algorithm that can update a $R = DV$ decomposition more efficiently than \emph{vineyards} in coarse time-varying contexts, making dynamic persistence more computationally tractable for a wider class of use-cases.
+Both the source code containing the algorithm we propose and the experiments performed in Section~\ref{sec:results} are available open source online.\footnote{For all accompanying software and materials, see: \url{https://github.com/peekxc/move_schedules}}
 
 \subsection{Related Work}\label{sec:related_work} 
-To the authors knowledge, work focused on ways of updating a  decomposition $R = DV$, for all homological dimensions, is limited: there is the vineyards algorithm~\cite{cohen2006vines} and the moves algorithm~\cite{busaryev2010tracking}, both of which are discussed extensively in section~\ref{sec:background}.
+To the author's knowledge, work focused on ways of updating a  decomposition $R = DV$, for all homological dimensions, is limited: there is the \emph{vineyards} algorithm~\cite{cohen2006vines} and the \emph{moves} algorithm~\cite{busaryev2010tracking}, both of which are discussed extensively in section~\ref{sec:background}.
 At the time of writing, we were made aware of very recent work~\cite{luo2021accelerating} that iteratively repairs a permuted decomposition via a column swapping and reduce strategy, which they call ``warm starts.'' Though their motivation is similar to our own, their approach relies on the reduction algorithm as a subprocedure, which is quite different from the strategy we employ here.
@@
 If  zeroth homology  is the only dimension of interest, computing and updating both the persistence and rank information  is greatly simplified.
-For example, if the edges of the graph are in filtered order a-priori, obtaining a tree representation fully characterizing the connectivity of the underlying space (also known as the \emph{incremental connectivity} problem) takes just $O(\alpha(n) n)$ time using the disjoint-set data structure, where $\alpha(n)$ is the extremely slow-growing inverse Ackermann function.
+For example, if the edges of the graph are in filtered order a priori, obtaining a tree representation fully characterizing the connectivity of the underlying space (also known as the \emph{incremental connectivity} problem) takes just $O(\alpha(n) n)$ time using the disjoint-set data structure, where $\alpha(n)$ is the extremely slow-growing inverse Ackermann function.
 Adapting this approach to the time-varying setting, Oesterling et al.~\cite{oesterling2015computing} give an algorithm that maintains a \emph{merge tree} with $e$ edges in $O(e)$ time per-update.
-If only Betti numbers are needed, the zeroth-dimension problem reduces even further to the \emph{dynamic connectivity problem}, which can be be efficiently solved in amortized $O(\log n)$ query and update times using either Link-cut trees or multi-level Euler tour trees~\cite{kapron2013dynamic}.
+If only Betti numbers are needed, the zeroth-dimension problem reduces even further to the \emph{dynamic connectivity problem}, which can be efficiently solved in amortized $O(\log n)$ query and update times using either Link-cut trees or multi-level Euler tour trees~\cite{kapron2013dynamic}.
 
@@
 
-Each image in the series is comprised of pixels whose intensities vary with time, upon which we build a simplicial complex using the \emph{Freudenthal} triangulation of the plane.
+Each image in the series consists of pixels whose intensities vary with time, upon which we build a simplicial complex using the \emph{Freudenthal} triangulation of the plane.
 For each complex, we create a filtration of simplices whose order is determined by the lower stars of pixel values.
@@
 Since it is unknown a priori when the persistent pairing function changes, one solution is to do $n$ independent persistence computations at $n$ evenly spaced points in the time domain.
-An alternative approach is to construct a homotopy between a pair of filtrations $(K, f)$, $(K,f')$ and then decompose this homotopy into adjacent transpositions based on the filtration order---the vineyards approach.
+An alternative approach is to construct a homotopy between a pair of filtrations $(K, f)$, $(K,f')$ and then decompose this homotopy into adjacent transpositions based on the filtration order---the \emph{vineyards} approach.
  We refer to the former as the \emph{discrete setting}, which is often used in practice, and the  latter as the \emph{continuous setting}.
@@
 It is known how to linearly interpolate $f \mapsto f'$ using $d \sim O(m^2)$ updates to the decomposition, where each update  requires at most two column operations~\cite{cohen2006vines}.
-Since each column operation takes $O(m)$, the complexity of reindexing $f \mapsto f'$ is $O(m^3)$, which is efficient if all $d$ decompositions are needed.
+Since each column operation takes $O(m)$, the complexity of re-indexing $f \mapsto f'$ is $O(m^3)$, which is efficient if all $d$ decompositions are needed.
 Otherwise, if only $(R', V')$ is needed, updating $R \mapsto R'$ using the approach from~\cite{cohen2006vines} matches the complexity of computing $R' = D' V'$ independently.
@@
 \end{equation}
-The quantities $\lvert \mathbb{I}_{k} \rvert$ and $\lvert \mathbb{J}_{k} \rvert$ depend on the sparsity of the $V_k$ and $R_k$ matrices, respectively, and $d \sim O(m)$ is a constant that depends on how similar $f$ and $f'$ are.
+The quantities $\lvert \mathbb{I}_{k} \rvert$ and $\lvert \mathbb{J}_{k} \rvert$ depend on the sparsity of the matrices $V_k$ and $R_k$, respectively, and $d \sim O(m)$ is a constant that depends on how similar $f$ and $f'$ are.
 The advantage of this result is that it depends explicitly on the sparsity pattern of the decomposition itself and is thus output sensitive, which we leverage in Section~\ref{sec:schedule_cost}.
@@
 %The complexity of determining $\mathcal{S}$ depends on the type of permutations considered; many edit distances exhibit tight quadratic growth rates and tend to require at least quadratic time (or worse) to compute~\cite{labarre2013lower}.
-By reinterpreting a special set of cyclic permutations as edit operations on strings, we find that a  \emph{smallest} such sequence mapping $f $ to $  f'$, has size  (Proposition~\ref{prop:schedule_size}) :
+By reinterpreting a special set of cyclic permutations as edit operations on strings, we find that any sequence mapping $f $ to $  f'$ of minimal size must have length (Proposition~\ref{prop:schedule_size}):
 \begin{equation}
@@
 In Section \ref{sec:move_schedules} we introduce move schedules and provide efficient algorithms to construct them. 
-In Section \ref{sec:results} we present applications of the proposed method, including the computation of crocker stacks from flock simulations and of a 2-dimensional persistence invariant on a data set of image patches derived from natural images.
+In Section \ref{sec:results} we present applications of the proposed method, including the computation of Crocker stacks from flock simulations and of a 2-dimensional persistence invariant on a data set of image patches derived from natural images.
 In Section \ref{sec:conclusion} we conclude the paper by discussing other possible applications and future work.
@@
 \subsection{The Reduction Algorithm}\label{sec:reduction}
-In this section we briefly recount the original reduction algorithm introduced in~\cite{zomorodian2005computing}, also sometimes called the \emph{standard} algorithm or more explicitly \texttt{pHcol}~\cite{de2011dualities}.
+In this section, we briefly recount the original reduction algorithm introduced in~\cite{zomorodian2005computing}, also sometimes called the \emph{standard} algorithm or more explicitly \texttt{pHcol}~\cite{de2011dualities}.
 The pseudocode is outlined in Algorithm~\ref{alg:reduce} in the appendix.
@@
 In this case, both before and after the permutation, $R_1$ is rendered non-reduced, requiring another column operation to restore the decomposition to a valid state.
-Starting with a valid reduction $R = DV$ and prior to performing the exchange, observe that that the highlighted entry in $V_1$ would render $V_1$ non-upper triangular after the exchange.
+Starting with a valid reduction $R = DV$ and prior to performing the exchange, observe that the highlighted entry in $V_1$ would render $V_1$ non-upper triangular after the exchange.
 This entry is removed by a left-to-right column operation, given by applying  $S_1^{2}$ on the right to $R_1$ and $V_1$.
@@
 However, several factors can affect the runtime efficiency of the vineyards algorithm. 
-On the positive side, as both the $V$ and $R$ matrices are often sparse, the cost of a given column operation is proportional to the number of non-zero  entries in the two columns being modified.
-Moreover, as a rule of thumb, its been observed that most transpositions require no column operations~\cite{edelsbrunner2000topological}.
+On the positive side, as both the matrices $R$ and $V$ are often sparse, the cost of a given column operation is proportional to the number of non-zero  entries in the two columns being modified.
+Moreover, as a rule of thumb, it has been observed that most transpositions require no column operations~\cite{edelsbrunner2000topological}.
 On the negative side, one needs to frequently query the non-zero status of various entries in $R$ and $V$ (consider evaluating e.g.
@@
 
-For completeness, we recapitulate the motivation of the moves algorithm from~\cite{busaryev2010tracking}. 
+For completeness, we recapitulate the motivation of the \emph{moves} algorithm from~\cite{busaryev2010tracking}. 
 Let $f: K \rightarrow [m]$ denote a filtration of size $m = \lvert K \rvert$ and $R = DV$ its decomposition. 
@@
 Note that the equivalent permutation using \emph{vineyards} requires $4$ column operations on both $R_1$ and $V_1$, respectively, whereas a single move operation accomplishes using only $2$ column operations per matrix.
-The pseudo-code for \emph{MoveRight} is given in Algorithm~\ref{alg:mr} and for \emph{MoveLeft} in Algorithm~\ref{alg:ml}.
+The pseudocode for \emph{MoveRight} is given in Algorithm~\ref{alg:mr} and for \emph{MoveLeft} in Algorithm~\ref{alg:ml}.
 
@@
 \begin{proposition}\label{prop:factor2}
-	Let $(K,f)$ denote a filtration of size $\lvert K \rvert = m$ with decomposition $R = D V$ and let $T$ denote the number of column operations needed by the vineyards algorithm to perform the sequence of transpositions:
+	Let $(K,f)$ denote a filtration of size $\lvert K \rvert = m$ with decomposition $R = D V$ and let $T$ denote the number of column operations needed by \emph{vineyards} to perform the sequence of transpositions:
 	$$ R = R_1 \overset{s_1}{\to} R_2 \overset{s_2}{\to} \dots \overset{s_k}{\to} R_{k-1} = R' $$
@@
 Now consider a single MoveRight($i$,$j$) outlined in Algorithm~\ref{alg:mr}. Here, the dominant cost again are the column operations (line 5). 
-Though we need an extra $O(m)$ storage allocation for the donor columns $d_\ast$ prior to the movement, notice that assignment to and from $d_\ast$ (lines (4), and (7) in RestoreRight of MoveRight, respectively) requires just $O(1)$ time via a pointer swapping argument.
+Though we need an extra $O(m)$ storage allocation for the donor columns $d_\ast$ prior to the movement, notice that assignment to and from $d_\ast$ (lines (4), and (7) in \textit{RestoreLeft} and \textit{MoveRight}, respectively) requires just $O(1)$ time via a pointer swapping argument.
 That is, when $d_\mathrm{low}' < d_\mathrm{low}$, instead of copying $\mathrm{col}_\ast(k)$ to $d_\ast'$---which takes $O(m)$ time---we instead swap their column pointers in $O(1)$   prior to column operations.
@@
 $f,f' : K \rightarrow [m]$ whose function $t\mapsto F(\sigma, t)$ is continuous and satisfies $f(\sigma) = F(\sigma, 0)$ and $f'(\sigma) = F(\sigma, 1)$ for every  $\sigma \in K$.
-Note this family includes the straight-line homotopy $F(\sigma, t) = (1 - t) f(\sigma) + t f'(\sigma)$,  studied in the original \emph{vineyards} paper~\cite{cohen2006vines}. 
+Note that this family includes the straight-line homotopy $F(\sigma, t) = (1 - t) f(\sigma) + t f'(\sigma)$,  studied in the original \emph{vineyards} paper~\cite{cohen2006vines}. 
 If we assume that each pair of curves $\big(t,  F(\cdot, t)\big) \subset [0,1]\times \mathbb{R}$  intersect in at most one point---at which they cross---the continuity and genericity assumptions on $F$ imply that for $ \sigma,\mu \in K$ distinct,  the curves   $t \mapsto F(\sigma ,t)$ and $t \mapsto F(\mu,t)$ intersect if and only if $f(\sigma) > f(\mu)$ and $f'(\sigma)  < f'(\mu)$, or $f(\sigma) < f(\mu)$ and $f'(\sigma)  > f'(\mu)$.
@@
 However, the coarsening depends entirely on the initial choice of $F$ and the quadratic upper bound remains---it is always possible that there are no contiguous subsequences to collapse.
-Suggesting one must either abandon the continuous  setting or make stronger assumptions on $F$ to have any hope of keeping $\lvert S_F \rvert \sim O(m)$ in size.
+This suggests one must either abandon the continuous  setting or make stronger assumptions on $F$ to have any hope of keeping $\lvert S_F \rvert \sim O(m)$ in size.
 % Indeed, although many transpositions require no column reductions in practice, detecting the existence of non-zero entries at specific positions in the sparse matrix takes $O(\log(m))$ time per transposition.
@@
 Contrasting the continuous-time setting, if we discard the use of a homotopy interpolation and allow move operations in any order, we obtain a trivial upper bound of $O(m)$ on the schedule size:  simply move each simplex in $K$ from  its position in the filtration given by $f$ to the position given by $f'$---which we call the \emph{naive strategy}.
-In particular, in losing the interpolation interpretation it is no longer clear the $O(m)$ bound is tight.
+In particular, in losing the interpolation interpretation, it is no longer clear the $O(m)$ bound is tight.
 Indeed, the ``intermediate'' filtrations need no longer even respect the face poset of the underlying complex $K$.
@@
 Let $S_m$ denote the symmetric group.
-Given two fixed permutations $p, q \in S_m$ and a set allowable permutations $\Sigma \subseteq S_m$,  a common  problem is to find a sequence of permutations $s_1, s_2, \dots, s_d \in \Sigma$ whose composition satisfies:
+Given two fixed permutations $p, q \in S_m$ and a set of allowable permutations $\Sigma \subseteq S_m$,  a common  problem is to find a sequence of permutations $s_1, s_2, \dots, s_d \in \Sigma$ whose composition satisfies:
 	\begin{equation}\label{eq:sorting}
@@
 
-Perhaps surprisingly, small changes to set of allowable edit operations $\Sigma$ dramatically affect both the size of $d_\Sigma(p,q)$ and the difficulty of obtaining a minimal sorting. 
+Perhaps surprisingly, small changes to the set of allowable edit operations $\Sigma$ dramatically affect both the size of $d_\Sigma(p,q)$ and the difficulty of obtaining a minimal sorting. 
 For example, while sorting by transpositions and reversals is NP-hard and sorting by prefix  transpositions is unknown, there are polynomial time algorithms for sorting by block interchanges, exchanges, and prefix exchanges~\cite{labarre2013lower}.
@@
 \begin{proposition}[Schedule Size]\label{prop:schedule_size}
-Let $(K, f), (K, f')$ denote two filtrations of size $\lvert K \rvert = m$. Then, the smallest move schedule $S^*$ reindexing $f \mapsto f'$ has size: 
+Let $(K, f), (K, f')$ denote two filtrations of size $\lvert K \rvert = m$. Then, the smallest move schedule $S^*$ re-indexing $f \mapsto f'$ has size: 
 $$ \lvert S^\ast \rvert = d = m - \lvert \mathrm{LCS}(f, f') \rvert $$ 
@@
 Since the LCS distance is a valid edit distance, if $\lvert \mathrm{LCS}(p, q) \rvert = k$, then $\lvert \mathrm{LCS}(p', \iota) \rvert = k$ as well.
-Notice that $\iota$ is strictly increasing and that any common subsequence $p'$ has with $\iota$ must also be strictly increasing.
+Notice that $\iota$ is strictly increasing, and that any common subsequence $p'$ has with $\iota$ must also be strictly increasing.
 The optimality of $d$ follows from the optimality of the well-studied LCS problem~\cite{kumar1987linear}.
@@
 For any pair of general string inputs of sizes $n$ and $m$, respectively, the LCS between them is computable in $O(mn)$ with dynamic programming, and there is substantial evidence that the complexity cannot be much lower than this~\cite{abboud2015tight}.
-In our setting, however, observe the total orders of any pair of simplexwise filtrations $(K, f), (K, f')$ of the same underlying complex $K$ may be thought of a permutations in $S_m$, in which case $d_{\mathrm{LCS}}$ reduces further to the \emph{permutation edit distance} problem.
+In our setting, however, we interpret the simplexwise filtrations $(K, f), (K, f')$ as \emph{permutations} of the same underlying complex $K$---in this setting, $d_{\mathrm{LCS}}$ reduces further to the \emph{permutation edit distance} problem.
 %Establishing a connection between the permutation edit distance and move scheduling allows us to exploit the combinatorial structure that comes from the developed theory on both LCS's and LIS's, which are both well-studied objects.
@@
 \begin{corollary}\label{cor:lis_schedule_complexity}
-Let $(K, f), (K, f')$ denote two filtrations of size $\lvert K \rvert = m$, and let $S^*$ denote schedule of minimal size reindexing $f \mapsto f'$.
+Let $(K, f), (K, f')$ denote two filtrations of size $\lvert K \rvert = m$, and let $S^*$ denote schedule of minimal size re-indexing $f \mapsto f'$.
 Then $\lvert S^\ast \rvert = d$ can be determined in $O(m \log \log m)$ time.
@@
 \begin{proof} \normalsize
-	By the same reduction from Proposition~\ref{prop:schedule_size}, the problem of computing $\mathrm{LCS}(f, f')$ reduces to the problem of computing the \emph{longest increasing subsequence} (LIS) of a particular permutation $p' \in S_m$, which can done in $O( m \log \log m)$ time using van Emde Boas trees~\cite{bespamyatnikh2000enumerating}
+	By the same reduction from Proposition~\ref{prop:schedule_size}, the problem of computing $\mathrm{LCS}(f, f')$ reduces to the problem of computing the \emph{longest increasing subsequence} (LIS) of a particular permutation $p' \in S_m$, which can be done in $O( m \log \log m)$ time using van Emde Boas trees~\cite{bespamyatnikh2000enumerating}
 \end{proof}
@@
 \begin{corollary}\label{cor:expectation}
-If $(K, f), (K,f')$ are random filtrations of a common complex $K$ of size $m$, then the expected size of longest common subsequence $\mathrm{LCS}(f, f')$ between $f,f'$ is no larger than $m - \sqrt{m}$, with probability $1$ as $m \to \infty$.
+If $(K, f), (K,f')$ are random filtrations of a common complex $K$ of size $m$, then the expected size of a longest common subsequence $\mathrm{LCS}(f, f')$ between $f,f'$ is no larger than $m - \sqrt{m}$, with probability $1$ as $m \to \infty$.
 \end{corollary}
@@
 \subsection{Constructing schedules}\label{sec:schedule_construction}
-While it is clear from the proof in Proposition~\ref{prop:schedule_size} that one may compute the LCS between two permutations $p, q \in S_m$ in $O(m \log \log m)$ time, it is not immediately clear how to obtain a sorting $p \mapsto q$ from a given $\mathcal{L} = \mathrm{LCS}(p,q)$ in an efficient way.
+While it is clear from the proof in Proposition~\ref{prop:schedule_size} that one may compute the LCS between two permutations $p, q \in S_m$ in $O(m \log \log m)$ time, it is not immediately clear how to obtain a sorting $p \mapsto q$ from a given $\mathcal{L} = \mathrm{LCS}(p,q)$ efficiently.
 We outline below a simple procedure which constructs such a sorting $\mathcal{S} = (s_1, \dots, s_d)$ in $O(dm\log m)$ time and $O(m)$ space, or $O(m \log m)$ time and $O(m)$ space per update in the online setting.
@@
 One advantage of the \emph{moves} framework is that the cost of a single move on a given $R = D V$ decomposition can be determined efficiently prior to any column operations.
-Thus, it is natural to consider whether one could minimize~\eqref{eq:schedule_cost} by greedily choosing the lowest cost move each step.
+Thus, it is natural to consider whether one could minimize~\eqref{eq:schedule_cost} by greedily choosing the lowest cost move in each step.
 Unfortunately, not only does this approach does not yield a minimal cost solution, we give a counter-example in the appendix (\ref{sec:greedy_counter_ex}) demonstrating such a greedy procedure may lead to arbitrarily bad behavior.
@@
 %Akin to how cyclic move permutations may be interpreted as a paired insertion-deletion operations,
-In other words, computing $F(m_{ij} \circ p, \iota)$ in $O(\log m)$ time corresponds to updating an arrays \emph{prefix sum} under element insertions and deletions, which is easily solved by many data structures (e.g.
+In other words, computing $F(m_{ij} \circ p, \iota)$ in $O(\log m)$ time corresponds to updating an array's \emph{prefix sum} under element insertions and deletions, which is easily solved by many data structures (e.g.
 segment trees).
@@
 %Thus, if we replace $\mathcal{A}$ with a data structure  supporting $O(\log m)$ access time to aggregate information and $O(\log m)$ modification time over $\lvert i - j \vert$ entries simultaneously, we could greedily choose the next index $k$ to minimize~\eqref{eq:greedy_step} in $O(\log m)$ time.
-Thus, $\mathcal{A}$ may be partitioned into at most four contiguous intervals each upon which the update is constant and we need only apply a constant number of \emph{range updates} to  $\mathcal{A}$.
-Such updates are known to require $O(\log m)$ time using e.g.
+Thus, $\mathcal{A}$ may be partitioned into at most four contiguous intervals, each upon which the update is constant.
+Such updates called \emph{range updates}, and are known to require $O(\log m)$ time using e.g.
 an \emph{implicit treap} data structure~\cite{blelloch1998fast}.
@@
 As one can see from the Figure, the cost of using the greedy heuristic tends to increase sub-linearly as a function of the data set size, suggesting the move scheduling approach is indeed quite scalable.
-Moreover, schedules with minimal size tended to be cheaper than otherwise, confirming our initial hypothesis that repairing the decomposition less can lead to substantial reductions at runtime.
+Moreover, schedules with minimal size tended to be cheaper than otherwise, confirming our initial hypothesis that repairing the decomposition less can lead to substantial reductions at run-time.
 
@@
 	\includegraphics[width=0.95\textwidth]{crocker_combo_1.png}
-	\caption{A crocker plot (right) depicts the evolution of dimension $p = 1$ Betti curves over time.
+	\caption{A Crocker plot (right) depicts the evolution of dimension $p = 1$ Betti curves over time.
 The green $\mathrm{X}$ marks correspond chronologically to the complexes (left), in row-major order.
@@
 
-Acknowledging this, Topaz et al.~\cite{topaz2015topological} proposed the use of a 2-dimensional summary visualization, called a \emph{crocker}\footnote{\emph{crocker} stands for ``Contour Realization Of Computed k-dimensional hole Evolution in the Rips complex.'' Although the acronym includes \emph{Rips complexes} in the name, in principle a crocker plot could just as easily be created using other types of triangulations (e.g.
+Acknowledging this, Topaz et al.~\cite{topaz2015topological} proposed the use of a 2-dimensional summary visualization, called a \emph{crocker}\footnote{\emph{crocker} stands for ``Contour Realization Of Computed k-dimensional hole Evolution in the Rips complex.'' Although the acronym includes \emph{Rips complexes} in the name, in principle a \emph{crocker} plot could just as easily be created using other types of triangulations (e.g.
 \v{C}ech filtrations).} plot.
-In brief, a crocker plot is a contour plot of a family of Betti curves.
+In brief, a \emph{crocker} plot is a contour plot of a family of Betti curves.
 Formally, given a filtration $K = K_0 \subseteq K_1 \subseteq \dots \subseteq K_m$, a $p$-dimensional \emph{Betti curve} $\beta_p^{\bullet}$ is defined as the ordered sequence of $p$-th dimensional Betti numbers:
 $$ \beta_p^\bullet = \{ \, \mathrm{rank}(H_p(K_0)), \, \mathrm{rank}(H_p(K_1)), \, \dots, \, \mathrm{rank}(H_p(K_m))\, \}$$
-Given a time-varying filtration $K(\tau)$, a crocker plot displays changes to $\beta_p^\bullet(\tau)$ as a function of $\tau$.
-A example of a crocker plot generated from the simulation described below is given in Figure~\ref{fig:crocker1}.
-Since only the Betti numbers at each simplex in the filtration are needed to generate these Betti curves, the persistence diagram is not directly needed to generate a crocker plot; it is sufficient to use e.g.
+Given a time-varying filtration $K(\tau)$, a \emph{crocker} plot displays changes to $\beta_p^\bullet(\tau)$ as a function of $\tau$.
+An example of a \emph{crocker} plot generated from the simulation described below is given in Figure~\ref{fig:crocker1}.
+Since only the Betti numbers at each simplex in the filtration are needed to generate these Betti curves, the persistence diagram is not directly needed to generate a \emph{crocker} plot; it is sufficient to use e.g.
 any of the specialized methods discussed in~\ref{sec:related_work}.
-This dependence only on the Betti numbers makes crocker plots easier to compute than standard persistence, however what one gains in efficiency one loses in stability; it is known that Betti curves are inherently unstable with respect to small fluctuations about the diagonal of the persistence diagram.
-
-Xian et al.~\cite{xian2020capturing} showed that crocker plots may be \emph{smoothed} to inherit the stability property of persistence diagrams and reduce noise in the visualization.
-That is, when applied to a time-varying persistence module $M = \{M_t\}_{t \in [0, T]}$ an $\alpha$-smoothed crocker plot for $\alpha \geq 0$ is the rank of the map $M_t(\epsilon - \alpha) \to M_t(\epsilon + \alpha)$ at time $t$ and scale $\epsilon$.
-For example, the standard crock plot is a $0$-smoothed crocker plot.
+This dependence only on the Betti numbers makes \emph{crocker} plots easier to compute than standard persistence, however what one gains in efficiency one loses in stability; it is known that Betti curves are inherently unstable with respect to small fluctuations about the diagonal of the persistence diagram.
+
+Xian et al.~\cite{xian2020capturing} showed that \emph{crocker} plots may be \emph{smoothed} to inherit the stability property of persistence diagrams and reduce noise in the visualization.
+That is, when applied to a time-varying persistence module $M = \{M_t\}_{t \in [0, T]}$, an $\alpha$-smoothed \emph{crocker} plot for $\alpha \geq 0$ is the rank of the map $M_t(\epsilon - \alpha) \to M_t(\epsilon + \alpha)$ at time $t$ and scale $\epsilon$.
+For example, the standard crock plot is a $0$-smoothed \emph{crocker} plot.
 Allowing all three parameters ($t, \epsilon, \alpha$) to vary continuously leads to 3D visualization called an $\alpha$\emph{-smoothed crocker stack}.
 \begin{definition}[crocker stack]
-	A crocker stack is a family of $\alpha$-smoothed crock plots which summarizes the topological information of a time-varying persistence module $M$ via the function $f_M : [0, T] \times [0, \infty) \times [0, \infty) \to \mathbb{N}$, where:
+	A \emph{crocker stack} is a family of $\alpha$-smoothed crock plots which summarizes the topological information of a time-varying persistence module $M$ via the function $f_M : [0, T] \times [0, \infty) \times [0, \infty) \to \mathbb{N}$, where:
 	$$ f_M(t,\epsilon, \alpha) = \mathrm{rank}(M_t(\epsilon - \alpha) \to M_t(\epsilon + \alpha)) $$
@@
 %\noindent A crocker stack is a sequence of $\alpha$-smoothed crocker plots that vary over $\alpha \geq 0$, satisfying $f_M(t,\epsilon,\alpha) \leq f_M(t,\epsilon, \alpha)$ for all $\alpha \geq \alpha'$.
-\noindent Note that, unlike crocker plots, applying this $\alpha$ smoothing efficiently \emph{requires} the persistence pairing.
-Indeed, it has been shown that crocker stacks and stacked persistence diagrams (i.e.
+\noindent Note that, unlike \emph{crocker} plots, applying this $\alpha$ smoothing efficiently \emph{requires} the persistence pairing.
+Indeed, it has been shown that \emph{crocker stacks} and stacked persistence diagrams (i.e.
 \emph{vineyards}) are equivalent to each other in the sense that either one contains the information needed to reconstruct the other~\cite{xian2020capturing}.
-Thus, computing crocker stacks reduces to computing the persistence of a (time-varying) family of filtrations.
+Thus, computing \emph{crocker stacks} reduces to computing the persistence of a (time-varying) family of filtrations.
 
@@
 
-To illustrate the applicability of our method, we test the efficiency of computing these crocker stacks using a spatiotemporal data set.
-Specifically, we ran a \emph{flocking} simulation similar to the one run in~\cite{topaz2015topological} with $m = 20$ vertices moving around on the unit square equipped with periodic boundary conditions (i.e.
+To illustrate the applicability of our method, we test the efficiency of computing these \emph{crocker stacks} using a spatio-temporal data set.
+Specifically, we ran a flocking simulation similar to the one run in~\cite{topaz2015topological} with $m = 20$ vertices moving around on the unit square equipped with periodic boundary conditions (i.e.
 $S^1 \times S^1$).
@@
 Such simulations are also called \emph{boid} simulations, and they have been extensively used as models to describe how the evolution of collective behavior over time can be described by simple sets of rules.
-The simulation is initialized with every vertex positioned randomly in the space; the positions of vertices over time is updated according to a set of rules related to the vertices acceleration, distance to other vertices, etc.
+The simulation is initialized with every vertex positioned randomly in the space; the positions of vertices over time is updated according to a set of rules related to the vertices' acceleration, distance to other vertices, etc.
 To get a sense of the time domain, we ran the simulation until a vertex made at least 5 rotations around the torus.
 
-Given this time-evolving data set, we computed the persistence diagram of the Rips filtration up to $\epsilon = 0.30$ at 60 evenly spaced time points using three approaches: the standard algorithm \texttt{pHcol} applied naïvely at each of the 60 time steps, the \emph{vineyards} algorithm applied to (linear) homotopy connecting filtrations adjacent in time, and our approach using \emph{moves}.
+Given this time-evolving data set, we computed the persistence diagram of the Rips filtration up to $\epsilon = 0.30$ at 60 evenly spaced time points using three approaches: the standard algorithm \texttt{pHcol} applied naively at each of the 60 time steps, the \emph{vineyards} algorithm applied to (linear) homotopy connecting filtrations adjacent in time, and our approach using \emph{moves}.
 The cumulative number of $O(m)$ column operations executed by three different approaches.
@@
 %	\includegraphics[width=\textwidth, height=16.15em]{boid_sim_results.png}
-	\caption{ One the left, the cumulative number of column operations (log-scale) of the three baseline approaches tested.
+	\caption{On the left, the cumulative number of column operations (log-scale) of the three baseline approaches tested.
 On the right, the normalized $K_\tau$ between adjacent filtrations depicts the coarseness of the discretization---about $5\%$ of the $\approx O(m^2)$ simplex pairs between adjacent filtrations are discordant.}
@@
 %Lesnick and Wright~\cite{lesnick2015interactive} introduced an elegant reparameterization that not only fully characterizes $\mathcal{B}(M)$, but also yields an efficient method for rendering $\mathcal{B}_L(M)$ in real time.
-Utilizing the equivalence between the rank and fibered barcode invariants, Lesnick and Wright~\cite{lesnick2015interactive} developed an elegant way of computing $\mathcal{B}(M)$ via a reparameterization using standard point-line duality.
+Utilizing the equivalence between the rank and fibered barcode invariants, Lesnick and Wright~\cite{lesnick2015interactive} developed an elegant way of computing $\mathcal{B}(M)$ via a re-parameterization using standard point-line duality.
 This clever technique effectively reduces the fibered barcode computation to a sequence of 1-D barcode computations at ``template points'' lying within the 2-cells of a particular planar subdivision $\mathcal{A}(M)$ of the half-plane $[0, \infty) \times \mathbb{R}$.
 This particular subdivision is induced by the arrangement of ``critical lines'' derived by the bigraded Betti numbers $\beta(M)$ of $M$.
-As the barcode of one template point $\mathcal{T}_e$ at the 2-cell $e \in \mathcal{A}(M)$ may be computed efficiently by re-using information from an adjacent template point $\mathcal{T}_{e'}$,~\cite{lesnick2015interactive} observed that computing the barcodes of all such template points (and thus, $\mathcal{B}(M)$) may be reduced to ordering the 2-cells in $\mathcal{A}(M)$ along a Eulerian path traversing the dual graph of $\mathcal{A}(M)$.
+As the barcode of one template point $\mathcal{T}_e$ at the 2-cell $e \in \mathcal{A}(M)$ may be computed efficiently by re-using information from an adjacent template point $\mathcal{T}_{e'}$,~\cite{lesnick2015interactive} observed that computing the barcodes of all such template points (and thus, $\mathcal{B}(M)$) may be reduced to ordering the 2-cells in $\mathcal{A}(M)$ along an Eulerian path traversing the dual graph of $\mathcal{A}(M)$.
 The full algorithm is out of scope for this effort; we include supplementary details for the curious reader in the appendix~\ref{app:2d_pers}. 
@@
 	\caption{Bipersistence example on an $8 \times 8$ coarsened grid.
-On left, the input data, colored by density.
+On the left, the input data, colored by density.
 In the middle, the bigraded Betti numbers $\beta_0(M)$ and $\beta_1(M)$ (green and red, respectively), the dimension function (gray), and a line $L$ emphasizing the persistence of features with high density.
@@
 	\includegraphics[width=0.98\textwidth]{natural_images}
-	\caption{Bipersistence example of natural images data set on an $12 \times 16$ coarsened grid.
-On the left, a projection the full data set is shown, along with the 15 landmark patches.
+	\caption{Bipersistence example of natural images data set on a $12 \times 16$ coarsened grid.
+On the left, a projection of the full data set is shown, along with the 15 landmark patches.
 (Middle) the bigraded Betti numbers and a fixed line $L$ over parameter space.
@@
 To demonstrate the benefit of 2-parameter persistence on the patch data, consider the (coarsened) fibered barcode computed from a standard Rips / codensity bifiltration on a representative sample of the image data from~\cite{hateren_schaaf_1998}, shown in Figure~\ref{fig:patch_data_dgm}.
-From the bigraded Betti number and the dimension function, one finds that a large area of dimension function is constant (highlighted as the blue portion in the middle of Figure~\ref{fig:patch_data_dgm}), wherein the first Betti number is 5.
+From the bigraded Betti number and the dimension function, one finds that a large area of the dimension function is constant (highlighted as the blue portion in the middle of Figure~\ref{fig:patch_data_dgm}), wherein the first Betti number is 5.
 Further inspection suggests one plausible candidate is the three-circle model $C_3$, which consists of three circles, two of which (say, $S_v$ and $S_h$) intersect the third (say, $S_\mathrm{lin}$) in exactly two points, but themselves do not intersect.
@@
 We combine the use of furthest-point sampling and proportionate allocation (stratified) sampling to sample landmarks $X \subset \mathcal{X}$ distributed within $n = 25$ strata.
-Each strata consists of the $(1/n)$-thick level set given the $k$-nearest neighbor density estimator $\rho_{15}$ with $k = 15$.
+Each stratum consists of the $(1/n)$-thick level set given the $k$-nearest neighbor density estimator $\rho_{15}$ with $k = 15$.
 The use of furthest-point sampling gives us certain coverage guarantees that the geometry is approximately preserved within each level set, whereas the stratification ensures the original density of is approximated preserved as well.
@@
 
-From an implementation perspective, one non-trivial complication of our approach is its heavy dependence on a particular sparse matrix data structure which permits permuting both the row and columns of a given matrix in at most $O(m)$ time~\cite{cohen2006vines}.
+From an implementation perspective, one non-trivial complication of our approach is its heavy dependence on a particular sparse matrix data structure, which permits permuting both the row and columns of a given matrix in at most $O(m)$ time~\cite{cohen2006vines}.
 As shown with the natural images example in section~\ref{sec:results}, there are often more permutation operations being applied than there are column reductions.
@@
 Another direction to explore would be the analysis of our approach under the cohomology computation~\cite{de2011dualities}, or the specialization of the move operations to specific types of filtrations such as Rips filtrations.
-Such adaptations may result in even larger reductions in the number of column operations, as have been observed in practice for the standard reduction algorithm~\cite{bauer2021ripser}.
+Such adaptations may result in even greater reductions in the number of column operations, as have been observed in practice for the standard reduction algorithm~\cite{bauer2021ripser}.
 Moreover, though we have carefully constructed an efficient greedy heuristic in section~\ref{sec:proxy_objective} and illustrated a different perspective with which to view our heuristic (via crossing minimization), it is an open question whether there exists a more structured reduction of~\eqref{eq:schedule_cost} or~\eqref{eq:sorting_additive} to a better-known problem.
@@
 % REQUIRED: siamplain format
-%\bibliographystyle{siamplain}
+% Looks like you have to just not specify the style with springer 
+% \bibliographystyle{siamplain} % springer only 
 %\bibliographystyle{unsrt}
-\bibliographystyle{plain} % arxiv only
+% \bibliographystyle{plain} % arxiv only
 \bibliography{move_schedules}
@@
 \subsubsection{Reduction Algorithm}
-The reduction algorithm, also called the ``standard algorithm,'' is the most often used modality  for computing persistence.
+The reduction algorithm, also called the ``standard algorithm,'' is the most widely used modality for computing persistence.
 While there exists other algorithms for computing persistence, they are typically not competitive with the reduction algorithm in practice.
-We outline the reduction algorithm below in Algorithm~\ref{alg:reduce}.
+We outline the reduction pseudocode in Algorithm~\ref{alg:reduce}.
 \begin{algorithm}[h!]
@@
 After setting $V$ is set to the identity, the algorithm proceeds with column operations on both $R$ and $V$, left to right, until the decomposition invariants are satisfied.
-Since each column operation takes $O(m)$ and there are potentially $O(k)$ columns in $D$ with identical low entries (line 4 in ~\ref{alg:reduce}, observe the reduction algorithm below clearly takes $O(m^2 k)$ time.
+Since each column operation takes $O(m)$ and there are potentially $O(k)$ columns in $D$ with identical low entries (line 4 in ~\ref{alg:reduce}, observe that the reduction algorithm below clearly takes $O(m^2 k)$ time.
 Since there exists complexes where $k \sim O(m)$, one concludes the bound of $O(m^3)$ is tight~\cite{morozov2005persistence}, though this seems to only be true on pathological inputs.
@@
 \subsubsection*{Move Algorithms}
-As we've covered the moves algorithm extensively in section~\ref{sec:moves}, we now record the algorithmic components of both \emph{MoveRight} and \emph{MoveLeft}.
+As we have covered the \emph{moves} algorithm extensively in Section~\ref{sec:moves}, we now record the algorithmic components of both \emph{MoveRight} and \emph{MoveLeft}.
 Though conceptually similar, note that there is an asymmetry between \emph{MoveRight} and \emph{MoveLeft}: moving a simplex upwards in the filtration requires removing non-zero entries along several columns of a particular row in $V$ so that the corresponding permutation does not render $V$ non-upper triangular.
@@
 This is due to the fact that higher entries in columns in $R$ (above the pivot entry) may very well introduce additional non-reduced columns after $R$ is permuted. 
-Since these operations always occur in a left-to-right fashion, its not immediately clear how to apply a donor column kind of concept.
+Since these operations always occur in a left-to-right fashion, it is not immediately clear how to apply a donor column kind of concept.
 Fortunately, like move right, we can still separate the algorithm into a reduction and restoration phase---see Algorithm~\ref{alg:ml}.
@@
 \subsubsection{LCS-Sort}
-Here we record explicitly the schedule construction algorithm outlined in section~\ref{sec:schedule_construction}.
-The algorithm is simple enough to derive using the rules discussed in section~\ref{sec:schedule_construction} (namely, equation~\eqref{eq:valid_move}, but nonetheless for posterity sake we record it here for the curious reader; the pseudocode is given in Algorithm~\ref{alg:sorting}.
+Here we record explicitly the schedule construction algorithm outlined in Section~\ref{sec:schedule_construction}.
+The algorithm is simple enough to derive using the rules discussed in section~\ref{sec:schedule_construction} (namely, equation~\eqref{eq:valid_move}, but nonetheless for posterity’s sake, we record it here for the curious reader; the pseudocode is given in Algorithm~\ref{alg:sorting}.
 
@@
 This relabeling preserves the LCS distance and has the additional advantage that $\bar{q} = \iota = [m]$ is a strictly increasing subsequence, and thus computing the LCS between $p,q \in S_m$ reduces to computing the LIS $\mathcal{L}$ of $\bar{p}$.
-By sorting $\bar{p} \mapsto \iota$ via operations which (strictly) increase the size of $\mathcal{L}$, we ensure that the size of the set of corresponding schedule is exactly $m - \lvert \mathcal{L} \rvert$.
+By sorting $\bar{p} \mapsto \iota$ by operations that (strictly) increase the size of $\mathcal{L}$, we ensure that the size of the corresponding schedule is exactly $m - \lvert \mathcal{L} \rvert$.
 
@@
 The algorithmic steps are as follows: given a LIS $\mathcal{L}$ has been computed from $\bar{p}$, since $\mathcal{L}$ is strictly increasing, the only elements left to permute are in $\mathcal{L} \setminus \bar{p}$, which we denote with $\mathcal{D}$.
-After choosing any $\sigma \in \mathcal{D}$, one then applies a cyclic permutation to $\bar{p}$ that moves $\sigma$ into any position that increases the size of $\mathcal{L}$.
+After choosing any $\sigma \in \mathcal{D}$, one then applies a cyclic permutation to $\bar{p}$ that moves $\sigma$ to any position that increases the size of $\mathcal{L}$.
 To do this efficiently, we use a set-like data structure $\mathcal{U}$ that supports efficient querying the successor and predecessor of any given $s \in \bar{p}$ with respect to $\mathcal{L}$ (such as a vEB tree).
-The simplified pseudocode also uses the inverse permutation $\bar{p}$ to query the position of a given element $\sigma \in \bar{p}$, though a more efficient representation can be used based off of an implicit treap of the \emph{displacements} array could also be used, see section~\ref{sec:proxy_objective}.
-After $\sigma$ is inserted into $\mathcal{L}$, we update $\bar{p}$, its inverse permutations $\bar{p}^{-1}$, $\mathcal{D}$ and $\mathcal{T}$ prior to the next move.
-The final set of permutations which sort $\bar{p} \to \iota$ (or equivalently, $p \mapsto q$) are stored in an array $\mathcal{S}$, which is then returned for further use.
+The simplified pseudocode also uses the inverse permutation $\bar{p}$ to query the position of a given element $\sigma \in \bar{p}$, although a more efficient representation can be used via an implicit treap of the \emph{displacements} array; see Section~\ref{sec:proxy_objective}.
+After $\sigma$ is inserted into $\mathcal{L}$, we update $\bar{p}$, its inverse permutations $\bar{p}^{-1}$, $\mathcal{D}$ and $\mathcal{U}$ prior to the next move.
+The final set of permutations to sort $\bar{p} \to \iota$ (or equivalently, $p \mapsto q$) are stored in an array $\mathcal{S}$, which is then returned for further use.
 
@@
 To connect this construction back to persistence, one requires the definition of bigraded Betti numbers.
-For our purposes, the $i^{\text{th}}$-graded Betti number of $M$ is simply a function $\beta_i(M): \mathbb{R}^2 \to \mathbb{N}$ whose values indicate the the number of elements at each degree in a basis of the $i^{\text{th}}$ module in a free resolution for $M$---the interested reader is referred to~\cite{lesnick2015interactive, carlsson2009theory} for a more precise algebraic definition.
+For our purposes, the $i^{\text{th}}$-graded Betti number of $M$ is simply a function $\beta_i(M): \mathbb{R}^2 \to \mathbb{N}$ whose values indicate the number of elements at each degree in a basis of the $i^{\text{th}}$ module in a free resolution for $M$---the interested reader is referred to~\cite{lesnick2015interactive, carlsson2009theory} for a more precise algebraic definition.
 Let $S = \mathrm{supp}\,\beta_0(M) \, \cup \, \mathrm{supp}\,\beta_1(M)$, where the functions $\beta_0(M), \beta_1(M)$ are $0^{\text{th}}$ and $1^{\text{st}}$ bigraded Betti numbers of $M$, respectively.

### 2023-10-13 move_schedules.tex
Final revision before release

@@
 We also show that, in expectation, the storage needed to employ this strategy is actually sublinear in $m$. 
-Exploiting this connection, we show experimentally that the decrease in operations to compute diagrams across a family of filtrations, is proportional to the difference between the expected quadratic number of states and the proposed sublinear coarsening.
+Exploiting this connection, we show experimentally that the decrease in operations to compute diagrams across a family of filtrations is proportional to the difference between the expected quadratic number of states and the proposed sublinear coarsening.
 Applications to video data, dynamic metric space data, and multiparameter persistence are also presented.
@@
 To the author's knowledge, work focused on ways of updating a  decomposition $R = DV$, for all homological dimensions, is limited: there is the \emph{vineyards} algorithm~\cite{cohen2006vines} and the \emph{moves} algorithm~\cite{busaryev2010tracking}, both of which are discussed extensively in section~\ref{sec:background}.
-At the time of writing, we were made aware of very recent work~\cite{luo2021accelerating} that iteratively repairs a permuted decomposition via a column swapping and reduce strategy, which they call ``warm starts.'' Though their motivation is similar to our own, their approach relies on the reduction algorithm as a subprocedure, which is quite different from the strategy we employ here.
+At the time of writing, we were made aware of very recent work~\cite{luo2021accelerating} that iteratively repairs a permuted decomposition via a column swapping strategy, which they call ``warm starts.'' Though their motivation is similar to our own, their approach relies on the reduction algorithm as a subprocedure, which is quite different from the strategy we employ here.
 
@@
 Two events critically change the persistence diagrams: the first occurs when the central connected component splits to form a cycle, and the second  when the annulus  splits into four components. 
-From left to right, the Betti numbers of the five evenly spaced `snapshots' of the filtration shown above are: $(\beta_0,\beta_1)  = (1,0), \; (1,1),  \; (1,1), \; (1,1), \; (4,0)$.
+From left to right, the $\epsilon$-persistent Betti numbers\footnote{By ``$\epsilon$-persistent Betti number'', we mean the number of persistent pairs lying above the diagonal with persistence at least $\epsilon$, for some suitable choice of $\epsilon > 0$.} of the five evenly spaced `snapshots' of the filtration shown in Figure~\ref{fig:vidExample} are: $(\beta_0,\beta_1)  = (1,0), \; (1,1),  \; (1,1), \; (1,1), \; (4,0)$.
 Thus, in this example, only a few persistence diagrams are needed to capture the major changes to the topology.
@@
  	\centering
- 	\includegraphics[width=0.75\textwidth]{circle_vineyards_vector.pdf}
+ 	\includegraphics[width=0.75\textwidth]{circle_vineyards.png}
  	\caption{The cumulative column operations needed to compute persistence across the time-varying filtration of grayscale images.
@@
 The quantities $\lvert \mathbb{I}_{k} \rvert$ and $\lvert \mathbb{J}_{k} \rvert$ depend on the sparsity of the matrices $V_k$ and $R_k$, respectively, and $d \sim O(m)$ is a constant that depends on how similar $f$ and $f'$ are.
-The advantage of this result is that it depends explicitly on the sparsity pattern of the decomposition itself and is thus output sensitive, which we leverage in Section~\ref{sec:schedule_cost}.
+As this result depends explicitly on the sparsity pattern of the decomposition itself, it is an output sensitive bound.
 %Surprisingly, due to a refined analysis of restore subprocedure from~\cite{busaryev2010tracking} we give in section~\ref{sec:continuous_setting}, we find that move operations are actually cheaper than vineyards, see Proposition~\ref{prop:factor2}.
@@
 			\State $h \gets \texttt{greedy}$ \Comment{Heuristic for picking moves}
-    		\State $\mathcal{S} \gets \mathrm{Schedule}(q, \mathrm{lis}_{q}, h)$ \Comment{$O(d^2\log m)$}
+    		\State $\mathcal{S} \gets \mathrm{Schedule}(q, \mathrm{lis}_{q}, h)$ \Comment{$O(d^2\log m)$, see Section~\ref{sec:schedule_cost}}
 %	    	\State $(R, V) \gets$ \Call{Reduction}{$D = \partial (K,f_1)$}
@@
 In light of the discussion above, we devise a \emph{proxy} objective function based on the Spearman distance~\cite{diaconis1977spearman} which we observed is both efficient to optimize and effective in practice.
-The \emph{Spearman footrule distance} $F(p,q)$ between two $p,q \in S_m$ is an $\ell_1$-type distance for measuring permutation disarrangement; it is classically defined as:
+The \emph{Spearman footrule distance} $F(p,q)$ between two $p,q \in S_m$ is an $\ell_1$-type distance for measuring permutation disarrangement:
 \begin{equation}\label{eq:spearman_dist}
@@
 \begin{equation}\label{eq:sorting_additive}
-\hat{F}_{\mathcal{S}}(p, \iota) = \sum\limits_{i=1}^{d-1} F(\hat{s}_i \circ p, \hat{s}_{i+1} \circ p) \geq F(p, \iota)
+\hat{F}_{\mathcal{S}}(p, \iota) = \sum\limits_{i=0}^{d-1} F(\hat{s}_i \circ p, \hat{s}_{i+1} \circ p) \geq F(p, \iota)
 \end{equation}
-where $\hat{s}_i =  s_i \circ \cdots \circ s_2 \circ s_1$ denote the composition of the first $i$ permutations of a sorting  $S = (s_1, \dots, s_d)$ that maps $p \mapsto \iota$.
+where $\hat{s}_i =  s_{i} \circ \cdots \circ s_2 \circ s_1$ denotes the composition of the first $i$ permutations of a sorting  $S = (s_1, \dots, s_d)$ that maps $p \mapsto \iota$ and $s_0 = \iota$.
 As a heuristic to minimize~\eqref{eq:sorting_additive}, we greedily select the optimal $k \in \mathcal{L}$ at each step which minimizes the Spearman distance to the identity permutation: 
@@
 $$ \lVert x \rVert_D = \sqrt{\sum_{i \sim j}(x_i - x_j)^2} = \sqrt{x^T D x}$$
-where $D$ is a fixed matrix which upon application $x^T D x$ to an image $x \in \mathbb{R}^9$ yields a value proportional to the sum of the differences between each pixels 4 connected neighbors (given above by the relation $i \sim j$).
-Their research was primarily motivated by discerning whether there existed clear qualitative differences in the distributions of patches extracted from images of different modalities, such optical and range images.
+where $D$ is a fixed matrix whose quadratic form $x^T D x$ applied to an image $x \in \mathbb{R}^9$ is proportional to the sum of the differences between each pixels 4 connected neighbors (given above by the relation $i \sim j$).
+Their research was primarily motivated by discerning whether there existed clear qualitative differences in the distributions of patches extracted from images of different modalities, such as optical and range images.
 By mean-centering, contrast normalizing, and``whitening'' the data via the Discrete Cosine Transform (DCT), they show a convenient basis for $D$ may be obtained via an expansion of 8 certain non-constant eigenvectors, shown below:
@@
 The algorithm begins by copying $D$ to a new matrix $R$, to be subsequently modified in-place.
-After setting $V$ is set to the identity, the algorithm proceeds with column operations on both $R$ and $V$, left to right, until the decomposition invariants are satisfied.
+After  $V$ is set to the identity, the algorithm proceeds with column operations on both $R$ and $V$, left to right, until the decomposition invariants are satisfied.
 Since each column operation takes $O(m)$ and there are potentially $O(k)$ columns in $D$ with identical low entries (line 4 in ~\ref{alg:reduce}, observe that the reduction algorithm below clearly takes $O(m^2 k)$ time.
-Since there exists complexes where $k \sim O(m)$, one concludes the bound of $O(m^3)$ is tight~\cite{morozov2005persistence}, though this seems to only be true on pathological inputs.
+Since there exist complexes where $k \sim O(m)$, one concludes the bound of $O(m^3)$ is tight~\cite{morozov2005persistence}, though this seems to only be true on pathological inputs.
 Indeed, a more refined analysis by Edelsbrunner et al.~\cite{edelsbrunner2000topological} shows the reduction algorithm scales by the sum of squares of the cycle persistences, which is an output-sensitive bound.

### 2023-10-14 misc/rivet_summary.tex
Reorganizing

\documentclass[10pt]{article}
	\usepackage[margin=0.75in]{geometry}
\usepackage{amsmath, amsthm, amssymb}
\usepackage{mathtools}
\usepackage{url}
\usepackage{enumitem}
\usepackage{setspace}
\usepackage{scalefnt}
\usepackage[utf8]{inputenc}
\usepackage[english]{babel}
\usepackage{algorithm}
%\usepackage{algorithm2e}
%\usepackage{algorithmicx}
\usepackage[noend]{algpseudocode}
\usepackage{bbm}
\usepackage{graphicx}

\usepackage{hyperref} % load hyperref last

\newtheorem{theorem}{Theorem}
\newtheorem{definition}{Definition}
\newtheorem{proposition}{Proposition}
\newtheorem{lemma}{Lemma}
\newtheorem{corollary}{Corollary}
% \newcommand{\CommentX}[1]{\unskip~\%~#1~\%}

\DeclareMathOperator*{\argmax}{arg\,max}
\DeclareMathOperator*{\argmin}{arg\,min}
\DeclarePairedDelimiter\ceil{\lceil}{\rceil}
\DeclarePairedDelimiter\floor{\lfloor}{\rfloor}

\begin{document}

	\subsection*{Multidimensional persistence}
A natural question is whether or not there exists a multi-dimensional generalization of persistence. The starting point for this generalization is a structure which filters the space in multiple dimensions simultaneously, a \emph{multifiltration}. Our goal is to use such a structure to identify persistent features by examining the entire multifiltration. Such a generalization has been thought to have enormous potential in terms of applications~\cite{}. 
For example, one of the often noted drawbacks of persistence is its instability with respect to strong outliers. That is, although persistence diagrams are stable with respect to their input and thus are robust to noise, the presence of outliers can artificially introduce topological features such as loops, obscuring the persistence of significant topological features~\cite{buchet2015topological}. Indeed, this obfuscation can hinder the ability of 1D persistence to uncover ``significant'' topological features, notably one of the most important applications of persistence. In theory, this formalism enables the practitioner to analyze even noisy data sets directly without resorting to sophisticated data preprocessing techniques which may or may not preserve the underlying topology of the data.

Unfortunately, invariants related to multidimensional persistence have proven challenging to use in practice due to the high algorithmic complexity of their corresponding computation. A number of computational strategies have been proposed to counter this, such as homology preserving multi-filtration simplification or parallelization schemes using shared-memory computation~\cite{}. Another strategy is to focus specifically on the case of 2-D persistence. Utilizing the equivalence between the rank and fibered barcode invariants, Lesnick and Wright~\cite{} developed an elegant way of computing the former via a reparameterization of the latter using standard point-line duality. This clever reparameterization reduces the computation of the rank invariant to sequence of 1-D barcode computations at ``template points'' lying within the 2-cells of particular planar subdivision of the half-plane $[0, \infty) \times \mathbb{R}$.

%Indeed, since algorithm~\ref{alg:schedule} was designed for precisely such a computation, the algorithm proposed by~\cite{lesnick2015interactive} is prototypical of the class of methods that stand to benefit from move scheduling. In what follows, we briefly recount this reparameterization

% --- For each section of the paper, consider writing a mini-introduction that says what its organization is, what is in each subpart, and how the parts relate to one another.

%To demonstrate how our scheduling approach can accelerate the multidimensional persistence computation, we establish the computational framework we will be considering. 
% These transforms end up being crucial in decomposing the definition of the fibered barcode (and thus, the rank invariant) into a discrete computation.

\subsubsection*{The Rank Invariant}

%We begin with a few definitions. An $n-D$ \emph{filtration} is a functor $\mathcal{F} : \mathbb{R}^n \to \textbf{Simp}$, the category of simplicial complexes, such that for all $a \leq b$, $\mathcal{F}(a,b) : F_a \to F_b$ is an inclusion. We will focus exclusively on the case where $n = 2$, where the object of study is a \emph{bifiltration}. Multifiltrations can be broadly categorized into two types: 1-critical and multi-critical filtrations. A \emph{1-critical filtration} $\mathcal{F}$ is a finite $n-D$ filtration each simplex $s \in K$ is associated with only one multi-grade (denoting its appearance in each of the filtered dimensions), otherwise $\mathcal{F}$ is said to \emph{multi-critical}. 
%An $n-D$ persistence module is a functor $M : \mathbb{R}^n \to \textbf{Vect}$, where $\textbf{Vect}$ denotes the category of $\mathbf{F}$-vector spaces and linear maps for some suitable field $\mathbb{F}$. A module $M$ is said to be \emph{pointwise finite dimensional} (p.f.d) if $\mathrm{dim}(M_a) < \infty$ for all $a \in \mathbb{R}^n$.
%For $i \geq 0$, let $H_i : \textbf{Simp} \to \textbf{Vect}$ denote the $i^{\text{th}}$ simplicial homology functor with coefficients in $\mathbb{F}$. 
%Using this notation, given a 1D filtration $\mathcal{F}$ of size $m$, each p.f.d 1D persistence module $M = H_i(\mathcal{F})$ for  may be associated with a multiset $\mathcal{B}(M)$ of intervals in $\mathbb{R}$ which record the isomorphism classes of indecomposable summands of $M$. The set $\mathcal{B}(M)$ is called the \emph{barcode}\footnote{Note that \emph{barcodes} and \emph{persistence diagrams} are interchangeable representations of the same multiset.} of $M$.

An \emph{invariant} of $n$-parameter persistence modules is a function from the collection of persistence modules to some set $S$ with the property that $f(M) = f(M')$ when $M \cong M'$. An invariant is said to be \emph{complete} if $M \cong M'$ whenever $f(M) = f(M')$. Thus, incomplete invariants are invariants which may have the same value on two non-isomorphic modules. 
\emph{The rank invariant} proposed by Carlsson et al~\cite{carlsson2009theory} is one such invariant that is complete when $n=1$ but is otherwise incomplete when $n \geq 2$. 
% The rank invariant is computable in polynomial time~\cite{carlsson2010computing}, that has been the subject of much interest lately. 
It is defined as follows: 
\begin{definition}[Rank Invariant]
	For $n \geq 1$, the \emph{rank invariant} of a $\mathbb{R}^n$-indexed persistence module $M$ over $\mathcal{H}^{n} \subseteq \mathbb{R}^n \times \mathbb{R}^n$ is the function:
	\begin{equation*}
		\begin{aligned}
		& \mathrm{rank}(M)  \; : & \mathcal{H}^n & \to \mathbb{N}  \\
		& & (a,b) & \mapsto \mathrm{rank} \, M(a,b) 
		\end{aligned}
	\end{equation*}
	for $a \leq b \in \mathcal{H}^n$. When $n = 1$, the rank invariant is complete, thus $\mathrm{rank}(M)$ and the persistence barcodes $\mathcal{B}(M)$ determine each other.
\end{definition}
\noindent Thus, in one dimension, persistent homology provides a complete invariant that fully captures the lifetime of topological features in a given filtration: 1-D persistent homology modules decompose in an essentially unique way into indecomposable summands~\cite{zomorodian2005computing}. In contrast, the theory of multidimensional persistence shows that no complete discrete invariant exists---the structure of the target for the invariant depends on the structure of the underlying field, and although multiparameter persistence modules also decompose in a similar way as the 1-D case, the set of isomorphism classes is known to be extremely complicated~\cite{carlsson2009theory, lesnick2015interactive}. 
Nonetheless, there are invariants related to multi-parameter persistence that can be quite useful in practice, even though they are incomplete. 
\noindent For example, even though the rank invariant does not encode the isomorphism type of the underlying module $M$ when $n > 1$, it does captures important ``first-order'' information about the modules structure.

Along this vein of research, Lesnick and Wright~\cite{lesnick2015interactive} identify three simple invariants associated with a multidimensional persistence module $M$: the \emph{dimension function of $M$}, the \emph{multigraded Betti numbers of $M$}, and the \emph{fibered barcode of $M$}. 
% The dimension function is simply the the function which maps every point $a \in \mathbb{R}^2$ to $\mathrm{dim}(M_a)$. It is a simple and easy to visualize invariant, but is unstable and yields no information about the persistent features of $M$. 
% The multigraded Betti numbers constitute a natural and important class of invariants associated with a multidimensional persistence module. There are many ways of defining these, with perhaps the simplest definition stemming from minimal resolutions: for a finitely generated $n$-parameter persistence module $M$ and $i \in \{0, 1, \dots, d \}$, the $i^{\text{th}}$ graded Betti number of $M$ at grade $z$, denoted as $\xi_i^M(z)$, is defined as the number of elements at grade $z$ in a basis of the $i^{\text{th}}$ module in a free resolution for $M$. 
% Thus, the $i^{\text{th}}$-graded Betti number of $M$ is a function $\xi_i(M): \mathbb{R}^n \to \mathbb{N}$. Being central objects in commutative algebra and algebraic geometry, both the theory and the computation of multi-graded Betti numbers is extensive and beyond the scope of this effort---the interested reader is referred to~\cite{carlsson2009theory, lesnick2015interactive} for more details. 
The fibered barcode, denoted as $\mathcal{B}(M)$, is defined as follows: 
\begin{definition}[Fibered barcode]
	%\in \overline{\mathcal{L}}
	The fibered barcode $\mathcal{B}(M)$ of a bipersistence module $M$ is the map which sends each line $L$ with non-negative slope to the barcode $\mathcal{B}_L(M)$: 
	$$ L \mapsto \mathcal{B}_L(M) $$
	Thus, the fibered barcode is simply a 2-parameter family of barcodes given by the collection of 1-D affine slices of $M$. 
\end{definition} 
\noindent Each line $L$ induces a 1-parameter restriction on $M$. The fibered barcode enjoys a number of stability properties readily expressed using interleaving distances, see~\cite{}. More important to this work, the fibered barcode $\mathcal{B}(M)$ and the rank invariant $\mathrm{rank}(M)$ determine each other~\cite{}.

\subsubsection*{Fibered barcode reparameterization}
Let $\overline{\mathcal{L}}$ denote the collection of all lines in $\mathbb{R}^2$ with non-negative slope, $\mathcal{L} \subset \overline{\mathcal{L}}$ the collection of all lines in $\mathbb{R}^2$ with non-negative, finite slope, and $\mathcal{L}^\circ$ the collection of all affine lines in $\mathbb{R}^2$ with positive, finite slope. There is a standard point-line duality that gives a parameterization of $\mathcal{L}$ with the half-plane $[0, \infty) \times \mathbb{R}$ that is convenient to work with here. Define the \emph{line} and \emph{point} dual transforms $\mathcal{D}_{\ell}$ and $\mathcal{D}_p$, respectively, as follows: 
\begin{equation}
	\begin{aligned}
\mathcal{D}_{\ell}: \mathcal{L} \rightarrow[0, \infty) \times \mathbb{R} & \quad \hfill \quad & \mathcal{D}_{p}:[0, \infty) \times \mathbb{R} \rightarrow \mathcal{L} \\
y=a x+b \mapsto(a,-b) &\quad \hfill \quad & (c, d) \mapsto y=c x-d
\end{aligned}
\end{equation}
% Use \multicolumn{1}{|r|}{text} to align
The transforms $\mathcal{D}_{\ell}$ and $\mathcal{D}_p$ are \emph{dual} to each other in the sense that for any point $a \in [0, \infty) \times \mathbb{R}$ and any line $L \in \mathcal{L}$, $a \in L$ if and only if $D_\ell(L) \in D_{p}(a)$. Now, for some fixed line $L$, define  the \emph{push map} $\mathrm{push}_L(a):  \mathbb{R}^2 \to L \cup \infty$ as: 
\begin{equation}
	\mathrm{push}_L(a) \mapsto \mathrm{min}\{ v \in L \mid a \leq v \}
\end{equation}
%\begin{equation*}
%	\begin{aligned}
%	& \mathrm{push}_L : & \mathbb{R}^2 & \to L \cup \infty \\
%	& & a \in \mathbb{R}^2 & \mapsto \mathrm{min}\{ v \in L \mid a \leq v \}
%	\end{aligned}
%\end{equation*}
The push map satisfies a number of useful properties. Namely: 
\begin{enumerate}
	\item For $r < s \in \mathbb{R}^2$, $\mathrm{push}_L(r) \leq \mathrm{push}_L(s)$
	\item For each $a \in \mathbb{R}^2$, $\mathrm{push}_L(a)$ is continuous on $\mathcal{L}^\circ$
	\item For $L \in \mathcal{L}^\circ$ and $S \subset \mathbb{R}^2$, $\mathrm{push}_L$ induces a totally ordered partition $S_L$ on $S$ 
\end{enumerate}
Property (1) asserts that for any $L \in \overline{\mathcal{L}}$, the partial order on $\mathbb{R}^2$ restricts to a total order on $L$. Properties (2) and (3) qualify the following definition:
\begin{definition}[Critical Lines]
	For some fixed $S \subset \mathbb{R}^2$, a line $L \in L^\circ$ is defined to be \emph{regular} if there is an open ball $B \in L^\circ$ containing $L$ such that $S_L = S_{L'}$ for all $L' \in B$. Otherwise, the line $L$ is defined as \emph{critical}. 
\end{definition}
\noindent The set of critical lines $\mathrm{crit}(M)$ with respect to some fixed set $S \subset \mathbb{R}^2$ fully characterizes a certain planar subdivision of the half plane $[0, \infty) \times \mathbb{R}$. Explicitly, define $\mathcal{A}^1(M)$ to the 1-skeleton of the line arrangement defined by: 
% \{ \mathcal{D}_p(\alpha) \text{ for } \alpha = \mathrm{LUB}(u,v), u,v \in S \} 
\begin{equation}\label{eq:arrangement_crit}
	\mathcal{A}^1(M) = \{ \mathcal{D}_{\ell}(\mathrm{crit}(M)) \cup (\{0\} \times \mathbb{R}) \}
\end{equation}
Expanding the 1-skeleton to a 2-D cell complex $\mathcal{A}(M)$ fully partitions the upper half plane.  
A consequence of equation~\ref{eq:arrangement_crit} that if the duals of two lines $L, L' \in \mathcal{L}$ are contained in the same $2$-cell in $\mathcal{A}(M)$, then $S_L = S_{L'}$, i.e. the partitions induced by $\mathrm{push}_L$ are equivalent. Indeed, the total order on $S_L$ is the pullback of the total order on $L$ with respect to the push map.
Since $\mathcal{A}(M)$ partitions the entire half-plane, the dual to every line $L \in \mathcal{L}$ is contained within $\mathcal{A}(M)$---the desired reparameterization.

Let $S = \mathrm{supp}\,\xi_0(M) \, \cup \, \mathrm{supp}\,\xi_1(M)$, where the functions $\xi_0(M), \xi_1(M)$ are $0^{\text{th}}$ and $1^{\text{st}}$ bigraded Betti numbers of $M$, respectively. 
The main mathematical result from~\cite{lesnick2015interactive} is a characterization of the barcodes $\mathcal{B}_L(M)$, for any $L \in \mathcal{L}$, in terms of a set of \emph{barcode templates} $\mathcal{T}$ computed at every 2-cell in $\mathcal{A}(M)$.
 More formally, for any line $L \in \overline{\mathcal{L}}$ and $e$ any 2-cell in $\mathcal{A}(M)$ whose closure contains the dual of $L$ under point-line duality, the 1-parameter restriction of the persistence module $M$ induced by $L$ is given by: 
 \begin{equation}
 \mathcal{B}_L(M) = \{ [\,\mathrm{push}_L(a), \, \mathrm{push}_L(b)\,) \mid (a,b) \in \mathcal{T}^e, \mathrm{push}_L(a) < \mathrm{push}_L(b)  \} 
\end{equation}
Minor additional conditions are needed for handling completely horizontal and vertical lines. 
The importance of this theorem for this effort lies in the fact that the fibered barcodes are completely defined from the precomputed barcode templates $\mathcal{T}$: once $\mathcal{T}$ has been computed and augmented onto $\mathcal{A}(M)$, $\mathcal{B}(M)$ is completely characterized, and the barcodes $\mathcal{B}_L(M)$ of any $L$-induced 1-D filtration can be efficiently computed via a point-location query on $\mathcal{A}(M)$ and a $O(\lvert \mathcal{B}_L(M) \rvert)$ application of the push map.

	\begin{figure}[h]
	\includegraphics[width=0.98\textwidth]{bifiltration_ex2}
	\caption{Bipersistence example on an $8 \times 8$ coarsened grid. On left, the input data, colored by density. In the middle, the bigraded Betti numbers $\xi_0(M)$ and $\xi_1(M)$ (green and red, respectively), the Hilbert function (gray), and a line $L$ emphasizing the persistence of features with high density. On the right, the line arrangement $\mathcal{A}(M)$ lying in the dual space $D_p(\alpha)$ derived from the $\xi(M)$. The barycenters of the 2-cells in $\mathcal{A}(M)$ depicting where the barcodes templates $\mathcal{T}$ are computed are drawn as orange points, and the dual graph $G$ of $\mathcal{A}(M)$ is drawn using dotted blue lines. The eulerian path through $G$ is shown by the solid colored lines. }
\end{figure}

\end{document}

### 2023-10-14 src/move_schedules.tex
Reorganizing

--- 2023-10-13 move_schedules.tex
@@
-%\documentclass{article} % for arxiv only
-%\usepackage{arxiv} % for arxiv only 
-%\pagestyle{plain} % arxiv only 
-%%\usepackage[margin=1in]{geometry} % arxiv only
-
-\PassOptionsToPackage{noend}{algorithmic}
-\PassOptionsToPackage{noend}{algpseudocode}
-\documentclass[sn-mathphys]{sn-jnl}
+\documentclass{article} % for arxiv only
+\usepackage{arxiv} % for arxiv only 
+\pagestyle{plain} % arxiv only 
+%\usepackage[margin=1in]{geometry} % arxiv only
+
+%\PassOptionsToPackage{noend}{algorithmic} % springer only 
+%\PassOptionsToPackage{noend}{algpseudocode} % springer only 
+%\documentclass[sn-mathphys]{sn-jnl} % springer only
 
@@
 
-%\usepackage[noend]{algorithmic}
+%\usepackage[noend]{algorithmic} % arxiv only
 
 % --- arxiv only ---- 
-%\usepackage{xcolor}
-%\usepackage{algorithm}% http://ctan.org/pkg/algorithms
-%\PassOptionsToPackage{noend}{algpseudocode}
-%\usepackage{algpseudocode}% http://ctan.org/pkg/algorithmicx
+\usepackage{xcolor}
+\usepackage{algorithm}% http://ctan.org/pkg/algorithms
+\PassOptionsToPackage{noend}{algpseudocode}
+\usepackage{algpseudocode}% http://ctan.org/pkg/algorithmicx
 
@@
 % --- Springer only ---
-\title[Fast Persistence Computations in Dynamic Settings]{
-	\vspace{-2.0em} Move Schedules: Fast persistence computations in coarse dynamic settings
-	\footnote{
-	\textbf{Funding:} This work was partially supported by the National Science Foundation through grants CCF-2006661 and CAREER award DMS-1943758.
-	}
-}
+%\title[Fast Persistence Computations in Dynamic Settings]{
+%	\vspace{-2.0em} Move Schedules: Fast persistence computations in coarse dynamic settings
+%	\footnote{
+%	\textbf{Funding:} This work was partially supported by the National Science Foundation through grants CCF-2006661 and CAREER award DMS-1943758.
+%	}
+%}
 
 % --- Arxiv only --- 
-%\title{\vspace{-0.0em} Move Schedules: Fast persistence \\ computations  in coarse dynamic settings\thanks{This work was partially supported by the National Science Foundation through grants CCF-2006661 and CAREER award DMS-1943758.}\vspace{-0.5em}}
-%\author{Matt Piekenbrock\thanks{Khoury College of Computer Sciences, Northeastern University.}\;\; and Jose A.
+\title{\vspace{-0.0em} Move Schedules: Fast persistence \\ computations  in coarse dynamic settings\thanks{This work was partially supported by the National Science Foundation through grants CCF-2006661 and CAREER award DMS-1943758.}\vspace{-0.5em}}
+\author{Matt Piekenbrock\thanks{Khoury College of Computer Sciences, Northeastern University.}\;\; and Jose A.
 Perea$^\dagger$\thanks{Department of Mathematics and Khoury College of Computer Sciences, Northeastern University.}}
-%\date{}
+\date{}
 
 % --- Springer only ---
-\author[1]{\fnm{Matthew} \sur{Piekenbrock}}\email{piekenbrock.m@northeastern.edu}
-\affil[1]{
-	\orgdiv{Khoury College of Computer Sciences}, 
-	\orgname{Northeastern University}}
+%\author[1]{\fnm{Matthew} \sur{Piekenbrock}}\email{piekenbrock.m@northeastern.edu}
+%\affil[1]{
+%	\orgdiv{Khoury College of Computer Sciences}, 
+%	\orgname{Northeastern University}}
 
 % --- Springer only --- 
-\author[2]{\fnm{Jose} \sur{A. Perea}}\email{j.pereabenitez@northeastern.edu}
-\affil[2]{
-	\orgdiv{Department of Mathematics and Khoury College of Computer Sciences},
-	\orgname{Northeastern University}}
+%\author[2]{\fnm{Jose} \sur{A. Perea}}\email{j.pereabenitez@northeastern.edu}
+%\affil[2]{
+%	\orgdiv{Department of Mathematics and Khoury College of Computer Sciences},
+%	\orgname{Northeastern University}}
 % \orgaddress{\street{360 Huntington Ave}, \city{Boston}, \postcode{02115}, \state{MA}}
@@
 
-%\maketitle  % arxiv only, ensures abstract follows title
+\maketitle  % arxiv only, ensures abstract follows title
 
 % REQUIRED - "Abstracts must be able to stand alone and so cannot contain citations to the paper's references, equations, etc. .... not exceeding 250 words"
-%\begin{abstract} % arxiv only 
-\abstract{ % springer only
+\begin{abstract} % arxiv only 
+%\abstract{ % springer only
 Matrix reduction is the standard procedure for computing the persistent homology of a filtered simplicial complex with $m$ simplices.
@@
 Applications to video data, dynamic metric space data, and multiparameter persistence are also presented.
-} % springer only
-%\end{abstract} % arxiv only
-
-\maketitle %  Springer only, ensures abstract follows title
+%} % springer only
+\end{abstract} % arxiv only
+
+%\maketitle %  Springer only, ensures abstract follows title
 
@@
 % --- Springer only (REQUIRED) ---
-\pacs[MSC Classification]{68T09, 55N31, 62R40}
+%\pacs[MSC Classification]{68T09, 55N31, 62R40}
 
@@
  	\centering
- 	\includegraphics[width=0.75\textwidth]{circle_vineyards.png}
+% 	\includegraphics[width=0.75\textwidth]{circle_vineyards.png} % springer only 
+	\includegraphics[width=0.70\textwidth]{circle_vineyards.png}
  	\caption{The cumulative column operations needed to compute persistence across the time-varying filtration of grayscale images.
@@
 
- The time complexity of \emph{vineyards} is determined entirely by the complexity of performing a single adjacent transposition.
-Formally, since column operations are the largest complexity operations needed and each column can have potentially $O(m)$ entries, the complexity of \emph{vineyards} is $O(m)$ per transposition. 
+ The time complexity of \emph{vineyards} is determined entirely by the complexity of performing  adjacent transpositions.
+Since column operations are the largest complexity operations needed and each column can have potentially $O(m)$ entries, the complexity of \emph{vineyards} is $O(m)$ per transposition. 
 Inspection of the individual cases of the algorithm from~\cite{cohen2006vines} shows that any single transposition requires at most two such operations on both $R$ and $V$.
@@
 We may zero these entries in $V$ using column operations $V(\ast)$, ensuring invariant I2~\eqref{inv:decomposition} is maintained, however these operations may render $R' = P R(\ast) P^T$ unreduced, breaking invariant I3.
-Of course, we could then reduce $R'$ with additional column operations, but the number of such operations scales $O(\lvert i - j\rvert^2)$, which no more efficient than simply performing the permutation and applying the reduction algorithm to columns $[i,j]$ in $R$.
+Of course, we could then reduce $R'$ with additional column operations, but the number of such operations scales $O(\lvert i - j\rvert^2)$, but this is no more efficient than simply performing the permutation and applying the reduction algorithm to columns $[i,j]$ in $R$.
 
@@
 As shown by example 2.3, \emph{moves} can be cheaper than \emph{vineyards} in terms of column operations.
-However, it is not clear that this is always the case upon inspection of~\ref{alg:mr}, as the usage of a donor column seemingly implies that many $O(m)$ copy operations need to be performed.
+However, it is not clear that this is always the case upon inspection of Algorithm~\ref{alg:mr}, as the usage of a donor column seemingly implies that many $O(m)$ copy operations need to be performed.
 It turns out that we may handle all such operations except the first in $O(1)$ time, which we formalize below.
@@
 
-Now consider a single MoveRight($i$,$j$) outlined in Algorithm~\ref{alg:mr}. Here, the dominant cost again are the column operations (line 5). 
+Now consider MoveRight($i$,$j$), outlined in Algorithm~\ref{alg:mr}. Here, the dominant cost again are the column operations (line 5). 
 Though we need an extra $O(m)$ storage allocation for the donor columns $d_\ast$ prior to the movement, notice that assignment to and from $d_\ast$ (lines (4), and (7) in \textit{RestoreLeft} and \textit{MoveRight}, respectively) requires just $O(1)$ time via a pointer swapping argument.
@@
 As $F$ determines the number of distinct filtrations encountered during the deformation from $f$ to $f'$, a natural question is whether such an interpolation can be modified so as to minimize $\lvert \mathcal{S}\rvert$---the number of times the decomposition is restored to a valid state.
-Towards explaining the phenomenon exhibited in Figure~\ref{fig:vineyards}, we begin by analyzing a particular class of interpolation schemes in order to establish an upper bound on this quantity.
+Towards explaining the phenomenon exhibited in Figure~\ref{fig:vineyards}, we begin by analyzing a class of interpolation schemes to establish an upper bound on this quantity.
 
@@
 If our goal is to decrease  $\lvert S_F \rvert$, one option is to \emph{coarsen} $S_F$ to a new schedule $\widetilde{S}_F$ by e.g.
-collapsing contiguous sequences of adjacent transpositions to moves, via the map:
-\begin{equation}\label{eq:tr_to_mv}
-	 (i, i+1)(i+1, i+2)\cdots(j-1, j) \mapsto (j, i+1, \cdots, j-1, i)  \quad \text{ if } i < j
-\end{equation}
+collapsing contiguous sequences of adjacent transpositions to moves, via the map $(i, i+1)(i+1, i+2)\cdots(j-1, j) \mapsto (j, i+1, \cdots, j-1, i)$ (if $i < j$).
+%\end{equation}
 Clearly $\lvert \widetilde{S}_F \rvert \leq \lvert S_F \rvert$ and the associated coarsened $\widetilde{S}_F$ requires just $O(m)$ time to compute.
@@
 \end{equation}
-In cycle notation, this corresponds to the cyclic permutation:
-\begin{equation}\label{eq:move_permutation}
-m_{ij} = (\enspace i \enspace j \enspace j\text{-}1 \enspace \dots \enspace i\text{+}2 \enspace i\text{+}1 \enspace )
-\end{equation}
+In cycle notation, this corresponds to the cyclic permutation $m_{ij} = (\enspace i \enspace j \enspace j\text{-}1 \enspace \dots \enspace i\text{+}2 \enspace i\text{+}1 \enspace )$.
+%\begin{equation}\label{eq:move_permutation}
+%m_{ij} = (\enspace i \enspace j \enspace j\text{-}1 \enspace \dots \enspace i\text{+}2 \enspace i\text{+}1 \enspace )
+%\end{equation}
 Observe that a move operation can be interpreted as a paired delete-and-insert operation, i.e.
@@
 
-First, we require a few definitions.
 Recall that a \emph{sorting} $\mathcal{S}$ with respect to two permutations $p, q \in S_m$ is an ordered sequence of permutations $\mathcal{S} = (s_1, s_2, \dots, s_d)$ satisfying $q = s_d \circ \dots s_1 \circ p$.
@@
 Thus, obtaining a sorting $p \mapsto q$ of size $d = m - \lvert \mathcal{L} \rvert$ reduces to applying a sequence of moves in the complement of $\mathcal{L}$.
- Formally, we define a permutation $s \in S_m$ as a \emph{valid} operation with respect to a fixed pair $p, q \in S_m$ if:
+ Formally, we define a permutation $s \in S_m$ as a \emph{valid} operation with respect to a fixed pair $p, q \in S_m$ if it satisfies:
  \begin{equation}\label{eq:valid_move}
@@
  \end{equation}
-  The problem of constructing a sorting $\mathcal{S}$ of size $d$ thus reduces to the problem of choosing a sequence of $d$ valid moves, which we call a \emph{valid sorting}. 
+  The problem of constructing a sorting $\mathcal{S}$ of size $d$ thus reduces to choosing a sequence of $d$ valid moves, which we call a \emph{valid sorting}. 
   To do this efficiently, let $\mathcal{U}$ denote an ordered set-like data structure that supports the following operations on elements $\sigma \in M$ from the set $M = \{0, 1, \dots, m+1\}$:
@@
 Thus, $\mathcal{A}$ may be partitioned into at most four contiguous intervals, each upon which the update is constant.
-Such updates called \emph{range updates}, and are known to require $O(\log m)$ time using e.g.
+Such updates are called \emph{range updates}, and are known to require $O(\log m)$ time using e.g.
 an \emph{implicit treap} data structure~\cite{blelloch1998fast}.
@@
 %thus the greedy solution to equation~\ref{eq:schedule_opt_succ} can be computed in $O(d^2 \log m)$ time. 
-\\
-\\
-\noindent \textbf{Example:} 
+We give an example of this partitioning in Figure~\ref{fig:displacement_proof}.
+%\\
+%\\
+%\noindent \textbf{Example:} 
 %valid moves in a corresponding $RV$ decomposition (not shown).    into new positions (given by the following row) 
-We re-use the previous example of sorting the sequence $p = (4, 2, 7, 1, 8, 6, 3, 5,0)$ to the identity $\iota = [m]$ using a precomputed LIS $\mathcal{L} = (1, 3, 5)$.
-The left table we show the same sorting as before, with elements chosen to move highlighted in yellow; on the right, a table showing the entries of $\mathcal{A}$ are recorded.
-\begin{figure}[htb!]\label{fig:schedule_construct}
+%We re-use the previous example of sorting the sequence $p = (4, 2, 7, 1, 8, 6, 3, 5,0)$ to the identity $\iota = [m]$ using a precomputed LIS $\mathcal{L} = (1, 3, 5)$.
+In the left table shown below, we display the same sorting as before, with elements chosen to move highlighted in yellow; on the right, a table showing the entries of $\mathcal{A}$ are recorded.
+\begin{figure}[htb!]
 	\center
-	\includegraphics[width=0.80\textwidth]{displacement_proof}
+	\caption{We re-use the previous example of sorting the sequence $p = (4, 2, 7, 1, 8, 6, 3, 5,0)$ to the identity $\iota = [m]$ using a precomputed LIS $\mathcal{L} = (1, 3, 5)$.
+In the left table shown below, we display the same sorting as before, with elements chosen to move highlighted in yellow; on the right, a table showing the entries of $\mathcal{A}$ are recorded.
+}
+	\includegraphics[width=0.70\textwidth]{displacement_proof}
+	\caption*{The colors in the right table group the range updates: green values are unchanged between rows, blue values are modified (and shifted) by $\pm 1$, and orange values are modified arbitrarily.
+As before, the $i$th entry of the column on the left-side shows $F(\hat{s}_i \circ p, \iota)$.
+Small black lines are used to show how the entries in $\mathcal{A}$ which change by $\pm 1$ move.
+}\label{fig:displacement_proof}
 %	\caption{Greedy schedule construction example.  }
 \end{figure}
-The colors of each entry in the right table indicate its partition: green values are unchanged between permutations, blue values are modified by $\pm 1$, and orange values are modified arbitrarily.
-As before, the $i$th entry of the column on the left-side shows $F(\hat{s}_i \circ p, \iota)$.
-Small black lines are used to show the movement of the entries in $\mathcal{A}$ which change by $\pm 1$.
 
@@
 
-In the second test, we aim to measure the asymptotics of our greedy LCS-based approach.
+In the second test, we measure the asymptotics of our greedy LCS-based approach.
 To do this, we generated 8 video data sets again of the expanding annulus outlined in section~\ref{sec:motivation}, each of increasing grid sizes of $5 \times 5$, $6 \times 6$, $\dots$, $12 \times 12$.
@@
 	\centering
-	\includegraphics[width=0.7\textwidth, height=12.15em]{boid_sim_results.png} % arxiv only
+	\includegraphics[width=0.6\textwidth, height=14.15em]{boid_sim_results.png} % arxiv only
 %	\includegraphics[width=\textwidth, height=16.15em]{boid_sim_results.png}
@@
 For example, better bounds or algorithms may be obtainable if stronger assumptions can be made on how the filtration is changing with time.
-Moreover, if the filtration $(K, f)$ shares little similarity to the ``target'' filtration $(L,f')$, then the overhead of reducing the simplices from $L \setminus K$ appended to the decomposition derived from $K$ may be large enough to motivate simply computing the decomposition at $L$ independently, especially if parallel processors are available.
+Moreover, if the filtration $(K, f)$ shares little similarity to the ``target'' filtration $(L,f')$, then the overhead of reducing the simplices from $L \setminus K$ appended to the decomposition derived from $K$ may be large enough to motivate simply computing the decomposition at $L$ independently.
 Our approach is primarily useful if the filtrations in the parameterized family are ``nearby'' in the combinatorial sense.
@@
 %\bibliographystyle{unsrt}
-% \bibliographystyle{plain} % arxiv only
+ \bibliographystyle{plain} % arxiv only
 \bibliography{move_schedules}
@@
 Since there exist complexes where $k \sim O(m)$, one concludes the bound of $O(m^3)$ is tight~\cite{morozov2005persistence}, though this seems to only be true on pathological inputs.
-Indeed, a more refined analysis by Edelsbrunner et al.~\cite{edelsbrunner2000topological} shows the reduction algorithm scales by the sum of squares of the cycle persistences, which is an output-sensitive bound.
+A more refined analysis by Edelsbrunner et al.~\cite{edelsbrunner2000topological} shows the reduction algorithm scales by the sum of squares of the cycle persistences.
 
@@
     		    % \State $p, p^{-1} \gets m_{ij} \circ p, p^{-1} \circ m_{ij}^{-1}$
-    		    \State $\bar{p}^{-1} \gets \bar{p}^{-1} \circ m_{ij}^{-1}$ where $m_{ij}$ is given by~\eqref{eq:move_permutation}\Comment{$O(m)$}
+    		    \State $\bar{p}^{-1} \gets \bar{p}^{-1} \circ m_{ij}^{-1}$ where $m_{ij}$ is given by~\eqref{eq:move_perm}\Comment{$O(m)$}
     		 \EndWhile
@@
 Drawing $(p, m_{ij} \circ p)$ as a bipartite graph $(U, V, E)$, observe that there are $\lvert i - j \rvert$ edge crossings in the graph, and thus minimizing ~\eqref{eq:upper_bound_move_cost} is akin to a  structured variation of the $k$-layered crossing minimization problem.
-We give an example visually relating the problem of minimizing \emph{net displacement} to the problem of minimizing crossing in a $k$-layered bipartite graph. 
-Let  $p= (1\;2\;3\;4\;5\;6\;7\;8\;9)$ and $q = (9\;4\;2\;7\;1\;8\;6\;3\;5)$.
+This is shown in an example below.
+\\
+\\
+\textbf{Example}: Let  $p= (1\;2\;3\;4\;5\;6\;7\;8\;9)$ and $q = (9\;4\;2\;7\;1\;8\;6\;3\;5)$.
 An example of three possible schedules, $S_1$, $S_2$, and $S_3$ sorting $p$ into $q$ is given in the figure below. 
@@
 and the edges track the movement of each element of the permutation. 
-Black/red vertices correspond to elements inside and outside the $\mathrm{LCS}$, respectively. 
+Red vertices track the $\mathrm{LCS}$ under each permutation. 
 All three schedules were generated from the same $\mathrm{LCS}(p,q) = (4\,7\,8)$ and each schedule transforms $p \mapsto q$ in $d = 6$ moves. 
 In this example, $S_1$ matches the minimal number of crossings amongst all possible schedules,  since $K_\tau(p,q) = 21$. 
+\\
+\\
+\noindent
+We note that the problem we seek to solve is more structured than crossing minimization, as we are restricted to performing \emph{valid} cyclic permutations (i.e.
+permutations which increase the size of the LIS).
 

### 2023-10-16 src/move_schedules.tex
Aarxiv version

@@
 many transpositions can be quite slow...
-it is sometimes much faster to simply recompute the $RU$-decomposition from scratch using the standard persistence algorithm.'' Indeed, they observe that maintaining the decomposition along a certain parameterized family is the most computationally demanding aspect of RIVET~\cite{rivet}, a software for computing and visualizing	 two-parameter persistent homology.
+it is sometimes much faster to simply recompute the $RU$-decomposition from scratch using the standard persistence algorithm.'' Indeed, they observe that maintaining the decomposition along a certain parameterized family is the most computationally demanding aspect of RIVET~\cite{rivet}, a software for computing two-parameter persistent homology.
 
@@
     \includegraphics[height=1.25in]{circle_complex.png}
-    \caption{Top: A video of an expanding annulus. Bottom: Sublevel-set filtrations, via (negative) pixel intensity, of a Freudenthal triangulation of the plane.}
+    \caption{Top: A video of an expanding annulus. Bottom: Sublevel-set filtrations, via negative pixel intensity, of a Freudenthal triangulation of the plane.}
     \label{fig:vidExample}
@@
 \textbf{Outline:} The paper is organized as follows: we review and establish the notations we will use to describe simplicial complexes, persistent homology, and dynamic persistence in Section~\ref{sec:background}.
-We also cover the reduction algorithm (designated here as \texttt{pHcol}), the vineyards algorithm, and the set of \emph{move}-related algorithms introduced in~\cite{busaryev2010tracking}, which serves as the starting point of this work.
-In Section \ref{sec:move_schedules} we introduce move schedules and provide efficient algorithms to construct them. 
-In Section \ref{sec:results} we present applications of the proposed method, including the computation of Crocker stacks from flock simulations and of a 2-dimensional persistence invariant on a data set of image patches derived from natural images.
+We also cover the reduction algorithm (designated here as \texttt{pHcol}), the \emph{vineyards} algorithm, and the set of \emph{move}-related algorithms introduced in~\cite{busaryev2010tracking}, which serves as the starting point of this work.
+In Section \ref{sec:move_schedules} we introduce \emph{move schedules}, and provide efficient algorithms to construct them. 
+In Section \ref{sec:results} we present applications of the proposed method, including applications related to dynamic metric spaces and 2-parameter persistence. 
 In Section \ref{sec:conclusion} we conclude the paper by discussing other possible applications and future work.
@@
 If a homology class is born at   $K_i$ and dies entering   $K_j$, the difference $\lvert i - j \rvert$ is called the \emph{persistence} of that class.
-In practice, filtrations often arise from triangulations parameterized by geometric scaling parameters, and the ``persistence'' of a homology class actually refers to its lifetime with respect to the scaling parameter.
+In practice, filtrations often arise from triangulations parameterized by geometric scaling parameters, and the ``persistence'' of a homology class refers to its lifetime with respect to the scaling parameter.
 %For this reason, we'll refer to the difference $\lvert i - j \rvert$ of a homology class $[c]$ as its \emph{index persistence}.
@@
 \textbf{ Decomposition Invariants:}
- \begin{enumerate}[labelsep=3pt, topsep=2pt,itemsep=-0.25ex,parsep=1.0ex]\label{inv:decomposition}
+ \begin{enumerate}[labelsep=3pt, topsep=1pt,itemsep=-0.25ex,parsep=1.0ex]\label{inv:decomposition}
  	\item[I1.] $R = D V$ where $D$ is the boundary matrix of the filtration $(K,f)$
@@
  	\end{enumerate} 
- \vspace*{1em}
+% \vspace*{1em} % springer only
  where $\mathrm{low}_R(i)$ denotes the largest row index of a non-zero entry in column $i$ of $R$. 
 We call the decomposition satisfying these three invariants \emph{valid}.
-The persistence diagrams of the corresponding filtration  can be determined from the lowest entries in $R$, once it has been reduced.
-Note that though the matrices $R$ and $V$ are not unique, the collection of persistent pairings are~\cite{zomorodian2005computing}.
+The persistence diagrams of the corresponding filtration  can be determined from the lowest entries in $R$.
+Note that though $R$ and $V$ are not unique, the collection of persistent pairings are~\cite{zomorodian2005computing}.
 
@@
 \noindent
-Inspection of the reduction algorithm from~\cite{edelsbrunner2000topological} suggests that a loose upper bound for the reduction is $O(m^3)$, where $m$ is the number of simplices of the filtration---this bound is in fact tight~\cite{morozov2005persistence}.
+Inspection of the algorithm from~\cite{edelsbrunner2000topological} suggests an upper bound for the reduction is $O(m^3)$, where $m$ is the number of simplices of the filtration---this bound is in fact tight~\cite{morozov2005persistence}.
 % the number of column operations has been observed to scale linearly in the size of the filtration in practice~\cite{bauer2021ripser}, due in part to the high sparsity and structure of $D$. 
-Despite this high algorithmic complexity, many variations and optimizations to Algorithm~\ref{alg:reduce} have been proposed over the past decade, see~\cite{bauer2021ripser, bauer2017phat, chen2011persistent} for an overview.
+Despite its high algorithmic complexity, many variations and optimizations to Algorithm~\ref{alg:reduce} have been proposed over the past decade, see~\cite{bauer2021ripser, bauer2017phat, chen2011persistent} for an overview.
 % Important corollaries of the persistence pairing lemma is that the 
@@
 
-To bypass this difficulty, Busaryev et al. observed that since $R$ is reduced, if it contains $s$ pivot entries in the columns $[i,j]$ of $R$, then $R'$ must also have $s$ pivots. 
-Thus, if column operations render some pivot-column $r_k$ of $R$ unreduced, then its pivot entry $\mathrm{low}_R(k)$ becomes \emph{free}\footnote{The process of donating pivot columns using $O(1)$ auxiliary storage is similar in spirit to the in-place sorting algorithm \emph{cycle sort}, which is often used to sort permutations in $O(n)$ time.}---if $r_{k}$ is copied prior to its modification, we may re-use or \emph{donate} its pivot entry to a later column $r_{k+1}, \dots, r_j$.
+To bypass this difficulty, Busaryev et al. observed that if $R$ contains $s$ pivot entries in the columns $[i,j]$, then since it is reduced, $R'$ must \emph{also} have $s$ pivots. 
+Thus, if column operations render some pivot-column $r_k$ of $R$ unreduced, then its pivot entry $\mathrm{low}_R(k)$ becomes \emph{free}\footnote{The process of donating pivot columns using $O(1)$ auxiliary storage is similar in spirit to the in-place sorting algorithm \emph{cycle sort}, which is often used to sort permutations in $O(n)$ time.}---if $r_{k}$ is copied prior to modification, we may re-use or \emph{donate} its pivot entry to a later column $r_{k+1}, \dots, r_j$.
 %possibly appearing as a pivot entry in some later column $r_{k+1}$.
@@
 \noindent Though similar to \emph{vineyards}, move operations confer additional advantages:
-\\	
- \begin{enumerate}[labelsep=5pt, topsep=0pt,itemsep=-0.25ex,parsep=1.2ex]
+ \begin{enumerate}[labelsep=5pt, topsep=1pt,itemsep=-0.25ex,parsep=1.2ex]
  	\item[M1:] Querying the non-zero status of entries in $R$ or $V$ occurs once per move.
@@
  \end{enumerate} 
- \vspace*{1em}
+% \vspace*{1em} % springer only
 First, consider property M1.
@@
 	\center
-	\includegraphics[width=0.90\textwidth]{greedy_schedule_construct}
+%	\includegraphics[width=0.90\textwidth]{greedy_schedule_construct} % springer only 
+	\includegraphics[width=0.80\textwidth]{greedy_schedule_construct} % arxiv only 
 %	\caption{Greedy schedule construction example.  }
@@
 As before, the $i$th entry of the column on the left-side shows $F(\hat{s}_i \circ p, \iota)$.
-Small black lines are used to show how the entries in $\mathcal{A}$ which change by $\pm 1$ move.
+Small black lines are used to show how the entries in $\mathcal{A}$ which change by $\pm 1$ each move.
 }\label{fig:displacement_proof}
@@
 As shown with the natural images example in section~\ref{sec:results}, there are often more permutation operations being applied than there are column reductions.
-In the more standard \emph{compressed} sparse matrix representations\footnote{
-By ``standard,'' we mean any of the common sparse representations used in scientific computing packages, like SciPy's sparse module (\url{https://docs.scipy.org/doc/scipy/reference/sparse.html})
-}, permuting both the rows and columns generally takes at most $O(Z)$ time, where $Z$ is the number of non-zero entries, which can be quite expensive if the particular filtration has many cycles.
+In the more standard \emph{compressed} sparse matrix representations,
+%\footnote{By ``standard,'' we mean any of the common sparse representations used in scientific computing packages, like SciPy's sparse module (\url{https://docs.scipy.org/doc/scipy/reference/sparse.html})},
+permuting both the rows and columns generally takes at most $O(Z)$ time, where $Z$ is the number of non-zero entries, which can be quite expensive if the particular filtration has many cycles.
 As a result, the more complex sparse matrix representation from~\cite{cohen2006vines} is necessary to be efficient in practice.
@@
 
-\subsubsection{Reduction Algorithm}
-The reduction algorithm, also called the ``standard algorithm,'' is the most widely used modality for computing persistence.
+\textbf{Reduction Algorithm: } The reduction algorithm, outlined in pseudocode in Algorithm~\ref{alg:reduce}, is the most widely used modality for computing persistence.
 While there exists other algorithms for computing persistence, they are typically not competitive with the reduction algorithm in practice.
-We outline the reduction pseudocode in Algorithm~\ref{alg:reduce}.
 \begin{algorithm}[h!]
@@
 A more refined analysis by Edelsbrunner et al.~\cite{edelsbrunner2000topological} shows the reduction algorithm scales by the sum of squares of the cycle persistences.
-
-\subsubsection*{Move Algorithms}
-As we have covered the \emph{moves} algorithm extensively in Section~\ref{sec:moves}, we now record the algorithmic components of both \emph{MoveRight} and \emph{MoveLeft}.
+\\
+\\
+\noindent \textbf{Move Left:}
 Though conceptually similar, note that there is an asymmetry between \emph{MoveRight} and \emph{MoveLeft}: moving a simplex upwards in the filtration requires removing non-zero entries along several columns of a particular row in $V$ so that the corresponding permutation does not render $V$ non-upper triangular.
@@
 For a supplementary description of the move algorithm, see~\cite{busaryev2010tracking}.
-
 %We recall an important claim given in~\cite{busaryev2010tracking} on the effect that move operations have on the status of simplices in the pairing.
@@
 If $i < j$ (respectively, $j < i$), all simplices at positions $k \in [i+1\;:\;j] $ are shifted down (respectively, up) by $1$.
-
-\subsubsection{LCS-Sort}
-Here we record explicitly the schedule construction algorithm outlined in Section~\ref{sec:schedule_construction}.
-The algorithm is simple enough to derive using the rules discussed in section~\ref{sec:schedule_construction} (namely, equation~\eqref{eq:valid_move}, but nonetheless for posterity’s sake, we record it here for the curious reader; the pseudocode is given in Algorithm~\ref{alg:sorting}.
-
-The high level idea of the algorithm is to first construct the LCS between two permutations $p,q \in S_m$.
-To do this efficiently, one re-labels $q \mapsto \iota$ to the identity permutation $\iota = [m]$ and applies a consistent re-labeling $p \mapsto \bar{p}$.
+%\subsubsection{LCS-Sort}
+\\
+\\
+\noindent \textbf{LCS-Sort:}
+The algorithm to construct a schedule from a given LIS is simple enough to derive using the rules discussed in section~\ref{sec:schedule_construction} (namely, equation~\eqref{eq:valid_move}, but nonetheless for posterity’s sake we record it here for the curious reader; the pseudocode is given in Algorithm~\ref{alg:sorting}.
+
+The high level idea of the algorithm is to first construct the LCS between two permutations $p,q \in S_m$ and then apply cyclic permutations successively to $p$, each step adding a new element extending the LCS.
+The algorithmic steps are as follows: first, one re-labels $q \mapsto \iota$ to the identity permutation $\iota = [m]$ and applies a consistent re-labeling $p \mapsto \bar{p}$.
 This relabeling preserves the LCS distance and has the additional advantage that $\bar{q} = \iota = [m]$ is a strictly increasing subsequence, and thus computing the LCS between $p,q \in S_m$ reduces to computing the LIS $\mathcal{L}$ of $\bar{p}$.
+Given $\mathcal{L}$ computed from $\bar{p}$, since $\mathcal{L}$ is strictly increasing, the only elements left to permute are in $\mathcal{L} \setminus \bar{p}$, which we denote with $\mathcal{D}$.
+After choosing any $\sigma \in \mathcal{D}$, one applies a cyclic permutation to $\bar{p}$ that moves $\sigma$ to any position that increases the size of $\mathcal{L}$, continuing this way until the identity sequence $\iota$ is recovered.
 By sorting $\bar{p} \mapsto \iota$ by operations that (strictly) increase the size of $\mathcal{L}$, we ensure that the size of the corresponding schedule is exactly $m - \lvert \mathcal{L} \rvert$. 
-
   \begin{algorithm}[h!]
@@
 \end{algorithm}
-The algorithmic steps are as follows: given a LIS $\mathcal{L}$ has been computed from $\bar{p}$, since $\mathcal{L}$ is strictly increasing, the only elements left to permute are in $\mathcal{L} \setminus \bar{p}$, which we denote with $\mathcal{D}$.
-After choosing any $\sigma \in \mathcal{D}$, one then applies a cyclic permutation to $\bar{p}$ that moves $\sigma$ to any position that increases the size of $\mathcal{L}$.
-To do this efficiently, we use a set-like data structure $\mathcal{U}$ that supports efficient querying the successor and predecessor of any given $s \in \bar{p}$ with respect to $\mathcal{L}$ (such as a vEB tree).
+
+To build the schedule of permutations efficiently, we use a set-like data structure $\mathcal{U}$ that supports efficient querying the successor and predecessor of any given $s \in \bar{p}$ with respect to $\mathcal{L}$ (such as a vEB tree).
 The simplified pseudocode also uses the inverse permutation $\bar{p}$ to query the position of a given element $\sigma \in \bar{p}$, although a more efficient representation can be used via an implicit treap of the \emph{displacements} array; see Section~\ref{sec:proxy_objective}.
@@
 The importance of this theorem lies in the fact that the fibered barcodes are completely defined from the precomputed barcode templates $\mathcal{T}$---once every barcode template $\mathcal{T}^e$ has been computed and augmented onto $\mathcal{A}(M)$, $\mathcal{B}(M)$ is completely characterized, and the barcodes $\mathcal{B}_L(M)$ associated to a 1-D filtration induced by \emph{any} choice of $L$ can be efficiently computed via a point-location query on $\mathcal{A}(M)$ and a $O(\lvert \mathcal{B}_L(M) \rvert)$ application of the push map.
-
-\subsubsection{Invariant computation}
-Computationally, the algorithm from~\cite{lesnick2019computing} can be summarized into three steps: 
+\\
+\\
+\noindent \textbf{Invariant computation: } Computationally, the algorithm from~\cite{lesnick2019computing} can be summarized into three steps: 
 \begin{enumerate}

### 2023-12-11 src/move_schedules.tex
Going to try to compile on overleaf for the bbl file
final revision pre-proof

@@
-\documentclass{article} % for arxiv only
-\usepackage{arxiv} % for arxiv only 
-\pagestyle{plain} % arxiv only 
+%\documentclass{article} % for arxiv only
+%\usepackage{arxiv} % for arxiv only 
+%\pagestyle{plain} % arxiv only 
 %\usepackage[margin=1in]{geometry} % arxiv only
 
-%\PassOptionsToPackage{noend}{algorithmic} % springer only 
-%\PassOptionsToPackage{noend}{algpseudocode} % springer only 
-%\documentclass[sn-mathphys]{sn-jnl} % springer only
+\PassOptionsToPackage{noend}{algorithmic} % springer only 
+\PassOptionsToPackage{noend}{algpseudocode} % springer only 
+\documentclass[sn-mathphys]{sn-jnl} % springer only
 
@@
 \usepackage{blkarray, bigstrut}
-\usepackage{epstopdf} % \DeclareGraphicsExtensions{.pdf,.png} % for pdf 
+%\usepackage{epstopdf} % \DeclareGraphicsExtensions{.pdf,.png} % for pdf 
 \usepackage{caption} 
@@
 \usepackage{enumitem}
-\usepackage{hyperref} % load hyperref last
+%\usepackage{hyperref} % load hyperref last
 
@@
 % --- arxiv only ---- 
-\usepackage{xcolor}
-\usepackage{algorithm}% http://ctan.org/pkg/algorithms
-\PassOptionsToPackage{noend}{algpseudocode}
-\usepackage{algpseudocode}% http://ctan.org/pkg/algorithmicx
+%\usepackage{xcolor}
+%\usepackage{algorithm}% http://ctan.org/pkg/algorithms
+%\PassOptionsToPackage{noend}{algpseudocode}
+%\usepackage{algpseudocode}% http://ctan.org/pkg/algorithmicx
+%\bibliographystyle{plain} % arxiv only; be sure to remove one before bibliography
 
@@
 % --- Springer only ---
-%\title[Fast Persistence Computations in Dynamic Settings]{
-%	\vspace{-2.0em} Move Schedules: Fast persistence computations in coarse dynamic settings
-%	\footnote{
-%	\textbf{Funding:} This work was partially supported by the National Science Foundation through grants CCF-2006661 and CAREER award DMS-1943758.
-%	}
-%}
+\title[Fast Persistence Computations in Dynamic Settings]{
+	\vspace{-2.0em} Move Schedules: Fast persistence computations in coarse dynamic settings
+	\footnote{
+	\textbf{Funding:} This work was partially supported by the National Science Foundation through grants CCF-2006661 and CAREER award DMS-1943758.
+	}
+}
 
 % --- Arxiv only --- 
-\title{\vspace{-0.0em} Move Schedules: Fast persistence \\ computations  in coarse dynamic settings\thanks{This work was partially supported by the National Science Foundation through grants CCF-2006661 and CAREER award DMS-1943758.}\vspace{-0.5em}}
-\author{Matt Piekenbrock\thanks{Khoury College of Computer Sciences, Northeastern University.}\;\; and Jose A.
+%\title{\vspace{-0.0em} Move Schedules: Fast persistence \\ computations  in coarse dynamic settings\thanks{This work was partially supported by the National Science Foundation through grants CCF-2006661 and CAREER award DMS-1943758.}\vspace{-0.5em}}
+%\author{Matt Piekenbrock\thanks{Khoury College of Computer Sciences, Northeastern University.}\;\; and Jose A.
 Perea$^\dagger$\thanks{Department of Mathematics and Khoury College of Computer Sciences, Northeastern University.}}
-\date{}
+%\date{}
 
 % --- Springer only ---
-%\author[1]{\fnm{Matthew} \sur{Piekenbrock}}\email{piekenbrock.m@northeastern.edu}
-%\affil[1]{
-%	\orgdiv{Khoury College of Computer Sciences}, 
-%	\orgname{Northeastern University}}
+\author[1]{\fnm{Matthew} \sur{Piekenbrock}}\email{piekenbrock.m@northeastern.edu}
+\affil[1]{
+	\orgdiv{Khoury College of Computer Sciences}, 
+	\orgname{Northeastern University}}
 
 % --- Springer only --- 
-%\author[2]{\fnm{Jose} \sur{A. Perea}}\email{j.pereabenitez@northeastern.edu}
-%\affil[2]{
-%	\orgdiv{Department of Mathematics and Khoury College of Computer Sciences},
-%	\orgname{Northeastern University}}
+\author[2]{\fnm{Jose} \sur{A. Perea}}\email{j.pereabenitez@northeastern.edu}
+\affil[2]{
+	\orgdiv{Department of Mathematics and Khoury College of Computer Sciences},
+	\orgname{Northeastern University}}
 % \orgaddress{\street{360 Huntington Ave}, \city{Boston}, \postcode{02115}, \state{MA}}
@@
 % REQUIRED - "Abstracts must be able to stand alone and so cannot contain citations to the paper's references, equations, etc. .... not exceeding 250 words"
-\begin{abstract} % arxiv only 
-%\abstract{ % springer only
+%\begin{abstract} % arxiv only 
+\abstract{ % springer only
 Matrix reduction is the standard procedure for computing the persistent homology of a filtered simplicial complex with $m$ simplices.
@@
 Applications to video data, dynamic metric space data, and multiparameter persistence are also presented.
-%} % springer only
-\end{abstract} % arxiv only
-
-%\maketitle %  Springer only, ensures abstract follows title
+} % springer only
+%\end{abstract} % arxiv only
+
+\maketitle %  Springer only, ensures abstract follows title
 
@@
 % --- Springer only (REQUIRED) ---
-%\pacs[MSC Classification]{68T09, 55N31, 62R40}
+\pacs[MSC Classification]{68T09, 55N31, 62R40}
 
@@
 \subsection{Related Work}\label{sec:related_work} 
-To the author's knowledge, work focused on ways of updating a  decomposition $R = DV$, for all homological dimensions, is limited: there is the \emph{vineyards} algorithm~\cite{cohen2006vines} and the \emph{moves} algorithm~\cite{busaryev2010tracking}, both of which are discussed extensively in section~\ref{sec:background}.
+To the authors' knowledge, work focused on ways of updating a  decomposition $R = DV$, for all homological dimensions, is limited: there is the \emph{vineyards} algorithm~\cite{cohen2006vines} and the \emph{moves} algorithm~\cite{busaryev2010tracking}, both of which are discussed extensively in section~\ref{sec:background}.
 At the time of writing, we were made aware of very recent work~\cite{luo2021accelerating} that iteratively repairs a permuted decomposition via a column swapping strategy, which they call ``warm starts.'' Though their motivation is similar to our own, their approach relies on the reduction algorithm as a subprocedure, which is quite different from the strategy we employ here.
@@
  	\centering
-% 	\includegraphics[width=0.75\textwidth]{circle_vineyards.png} % springer only 
-	\includegraphics[width=0.70\textwidth]{circle_vineyards.png}
+ 	\includegraphics[width=0.75\textwidth]{circle_vineyards.png} % springer only 
+%	\includegraphics[width=0.70\textwidth]{circle_vineyards.png}
  	\caption{The cumulative column operations needed to compute persistence across the time-varying filtration of grayscale images.
@@
 %The complexity of determining $\mathcal{S}$ depends on the type of permutations considered; many edit distances exhibit tight quadratic growth rates and tend to require at least quadratic time (or worse) to compute~\cite{labarre2013lower}.
-By reinterpreting a special set of cyclic permutations as edit operations on strings, we find that any sequence mapping $f $ to $  f'$ of minimal size must have length (Proposition~\ref{prop:schedule_size}):
+By reinterpreting a special set of cyclic permutations as edit operations on strings, we find that any sequence mapping $f$ to $f'$ via \emph{move} operations must have size at least $d$ (Proposition~\ref{prop:schedule_size}), where:
 \begin{equation}
@@
 \end{equation}
-where $\mathrm{LCS}(f,f')$ refers to the size of the longest common subsequence between the simplexwise filtrations $(K, f)$ and $(K,f')$ (see section~\ref{sec:moves_setting} for more details).
-We also show that the information needed to construct any $\mathcal{S}$ with optimal size can be computed in $O(m \log \log m)$ preprocessing time and $O(m)$ memory.
-We provide evidence  that $d \sim m - \sqrt{m}$ in expectation for random filtrations (Corollary~\ref{cor:expectation}).
+and where $\mathrm{LCS}(f,f')$ refers to the size of the longest common subsequence between the simplexwise filtrations $(K, f)$ and $(K,f')$ (see section~\ref{sec:moves_setting} for more details).
+Moreover, we provide evidence  that $d \sim m - \sqrt{m}$ in expectation for random filtrations (Corollary~\ref{cor:expectation}).
 Although this implies $d$ can be $O(m)$ for pathological inputs, we give empirical results suggesting $d$  can be much smaller in practice.
+Finally, we show that the information needed to construct any $\mathcal{S}$ with optimal size can be computed in $O(m \log \log m)$ preprocessing time and $O(m)$ memory. 
 %Finally, we perform extensive experiments 
@@
  	\end{enumerate} 
-% \vspace*{1em} % springer only
+ \vspace*{1em} % springer only
  where $\mathrm{low}_R(i)$ denotes the largest row index of a non-zero entry in column $i$ of $R$. 
@@
 The persistence diagrams of the corresponding filtration  can be determined from the lowest entries in $R$.
-Note that though $R$ and $V$ are not unique, the collection of persistent pairings are~\cite{zomorodian2005computing}.
+Note that though $R$ and $V$ are not unique, the persistent pairing is unique~\cite{zomorodian2005computing}.
 
@@
 		\begin{block}{c[ccc]}
-  		b \kern\rowsp  & \; 1 & 1 & 1 \; \topstrut \\
-  		a \kern\rowsp & \;     & 1 & 1 \; \\
+  		a \kern\rowsp  & \; 1 & 1 & 1 \; \topstrut \\
+  		b \kern\rowsp & \;     & 1 & 1 \; \\
   		c \kern\rowsp & \;     &    & 1 \; \botstrut \\
@@
  \end{enumerate} 
-% \vspace*{1em} % springer only
+ \vspace*{1em} % springer only
 First, consider property M1.
@@
 As shown by example 2.3, \emph{moves} can be cheaper than \emph{vineyards} in terms of column operations.
-However, it is not clear that this is always the case upon inspection of Algorithm~\ref{alg:mr}, as the usage of a donor column seemingly implies that many $O(m)$ copy operations need to be performed.
+However, it is not clear that this is always the case upon inspection of Algorithm~\ref{alg:mr}, as the usage of a donor column suggests that many $O(m)$ copy operations are required.
 It turns out that we may handle all such operations except the first in $O(1)$ time, which we formalize below.
@@
 \end{proof}
+Though conceptually similar, there is an asymmetry between \emph{MoveRight} and \emph{MoveLeft}: moving a simplex upwards in the filtration (moving right) requires removing non-zero entries along a particular \emph{row} of $V$, whereas the converse requires removing non-zero entries along a particular \emph{column}.
+These entries must be removed to ensure the corresponding permutation does not render $V$ non-upper triangular.
+The key insight of the algorithm presented in~\cite{busaryev2010tracking} is that $R$ can actually be maintained in all but one column during this procedure (by employing the \emph{donor} column). 
+While in the \emph{MoveLeft} algorithm $R$ does stay reduced during this cancellation procedure in all but one column (like \emph{MoveRight}), more than one column operation per pivot may be needed  due to the fact that entries above the pivot may introduce additional non-zeros entries.
+%$R$ may cascade into a larger chain of column operations . 
+%This is due to the fact that higher entries in columns in $R$ (above the pivot entry) may very well introduce additional non-reduced columns after $R$ is permuted. 
+As a result, it is not immediately clear how to adapt the donor column concept to \emph{MoveLeft}. 
+Fortunately, like \emph{MoveRight}, we can still separate the algorithm into a reduction and restoration phase---see Algorithm~\ref{alg:ml}.
+Moreover, since $PRP^T$ is reduced in all but one column, the number of columns to reduce still depends directly on the sparsity of both $R$ and $V$ (e.g.
+see line 6 in Algorithm~\ref{alg:ml}).
+For a supplementary description of the move algorithm, see~\cite{busaryev2010tracking}.
+
 As a final remark, we note that the combination of \emph{MoveRight} and \emph{MoveLeft} enable efficient simplex additions or deletions to the underlying complex.
@@
 We pass $q$, $\mathrm{LIS}(q)$, and a ``greedy'' heuristic for picking moves $h$ to our scheduling algorithm (to be defined later), which returns as output an ordered set of move permutations  $\mathcal{S}$ of minimum size satisfying~\eqref{eq:rv_seq_schedule}---a \emph{schedule} for the pair $(f, f')$.
-These sequence of steps is outlined in Algorithm~\ref{alg:schedule} below.
+This sequence is outlined in Algorithm~\ref{alg:schedule} below.
 %Note one need not explicitly store the family of filtrations nor the schedules between them---Algorithm~\ref{alg:schedule} may be easily modified to be completely \emph{online}, keeping at most two filtrations and one decomposition in memory at any given time.
@@
 For any pair of general string inputs of sizes $n$ and $m$, respectively, the LCS between them is computable in $O(mn)$ with dynamic programming, and there is substantial evidence that the complexity cannot be much lower than this~\cite{abboud2015tight}.
-In our setting, however, we interpret the simplexwise filtrations $(K, f), (K, f')$ as \emph{permutations} of the same underlying complex $K$---in this setting, $d_{\mathrm{LCS}}$ reduces further to the \emph{permutation edit distance} problem.
+However, in the simplexwise setting, we interpret a pair of filtrations $(K, f), (K, f')$ as \emph{permutations} of the same underlying complex $K$---thus, $d_{\mathrm{LCS}}$ can be specialized to the \emph{permutation edit distance} problem.
 %Establishing a connection between the permutation edit distance and move scheduling allows us to exploit the combinatorial structure that comes from the developed theory on both LCS's and LIS's, which are both well-studied objects.
@@
 $q$).
-Thus, obtaining a sorting $p \mapsto q$ of size $d = m - \lvert \mathcal{L} \rvert$ reduces to applying a sequence of moves in the complement of $\mathcal{L}$.
+Thus, obtaining a sorting $p \mapsto q$ of size $d = m - \lvert \mathcal{L} \rvert$ reduces to applying a sequence of moves to $\mathcal{D} = p \setminus \mathcal{L}$---the complement of $\mathcal{L}$.
  Formally, we define a permutation $s \in S_m$ as a \emph{valid} operation with respect to a fixed pair $p, q \in S_m$ if it satisfies:
@@
 Similarly, if  $p^{-1}(\mathcal{U}_{\mathrm{succ}}(\sigma)) < p^{-1}(\sigma)$, then we must move $\sigma$ left of its successor in $p$.
-Assuming the structure $\mathcal{U}$ supports all of the above operations in $O(\log m)$ time, we easily deduce a $O(d m\log m)$ algorithm for obtaining a valid sorting.
+Assuming the structure $\mathcal{U}$ supports all of the above operations in $O(\log m)$ time, we easily deduce a $O(d m\log m)$ algorithm for obtaining a valid sorting (see Algorithm~\ref{alg:sorting} in Section~\ref{app:algorithms} of the appendix for the explicit pseudocode).
 %  which for completeness is shown via Algorithm~\ref{alg:sorting} in the appendix. 
@@
 Thus, it is natural to consider whether one could minimize~\eqref{eq:schedule_cost} by greedily choosing the lowest cost move in each step.
-Unfortunately, not only does this approach does not yield a minimal cost solution, we give a counter-example in the appendix (\ref{sec:greedy_counter_ex}) demonstrating such a greedy procedure may lead to arbitrarily bad behavior.
+Unfortunately, not only does this approach not yield a minimal cost solution, we give a counter-example in the appendix (\ref{sec:greedy_counter_ex}) demonstrating such a greedy procedure may lead to arbitrarily bad behavior.
 %---no more than $O(m)$ time with row-oriented sparse matrices. 
@@
 
-In light of section~\ref{sec:greedy}, we seek an alternative objective that correlates with~\eqref{eq:schedule_cost} and does not depend on the entries in the decomposition.
+Since minimizing~\eqref{eq:schedule_cost} is costly and the greedy approach is inefficient, we seek an alternative objective that correlates with~\eqref{eq:schedule_cost} and does not depend on the entries in the decomposition.
 Given a pair of filtrations $(K, f), (K,f')$, a natural schedule $\mathcal{S} \in \Phi(f,f',d)$ of cyclic permutations $(i_1, j_1), (i_2, j_2), \dots, (i_d, j_d)$ is one minimizing the upper bound:
@@
 Our motivation for considering the footrule distance is motivated by the fact that $F$ recovers $\mathrm{\widetilde{cost}}(\mathcal{S})$ when $(p,q)$ differ by a cyclic permutation (i.e.
-$F(p, m_{ij} \circ p) = 2 \lvert i - j \rvert$) and by its usage on similar\footnote{
+$F(p, m_{ij} \circ p) = 2 \lvert i - j \rvert$) and by its usage on similar combinatorial optimization\footnote{
 $F$ is often used to approximate $K_\tau$ in \emph{rank aggregation} problems due to the fact that computing the [Kemini] optimal rank aggregation is NP-hard with respect to $K_\tau$, but only polynomial time with respect to $F$~\cite{dinu2006efficient}.
 Note $F$ bounds $K_\tau$ via the inequalities $K_\tau(p, q) \leq F(p, q) \leq 2 K_\tau(p, q)$~\cite{diaconis1977spearman}.
-} combinatorial optimization problems. 
+} problems. 
 To adapt $F$ to sortings, we decompose $F$ additively via the bound:
@@
 where $\hat{s}_i =  s_{i} \circ \cdots \circ s_2 \circ s_1$ denotes the composition of the first $i$ permutations of a sorting  $S = (s_1, \dots, s_d)$ that maps $p \mapsto \iota$ and $s_0 = \iota$.
-As a heuristic to minimize~\eqref{eq:sorting_additive}, we greedily select the optimal $k \in \mathcal{L}$ at each step which minimizes the Spearman distance to the identity permutation: 
+As a heuristic to minimize~\eqref{eq:sorting_additive}, we greedily select the optimal $k \in p \setminus \mathcal{L}$ at each step which minimizes the Spearman distance to the identity permutation: 
 \begin{equation}\label{eq:greedy_step}
-	k_{\text{greedy}} = \argmin_{k \in \mathcal{D}} F(m_{ij} \circ p, \iota), \quad i = p^{-1}(k)
+	k_{\text{greedy}} = \argmin_{k \in p \setminus \mathcal{L}} F(m_{ij} \circ p, \iota), \quad i = p^{-1}(k)
 \end{equation}  
-Note that equality between $F$ and $\hat{F}_{\mathcal{S}}$~\eqref{eq:sorting_additive} is achieved when the displacement of each $\sigma \in p$ between its initial position in $p$ to its value is non-increasing with every application of $s_i$, which in general not guaranteed using the schedule construction method in section~\ref{sec:move_schedules}.
-To build intuition for how this heuristic interacts with Algorithm~\ref{alg:sorting}, we show a purely combinatorial example below.
+where $j \neq i$ is the index closest to $i$ whose corresponding permutation $m_{ij}$ satisfies~\eqref{eq:valid_move}---by definition, this will either be the successor or predecessor of $k$ in $p$, which is easily determined using the data structure discussed in Section~\ref{sec:schedule_construction}.
+Note that equality between $F$ and $\hat{F}_{\mathcal{S}}$~\eqref{eq:sorting_additive} is achieved when the displacement of each $\sigma \in p$ to its value is non-increasing with every application of $s_i$, which in general not guaranteed using the schedule construction method in section~\ref{sec:move_schedules}.
+To build intuition for how this heuristic interacts with Algorithm~\ref{alg:sorting}, a purely combinatorial example is given below.
 \\
@@
 	\center
-%	\includegraphics[width=0.90\textwidth]{greedy_schedule_construct} % springer only 
-	\includegraphics[width=0.80\textwidth]{greedy_schedule_construct} % arxiv only 
+	\includegraphics[width=0.90\textwidth]{greedy_schedule_construct} % springer only 
+%	\includegraphics[width=0.80\textwidth]{greedy_schedule_construct} % arxiv only 
 %	\caption{Greedy schedule construction example.  }
@@
 On the left, a table records permuted sequences $\{ \hat{s}_i \circ p \}_{i=0}^d$, given on rows $i \in \{0, 1, \dots, 6\}$, for the moved elements $(7, 0, 8, 4, 6, 2)$ (highlighted in yellow). 
- In the middle, we show the Spearman distance cost~\eqref{eq:greedy_step} associated with both each permuted sequence (left column) and with each candidate permutation $m_{ij} \circ p$ (rows).
-Note that elements $\sigma \in \mathcal{L}$ (red) induce identity permutations that do not modify the cost, and thus the minimization is restricted to elements $\sigma \in p \setminus \mathcal{L}$ (black).
+ In the middle, we show the Spearman distance cost~\eqref{eq:greedy_step} associated with both each permuted sequence (left column) and each candidate permutation $m_{ij} \circ p$ (rows).
+Note that elements $\sigma \in \mathcal{L}$ (red) induce identity permutations that do not modify the cost, and thus the minimization is restricted to elements $\sigma \in \mathcal{D} = p \setminus \mathcal{L}$ (black).
  The right plot shows the connection to the bipartite crossing minimization problem discussed in section~\ref{sec:greedy}.
@@
 Thus, the problem of constructing a schedule reduces to choosing  an order of elements in $\mathcal{D} = [m] \setminus \mathcal{L}(p)$ to move.
-First, consider an array $\mathcal{A}$ of size $m$ which provides $O(1)$ access and modification, initialized with the \emph{signed displacement} of every element in $p$ to its corresponding position in $q$.
-In the case where $q = \iota$, note $\mathcal{A}(i) = i - p(i)$, thus $F(p, \iota)$ is given by the sum of the absolute values of the elements in $\mathcal{A}$.
+First, consider an array $\mathcal{A}_p$ of size $m$ which provides $O(1)$ access and modification, initialized with the \emph{signed displacement} of every element in $p$ to its corresponding position in $q$.
+In the case where $q = \iota$, note $\mathcal{A}_p[i] = i - p(i)$, thus $F(p, \iota)$ is given by the sum of the absolute values of the elements in $\mathcal{A}$.
 %$\lVert \mathcal{A} \rVert_1$. 
@@
 
-Because the values of $\mathcal{A}$ are signed, the reduction to prefix sums is not exact: it is not immediately clear how to modify $\lvert i - j \rvert$ elements in $O(\log m)$ time. 
-To address this, observe that at any point during the execution of Algorithm~\ref{alg:sorting}, a cyclic permutation changes each value of $\mathcal{A}$ in at most three different ways: 
+Because the values of $\mathcal{A}_p$ are signed, the reduction to prefix sums is not exact: it is not immediately clear how to modify $\lvert i - j \rvert$ elements in $O(\log m)$ time. 
+To address this, observe that at any point during the execution of Algorithm~\ref{alg:sorting}, a cyclic permutation changes each value of $\mathcal{A}_p$ in at most three different ways: 
 \[
-\mathcal{A}(m_{ij} \circ p) = 
+\mathcal{A}_{m_{ij} \circ p}[k] = 
 \begin{cases} 
-	 \mathcal{A}(k) \pm \lvert i - j \rvert & p^{-1}(k) = i \\
-	 \mathcal{A}(k) \pm 1 & i < p^{-1}(k) \leq j \\
-	 \mathcal{A}(k) & \text{ otherwise }
+	 \mathcal{A}_p[k] \pm \lvert i - j \rvert & p^{-1}(k) = i \\
+	 \mathcal{A}_p[k] \pm 1 & i < p^{-1}(k) \leq j \\
+	 \mathcal{A}_p[k] & \text{ otherwise }
 \end{cases}
@@
 	\caption{We re-use the previous example of sorting the sequence $p = (4, 2, 7, 1, 8, 6, 3, 5,0)$ to the identity $\iota = [m]$ using a precomputed LIS $\mathcal{L} = (1, 3, 5)$.
-In the left table shown below, we display the same sorting as before, with elements chosen to move highlighted in yellow; on the right, a table showing the entries of $\mathcal{A}$ are recorded.
+In the left table shown below, we display the same sorting as before, with elements chosen to move highlighted in yellow; on the right, a table showing the entries of $\mathcal{A}$ (at every step) are recorded.
 }
@@
 As before, the $i$th entry of the column on the left-side shows $F(\hat{s}_i \circ p, \iota)$.
-Small black lines are used to show how the entries in $\mathcal{A}$ which change by $\pm 1$ each move.
+Small black lines are used to show the entries in $\mathcal{A}$ which change by $\pm 1$ each move.
 }\label{fig:displacement_proof}
@@
 	\centering
-	\includegraphics[width=0.6\textwidth, height=14.15em]{boid_sim_results.png} % arxiv only
-%	\includegraphics[width=\textwidth, height=16.15em]{boid_sim_results.png}
+%	\includegraphics[width=0.6\textwidth, height=14.15em]{boid_sim_results.png} % arxiv only
+	\includegraphics[width=\textwidth, height=16.15em]{boid_sim_results.png} % springer only
 	\caption{On the left, the cumulative number of column operations (log-scale) of the three baseline approaches tested.
@@
 	\caption{Bipersistence example on an $8 \times 8$ coarsened grid.
-On the left, the input data, colored by density.
-In the middle, the bigraded Betti numbers $\beta_0(M)$ and $\beta_1(M)$ (green and red, respectively), the dimension function (gray), and a line $L$ emphasizing the persistence of features with high density.
+On the left, the data, colored by density.
+In the middle, the bigraded Betti numbers $\beta_0(M)$ and $\beta_1(M)$ (green and red, resp.), the dimension function (gray), and a line $L$ emphasizing the persistence of features with high density.
 On the right, the line arrangement $\mathcal{A}(M)$ lying in the dual space derived from the $\beta(M)$.
@@
 If one was not aware of the analysis done by~\cite{lee2003nonlinear, hateren_schaaf_1998, carlsson2008local, perea2014klein}, it is not immediately clear a priori that the Klein bottle model is  a good candidate for capturing the non-linearity of image patches.
-Indeed, armed with a refined topological intuition, Carlsson still needed to perform extensive sampling, preprocessing, and model fitting techniques in order to reveal the underlying the topological space with persistent homology~\cite{carlsson2008local}.
+Indeed, armed with a refined topological intuition, Carlsson still needed to perform extensive sampling, preprocessing, and model fitting techniques in order to reveal the underlying topological space with persistent homology~\cite{carlsson2008local}.
 One reason such preprocessing is needed is due to persistent homology's aforementioned instability with respect to strong outliers. 
@@
 % Looks like you have to just not specify the style with springer 
-% \bibliographystyle{siamplain} % springer only 
+ \bibliographystyle{siamplain} % springer only 
 %\bibliographystyle{unsrt}
- \bibliographystyle{plain} % arxiv only
+% \bibliographystyle{plain} % arxiv only
 \bibliography{move_schedules}
@@
 
-\subsection{Algorithms}
+\subsection{Algorithms}\label{app:algorithms}
 
@@
 A more refined analysis by Edelsbrunner et al.~\cite{edelsbrunner2000topological} shows the reduction algorithm scales by the sum of squares of the cycle persistences.
-\\
-\\
-\noindent \textbf{Move Left:}
-Though conceptually similar, note that there is an asymmetry between \emph{MoveRight} and \emph{MoveLeft}: moving a simplex upwards in the filtration requires removing non-zero entries along several columns of a particular row in $V$ so that the corresponding permutation does not render $V$ non-upper triangular.
-The key insight of the algorithm presented in~\cite{busaryev2010tracking} is that $R$ can actually be maintained in all but one column during this procedure (by employing the \emph{donor} column).
-In contrast, moving a simplex to an earlier time in the filtration requires removing non-zero entries along several rows of a particular column of $V$.
-As before, though $R$ stays reduced during this cancellation procedure in all but one column, the subsequent permutation to $R$ requires reducing a pair of columns which may cascade into a larger chain of column operations to keep $R$ reduced.
-This is due to the fact that higher entries in columns in $R$ (above the pivot entry) may very well introduce additional non-reduced columns after $R$ is permuted. 
-Since these operations always occur in a left-to-right fashion, it is not immediately clear how to apply a donor column kind of concept.
-Fortunately, like move right, we can still separate the algorithm into a reduction and restoration phase---see Algorithm~\ref{alg:ml}.
-Moreover, since $R$ is reduced in all but one column by line 6 in~Algorithm~\ref{alg:ml}, we can still guarantee the number of low entries to reduce in $R$ will be at most $\lvert i - j \rvert$.
-For a supplementary description of the move algorithm, see~\cite{busaryev2010tracking}.
+%\\
+%\\
+%\noindent \textbf{Move Left:}
 %We recall an important claim given in~\cite{busaryev2010tracking} on the effect that move operations have on the status of simplices in the pairing.
